### 2020-02-20 Thesis-Chapter-Intro.tex
Build: fix live-update functionality
Have VScode rely on the external build process, but still live-update
the file.

\chapter{Introduction}

\section{Premises}

\section{Purpose}

\section{Contribution}

### 2020-02-20 scripts/wait-till-pdf-updated.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+PDF=$(git rev-parse --show-toplevel)/build/Thesis.pdf
+
+function getTs() {
+    stat --format=%Y $PDF || echo 0
+}
+
+START_TS=$(getTs)
+while [[ $START_TS == $(getTs) ]]; do
+    sleep 1
+done

### 2020-02-20 Thesis-Chapter-Background.tex
Begin writing introduction

\chapter{Background}

\section{Arguments For and Against}

\subsection{Arguments For}

\subsection{Arguments Against}

\section{Technical State of the Art}

\subsection{Historical Approaches}

\subsection{Modern Proposals}

\section{Paths Ahead}

\subsection{Legislative Priorities}

\subsection{Technical Research}

### 2020-02-20 Thesis-Chapter-Intro.tex
Begin writing introduction

@@
 \chapter{Introduction}
+
+Exceptional access is a wicked problem. Characterized by a dynamic technological environment, disagreement about
+underlying values, and resistance to a clear solution, the debate on exceptional access refuses to end. This paper will
+not attempt to end the debate, but to inform it with additional structure.
+
+\section{Motivation}
+
+\begin{itemize}
+    \item In the pursuit of data privacy, legislative action is part of the threat model
+\end{itemize}
 
@@
 
-\section{Purpose}
+\begin{itemize}
+    \item The lack of convincing of legislative bodies and law enforcement agencies demands that the technical community continue the discussion
+    \item The ability to access digital data is important and will be increasingly important for investigations
+    \item Various forms of encryption are foundational all elements of online security
+    \item Mass surveillance is a grave danger that representative governments are not immune from
+    \item International and human rights considerations significantly constrain workable solutions
+    \item Transparency to data access is critical for public trust
+\end{itemize}
 
 \section{Contribution}
+
+\begin{itemize}
+    \item Argument maps formalizing the debate
+    \item A broad threat model describing the environment
+    \item A structure for organizing future progress
+\end{itemize}

### 2020-02-20 scripts/bib-gen.sh

@@ -1,10 +1,7 @@
 #!/bin/bash
 
-pushd $(git rev-parse --show-toplevel)/reading &>/dev/null
-function cleanup() {
-  popd &>/dev/null
-}
-trap cleanup EXIT
+set -e
+cd $READING_DIR
 
 OUTFILE="../Thesis.bib"
 

### 2020-02-20 scripts/new-source.sh

@@ -9,13 +9,11 @@ SOURCE - The DOI or URL of the new material
 "
 
 ORIG_DIR=$(pwd)
-pushd $(git rev-parse --show-toplevel)/reading &>/dev/null
+cd $READING_DIR
 function cleanup() {
     if [[ 2 == $? ]]; then
         printf "$USAGE"
     fi
-    set +x
-    popd &>/dev/null
 }
 trap cleanup EXIT
 

### 2020-02-20 scripts/show-source.sh

@@ -2,7 +2,7 @@
 
 # set -x
 ORIG_DIR=$(pwd)
-cd $(git rev-parse --show-toplevel)/reading
+cd $READING_DIR
 
 if [[ 1 == $# && -f $ORIG_DIR/$1 ]]; then
     FILE=$ORIG_DIR/$1

### 2020-02-20 scripts/sourceme.sh
Scripts: export variables in sourceme.sh

@@ -1,7 +1,7 @@
 # Source this file
 
-READING_DIR="$(git rev-parse --show-toplevel)/reading"
-SCRIPTS_DIR="$(git rev-parse --show-toplevel)/scripts"
+export READING_DIR="$(git rev-parse --show-toplevel)/reading"
+export SCRIPTS_DIR="$(git rev-parse --show-toplevel)/scripts"
 
 alias new="$SCRIPTS_DIR/new-source.sh"
 alias show="$SCRIPTS_DIR/show-source.sh"

### 2020-02-26 Thesis-Chapter-Background.tex
Work on intro/background

@@
 \chapter{Background}
+
+Private use of encryption has a contentious history. [Some highlights.]
+
+Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
+human rights. [This is where I'd like to understand more of the legal side. With a historical perspective.]
+
+Privacy is not encryption's only purpose, however. Encryption is the technical foundation for many forms of computer and
+network security. [Properties: confidentiality, authentication, integrity, non-repudiation; primitives: hashed
+passwords, digital signatures, ...; technologies: PKI, TLS, SSO/Kerberos, VPNs, disk encryption, blockchains (might get
+eye-rolls).]
+
+Electronic encryption's dual role as enabler of perfect privacy and cornerstone of computer security may be the central
+conflict in the exceptional access debate. Those who focus on its role in privacy object to its absolute power. Those
+who focus on its role in security rely on its absolute power.
+
+% Encryption requires absolute cryptographic fidelity in order to provide security.
 

### 2020-02-26 Thesis-Chapter-Intro.tex
Work on intro/background

@@
     \item In the pursuit of data privacy, legislative action is part of the threat model
+    \item Legislative action is a highly likely
+    \item We as the technical community need to be prepared
 \end{itemize}
@@
 \begin{itemize}
-    \item The lack of convincing of legislative bodies and law enforcement agencies demands that the technical community continue the discussion
+    \item The failure to convince legislative bodies and law enforcement agencies demands that the technical community continue the discussion
     \item The ability to access digital data is important and will be increasingly important for investigations

### 2020-02-27 Thesis-Chapter-Background.tex
Outline background

@@
 
-\section{Arguments For and Against}
+\section{Exceptional Access: Arguments For and Against}
 
@@
 
+Primary reasons (why EA itself is beneficial):
+\begin{itemize}
+    \item Absolute privacy is not a right
+    \item Encryption gives too much power to the individual
+    \item EA would allow more crimes to be prevented and prosecuted
+\end{itemize}
+
+Secondary reasons (why studying EA is beneficial):
+\begin{itemize}
+    \item EA done well would be more secure and privacy preserving than EA done poorly
+    \item The threat of legislative action creates urgency
+\end{itemize}
+
 \subsection{Arguments Against}
+
+Primary reasons (why EA itself is bad):
+\begin{itemize}
+    \item Privacy is an absolute right
+    \item Exceptional access gives too much power to the state
+    \item Exceptional access is not possible in an acceptably secure manner
+\end{itemize}
+
+Secondary reasons (why studying EA is non-beneficial):
+\begin{itemize}
+    \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
+            them
+    \item The technical community should not entertain delusions about what is truly possible
+\end{itemize}
 
@@
 
+Historical approaches and proposals out of the first crypto war that saw various amounts of support.
+
+\begin{itemize}
+    \item Clipper chip
+    \item Government key escrow
+    \item Oblivious key escrow \cite{goos_oblivious_1996}
+    \item Partial key escrow (see \cite{denning_taxonomy_1996})
+\end{itemize}
+
 \subsection{Modern Proposals}
+
+Recent proposals out of the second crypto war that have various levels of support.
+
+\begin{itemize}
+    \item CLEAR \cite{ozzie_2018}
+    \item Device escrow with HW \cite{savage_lawful_2018}
+    \item Computational, blockchain approaches \cite{phan_key_2017}
+    \item Lawful hacking \cite{nguyen_lawful_2017} \cite{soesanto_2018}
+    \item Forced password disclosure \cite{bittenbender_2019}, see \cite{kerr_encryption_2017}
+\end{itemize}
 
@@
 
+\begin{itemize}
+    \item Reject EA until technical capabilities are ready \cite{varia_2018}
+    \item Be forthcoming with data so that we know what to work on \cite{armerding_2017}
+    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
+    \item Fund the research \cite{varia_2018}
+\end{itemize}
+
 \subsection{Technical Research}
+
+\begin{itemize}
+    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
+    \item Do the research
+    \item Focus on principles \cite{levy_robinson_2018}
+    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
+            \cite{abelson_2015}), and clearly define their interfaces with principles
+            \cite{matyas_incommensurability_2018}
+\end{itemize}

### 2020-02-28 Thesis-Chapter2-Background.tex
Improve file naming scheme, work on background

--- 2020-02-27 Thesis-Chapter-Background.tex
@@
 \chapter{Background}
+\label{chap-background}
 
@@
 \section{Exceptional Access: Arguments For and Against}
+\label{sec-argsbackground}
+
+The debate over exceptional access takes place on two fundamental levels:
+\begin{enumerate}
+    \item Whether cryptosystems with exceptional access represent a net benefit to society
+    \item Whether and how to conduct exceptional access research
+\end{enumerate}
+
+These arguments are introduced here and analyzed in detail in Chapter \ref{chap-arguments}.
 
@@
 
-Primary reasons (why EA itself is beneficial):
+Why EA is a benefit to society:
 \begin{itemize}
@@
 
-Secondary reasons (why studying EA is beneficial):
+Why to research EA:
 \begin{itemize}
@@
 
-Primary reasons (why EA itself is bad):
+Why EA is a danger to society:
 \begin{itemize}
@@
 
-Secondary reasons (why studying EA is non-beneficial):
+Why not to research EA:
 \begin{itemize}

### 2020-02-28 Thesis-Chapter3-Arguments.tex
Improve file naming scheme, work on background

\chapter{Arguments}
\label{chap-arguments}

In this chapter, develop the arguments introduced in \ref{sec-argsbackground}. Make use of argument maps to break down
and visualize the debate.

### 2020-02-28 Thesis-Chapter4-ThreatModel.tex
Improve file naming scheme, work on background

\chapter{Threat Model}
\label{chap-threatmodel}

In this chapter, develop a threat model using data flow diagrams. May need separate ones for different scenarios (data
at rest on encrypted devices, data in motion using E2E), but want at least one high level diagram encapsulating
everything.

### 2020-02-29 Thesis-Chapter3-Arguments.tex
Initial stab at building and including Argdown maps
Improve Argdown build process and image imports

@@
 In this chapter, develop the arguments introduced in \ref{sec-argsbackground}. Make use of argument maps to break down
-and visualize the debate.
+and visualize the debate, as shown in \Figure~\fref{fig-args-top-level}.
+
+% \includepdf[]{arguments/build/top-level.pdf}
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{arguments/build/top-level.pdf}
+    \caption[Argument: Top Level]{Top level arguments for and against exceptional access.}
+    \label{fig-args-top-level}
+\end{figure}

### 2020-02-29 scripts/bib-gen.sh

@@ -1,6 +1,7 @@
 #!/bin/bash
 
 set -e
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
 cd $READING_DIR
 
 OUTFILE="../Thesis.bib"

### 2020-02-29 scripts/install-deps.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+# For Bibtex scripts, must have Docker installed. Not covered here.
+# This is not necessarily fully-inclusive.
+
+sudo apt-get install -y \
+    make \
+    texlive-latex-extra \
+    texlive-science \
+    texlive-xetex \
+    texlive-generic-extra \
+    texlive-publishers \
+    latexmk \
+    inkscape \
+    libcanberra-gtk-module \
+    libcanberra-gtk3-module
+

### 2020-02-29 scripts/new-source.sh

@@ -1,5 +1,7 @@
 #!/bin/bash
 
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+
 ##################################################################### SCRIPT SETUP
 USAGE="
 $0 SUBDIR SOURCE

### 2020-02-29 scripts/show-source.sh

@@ -1,5 +1,6 @@
 #!/bin/bash
 
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
 # set -x
 ORIG_DIR=$(pwd)
 cd $READING_DIR

### 2020-02-29 Thesis-Chapter3-Arguments.tex
Add \my{chap/sec/fig/tab} commands

@@
 
-In this chapter, develop the arguments introduced in \ref{sec-argsbackground}. Make use of argument maps to break down
-and visualize the debate, as shown in \Figure~\fref{fig-args-top-level}.
+In this chapter, develop the arguments introduced in \mysec{sec-argsbackground}. Make use of argument maps to break down
+and visualize the debate, as shown in \myfig{fig-args-top-level}.
 

### 2020-03-01 Thesis-Chapter2-Background.tex
Add \my{chap/sec/fig/tab} commands
Add some new sources

@@
 
-These arguments are introduced here and analyzed in detail in Chapter \ref{chap-arguments}.
+These arguments are introduced here and analyzed in detail in \mychap{chap-arguments}.
 
@@
     \item Reject EA until technical capabilities are ready \cite{varia_2018}
-    \item Be forthcoming with data so that we know what to work on \cite{armerding_2017}
+    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
     \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}

### 2020-03-01 scripts/new-source.sh

@@ -54,7 +54,7 @@ fi
 
 ##################################################################### DATA FETCHING
 FORMAT=bibtex
-CURL_TIMEOUT=5
+CURL_TIMEOUT=20
 
 function bad_json() {
     [[ 0 == $# ]] || [[ $(echo $@ | grep -v "{") ]]

### 2020-03-15 Thesis-Chapter1-Introduction.tex
Improve file naming scheme, work on background
Update background; add words to folder dictionary

--- 2020-02-26 Thesis-Chapter-Intro.tex
@@
 \chapter{Introduction}
+\label{chap-introduction}
 
@@
 \begin{itemize}
-    \item In the pursuit of data privacy, legislative action is part of the threat model
-    \item Legislative action is a highly likely
-    \item We as the technical community need to be prepared
+    \item In the pursuit of data privacy, legislative action is part of the threat model--and action is likely
+    \item If the technical community acts, we can not only be prepared for regulation, but lead and shape it
 \end{itemize}

### 2020-03-15 Thesis-Chapter2-Background.tex
Update background; add words to folder dictionary

@@
 network security. [Properties: confidentiality, authentication, integrity, non-repudiation; primitives: hashed
-passwords, digital signatures, ...; technologies: PKI, TLS, SSO/Kerberos, VPNs, disk encryption, blockchains (might get
-eye-rolls).]
+passwords, digital signatures, ...; technologies: PKI, TLS, SSO/Kerberos, VPNs, disk encryption, blockchains, ...]
 
@@
 These arguments are introduced here and analyzed in detail in \mychap{chap-arguments}.
+[This section is currently written in outline-format, but will be drafted in a narrative, introductory style, as these
+arguments are the focus of a later chapter.]
 
@@
 
-\section{Technical State of the Art}
+\section{Regulatory Environment}
+
+\subsection{US Efforts to Regulate Encryption}
+
+\subsection{Regulations around the World}
+
+\section{Technical Approaches}
 
@@
 \end{itemize}
-
-\section{Paths Ahead}
-
-\subsection{Legislative Priorities}
-
-\begin{itemize}
-    \item Reject EA until technical capabilities are ready \cite{varia_2018}
-    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
-    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
-    \item Fund the research \cite{varia_2018}
-\end{itemize}
-
-\subsection{Technical Research}
-
-\begin{itemize}
-    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
-    \item Do the research
-    \item Focus on principles \cite{levy_robinson_2018}
-    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
-            \cite{abelson_2015}), and clearly define their interfaces with principles
-            \cite{matyas_incommensurability_2018}
-\end{itemize}

### 2020-03-15 Thesis-Chapter5-PathsAhead.tex
Update background; add words to folder dictionary

--- 2020-02-20 Thesis-Chapter-Background.tex
@@
-\chapter{Background}
+\chapter{Paths Ahead}
+\label{chap-pathsahead}
 
-\section{Arguments For and Against}
+In this chapter, describe actions to take in discussion, regulation, and research.
 
-\subsection{Arguments For}
+\section{Legislative Priorities}
 
-\subsection{Arguments Against}
+\begin{itemize}
+    \item Reject EA until technical capabilities are ready \cite{varia_2018}
+    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
+    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
+    \item Fund the research \cite{varia_2018}
+\end{itemize}
 
-\section{Technical State of the Art}
+\section{Technical Research}
 
-\subsection{Historical Approaches}
-
-\subsection{Modern Proposals}
-
-\section{Paths Ahead}
-
-\subsection{Legislative Priorities}
-
-\subsection{Technical Research}
+\begin{itemize}
+    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
+    \item Do the research
+    \item Focus on principles \cite{levy_robinson_2018}
+    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
+            \cite{abelson_2015}), and clearly define their interfaces with principles
+            \cite{matyas_incommensurability_2018}
+\end{itemize}

### 2020-03-18 Thesis-Chapter1-Introduction.tex
Refactor background and intro

@@
 \begin{itemize}
-    \item In the pursuit of data privacy, legislative action is part of the threat model--and action is likely
+    \item In the pursuit of data privacy, regulatory action is part of the threat model--and action is likely
     \item If the technical community acts, we can not only be prepared for regulation, but lead and shape it
@@
     \item Transparency to data access is critical for public trust
+    % \item In pluralistic societies, government is obligated to respect and balance opposing viewpoints
 \end{itemize}
@@
 
-\begin{itemize}
-    \item Argument maps formalizing the debate
-    \item A broad threat model describing the environment
-    \item A structure for organizing future progress
-\end{itemize}
+The paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
+brief history of encryption regulation, technology, and conflict, it introduces argument maps and threat models using
+data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and creates a
+multi-level threat model, broken down by the significantly different challenges posed by data at rest and data in
+motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
+and threats.
+
+% \begin{itemize}
+%     \item Argument maps formalizing the debate
+%     \item A broad threat model describing the environment
+%     \item A structure for organizing future progress
+% \end{itemize}

### 2020-03-18 Thesis-Chapter2-Background.tex
Refactor background and intro

@@
 \label{chap-background}
+
+This chapter summarizes the history of encryption and the exceptional access debate, introduces argument maps, and
+introduces threat modeling with data flow diagrams.
+
+\section{A Brief History of Encryption}
+\label{sec-history}
 
@@
 
-\section{Exceptional Access: Arguments For and Against}
-\label{sec-argsbackground}
+\subsection{Regulatory Environment}
+\label{sec-reg-environment}
 
-The debate over exceptional access takes place on two fundamental levels:
-\begin{enumerate}
-    \item Whether cryptosystems with exceptional access represent a net benefit to society
-    \item Whether and how to conduct exceptional access research
-\end{enumerate}
+\subsubsection*{US Efforts to Regulate Encryption}
 
-These arguments are introduced here and analyzed in detail in \mychap{chap-arguments}.
-[This section is currently written in outline-format, but will be drafted in a narrative, introductory style, as these
-arguments are the focus of a later chapter.]
+\subsubsection*{Regulations around the World}
 
-\subsection{Arguments For}
+\subsection{Technical Approaches}
+\label{sec-tech-approaches}
 
-Why EA is a benefit to society:
-\begin{itemize}
-    \item Absolute privacy is not a right
-    \item Encryption gives too much power to the individual
-    \item EA would allow more crimes to be prevented and prosecuted
-\end{itemize}
-
-Why to research EA:
-\begin{itemize}
-    \item EA done well would be more secure and privacy preserving than EA done poorly
-    \item The threat of legislative action creates urgency
-\end{itemize}
-
-\subsection{Arguments Against}
-
-Why EA is a danger to society:
-\begin{itemize}
-    \item Privacy is an absolute right
-    \item Exceptional access gives too much power to the state
-    \item Exceptional access is not possible in an acceptably secure manner
-\end{itemize}
-
-Why not to research EA:
-\begin{itemize}
-    \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
-            them
-    \item The technical community should not entertain delusions about what is truly possible
-\end{itemize}
-
-\section{Regulatory Environment}
-
-\subsection{US Efforts to Regulate Encryption}
-
-\subsection{Regulations around the World}
-
-\section{Technical Approaches}
-
-\subsection{Historical Approaches}
+\subsubsection*{Historical Approaches}
 
@@
 
-\subsection{Modern Proposals}
+\subsubsection*{Modern Proposals}
 
@@
 \end{itemize}
+
+\section{Argument Maps}
+\label{sec-arg-maps-intro}
+
+\section{Threat Models and Data Flow Diagrams}
+\label{sec-threat-model-intro}

### 2020-03-18 Thesis-Chapter3-Arguments.tex
Add some new sources
Refactor background and intro

@@
 
-In this chapter, develop the arguments introduced in \mysec{sec-argsbackground}. Make use of argument maps to break down
-and visualize the debate, as shown in \myfig{fig-args-top-level}.
+In this chapter, develop the arguments employed in EA debates. Make use of argument maps to break down and visualize the
+debate, as shown in \myfig{fig-args-top-level}.
 
-% \includepdf[]{arguments/build/top-level.pdf}
 \begin{figure}[h]
@@
 \end{figure}
+
+The debate over exceptional access takes place on two fundamental levels:
+\begin{enumerate}
+    \item Whether cryptosystems with exceptional access represent a net benefit to society
+    \item Whether and how to conduct exceptional access research
+\end{enumerate}
+
+\subsection{Arguments For}
+
+Why EA is a benefit to society:
+\begin{itemize}
+    \item Absolute privacy is not a right
+    \item Encryption gives too much power to the individual
+    \item EA would allow more crimes to be prevented and prosecuted
+\end{itemize}
+
+Why to research EA:
+\begin{itemize}
+    \item EA done well would be more secure and privacy preserving than EA done poorly
+    \item The threat of legislative action creates urgency
+\end{itemize}
+
+\subsection{Arguments Against}
+
+Why EA is a danger to society:
+\begin{itemize}
+    \item Privacy is an absolute right
+    \item Exceptional access gives too much power to the state
+    \item Exceptional access is not possible in an acceptably secure manner
+\end{itemize}
+
+Why not to research EA:
+\begin{itemize}
+    \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
+            them
+    \item The technical community should not entertain delusions about what is truly possible
+\end{itemize}

### 2020-03-25 scripts/show-source.sh
Scripts: show prioitizes bibtex keys

@@ -6,11 +6,18 @@ ORIG_DIR=$(pwd)
 cd $READING_DIR
 
 if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
+    # Show all
     eval "grep -riIsl \"${@:2}\""
     exit 0
 elif [[ 1 == $# && -f $ORIG_DIR/$1 ]]; then
+    # Show a file based on path
     FILE=$ORIG_DIR/$1
+elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
+    # Show a file based on bibtex ref
+    FILE="$(grep -riIsl "@*{$1," . | head -1)"
+    FILE="$(readlink -f "$FILE")"
 else
+    # Show the file with the most string matches
     FILE="$(grep -riIsc "$@" . | sort -t: -n -k2 | tail -1 | cut -d: -f1)"
     FILE="$(readlink -f "$FILE")"
 fi

### 2020-03-25 Thesis-Chapter2-Background.tex
Add background on crypto wars

@@
 
-This chapter summarizes the history of encryption and the exceptional access debate, introduces argument maps, and
-introduces threat modeling with data flow diagrams.
+This chapter summarizes the history of encryption and the exceptional access debate, introduces argument maps as a
+logical modeling tool, and introduces threat modeling with data flow diagrams.
 
@@
 
-Private use of encryption has a contentious history. [Some highlights.]
+Electronic encryption is a new technology, but the field of cryptography is old, as demand for privacy is as old as
+communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
+the character transposition technique Julius Caesar used to protect private correspondence
+\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
+in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
+computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
+analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
+research on public key cryptography in 1976 \cite{diffie_new_1976}.
 
-Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
-human rights. [This is where I'd like to understand more of the legal side. With a historical perspective.]
+Public key cryptography, paired with computer networking and personal computing advances in the 1980s, for the first
+time put the power of strong encryption in the hands of ordinary people. However, the privacy provided by cryptography
+is a powerful tool, and privacy is a property governments strive to regulate. The use of encryption for for private
+purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason based on evidence from
+decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force testimony from his secretary on
+the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it
+inherently uncomfortable with the rapid increase in availability of strong encryption.
 
-Privacy is not encryption's only purpose, however. Encryption is the technical foundation for many forms of computer and
-network security. [Properties: confidentiality, authentication, integrity, non-repudiation; primitives: hashed
-passwords, digital signatures, ...; technologies: PKI, TLS, SSO/Kerberos, VPNs, disk encryption, blockchains, ...]
+% Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
+% human rights. [This is where I'd like to understand more of the legal side. With a historical perspective.]
 
-Electronic encryption's dual role as enabler of perfect privacy and cornerstone of computer security may be the central
+Privacy is not encryption's only purpose, however.
+
+% Spoofing                  Authentication
+% Tampering                 Integrity
+% Repudiation               Non- Repudiation
+% Information Disclosure    Confidentiality
+% Denial of Service         Availability
+% Elevation of Privilege    Authorization
+
+Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
+often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
+and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
+from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
+
+Cryptographic primitives include symmetric key encryption, public key encryption, one-way hash functions, and digital
+signatures. A few technologies that directly rely on these primitives include TLS (transport layer security), PKI
+(public key infrastructure), SSO (single sign on, e.g. Kerberos), VPNs (virtual private networks), secure password
+storage, disk encryption tools, and secure- and trusted-boot processes. A few systems that directly rely on these
+technologies include the financial system, local and global; industrial control systems and critical infrastructure;
+health information systems; private, government, and military communications; and the integrity of the internet as a
+whole.
+
+If the cryptographic primitives that compose these systems were to be compromised, the systems they compose could
+collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
+infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
+whole loses integrity, inconceivably large damage would occur.
+
+% Let us break security down into three scales: primitives, technologies, and systems. A primitive is a conceptual
+% building block that serves a single technical function. A technology is a set of implemented primitives glued together
+% to provide useful features. A system is a collection of technologies that collectively serve the ultimate outcome.
+% The security properties critical at the systems level depend
+
+% Primitives: one-way hashes enable integrity checks, block ciphers enable efficient confidential data storage, and
+% digital signature enable non-repudiation
+
+% Technologies: Transport Layer Security (TLS) enables confidentiality and integrity over the internet, Kerberos enables
+% convenient authentication, and virtual private networks (VPNs) enable confidential and authenticated remote network
+% connections
+
+% Systems: ...
+
+Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
 conflict in the exceptional access debate. Those who focus on its role in privacy object to its absolute power. Those
@@
 % Encryption requires absolute cryptographic fidelity in order to provide security.
+
+All these realities--government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
+and a blossoming computer industry foundationally reliant on cryptography--came to a head in what has become known as
+the First Crypto War.
+
+In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passes
+the Arms Export Control Act and declares that strong cryptography is subject to export controls \cite{kehl_right_2015}.
+In 1991, the U.S. Senate introduces, but does not pass, a bill mandating access to plain text contents when authorized
+by law; in response, Philip Zimmermann releases PGP (Pretty Good Privacy), a public key encryption tool to secure email,
+in order that strong cryptography be ``made available to the American public before it became illegal to use''
+\cite{zimmermann_1996}. In 1993, the Clinton administration introduces Clipper Chip with the goal of ``providing the
+public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
+access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
+in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
+reacted negatively to the chip \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
+discovered flaws that allowed users to subvert the Clipper Chip mechanisms \cite{blaze_protocol_1994}, the proposal
+died.
+
+Despite the failure of the Clipper Chip, the debate over access to strong encryption continued, primarily focused on
+various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
+it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryptions software \cite{burns_s1726_1996}.
+In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
+activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have
+barred the government from creating a mandatory key escrow system and would also have removed export restrictions on
+most generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of
+PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
+Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. The the Pro-C0DE Act passed, but the SAFE Act did not.
+
+From 1996 to 1999, the SAFE Act was proposed several times, had hearings, and gained support \cite{kehl_right_2015}. By
+this time, there was "an overwhelming amount of evidence against moving ahead with any key escrow schemes"
+\cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all SAFE Act
+proposals \cite{kehl_right_2015}. This development marks the end of the First Crypto War.
+
+Two developments sewed the seeds for the Second Crypto War, and three set if off. The first seed was actually a
+non-development: when the Clinton administration changed its encryption policy, the U.S. House dropped the SAFE Act
+\cite{kehl_right_2015}. This meant that rather than being written in the pen of law, encryption policy was written in
+the pencil of executive order. The second seed was the events of and reaction to morning of September 11, 2001.
+
+Three developments have triggered the Second Crypto War.
+(1) The 2013 Snowden revelations
+(2) The mid-2010s spread of default device encryption in smartphones
+(3) The late-2010s spread of default E2E encryption in messaging services
 
@@
 \begin{itemize}
-    \item Clipper chip
-    \item Government key escrow
+    \item Clipper chip/third-party key escrow \cite{blaze_protocol_1994}
     \item Oblivious key escrow \cite{goos_oblivious_1996}
@@
     \item Computational, blockchain approaches \cite{phan_key_2017}
-    \item Lawful hacking \cite{nguyen_lawful_2017} \cite{soesanto_2018}
-    \item Forced password disclosure \cite{bittenbender_2019}, see \cite{kerr_encryption_2017}
+    \item Lawful hacking \cite{nguyen_lawful_2017} \cite{soesanto_2018} \cite{kerr_encryption_2017}
+    \item Forced password disclosure \cite{bittenbender_2019} \cite{kerr_encryption_2017}
+    \item Alternative paths to plaintext (access device while unlocked; find copies elsewhere)
+        \cite{kerr_encryption_2017}
 \end{itemize}

### 2020-03-25 scripts/show-source.sh
Scripts: if no matches, show says so and exits

@@ -18,7 +18,13 @@ elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
     FILE="$(readlink -f "$FILE")"
 else
     # Show the file with the most string matches
-    FILE="$(grep -riIsc "$@" . | sort -t: -n -k2 | tail -1 | cut -d: -f1)"
+    FILE_W_NUM="$(grep -riIsc "$@" . | sort -t: -n -k2 | tail -1)"
+    NUM="$(echo $FILE_W_NUM | cut -d: -f2)"
+    if (( $NUM == 0 )); then
+        echo "No matches"
+        exit 1
+    fi
+    FILE="$(echo $FILE_W_NUM | cut -d: -f1)"
     FILE="$(readlink -f "$FILE")"
 fi
 

### 2020-03-25 Thesis-Chapter1-Introduction.tex
Add to background

@@
 The paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief history of encryption regulation, technology, and conflict, it introduces argument maps and threat models using
-data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and creates a
-multi-level threat model, broken down by the significantly different challenges posed by data at rest and data in
-motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
+brief U.S.-focused history of encryption regulation, technology, and conflict, it introduces argument maps and threat
+models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
+creates a multi-level threat model, broken down by the significantly different challenges posed by data at rest and data
+in motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
 and threats.

### 2020-03-25 Thesis-Chapter2-Background.tex
Add to background

@@
 
-\section{A Brief History of Encryption}
+\section{Historical Context}
 \label{sec-history}
+
+\subsection{Encryption History and the First Crypto War}
 
@@
 
-If the cryptographic primitives that compose these systems were to be compromised, the systems they compose could
+If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
 collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
@@
 
-In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passes
+In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
 the Arms Export Control Act and declares that strong cryptography is subject to export controls \cite{kehl_right_2015}.
-In 1991, the U.S. Senate introduces, but does not pass, a bill mandating access to plain text contents when authorized
-by law; in response, Philip Zimmermann releases PGP (Pretty Good Privacy), a public key encryption tool to secure email,
-in order that strong cryptography be ``made available to the American public before it became illegal to use''
-\cite{zimmermann_1996}. In 1993, the Clinton administration introduces Clipper Chip with the goal of ``providing the
-public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
-access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
-in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
-reacted negatively to the chip \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
-discovered flaws that allowed users to subvert the Clipper Chip mechanisms \cite{blaze_protocol_1994}, the proposal
-died.
+In 1991, the U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by
+law; in response, Philip Zimmermann released PGP (Pretty Good Privacy), a public key encryption tool to secure email, in
+order that strong cryptography be ``made available to the American public before it became illegal to use''
+\cite{zimmermann_1996}. In 1993, the Clinton administration introduced Clipper Chip \cite{press_1993} with the goal of
+``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement and
+intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
+foundational role of cryptography in security--and the potential of human rights abuses for compromised
+privacy--industry and prominent technical leaders reacted negatively to the chip \cite{kehl_right_2015}
+\cite{zimmermann_1996}. When security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper
+Chip mechanisms \cite{blaze_protocol_1994}, the proposal died.
 
@@
 
-Two developments sewed the seeds for the Second Crypto War, and three set if off. The first seed was actually a
-non-development: when the Clinton administration changed its encryption policy, the U.S. House dropped the SAFE Act
-\cite{kehl_right_2015}. This meant that rather than being written in the pen of law, encryption policy was written in
-the pencil of executive order. The second seed was the events of and reaction to morning of September 11, 2001.
+\subsection{The Second Crypto War and the Current Context}
+
+Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
+administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. This meant
+that rather than being written in the pen of law, encryption policy was written in the pencil of executive order. The
+second development was the events of and reaction to morning of September 11, 2001.
 
 Three developments have triggered the Second Crypto War.
-(1) The 2013 Snowden revelations
-(2) The mid-2010s spread of default device encryption in smartphones
-(3) The late-2010s spread of default E2E encryption in messaging services
+\begin{enumerate}
+    \item The 2013 Snowden revelations
+    \item The mid-2010s spread of default device encryption in smartphones
+    \item The late-2010s spread of default E2E encryption in messaging services
+\end{enumerate}
+
+Characteristic elements of the Second Crypto War include:
+\begin{enumerate}
+    \item The Apple v. FBI case over access to the San Bernardino shooter's iPhone \cite{schulze_clipper_2017}
+    \item A return to the Four Horsemen \cite{schneier_scaring_2019}, with a (warranted \cite{keller_internet_2019})
+            focus currently on CSAM \cite{geller_2019}
+    \item "Security vs Security" vs "Security vs Privacy" \cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}
+    \item Pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}
+    \item A political landscape marked by uncertainty, extremism, misinformation, and most recently, a pandemic that
+            will have unknown impacts
+\end{enumerate}
+
+That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
+U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
+like ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020, the
+Republican atop the Senate Judiciary Committee introduced S.3398, the EARN IT Act of 2020, which seeks to establish a
+16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
+committee determines the platform is not doing enough to fight child sexual abuse material (CSAM), a legal synonym for
+child pornography \cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
+\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
+largely see the EARN IT Act as an attempt to indirectly yet effectively outlaw E2E \cite{newman_2020}
+\cite{pfefferkorn_2020}.
+
+% TODO: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
+
+The following sections summarize the encryption regulatory environment and technical approaches to exceptional access.
 
@@
 
+The following laws and regulations have been applied to cases involving cryptography and access to encrypted
+information:
+
+\begin{itemize}
+    \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
+    \item The Arms Export Control Act (AECA) of 1976 established export controls over strong encryption techniques
+        introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
+    \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
+    \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
+    \item \dots
+\end{itemize}
+
+The following are proposed laws and regulations that did not go into effect or have yet to be voted on:
+
+\begin{itemize}
+    \item The Clipper Chip \cite{press_1993} \cite{thompson_2015}
+    \item The SAFE Act \cite{goodlatte_hr3011_1996}
+    \item The EARN IT Act \cite{graham_s3398_2020}
+    \item \dots
+\end{itemize}
+
 \subsubsection*{Regulations around the World}
+
+The following is a very brief account of the regulatory environment in various jurisdictions around the world:
+
+\begin{itemize}
+    \item Australia: in 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
+    \item UK: Investigatory Powers Act \cite{legislature_2016}
+    \item EU: GDPR
+    \item China: \cite{donahue_comparative_2018}
+    \item Israel: \cite{donahue_comparative_2018}
+    \item Japan, Korea? -- can't find anything
+    \item \dots
+\end{itemize}
 

### 2020-03-26 Thesis-Chapter2-Background.tex
Add more sources and notes on laws

@@
 foundational role of cryptography in security--and the potential of human rights abuses for compromised
-privacy--industry and prominent technical leaders reacted negatively to the chip \cite{kehl_right_2015}
+privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
 \cite{zimmermann_1996}. When security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper
@@
 From 1996 to 1999, the SAFE Act was proposed several times, had hearings, and gained support \cite{kehl_right_2015}. By
-this time, there was "an overwhelming amount of evidence against moving ahead with any key escrow schemes"
+this time, there was ``an overwhelming amount of evidence against moving ahead with any key escrow schemes''
 \cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all SAFE Act
@@
             focus currently on CSAM \cite{geller_2019}
-    \item "Security vs Security" vs "Security vs Privacy" \cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}
+    \item ``Security vs Security'' vs ``Security vs Privacy'' \cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}
     \item Pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}
@@
 \begin{itemize}
+    \item The All Writs Act \cite{congress_1789}
     \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}

### 2020-03-28 scripts/show-source.sh
Fix 'show' script

@@ -18,7 +18,7 @@ elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
     FILE="$(readlink -f "$FILE")"
 else
     # Show the file with the most string matches
-    FILE_W_NUM="$(grep -riIsc "$@" . | sort -t: -n -k2 | tail -1)"
+    FILE_W_NUM=$(eval "grep -riIsc \"$@\" . | sort -t: -n -k2 | tail -1")
     NUM="$(echo $FILE_W_NUM | cut -d: -f2)"
     if (( $NUM == 0 )); then
         echo "No matches"

### 2020-03-28 scripts/count-citations.sh
Update scripts: add 'cite', export 'exitprint()'

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+cd $(git rev-parse --show-toplevel)
+source ./scripts/sourceme.sh
+# set -x
+
+LIGHT_GREEN='\e[1;32m'
+LIGHT_CYAN='\e[1;36m'
+NO_COLOR='\e[0m'
+
+function print_number() {
+    echo $(grep $1 *.tex | wc -l)
+}
+
+function print_count() {
+    COUNT=$(print_number $1)
+    echo -e "$LIGHT_GREEN$1$NO_COLOR is cited $COUNT time$(((1 == COUNT)) || echo s)."
+}
+
+if (( 2 <= $# )) && [[ "-c" == "$1" ]]; then
+    # Print count
+    print_count $2
+elif (( 2 <= $# )) && [[ "-n" == "$1" ]]; then
+    # Print number only
+    print_number $2
+else
+    print_count $1
+    echo
+    grep --color=auto -n $1 *.tex
+fi

### 2020-03-28 scripts/new-source.sh

@@ -19,13 +19,8 @@ function cleanup() {
 }
 trap cleanup EXIT
 
-function printexit() {
-    echo "${@:2}"
-    exit $1
-}
-
 ##################################################################### DEPENDENCIES
-[[ 0 < $(which docker | wc -l) ]] || printexit 1 "Docker must be installed."
+[[ 0 < $(which docker | wc -l) ]] || exitprint 1 "Docker must be installed."
 if [[ ! $(which getdoi) ]]; then
     echo "Installing scholarref tools..."
     git clone git://src.adamsgaard.dk/scholarref ../../scholarref &>/dev/null
@@ -35,12 +30,12 @@ if [[ ! $(which getdoi) ]]; then
 fi
 
 ##################################################################### ARGUMENTS
-(( 2 <= $# )) || printexit 2 "Incorrect args"
+(( 2 <= $# )) || exitprint 2 "Incorrect args"
 case "$1" in
     /*) SUBDIR=$(echo $1 | sed 's:/*$::') ;;
     *)  SUBDIR=$ORIG_DIR/$(echo $1 | sed 's:/*$::') ;;
 esac
-[ -d $SUBDIR ] || printexit 2 "SUBDIR=$SUBDIR must exist"
+[ -d $SUBDIR ] || exitprint 2 "SUBDIR=$SUBDIR must exist"
 SOURCE="${@:2}"
 
 ##################################################################### ZOTERO CONTAINER
@@ -76,18 +71,18 @@ if bad_json $JSON; then
                     127.0.0.1:$DOCKER_PORT/search 2>/dev/null)
 fi
 if bad_json $JSON; then
-    printexit 1 "Could not identify data source from $SOURCE"
+    exitprint 1 "Could not identify data source from $SOURCE"
 fi
 BIBTEX_CITATION=$(curl -v -d "$JSON" -H "Content-Type: application/json" \
                     "127.0.0.1:$DOCKER_PORT/export?format=$FORMAT" 2>/dev/null)
-[[ "" != "$BIBTEX_CITATION" ]] || printexit 1 "Could not extract $FORMAT citation from $SOURCE"
+[[ "" != "$BIBTEX_CITATION" ]] || exitprint 1 "Could not extract $FORMAT citation from $SOURCE"
 
 # Remove empty lines
 BIBTEX_CITATION="$(echo "$BIBTEX_CITATION" | sed '/^[[:space:]]*$/d')"
 
[…]

### 2020-03-28 scripts/show-source.sh

@@ -37,6 +37,7 @@ NAME="$(cat "$FILE" | head -1 | sed 's/# //')"
 AUTHOR="$(cat "$FILE" | grep -A1 "\sauthor =" | tr '\n' ' ' | cut -d{ -f2 | cut -d} -f1)"
 REF="$(cat "$FILE" | grep -A1 "\`bib" | tr '\n' ' ' | cut -d{ -f2 | cut -d, -f1)"
 URL=$(cat "$FILE" | grep -A1 "\surl =" | tr '\n' ' ' | cut -d{ -f2 | cut -d} -f1)
+CIT_COUNT=$($SCRIPTS_DIR/count-citations.sh -n $REF)
 
 YELLOW='\e[0;33m'
 LIGHT_GREEN='\e[1;32m'
@@ -51,5 +52,5 @@ echo
 echo -e "$BLUE$URL$NO_COLOR"
 echo -e "$FILE"
 echo
-echo -e "cite as $LIGHT_GREEN$REF$NO_COLOR"
+echo -e "cite as $LIGHT_GREEN$REF$NO_COLOR (cited in $CIT_COUNT place$(((1 == CIT_COUNT)) || echo s))"
 echo

### 2020-03-28 scripts/sourceme.sh

@@ -3,6 +3,13 @@
 export READING_DIR="$(git rev-parse --show-toplevel)/reading"
 export SCRIPTS_DIR="$(git rev-parse --show-toplevel)/scripts"
 
+function exitprint() {
+    echo "${@:2}"
+    exit $1
+}
+export -f exitprint
+
 alias new="$SCRIPTS_DIR/new-source.sh"
 alias show="$SCRIPTS_DIR/show-source.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
+alias cite="$SCRIPTS_DIR/count-citations.sh"

### 2020-03-28 scripts/count-citations.sh
Scripts: move colors to common, improve 'cite'

@@ -4,17 +4,13 @@ cd $(git rev-parse --show-toplevel)
 source ./scripts/sourceme.sh
 # set -x
 
-LIGHT_GREEN='\e[1;32m'
-LIGHT_CYAN='\e[1;36m'
-NO_COLOR='\e[0m'
-
 function print_number() {
-    echo $(grep $1 *.tex | wc -l)
+    echo $(grep "\\cite{$1}" *.tex | wc -l)
 }
 
 function print_count() {
     COUNT=$(print_number $1)
-    echo -e "$LIGHT_GREEN$1$NO_COLOR is cited $COUNT time$(((1 == COUNT)) || echo s)."
+    echo -e "$LIGHT_CYAN$1$NO_COLOR is cited $COUNT time$(((1 == COUNT)) || echo s)."
 }
 
 if (( 2 <= $# )) && [[ "-c" == "$1" ]]; then
@@ -26,5 +22,5 @@ elif (( 2 <= $# )) && [[ "-n" == "$1" ]]; then
 else
     print_count $1
     echo
-    grep --color=auto -n $1 *.tex
+    grep --color=auto -n "\\cite{$1}" *.tex
 fi

### 2020-03-28 scripts/show-source.sh

@@ -38,12 +38,10 @@ AUTHOR="$(cat "$FILE" | grep -A1 "\sauthor =" | tr '\n' ' ' | cut -d{ -f2 | cut 
 REF="$(cat "$FILE" | grep -A1 "\`bib" | tr '\n' ' ' | cut -d{ -f2 | cut -d, -f1)"
 URL=$(cat "$FILE" | grep -A1 "\surl =" | tr '\n' ' ' | cut -d{ -f2 | cut -d} -f1)
 CIT_COUNT=$($SCRIPTS_DIR/count-citations.sh -n $REF)
-
-YELLOW='\e[0;33m'
-LIGHT_GREEN='\e[1;32m'
-LIGHT_CYAN='\e[1;36m'
-BLUE='\e[0;34m'
-NO_COLOR='\e[0m'
+CIT_MESSAGE=""
+if ((0 < CIT_COUNT)); then
+    CIT_MESSAGE="(cited in $CIT_COUNT place$(((1 == CIT_COUNT)) || echo s))"
+fi
 
 echo
 echo -e "$LIGHT_CYAN$NAME$NO_COLOR"
@@ -52,5 +50,5 @@ echo
 echo -e "$BLUE$URL$NO_COLOR"
 echo -e "$FILE"
 echo
-echo -e "cite as $LIGHT_GREEN$REF$NO_COLOR (cited in $CIT_COUNT place$(((1 == CIT_COUNT)) || echo s))"
+echo -e "cite as $LIGHT_GREEN$REF$NO_COLOR $CIT_MESSAGE"
 echo

### 2020-03-28 scripts/sourceme.sh

@@ -3,6 +3,12 @@
 export READING_DIR="$(git rev-parse --show-toplevel)/reading"
 export SCRIPTS_DIR="$(git rev-parse --show-toplevel)/scripts"
 
+export YELLOW='\e[0;33m'
+export LIGHT_GREEN='\e[1;32m'
+export LIGHT_CYAN='\e[1;36m'
+export BLUE='\e[0;34m'
+export NO_COLOR='\e[0m'
+
 function exitprint() {
     echo "${@:2}"
     exit $1

### 2020-03-28 Thesis-Chapter2-Background.tex
Add reading sources; update background

@@
 that rather than being written in the pen of law, encryption policy was written in the pencil of executive order. The
-second development was the events of and reaction to morning of September 11, 2001.
+second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
+\cite{shamsi_2011}, \cite{bloss_escalating_2007}.)
 
@@
 \begin{enumerate}
-    \item The 2013 Snowden revelations
-    \item The mid-2010s spread of default device encryption in smartphones
-    \item The late-2010s spread of default E2E encryption in messaging services
+    \item The 2013 Snowden revelations \cite{thorsen_cryptic_2017} \cite{treguer_us_2018}
+    \item The mid-2010s spread of default device encryption in smartphones \cite{treguer_us_2018}
+    \item The late-2010s spread of default E2E encryption in messaging services \cite{treguer_us_2018}
 \end{enumerate}
@@
 
-The following laws and regulations have been applied to cases involving cryptography and access to encrypted
-information:
+The following laws and regulations have been applied to cases involving surveillance, cryptography, and access to
+encrypted information:
 
@@
         introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
+    \item Foreign Intelligence Surveillance (FISA) Act \cite{rodino_1978}
     \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
+    \item Electronic Communications Privacy Act (ECPA) of 1986 \cite{kastenmeier_hr4952_1986}
     \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
+    \item Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism
+        (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
+    \item FISA Amendments Act of 2008 \cite{reyes_hr6304_2008}
+    \item Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring
+        (USA FREEDOM) Act of 2015 \cite{sensenbrenner_2015}
     \item \dots
@@
     \item Japan, Korea? -- can't find anything
-    \item \dots
+    \item \dots General, see \cite{budish_encryption_2018}
 \end{itemize}

### 2020-03-28 Thesis-Chapter5-PathsAhead.tex
Add reading sources; update background

@@
     \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
+        (e.g., ECPA became outdated because was too tied to specific technology \cite{shamsi_2011})
     \item Fund the research \cite{varia_2018}

### 2020-03-30 Thesis-Chapter2-Background.tex
Cite a few U.S. laws and attempted laws

@@
 various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
-it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryptions software \cite{burns_s1726_1996}.
+it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryption software \cite{burns_s1726_1996}.
 In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
@@
 PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
-Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. The the Pro-C0DE Act passed, but the SAFE Act did not.
+Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. 
+% TODO: "The the Pro-C0DE Act passed, but the SAFE Act did not." -- WRONG! Neither passed, this was also achieved by executive order!
 
@@
     \item The SAFE Act \cite{goodlatte_hr3011_1996}
+    \item Encrypted Communications Privacy Act (mix of SE protection and EA guidelines?) \cite{leahy_s376_1997}
+    \item Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act
+        (like above?) \cite{ashcroft_s2067_1998}
+    \item Encryption for the National Interest Act and Tax Relief for Responsible Encryption Act (introduced by same rep
+        on same day) \cite{goss_hr2616_1999} \cite{goss_hr2617_1999}
+    \item Compiance with Court Orders Act of 2016 (discussion draft) \cite{burr_2016}
+    \item The Secure Data Act (federal SE protections) \cite{lofgren_hr5823_2018}
+    \item The ENCRYPT Act (state SE protections) \cite{lieu_hr4170_2019}
     \item The EARN IT Act \cite{graham_s3398_2020}
@@
     \item UK: Investigatory Powers Act \cite{legislature_2016}
-    \item EU: GDPR
+    \item EU: GDPR \cite{parliament_2016}
     \item China: \cite{donahue_comparative_2018}

### 2020-04-01 Thesis-Chapter2-Background.tex
Finish section of first crypto war

@@
 PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
-Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. 
-% TODO: "The the Pro-C0DE Act passed, but the SAFE Act did not." -- WRONG! Neither passed, this was also achieved by executive order!
+Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. In November of that year, the Clinton administration
+released an executive order effectively removing export controls on strong encryption products, along the lines of the
+Pro-C0DE Act \cite{clinton_1996}.
 
@@
 Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
-administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. This meant
-that rather than being written in the pen of law, encryption policy was written in the pencil of executive order. The
-second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
+administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. Also un-passed
+were the related Pro-C0DE Act, Encrypted Communications Privacy Act (not to be confused with the \textit{Electronic}
+Communications Privacy Act (ECPA) of 1986), and Encryption Protects the Rights of Individuals from Violation and Abuse
+in CYberspace (E-PRIVACY) Act \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998}. This legislative failure meant that
+rather than being written in the pen of law, encryption policy was written in the pencil of executive order.
+
+The second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
 \cite{shamsi_2011}, \cite{bloss_escalating_2007}.)
@@
     \item Foreign Intelligence Surveillance (FISA) Act \cite{rodino_1978}
+        (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
+        % (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
+        %     https://www.washingtontimes.com/news/2020/mar/15/adam-schiff-fisa-reform-intervention-protects-fbi-/,
+        %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
     \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}

### 2020-04-01 scripts/bib-gen.sh
Scripts: sanitize language field when generating bib

@@ -14,13 +14,21 @@ function check_for_duplicate_refname() {
   fi
 }
 
+function sanitize_language_field() {
+  if [[ "$REF" =~ "language" ]]; then
+    for PATTERN in "n" "ng" "nglish" "n_GB" "n_US" "n\\\\_GB" "n\\\\_US" "n-GB" "n-US"; do
+      REF=${REF/\{[Ee]$PATTERN\}/\{english\}}
+    done
+  fi
+}
+
 echo "% This is an auto-generated bibliography file built by $(basename $0)" > $OUTFILE
-echo "" >> $OUTFILE
+echo >> $OUTFILE
 
 for FILE in $(find . -type f -name "*.md"); do
   ROWS=( $(grep -n '```' $FILE | cut -d: -f1) )
   REF="$(sed -n $(( ${ROWS[0]} + 1 )),$(( ${ROWS[1]} - 1 ))p $FILE)"
   check_for_duplicate_refname
+  sanitize_language_field
   echo "$REF" >> $OUTFILE
-  echo >> $OUTFILE
 done

### 2020-04-01 Thesis-Chapter2-Background.tex
Flesh out Second Crypto War section

@@
 \subsection{Encryption History and the First Crypto War}
+\label{sec-crypto-cwi}
 
@@
 \subsection{The Second Crypto War and the Current Context}
+\label{sec-cwii-context}
 
-Two developments sewed the seeds for the Second Crypto War. The first was actually a non-development: when the Clinton
+Two developments paved the way for the Second Crypto War. The first was actually a non-development: when the Clinton
 administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. Also un-passed
@@
 
-The second development was the events of and reaction to the morning of September 11, 2001. (resources on 9/11:
-\cite{shamsi_2011}, \cite{bloss_escalating_2007}.)
+The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
+been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
+of the September 11 attacks to some justified yet greater investigatory powers \cite{bloss_escalating_2007}. October of
+2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
+FISA requests on U.S. citizens; it expanded the scope of what could be compelled via FISA orders; it authorized
+``roving'' wiretaps; and it increased the power of National Security Letters (NSLs), which can be used without judicial
+review to compel information from electronic service providers while precluding any public release of the event
+\cite{sensenbrenner_2001} \cite{shamsi_2011}.
 
-Three developments have triggered the Second Crypto War.
-\begin{enumerate}
-    \item The 2013 Snowden revelations \cite{thorsen_cryptic_2017} \cite{treguer_us_2018}
-    \item The mid-2010s spread of default device encryption in smartphones \cite{treguer_us_2018}
-    \item The late-2010s spread of default E2E encryption in messaging services \cite{treguer_us_2018}
-\end{enumerate}
+If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
+war, and increased government surveillance set events in motion toward conflict, it was the 2013 Snowden revelations
+that constituted the crossing of the threshold. Though policies such as the U.S. National Security Agency (NSA)'s
+warrantless wiretapping caused a stir, it was the mass data collection under the agency's PRISM and related programs
+that caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy,
+encryption promised a technical solution. Under pressure from both the human rights field and from ordinary users
+demanding improved digital privacy, U.S. tech companies responded by introducing default device encryption for
+data-at-rest (DAR) and end-to-end (E2E) encryption for data-in-motion (DIM) \cite{treguer_us_2018}.
 
-Characteristic elements of the Second Crypto War include:
-\begin{enumerate}
-    \item The Apple v. FBI case over access to the San Bernardino shooter's iPhone \cite{schulze_clipper_2017}
-    \item A return to the Four Horsemen \cite{schneier_scaring_2019}, with a (warranted \cite{keller_internet_2019})
-            focus currently on CSAM \cite{geller_2019}
-    \item ``Security vs Security'' vs ``Security vs Privacy'' \cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}
-    \item Pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}
-    \item A political landscape marked by uncertainty, extremism, misinformation, and most recently, a pandemic that
-            will have unknown impacts
-\end{enumerate}
+Strong encryption's transition into being caused the First Crypto War; likewise, strong encryption's transition into
+ubiquity is causing the Second Crypto War.
+
+The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
+their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
+basic digital security. Recall encryption and cryptography's critical role in security as described in
+\mysec{sec-crypto-cwi}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
+security is a central conflict in the exceptional access debate. When questioned by lawmakers, technical leaders and
+executives have repeatedly appealed to the fact that strong cryptography is a necessity for security
+\cite{schulze_clipper_2017}.
+
+Several events and arguments have come to characterize the Second Crypto War.
+
+\paragraph*{Apple vs. FBI} The Apple v. FBI case over access to the San Bernardino shooter's iPhone
+\cite{schulze_clipper_2017}.
+
+\paragraph*{The Four Horsemen} A return to the Four Horsemen \cite{schneier_scaring_2019}, with a warranted focus
+currently on child sexual abuse material \cite{keller_internet_2019} \cite{geller_2019}.
+
+\paragraph*{``Security vs Security''} ``Security vs Security'' vs ``Security vs Privacy''
+\cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}.
+
+\paragraph*{DIM vs. DAR} A pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}.
+
+\paragraph*{Volatile Politics} A political landscape marked by uncertainty, extremism, misinformation, and most
+recently, a pandemic that will have unknown impacts.
 
@@
 
+% TODO: keep up to date
 % TODO: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
@@
         (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
+    \item Protect America Act (PAA) of 2007 \cite{mcconnell_s1927_2007}
     \item FISA Amendments Act of 2008 \cite{reyes_hr6304_2008}

### 2020-04-02 Thesis-Chapter2-Background.tex
Expand on current context

@@
 Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
-conflict in the exceptional access debate. Those who focus on its role in privacy object to its absolute power. Those
+conflict in the EA debate. Those who focus on its role in privacy object to its absolute power. Those
 who focus on its role in security rely on its absolute power.
@@
 review to compel information from electronic service providers while precluding any public release of the event
-\cite{sensenbrenner_2001} \cite{shamsi_2011}.
+\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
+weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.
 
@@
 \mysec{sec-crypto-cwi}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
-security is a central conflict in the exceptional access debate. When questioned by lawmakers, technical leaders and
+security is a central conflict in the EA debate. When questioned by lawmakers, technical leaders and
 executives have repeatedly appealed to the fact that strong cryptography is a necessity for security
@@
 
-\paragraph*{Apple vs. FBI} The Apple v. FBI case over access to the San Bernardino shooter's iPhone
-\cite{schulze_clipper_2017}.
+\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as antagonistic,
+an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown that the
+debate is more about ``Security vs. Security'' than ``Security vs. Privacy'' \cite{stalla_bourdillon_privacy_2014}. In
+this context, there are two types of security: public security, the pursuit of law enforcement, and cybersecurity, the
+pursuit of tech professionals. Both sides of the debate have by now acknowledged that privacy and public security are
+not always in conflict, and in fact EA's effects on public security vs. cybersecurity--which in return affects public
+security--may be more important \cite{schneier_2019}.
+
+\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
+locked iPhone, resulted in the first major battle of the Second Crypto War. The FBI issued an order under the All Writs
+Act to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
+objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
+\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
+February 16, 2016, when the warrant was issued, to March 28, when the FBI announced they had gained access to the phone
+through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
+\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto wars,
+terrorism and device encryption. Shortly after the case ended, Senators Dianne Feinstein and Richard Burr released a
+discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the
+floor, and the unresolved nature of the court case left the legal debate wide open.
 
@@
 currently on child sexual abuse material \cite{keller_internet_2019} \cite{geller_2019}.
-
-\paragraph*{``Security vs Security''} ``Security vs Security'' vs ``Security vs Privacy''
-\cite{stalla_bourdillon_privacy_2014} \cite{schneier_2019}.
 

### 2020-04-04 Thesis-Glossary.tex
Add glossaries and glossaries-extra

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       GLOSSARY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Reference using the schema
%
%   \ac{}   -- general: abbreviation except first time is full
%   \acs{}  -- short: abbreviation
%   \acl{}  -- long: definition
%   \acf{}  -- full: definition (abbreviation)
%
%   pluralization       -- append "p", e.g. \acp{}
%   capitalize first    -- capitalize first of command, e.g. \Ac{}
%   capitalize all      -- capitalize all of command, e.g. \AC{}
%
%   See page 85 of the documentation:
%   http://ctan.math.utah.edu/ctan/tex-archive/macros/latex/contrib/glossaries-extra/glossaries-extra-manual.pdf

\makeglossaries

\newcommand{\newterm}[3]{
    \newglossaryentry{#1}{
        name={#2},
        description={#3}
    }
}
\newcommand{\term}{\ac}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          TERMS

% NOTE: Do not end with a period. The package adds this for you.
% TODO: auto-update glossary.md. This should be the master, because it is more fully specified

% Start terms
\newterm{clipperchip}{Clipper Chip}{
    A 1993 initiative by the Clinton administration to provide ``the public with strong cryptographic tools without
    sacrificing the ability of law enforcement and intelligence agencies to access unencrypted versions of those
    communications'' via hardware additions to consumer electronics. The proposal faced opposition, and died in 1994
    when security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper Chip mechanisms
}

\newterm{PRISM}{PRISM}{
    An NSA program unveiled by Edward Snowden involving the mass collection of  internet communications data of non-U.S.
    persons and those communicating with them
}

\newterm{horsemen}{The Four Horsemen}{
    Four reasons that governments and law enforcement agencies use to undercut public support for strong encryption. The
    four horsemen are terrorists, pedophiles (child pornographers), drug dealers (and traffickers), and money launderers
    (or kidnappers)
}

\newterm{horsemen-full}{The Four Horsemen of the Infocalypse}{
    The long name for ``The Four Horsemen''
}
% End terms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                       ACRONYMS

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{EA}{EA}{exceptional (or extraordinary) access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                    UN-OFFICIAL

\newcommand{\US}{U.S. } % Not good; doesn't handle spacing, periods at ends of sentences very well

### 2020-04-04 scripts/acronyms.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+set -eEuo pipefail
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+# set -x
+
+INFILE="$BASE_DIR/notes/glossary.md"
+
+function lineof_with_offset() {
+  echo $(( $(grep -n "$1" $INFILE | cut -d: -f1) + $2 ))
+}
+
+function cat_acronym_table() {
+  START_LINE=$(lineof_with_offset "<!-- start acronyms" 4)
+  END_LINE=$(lineof_with_offset "<!-- end acronyms" -2)
+
+  sed -n ${START_LINE},${END_LINE}p $INFILE
+}
+
+function to_lower() {
+  echo "$1" | tr '[:upper:]' '[:lower:]'
+}
+
+function convert_to_latex() {
+  TERM="$(echo $1 | cut -d\| -f 2 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
+  DEFN="$(echo $1 | cut -d\| -f 3 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
+
+  echo "\newabbreviation{$TERM}{$TERM}{$DEFN}"
+}
+
+while read -r ACR; do
+    convert_to_latex "$ACR"
+done <<< $(cat_acronym_table)

### 2020-04-04 scripts/sourceme.sh

@@ -1,7 +1,8 @@
 # Source this file
 
-export READING_DIR="$(git rev-parse --show-toplevel)/reading"
-export SCRIPTS_DIR="$(git rev-parse --show-toplevel)/scripts"
+export BASE_DIR="$(git rev-parse --show-toplevel)"
+export READING_DIR="$BASE_DIR/reading"
+export SCRIPTS_DIR="$BASE_DIR/scripts"
 
 export YELLOW='\e[0;33m'
 export LIGHT_GREEN='\e[1;32m'
@@ -19,3 +20,4 @@ alias new="$SCRIPTS_DIR/new-source.sh"
 alias show="$SCRIPTS_DIR/show-source.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
+alias todo="(cd $BASE_DIR && grep -n TODO *.tex)"

### 2020-04-04 scripts/update-acronyms.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+set -eEuo pipefail
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+#set -x
+
+TEX_FILE="$BASE_DIR/Thesis-Glossary.tex"
+
+function lineof_with_offset() {
+  echo $(( $(grep -n "$1" $TEX_FILE | cut -d: -f1) + $2 ))
+}
+
+START_LINE=$(lineof_with_offset "% Start acronyms" 0)
+END_LINE=$(lineof_with_offset "% End acronyms" 0)
+EOF_LINE=$(wc -l $TEX_FILE | awk '{print $1}')
+
+BEGINNING="$(sed -n 1,${START_LINE}p         $TEX_FILE)"
+END="$(      sed -n ${END_LINE},${EOF_LINE}p $TEX_FILE)"
+
+echo "$BEGINNING"         >  $TEX_FILE
+$SCRIPTS_DIR/acronyms.sh  >> $TEX_FILE
+echo "$END"               >> $TEX_FILE
+

### 2020-04-05 scripts/show-source.sh
Scripts: show -a yields absolute paths

@@ -7,7 +7,9 @@ cd $READING_DIR
 
 if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
     # Show all
-    eval "grep -riIsl \"${@:2}\""
+    for FILE in $(eval "grep -riIsl \"${@:2}\""); do
+        readlink -f $FILE
+    done
     exit 0
 elif [[ 1 == $# && -f $ORIG_DIR/$1 ]]; then
     # Show a file based on path

### 2020-04-05 Thesis-Chapter1-Introduction.tex
Flesh out Second Crypto War section
Add glossaries and glossaries-extra
Work on the introduction

@@
 
-Exceptional access is a wicked problem. Characterized by a dynamic technological environment, disagreement about
-underlying values, and resistance to a clear solution, the debate on exceptional access refuses to end. This paper will
-not attempt to end the debate, but to inform it with additional structure.
+Difficult policy matters are sometimes classified as wicked problems. Wicked problems are defined not by their moral
+nature, but by their vicious persistence. These problems have multiple causes and no clear solution; attempted solutions
+have unintended consequences, and the complex environment results in chronic policy failure
+\cite{commission_tackling_2018}.
+
+\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption reserved for law
+enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
+resistance to a clear solution, the debate on \acl{EA} refuses to end. This paper will not attempt to end the debate,
+but to inform it with additional structure.
 
 \section{Motivation}
+\label{sec-motivation}
 
-\begin{itemize}
-    \item In the pursuit of data privacy, regulatory action is part of the threat model--and action is likely
-    \item If the technical community acts, we can not only be prepared for regulation, but lead and shape it
-\end{itemize}
+% \begin{itemize}
+%     % TODO: Why strong encryption is important + Why some want exceptional access
+%     \item In the pursuit of data privacy, regulatory action is part of the threat model--and action is likely
+%     \item If the technical community acts, we can not only be prepared for regulation, but lead and shape it
+% \end{itemize}
+
+% TODO: clarify structure of argument in this whole section; it is a little hacked together and logically scattered
+
+The conflict at the heart of the debate is that the same cryptographic and design principles that underlie all digital
+security also enable an unprecedented degree of individual privacy. Encryption is a foundational tool to the integrity
+and confidentiality of all connected computing systems. Its increasing ubiquity in communications and storage provides
+clear benefits and should be welcomed by all. In a world where information security is frightfully poor yet increasingly
+important, strong encryption's necessity can not be understated. However, the privacy afforded by certain encryption
+technologies hampers law enforcement investigations and hides wrongdoing. In a society that cares about bringing
+dangerous people to justice, this risk should not be ignored.
+
+The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
+access encrypted data to perform their duties. Human rights groups and technical leaders counter that a weakened
+encryption environment would fatally compromise privacy and security as we know it. As the debate continues, sides have
+become entrenched, and both sides have engaged in disingenuous attacks against their perceived opponents.
+
+It is time that the technical community moves the debate forward. Increased regulation of the tech industry is
+inevitable. Despite strong encryption's benefits, regulatory interest in the subject has not subdued. The form that
+regulation will take in the field of encryption depends on the good faith efforts made to equitably balance all the
+benefits and risks involved in deploying a crypto system. If the technical community acts, we can not only be prepared
+for regulation, but lead and shape it.
+
+Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model.
 
 \section{Premises}
+\label{sec-premises}
 
-\begin{itemize}
-    \item The failure to convince legislative bodies and law enforcement agencies demands that the technical community continue the discussion
-    \item The ability to access digital data is important and will be increasingly important for investigations
-    \item Various forms of encryption are foundational all elements of online security
-    \item Mass surveillance is a grave danger that representative governments are not immune from
-    \item International and human rights considerations significantly constrain workable solutions
-    \item Transparency to data access is critical for public trust
-    % \item In pluralistic societies, government is obligated to respect and balance opposing viewpoints
-\end{itemize}
+This paper accepts and builds on the following premises, some of which were introduced in \mysec{sec-motivation}.
+
+% TODO: flesh out, probably add some references
+
+\paragraph*{The Discussion Must Go On} The failure of the technical community to convince legislative bodies and law
+enforcement agencies to stop pursuing \ac{EA} demands that the technical community continues the discussion.
+
+\paragraph*{Investigators Need Data} The ability to access digital data is important and will be increasingly important
+for investigations.
+
+\paragraph*{Cryptography is Foundational} Various forms of cryptography are foundational all elements of online
+security.
+
+\paragraph*{Mass Surveillance is a Danger} Mass surveillance is a grave danger from which representative governments are
+not immune.
+
+\paragraph*{\ac{EA} is an International Problem} The internet is largely borderless, and international and human rights
+considerations significantly constrain workable \ac{EA} solutions.
+
+\paragraph*{Transparency is Necessary} Transparency to data access is critical for public trust.
+
+% \begin{itemize}
+%     \item The failure to convince legislative bodies and law enforcement agencies demands that the technical community continue the discussion
+%     \item The ability to access digital data is important and will be increasingly important for investigations
+%     \item Various forms of encryption are foundational all elements of online security
+%     \item Mass surveillance is a grave danger that representative governments are not immune from
+%     \item International and human rights considerations significantly constrain workable solutions
+%     \item Transparency to data access is critical for public trust
+%     % \item In pluralistic societies, government is obligated to respect and balance opposing viewpoints
+% \end{itemize}
 
 \section{Contribution}
+\label{sec-contribution}
 
@@
 models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
-creates a multi-level threat model, broken down by the significantly different challenges posed by data at rest and data
-in motion. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
+creates a multi-level threat model, broken down by the significantly different challenges posed by \ac{DAR} and
+\ac{DIM}. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
 and threats.
-
-% \begin{itemize}
-%     \item Argument maps formalizing the debate
-%     \item A broad threat model describing the environment
-%     \item A structure for organizing future progress
-% \end{itemize}

### 2020-04-05 Thesis-Glossary.tex
Work on the introduction
Work on background; rearrange sections a bit

@@
 \newcommand{\term}{\ac}
+\newcommand{\Term}{\Ac}
 
@@
 % NOTE: Do not end with a period. The package adds this for you.
-% TODO: auto-update glossary.md. This should be the master, because it is more fully specified
+% IDEA: auto-update glossary.md. This should be the master, because it is more fully specified
 
@@
 
-\newterm{horsemen}{The Four Horsemen}{
+\newterm{horsemen}{the Four Horsemen}{
     Four reasons that governments and law enforcement agencies use to undercut public support for strong encryption. The
@@
 }
+\newterm{horsemen-full}{The Four Horsemen of the Infocalypse}{The long name for ``The Four Horsemen''}
 
-\newterm{horsemen-full}{The Four Horsemen of the Infocalypse}{
-    The long name for ``The Four Horsemen''
+\newterm{plaintext}{plaintext}{Unencrypted data}
+\newterm{ciphertext}{ciphertext}{Encrypted data}
+
+\newterm{cryptography}{cryptography}{
+    The study of techniques for communicating secretly in the presence of third parties
+}
+
+\newterm{cipher}{cipher}{
+    A tool or algorithm to perform encryption, translating between plaintext and ciphertext
+}
+\newterm{key}{key}{Secret information required to operate the cipher and perform encryption and decryption}
+
+\newterm{encryption}{encryption}{The process of translating from plaintext to ciphertext}
+\newterm{decryption}{decryption}{The process of translating from ciphertext to plaintext}
+
+\newterm{public-key-crypto}{public key cryptography}{
+    Another name for asymmetric encryption
+}
+\newterm{asym-crypto}{asymmetric cryptography}{
+    A cipher scheme where encryption and decryption are performed with separate, paired keys, called the public key and
+    the private key
+}
+\newterm{sym-crypto}{symmetric cryptography}{
+    A cipher scheme where encryption and decryption are performed with the same key
 }
@@
 \newabbreviation{E2E}{E2E}{end to end}
-\newabbreviation{EA}{EA}{exceptional (or extraordinary) access}
+\newabbreviation{EA}{EA}{exceptional access}
 \newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}

### 2020-04-05 Thesis-Glossary.tex
Write more background

@@
 % Start terms
+\newterm{the-cw1}{the first crypto war}{
+    The flurry of evolution and conflict in encryption policy during the 1990s, characterized by episodes around export
+    controls, \ac{PGP}, the \ac{clipperchip}, and early key escrow proposals
+}
+\newterm{the-cw2}{the second crypto war}{
+    The flurry of conflict in encryption policy during the late 2010s to present, characterized by angst from the
+    Snowden revelations, the proliferation of strong encryption, and the Apple vs. FBI case
+}
+
 \newterm{clipperchip}{Clipper Chip}{
@@
 }
+
+\newterm{forward-secrecy}{forward secrecy}{
+    A property of \ac{DIM} encryption protocols that ensures that a leaked private key or session key does not
+    compromise any other private key or session key
+}
+\newterm{replay-protection}{replay protection}{
+    A property of \ac{DIM} encryption protocols that ensures that messages cannot be replayed by an attacker without
+    detection
+}
+
+\newterm{disk-encryption}{disk encryption}{
+    An application of ecnryption for \ac{DAR} that involves encrypting the contents
+}
 % End terms
@@
 \newabbreviation{E2E}{E2E}{end to end}
+\newabbreviation{E2EE}{E2EE}{end to end encryption}
 \newabbreviation{EA}{EA}{exceptional access}

### 2020-04-05 scripts/show-source.sh

@@ -12,8 +12,11 @@ if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
     done
     exit 0
 elif [[ 1 == $# && -f $ORIG_DIR/$1 ]]; then
-    # Show a file based on path
+    # Show a file based on relative path
     FILE=$ORIG_DIR/$1
+elif [[ 1 == $# && -f $1 ]]; then
+    # Show a file based on absolute path
+    FILE=$1
 elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
     # Show a file based on bibtex ref
     FILE="$(grep -riIsl "@*{$1," . | head -1)"

### 2020-04-05 Thesis-Chapter1-Introduction.tex
Orgranize the sections, update the TODOs
Work on introduction

@@
 
-% \begin{itemize}
-%     % TODO: Why strong encryption is important + Why some want exceptional access
-%     \item In the pursuit of data privacy, regulatory action is part of the threat model--and action is likely
-%     \item If the technical community acts, we can not only be prepared for regulation, but lead and shape it
-% \end{itemize}
-
-% TODO: clarify structure of argument in this whole section; it is a little hacked together and logically scattered
-
-The conflict at the heart of the debate is that the same cryptographic and design principles that underlie all digital
-security also enable an unprecedented degree of individual privacy. Encryption is a foundational tool to the integrity
-and confidentiality of all connected computing systems. Its increasing ubiquity in communications and storage provides
-clear benefits and should be welcomed by all. In a world where information security is frightfully poor yet increasingly
-important, strong encryption's necessity can not be understated. However, the privacy afforded by certain encryption
-technologies hampers law enforcement investigations and hides wrongdoing. In a society that cares about bringing
-dangerous people to justice, this risk should not be ignored.
+The conflict at the heart of the encryption and \acl{EA} is that the same cryptographic and design principles that
+underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
+foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
+communications and storage provides clear benefits and should be welcomed by all. In a world where information security
+is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
+privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. In a
+society that cares about bringing wrongdoers to justice, this risk should not be ignored.
 
@@
 
-It is time that the technical community moves the debate forward. Increased regulation of the tech industry is
-inevitable. Despite strong encryption's benefits, regulatory interest in the subject has not subdued. The form that
+In the long term, increased regulation of the tech industry in general is inevitable, and regulation of encryption is
+possible. Despite strong encryption's benefits, regulatory interest in the subject has not subdued. The form that
 regulation will take in the field of encryption depends on the good faith efforts made to equitably balance all the
-benefits and risks involved in deploying a crypto system. If the technical community acts, we can not only be prepared
-for regulation, but lead and shape it.
+benefits and risks involved in deploying a cryptographic system. If the technical community acts, it can not only be
+prepared for regulation, but lead and shape it.
 
-Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model.
+Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
+community fails to act but regulators move forward, everyone may become subject to technically misguided laws that do
+great harm. Bad \ac{EA} policy is just as much a threat as poor passwords.
+
+For all these reasons, it is time that the technical community moves the debate forward.
 
@@
 
-This paper accepts and builds on the following premises, some of which were introduced in \mysec{sec-motivation}.
+This paper accepts and builds on the following premises, some of which were introduced in \mysec{sec-motivation}. These
+premises are not principles for potential \ac{EA} designs, but the starting place for discussing the \ac{EA} debate.
 
-% TODO: flesh out, probably add some references
+\eestart
 
-\paragraph*{The Discussion Must Go On} The failure of the technical community to convince legislative bodies and law
-enforcement agencies to stop pursuing \ac{EA} demands that the technical community continues the discussion.
+\ee{The discussion must go on} As long as regulators continue pursuing \ac{EA} regulation, it is the responsibility of
+the technical community to engage and respect the valid arguments being presented.
 
-\paragraph*{Investigators Need Data} The ability to access digital data is important and will be increasingly important
-for investigations.
+\ee{Cryptography is foundational} Strong cryptography in its various forms is foundational to digital security.
 
-\paragraph*{Cryptography is Foundational} Various forms of cryptography are foundational all elements of online
-security.
+\ee{Investigators need data} The ability to access digital data for investigation and prosecution is important today and
+will be increasingly important in the future.
 
-\paragraph*{Mass Surveillance is a Danger} Mass surveillance is a grave danger from which representative governments are
-not immune.
+\ee{Mass surveillance is a danger} Mass surveillance, which \ac{EA} makes easier, is a grave danger from which
+representative governments are not immune.
 
-\paragraph*{\ac{EA} is an International Problem} The internet is largely borderless, and international and human rights
+\ee{Transparency is necessary} Transparency of the law enforcement and intelligence communities' data access capability
+and usage is critical for public trust.
+
+\ee{\Ac{EA} is an international problem} The internet is largely borderless, and international and human rights
 considerations significantly constrain workable \ac{EA} solutions.
 
-\paragraph*{Transparency is Necessary} Transparency to data access is critical for public trust.
+\ee{Perfection is not the goal} To loosely quote the ``insecurity axioms,'' insecurity cannot be destroyed, only moved
+around \cite{nrc_schneider_1999}. \Ac{EA}, like all security, is a matter of finding the right balance of risk.
 
-% \begin{itemize}
-%     \item The failure to convince legislative bodies and law enforcement agencies demands that the technical community continue the discussion
-%     \item The ability to access digital data is important and will be increasingly important for investigations
-%     \item Various forms of encryption are foundational all elements of online security
-%     \item Mass surveillance is a grave danger that representative governments are not immune from
-%     \item International and human rights considerations significantly constrain workable solutions
-%     \item Transparency to data access is critical for public trust
-%     % \item In pluralistic societies, government is obligated to respect and balance opposing viewpoints
-% \end{itemize}
+\eeend
 

### 2020-04-05 scripts/sourceme.sh
Scripts: add some more helper aliases

@@ -21,3 +21,7 @@ alias show="$SCRIPTS_DIR/show-source.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
 alias todo="(cd $BASE_DIR && grep -n TODO *.tex)"
+
+alias pdf="xdg-open build/Thesis.pdf"
+alias gvsu="xdg-open build/Thesis-GVSU.pdf"
+alias start="(codium .) && (xdg-open build/Thesis.pdf) && make live"

### 2020-04-08 Thesis-Chapter1-Introduction.tex
Work on background, add demo argmap and dfd

@@
 
-Difficult policy matters are sometimes classified as wicked problems. Wicked problems are defined not by their moral
-nature, but by their vicious persistence. These problems have multiple causes and no clear solution; attempted solutions
-have unintended consequences, and the complex environment results in chronic policy failure
-\cite{commission_tackling_2018}.
+Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
+\ac{wicked-prob} uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
+\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
+answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or metrics. Each attempt at a solution matters,
+has unintended consequences, and often exists in a pattern of chronic policy failure.
 

### 2020-04-08 Thesis-Chapter2-Background.tex
Work on background, add demo argmap and dfd

@@
 
-This chapter summarizes the history of encryption and the exceptional access debate, introduces argument maps as a
-logical modeling tool, and introduces threat modeling with data flow diagrams.
+The problem of strong encryption and exceptional access exists in a technical, historical, and regulatory context. This
+chapter introduces cryptography, summarizes the history of encryption and the crypto wars, lists relevant encryption
+regulations, and lists prominent technical \ac{EA} proposals. It then introduces argument maps as a logical modeling
+tool, and introduces threat modeling with data flow diagrams.
 
-\section{Historical Context}
-\label{sec-history}
+\section{Cryptography Basics}
+\label{sec-crypto-basics}
 
-\subsection{Encryption History and the First Crypto War}
-\label{sec-crypto-cwi}
+\Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
+performed by using some \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. The cipher is the tool or
+algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
+processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
+cipher ensures that only those with the correct secret information, the \ac{key}, can perform encryption or decryption
+on the text.
 
-Electronic encryption is a new technology, but the field of cryptography is old, as demand for privacy is as old as
-communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
-the character transposition technique Julius Caesar used to protect private correspondence
-\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
-in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
-computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
-analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
-research on public key cryptography in 1976 \cite{diffie_new_1976}.
+% There are two classes of cryptographic ciphers--stream and block
 
-Public key cryptography, paired with computer networking and personal computing advances in the 1980s, for the first
-time put the power of strong encryption in the hands of ordinary people. However, the privacy provided by cryptography
-is a powerful tool, and privacy is a property governments strive to regulate. The use of encryption for for private
-purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason based on evidence from
-decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force testimony from his secretary on
-the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it
-inherently uncomfortable with the rapid increase in availability of strong encryption.
+There are two major cryptographic protocol families--symmetric and asymmetric. In traditional \ac{sym-crypto},
+encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
+securely. In \ac{asym-crypto}, encryption and decryption are performed with separate, paired keys, called the public key
+and the private key. The innovation of \ac{public-key-crypto} is that the public key is not secret, and using this
+technique, two parties can communicate securely without requiring an a previously agreed upon shared secret.
 
-% Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
-% human rights. [This is where I'd like to understand more of the legal side. With a historical perspective.]
+One can also define two major cryptographic applications--securing \acf{DIM} and \acf{DAR}. Each application is a
+different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
+perform authentication and establish short-lived \ac{sym-crypto} sessions keys; the sessions keys perform the actual
+encryption of the data in transit. Important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}, which
+ensures that a leaked private key or session key does not compromise any other private key or session key, and
+\ac{replay-protection}, which ensures that messages cannot be replayed by an attacker without detection
+\cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for \ac{DIM}.
 
-Privacy is not encryption's only purpose, however.
+\Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
+encryption-time, as \ac{DIM} session keys are, these keys are either derived from a password stored in human memory or
+are stored somewhere in computer memory, which often takes place with the assistance of dedicated hardware.
+\Ac{disk-encryption} for laptops and mobile devices is an example of encryption for \ac{DAR}.
+
+Data secrecy is not cryptography's only purpose, however.
+
+Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
+often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
+and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
+from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
 
@@
 
-Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
-often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
-and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
-from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
-
-Cryptographic primitives include symmetric key encryption, public key encryption, one-way hash functions, and digital
-signatures. A few technologies that directly rely on these primitives include TLS (transport layer security), PKI
-(public key infrastructure), SSO (single sign on, e.g. Kerberos), VPNs (virtual private networks), secure password
-storage, disk encryption tools, and secure- and trusted-boot processes. A few systems that directly rely on these
-technologies include the financial system, local and global; industrial control systems and critical infrastructure;
-health information systems; private, government, and military communications; and the integrity of the internet as a
-whole.
-
-If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
-collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
-infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
-whole loses integrity, inconceivably large damage would occur.
+Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
+signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
+Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
+systems that directly rely on these technologies include the financial system, local and global; industrial control
+systems and critical infrastructure; health information systems; private, government, and military communications; and
+the integrity of the internet as a whole.
 
@@
 
-% Systems: ...
+If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
+collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
+infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
+whole loses integrity, inconceivably large damage would occur.
 
+% FIXME: Kristen doesn't like the wording; also remove word "electronic"?
 Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
-conflict in the EA debate. Those who focus on its role in privacy object to its absolute power. Those
-who focus on its role in security rely on its absolute power.
+conflict in the EA debate. Those who focus on its role in privacy object to its absolute power. Those who focus on its
+role in security rely on its absolute power.
 
-% Encryption requires absolute cryptographic fidelity in order to provide security.
+\section{Encryption History}
+\label{sec-crypto-history}
+
+The section provides an overview of encryption from ancient to contemporary times.
+
+\subsection{Encryption In the Past}
+\label{sec-history-old}
+
+Computerized encryption is a new technology, but the field of cryptography is old, as demand for privacy is as old as
+communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
+the character transposition technique Julius Caesar used to protect private correspondence
+\cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
+in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
+computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
+analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
+research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}.
+
+The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
+computing advances in the 1980s, it for the first time put the power of strong encryption in the hands of ordinary
+people. However, the privacy provided by cryptography is a powerful tool, and privacy is a property governments strive
+to regulate. The use of encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was
+convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason
+tried to force testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}.
+Government's natural stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of
+strong encryption.
+
+% Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
+% human rights.
+
+\subsection{The First Crypto War}
+\label{sec-history-cw1}
 
@@
 and a blossoming computer industry foundationally reliant on cryptography--came to a head in what has become known as
-the First Crypto War.
+\ac{the-cw1}.
 
 In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
-the Arms Export Control Act and declares that strong cryptography is subject to export controls \cite{kehl_right_2015}.
-In 1991, the U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by
-law; in response, Philip Zimmermann released PGP (Pretty Good Privacy), a public key encryption tool to secure email, in
-order that strong cryptography be ``made available to the American public before it became illegal to use''
-\cite{zimmermann_1996}. In 1993, the Clinton administration introduced Clipper Chip \cite{press_1993} with the goal of
-``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement and
-intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
-foundational role of cryptography in security--and the potential of human rights abuses for compromised
-privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
-\cite{zimmermann_1996}. When security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper
-Chip mechanisms \cite{blaze_protocol_1994}, the proposal died.
+the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. In 1991, the
+U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by law; in
+response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to secure email, in order that strong
+cryptography be ``made available to the American public before it became illegal to use'' \cite{zimmermann_1996}. In
+1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with the goal of ``providing the
+public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
+access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
+in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
+reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
+discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
+died.
 
-Despite the failure of the Clipper Chip, the debate over access to strong encryption continued, primarily focused on
+Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
 various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
@@
 In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
-activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have
-barred the government from creating a mandatory key escrow system and would also have removed export restrictions on
-most generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of
-PGP, who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
+activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have barred
+the government from creating a mandatory key escrow system and would also have removed export restrictions on most
+generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of PGP,
+who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
 Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. In November of that year, the Clinton administration
@@
 \cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all SAFE Act
-proposals \cite{kehl_right_2015}. This development marks the end of the First Crypto War.
+proposals \cite{kehl_right_2015}. This development marks the end of \ac{the-cw1}.
 
-\subsection{The Second Crypto War and the Current Context}
-\label{sec-cwii-context}
+\subsection{The Second Crypto War}
+\label{sec-history-cw2}
 
-Two developments paved the way for the Second Crypto War. The first was actually a non-development: when the Clinton
+Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
 administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. Also un-passed
 were the related Pro-C0DE Act, Encrypted Communications Privacy Act (not to be confused with the \textit{Electronic}
-Communications Privacy Act (ECPA) of 1986), and Encryption Protects the Rights of Individuals from Violation and Abuse
-in CYberspace (E-PRIVACY) Act \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998}. This legislative failure meant that
-rather than being written in the pen of law, encryption policy was written in the pencil of executive order.
+Communications Privacy Act \acs{ECPA} of 1986), and Encryption Protects the Rights of Individuals from Violation and
+Abuse in CYberspace (E-PRIVACY) Act \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998}. This legislative failure meant
+that rather than being written in the pen of law, encryption policy was written in the pencil of executive order.
 
@@
 2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
-FISA requests on U.S. citizens; it expanded the scope of what could be compelled via FISA orders; it authorized
-``roving'' wiretaps; and it increased the power of National Security Letters (NSLs), which can be used without judicial
-review to compel information from electronic service providers while precluding any public release of the event
+\ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
+authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
+compel information from digital service providers while precluding any public disclosure of the event
 \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
@@
 war, and increased government surveillance set events in motion toward conflict, it was the 2013 Snowden revelations
-that constituted the crossing of the threshold. Though policies such as the U.S. National Security Agency (NSA)'s
-warrantless wiretapping caused a stir, it was the mass data collection under the agency's PRISM and related programs
-that caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy,
-encryption promised a technical solution. Under pressure from both the human rights field and from ordinary users
-demanding improved digital privacy, U.S. tech companies responded by introducing default device encryption for
-data-at-rest (DAR) and end-to-end (E2E) encryption for data-in-motion (DIM) \cite{treguer_us_2018}.
+that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
+caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
+public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
+technical solution. Under pressure from both the human rights field and from ordinary users demanding improved digital
+privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and \acl{E2E} encryption
+for \acl{DIM} \cite{treguer_us_2018}.
 
-Strong encryption's transition into being caused the First Crypto War; likewise, strong encryption's transition into
-ubiquity is causing the Second Crypto War.
+The emergence of strong encryption caused \ac{the-cw1}; its proliferation is causing the second.
 
@@
 basic digital security. Recall encryption and cryptography's critical role in security as described in
-\mysec{sec-crypto-cwi}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
-security is a central conflict in the EA debate. When questioned by lawmakers, technical leaders and
-executives have repeatedly appealed to the fact that strong cryptography is a necessity for security
-\cite{schulze_clipper_2017}.
+\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
+security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
+have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.
 
-Several events and arguments have come to characterize the Second Crypto War.
+\subsection{Current Context}
+\label{sec-history-current}
 
-\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as antagonistic,
-an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown that the
-debate is more about ``Security vs. Security'' than ``Security vs. Privacy'' \cite{stalla_bourdillon_privacy_2014}. In
-this context, there are two types of security: public security, the pursuit of law enforcement, and cybersecurity, the
-pursuit of tech professionals. Both sides of the debate have by now acknowledged that privacy and public security are
-not always in conflict, and in fact EA's effects on public security vs. cybersecurity--which in return affects public
-security--may be more important \cite{schneier_2019}.
+Several events and arguments have come to characterize \ac{the-cw2}.
+
+\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
+antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
+that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
+\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
+law enforcement, and cybersecurity, the pursuit of tech professionals. Both sides of the debate have by now acknowledged
+that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public security vs.
+cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.
 
 \paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
-locked iPhone, resulted in the first major battle of the Second Crypto War. The FBI issued an order under the All Writs
-Act to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
+locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
+to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
 objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
 \cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
-February 16, 2016, when the warrant was issued, to March 28, when the FBI announced they had gained access to the phone
-through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
+February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
+phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
 \cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto wars,
-terrorism and device encryption. Shortly after the case ended, Senators Dianne Feinstein and Richard Burr released a
-discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the
-floor, and the unresolved nature of the court case left the legal debate wide open.
+terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators jointly
+released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it
+to the floor, and the unresolved nature of the court case left the legal debate wide open.
 
-\paragraph*{The Four Horsemen} A return to the Four Horsemen \cite{schneier_scaring_2019}, with a warranted focus
-currently on child sexual abuse material \cite{keller_internet_2019} \cite{geller_2019}.
+\paragraph*{\Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
+describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
+\cite{may_1994}. The traditional list includes terrorists, pedophiles, drug dealers, and money launderers, though
+sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
+\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography
+\cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
+terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on child
+pornography is likely a result of the growth of the ``dark web'' and a severe, in-depth 2019 investigation by the
+\textit{New York Times} \cite{keller_internet_2019}.
 
-\paragraph*{DIM vs. DAR} A pendulating emphasis on DIM/E2E and DAR/DE \cite{schneier_2019}.
+\paragraph*{\ac{DIM} vs. \ac{DAR}} As noted in \mysec{sec-crypto-basics}, securing \acl{DIM} and \acl{DAR} are two
+different problems that result in different solutions. Therefore, both technical and policy proposals often split their
+recommendations along these lines \cite{group_2019} \cite{owen_law_2018}. The focus from law enforcement in the U.S. has
+pendulated between each of these applications a few times \cite{schneier_2019}. The typical \ac{DIM} application is
+\ac{E2EE} in instant messaging applications, whereas the typical \ac{DAR} application is \ac{disk-encryption} of mobile
+phones. A research group at Carnegie Mellon with participants from both sides of the debate has identified \ac{EA} for
+\ac{DAR} as the more tractable problem \cite{group_2019}.
 
-\paragraph*{Volatile Politics} A political landscape marked by uncertainty, extremism, misinformation, and most
-recently, a pandemic that will have unknown impacts.
+\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
+uncertainty, extremism, misinformation, and most recently, a pandemic that will have unknown impacts. Since legislature
+is part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
+leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
+community to engage productively with policy makers \cite{granick_2018}.
 
@@
 16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
-committee determines the platform is not doing enough to fight child sexual abuse material (CSAM), a legal synonym for
+committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for
 child pornography \cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
 \cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
-largely see the EARN IT Act as an attempt to indirectly yet effectively outlaw E2E \cite{newman_2020}
+largely see the EARN IT Act as an attempt to indirectly yet effectively outlaw \ac{E2E} encryption \cite{newman_2020}
 \cite{pfefferkorn_2020}.
 
-% TODO: keep up to date
-% TODO: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
+% NOTE: keep up to date, including the section above mentioning the pandemic
+% IDEA: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
 
@@
 
-\subsection{Regulatory Environment}
+\section{Regulatory Environment}
 \label{sec-reg-environment}
 
-\subsubsection*{US Efforts to Regulate Encryption}
+\subsection{U.S. Efforts to Regulate Encryption}
 
@@
 \begin{itemize}
-    \item The All Writs Act \cite{congress_1789}
-    \item The Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
-    \item The Arms Export Control Act (AECA) of 1976 established export controls over strong encryption techniques
+    \item \acf{AWA} of 1789 \cite{congress_1789}
+    \item Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
+    \item \acf{AECA} of 1976 established export controls over strong encryption techniques
         introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
-    \item Foreign Intelligence Surveillance (FISA) Act \cite{rodino_1978}
+    \item \acf{FISA} of 1978 \cite{rodino_1978}
         (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
@@
         %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
-    \item The Computer Fraud and Abuse Act (CFAA) of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
-    \item Electronic Communications Privacy Act (ECPA) of 1986 \cite{kastenmeier_hr4952_1986}
-    \item The Communications Assistance for Law Enforcement Act (CALEA) of 1994 \cite{edwards_hr4922_1994}
+    \item \acf{CFAA} of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
+    \item \acf{ECPA} of 1986 \cite{kastenmeier_hr4952_1986}
+    \item \acf{CALEA} of 1994 \cite{edwards_hr4922_1994}
     \item Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism
         (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
-    \item Protect America Act (PAA) of 2007 \cite{mcconnell_s1927_2007}
-    \item FISA Amendments Act of 2008 \cite{reyes_hr6304_2008}
+    \item \acf{PAA} of 2007 \cite{mcconnell_s1927_2007}
+    \item \acf{FISAAA} of 2008 \cite{reyes_hr6304_2008}
     \item Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring
@@
 \begin{itemize}
-    \item The Clipper Chip \cite{press_1993} \cite{thompson_2015}
+    \item The \ac{clipperchip} \cite{press_1993} \cite{thompson_2015}
     \item The SAFE Act \cite{goodlatte_hr3011_1996}
@@
         on same day) \cite{goss_hr2616_1999} \cite{goss_hr2617_1999}
-    \item Compiance with Court Orders Act of 2016 (discussion draft) \cite{burr_2016}
+    \item Compliance with Court Orders Act of 2016 (discussion draft) \cite{burr_2016}
     \item The Secure Data Act (federal SE protections) \cite{lofgren_hr5823_2018}
@@
 
-\subsubsection*{Regulations around the World}
+\subsection{Regulations around the World}
 
@@
 
-\subsection{Technical Approaches}
+\section{Technical Approaches}
 \label{sec-tech-approaches}
 
-\subsubsection*{Historical Approaches}
+\subsection{Historical Approaches}
 
-Historical approaches and proposals out of the first crypto war that saw various amounts of support.
+Historical approaches and proposals out of \ac{the-cw1} that saw various amounts of support.
 
@@
 
-\subsubsection*{Modern Proposals}
+\subsection{Modern Proposals}
 
-Recent proposals out of the second crypto war that have various levels of support.
+Recent proposals out of \ac{the-cw2} that have various levels of support.
 
@@
 
+In order to advance the \ac{EA} debate, it helps to organize the all the arguments used in a cohesive manner. One way to
+organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and
+relationships between statements, premises, and conclusions. Argument maps are an outgrowth of the same research by
+Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
+down \acp{wicked-prob} and document the reasoned approach that goes into solving them \cite{kunz_issues_1970}. Interest
+in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
+\cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006} tools designed to produce graphical
+representations. Argdown is a recent tool that generates argument maps from structures specified in a Markdown-based
+language \cite{voigt_argdown_2018}.
+
+This paper uses Argdown to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the basic
+structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be addressed
+through argument. Statements and arguments may support or attack one another, and arguments support or undercut other
+arguments.
+
+\begin{figure}[ht]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{arguments/build/demo.pdf}
+    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
+    \label{fig-args-demo}
+\end{figure}
+
+% IDEA:
+% Use the slightly more complete IBIS notation, which includes ISSUES and POSITIONS as well as ARGUMENTS.
+% (Argdown) basically just has arguments and positions. The additional types may make the total picture clearer.
+% Either use styling with #hashtags (https://argdown.org/syntax/#hashtags with
+%                                    https://argdown.org/guide/colorizing-maps.html#the-colorscheme-setting)
+% Or add a plugin (https://argdown.org/guide/extending-argdown.html)
+
 \section{Threat Models and Data Flow Diagrams}
 \label{sec-threat-model-intro}
+
+Threat modeling is an important step in analyzing the security at the systems level. Using models abstracts away fine
+details and focuses on the architecture, processes, and dataflows in a system. Models assist in the understanding of
+current systems and the prevention of problems in new systems, both of which are important when facing the prospect of
+designing an \ac{EA} mechanism built on top of a complex and aging technology stack.
+
+Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
+\cite{shostack_threat_2014}. Answering the first question accurately is crucial, particularly for the field of
+cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. Answering the second
+question depends on the types of attackers to worry about, and determines the scope of threats to be considered. ``What
+are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go wrong?''
+cannot be faithfully answered without knowing what is being built.
+
+\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
+external entities. \Acp{DFD} well suited for threat modeling because security vulnerabilities tend to follow data flow,
+not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
+primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. This paper creates \acp{DFD} in this style
+to analyze the threats already present and those that \ac{EA} would introduce.
+
+\begin{figure}[ht]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/DFD-demo.png}
+    \caption[\Acs{DFD} Demo]{A demonstrative example of a \acf{DFD}.}
+    \label{fig-dfd-demo}
+\end{figure}

### 2020-04-08 Thesis-Glossary.tex
Remove \term->\ac tex alias
Work on introduction
Work on background, add demo argmap and dfd

@@
 }
-\newcommand{\term}{\ac}
-\newcommand{\Term}{\Ac}
 
@@
 % Start terms
+\newterm{wicked-prob}{wicked problem}{
+    A problem
+}
+
 \newterm{the-cw1}{the first crypto war}{
@@
 }
-\newterm{horsemen-full}{The Four Horsemen of the Infocalypse}{The long name for ``The Four Horsemen''}
+\newterm{horsemen-full}{the Four Horsemen of the Infocalypse}{The long name for ``The Four Horsemen''}
 
@@
 \newabbreviation{CSAM}{CSAM}{child sexual abuse material}
+\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
 \newabbreviation{DE}{DE}{device encryption}
+\newabbreviation{DFD}{DFD}{data flow diagram}
 \newabbreviation{DIM}{DIM}{data in motion}
@@
 \newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
+\newabbreviation{IBIS}{IBIS}{issue-based information system}
 \newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}

### 2020-04-08 scripts/install-deps.sh

@@ -15,3 +15,4 @@ sudo apt-get install -y \
     libcanberra-gtk-module \
     libcanberra-gtk3-module
 
+snap install drawio

### 2020-04-09 Thesis-Chapter2-Background.tex
Work on background

@@
 
-\begin{itemize}
-    \item \acf{AWA} of 1789 \cite{congress_1789}
-    \item Bill of Rights: Amendments I, IV, and V \cite{madison_1791}
-    \item \acf{AECA} of 1976 established export controls over strong encryption techniques
-        introduced ITAR (International Traffic in Arms Regulations) \cite{morgan_hr13680_1976} \cite{kehl_right_2015}
-    \item \acf{FISA} of 1978 \cite{rodino_1978}
-        (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
+\newcommand{\lawstart}[0]{\begin{itemize}}
+\newcommand{\law}[3]{\item #1: \textbf{#2} \cite{#3}} % Year, name, citation
+\newcommand{\lawsend}{\end{itemize}}
+
+\lawstart
+    \law{1789}{\acf{AWA}}{congress_1789}
+    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}
+    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}
+% Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
+% Regulations) \cite{kehl_right_2015}
+    \law{1978}{\acf{FISA}}{rodino_1978}
+% (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
 % (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
@@
 %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
-    \item \acf{CFAA} of 1986 \cite{hughes_hr4718_1986} \cite{wolff_computer_2016}
-    \item \acf{ECPA} of 1986 \cite{kastenmeier_hr4952_1986}
-    \item \acf{CALEA} of 1994 \cite{edwards_hr4922_1994}
-    \item Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism
-        (USA PATRIOT ACT) Act of 2001 \cite{sensenbrenner_2001}
-    \item \acf{PAA} of 2007 \cite{mcconnell_s1927_2007}
-    \item \acf{FISAAA} of 2008 \cite{reyes_hr6304_2008}
-    \item Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring
-        (USA FREEDOM) Act of 2015 \cite{sensenbrenner_2015}
-    \item \dots
-\end{itemize}
+    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}
+% \cite{wolff_computer_2016}
+    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}
+    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}
+    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}
+    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}
+    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}
+    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}
+\lawsend
 
@@
 
-\begin{itemize}
-    \item The \ac{clipperchip} \cite{press_1993} \cite{thompson_2015}
-    \item The SAFE Act \cite{goodlatte_hr3011_1996}
-    \item Encrypted Communications Privacy Act (mix of SE protection and EA guidelines?) \cite{leahy_s376_1997}
-    \item Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act
-        (like above?) \cite{ashcroft_s2067_1998}
-    \item Encryption for the National Interest Act and Tax Relief for Responsible Encryption Act (introduced by same rep
-        on same day) \cite{goss_hr2616_1999} \cite{goss_hr2617_1999}
-    \item Compliance with Court Orders Act of 2016 (discussion draft) \cite{burr_2016}
-    \item The Secure Data Act (federal SE protections) \cite{lofgren_hr5823_2018}
-    \item The ENCRYPT Act (state SE protections) \cite{lieu_hr4170_2019}
-    \item The EARN IT Act \cite{graham_s3398_2020}
-    \item \dots
-\end{itemize}
+\lawstart
+    \law{1993}{The \ac{clipperchip}}{press_1993}
+% \cite{thompson_2015}
+    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}
+    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}
+% mix of SE protection and EA guidelines?
+    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}
+% like above?
+    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}
+    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}
+% ^^ these two introduced by same rep on same day
+    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}
+    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}
+% federal SE protections
+    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}
+% state SE protections
+    \law{2020}{The EARN IT Act}{graham_s3398_2020}
+\lawsend
 
@@
 
-\begin{itemize}
-    \item Australia: in 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
-    \item UK: Investigatory Powers Act \cite{legislature_2016}
-    \item EU: GDPR \cite{parliament_2016}
-    \item China: \cite{donahue_comparative_2018}
-    \item Israel: \cite{donahue_comparative_2018}
-    \item Japan, Korea? -- can't find anything
-    \item \dots General, see \cite{budish_encryption_2018}
-\end{itemize}
+\newcommand{\countrystart}[0]{\begin{itemize}}
+\newcommand{\country}[1]{\item \textbf{#1:}} % Country
+\newcommand{\countryend}{\end{itemize}}
+
+\countrystart
+    \country{Australia} In 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
+    \country{UK} Investigatory Powers Act \cite{legislature_2016}
+    \country{EU} GDPR \cite{parliament_2016}
+    \country{China} See \cite{donahue_comparative_2018}
+    \country{Israel} See \cite{donahue_comparative_2018}
+    \country{Japan, Korea} can't find anything
+    \country{General} See \cite{budish_encryption_2018}
+\countryend
 
@@
 
-\begin{itemize}
-    \item Clipper chip/third-party key escrow \cite{blaze_protocol_1994}
-    \item Oblivious key escrow \cite{goos_oblivious_1996}
-    \item Partial key escrow (see \cite{denning_taxonomy_1996})
-\end{itemize}
+\newcommand{\propstart}[0]{\begin{itemize}}
+\newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
+\newcommand{\propend}{\end{itemize}}
+
+\propstart
+    \prop{Clipper chip/third-party key escrow}{press_1993}
+% toasted by \cite{blaze_protocol_1994}
+    \prop{Oblivious key escrow}{goos_oblivious_1996}
+    \prop{Partial key escrow}{denning_taxonomy_1996}
+\propend
 
@@
 
-\begin{itemize}
-    \item CLEAR \cite{ozzie_2018}
-    \item Device escrow with HW \cite{savage_lawful_2018}
-    \item Computational, blockchain approaches \cite{phan_key_2017}
-    \item Lawful hacking \cite{nguyen_lawful_2017} \cite{soesanto_2018} \cite{kerr_encryption_2017}
-    \item Forced password disclosure \cite{bittenbender_2019} \cite{kerr_encryption_2017}
-    \item Alternative paths to plaintext (access device while unlocked; find copies elsewhere)
-        \cite{kerr_encryption_2017}
-\end{itemize}
+\propstart
+    \prop{CLEAR}{ozzie_2018}
+    \prop{Device escrow with HW}{savage_lawful_2018}
+    \prop{Computational, blockchain approaches}{phan_key_2017}
+    \prop{Lawful hacking}{nguyen_lawful_2017}
+% see also \cite{soesanto_2018} \cite{kerr_encryption_2017}
+    \prop{Forced password disclosure}{kerr_encryption_2017}
+% see also \cite{bittenbender_2019}
+    \prop{Alternative paths to plaintext}{kerr_encryption_2017}
+% access device while unlocked; find copies elsewhere
+\propend
 

### 2020-04-09 Thesis-Glossary.tex
Work on background

@@
 \newterm{wicked-prob}{wicked problem}{
-    A problem
+    A problem lacking clear formulations, causes, resolutions, or metrics. Consistenly difficult to solve, and often
+    social or political in nature
 }
@@
 
+% Custom acronyms
+
+\newabbreviation{USAPATRIOT}{USA PATRIOT}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism}
+\newabbreviation{USAFREEDOM}{USA FREEDOM}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-09 scripts/count-citations.sh

@@ -5,7 +5,7 @@ source ./scripts/sourceme.sh
 # set -x
 
 function print_number() {
-    echo $(grep "\\cite{$1}" *.tex | wc -l)
+    echo $(grep "{$1}" *.tex | wc -l)
 }
 
 function print_count() {

### 2020-04-09 Thesis-Chapter1-Introduction.tex
Move EA alternatives to own section

@@
 
-\eestart
+\newcommand{\principlesstart}{\begin{enumerate}}
+\newcommand{\principle}[1]{\item \textbf{#1.}}
+\newcommand{\principlesend}{\end{enumerate}}
 
-\ee{The discussion must go on} As long as regulators continue pursuing \ac{EA} regulation, it is the responsibility of
-the technical community to engage and respect the valid arguments being presented.
+\principlesstart
 
-\ee{Cryptography is foundational} Strong cryptography in its various forms is foundational to digital security.
+\principle{The discussion must go on} As long as regulators continue pursuing \ac{EA} regulation, it is the
+responsibility of the technical community to engage and respect the valid arguments being presented.
 
-\ee{Investigators need data} The ability to access digital data for investigation and prosecution is important today and
-will be increasingly important in the future.
+\principle{Cryptography is foundational} Strong cryptography in its various forms is foundational to digital security.
 
-\ee{Mass surveillance is a danger} Mass surveillance, which \ac{EA} makes easier, is a grave danger from which
+\principle{Investigators need data} The ability to access digital data for investigation and prosecution is important
+today and will be increasingly important in the future.
+
+\principle{Mass surveillance is a danger} Mass surveillance, which \ac{EA} makes easier, is a grave danger from which
 representative governments are not immune.
 
-\ee{Transparency is necessary} Transparency of the law enforcement and intelligence communities' data access capability
-and usage is critical for public trust.
+\principle{Transparency is necessary} Transparency of the law enforcement and intelligence communities' data access
+capability and usage is critical for public trust.
 
-\ee{\Ac{EA} is an international problem} The internet is largely borderless, and international and human rights
+\principle{\Ac{EA} is an international problem} The internet is largely borderless, and international and human rights
 considerations significantly constrain workable \ac{EA} solutions.
 
-\ee{Perfection is not the goal} To loosely quote the ``insecurity axioms,'' insecurity cannot be destroyed, only moved
-around \cite{nrc_schneider_1999}. \Ac{EA}, like all security, is a matter of finding the right balance of risk.
+\principle{Perfection is not the goal} To loosely quote the ``insecurity axioms,'' insecurity cannot be destroyed, only
+moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security, is a matter of finding the right balance of risk.
 
-\eeend
+\principlesend
 

### 2020-04-09 scripts/install-deps.sh
Small build system updates

@@ -1,6 +1,7 @@
 #!/bin/bash
 
 # For Bibtex scripts, must have Docker installed. Not covered here.
+# For Argdown CLI, must have NPM installed. Not covered here.
 # This is not necessarily fully-inclusive.
 
 sudo apt-get install -y \
@@ -16,3 +17,5 @@ sudo apt-get install -y \
     libcanberra-gtk3-module
 
 snap install drawio
+
+npm install -g @argdown/cli

### 2020-04-11 scripts/install-deps.sh
Scripts: update install-deps

@@ -1,11 +1,15 @@
 #!/bin/bash
 
-# For Bibtex scripts, must have Docker installed. Not covered here.
-# For Argdown CLI, must have NPM installed. Not covered here.
-# This is not necessarily fully-inclusive.
+# For Zotero container used in some helper scripts, must have Docker installed. Not covered here.
+
+# Also see the Dockerfile for a self-contained build environment.
 
 sudo apt-get install -y \
     make \
+    curl \
+    wget \
+    git \
+    apt-utils \
     texlive-latex-extra \
     texlive-science \
     texlive-xetex \
@@ -14,8 +18,19 @@ sudo apt-get install -y \
     latexmk \
     inkscape \
     libcanberra-gtk-module \
-    libcanberra-gtk3-module
+    libcanberra-gtk3-module \
+    npm
 
-snap install drawio
+curl -sL https://deb.nodesource.com/setup_10.x | bash -
+sudo apt-get install -y --no-install-recommends \
+    nodejs
+sudo npm install -g @argdown/cli
 
-npm install -g @argdown/cli
+if which snap &>/dev/null; then
+    snap install drawio
+else
+    URL=$(curl -sL https://github.com/jgraph/drawio-desktop/releases/latest | grep "\.deb" | head -1  | cut -d\" -f2)
+    wget -q $URL
+    sudo dpkg -i ./*.deb || sudo apt-get -fy install
+    rm ./*.deb
+fi

### 2020-04-11 Thesis-Chapter2-Background.tex
Move EA alternatives to own section
Work on intro, background

@@
 
-\newcommand{\lawstart}[0]{\begin{itemize}}
-\newcommand{\law}[3]{\item #1: \textbf{#2} \cite{#3}} % Year, name, citation
+\newcommand{\lawsstart}{\begin{itemize}}
+\newcommand{\law}[4]{ % Year, name, citation, description
+    \item #1: \textbf{#2} \cite{#3} \newline
+    \parbox{\linewidth}{#4} \newline
+}
 \newcommand{\lawsend}{\end{itemize}}
 
-\lawstart
-    \law{1789}{\acf{AWA}}{congress_1789}
-    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}
-    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}
-% Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
-% Regulations) \cite{kehl_right_2015}
-    \law{1978}{\acf{FISA}}{rodino_1978}
+\lawsstart
+    \law{1789}{\acf{AWA}}{congress_1789}{}
+
+    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}{}
+
+    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}{
+
+Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
+Regulations) \cite{kehl_right_2015}
+
+    }
+
+    \law{1978}{\acf{FISA}}{rodino_1978}{}
+
 % (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
@@
 %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
-    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}
+
+    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{}
+
 % \cite{wolff_computer_2016}
-    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}
-    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}
-    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}
-    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}
-    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}
-    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}
+
+    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{}
+
+    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{}
+
+    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{}
+
+    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{}
+
+    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{}
+
+    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{}
+
 \lawsend
@@
 
-\lawstart
-    \law{1993}{The \ac{clipperchip}}{press_1993}
+\lawsstart
+    \law{1993}{The \ac{clipperchip}}{press_1993}{}
 % \cite{thompson_2015}
-    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}
-    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}
+    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{}
+    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{}
 % mix of SE protection and EA guidelines?
-    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}
+    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}{}
 % like above?
-    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}
-    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}
+    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{}
+    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{}
 % ^^ these two introduced by same rep on same day
-    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}
-    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}
+    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{}
+    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{}
 % federal SE protections
-    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}
+    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{}
 % state SE protections
-    \law{2020}{The EARN IT Act}{graham_s3398_2020}
+    \law{2020}{The EARN IT Act}{graham_s3398_2020}{}
 \lawsend
@@
 
-\newcommand{\countrystart}[0]{\begin{itemize}}
+\newcommand{\countriesstart}{\begin{itemize}}
 \newcommand{\country}[1]{\item \textbf{#1:}} % Country
-\newcommand{\countryend}{\end{itemize}}
+\newcommand{\countriesend}{\end{itemize}}
 
-\countrystart
+\countriesstart
     \country{Australia} In 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
@@
     \country{General} See \cite{budish_encryption_2018}
-\countryend
+\countriesend
 
-\section{Technical Approaches}
+\section{Approaches to Exceptional Access}
 \label{sec-tech-approaches}
+
+% There are several types of approaches:
+% - Allow strong encryption (anyone with the users' private keys can see it)
+% - Disallow encryption (anyone can see it)
+% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
+% - Require use of backdoored algorithms, security via obscurity (anyone with the secret knowledge can see it)
+% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
+%   - Store extra in gov't DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
+%   - Split the key and store it multiple places
+% - Use some alternative path
+%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking
 
@@
 
-\newcommand{\propstart}[0]{\begin{itemize}}
+\newcommand{\propsstart}[0]{\begin{itemize}}
 \newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
-\newcommand{\propend}{\end{itemize}}
+\newcommand{\propsend}{\end{itemize}}
 
-\propstart
+\propsstart
     \prop{Clipper chip/third-party key escrow}{press_1993}
@@
     \prop{Partial key escrow}{denning_taxonomy_1996}
-\propend
+\propsend
 
@@
 
-\propstart
+\propsstart
     \prop{CLEAR}{ozzie_2018}
@@
     \prop{Computational, blockchain approaches}{phan_key_2017}
+\propsend
+
+\subsection{Alternative Approaches}
+
+The following approaches do not use \ac{EA} at all, but are alternative technical or legal mechanisms that can be used
+towards the same goal.
+
+\propsstart
     \prop{Lawful hacking}{nguyen_lawful_2017}
@@
 % access device while unlocked; find copies elsewhere
-\propend
+\propsend
 

### 2020-04-12 Thesis-Chapter1-Introduction.tex
Work on intro, background
Simplify premises/principles

@@
 
-\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption reserved for law
+\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption intended for law
 enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
@@
 community fails to act but regulators move forward, everyone may become subject to technically misguided laws that do
-great harm. Bad \ac{EA} policy is just as much a threat as poor passwords.
+great harm. Bad \ac{EA} policy is just as much a threat as poor passwords. Due to this threat, as long as lawmakers
+continue pursuing \ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and
+respond to the arguments being presented.
 
@@
 
-This paper accepts and builds on the following premises, some of which were introduced in \mysec{sec-motivation}. These
-premises are not principles for potential \ac{EA} designs, but the starting place for discussing the \ac{EA} debate.
+This paper accepts and builds on the following premises. These premises are not principles for potential \ac{EA}
+designs, but the starting place for discussing the \ac{EA} debate.
 
@@
 
-\principle{The discussion must go on} As long as regulators continue pursuing \ac{EA} regulation, it is the
-responsibility of the technical community to engage and respect the valid arguments being presented.
+\principle{Cybersecurity is critical} With current levels of reliance on computer information systems, cybersecurity is
+critical to the wellbeing of society. Two important elements of cybersecurity are cryptography and system architecture.
+Policy which supports security does not undercut cryptographic integrity or require high-risk architectures.
 
-\principle{Cryptography is foundational} Strong cryptography in its various forms is foundational to digital security.
+\principle{Absolute privacy is not an absolute right} Certain rights supersede the legitimate claims of government, but
+absolute privacy is not one of them. In some contexts, absolute privacy is proper, and in almost all contexts,
+reasonable privacy should be presumed. However, privacy is not an absolute right, and investigators should have access
+to some classes of data. Access to these classes of data is important today and will be increasingly important in the
+future.
 
-\principle{Investigators need data} The ability to access digital data for investigation and prosecution is important
-today and will be increasingly important in the future.
+\principle{\Ac{EA} is an inherently complex problem} The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
+proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
+potential for abuse, international reception, economic impact, and need for transparency.
 
-\principle{Mass surveillance is a danger} Mass surveillance, which \ac{EA} makes easier, is a grave danger from which
-representative governments are not immune.
-
-\principle{Transparency is necessary} Transparency of the law enforcement and intelligence communities' data access
-capability and usage is critical for public trust.
-
-\principle{\Ac{EA} is an international problem} The internet is largely borderless, and international and human rights
-considerations significantly constrain workable \ac{EA} solutions.
-
-\principle{Perfection is not the goal} To loosely quote the ``insecurity axioms,'' insecurity cannot be destroyed, only
-moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security, is a matter of finding the right balance of risk.
+\principle{Perfection is not the standard} Wicked policy problems do not have perfect solutions, or even verifiably
+optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
+destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
+right balance of risk given the threats under consideration--which includes regulatory action.
 
@@
 models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
-creates a multi-level threat model, broken down by the significantly different challenges posed by \ac{DAR} and
-\ac{DIM}. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
+creates a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and
+\acl{DIM}. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
 and threats.

### 2020-04-12 Thesis-Chapter2-Background.tex
Work on EA regulations section

@@
 % FIXME: Kristen doesn't like the wording; also remove word "electronic"?
-Electronic encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central
-conflict in the EA debate. Those who focus on its role in privacy object to its absolute power. Those who focus on its
-role in security rely on its absolute power.
+Encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central conflict in
+the EA debate. Those who focus on its role in privacy object to its absolute power. Those who focus on its role in
+security rely on its absolute power.
 
@@
 
+The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
+problem, as digital technology has little respect for borders. This paper is U.S.-focused, but briefly touches on the
+regulatory environment of other countries.
+
 \subsection{U.S. Efforts to Regulate Encryption}
@@
 \lawsstart
-    \law{1789}{\acf{AWA}}{congress_1789}{}
+    \law{1789}{\acf{AWA}}{congress_1789}{
 
-    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}{}
+A short law designed to grant federal courts the right to ``issue all writs necessary or appropriate in aid of their
+respective jurisdictions and agreeable to the usages and principles of law,'' creating a legal framework for providing
+data to law enforcement. The \acs{AWA} was famously cited in the Apple vs. FBI case, though its applicability to \ac{EA}
+in general is unsettled \cite{matyas_incommensurability_2018}.
+
+}
+
+    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}{
+
+The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: security against
+unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.
+
+}
 
@@
 
-Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
-Regulations) \cite{kehl_right_2015}
+Expanded arms export controls and created the \ac{ITAR} regulations framework. Encryption products were classified as
+arms and subjected to export controls \cite{kehl_right_2015}.
 
@@
 
-    \law{1978}{\acf{FISA}}{rodino_1978}{}
+    \law{1978}{\acf{FISA}}{rodino_1978}{
 
-% (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
+Allows the executive branch ``to authorize electronic surveillance for foreign intelligence purposes without a court
+order, in some circumstances.'' \acs{FISA} applications take place in secret, and may be conducted against U.S. citizens
+if foreign intelligence is the ``primary purpose'' \cite{shamsi_2011}. \acs{FISA} has been the subject of frequent FBI
+abuse \cite{shamsi_2011} \cite{tucker_2020}.
+
+}
+
 % (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
@@
 % There are several types of approaches:
+% TODO: after defense, create table of these and reference it in \prop lists below
 % - Allow strong encryption (anyone with the users' private keys can see it)

### 2020-04-12 Thesis-Chapter1-Introduction.tex
Incorporate feedback from Jesse
https://github.com/Jesse-Millwood -- Thanks!

@@
 Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
-\ac{wicked-prob} uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
+``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
 \cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
@@
 enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
-resistance to a clear solution, the debate on \acl{EA} refuses to end. This paper will not attempt to end the debate,
-but to inform it with additional structure.
+resistance to a clear solution, the debate on \ac{EA} refuses to end. This paper will not attempt to end the debate, but
+to inform it with additional structure.
 
@@
 
-The conflict at the heart of the encryption and \acl{EA} is that the same cryptographic and design principles that
+The conflict at the heart of the encryption and \ac{EA} is that the same cryptographic and design principles that
 underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
@@
 regulation will take in the field of encryption depends on the good faith efforts made to equitably balance all the
-benefits and risks involved in deploying a cryptographic system. If the technical community acts, it can not only be
-prepared for regulation, but lead and shape it.
+benefits and risks involved in deploying a cryptographic system. If the technical community acts, it can lead and shape
+legislation instead of being subjected to it.
 
@@
 
-For all these reasons, it is time that the technical community moves the debate forward.
+For these reasons, it is time that the technical community moves the debate forward.
 
@@
 
-The paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
+This paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
 brief U.S.-focused history of encryption regulation, technology, and conflict, it introduces argument maps and threat

### 2020-04-12 Thesis-Chapter2-Background.tex
Incorporate feedback from Jesse
https://github.com/Jesse-Millwood -- Thanks!

@@
 
-The problem of strong encryption and exceptional access exists in a technical, historical, and regulatory context. This
-chapter introduces cryptography, summarizes the history of encryption and the crypto wars, lists relevant encryption
+The problem of strong encryption and \acl{EA} exists in a technical, historical, and regulatory context. This chapter
+introduces cryptography, summarizes the history of encryption and the ``crypto wars,'' lists relevant encryption
 regulations, and lists prominent technical \ac{EA} proposals. It then introduces argument maps as a logical modeling
@@
 processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
-cipher ensures that only those with the correct secret information, the \ac{key}, can perform encryption or decryption
-on the text.
+cipher ensures that only those parties with the correct secret information, the \ac{key}, can perform encryption or
+decryption on the text.
 
@@
 and a blossoming computer industry foundationally reliant on cryptography--came to a head in what has become known as
-\ac{the-cw1}.
+the first ``crypto war.''
+% \ac{the-cw1}.
 
@@
 
-The following sections summarize the encryption regulatory environment and technical approaches to exceptional access.
+The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.
 
@@
 
-In order to advance the \ac{EA} debate, it helps to organize the all the arguments used in a cohesive manner. One way to
+In order to advance the \ac{EA} debate, it helps to organize all the arguments used in a cohesive manner. One way to
 organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and

### 2020-04-15 Thesis-Chapter1-Introduction.tex
Write some more; add international EA resource

@@
 enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
-resistance to a clear solution, the debate on \ac{EA} refuses to end. This paper will not attempt to end the debate, but
-to inform it with additional structure.
+resistance to a clear solution, the debate on \ac{EA} refuses to end. This thesis will not attempt to end the debate,
+but to inform it with additional structure.
 
@@
 
-The conflict at the heart of the encryption and \ac{EA} is that the same cryptographic and design principles that
+The conflict at the heart of the encryption and \ac{EA} debate is that the same cryptographic and design principles that
 underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
@@
 is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
-privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. In a
+privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. 
+% TODO: citation
+In a
 society that cares about bringing wrongdoers to justice, this risk should not be ignored.
@@
 The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
-access encrypted data to perform their duties. Human rights groups and technical leaders counter that a weakened
-encryption environment would fatally compromise privacy and security as we know it. As the debate continues, sides have
+access encrypted data to perform their duties. 
+% TODO: citation
+Human rights groups and technical leaders counter that a weakened
+encryption environment would fatally compromise privacy and security as we know it. 
+% TODO: citation
+As the debate continues, sides have
 become entrenched, and both sides have engaged in disingenuous attacks against their perceived opponents.
@@
 
-This paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief U.S.-focused history of encryption regulation, technology, and conflict, it introduces argument maps and threat
-models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
-creates a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and
-\acl{DIM}. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
-and threats.
+This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
+brief U.S.-focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
+models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
+a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
+thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.
+
+% TODO: use first person voice? Originally had wording like "this thesis introduces", but Kalafut says not to be afraid
+% of "I". Could also use passive voice.

### 2020-04-15 Thesis-Chapter2-Background.tex
Write some more; add international EA resource

@@
 encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
-securely. In \ac{asym-crypto}, encryption and decryption are performed with separate, paired keys, called the public key
-and the private key. The innovation of \ac{public-key-crypto} is that the public key is not secret, and using this
-technique, two parties can communicate securely without requiring an a previously agreed upon shared secret.
+securely. In \ac{asym-crypto}, also called \ac{public-key-crypto}, encryption and decryption are performed with two
+paired keys, called the public key and the private key. The innovation of \ac{public-key-crypto} is that the public key
+is not secret, and using this technique, two parties can communicate securely without requiring an a previously agreed
+upon shared secret.
 
@@
 The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
-problem, as digital technology has little respect for borders. This paper is U.S.-focused, but briefly touches on the
+problem, as digital technology has little respect for borders. This thesis is U.S.-focused, but briefly touches on the
 regulatory environment of other countries.
@@
 \lawsstart
+
     \law{1789}{\acf{AWA}}{congress_1789}{
@@
 
+% IDEA: go into
+%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
+%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
+%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
+% etc?
+
     \law{1976}{\acf{AECA}}{morgan_hr13680_1976}{
@@
 
-}
-
 % (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
@@
 
-    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{}
+}
 
-% \cite{wolff_computer_2016}
+    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
 
-    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{}
+% FIXME: is this actually related enough to keep?
 
-    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{}
+Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
+violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
+the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
+% --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
+whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
+\cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
+activities.
 
-    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{}
+}
 
-    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{}
+    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{
 
-    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{}
+% Amends Wiretap Act, Stored Communications Act, Pen Register Act. Deep history, but picking up here.
 
-    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{}
+}
+
+    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{
+
+}
+
+    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{
+
+}
+
+    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
+
+}
+
+    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{
+
+}
+
+    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{
+
+}
 
@@
 \lawsstart
-    \law{1993}{The \ac{clipperchip}}{press_1993}{}
+
+    \law{1993}{The \ac{clipperchip}}{press_1993}{
+
+}
 % \cite{thompson_2015}
-    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{}
-    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{}
+
+    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{
+
+}
+
+    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{
+
 % mix of SE protection and EA guidelines?
-    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}{}
+
+}
+
+    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}
+        {ashcroft_s2067_1998}{
+
 % like above?
-    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{}
-    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{}
+
+}
+
+    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{
+
+}
+
+    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{
+
 % ^^ these two introduced by same rep on same day
-    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{}
-    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{}
+
+}
+
+    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{
+
+}
+
+    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{
+
 % federal SE protections
-    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{}
+
+}
+
+    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{
+
 % state SE protections
-    \law{2020}{The EARN IT Act}{graham_s3398_2020}{}
+
+}
+
+    \law{2020}{The EARN IT Act}{graham_s3398_2020}{
+
+}
+
 \lawsend
@@
 Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
-down \acp{wicked-prob} and document the reasoned approach that goes into solving them \cite{kunz_issues_1970}. Interest
+down \acp{wicked-prob} and document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest
 in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
@@
 
-This paper uses Argdown to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the basic
-structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be addressed
-through argument. Statements and arguments may support or attack one another, and arguments support or undercut other
-arguments.
+% TODO: reference forward chapter
+Argdown is used in this thesis to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the
+basic structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be
+addressed through argument. Statements and arguments may support or attack one another, and arguments support or
+undercut other arguments.
 
@@
 not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
-primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. This paper creates \acp{DFD} in this style
-to analyze the threats already present and those that \ac{EA} would introduce.
+primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in this
+style to analyze the threats already present and those that \ac{EA} would introduce.
+% TODO: reference forward chapter
 

### 2020-04-15 scripts/show-source.sh
Scripts: with 'show -a', list number of matches

@@ -7,7 +7,7 @@ cd $READING_DIR
 
 if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
     # Show all
-    for FILE in $(eval "grep -riIsl \"${@:2}\""); do
+    for FILE in $(eval "grep -riIsc \"${@:2}\" | grep -v :0 | sort -t: -n -k2 -r"); do
         readlink -f $FILE
     done
     exit 0

### 2020-04-15 Thesis-Chapter2-Background.tex
Add stuff on US laws

@@
 If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
-war, and increased government surveillance set events in motion toward conflict, it was the 2013 Snowden revelations
-that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
-caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
-public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
-technical solution. Under pressure from both the human rights field and from ordinary users demanding improved digital
-privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and \acl{E2E} encryption
-for \acl{DIM} \cite{treguer_us_2018}.
+war, and increased government surveillance set events in motion toward conflict, it was the 2013 \ac{snowden}
+revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
+wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
+caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
+promised a technical solution. Under pressure from both the human rights field and from ordinary users demanding
+improved digital privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and
+\acl{E2E} encryption for \acl{DIM} \cite{treguer_us_2018}.
 
@@
 
-The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
-their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
-basic digital security. Recall encryption and cryptography's critical role in security as described in
+The U.S. corporate response to \ac{snowden} should not be overly construed as a morally motivated defense of civil
+rights, as their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the
+pursuit of basic digital security. Recall encryption and cryptography's critical role in security as described in
 \mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
@@
 
-% Amends Wiretap Act, Stored Communications Act, Pen Register Act. Deep history, but picking up here.
+Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
+electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
+access to certain data as well; some terminology has aged poorly with technological evolution, making the access
+increasingly permissive \cite{shamsi_2011}.
 
@@
 
-}
-
-    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{
+Mandated that telecommunications carriers enable the government to ``intercept all of the subscriber's wire and
+electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized to
+be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant that
+the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
+specifically excludes ``information services'' from its interception requirement which has included internet platforms,
+though that exclusion is currently under attack \cite{pfefferkorn_2019}.
 
@@
 
-    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
+    \law{2001}{\ac{USAPATRIOTAct}}{sensenbrenner_2001}{
+
+Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers
+in 25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope
+(Section 206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business
+records, but the \ac{snowden} leaks revealed that the government had been secretly using a non-standard definition of
+``business records'' in order ``to justify requests for domestic telephone metadata delivered in bulk, not
+individualized requests'' \cite{landau_making_2013}.
 
@@
 
+%     \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
+%
+% Retroactively approved of GWB's warrentless wiretapping policy, a loose intepretation of FISA
+% \cite{landau_making_2013}
+%
+% }
+
     \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{
+
+Like the \ac{USAPATRIOTAct} before it, \acs{FISAAA} expanded \ac{FISA} powers once again. It additionally provided
+``telephone companies retroactive immunity for participating in the warrantless surveillance'' on any international
+communication, a policy established in 2001 under another loose, secret interpretation of \ac{FISA}, ``as well as
+creating prospective immunity for FISAAA activities'' \cite{landau_making_2013}.
 
@@
 
-    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{
+    \law{2015}{\ac{USAFREEDOMAct}}{sensenbrenner_2015}{
+
+The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the senate
+almost two years to the day after \ac{snowden}'s leaks began. It contains many reforms to \ac{FISA} and the use of
+\acp{NSL}. Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass
+collection of telephone metadata.
 
@@
 % state SE protections
+
+}
+
+\law{2019}{Technology in Criminal Justice Act}{demings_hr5227_2019}{
+
+% state SE protections
+
+% pfefferkorn_2019
+% And in fact, there’s already proposed legislation based on CSIS’s recommendations: the Technology in Criminal Justice
+% Act of 2019, introduced in the House last month by a bipartisan group of co-sponsors led by Rep. Val Demings of Florida.
+% The bill seeks to close the resource, training, and information-sharing gap for digital evidence, and to improve the
+% frayed relationship between law enforcement and technology companies. Notably absent from the bill is any language that
+% would roll back the CALEA compromise or mandate encryption backdoors. To be clear, I’m not advocating for or against
+% Rep. Demings’ bill. I just want to point out that it offers an alternative to whatever backdoor nightmare the Senate
+% Judiciary Committee is surely drafting up right now.
 

### 2020-04-15 Thesis-Glossary.tex
Add stuff on US laws

--- 2020-04-08 Thesis-Glossary.tex
@@
 }
+
+\newterm{snowden}{Snowden}{
+    Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
+    mass collection of internet communications for surveillance
+}
 % End terms
@@
 
+% Custom acronyms
+
+\newabbreviation{USAPATRIOTAct}{USA PATRIOT Act}{Uniting and Strengthening America by Providing Appropriate Tools Required to Intercept and Obstruct Terrorism Act}
+\newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-16 Thesis-Chapter2-Background.tex
Work on proposed US laws

@@
 
+    % TODO: working here
+The \ac{clipperchip} was a voluntary initiative from the White House with the goal of ``providing the public
+with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to access
+unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography in
+security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
+reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
+discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
+died.
+
 }
@@
 % ^^ these two introduced by same rep on same day
+
+}
+
+    \law{1999}{Cyberspace Electronic Security Act}{administration_CESA_1999}{
+
+% Bill by the White House
+
+}
+
+    \law{1999}{Enhancement of Privacy and Public Safety in Cyberspace Act}{leahy_s3083_2000}{
+
+% Bill _on behalf of_ the White House; watered down CESA, mostly CFAA updates
 

### 2020-04-18 Thesis-Chapter1-Introduction.tex
Formatting improvements

@@
 \newcommand{\principlesstart}{\begin{enumerate}}
-\newcommand{\principle}[1]{\item \textbf{#1.}}
+\newcommand{\principle}[2]{\item \textbf{#1.} #2}
 \newcommand{\principlesend}{\end{enumerate}}
@@
 
-\principle{Cybersecurity is critical} With current levels of reliance on computer information systems, cybersecurity is
+\principle{Cybersecurity is critical}{With current levels of reliance on computer information systems, cybersecurity is
 critical to the wellbeing of society. Two important elements of cybersecurity are cryptography and system architecture.
-Policy which supports security does not undercut cryptographic integrity or require high-risk architectures.
+Policy which supports security does not undercut cryptographic integrity or require high-risk architectures.}
 
-\principle{Absolute privacy is not an absolute right} Certain rights supersede the legitimate claims of government, but
+\principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
 absolute privacy is not one of them. In some contexts, absolute privacy is proper, and in almost all contexts,
@@
 to some classes of data. Access to these classes of data is important today and will be increasingly important in the
-future.
+future.}
 
-\principle{\Ac{EA} is an inherently complex problem} The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
+\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
 proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
-potential for abuse, international reception, economic impact, and need for transparency.
+potential for abuse, international reception, economic impact, and need for transparency.}
 
-\principle{Perfection is not the standard} Wicked policy problems do not have perfect solutions, or even verifiably
+\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
 optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
 destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
-right balance of risk given the threats under consideration--which includes regulatory action.
+right balance of risk given the threats under consideration--which includes regulatory action.}
 

### 2020-04-18 Thesis-Chapter2-Background.tex
Formatting improvements

@@
 
+% IDEA: Create an appendix to go into greater detail?
+% Could go into these, as well as the commented-out ones below
+%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
+%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
+%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
+
 \subsection{U.S. Efforts to Regulate Encryption}
@@
 \newcommand{\law}[4]{ % Year, name, citation, description
-    \item #1: \textbf{#2} \cite{#3} \newline
-    \parbox{\linewidth}{#4} \newline
+    \item #1: \textbf{#2} \cite{#3}
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#4} \vspace{0.5\baselineskip}
 }
@@
 
-% IDEA: go into
-%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
-%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
-%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
-% etc?
-
     \law{1976}{\acf{AECA}}{morgan_hr13680_1976}{
@@
 
-    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
-
-% FIXME: is this actually related enough to keep?
-
-Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
-violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
-the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
-% --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
-whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
-\cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
-activities.
-
-}
+%     \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
+% Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
+% violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
+% the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
+% % --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
+% whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
+% \cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
+% activities.
+% }
 
@@
 %     \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
-%
-% Retroactively approved of GWB's warrentless wiretapping policy, a loose intepretation of FISA
-% \cite{landau_making_2013}
-%
+% Retroactively approved of GWB's warrantless wiretapping policy, a loose interpretation of FISA
+% \cite{landau_making_2013}.
 % }
@@
 }
-% \cite{thompson_2015}
 
@@
 
-    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{
+%     \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{
+% % mix of SE protection and EA guidelines?
+% }
 
-% mix of SE protection and EA guidelines?
+%     \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}
+%         {ashcroft_s2067_1998}{
+% % like above?
+% }
 
-}
+%     \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{
+% }
 
-    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}
-        {ashcroft_s2067_1998}{
-
-% like above?
-
-}
-
-    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{
-
-}
-
-    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{
-
-% ^^ these two introduced by same rep on same day
-
-}
+%     \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{
+% % ^^ these two introduced by same rep on same day
+% }
 
@@
 
-    \law{1999}{Enhancement of Privacy and Public Safety in Cyberspace Act}{leahy_s3083_2000}{
+    \law{2000}{Enhancement of Privacy and Public Safety in Cyberspace Act}{leahy_s3083_2000}{
 
@@
 
-\law{2019}{Technology in Criminal Justice Act}{demings_hr5227_2019}{
-
-% state SE protections
-
-% pfefferkorn_2019
-% And in fact, there’s already proposed legislation based on CSIS’s recommendations: the Technology in Criminal Justice
-% Act of 2019, introduced in the House last month by a bipartisan group of co-sponsors led by Rep. Val Demings of Florida.
-% The bill seeks to close the resource, training, and information-sharing gap for digital evidence, and to improve the
-% frayed relationship between law enforcement and technology companies. Notably absent from the bill is any language that
-% would roll back the CALEA compromise or mandate encryption backdoors. To be clear, I’m not advocating for or against
-% Rep. Demings’ bill. I just want to point out that it offers an alternative to whatever backdoor nightmare the Senate
-% Judiciary Committee is surely drafting up right now.
-
-}
+% \law{2019}{Technology in Criminal Justice Act}{demings_hr5227_2019}{
+% % pfefferkorn_2019
+% % And in fact, there’s already proposed legislation based on CSIS’s recommendations: the Technology in Criminal Justice
+% % Act of 2019, introduced in the House last month by a bipartisan group of co-sponsors led by Rep. Val Demings of Florida.
+% % The bill seeks to close the resource, training, and information-sharing gap for digital evidence, and to improve the
+% % frayed relationship between law enforcement and technology companies. Notably absent from the bill is any language that
+% % would roll back the CALEA compromise or mandate encryption backdoors. To be clear, I’m not advocating for or against
+% % Rep. Demings’ bill. I just want to point out that it offers an alternative to whatever backdoor nightmare the Senate
+% % Judiciary Committee is surely drafting up right now.
+% }
 

### 2020-04-18 Thesis-Chapter2-Background.tex
Finish section on US regs

@@
 the first ``crypto war.''
-% \ac{the-cw1}.
 
@@
 in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
-reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
+reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher
 discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
@@
 various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
-it was S.1726, the Pro-C0DE Act of 1996, to abolish the export controls on encryption software \cite{burns_s1726_1996}.
-In the House, it was HR.3011, the SAFE Act of 1996, to explicitly allow arbitrarily strong encryption for all legal
-activity \cite{goodlatte_hr3011_1996}. Indeed, the SAFE Act went further, including ``provisions that would have barred
-the government from creating a mandatory key escrow system and would also have removed export restrictions on most
-generally available software and hardware containing encryption'' \cite{kehl_right_2015}. Zimmermann, the author of PGP,
-who had by this time endured an investigation by the Customs Service for publishing his work, testified before the
-Senate in favor of the Pro-C0DE Act \cite{zimmermann_1996}. In November of that year, the Clinton administration
-released an executive order effectively removing export controls on strong encryption products, along the lines of the
-Pro-C0DE Act \cite{clinton_1996}.
+it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
+\cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
+encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went further, including
+``provisions that would have barred the government from creating a mandatory key escrow system and would also have
+removed export restrictions on most generally available software and hardware containing encryption''
+\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
+Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
+November of that year, the Clinton administration released an executive order effectively removing export controls on
+strong encryption products, along the lines of the \ac{PROCODEAct} \cite{clinton_1996}.
 
-From 1996 to 1999, the SAFE Act was proposed several times, had hearings, and gained support \cite{kehl_right_2015}. By
-this time, there was ``an overwhelming amount of evidence against moving ahead with any key escrow schemes''
-\cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all SAFE Act
+From 1996 to 1999, the \ac{SAFEAct} was proposed several times, had hearings, and gained support \cite{kehl_right_2015}.
+By this time, there was ``an overwhelming amount of evidence against moving ahead with any key escrow schemes''
+\cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all \ac{SAFEAct}
 proposals \cite{kehl_right_2015}. This development marks the end of \ac{the-cw1}.
@@
 Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
-administration changed its encryption policy, the U.S. House dropped the SAFE Act \cite{kehl_right_2015}. Also un-passed
-were the related Pro-C0DE Act, Encrypted Communications Privacy Act (not to be confused with the \textit{Electronic}
-Communications Privacy Act \acs{ECPA} of 1986), and Encryption Protects the Rights of Individuals from Violation and
-Abuse in CYberspace (E-PRIVACY) Act \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998}. This legislative failure meant
-that rather than being written in the pen of law, encryption policy was written in the pencil of executive order.
+administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
+\ac{PROCODEAct} and other encryption-related bills went un-passed as well \cite{burns_s1726_1996} \cite{leahy_s376_1997}
+\cite{ashcroft_s2067_1998} \cite{goss_hr2616_1999} \cite{goss_hr2617_1999} \cite{administration_CESA_1999}
+\cite{leahy_s3083_2000}. This legislative failure meant that rather than being written in the pen of law, encryption
+policy was written in the pencil of executive order.
+
+% TODO: add reference to future, comprehensive Appendix of all proposed encryption laws
 
@@
 like ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020, the
-Republican atop the Senate Judiciary Committee introduced S.3398, the EARN IT Act of 2020, which seeks to establish a
+Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to establish a
 16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
-committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for
-child pornography \cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
+committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child pornography
+\cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
 \cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
-largely see the EARN IT Act as an attempt to indirectly yet effectively outlaw \ac{E2E} encryption \cite{newman_2020}
+largely see the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
 \cite{pfefferkorn_2020}.
@@
 \newcommand{\law}[4]{ % Year, name, citation, description
-    \item #1: \textbf{#2} \cite{#3}
+    \item #1: \textbf{#2} \cite{#3} \nopagebreak
 
@@
 
-The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the senate
+The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the Senate
 almost two years to the day after \ac{snowden}'s leaks began. It contains many reforms to \ac{FISA} and the use of
@@
 
-    % TODO: working here
-The \ac{clipperchip} was a voluntary initiative from the White House with the goal of ``providing the public
-with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to access
-unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography in
-security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
-reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When security researcher Matt Blaze
-discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
-died.
+As described in \mysec{sec-history-cw1}, the \ac{clipperchip} was a voluntary \ac{EA} initiative from the White House.
+The initiative targeted \ac{DIM} and established the Escrowed Encryption Standard, which includes an encrypted copy of
+the session key in the message \cite{blaze_protocol_1994}. The initiative received a mixed reception and prompted much
+debate \cite{kehl_right_2015}. When a security researcher discovered flaws that allowed users to subvert the
+\ac{clipperchip} mechanisms, the proposal died \cite{blaze_protocol_1994}.
 
@@
 
-    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{
+    \law{1996}{The \acf{SAFEAct}}{goodlatte_hr3011_1996}{
+
+The \ac{SAFEAct} would have broadly protected ``use any encryption regardless of the encryption algorithm selected,
+encryption key length chosen, or implementation technique or medium used,'' excepting ``the unlawful use of encryption
+in furtherance of a criminal act.'' It also would have lifted export controls. The bill grew support as it was
+reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's policies in
+1999 \cite{kehl_right_2015}.
+
+}
+
+    \law{1996}{The \acf{PROCODEAct}}{burns_s1726_1996}{
+
+The Senate's \ac{PROCODEAct} was very similar to House's \ac{SAFEAct}, though it focused primarily on removing export
+controls, and used slightly weaker language protecting all forms of encryption. Like the \ac{SAFEAct}, it was abandoned
+when the Clinton administration adopted the act's policies in 1996 \cite{thompson_2015}.
 
@@
 
-    \law{1999}{Cyberspace Electronic Security Act}{administration_CESA_1999}{
+    \law{1999}{\ac{CESA}}{administration_CESA_1999}{
 
-% Bill by the White House
+\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
+led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely--encryption is an
+important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
+child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
+fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
+``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
+a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The legislation never got a legislative
+sponsor to be officially considered.
 
@@
 
-% Bill _on behalf of_ the White House; watered down CESA, mostly CFAA updates
+After \ac{CESA} failed, the White House proposed watered-down legislation that this time was successfully sponsored by a
+senator. This bill would have amended the \ac{ECPA} and \ac{CFAA}. It would have made relatively modest changes to
+government data access requirements. Though it received congressional sponsorship, even the sponsor was critical of
+parts of the bill, and it died in committee \cite{senate_2000}.
 
@@
 
+Though this bill was not officially proposed, it represents one of the first legislative attempts of \ac{the-cw2} to
+regulate encryption. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
+Apple to unlock the phone. Though it would not mandate a technical mechanism, the bill would have mandated that
+manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
+proposal in \ac{the-cw1}. Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the bill's authors
+both used the phrase ``above the law'' in promotion of the bill.
+
 }
@@
 
-% federal SE protections
+Proposed in reaction to the Compliance with Court Orders Act and anti-encryption sentiment, this bill would have
+provided protections against ``mandating the deployment of vulnerabilities in data security technologies'' at the
+federal level. It stopped short of explicitly protecting all forms of encryption.
 
@@
 
-    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{
+    \law{2019}{The \ac{ENCRYPTAct}}{lieu_hr4170_2019}{
 
-% state SE protections
+This bill is designed to ``preempt State data security vulnerability mandates and decryption requirements,'' being a
+state-level version of the Secure Data Act. This bill goes further to explicitly disallow \ac{EA} requirements in state
+law.
+
+% TODO: as of 4/18/20, this bill is still active. Follow up.
 
@@
 
-    \law{2020}{The EARN IT Act}{graham_s3398_2020}{
+    \law{2020}{The \acf{EARNITAct}}{graham_s3398_2020}{
+
+Introduced by the chair of the Senate Judiciary Committee and one of the outspoken encryption critics during the 2019
+Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct}, designed to address \ac{CSAM}, has received attention as a
+threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the bill
+would revoke online platforms' liability protection for user-submitted content if is does not doing enough to combat
+\ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one popular
+\ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass \cite{lund_2020}.
+
+% TODO: as of 4/18/20, this bill is still active. Follow up.
 

### 2020-04-18 Thesis-Glossary.tex
Finish section on US regs

@@
 \newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
+\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
 \newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
@@
 \newabbreviation{USAFREEDOMAct}{USA FREEDOM Act}{Uniting and Strengthening America by Fulfilling Rights and Ensuring Effective Discipline Over Monitoring Act}
+\newabbreviation{SAFEAct}{SAFE Act}{Security and Freedom Through Encryption Act}
+\newabbreviation{PROCODEAct}{Pro-C0DE Act}{Promotion of Commerce On-Line in the Digital Era Act}
+\newabbreviation{ENCRYPTAct}{ENCRYPT Act}{Ensuring National Constitutional Rights for Your Private Telecommunications Act}
+\newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
 

### 2020-04-18 Thesis-Chapter2-Background.tex
Work on international regs

@@
 
-The following is a very brief account of the regulatory environment in various jurisdictions around the world:
+This thesis is U.S.-focused, but \ac{EA} technologies and policies have international impacts. The following is a very
+brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
+by the Library of Congress \cite{acosta_government_2016}.
 
 \newcommand{\countriesstart}{\begin{itemize}}
-\newcommand{\country}[1]{\item \textbf{#1:}} % Country
+\newcommand{\country}[3]{ % Name, citation, description
+    \item \textbf{#1} \cite{#2} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#3} \vspace{0.5\baselineskip}
+}
 \newcommand{\countriesend}{\end{itemize}}
@@
 \countriesstart
-    \country{Australia} In 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
-    \country{UK} Investigatory Powers Act \cite{legislature_2016}
-    \country{EU} GDPR \cite{parliament_2016}
-    \country{China} See \cite{donahue_comparative_2018}
-    \country{Israel} See \cite{donahue_comparative_2018}
-    \country{Japan, Korea} can't find anything
-    \country{General} See \cite{budish_encryption_2018}
+
+    \country{Five-Eyes Countries}{acosta_government_2016}{
+% UK, Australia, Canada
+% UK Investigatory Powers Act \cite{legislature_2016}
+% Australia In 2018 passed its Assistance and Access bill \cite{ag_2018}
+
+}
+
+    \country{European Countries}{acosta_government_2016}{
+% France, Belgium, Germany, Sweden, EU
+% EU GDPR \cite{parliament_2016}
+
+}
+
+    \country{Asian Countries}{acosta_government_2016}{
+% China, Japan, Taiwan
+% China Counterterrorism Law \cite{donahue_comparative_2018}
+
+}
+
+    \country{Other Countries}{acosta_government_2016}{
+% Brazil, Israel, South Africa
+% Israel cybersecurity forensics and parnerships \cite{donahue_comparative_2018}
+
+}
+
 \countriesend
+
+% For more, read donahue_comparative_2018 (Israel and China) and budish_encryption_2018 (framework for analysis
 

### 2020-04-19 Thesis-Chapter2-Background.tex
Work on international regs

@@
 
+This section includes both successful and failed cryptography regulation attempts in the U.S.
+
+\subsubsection{Enacted U.S. Regulations}
+
 The following laws and regulations have been applied to cases involving surveillance, cryptography, and access to
@@
 \lawsend
+
+\subsubsection{Failed U.S. Regulation Attempts}
 
@@
 brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
-by the Library of Congress \cite{acosta_government_2016}.
+by the Law Library of Congress \cite{acosta_government_2016}.
 
@@
 
+% What are the most relevant questions? Can the government--
+% - Compel access to--
+%   - all data?
+%   - data encrypted by a service provider? Who are covered service providers?
+% - Compel decryption of--
+%   - all data?
+%   - data encrypted by a service provider? Who are covered service providers?
+% - Perform mass collection?
+% - What legal limits are in place (e.g. warrants)?
+% - What oversight is in place (e.g. inspectors general, public transparency)?
+% - Does the government prescribe technical approaches?
+% - Is the government explicitly allowed to engage in hacking?
+
 \countriesstart
 
-    \country{Five-Eyes Countries}{acosta_government_2016}{
-% UK, Australia, Canada
-% UK Investigatory Powers Act \cite{legislature_2016}
-% Australia In 2018 passed its Assistance and Access bill \cite{ag_2018}
+    \country{Five Eyes Countries}{acosta_government_2016}{
+
+In 1946 The U.S. formalized an intelligence sharing operation with the U.K. in the BRUSA (now UKUSA) Agreement; over the
+following 10 years, Australia, Canada, and New Zealand were made full partners to this agreement, which has become known
+as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
+these countries, the member's encryption policies are relevant to one another.
+
+% IDEA: put this in an appendix too? Or its own chapter? Too much detail for the background section.
+%
+% Looking at it now, even what I have seems like too much. Maybe move it all to an appendix.
+%
+% For more, read donahue_comparative_2018 (Israel and China) and budish_encryption_2018 (framework for analysis)
+%
+% Since the enactment of the Regulation of Investigatory Powers Act 2000, the U.K. government had the ability to serve
+% written notices requiring mandated decryption. Law enforcement and intelligence agencies also have explicit permission
+% to perform ``computer network exploitation,'' and approach this thesis refers to as ``\ac{lawful-hacking}.'' 2016's
+% Investigatory Powers Act expanded these capabilities and, among other additions, in section 253 created ``technical
+% capability notices'' that the Secretary of State can issue to telecommunications operators mandating \ac{EA}
+% \cite{legislature_2016}.
+
+The U.K. government has the explicit ability to force decryption, perform \ac{lawful-hacking}, and mandate \ac{EA}
+though ``technical capability notices'' introduced in 2016's Investigatory Powers Act \cite{legislature_2016}. Australia
+has weaker powers through a variety of laws, and in 2018 passed the Assistance and Access Act which created very similar
+``technical capability notices'' \cite{ag_2018}. Canadian law requires cooperation between telecommunications providers
+and law enforcement but does not directly address encryption. Canada has traditionally supported strong encryption, but
+the pressure of \ac{the-cw2} is causing strain \cite{parsons_2019}.
 
@@
 
+% TODO: it is very important to know who these requirements apply to. Traditional telecoms, obviously. Facebook and
+% Google? Signal and WhatsApp?
+
+France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
+France, Belgium, and Germany, that includes decrypting network traffic when possible. For Belgium, that decryption
+capability is mandatory as of 2011. France, Germany, and Sweden each have some level of \ac{lawful-hacking} capability.
+There is no E.U. legislation that requires key disclosure or decryption of network traffic.
+
 }
@@
     \country{Asian Countries}{acosta_government_2016}{
-% China, Japan, Taiwan
-% China Counterterrorism Law \cite{donahue_comparative_2018}
+
+Japanese authorities require cooperation and can request access to decrypted data, but subjects are not punished for
+declining such requests. Taiwanese regulation does not specifically address encryption, but does mandate that
+telecommunications providers provide interfaces ``with functions that can cooperate with interception.'' Chinese
+authorities have great access to plaintext data by virtue of the political structure of the state, but its 2015
+counter-terrorism law explicitly compels access to plaintext data upon request \cite{donahue_comparative_2018}.
+
+% TODO: but you can still get iPhones in China, right... ? Again, scope?
 
@@
     \country{Other Countries}{acosta_government_2016}{
-% Brazil, Israel, South Africa
-% Israel cybersecurity forensics and parnerships \cite{donahue_comparative_2018}
+
+Brazil, South Africa, and Israel require cooperation between telecommunications providers and law enforcement. Brazil
+law does not specifically address encryption, though it does require ``the technological resources necessary to suspend
+telecommunications confidentiality in accordance with the law.'' South Africa mandates decryption when possible. Israeli
+authorities can issue warrants for access to data or use warrantless orders similar to U.S. \ac{FISA} mechanisms. Israel
+also has well developed \ac{lawful-hacking} capabilities and a centralized forensics laboratory
+\cite{donahue_comparative_2018}.
 
@@
 
-% For more, read donahue_comparative_2018 (Israel and China) and budish_encryption_2018 (framework for analysis
-
 \section{Approaches to Exceptional Access}
@@
 % There are several types of approaches:
-% TODO: after defense, create table of these and reference it in \prop lists below
+% TODO: after defense, create table of these and reference it in \prop lists below. This is better than sorting by
+% historical origin.
 % - Allow strong encryption (anyone with the users' private keys can see it)
@@
 
+There are several technical approaches to \ac{EA}. This section breaks down these approaches according to those
+originating from \ac{the-cw1}, \ac{the-cw2}, and those that side-step the issue by using non-technical means to acquire
+plaintext.
+
 \subsection{Historical Approaches}
@@
 \newcommand{\propsstart}[0]{\begin{itemize}}
-\newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
+% \newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
+\newcommand{\prop}[3]{ % Name, citation, description
+    \item \textbf{#1} \cite{#2} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#3} \vspace{0.5\baselineskip}
+}
 \newcommand{\propsend}{\end{itemize}}
@@
 \propsstart
-    \prop{Clipper chip/third-party key escrow}{press_1993}
+    \prop{Clipper chip/third-party key escrow}{press_1993}{}
 % toasted by \cite{blaze_protocol_1994}
-    \prop{Oblivious key escrow}{goos_oblivious_1996}
-    \prop{Partial key escrow}{denning_taxonomy_1996}
+    \prop{Oblivious key escrow}{goos_oblivious_1996}{}
+    \prop{Partial key escrow}{denning_taxonomy_1996}{}
 \propsend
@@
 \propsstart
-    \prop{CLEAR}{ozzie_2018}
-    \prop{Device escrow with HW}{savage_lawful_2018}
-    \prop{Computational, blockchain approaches}{phan_key_2017}
+    \prop{CLEAR}{ozzie_2018}{}
+    \prop{Device escrow with HW}{savage_lawful_2018}{}
+    \prop{Computational, blockchain approaches}{phan_key_2017}{}
 \propsend
@@
 \propsstart
-    \prop{Lawful hacking}{nguyen_lawful_2017}
+    \prop{\Ac{lawful-hacking}}{nguyen_lawful_2017}{}
 % see also \cite{soesanto_2018} \cite{kerr_encryption_2017}
-    \prop{Forced password disclosure}{kerr_encryption_2017}
+    \prop{Forced password disclosure}{kerr_encryption_2017}{}
 % see also \cite{bittenbender_2019}
-    \prop{Alternative paths to plaintext}{kerr_encryption_2017}
+    \prop{Alternative paths to plaintext}{kerr_encryption_2017}{}
 % access device while unlocked; find copies elsewhere

### 2020-04-21 Thesis-Chapter2-Background.tex
Work on EA types

@@
 % There are several types of approaches:
-% TODO: after defense, create table of these and reference it in \prop lists below. This is better than sorting by
-% historical origin.
+% TODO: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
+% properties below, similar to denning_taxonomy_1996
 % - Allow strong encryption (anyone with the users' private keys can see it)
@@
 %   - Split the key and store it multiple places
+%   - See denning_taxonomy_1996
 % - Use some alternative path
@@
 
-There are several technical approaches to \ac{EA}. This section breaks down these approaches according to those
-originating from \ac{the-cw1}, \ac{the-cw2}, and those that side-step the issue by using non-technical means to acquire
-plaintext.
+% Lesson: the problem is _not_ cryptographically storing and splitting key recovery information. It's keeping those keys
+%   safe! If you can sit on recorded data and wait for the keys to get leaked, which they will...
 
-\subsection{Historical Approaches}
+% Moving forward, what are the elements that are needed for a feasible EA system? (1) White-box-secure ways of producing
+% key recovery information. Do the protocols exist? EES failed. Needs to be rock solid. (2) Dispersed key recovery
+% information. (3) A robust, antifragile, inherently secure means of protecting those keys, and perhaps decaying them over
+% time (https://en.wikipedia.org/wiki/Data_degradation).
 
-Historical approaches and proposals out of \ac{the-cw1} that saw various amounts of support.
+There are several technical approaches to \ac{EA}. This section breaks down these approaches according to their basic
+approach to \ac{key-escrow}, to cryptographic recovery without a key, or to side-stepping the issue by using
+non-technical means to acquire plaintext.
 
 \newcommand{\propsstart}[0]{\begin{itemize}}
-% \newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
-\newcommand{\prop}[3]{ % Name, citation, description
-    \item \textbf{#1} \cite{#2} \nopagebreak
+\newcommand{\prop}[2]{ % Name, description
+    \item \textbf{#1} \nopagebreak
 
-    \vspace{0.5\baselineskip} \parbox{\linewidth}{#3} \vspace{0.5\baselineskip}
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
 }
@@
 
+\subsection{Types of Key Escrow}
+
+The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys, or information that can be
+used to derive additional copies. This approach is called \ac{key-escrow}. There are several classes of key escrow.
+
 \propsstart
-    \prop{Clipper chip/third-party key escrow}{press_1993}{}
-% toasted by \cite{blaze_protocol_1994}
-    \prop{Oblivious key escrow}{goos_oblivious_1996}{}
-    \prop{Partial key escrow}{denning_taxonomy_1996}{}
+    \prop{Trusted-party key escrow}{
+
+\cite{denning_taxonomy_1996}
+\ac{clipperchip} \cite{blaze_protocol_1994}
+% The technical approach used by the \ac{clipperchip} secret protocol hardware enabled split-key government escrow.
+
+}
+
+    \prop{Distributed key escrow}{
+
+\cite{goos_oblivious_1996}
+\cite{phan_key_2017}
+
+}
+
+    \prop{Device key escrow}{
+
+\cite{savage_lawful_2018}
+\cite{ozzie_2018}
+
+}
 \propsend
 
-\subsection{Modern Proposals}
+\subsection{Non-Escrow Cryptographic Recovery}
 
-Recent proposals out of \ac{the-cw2} that have various levels of support.
+Though key recovery is the most discussed \ac{EA} approach, alternative technical means have been proposed.
 
 \propsstart
-    \prop{CLEAR}{ozzie_2018}{}
-    \prop{Device escrow with HW}{savage_lawful_2018}{}
-    \prop{Computational, blockchain approaches}{phan_key_2017}{}
+    \prop{Translucent Cryptography}{
+
+\cite{bellare_translucent_1996}
+
+}
 \propsend
@@
 \propsstart
-    \prop{\Ac{lawful-hacking}}{nguyen_lawful_2017}{}
-% see also \cite{soesanto_2018} \cite{kerr_encryption_2017}
-    \prop{Forced password disclosure}{kerr_encryption_2017}{}
-% see also \cite{bittenbender_2019}
-    \prop{Alternative paths to plaintext}{kerr_encryption_2017}{}
+    \prop{\Ac{lawful-hacking}}{
+
+\cite{nguyen_lawful_2017}
+\cite{soesanto_2018}
+\cite{kerr_encryption_2017}
+
+}
+
+    \prop{Forced password disclosure}{
+
+\cite{bittenbender_2019}
+\cite{kerr_encryption_2017}
+
+}
+
+    \prop{Alternative paths to plaintext}{
+
 % access device while unlocked; find copies elsewhere
+\cite{kerr_encryption_2017}
+
+}
 \propsend

### 2020-04-21 Thesis-Glossary.tex
Work on international regs
Work on EA types

@@
 }
+
+\newterm{lawful-hacking}{lawful hacking}{
+    An approach to digital evidence and intelligence gathering that involves deliberate computer and network
+    manipulation which is ordinarily illegal but has been cleared for law enforcement and intelligence purposes
+}
+
+\newterm{key-escrow}{key escrow}{
+    An approach to \ac{EA} that relies on storing additional copies of the encryption key or storing information that
+    can be used to derive additional additional copies
+}
 % End terms

### 2020-04-21 Thesis-Chapter2-Background.tex
Finish EA types brief overview

@@
 % TODO: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
-% properties below, similar to denning_taxonomy_1996
+% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
 % - Allow strong encryption (anyone with the users' private keys can see it)
@@
 
+% TODO: need an aside on shamir m/n secret sharing? Probably in order to understand split key approaches at all
+
 The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys, or information that can be
@@
 
-\cite{denning_taxonomy_1996}
-\ac{clipperchip} \cite{blaze_protocol_1994}
-% The technical approach used by the \ac{clipperchip} secret protocol hardware enabled split-key government escrow.
+Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
+of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
+a wide variety of implementation options within this category \cite{denning_taxonomy_1996}--secret or open protocol,
+hardware-enabled or software-only, single key or split key, government or third-party escrow--the foundation of the
+effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
+a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.
 
@@
 
-\cite{goos_oblivious_1996}
-\cite{phan_key_2017}
+Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
+information. In this arrangement, trust is laid in the distrubuted system, not in any single party. Such systems offer
+high data availability while making it difficult for access requests to be secret or to subvert access policies. This
+approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the application of
+blockchain \cite{phan_key_2017}. These approaches are not as developed as trusted-party \ac{key-escrow}.
 
@@
 
-\cite{savage_lawful_2018}
-\cite{ozzie_2018}
+A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
+trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
+``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
+\ac{DAR}, and include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
+\cite{ozzie_2018}.
 
@@
 \propsstart
-    \prop{Translucent Cryptography}{
+    \prop{Translucent cryptography}{
 
-\cite{bellare_translucent_1996}
+Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
+some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
+scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction--say, 40\%--of plaintext data.
+That fraction would be designed to balance privacy, digital security, and public security. This research, though
+decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
+escrow do exist \cite{bellare_translucent_1996}.
 
@@
 
-\cite{nguyen_lawful_2017}
-\cite{soesanto_2018}
-\cite{kerr_encryption_2017}
+As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Instead of having
+access to plaintext data via an alternative means of decryption, law enforcement is allowed to perform otherwise-illegal
+hacking activity in order to find the key, compromise a device, or intercept plaintext versions of the data
+\cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
+enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
+general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.
 
@@
 
-\cite{bittenbender_2019}
-\cite{kerr_encryption_2017}
+In some cases, a suspect in custody may have the ability to access the desired encrypted information via key, \ac{PIN},
+or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S., this
+approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of course,
+this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.
+
+% these look good: sacharoff_unlocking_2018, fredona_fifth_2016
+% this probably has something to say: kerr_encryption_2017
 
@@
 
-% access device while unlocked; find copies elsewhere
-\cite{kerr_encryption_2017}
+Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
+authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
+enforcement is traditionally good at--investigation and sting operations--and is used successfully. In 2017, an
+international law enforcement operation used this technique to take down a large illicit online drug market
+\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
 

### 2020-04-21 Thesis-Chapter2-Background.tex
CW2 intro reword

@@
 administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
-\ac{PROCODEAct} and other encryption-related bills went un-passed as well \cite{burns_s1726_1996} \cite{leahy_s376_1997}
-\cite{ashcroft_s2067_1998} \cite{goss_hr2616_1999} \cite{goss_hr2617_1999} \cite{administration_CESA_1999}
-\cite{leahy_s3083_2000}. This legislative failure meant that rather than being written in the pen of law, encryption
-policy was written in the pencil of executive order.
+\ac{PROCODEAct} and other encryption-related bills went un-passed as well, as elaborated in \mysec{sec-us-reg-fails}.
+This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
+pencil of executive order.
 
@@
 \subsubsection{Failed U.S. Regulation Attempts}
+\label{sec-us-reg-fails}
 
 The following are proposed laws and regulations that did not go into effect or have yet to be voted on:
+
+% \cite{burns_s1726_1996} \cite{leahy_s376_1997} \cite{ashcroft_s2067_1998} \cite{goss_hr2616_1999}
+% \cite{goss_hr2617_1999} \cite{administration_CESA_1999} \cite{leahy_s3083_2000}
 

### 2020-06-29 Thesis-Chapter2-Background.tex
Add a note on CFAA follow up
Add article on Chinese intelligence law
Resume work!

@@
 
-% FIXME: Kristen doesn't like the wording; also remove word "electronic"?
+% FIXME: Kristen doesn't like the wording
 Encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central conflict in
@@
 %     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
+% May 27: renewal in great doubt; Trump opposed because he sees himself as a victim of FISA abuse
+%     (https://apnews.com/c04a5e3bb156774af83b686bbe2ea872)
 
@@
 % activities.
+%
+% Follow up: CFAA going before supreme court: https://www.cyberscoop.com/cfaa-will-soon-day-supreme-court/
 % }
@@
 \lawsend
+
+\subsubsection{The Judicial Environment}
+\label{sec-us-judicial}
+
+The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
+and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
+judiciary branch exercises power through a framework based on AAAA and BBBB. In this section we analyze several
+important cases and developments in the last few years.
+
+% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998 and gizzi_fourth_2016. Update the
+%   "AAAA and BBBB" in the intro.
+
+% TODO: Review historically important cases on the 1st, 4th, and 5th amendments, e.g. Katz v. United States, 1969.
+
+% TODO: Review recent cases. See orenstein_2016, bittenbender_2019, vaas_2019.
 
@@
 counter-terrorism law explicitly compels access to plaintext data upon request \cite{donahue_comparative_2018}.
+
+% TODO: China, see also blanchard_2017
 

### 2020-07-02 Thesis-Chapter1-Introduction.tex
Writing and editing comments

@@
 \cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
-answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or metrics. Each attempt at a solution matters,
-has unintended consequences, and often exists in a pattern of chronic policy failure.
+answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
+permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.
 
@@
 
-For these reasons, it is time that the technical community moves the debate forward.
+For these reasons, it is important that the technical community keeps moving the debate forward.
 
@@
 \principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
-absolute privacy is not one of them. In some contexts, absolute privacy is proper, and in almost all contexts,
-reasonable privacy should be presumed. However, privacy is not an absolute right, and investigators should have access
-to some classes of data. Access to these classes of data is important today and will be increasingly important in the
-future.}
+absolute privacy in all contexts is not one of them. In some contexts, absolute privacy is proper, and in all contexts,
+reasonable privacy should be presumed. However, absolute privacy does not always apply, and investigators should have
+access to some classes of data. Access to certain classes of digital data is important today and will be increasingly
+important in the future.}
 
@@
 proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
-potential for abuse, international reception, economic impact, and need for transparency.}
+potential for abuse, international consequences, economic impact, and need for transparency.}
 
@@
 destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
-right balance of risk given the threats under consideration--which includes regulatory action.}
+right balance of risk given the threats under consideration--including the threat of regulatory action.}
 

### 2020-07-02 Thesis-Chapter2-Background.tex
Writing and editing comments

@@
 \Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
-performed by using some \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. The cipher is the tool or
+performed by using a \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. A cipher is a tool or
 algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
@@
 
-There are two major cryptographic protocol families--symmetric and asymmetric. In traditional \ac{sym-crypto},
-encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
-securely. In \ac{asym-crypto}, also called \ac{public-key-crypto}, encryption and decryption are performed with two
-paired keys, called the public key and the private key. The innovation of \ac{public-key-crypto} is that the public key
-is not secret, and using this technique, two parties can communicate securely without requiring an a previously agreed
-upon shared secret.
+There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was discovered
+long ago, encryption and decryption are performed with the same key, and both parties must have this key in order to
+communicate securely. In \ac{asym-crypto}, which was discovered recently and is also called \ac{public-key-crypto},
+encryption and decryption are performed with two paired keys, called the public key and the private key. The innovation
+of \ac{public-key-crypto} is that the public key is not secret, and using this technique, two parties can communicate
+securely without requiring a previously agreed upon shared secret.
 
@@
 different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
-perform authentication and establish short-lived \ac{sym-crypto} sessions keys; the sessions keys perform the actual
-encryption of the data in transit. Important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}, which
-ensures that a leaked private key or session key does not compromise any other private key or session key, and
-\ac{replay-protection}, which ensures that messages cannot be replayed by an attacker without detection
-\cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for \ac{DIM}.
+perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
+encryption of the data in transit. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}
+and \ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise
+any other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an
+attacker without detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of
+encryption for \ac{DIM}.
 
 \Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
-encryption-time, as \ac{DIM} session keys are, these keys are either derived from a password stored in human memory or
-are stored somewhere in computer memory, which often takes place with the assistance of dedicated hardware.
-\Ac{disk-encryption} for laptops and mobile devices is an example of encryption for \ac{DAR}.
+encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
+somewhere in computer memory, which often takes place with the assistance of dedicated hardware. \Ac{disk-encryption}
+for laptops and mobile devices is an example of encryption for \ac{DAR}.
 
@@
 from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
+% FIXME: reading note: paragraph makes sense, but need to read it multiple times for it to sink in. Improve.
 
@@
 Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
-systems that directly rely on these technologies include the financial system, local and global; industrial control
+systems that directly rely on these technologies include the local and global financial systems; industrial control
 systems and critical infrastructure; health information systems; private, government, and military communications; and
 the integrity of the internet as a whole.
+% FIXME: Jumping rather quickly from primitives to huge systems. Perhaps scale it more smoothly. Connect little
+% primitives to higher technologies, higher technologies to underpinnings of network computer systems, and network
+% computer systems as critical to these industries.
 
@@
 
-The section provides an overview of encryption from ancient to contemporary times.
+The section provides a brief overview of encryption from ancient to modern times.
 
@@
 
-Computerized encryption is a new technology, but the field of cryptography is old, as demand for privacy is as old as
+Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
 communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
@@
 The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
-computing advances in the 1980s, it for the first time put the power of strong encryption in the hands of ordinary
-people. However, the privacy provided by cryptography is a powerful tool, and privacy is a property governments strive
-to regulate. The use of encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was
-convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason
-tried to force testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}.
-Government's natural stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of
-strong encryption.
+computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
+governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
+property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
+Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
+Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
+\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
+rapid increase in availability of strong encryption.
+% FIXME: The point is made--these developments made gov'ts uncomfortable (just look at these examples!)--but the line of
+% thought is not clear enough to the reader.
 
@@
 In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
-the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. In 1991, the
-U.S. Senate introduced, but did not pass, a bill mandating access to plain text contents when authorized by law; in
-response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to secure email, in order that strong
-cryptography be ``made available to the American public before it became illegal to use'' \cite{zimmermann_1996}. In
-1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with the goal of ``providing the
-public with strong cryptographic tools without sacrificing the ability of law enforcement and intelligence agencies to
-access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the foundational role of cryptography
-in security--and the potential of human rights abuses for compromised privacy--industry and prominent technical leaders
-reacted negatively to the initiative \cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher
-discovered flaws that allowed users to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal
-died.
+the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
+shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
+text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
+secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
+use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
+the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
+and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
+foundational role of cryptography in security--and the potential of human rights abuses for compromised
+privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
+\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
+\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.
 
 Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
-various key escrow proposals \cite{thompson_2015}. In 1996, two bills were introduced in the US Congress. In the Senate,
-it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
+various key escrow proposals \cite{thompson_2015}.
+% FIXME: export controls were a hot topic of the time, too. Can't just drop that thread.
+In 1996, two pro-encryption bills were introduced in the US Congress.
+In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
 \cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
@@
 November of that year, the Clinton administration released an executive order effectively removing export controls on
-strong encryption products, along the lines of the \ac{PROCODEAct} \cite{clinton_1996}.
+strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.
 
@@
 been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
-of the September 11 attacks to some justified yet greater investigatory powers \cite{bloss_escalating_2007}. October of
+of the September 11 attacks seemed to justify yet greater investigatory powers \cite{bloss_escalating_2007}. October of
 2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
 \ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
+% FIXME: introduce FISA above, in the "had already been expanding" section. Then we have more of a narrative of
+% expanding power.
 authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
@@
 
-If the weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
-war, and increased government surveillance set events in motion toward conflict, it was the 2013 \ac{snowden}
+The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
+war, and increased government surveillance set events in motion toward conflict, but it was the 2013 \ac{snowden}
 revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
@@
 caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
-promised a technical solution. Under pressure from both the human rights field and from ordinary users demanding
-improved digital privacy, U.S. tech companies responded by introducing default device encryption for \acl{DAR} and
-\acl{E2E} encryption for \acl{DIM} \cite{treguer_us_2018}.
+promised a technical solution. Benefiting from advancements in underlying technologies and under pressure from both the
+human rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing
+default device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.
 
@@
 antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
-that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
-\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
-law enforcement, and cybersecurity, the pursuit of tech professionals. Both sides of the debate have by now acknowledged
-that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public security vs.
-cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.
+% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
+Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
+pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
+now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
+security vs. cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.
 
@@
 phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
-\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto wars,
-terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators jointly
-released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not make it
-to the floor, and the unresolved nature of the court case left the legal debate wide open.
+\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
+wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
+% FIXME: determine correct usage of -- vs ---
+jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
+make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.
 
@@
 terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on child
-pornography is likely a result of the growth of the ``dark web'' and a severe, in-depth 2019 investigation by the
-\textit{New York Times} \cite{keller_internet_2019}.
+pornography is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
+York Times} \cite{keller_internet_2019}.
 
@@
 U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
-like ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020, the
-Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to establish a
-16-chair committee with the power to revoke online platforms' liability protection for user-submitted content if the
-committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child pornography
-\cite{graham_s3398_2020}. As illustrated by an in-depth article by the \textit{New York Times}
+such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
+the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
+establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
+content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
+pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
 \cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
@@
 The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.
+% FIXME: remove this sentence?
 

### 2020-07-04 Thesis-Chapter2-Background.tex
Make more writing and editing improvements

@@
 % FIXME: determine correct usage of -- vs ---
+% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
 jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
@@
 
-The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
-problem, as digital technology has little respect for borders. This thesis is U.S.-focused, but briefly touches on the
-regulatory environment of other countries.
+The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
+U.S.-focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
+problem, as digital technology has little respect for political borders.
 
@@
 \subsubsection{Enacted U.S. Regulations}
+% FIXME: distinguish statutes vs. regulations; rename "failed" section
 
@@
 
-The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: security against
+The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: protection from
 unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.
@@
 Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
+% FIXME: Briefly introduce these. Even just "Amended the switch-board-era"...
 electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
@@
 
-The following are proposed laws and regulations that did not go into effect or have yet to be voted on:
+The following are proposed laws and regulations that did not go into effect or are still in process:
 
@@
 The initiative targeted \ac{DIM} and established the Escrowed Encryption Standard, which includes an encrypted copy of
-the session key in the message \cite{blaze_protocol_1994}. The initiative received a mixed reception and prompted much
-debate \cite{kehl_right_2015}. When a security researcher discovered flaws that allowed users to subvert the
+the session key in the message \cite{blaze_protocol_1994}. The initiative prompted debate but received a largely hostile
+reception \cite{kehl_right_2015}. When a security researcher discovered flaws that allowed users to subvert the
 \ac{clipperchip} mechanisms, the proposal died \cite{blaze_protocol_1994}.
@@
 in furtherance of a criminal act.'' It also would have lifted export controls. The bill grew support as it was
-reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's policies in
-1999 \cite{kehl_right_2015}.
+reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's
+pro-encryption policies in 1999 \cite{kehl_right_2015}.
 
@@
 controls, and used slightly weaker language protecting all forms of encryption. Like the \ac{SAFEAct}, it was abandoned
-when the Clinton administration adopted the act's policies in 1996 \cite{thompson_2015}.
+when the Clinton administration adopted the act's export control policies in 1996 \cite{thompson_2015}.
 
@@
 ``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
-a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The legislation never got a legislative
-sponsor to be officially considered.
+a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
+member of Congress to officially sponsor the bill.
 
@@
 
-Though this bill was not officially proposed, it represents one of the first legislative attempts of \ac{the-cw2} to
-regulate encryption. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
+Though this bill was not officially proposed, it represents one of the first legislative attempts to regulate encryption
+in \ac{the-cw2}. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
 Apple to unlock the phone. Though it would not mandate a technical mechanism, the bill would have mandated that
 manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
-proposal in \ac{the-cw1}. Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the bill's authors
-both used the phrase ``above the law'' in promotion of the bill.
+proposals in \ac{the-cw1}.
+% TODO: how so?
+Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the draft's
+authors both used the phrase ``above the law'' in promotion of the bill.
+% TODO: how does this differ from the new LEAD Act?
 
@@
 
-Introduced by the chair of the Senate Judiciary Committee and one of the outspoken encryption critics during the 2019
-Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct}, designed to address \ac{CSAM}, has received attention as a
-threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the bill
-would revoke online platforms' liability protection for user-submitted content if is does not doing enough to combat
-\ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one popular
-\ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass \cite{lund_2020}.
+Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
+Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
+as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
+bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
+combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
+popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
+\cite{lund_2020}.
 
 % TODO: as of 4/18/20, this bill is still active. Follow up.
+
+}
+
+    \law{2020}{The \acf{LEADAct}}{graham_2020}{
+
+Introduced by the same senator as the \ac{EARNITAct}, the \ac{LEADAct} is a direct attack on encryption. The
+\ac{LEADAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
+developers to implement the ability to decrypt all data stored or passing through the device or software system
+``concurrently with their transmission'' ``unless the independent actions of an unaffiliated entity make it technically
+impossible to do so.'' The act neither mandates nor recommends any technical approach to providing this capability.
+
+% TODO: how does this differ from the old "Compliance with Court Orders Act" draft?
 
@@
 France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
-France, Belgium, and Germany, that includes decrypting network traffic when possible. For Belgium, that decryption
-capability is mandatory as of 2011. France, Germany, and Sweden each have some level of \ac{lawful-hacking} capability.
-There is no E.U. legislation that requires key disclosure or decryption of network traffic.
+France, Belgium, and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each
+have some level of \ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or
+decryption of network traffic.
 
@@
 
-There are several technical approaches to \ac{EA}. This section breaks down these approaches according to their basic
-approach to \ac{key-escrow}, to cryptographic recovery without a key, or to side-stepping the issue by using
-non-technical means to acquire plaintext.
+There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
+basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
+plaintext.
 
@@
 
-The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys, or information that can be
-used to derive additional copies. This approach is called \ac{key-escrow}. There are several classes of key escrow.
+The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
+information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
+classes of key escrow.
 
@@
 Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
-information. In this arrangement, trust is laid in the distrubuted system, not in any single party. Such systems offer
-high data availability while making it difficult for access requests to be secret or to subvert access policies. This
-approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the application of
-blockchain \cite{phan_key_2017}. These approaches are not as developed as trusted-party \ac{key-escrow}.
+information. In this arrangement, trust is put in the distrubuted system, not in any single party. Such systems offer
+high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
+policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the
+application of blockchain \cite{phan_key_2017}. These approaches are not as thoroughly explored as trusted-party
+\ac{key-escrow}.
 
@@
 ``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
-\ac{DAR}, and include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
+\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
 \cite{ozzie_2018}.
@@
 
-As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Instead of having
-access to plaintext data via an alternative means of decryption, law enforcement is allowed to perform otherwise-illegal
-hacking activity in order to find the key, compromise a device, or intercept plaintext versions of the data
-\cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
-enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
-general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.
+As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Under
+\ac{lawful-hacking}, instead of having access to plaintext data via an alternative means of decryption, law enforcement
+is allowed to perform otherwise-illegal hacking activity in order to find the key, compromise a device, or intercept
+plaintext versions of the data \cite{kerr_encryption_2017}. While some see this as a viable middle ground
+\cite{nguyen_lawful_2017}, encouraging law enforcement to exploit vulnerabilities creates mis-aligned incentives. If law
+enforcement opposes strong security in general, it could foster the exploit market while costing ordinary users money
+and security \cite{soesanto_2018}.
 
@@
 
-In some cases, a suspect in custody may have the ability to access the desired encrypted information via key, \ac{PIN},
-or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S., this
-approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of course,
-this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.
+In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
+\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
+this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of
+course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.
 
@@
 
-In order to advance the \ac{EA} debate, it helps to organize all the arguments used in a cohesive manner. One way to
-organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and
-relationships between statements, premises, and conclusions. Argument maps are an outgrowth of the same research by
-Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
-down \acp{wicked-prob} and document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest
-in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
-\cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006} tools designed to produce graphical
-representations. Argdown is a recent tool that generates argument maps from structures specified in a Markdown-based
-language \cite{voigt_argdown_2018}.
+The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
+help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create argument maps.
+Argument maps are graphical representations of the logical structures and relationships between statements, premises,
+and conclusions. Argument maps are an outgrowth of the same research by Horst Rittel that originated the idea of
+\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
+reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
+software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
+\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
+generates argument maps from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.
 
@@
 \Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
-external entities. \Acp{DFD} well suited for threat modeling because security vulnerabilities tend to follow data flow,
-not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
-primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in this
-style to analyze the threats already present and those that \ac{EA} would introduce.
+external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
+flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is
+the primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in
+this style to analyze the threats already present and those that \ac{EA} would introduce.
 % TODO: reference forward chapter

### 2020-07-04 scripts/sourceme.sh

@@ -20,7 +20,7 @@ alias new="$SCRIPTS_DIR/new-source.sh"
 alias show="$SCRIPTS_DIR/show-source.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
-alias todo="(cd $BASE_DIR && grep -n TODO *.tex)"
+alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
 
 alias pdf="xdg-open build/Thesis.pdf"
 alias gvsu="xdg-open build/Thesis-GVSU.pdf"

### 2020-07-04 Thesis-Chapter1-Introduction.tex
Fix usage of hyphen, en dash, and em dash

@@
 destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
-right balance of risk given the threats under consideration--including the threat of regulatory action.}
+right balance of risk given the threats under consideration---including the threat of regulatory action.}
 
@@
 This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief U.S.-focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
+brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
 models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create

### 2020-07-04 Thesis-Chapter2-Background.tex
Fix usage of hyphen, en dash, and em dash

@@
 
-% There are two classes of cryptographic ciphers--stream and block
+% There are two classes of cryptographic ciphers---stream and block
 
@@
 
-One can also define two major cryptographic applications--securing \acf{DIM} and \acf{DAR}. Each application is a
+One can also define two major cryptographic applications---securing \acf{DIM} and \acf{DAR}. Each application is a
 different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
@@
 % Tampering                 Integrity
-% Repudiation               Non- Repudiation
+% Repudiation               Non-Repudiation
 % Information Disclosure    Confidentiality
@@
 rapid increase in availability of strong encryption.
-% FIXME: The point is made--these developments made gov'ts uncomfortable (just look at these examples!)--but the line of
-% thought is not clear enough to the reader.
+% FIXME: The point is made---these developments made gov'ts uncomfortable (just look at these examples!)---but the line
+% of thought is not clear enough to the reader.
 
@@
 
-All these realities--government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
-and a blossoming computer industry foundationally reliant on cryptography--came to a head in what has become known as
+All these realities---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
+and a blossoming computer industry foundationally reliant on cryptography---came to a head in what has become known as
 the first ``crypto war.''
@@
 and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
-foundational role of cryptography in security--and the potential of human rights abuses for compromised
-privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
+foundational role of cryptography in security---and the potential of human rights abuses for compromised
+privacy---industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
 \cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
@@
 administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
-\ac{PROCODEAct} and other encryption-related bills went un-passed as well, as elaborated in \mysec{sec-us-reg-fails}.
+\ac{PROCODEAct} and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}.
 This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
@@
 now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
-security vs. cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.
+security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.
 
@@
 \cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
-wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
-% FIXME: determine correct usage of -- vs ---
+wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
 % (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
@@
 The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
-U.S.-focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
+U.S.--focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
 problem, as digital technology has little respect for political borders.
@@
 % the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
-% % --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
+% % ---what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices---
 % whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
@@
 \Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
-led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely--encryption is an
+led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely---encryption is an
 important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
@@
 
-This thesis is U.S.-focused, but \ac{EA} technologies and policies have international impacts. The following is a very
+This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
 brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
@@
 
-% What are the most relevant questions? Can the government--
-% - Compel access to--
+% What are the most relevant questions? Can the government---
+% - Compel access to---
 %   - all data?
 %   - data encrypted by a service provider? Who are covered service providers?
-% - Compel decryption of--
+% - Compel decryption of---
 %   - all data?
@@
 of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
-a wide variety of implementation options within this category \cite{denning_taxonomy_1996}--secret or open protocol,
-hardware-enabled or software-only, single key or split key, government or third-party escrow--the foundation of the
+a wide variety of implementation options within this category \cite{denning_taxonomy_1996}---secret or open protocol,
+hardware-enabled or software-only, single key or split key, government or third-party escrow---the foundation of the
 effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
@@
 some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
-scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction--say, 40\%--of plaintext data.
-That fraction would be designed to balance privacy, digital security, and public security. This research, though
+scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
+data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
 decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
@@
 authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
-enforcement is traditionally good at--investigation and sting operations--and is used successfully. In 2017, an
+enforcement is traditionally good at---investigation and sting operations---and is used successfully. In 2017, an
 international law enforcement operation used this technique to take down a large illicit online drug market

### 2020-07-06 Thesis-Chapter1-Introduction.tex
Add 4thAm in Flux notes

@@
 \principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
-absolute privacy in all contexts is not one of them. In some contexts, absolute privacy is proper, and in all contexts,
-reasonable privacy should be presumed. However, absolute privacy does not always apply, and investigators should have
-access to some classes of data. Access to certain classes of digital data is important today and will be increasingly
+privacy in all contexts is not one of them. While individuals under a limited government are entitled to an expectation
+of privacy, an absolute right to privacy does not apply in all circumstances. Investigators should have access to some
+classes of data. In particular, access to certain classes of digital data is important today and will be increasingly
 important in the future.}

### 2020-07-10 scripts/sourceme.sh
Add fixme alias

@@ -21,6 +21,7 @@ alias show="$SCRIPTS_DIR/show-source.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
 alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
+alias fixme="(cd $BASE_DIR && grep -nE 'FIXME' *.tex)"
 
 alias pdf="xdg-open build/Thesis.pdf"
 alias gvsu="xdg-open build/Thesis-GVSU.pdf"

### 2020-07-10 Thesis-Chapter2-Background.tex
Add 4thAm in Flux notes
Address a number of 'fixme's; improve writing

@@
 
-Cryptography is the technical foundation for many forms of computer and network security, at every scale. Security is
-often defined in terms of several properties: authentication, integrity, non-repudiation, confidentiality, availability,
-and authorization \cite{shostack_threat_2014}. Security properties are systems properties, but systems properties emerge
-from the composition of their parts, and innumerable primitives and technologies rely on cryptography.
-% FIXME: reading note: paragraph makes sense, but need to read it multiple times for it to sink in. Improve.
+Cryptography is the technical foundation for many forms of computer and network security. Security is defined in terms
+of several properties: authentication, integrity, non-repudiation, confidentiality, availability, and authorization
+\cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
+(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
+architecture of its subcomponents and the security of the subcompenents themselves. At the bottom of this analysis, the
+security of primitive components commonly relies on cryptography.
+
+\mytab{table:security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
+cryptographic foundations, availability indirectly relies on each other property, and authorization schemes are
+typically rooted in authentication.
+
+% TODO: clarify relationship between encryption and cryptography?
 
@@
 % Elevation of Privilege    Authorization
+\begin{table}[h]
+    \caption[Security Properties and Cryptography]{The Cryptographic Basis of Security Properties}
+    \label{table:security-and-crypto}
+    \centering
+    \begin{tabular}{|l|l|}
+        \hline
+        \thead{Property} & \thead{Cryptographic Basis} \\ \hline
+        Authentication  & Digital certificates \\ \hline
+        Integrity       & Cryptographic hashes \\ \hline
+        Non-repudiation & Digital signatures \\ \hline
+        Confidentiality & Encryption \\ \hline
+        Availability    & \small{\ii{based on architecture}} \\ \hline
+        Authorization   & \small{\ii{rooted in authentication}} \\ \hline
+    \end{tabular}
+\end{table}
 
-Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
-signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
-Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
-systems that directly rely on these technologies include the local and global financial systems; industrial control
-systems and critical infrastructure; health information systems; private, government, and military communications; and
-the integrity of the internet as a whole.
-% FIXME: Jumping rather quickly from primitives to huge systems. Perhaps scale it more smoothly. Connect little
-% primitives to higher technologies, higher technologies to underpinnings of network computer systems, and network
-% computer systems as critical to these industries.
+% Cryptographic primitives include symmetric key encryption, symmetric key encryption, one-way hash functions, and digital
+% signatures. A few technologies that directly rely on these primitives include \ac{TLS}, \ac{PKI}, \ac{SSO} (e.g.
+% Kerberos), \acp{VPN}, secure password storage, disk encryption tools, and secure- and trusted-boot processes. A few
+% systems that directly rely on these technologies include the local and global financial systems; industrial control
+% systems and critical infrastructure; health information systems; private, government, and military communications; and
+% the integrity of the internet as a whole.
+% % If I bring this back---this is jumping rather quickly from primitives to huge systems. Scale it more smoothly.
+% % Connect little primitives to higher technologies, higher technologies to underpinnings of network computer systems,
+% % and network computer systems as critical to these industries.
 
-% Let us break security down into three scales: primitives, technologies, and systems. A primitive is a conceptual
-% building block that serves a single technical function. A technology is a set of implemented primitives glued together
-% to provide useful features. A system is a collection of technologies that collectively serve the ultimate outcome.
-% The security properties critical at the systems level depend
+Violations of cryptographic integrity in primitive components, introduced by design or by accident, could have
+catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
+information disclosure done both for their own sake and as tools in general attacks. The importance of taking these
+tools away is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
+disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
+\cite{goodin_ransomware_2019}.
 
-% Primitives: one-way hashes enable integrity checks, block ciphers enable efficient confidential data storage, and
-% digital signature enable non-repudiation
+% NOTE: tried to find examples of high profile bugs or hacks that relied on crypto failures, but most are still
+% architectural (exposed databases, default credentials (Mirai botnet)), human (BEC), or insecure code (buffer
+% overflows--key to EternalBlue (wannacry and notpetya) and MS08-067 (Conficker)).
 
-% Technologies: Transport Layer Security (TLS) enables confidentiality and integrity over the internet, Kerberos enables
-% convenient authentication, and virtual private networks (VPNs) enable confidential and authenticated remote network
-% connections
-
-If the cryptographic primitives that compose these systems were to be compromised, the systems they comprise could
-collapse. If the financial system loses integrity, health information systems lose confidentiality, critical
-infrastructure loses availability, government communications lose authenticity and non-repudiation, or the internet as a
-whole loses integrity, inconceivably large damage would occur.
-
-% FIXME: Kristen doesn't like the wording
-Encryption's dual role as enabler of strong privacy and cornerstone of computer security may be the central conflict in
-the EA debate. Those who focus on its role in privacy object to its absolute power. Those who focus on its role in
-security rely on its absolute power.
+Encryption's dual role as enabler of privacy and cornerstone of security is at the heart of the EA debate. Cryptography
+has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary in the realm
+of security.
 
@@
 analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
-research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}.
+research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
+important advance. Combined with computer networking and personal computing advances in the 1980s, it for the first time
+put the power of strong encryption not just in the hands of governments and militaries, but of ordinary people.
 
-The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
-computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
-governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
-property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
-Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
-Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
-\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
-rapid increase in availability of strong encryption.
-% FIXME: The point is made---these developments made gov'ts uncomfortable (just look at these examples!)---but the line
-% of thought is not clear enough to the reader.
+However, this power enables very strong privacy, and privacy is a property governments aim to regulate. The use of
+encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason
+based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force
+testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural
+stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of strong encryption.
 
@@
 
-Despite the failure of the \ac{clipperchip}, the debate over access to strong encryption continued, primarily focused on
-various key escrow proposals \cite{thompson_2015}.
-% FIXME: export controls were a hot topic of the time, too. Can't just drop that thread.
-In 1996, two pro-encryption bills were introduced in the US Congress.
-In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls on encryption software
-\cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly allow arbitrarily strong
-encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went further, including
-``provisions that would have barred the government from creating a mandatory key escrow system and would also have
-removed export restrictions on most generally available software and hardware containing encryption''
+Despite the failure of the \ac{clipperchip}, the debate over export controls and access to strong encryption continued,
+primarily focused on various key escrow proposals \cite{thompson_2015}. In 1996, two pro-encryption bills were
+introduced in the US Congress. In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls
+on encryption software \cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly
+allow arbitrarily strong encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went
+further, including ``provisions that would have barred the government from creating a mandatory key escrow system and
+would also have removed export restrictions on most generally available software and hardware containing encryption''
 \cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
@@
 
-The second development was the events of and reaction to the morning of September 11, 2001. When the U.S. had already
-been expanding police surveillance activities to combat communism, crime, drugs, and terrorism, the scale and audacity
-of the September 11 attacks seemed to justify yet greater investigatory powers \cite{bloss_escalating_2007}. October of
-2001 saw passage of the USA PATRIOT Act, which had several significant impacts: it weakened the limitation for using
-\ac{FISA} requests on U.S. citizens; it expanded the scope of what could be compelled via \ac{FISA} orders; it
-% FIXME: introduce FISA above, in the "had already been expanding" section. Then we have more of a narrative of
-% expanding power.
-authorized ``roving'' wiretaps; and it increased the power of \acp{NSL}, which can be used without judicial review to
-compel information from digital service providers while precluding any public disclosure of the event
-\cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was performing more surveillance than ever, and even these
-weakened limitations would be repeatedly violated \cite{shamsi_2011} \cite{tucker_2020}.
+The second development was the events of and reaction to the morning of September 11, 2001. These events occurred in the
+context of decades of all three branches of government expanding law enforcement powers. Presidents promoted the idea of
+being ``tough on crime'' and launched a war on drugs; legislators passed bills such as 1978's \ac{FISA}, which provided
+broad surveillance powers over foreign nationals, and in some cases U.S. citizens \cite{rodino_1978}; and the judiciary
+practiced a ``jurisprudence of crime control'' that gave police broad leeway \cite{gizzi_fourth_2016}.
+
+In this context of expanded surveillance, the scale and audacity of the September 11 attacks seemed to justify yet
+greater investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had
+several significant impacts: it weakened the limitation for using \ac{FISA} requests on U.S. citizens; it expanded the
+scope of what could be compelled via \ac{FISA} orders; it authorized ``roving'' wiretaps; and it increased the power of
+\acp{NSL}, which can be used without judicial review to compel information from digital service providers while
+precluding any public disclosure of the event \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was
+performing more surveillance than ever, and even these weakened limitations would be repeatedly violated
+\cite{shamsi_2011} \cite{tucker_2020}.
 
@@
 antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
-% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
-Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
-pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
-now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
+that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
+\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
+law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by now
+acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
 security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.
@@
 
-The following sections summarize the encryption regulatory environment and technical approaches to \ac{EA}.
-% FIXME: remove this sentence?
-
 \section{Regulatory Environment}
@@
 %  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
+% TODO: I _really_ need to understand these and how they're used today.
 

### 2020-07-10 Thesis-Chapter1-Introduction.tex
Address several more todos; add some reading sources

@@
 is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
-privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. 
-% TODO: citation
-In a
-society that cares about bringing wrongdoers to justice, this risk should not be ignored.
+privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing
+\cite{cox_2020} \cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk
+should not be ignored.
 
 The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
-access encrypted data to perform their duties. 
-% TODO: citation
-Human rights groups and technical leaders counter that a weakened
-encryption environment would fatally compromise privacy and security as we know it. 
-% TODO: citation
-As the debate continues, sides have
-become entrenched, and both sides have engaged in disingenuous attacks against their perceived opponents.
+access encrypted data to perform their duties \cite{ministerial_2018} \cite{fbi_2020}. Human rights groups and technical
+leaders counter that a weakened encryption environment would fatally compromise privacy and security as we know it
+\cite{ruiz_there_2018} \cite{abelson_2015}. As the debate continues, sides have become entrenched, and both sides have
+engaged in disingenuous attacks against their opponents.
 
@@
 thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.
-
-% TODO: use first person voice? Originally had wording like "this thesis introduces", but Kalafut says not to be afraid
-% of "I". Could also use passive voice.

### 2020-07-10 Thesis-Chapter2-Background.tex
Address several more todos; add some reading sources

@@
 
-% TODO: clarify relationship between encryption and cryptography?
+% TODO: clarify relationship between encryption and cryptography. Use appropriately throughout.
 
@@
 pencil of executive order.
-
-% TODO: add reference to future, comprehensive Appendix of all proposed encryption laws
 
@@
 reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's
-pro-encryption policies in 1999 \cite{kehl_right_2015}.
+pro-encryption policies in 1999 \cite{:kehl_right_2015}.
 
@@
 important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
-child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
+child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``appropriate means must be available to
 fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
@@
 manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
-proposals in \ac{the-cw1}.
-% TODO: how so?
-Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the draft's
-authors both used the phrase ``above the law'' in promotion of the bill.
-% TODO: how does this differ from the new LEAD Act?
+proposals in \ac{the-cw1} by forcing platform and application developers to comply. Previewing a common phrase in recent
+\ac{EA} arguments \cite{geller_2019}, the draft's authors both used the phrase ``above the law'' in promotion of the
+bill.
 
@@
 
-% TODO: as of 4/18/20, this bill is still active. Follow up.
+% TODO: as of 7/10/20, this bill is still active. Follow up.
 
 }
-
-% \law{2019}{Technology in Criminal Justice Act}{demings_hr5227_2019}{
-% % pfefferkorn_2019
-% % And in fact, there’s already proposed legislation based on CSIS’s recommendations: the Technology in Criminal Justice
-% % Act of 2019, introduced in the House last month by a bipartisan group of co-sponsors led by Rep. Val Demings of Florida.
-% % The bill seeks to close the resource, training, and information-sharing gap for digital evidence, and to improve the
-% % frayed relationship between law enforcement and technology companies. Notably absent from the bill is any language that
-% % would roll back the CALEA compromise or mandate encryption backdoors. To be clear, I’m not advocating for or against
-% % Rep. Demings’ bill. I just want to point out that it offers an alternative to whatever backdoor nightmare the Senate
-% % Judiciary Committee is surely drafting up right now.
-% }
 
@@
 
-% TODO: as of 4/18/20, this bill is still active. Follow up.
+% TODO: Update based on new amendment.
 
@@
 %
-% Looking at it now, even what I have seems like too much. Maybe move it all to an appendix.
-%
 % For more, read donahue_comparative_2018 (Israel and China) and budish_encryption_2018 (framework for analysis)
@@
     \country{European Countries}{acosta_government_2016}{
-% France, Belgium, Germany, Sweden, EU
-% EU GDPR \cite{parliament_2016}
 
 % TODO: it is very important to know who these requirements apply to. Traditional telecoms, obviously. Facebook and
-% Google? Signal and WhatsApp?
+% Google? Signal and WhatsApp? Telecoms, not platforms or apps. This is the big differentiator. This can be expanded
+% upon when I talk a bit about pen registers, etc.
 
@@
 counter-terrorism law explicitly compels access to plaintext data upon request \cite{donahue_comparative_2018}.
-
-% TODO: China, see also blanchard_2017
-
-% TODO: but you can still get iPhones in China, right... ? Again, scope?
 
@@
 % There are several types of approaches:
-% TODO: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
+% IDEA: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
 % properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
@@
 \subsection{Types of Key Escrow}
-
-% TODO: need an aside on shamir m/n secret sharing? Probably in order to understand split key approaches at all
 
@@
 
-% TODO: reference forward chapter
-Argdown is used in this thesis to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the
-basic structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be
-addressed through argument. Statements and arguments may support or attack one another, and arguments support or
-undercut other arguments.
+Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo}
+shows the basic structure of arguments. Statements are positions on issues. Some statements are based on values that
+cannot be addressed through argument. Statements and arguments may support or attack one another, and arguments support
+or undercut other arguments.
 
@@
 flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is
-the primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in
-this style to analyze the threats already present and those that \ac{EA} would introduce.
-% TODO: reference forward chapter
+the primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
+\acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.
 

### 2020-07-31 scripts/sourceme.sh
Update 'notes' alias to open selected file

@@ -22,9 +22,14 @@ alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
 alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
 alias fixme="(cd $BASE_DIR && grep -nE 'FIXME' *.tex)"
-alias notes="(cd $BASE_DIR/reading && fzf --layout=reverse --info=inline --border --preview-window=left:120:wrap \
-    --preview='$SCRIPTS_DIR/show-source.sh {} && printf \"====================\n\n\" && bat --plain --color=always {}')"
-
 alias pdf="xdg-open build/Thesis.pdf"
 alias gvsu="xdg-open build/Thesis-GVSU.pdf"
 alias start="(codium .) && (xdg-open build/Thesis.pdf) && make live"
+
+function notes {
+    pushd $BASE_DIR/reading >/dev/null
+    local FZF_PRV_SET="--layout=reverse --info=inline --border --preview-window=left:120:wrap"
+    local FZF_PRV_CMD="$SCRIPTS_DIR/show-source.sh {} && printf '====================\n\n' && bat --plain --color=always {}"
+    codium $(fzf $FZF_PRV_SET --preview="$FZF_PRV_CMD")
+    popd >/dev/null
+}

### 2020-08-27 Thesis-Chapter2-Background.tex
Additional reading. 'Wicked Crypto' is an excellent source
Several build updates; add arguments
Update to Node 12.x; add 'fallacies' argument map

@@
 % TODO: Review recent cases. See orenstein_2016, bittenbender_2019, vaas_2019.
+% NOTE: _Carpenter_ was a recent important Supreme Court case limiting the scope of the third-party doctrine!
+% See franklin_2018 and rozenshtein_2018
 
@@
 reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's
-pro-encryption policies in 1999 \cite{:kehl_right_2015}.
+pro-encryption policies in 1999 \cite{kehl_right_2015}.
 
@@
 }
+
+% TODO: add wright_crypto_2018 (I _think_ it goes here)
+
 \propsend

### 2020-08-27 Thesis-Chapter3-Arguments.tex
Several build updates; add arguments
Update to Node 12.x; add 'fallacies' argument map

@@
 
-In this chapter, develop the arguments employed in EA debates. Make use of argument maps to break down and visualize the
-debate, as shown in \myfig{fig-args-top-level}.
+% In this chapter, develop the arguments employed in EA debates. Make use of argument maps to break down and visualize the
+% debate, as shown in \myfig{fig-args-top-level}.
 
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{arguments/build/top-level.pdf}
-    \caption[Argument: Top Level]{Top level arguments for and against exceptional access.}
-    \label{fig-args-top-level}
-\end{figure}
+% \begin{figure}[h]
+%     \centering\CaptionFontSize
+%     \includegraphics[width=\linewidth]{arguments/build/top-level.pdf}
+%     \caption[Argument: Top Level]{Top level arguments for and against exceptional access.}
+%     \label{fig-args-top-level}
+% \end{figure}
 
@@
 \end{itemize}
+
+\subsection{Fallacious Arguments}
+
+Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}.

### 2020-08-27 arguments/argdown.config.js

@@ -45,6 +45,17 @@ module.exports = {
         selectedTags: ['ea-itself', 'anti-ea'],
         selectElementsWithoutTag: true,
       }
+    },
+    fallacies: {
+      process: 'export-svg',
+      outputSuffix: ".fallacies",
+      selection: {
+        selectedSections: [
+          'Central Issues', 'Response Measures', 'Exceptional Access', 'Fallacies', 'Fallacious Arguments'
+        ],
+        selectedTags: ['core', 'ea-itself', 'fallacy', 'fallacious'],
+        selectElementsWithoutTag: true
+      }
     }
   }
 };

### 2020-08-27 scripts/argdown-insert-citations.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-set -eEuo pipefail
-
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
-cd $BASE_DIR
-
-function get_ref_from_citation {
-  grep "bibcite{$1}" build/Thesis.aux | cut -d'{' -f3 | cut -d'}' -f1
-}
-
-function get_full_citations_in_file {
-  grep -o "#cite_[a-zA-Z_0-9]*" $1 | uniq
-}
-
-function parse_infile_to_outfile {
-  # echo "IN: $1; OUT: $2"
-  cp $1 $2
-  for CITE in $(get_full_citations_in_file $1); do
-    local PLAIN_CITE=${CITE/\#cite_/}
-    # local ACTUAL_CITE=${CITE/+/ }
-    local REF="$(get_ref_from_citation $PLAIN_CITE)"
-    [[ "" == "$REF" ]] && exitprint 1 "Citation for $CITE ($PLAIN_CITE) not found"
-    # echo "CITE: $PLAIN_CITE; REF: $REF"
-    sed -i "s/$CITE/\\\\[$REF\\\\]/g" $2
-  done
-}
-
-mkdir -p arguments/build/w_citations
-
-FILES=${@:-$(cd arguments; ls *.argdown)}
-for ARG_FILE in $FILES; do
-  parse_infile_to_outfile arguments/$ARG_FILE arguments/build/w_citations/$ARG_FILE
-done

### 2020-08-27 scripts/argdown-preprocess.sh

@@ -1,0 +1,43 @@
+#!/bin/bash
+
+set -eEuo pipefail
+
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+cd $BASE_DIR
+
+function get_ref_from_citation {
+  grep "bibcite{$1}" build/Thesis.aux | cut -d'{' -f3 | cut -d'}' -f1
+}
+
+function get_full_citations_in_file {
+  grep -o "(cite [a-zA-Z_0-9\\]*)" $1 | tr -d ")\\\\" | awk '{print $2}' | uniq
+}
+
+function parse_infile_to_outfile {
+  # Process citations
+  # echo "IN: $1; OUT: $2"
+  cp $1 $2
+  # echo $(get_full_citations_in_file $1)
+  for CITE in $(get_full_citations_in_file $1); do
+    local FULL_CITE="(cite ${CITE/_/\\\\_})"
+    local REF="$(get_ref_from_citation $CITE)"
+    [[ "" == "$REF" ]] && exitprint 1 "${RED}Bibtex citation for $CITE not found$NO_COLOR"
+    # echo "CITE: $CITE; REF: $REF"
+    sed -i "s/$FULL_CITE/\\\\[$REF\\\\]/g" $2
+  done
+
+  # Remove preview configs
+  if grep -q "Preview configs start" $2 && grep -q "Preview configs end" $2; then
+    START_LINE=$(grep -n "Preview configs start" $2 | cut -d: -f1)
+    END_LINE=$(grep -n "Preview configs end" $2 | cut -d: -f1)
+    echo "Cutting $START_LINE to $END_LINE of $2"
+    sed -i ${START_LINE},${END_LINE}d $2
+  fi
+}
+
+mkdir -p arguments/build/w_citations
+
+FILES=${@:-$(cd arguments; ls *.ad)}
+for ARG_FILE in $FILES; do
+  parse_infile_to_outfile arguments/$ARG_FILE arguments/build/w_citations/$ARG_FILE
+done

### 2020-08-27 scripts/count-citations.sh

@@ -9,14 +9,23 @@ function print_number_in_tex() {
 }
 
 function print_number_in_argdown() {
-    echo $(grep "cite_$1" arguments/*.argdown | wc -l)
+    echo $(grep "(cite ${1/_/\\\\_})" arguments/*.ad | wc -l)
+}
+
+function print_colored_number() {
+    (( $1 > 0 )) && COLOR=$LIGHT_GREEN || COLOR=$YELLOW
+    printf "$COLOR$1$NO_COLOR"
+}
+
+function pluralize() {
+    (( $1 != 1 )) && printf s
 }
 
 function print_count() {
     COUNT_TEX=$(print_number_in_tex $1)
     COUNT_ARGDOWN=$(print_number_in_argdown $1)
-    printf "$LIGHT_CYAN$1$NO_COLOR is cited $LIGHT_GREEN$COUNT_TEX$NO_COLOR time$(((1 == COUNT_TEX)) || echo s) in writing"
-    printf " and $LIGHT_GREEN$COUNT_ARGDOWN$NO_COLOR time$(((1 == COUNT_ARGDOWN)) || echo s) in arguments.\n"
+    printf "$LIGHT_CYAN$1$NO_COLOR is cited $(print_colored_number $COUNT_TEX) time$(pluralize $COUNT_TEX) in writing"
+    printf " and $(print_colored_number $COUNT_ARGDOWN) time$(pluralize $COUNT_ARGDOWN) in arguments.\n"
 }
 
 if (( 2 <= $# )) && [[ "-c" == "$1" ]]; then
@@ -28,6 +37,6 @@ elif (( 2 <= $# )) && [[ "-n" == "$1" ]]; then
 else
     print_count $1
     echo
-    grep --color=auto -n "\\\\cite{$1}" *.tex
-    grep --color=auto -n "#cite_$1" arguments/*.argdown
+    grep --color=auto -rn --include \*.tex "\\\\cite{$1}" .
+    grep --color=auto -rn --include \*.ad "(cite ${1/_/\\\\_})" .
 fi

### 2020-08-27 scripts/sourceme.sh

@@ -4,6 +4,7 @@ export BASE_DIR="$(git rev-parse --show-toplevel)"
 export READING_DIR="$BASE_DIR/reading"
 export SCRIPTS_DIR="$BASE_DIR/scripts"
 
+export RED='\e[1;31m'
 export YELLOW='\e[0;33m'
 export LIGHT_GREEN='\e[1;32m'
 export LIGHT_CYAN='\e[1;36m'
@@ -11,7 +12,7 @@ export BLUE='\e[0;34m'
 export NO_COLOR='\e[0m'
 
 function exitprint() {
-    echo "${@:2}"
+    echo -e "${@:2}"
     exit $1
 }
 export -f exitprint

### 2020-08-27 arguments/argdown.config.js
Add more fallacious arguments

@@ -4,11 +4,8 @@ module.exports = {
   author: 'Kevin Kredit',
   inputPath: './build/w_citations/comprehensive.ad',
   model: {
-    mode: 'loose',
+    mode: 'strict',
     removeTagsFromText: true
-  },
-  selection: {
-    excludeDisconnected: false
   },
   svg: {
     outputDir: './build'

### 2020-08-27 scripts/argdown-preprocess.sh

@@ -9,17 +9,17 @@ function get_ref_from_citation {
   grep "bibcite{$1}" build/Thesis.aux | cut -d'{' -f3 | cut -d'}' -f1
 }
 
-function get_full_citations_in_file {
-  grep -o "(cite [a-zA-Z_0-9\\]*)" $1 | tr -d ")\\\\" | awk '{print $2}' | uniq
+function get_citations_in_file {
+  grep -o "(cite:[a-zA-Z_0-9\\]*)" $1 | tr -d ")\\\\" | cut -d: -f2 | uniq
 }
 
 function parse_infile_to_outfile {
   # Process citations
   # echo "IN: $1; OUT: $2"
   cp $1 $2
-  # echo $(get_full_citations_in_file $1)
-  for CITE in $(get_full_citations_in_file $1); do
-    local FULL_CITE="(cite ${CITE/_/\\\\_})"
+  # echo $(get_citations_in_file $1)
+  for CITE in $(get_citations_in_file $1); do
+    local FULL_CITE="(cite:${CITE/_/\\\\_})"
     local REF="$(get_ref_from_citation $CITE)"
     [[ "" == "$REF" ]] && exitprint 1 "${RED}Bibtex citation for $CITE not found$NO_COLOR"
     # echo "CITE: $CITE; REF: $REF"

### 2020-08-27 scripts/count-citations.sh

@@ -9,7 +9,7 @@ function print_number_in_tex() {
 }
 
 function print_number_in_argdown() {
-    echo $(grep "(cite ${1/_/\\\\_})" arguments/*.ad | wc -l)
+    echo $(grep "(cite:${1/_/\\\\_})" arguments/*.ad | wc -l)
 }
 
 function print_colored_number() {
@@ -38,5 +38,5 @@ else
     print_count $1
     echo
     grep --color=auto -rn --include \*.tex "\\\\cite{$1}" .
-    grep --color=auto -rn --include \*.ad "(cite ${1/_/\\\\_})" .
+    grep --color=auto -rn --include \*.ad "(cite:${1/_/\\\\_})" .
 fi

### 2020-09-05 scripts/argdown-preprocess.sh
Improve argdown map generation build process

@@ -13,6 +13,13 @@ function get_citations_in_file {
   grep -o "(cite:[a-zA-Z_0-9\\]*)" $1 | tr -d ")\\\\" | cut -d: -f2 | uniq
 }
 
+# There are two methods of resolving citations.
+# 1. Manually using the Thesis.aux file.
+# 2. Automatically using TeX.
+
+TEX=false
+# TEX=true
+
 function parse_infile_to_outfile {
   # Process citations
   # echo "IN: $1; OUT: $2"
@@ -20,17 +27,25 @@ function parse_infile_to_outfile {
   # echo $(get_citations_in_file $1)
   for CITE in $(get_citations_in_file $1); do
     local FULL_CITE="(cite:${CITE/_/\\\\_})"
-    local REF="$(get_ref_from_citation $CITE)"
-    [[ "" == "$REF" ]] && exitprint 1 "${RED}Bibtex citation for $CITE not found$NO_COLOR"
-    # echo "CITE: $CITE; REF: $REF"
-    sed -i "s/$FULL_CITE/\\\\[$REF\\\\]/g" $2
+    if ! $TEX; then
+      local REF="$(get_ref_from_citation $CITE)"
+      if [[ "" == "$REF" ]]; then
+          printf "${RED}Bibtex citation for $CITE not found$NO_COLOR\n"
+          REF="E\\\\_AGAIN"
+      fi
+      # echo "CITE: $CITE; REF: $REF"
+      sed -i "s/$FULL_CITE/\\\\[$REF\\\\]/g" $2
+    else
+      TEX_CITE="\\\\\\\\cite\\\\{${CITE/_/\\\\_}\\\\}"
+      sed -i "s/$FULL_CITE/$TEX_CITE/g" $2
+    fi
   done
 
   # Remove preview configs
   if grep -q "Preview configs start" $2 && grep -q "Preview configs end" $2; then
     START_LINE=$(grep -n "Preview configs start" $2 | cut -d: -f1)
     END_LINE=$(grep -n "Preview configs end" $2 | cut -d: -f1)
-    echo "Cutting $START_LINE to $END_LINE of $2"
+    # echo "Cutting $START_LINE to $END_LINE of $2"
     sed -i ${START_LINE},${END_LINE}d $2
   fi
 }

### 2020-09-05 scripts/bib-gen.sh

@@ -16,7 +16,7 @@ function check_for_duplicate_refname() {
 
 function sanitize_language_field() {
   if [[ "$REF" =~ "language" ]]; then
-    for PATTERN in "n" "ng" "nglish" "n_GB" "n_US" "n\\\\_GB" "n\\\\_US" "n-GB" "n-US"; do
+    for PATTERN in "n" "ng" "nglish" "n_GB" "n_US" "n\\\\_GB" "n\\\\_US" "n-GB" "n-US" "n-us"; do
       REF=${REF/\{[Ee]$PATTERN\}/\{english\}}
     done
   fi

### 2020-09-05 arguments/argdown.config.js
Update arguments

@@ -44,6 +44,11 @@ module.exports = {
       }
     },
     fallacies: {
+      dot: {
+        graphVizSettings: {
+          rankdir: 'LR'
+        }
+      },
       process: 'export-svg',
       outputSuffix: ".fallacies",
       selection: {

### 2020-09-06 scripts/sourceme.sh
Add 'term' script

@@ -19,6 +19,7 @@ export -f exitprint
 
 alias new="$SCRIPTS_DIR/new-source.sh"
 alias show="$SCRIPTS_DIR/show-source.sh"
+alias term="$SCRIPTS_DIR/term.sh"
 alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
 alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"

### 2020-09-06 scripts/term.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+set -eEuo pipefail
+
+source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+cd $BASE_DIR
+
+WHOLE="$(tr -d '\n' < Thesis-Glossary.tex | grep -iEo "newabbreviation{$1}{[^}]*}{[^}]*}" | tr -d '}')"
+SHORT="$(echo $WHOLE | cut -d{ -f2)"
+LONG="$(echo $WHOLE | cut -d{ -f3)"
+DEF="$(echo $WHOLE | cut -d{ -f4)"
+
+echo -e "$LIGHT_CYAN$LONG$NO_COLOR ($SHORT)"
+echo -e "$LIGHT_GREEN$DEF$NO_COLOR"

### 2020-09-13 Thesis-Chapter1-Introduction.tex
Additional reading. 'Wicked Crypto' is an excellent source
Address some TODOs, fill in judicial role

@@
 
-Difficult policy matters are \acp{wicked-prob}. Originally coming from the field of design theory, the term
+Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
 ``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
@@
 
-\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption intended for law
-enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
-resistance to a clear solution, the debate on \ac{EA} refuses to end. This thesis will not attempt to end the debate,
-but to inform it with additional structure.
+\Acl{EA} is a wicked problem \cite{rozenshtein_wicked_2018}. In encryption policy, \ac{EA} is alternative means of
+decryption intended for law enforcement use. Characterized by a dynamic technological environment, disagreement about
+underlying values, and resistance to a clear solution, the debate on \ac{EA} will not go away. This thesis does not
+attempt to end the debate, but to structure and analyze it.
 

### 2020-09-13 Thesis-Chapter2-Background.tex
Improve argdown map generation build process
Create Latex macro for argument map includes
Address some TODOs, fill in judicial role

@@
 
-Data secrecy is not cryptography's only purpose, however.
+Data secrecy is not cryptography's only use, however. Precisely speaking, data secrecy is the purpose of encryption.
+Encryption is cryptography's ``killer app,'' but it is only a part of cryptography's usefulness to security.
 
@@
 (architecture) and the security of the components themselves; likewise, the security of each component emerges from the
-architecture of its subcomponents and the security of the subcompenents themselves. At the bottom of this analysis, the
-security of primitive components commonly relies on cryptography.
+architecture of its subcomponents and the security of the subcompenents themselves. At the end of this chain of
+analysis, the security of primitive components commonly relies on cryptography.
 
@@
 typically rooted in authentication.
-
-% TODO: clarify relationship between encryption and cryptography. Use appropriately throughout.
 
@@
 The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
-U.S.--focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
-problem, as digital technology has little respect for political borders.
+U.S.--focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an
+international problem, as digital technology has little respect for political borders.
 
@@
 \subsubsection{Enacted U.S. Regulations}
-% FIXME: distinguish statutes vs. regulations; rename "failed" section
 
@@
 Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
-% FIXME: Briefly introduce these. Even just "Amended the switch-board-era"...
+% TODO: Briefly introduce these. Even just "Amended the switch-board-era"...
 electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
@@
 
-\subsubsection{The Judicial Environment}
-\label{sec-us-judicial}
-
-The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
-and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
-judiciary branch exercises power through a framework based on AAAA and BBBB. In this section we analyze several
-important cases and developments in the last few years.
-
-% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998 and gizzi_fourth_2016. Update the
-%   "AAAA and BBBB" in the intro.
-
-% TODO: Review historically important cases on the 1st, 4th, and 5th amendments, e.g. Katz v. United States, 1969.
-
-% TODO: Review recent cases. See orenstein_2016, bittenbender_2019, vaas_2019.
-% NOTE: _Carpenter_ was a recent important Supreme Court case limiting the scope of the third-party doctrine!
-% See franklin_2018 and rozenshtein_2018
-
-\subsubsection{Failed U.S. Regulation Attempts}
+\subsubsection{Unsuccessful U.S. Regulation Attempts}
 \label{sec-us-reg-fails}
@@
 
+\subsubsection{The Judicial Environment}
+\label{sec-us-judicial}
+
+The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
+and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
+judiciary branch exercises power through a framework based on AAAA and BBBB. In this section I describe this framework,
+review how it has influenced cryptography policy, and analyze where cryptographic jurisprudence may go from here.
+
+% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998. Update the "AAAA and BBBB" in the
+% intro and the todo in the conclusion.
+
+Judicial influence on encryption policy is primarily mediated through Fourth Amendment jurisprudence, particularly
+regarding the interpretation of a person's ``effects,'' what is ``reasonable,'' and what is a ``search.'' The Supreme
+Court's interpretation and implementation of the Amendment's crime control-- and civil liberty--balancing principles has
+changed over time. An in-depth analysis of the Court's jurisprudence from the 1960's Warren Court, through the
+influential 1980s Rehnquist Court, to the mid-2010s Roberts Court found a deep proclivity towards crime control that is
+slowly thawing \cite{gizzi_fourth_2016}. This thawing, caused by evolution of technology, of the court, and of the
+justices themselves, has been favorable for data privacy. Recent cases have increasingly sided with defendants, have
+caused justices to strain to show that they understand technology, and have increased the categories of protected
+digital data \cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter v. United States} may be the most
+impactful yet; though it only resulted in a narrow decision against unwarranted cell phone location data tracking, it
+does undermine the pre-digital third-party doctrine, which holds that Fourth Amendment protection is forfeit when the
+data is willingly given to a third-party \cite{franklin_2018}.
+
+As this relates to \ac{EA}, Fourth Amendment interpretation practically determines the threshold at which authorities
+require a warrant, not what demands they can or cannot make once they have one. The second most salient legal issue is
+the Fifth Amendment right against self-incrimination, which comes up in cases of compelled password disclosure. This
+question has not yet reached the Supreme Court, and state courts have ruled compelled disclosure both constitutional
+\cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional \cite{lee_its_2020} \cite{vaas_2019}.
+
+% Paragraph interpreting this change as principles vs rules in a wicked problem
+
+That judicial influence on privacy and technology is changing is to be expected. To explain the fluctuating state of
+legal interpretation as it relates to \ac{EA}, Matyas et al. make the helpful distinction between legal rules and
+principles \cite{matyas_incommensurability_2018}:
+
+\begin{displayquote}
+The tension between technical and legal views of sensitive issues such as encryption and surveillance is illuminated by
+applying the jurisprudential lens. \dots~Liberal legal systems, manifesting what is generally understood as ``the rule
+of law,'' are actually composed of both rules and principles. Legal ``rules'' can be understood as logical propositions
+that are expected to yield answers about what is and is not permitted using formal reasoning capabilities. By contrast,
+legal ``principles'' articulate values and policies that must be reflected in a legal system but do not necessarily
+dictate an unambiguous outcome in any given case.
+\end{displayquote}
+
+Rules are the low-level implementation of principles. When the underlying facts of the situation change, as they do by
+definition for wicked problems, the rules become out of sync with the principles that bade them. Recent flux in judicial
+rulings surrounding surveillance and technology are evidence that the courts are in the process of updating
+interpretations of rules in light of the new facts; which principles they ultimately choose to favor remains to be seen,
+but rulings like \ii{Carpenter} show a willingness to protect privacy.
+% TODO: Connect back to system of judicial influence--this willingness uses that framework to support encryption.
+
 \subsection{Regulations around the World}
@@
 This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
-brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
-by the Law Library of Congress \cite{acosta_government_2016}.
+brief account of the regulatory environment in various geographical jurisdictions around the world based primarily on a
+2016 analysis by the Law Library of Congress \cite{acosta_government_2016}. Globally, a 2017 study found that more than
+half the world's internet users lived in a country where some form of plaintext recovery is mandated \cite{lewis_2017}.
 
@@
 
-France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
-France, Belgium, and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each
-have some level of \ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or
-decryption of network traffic.
+Among European countriues, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
+Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for France, Belgium,
+and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each have some level of
+\ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or decryption of network
+traffic.
 
@@
 telecommunications providers provide interfaces ``with functions that can cooperate with interception.'' Chinese
-authorities have great access to plaintext data by virtue of the political structure of the state, but its 2015
-counter-terrorism law explicitly compels access to plaintext data upon request \cite{donahue_comparative_2018}.
+authorities have great access to plaintext data by virtue of the political structure of the state, but also through
+explicit requirements in recent anti-terrorism and cybersecurity laws \cite{lewis_2017}.
 
@@
 
-% TODO: add wright_crypto_2018 (I _think_ it goes here)
+    \prop{Cryptographic ``crumple zones''}{
 
+A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
+construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
+human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
+surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
+that makes them inherently recoverable only after a large up front and marginal cost.
+
+}
 \propsend
@@
 
-    \prop{Forced password disclosure}{
+    \prop{Compelled password disclosure}{
 
@@
 \ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
-this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of
-course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.
+this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}
+\cite{sobel_2019}. Of course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is
+not in custody.
 
@@
     \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{arguments/build/demo.pdf}
+    \myincludeargument{demo}
     \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}

### 2020-09-19 article/Article.tex
Add conference article template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SETUP %
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\tt#1{\mbox{\texttt{#1}}}

\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %

\title{Wicked Exceptional Access}

\author{\IEEEauthorblockN{Kevin Kredit}
\IEEEauthorblockA{\textit{GVSU ACS Department} \\
Grand Rapids, Michigan \\
Email: k.kredit.us@ieee.org}
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %

\begin{abstract}

The debate has resumed over whether encryption systems should support alternative means of decryption intended for law
enforcement use, called exceptional access (EA). Recent events have renewed interest in EA from the U.S. legislature.
Although EA would inevitably reduce digital security and privacy compared to free use of encryption, the quality of the
technical and regulatory approach can make a substantial difference. If one considers legislative action as part of the
threat model, one has no choice but to engage the topic. In this thesis I aim to clarify and analyze the debate. I begin
by outlining the issue's technical and historical context. In the following chapters I use argument maps to break down
the arguments used by both sides and use data flow diagrams to create threat models for the separate problems of EA for
data in motion and data at rest.

\end{abstract}

\begin{IEEEkeywords}
encryption, exceptional access, policy, wicked problems
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PAPER %

\section{Introduction}

Qwer \cite{flawfinder}.

\section{Conclusion}

Asdf.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %

\begin{thebibliography}{00}

\bibitem{flawfinder} D. Wheeler, "Flawfinder Home Page", \textit{Dwheeler.com}, 2019. [Online].
Available: https://dwheeler.com/Flawfinder/. [Accessed: 02- Nov- 2019].

\end{thebibliography}

\end{document}

### 2020-09-20 Thesis-Chapter3-Arguments.tex
Add more fallacious arguments
Improve argdown map generation build process
Create Latex macro for argument map includes
Flesh out outline and plan
Address some TODOs, fill in judicial role
Update TODO list and add JPMATMS notes
Add argumentation notes

@@
 \label{chap-arguments}
+
+\section{Wicked Problems}
+
+\subsection{Encryption and Exceptional Access as a Wicked Problem}
+
+- rozenshtein_wicked_2018 convincingly argues the point at length
+
+\subsection{Approaches to Solving Wicked Problems}
+
+\subsubsection{Failure of the Classical Analytic Method}
+
+\subsubsection{Muddling Through}
+
+% Mention hermeneutics and complex adaptive systems?
+
+\subsubsection{Pitfalls to Avoid}
+
+- Easy answers
+- Fallacies!
+
+\section{Desireable Outcomes}
+
+\section{Proposed Alternatives}
+
+% - But we're _exactly_ following the classical analytic method! Well, yes, but in a cooperative manner, knowing it will
+%     require iteration, and being careful about side effects. It's an agile approach to policy making, but it's still
+%     policy making.
+
+commission_tackling_2018:
+- Strategies
+  - Authoritative
+  - Competitive
+  - Collaborative
+- Approaches
+  - Linear problem solving
+    - Ineffective! Several reasons, good breakdown. Dovetails with rozenshtein_wicked_2018.
+  - Holistic problem solving: adaptive, innovative, flexible, cross-organizational
+- Avoid
+  - Narrow solutions
+
+rozenshtein_wicked_2018:
+- Approaches
+  - Muddling through (he cites Charles E. Lindblom, just like feeley_judicial_2000!)
+  - Risk management (--finally, something familiar to security experts!)
+    - So, a broader view of security, inclusive of both public safety and information security
+    - In this model, threat actors now include terrorists, CSAM distributors, organized crime
+    - I have already suggested adding government mandates to the traditional IS threat model
+  - Imperfect solutions, like lawful hacking
+  - Meanwhile, invest in knowledge production and improved relationships (like "cross-organizational" from
+      commission_tackling_2018)
+- Avoid
+  - Easy answers ("narrow solutions" from commission_tackling_2018)
+    - Such as EA mandates or "golden age of surveillance"-based denials
+
+matyas_incommensurability_2018:
+- Approaches to reducing incommensurability
+  - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
+  - Use transparency and accountability (through technical means)
+
+Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
+effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
+decrypted! So must be very careful to avoid solution attempts that leave long scars.
+
+feeley_judicial_2000:
+- Classical analytic method ("linear" in commission_tackling_2018)
+- Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)
+- The hermeneutic circle
+
+% How to tackle wicked problems in general? Take some tips from commission_tackling_2018, rozenshtein_wicked_2018,
+% feeley_judicial_2000, and matyas_incommensurability_2018:
+% Correctly state the problem--identify the root causes. They aren't going away, there by definition is no clear long
+% term solution. Put the proposed solutions in context--none are perfect, none are permanent. For EA in particular,
+% consider balance of principles and rules.
+
+% Subject one: do we even want EA?
+%  --> no? exit here
+%  --> yes? carry on
+%  --> pragmatist? carry on
+% include fallacies argument map here
+% see rozenshtein_wicked_2018, benaloh_what_2018, soesanto_2018, ministerial_2018, fbi_2020, brantley_2018, lund_2020
+% ... group_2019, ruiz_there_2018, rozenshtein_2019
+
+% Subject two: what properties would good EA have?
+% easy for LEAs to use? hard? expensive? voluntary? unable to be used for mass surveillance? transparent (attempts are
+%   visible)? alignment of values with security (e.g. discourages bug hoarding)? adaptable (since solution won't be
+%   permanent, either technically or according to collective will as things change rozenshtein_wicked_2018)? suited to
+%   rules and principles (matyas_incommensurability_2018)?
+
+% Subject three: what options do we have?
+% map of properties onto proposals--leads into threat modeling, which will check the veracity of claims
+
+%%%%%%%%%%%%
+% Chapter 4, threat models
+% (how to be specific enough to be useful but generic enough to apply to many systems? Focus on the EA type (DAR v DIM),
+%   then apply to specific systems)
+% DAR, DIM threat models (make sure to cover abelson_2015)
+% Basic proposals
+%  --> mappings onto threat models
+% At work: 3 example systems analyzed with this threat model
+% 1. IM (E2E IM too?) -- in motion
+% 2. Cloud storage -- at rest
+% 3. Mobile phone -- at rest, but way different scenario than cloud
+%  --> include Ozzie's CLEAR, since it was higher profile
+%%%%%%%%%%%%
+
+% Leave for conclusion chapter:
+% Subject four: now what?
+%  --> How current proposals stack up against desired properties, threat model
+%  --> security expert? do research, provide clear security advice, be vigilant
+%  --> policymaker? fund research, don't mislead, don't be premature
+%  --> either? know that there is no stable solution, balance rules and principles, serve your neighbor
+% see varia_2018
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}.
+Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}. Also cited:
+\cite{hanna_2019}.
+
+Fallacies are as shown in \myfig{fig-args-fallacies}.
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \myincludeargument{comprehensive.fallacies}
+    \caption[Fallacious Arguments]{Fallacious arguments used in the exceptional access debate.}
+    \label{fig-args-fallacies}
+\end{figure}

### 2020-09-20 scripts/term.sh

@@ -5,7 +5,13 @@ set -eEuo pipefail
 source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
 cd $BASE_DIR
 
-WHOLE="$(tr -d '\n' < Thesis-Glossary.tex | grep -iEo "newabbreviation{$1}{[^}]*}{[^}]*}" | tr -d '}')"
+GLOS=Thesis-Glossary.tex
+
+if grep -iEoq "newabbreviation{$1}" $GLOS; then
+    WHOLE="$(tr -d '\n' < $GLOS | grep -iEo "newabbreviation{$1}{[^}]*}{[^}]*}" | tr -d '}')"
+else
+    WHOLE="$(tr -d '\n' < $GLOS | grep -iEo "newterm{$1}{[^}]*}{[^}]*}" | tr -d '}')"
+fi
 SHORT="$(echo $WHOLE | cut -d{ -f2)"
 LONG="$(echo $WHOLE | cut -d{ -f3)"
 DEF="$(echo $WHOLE | cut -d{ -f4)"

### 2020-09-20 Thesis-Chapter3-Arguments.tex
Update Ch3 outline and add CSAM sources

@@
 
-\section{Desireable Outcomes}
-
-\section{Proposed Alternatives}
+\subsection{Plan for this Chapter}
 
@@
 %     policy making.
+% Already covered defining the problem. Next will discuss goals and list alternatives. That will feed into threat
+% modeling.
+
+\section{Desireable Outcomes}
+
+\section{Proposed Alternatives}
 
@@
     - In this model, threat actors now include terrorists, CSAM distributors, organized crime
+      - Add a new character? Evil Alice and evil Bob (Alex and Becca?)? This automatically raises to question of how to
+          identify bad actors--AI (lee_detecting_2020)? PhotoDNA (microsoft_2020)? Government orders (warrants)? Could
+          be useful.
     - I have already suggested adding government mandates to the traditional IS threat model
@@
 
-%%%%%%%%%%%%
-% Chapter 4, threat models
-% (how to be specific enough to be useful but generic enough to apply to many systems? Focus on the EA type (DAR v DIM),
-%   then apply to specific systems)
-% DAR, DIM threat models (make sure to cover abelson_2015)
-% Basic proposals
-%  --> mappings onto threat models
-% At work: 3 example systems analyzed with this threat model
-% 1. IM (E2E IM too?) -- in motion
-% 2. Cloud storage -- at rest
-% 3. Mobile phone -- at rest, but way different scenario than cloud
-%  --> include Ozzie's CLEAR, since it was higher profile
-%%%%%%%%%%%%
-
-% Leave for conclusion chapter:
-% Subject four: now what?
-%  --> How current proposals stack up against desired properties, threat model
-%  --> security expert? do research, provide clear security advice, be vigilant
-%  --> policymaker? fund research, don't mislead, don't be premature
-%  --> either? know that there is no stable solution, balance rules and principles, serve your neighbor
-% see varia_2018
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The debate over exceptional access takes place on two fundamental levels:
-\begin{enumerate}
-    \item Whether cryptosystems with exceptional access represent a net benefit to society
-    \item Whether and how to conduct exceptional access research
-\end{enumerate}
+% The debate over exceptional access takes place on two fundamental levels:
+% \begin{enumerate}
+%     \item Whether cryptosystems with exceptional access represent a net benefit to society
+%     \item Whether and how to conduct exceptional access research
+% \end{enumerate}
 
-\subsection{Arguments For}
+% \subsection{Arguments For}
 
-Why EA is a benefit to society:
-\begin{itemize}
-    \item Absolute privacy is not a right
-    \item Encryption gives too much power to the individual
-    \item EA would allow more crimes to be prevented and prosecuted
-\end{itemize}
+% Why EA is a benefit to society:
+% \begin{itemize}
+%     \item Absolute privacy is not a right
+%     \item Encryption gives too much power to the individual
+%     \item EA would allow more crimes to be prevented and prosecuted
+% \end{itemize}
 
-Why to research EA:
-\begin{itemize}
-    \item EA done well would be more secure and privacy preserving than EA done poorly
-    \item The threat of legislative action creates urgency
-\end{itemize}
+% Why to research EA:
+% \begin{itemize}
+%     \item EA done well would be more secure and privacy preserving than EA done poorly
+%     \item The threat of legislative action creates urgency
+% \end{itemize}
 
-\subsection{Arguments Against}
+% \subsection{Arguments Against}
 
-Why EA is a danger to society:
-\begin{itemize}
-    \item Privacy is an absolute right
-    \item Exceptional access gives too much power to the state
-    \item Exceptional access is not possible in an acceptably secure manner
-\end{itemize}
+% Why EA is a danger to society:
+% \begin{itemize}
+%     \item Privacy is an absolute right
+%     \item Exceptional access gives too much power to the state
+%     \item Exceptional access is not possible in an acceptably secure manner
+% \end{itemize}
 
-Why not to research EA:
-\begin{itemize}
-    \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
-            them
-    \item The technical community should not entertain delusions about what is truly possible
-\end{itemize}
+% Why not to research EA:
+% \begin{itemize}
+%     \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
+%             them
+%     \item The technical community should not entertain delusions about what is truly possible
+% \end{itemize}
 
-\subsection{Fallacious Arguments}
+% \subsection{Fallacious Arguments}
 
-Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}. Also cited:
-\cite{hanna_2019}.
+% Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}. Also cited:
+% \cite{hanna_2019}.
 
-Fallacies are as shown in \myfig{fig-args-fallacies}.
+% Fallacies are as shown in \myfig{fig-args-fallacies}.
 
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \myincludeargument{comprehensive.fallacies}
-    \caption[Fallacious Arguments]{Fallacious arguments used in the exceptional access debate.}
-    \label{fig-args-fallacies}
-\end{figure}
+% \begin{figure}[h]
+%     \centering\CaptionFontSize
+%     \myincludeargument{comprehensive.fallacies}
+%     \caption[Fallacious Arguments]{Fallacious arguments used in the exceptional access debate.}
+%     \label{fig-args-fallacies}
+% \end{figure}

### 2020-09-24 Thesis-Chapter3-Arguments.tex
Work on DFDs

@@
 \section{Wicked Problems}
+
+Wicked problems
 
@@
 - Easy answers
+  - EA mandates, complexity denial
 - Fallacies!
+- Bad results with long technical tails
+
+% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
+% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
+% decrypted! So must be very careful to avoid solution attempts that leave long scars.
 
@@
   - Use transparency and accountability (through technical means)
-
-Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
-effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
-decrypted! So must be very careful to avoid solution attempts that leave long scars.
 

### 2020-09-26 Thesis-Abstract.tex
Update abstract and TODOs

Public debate has resumed over whether encryption systems should support alternative means of decryption intended for
law enforcement use, called exceptional access (EA). Rather than a renege on a solemn promise made at the end of the
1990s ``crypto war,'' this represents a valid reassessment of optimal policy in light of changing circumstances.
Achieving proper balance between privacy and access in the context of constantly changing society and technology is a
wicked problem that has and will evade a permanent solution. As policymakers consider next steps, it behooves the
technical community to stay engaged. Although the introduction of EA would inevitably introduce risk, the quality of the
technical and regulatory approach can make a substantial difference. Furthermore, if one considers ham-fisted
legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall
risk.

The root of the conflict lies in cryptography's dual role as enabler of radical privacy and cornerstone of security. The
emergence of strong encryption triggered the first crypto war, and its proliferation is causing the second. In the face
of both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked problems and promote
an iterative approach to the problem of encryption and EA. Along the way I illustrate the components of the debate in
argument maps and demonstrate the security risks with data flow diagrams, making clear distinctions for the separate
problems of EA for data in motion and data at rest.

### 2020-09-30 Thesis-Chapter2-Background.tex
Finish judicial influence section

@@
 and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
-judiciary branch exercises power through a framework based on AAAA and BBBB. In this section I describe this framework,
-review how it has influenced cryptography policy, and analyze where cryptographic jurisprudence may go from here.
+judiciary branch influences policy through interpretation of laws and creation of legal doctrine. In this section I
+describe this influence, review how it has affected cryptography policy, and analyze where cryptographic jurisprudence
+may go from here.
 
-% TODO: Framework for understanding judicial influence. See bergerson_judicial_1998. Update the "AAAA and BBBB" in the
-% intro and the todo in the conclusion.
+The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
+authoritative legal texts. However, whether it is condemned, denied, seen as a stark reality, or seen as a valid
+function, the judicial branch also establishes legal doctrine and actively enforces rulings in a manner that amounts to
+creation of public policy \cite{feeley_judicial_2000}. Due to the clear congressional authority over telecommunications
+and federal investigatory powers, the courts are most likely to exert influence in this area through interpretation of
+current law. This interpretation will both determine how present law is enforced and draw a constitutional fence around
+any actions the executive or legislative branches undertake.
 
-Judicial influence on encryption policy is primarily mediated through Fourth Amendment jurisprudence, particularly
-regarding the interpretation of a person's ``effects,'' what is ``reasonable,'' and what is a ``search.'' The Supreme
-Court's interpretation and implementation of the Amendment's crime control-- and civil liberty--balancing principles has
-changed over time. An in-depth analysis of the Court's jurisprudence from the 1960's Warren Court, through the
-influential 1980s Rehnquist Court, to the mid-2010s Roberts Court found a deep proclivity towards crime control that is
-slowly thawing \cite{gizzi_fourth_2016}. This thawing, caused by evolution of technology, of the court, and of the
-justices themselves, has been favorable for data privacy. Recent cases have increasingly sided with defendants, have
-caused justices to strain to show that they understand technology, and have increased the categories of protected
-digital data \cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter v. United States} may be the most
-impactful yet; though it only resulted in a narrow decision against unwarranted cell phone location data tracking, it
-does undermine the pre-digital third-party doctrine, which holds that Fourth Amendment protection is forfeit when the
-data is willingly given to a third-party \cite{franklin_2018}.
+Judicial influence on encryption policy is primarily mediated through the Fourth Amendment, particularly regarding the
+interpretation of a person's ``effects,'' what is ``reasonable,'' and what is a ``search.'' The Supreme Court's
+interpretation and implementation of the Amendment's crime control-- and civil liberty--balancing principles has changed
+over time. An in-depth analysis of the Court's jurisprudence from the 1960's Warren Court, through the influential 1980s
+Rehnquist Court, to the mid-2010s Roberts Court found a deep proclivity towards crime control that is slowly thawing
+\cite{gizzi_fourth_2016}. This thawing, caused by evolution of technology, of the court, and of the justices themselves,
+has been favorable for data privacy. Recent cases have increasingly sided with defendants, have caused justices to
+strain to show that they understand technology, and have increased the categories of protected digital data
+\cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter v. United States} may be the most impactful yet;
+though it only resulted in a narrow decision against unwarranted cell phone location data tracking, it does undermine
+the pre-digital third-party doctrine, which holds that Fourth Amendment protection is forfeit when the data is willingly
+given to a third-party \cite{franklin_2018}.
 
-As this relates to \ac{EA}, Fourth Amendment interpretation practically determines the threshold at which authorities
-require a warrant, not what demands they can or cannot make once they have one. The second most salient legal issue is
-the Fifth Amendment right against self-incrimination, which comes up in cases of compelled password disclosure. This
-question has not yet reached the Supreme Court, and state courts have ruled compelled disclosure both constitutional
-\cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional \cite{lee_its_2020} \cite{vaas_2019}.
+As this relates to \ac{EA}, Fourth Amendment interpretation determines the threshold at which authorities require a
+warrant, not what demands they can or cannot make once they have one (e.g. demands for access to \ac{plaintext}). As
+mentioned above, the FBI argued in the San Bernardino case that the \ac{AWA} grants it authority to demand access,
+though it is strongly argued that \acs{CALEA} explicitly precludes this kind of demand \cite{gidari_2016}. Since the FBI
+dropped the case, the matter is not formally settled, though the consensus is that legislation such as the \ac{LEADAct}
+is required before courts would uphold that command.
 
-% Paragraph interpreting this change as principles vs rules in a wicked problem
+The second most salient legal issue is the Fifth Amendment right against self-incrimination, which comes up in cases of
+compelled password disclosure. This question has not yet reached the Supreme Court, and state courts have ruled
+compelled disclosure both constitutional \cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional \cite{lee_its_2020}
+\cite{vaas_2019}.
 
@@
 definition for wicked problems, the rules become out of sync with the principles that bade them. Recent flux in judicial
-rulings surrounding surveillance and technology are evidence that the courts are in the process of updating
-interpretations of rules in light of the new facts; which principles they ultimately choose to favor remains to be seen,
-but rulings like \ii{Carpenter} show a willingness to protect privacy.
-% TODO: Connect back to system of judicial influence--this willingness uses that framework to support encryption.
+rulings surrounding surveillance and technology, particularly around the Fourth Amendment, are evidence that the courts
+are in the process of updating interpretations of rules in light of the new facts; which principles they ultimately
+choose to favor remains to be seen, but rulings like \ii{Carpenter} show a willingness to protect privacy.
+
+Looking ahead, there are three types of cases through which the judicial branch could have a strong effect on digital
+privacy policy. First, Fourth Amendment cases will establish classes of protected data and thresholds for warrants.
+Second, a Supreme Court Fifth Amendment case could settle under what circumstances, if any, a password could be
+compelled. Third, a surprise \ac{AWA} ruling could open the door to demands for access, though legislation like
+\ac{LEADAct} is more likely to have this effect.
 

### 2020-10-03 Thesis-Abstract.tex
Update abstract

@@
 
-The root of the conflict lies in cryptography's dual role as enabler of radical privacy and cornerstone of security. The
-emergence of strong encryption triggered the first crypto war, and its proliferation is causing the second. In the face
-of both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked problems and promote
-an iterative approach to the problem of encryption and EA. Along the way I illustrate the components of the debate in
-argument maps and demonstrate the security risks with data flow diagrams, making clear distinctions for the separate
-problems of EA for data in motion and data at rest.
+The root of the conflict lies in cryptography's dual role as enabler of unprecedented privacy and cornerstone of
+security. The emergence of strong encryption triggered the first crypto war, and its proliferation is causing the
+second. In response to both polarized and conciliatory voices, I will analyze strategies that do and do not work on
+wicked problems and promote an iterative approach to the case of encryption and EA. Along the way I will illustrate the
+components of the debate in argument maps and demonstrate the security risks with data flow diagrams, making clear
+distinctions for the separate problems of EA for data in motion and data at rest.

### 2020-10-10 Thesis-Chapter3-Arguments.tex
Add signal x3dh protocol DFD
Add OODA loop diagram and notes; update outline

@@
 
-Wicked problems
+Wicked problems. \ac{OODAloop}.
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
+    \caption[The OODA Loop]{The OODA Loop}
+    \label{fig-ooda-loop}
+\end{figure}
 
@@
 % Mention hermeneutics and complex adaptive systems?
+% As feeley_judicial_2000 points out, muddling through is too focused on, or even consists entirely of, the
+% implementation step. Because of the long tail effects of any implementation in a wicked scenario, and especially in
+% the context of encryption, mere muddling through is too risky. Therefore I suggest iteration. Classical methods, but
+% applied in a cyclical manner.
 
@@
 
+% See committee_decrypting_2018!
+% For a wholistic threat model, include both the threats of government (legislation) and of criminal abusers of your
+% system.
+
 \section{Proposed Alternatives}
+
+% See committee_decrypting_2018!
 
@@
 - Classical analytic method ("linear" in commission_tackling_2018)
+% 1. Classical analytic method; each step has its own approaches, e.g. cost-benefit analysis (loosely based on natural
+% sciences)
+%   (missing: Observe)
+%   a. Define the problem (Orient)
+%   b. Identify a goal (Orient)
+%   c. Generate a range of alternatives for achieving that goal (Orient)
+%   d. Select the alternative that seems most promising (Decide)
+%   e. Implement the selected alternative (Act)
+%   (missing: loop)
 - Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)

### 2020-10-14 Thesis-Chapter2-Background.tex
Update title and README
Update outline and renumber regulations sections

@@
 international problem, as digital technology has little respect for political borders.
+\mysecrange{sec-us-regs}{sec-us-judicial} describe the regulatory environment in the U.S. and \mysec{sec-world-regs}
+describes regulations around the world.
 
@@
 
-\subsection{U.S. Efforts to Regulate Encryption}
-
-This section includes both successful and failed cryptography regulation attempts in the U.S.
-
-\subsubsection{Enacted U.S. Regulations}
+\subsection{Enacted U.S. Regulations}
+\label{sec-us-regs}
 
@@
 
-\subsubsection{Unsuccessful U.S. Regulation Attempts}
+\subsection{Unsuccessful U.S. Regulation Attempts}
 \label{sec-us-reg-fails}
@@
 
-\subsubsection{The Judicial Environment}
+\subsection{The U.S. Judicial Environment}
 \label{sec-us-judicial}
@@
 and federal investigatory powers, the courts are most likely to exert influence in this area through interpretation of
-current law. This interpretation will both determine how present law is enforced and draw a constitutional fence around
+current law. This interpretation will both determine how present law is enforced and draw a constitutional line around
 any actions the executive or legislative branches undertake.
@@
 
+% Prof. Walsh:
+% Due to the dynamic nature of future judicial rulings that may expand or retract the application of exceptional access
+% it is beyond the scope of this research to comment further...
+
 \subsection{Regulations around the World}
+\label{sec-world-regs}
 
@@
 \ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
-this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}
+this approach draws debates about the Fifth Amendment right against self-incrimination \cite{bittenbender_2019}
 \cite{sobel_2019}. Of course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is

### 2020-10-14 Thesis-Chapter1-Introduction.tex
Update notes on five-eyes communiques

@@
 The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
-access encrypted data to perform their duties \cite{ministerial_2018} \cite{fbi_2020}. Human rights groups and technical
-leaders counter that a weakened encryption environment would fatally compromise privacy and security as we know it
-\cite{ruiz_there_2018} \cite{abelson_2015}. As the debate continues, sides have become entrenched, and both sides have
-engaged in disingenuous attacks against their opponents.
+access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{fbi_2020}. Human rights
+groups and technical leaders counter that a weakened encryption environment would fatally compromise privacy and
+security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate continues,
+sides have become entrenched, and both sides have engaged in disingenuous attacks against their opponents.
 

### 2020-10-14 Thesis-Chapter3-Arguments.tex
Reoutline Chapter 3

@@
-\chapter{Arguments}
+\chapter{Wicked Problems and the EA Debate}
 \label{chap-arguments}
+
+Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
+greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
+to tackling \acp{wicked-prob} succeed and fail? How can we collectively situation ourselves to make real progress? This
+chapter seeks to answer these questions.
 
@@
 
-Wicked problems. \ac{OODAloop}.
+\subsection{Characteristics}
 
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
-    \caption[The OODA Loop]{The OODA Loop}
-    \label{fig-ooda-loop}
-\end{figure}
-
-\subsection{Encryption and Exceptional Access as a Wicked Problem}
+\subsection{Encryption and EA as a Wicked Problem}
 
@@
 
-\subsection{Approaches to Solving Wicked Problems}
+\section{Failure of Current Policymaking Approaches}
 
-\subsubsection{Failure of the Classical Analytic Method}
+\subsection{The Classical Analytical Method}
 
-\subsubsection{Muddling Through}
+\subsection{Muddling Through}
 
@@
 
-\subsubsection{Pitfalls to Avoid}
+\subsection{Lessons}
 
+- Avoid pitfalls of
   - Easy answers
@@
 
-\subsection{Plan for this Chapter}
+\section{Proposal: The OODA Loop for Policymaking}
 
-% - But we're _exactly_ following the classical analytic method! Well, yes, but in a cooperative manner, knowing it will
-%     require iteration, and being careful about side effects. It's an agile approach to policy making, but it's still
-%     policy making.
+\ac{OODAloop}.
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
+    \caption[The OODA Loop]{The OODA Loop}
+    \label{fig-ooda-loop}
+\end{figure}
+
 % Already covered defining the problem. Next will discuss goals and list alternatives. That will feed into threat
-% modeling.
+% modeling
 
-\section{Desireable Outcomes}
+\section{Mapping the EA Debate}
+
+\subsection{Desireable Outcomes}
 
@@
 
-\section{Proposed Alternatives}
+\subsection{Proposed Alternatives}
 

### 2020-10-17 Thesis-Chapter1-Introduction.tex
Update OODA model and begin writing chap 3

@@
 ``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
-\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science and engineering are well equipped to
-answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each attempted solution has
-permanent and often unintended consequences, and likely exists in a pattern of chronic policy failure.
+\cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
+are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
+attempted solution has permanent and often unintended consequences, and likely exists in a pattern of chronic policy
+failure.
 

### 2020-10-17 Thesis-Chapter3-Arguments.tex
Update OODA model and begin writing chap 3

@@
 greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
-to tackling \acp{wicked-prob} succeed and fail? How can we collectively situation ourselves to make real progress? This
+to tackling \acp{wicked-prob} succeed and fail? How can we collectively situate ourselves to make real progress? This
 chapter seeks to answer these questions.
@@
 
+\Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
+In this section we analyze the nature of wicked problems and strategies for tackling them.
+
 \subsection{Characteristics}
+
+Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to classical problem solving
+methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the sweeping
+application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
+structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
+twentieth century, believers in the power of reason thought this progress would march on in the realm of public
+planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
+solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
+primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
+context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:
+
+\begin{displayquote}
+A great many barriers keep us from perfecting such a planning/governing system: theory is inadequate for decent
+forecasting; our intelligence is insufficient to our tasks; plurality of objectives held by pluralities of politics
+makes it impossible to pursue unitary aims; and so on. The difficulties attached to rationality are tenacious, and we
+have so far been unable to get untangled from their web. This is partly because the classical paradigm of science and
+engineering---the paradigm that has underlain modern professionalism---is not applicable to the problems of open
+societal systems. One reason the publics have been attacking the social professions, we believe, is that the cognitive
+and occupational styles of the professions---mimicking the cognitive style of science and the occupational style of
+engineering---have just not worked on a wide array of social problems. The lay customers are complaining because
+planners and other professionals have not succeeded in solving the problems they claimed they could solve. We shall want
+to suggest that the social professions were misled somewhere along the line into assuming they could be applied
+scientists---that they could solve problems in the ways scientists can solve their sorts of problems. The error has been
+a serious one. \cite{rittel_dilemmas_1973}
+\end{displayquote}
+
+When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
+alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation with countless
+cultures and subcultures which are effectively represented by two mutually hostile political parties, agreeing on goals
+is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that defy
+the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
+\cite{rittel_dilemmas_1973}:
+
+\begin{displayquote}
+  \begin{enumerate}
+    \item There is no definitive formulation of a wicked problem [and each formulation presupposes a solution].
+    \item Wicked problems have no stopping rule.
+    \item Solutions to wicked problems are not true-or-false, but good-or-bad.
+    \item There is no immediate and no ultimate test of a solution to a wicked problem.
+    \item Every solution to a wicked problem is a ``one-shot operation''; because there is no opportunity to learn by
+          trial-and-error, every attempt counts significantly.
+    \item Wicked problems do not have an enumerable (or an exhaustively describable) set of potential solutions, nor is
+          there a well-described set of permissible operations that may be incorporated into the plan.
+    \item Every wicked problem is essentially unique.
+    \item Every wicked problem can be considered to be a symptom of another problem.
+    \item The existence of a discrepancy representing a wicked problem can be explained in numerous ways. The choice of
+          explanation determines the nature of the problem's resolution.
+    \item The planner has no right to be wrong [i.e., they are liable for the consequences of their decisions].
+  \end{enumerate}
+\end{displayquote}
+
+Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
+infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
+analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
+conditions.
+
+Several of these characteristics are results of the fact that we have no accurate predictive model of the world and
+human behavior at large. This does not need to be expounded upon.
+
+The more interesting insight is that problem definition is not self-contained. There is no way to frame the problem that
+provides all the information required for a solution...
+
+% Cite commission_tackling_2018 and some other wicked problems
 

### 2020-10-21 Thesis-Chapter3-Arguments.tex
Finish section on wicked prob background

@@
 
-The more interesting insight is that problem definition is not self-contained. There is no way to frame the problem that
-provides all the information required for a solution...
+The more important insight is one step removed: precisely because there is an inexhaustable set of potential solutions,
+the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
+proposed solution demands research and context to be added to the problem definition. This feedback from proposal to
+definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals to
+a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
+themselves.
 
-% Cite commission_tackling_2018 and some other wicked problems
+A 2018 study by the Australian government uses climate change as an example of a \ac{wicked-prob}
+\cite{commission_tackling_2018}. Say the problem is initially formulated as long-term change to the environment caused
+by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms: profligate behavior
+in consumeristic societies must be reigned in at a local and personal level; global inter-governmental coordination is
+the only solution, as individual changes will have no impact; or the situation is overblown by idealists and
+power-mongers, and technological progress and adaptive markets will handle any negative effects that come to pass
+\cite{commission_tackling_2018}. Which response is correct? One cannot tell from the problem statement. How much do
+individual choices contribute to greenhouse gas emissions? How effective are international accords? How costly will the
+changes be, and how capable is technology to respond? Evaluating the validity of each proposal requires updates to the
+problem statement itself.
+
+Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
+amount will result in a certain difference in response with a certain confidence level. Combining the tame elements in a
+cohesive solution is the ``wicked'' part. Despite certain tame elements, one cannot verify the effectiveness of a plan
+ahead of time, port a solution from a different domain, or even definitively evaluate its performance after the fact.
 
@@
 
-- rozenshtein_wicked_2018 convincingly argues the point at length
+Dogged by concerns over privacy, security, safety, and trust, \ac{encryption} and the presupposed solution of \ac{EA} is
+a \ac{wicked-prob}. It has each of the characteristics from Rittel's list above:
+
+\begin{enumerate}
+  \item There is no formulation that encapsulates the problem of encryption's interplay with privacy, security,
+        safety, and trust.
+  \item Balancing each value in the face of constantly evolving technology is never ending.
+  \item \ac{EA} or other proposals cannot definitively solve the problem.
+  \item \ac{EA} or other proposals cannot be objectively tested.
+  \item Every policy implementation has irreversible impacts.
+  \item There is an inexhaustable set of potential solutions to the problem.
+  \item Solutions from other domains do not apply directly to this problem.
+  \item The need for \ac{EA} or some alternative is a symptom of differing values, rapid technological change, and
+        criminal behavior.
+  \item The problem can be framed as insufficient investigatory access to digital data (presupposing \ac{EA} as the
+        solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions), and
+        more.
+  \item The decisions of regulators and technologists alike have real impacts in the world today.
+\end{enumerate}
+
+Cybersecurity law scholar Alan Rozenshtein argues this point at length \cite{rozenshtein_wicked_2018}. He breaks the
+root issues down into three categories. First, that there is disagreement on goals (or even basic premises). Sides can't
+agree whether things are ``going dark'' or we're in the ``golden age of surveillance'' or how much to value competing
+notions of security. Second, that information is ``uncertain and diffuse.'' There are insufficient statistics on
+encryption's effect on investigations, and though the consensus is that we are not presently capable of acceptably
+secure \ac{EA}, that is not certain to always be the case, especially as the area is under-researched. Third, that the
+problem cannot be definitively solved. Evolving values and technology mean that the subject is always up for
+renegotiation.
+
+Rozenshtein reflects on the diagnosis positively:
+
+\begin{displayquote}
+Recognizing that something is a wicked problem is not an admission of its insolubility; rather, it’s just a realistic
+appreciation of its challenges. Progress on difficult social problems reflects, almost by definition, progress on wicked
+problems, whether economic inequality, environmental degradation, or government access to data. Progress can be made,
+but it first requires a clear-eyed appreciation of the nature of the problem and the nature of its challenges.
+\cite{rozenshtein_wicked_2018}
+\end{displayquote}
+
+Embracing reality is the first step to dealing with it \cite{baker_2019}. In the next two sections, we look at
+strategies for dealing with \acp{wicked-prob}.
 

### 2020-10-22 Thesis-Chapter3-Arguments.tex
Develop classical method section

@@
 \subsection{Characteristics}
+\label{wicked-characteristics}
 
@@
 
-Rozenshtein reflects on the diagnosis positively:
+Rozenshtein reflects on the diagnosis optimistically:
 
@@
 
-Embracing reality is the first step to dealing with it \cite{baker_2019}. In the next two sections, we look at
-strategies for dealing with \acp{wicked-prob}.
+Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
+\acp{wicked-prob}. In the next two sections, we look at strategies for dealing with it.
 
@@
 
+This section describes the two common current policymaking approaches, the classical analytical method and ``muddling
+through.''
+
 \subsection{The Classical Analytical Method}
 
+The classical analytical method \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
+\cite{rittel_dilemmas_1973}, ``traditional policy analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking''
+\cite{commission_tackling_2018}) has already been introduced. It is the reason-based method that functions by setting
+goals, identifying problems, evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct
+errors. \myfig{fig-classical-method} illustrates the approach in the context of encryption and \ac{EA}. Per its name, it
+has a linear flow from problem to solution. The refinement step reflects the fact that this method tracks outcomes and
+tweaks solutions over time; however, refinement is always in-kind---it represents a reinforcement, as opposed to a
+reassessment, of the chosen solution.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{dfds/build/OODA-classical.png}
+  \caption[The Classical Analytical Method]{The Classical Analytical Method}
+  \label{fig-classical-method}
+\end{figure}
+
+The shortcomings of the classical analytical method were discussed in \mysec{wicked-characteristics}. It fails due to
+disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
+flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.
+
 \subsection{Muddling Through}
+
+``Muddling through'' represents an alternative policy making strategy.
+
+See \myfig{fig-muddling-through}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
+  \caption[The ``Muddling Through'' Method]{The ``Muddling Through'' Method}
+  \label{fig-muddling-through}
+\end{figure}
 

### 2020-10-24 Thesis-Chapter3-Arguments.tex
Write incrementalism section

@@
 
-Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to classical problem solving
-methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the sweeping
-application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
+Rittel's categorization of \acp{wicked-prob} grew out of frustration with the problems' resistance to classical problem
+solving methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the
+sweeping application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
 structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
@@
 
-This section describes the two common current policymaking approaches, the classical analytical method and ``muddling
-through.''
+This section describes two common policymaking approaches, the \ac{classical-method} and \ac{incrementalism}.
 
@@
 
-The classical analytical method \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
-\cite{rittel_dilemmas_1973}, ``traditional policy analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking''
-\cite{commission_tackling_2018}) has already been introduced. It is the reason-based method that functions by setting
-goals, identifying problems, evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct
-errors. \myfig{fig-classical-method} illustrates the approach in the context of encryption and \ac{EA}. Per its name, it
-has a linear flow from problem to solution. The refinement step reflects the fact that this method tracks outcomes and
-tweaks solutions over time; however, refinement is always in-kind---it represents a reinforcement, as opposed to a
-reassessment, of the chosen solution.
+The \ac{classical-method} \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
+\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_science_1959}, ``traditional policy
+analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking'' \cite{commission_tackling_2018}) has already been
+introduced. It is the reason-based method that functions by setting goals, identifying problems, evaluating
+alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
+illustrates the approach in the context of encryption and \ac{EA}. Per its name, it has a linear flow from problem to
+solution. The refinement step reflects the fact that this method tracks outcomes and tweaks solutions over time;
+however, refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
+solution.
 
@@
 
-The shortcomings of the classical analytical method were discussed in \mysec{wicked-characteristics}. It fails due to
+The shortcomings of the \ac{classical-method} were discussed in \mysec{wicked-characteristics}. It fails due to
 disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
@@
 
-\subsection{Muddling Through}
+\subsection{Incrementalism}
 
-``Muddling through'' represents an alternative policy making strategy.
+\Ac{incrementalism} is an intuitive and iterative approach posed as an alternative to the \ac{classical-method}.
+\Ac{incrementalism} as a policymaking strategy is often referred to as ``\ac{muddling-through}'' after political
+scientist Charles Lindblom's classic paper defining and defending the approach \cite{lindblom_muddling_1959}. Written
+before Rittel developed the idea of ``\acp{wicked-prob},'' Lindblom nonetheless identified many of the same shortcomings
+of the analytical method and sought to formalize the process policymakers were already often using.
 
-See \myfig{fig-muddling-through}.
+Lindblom's ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis. The
+alternatives selected for comparison are defined relative to the status quo and are close enough to one another to be
+able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
+predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
+impossible to impose in a democratic system \cite{lindblom_muddling_1959}. \myfig{fig-muddling-through} illustrates this
+method.
+
+%Footnote: see Lindblom's bracing quote: ``Party behavior is in turn rooted in public attitudes, and political theorists
+%cannot conceive of democracy's surviving in the United States in the absence of fundamental agreement on potentially
+%disruptive issues, with consequent limitation of policy debates to relatively small differences in policy.''
 
@@
   \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
-  \caption[The ``Muddling Through'' Method]{The ``Muddling Through'' Method}
+  \caption[The Incrementalist Method]{The Incrementalist Method}
   \label{fig-muddling-through}
@@
 
-% Mention hermeneutics and complex adaptive systems?
-% As feeley_judicial_2000 points out, muddling through is too focused on, or even consists entirely of, the
-% implementation step. Because of the long tail effects of any implementation in a wicked scenario, and especially in
-% the context of encryption, mere muddling through is too risky. Therefore I suggest iteration. Classical methods, but
-% applied in a cyclical manner.
+\Ac{incrementalism} has several advantages over the \ac{classical-method} for handling \acp{wicked-prob}. It is rooted
+in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved,
+emphasizing iteration (``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
+process of successive approximation to some desired objectives in which what is desired itself continues to change under
+reconsideration.'' \cite{lindblom_muddling_1959}). Most importantly, it eschews the linear
+problem-definition--alternatives-analysis steps out of respect that this is, indeed, impossible. Lindblom argues at
+length that policy ends and means are interlinked, eventually concluding:
+
+\begin{displayquote}
+As to whether the attempt to clarify objectives in advance of policy selection is more or less rational than the close
+intertwining of marginal evaluation and empirical analysis, the principal difference established is that for complex
+[i.e., wicked] problems the first is impossible and irrelevant, and the second is both possible and relevant. The second
+is possible because the administrator need not try to analyze any values except the values by which alternative policies
+differ and need not be concerned with them except as they differ marginally. His need for information on values or
+objectives is drastically reduced as compared with the root [i.e., classical analytic] method; and his capacity for
+grasping, comprehending, and relating values to one another is not strained beyond the breaking point.
+\cite{lindblom_muddling_1959}
+\end{displayquote}
+
+Despite its strengths, \ac{incrementalism} has its weaknesses at addressing \acp{wicked-prob}. Its main weakness is its
+lack of high level strategic analysis. It is incapable of drastic change, which is sometimes necessary. By Lindblom's
+admission, it lacks a safeguard for consideration of all relevant values and may ``overlook excellent policies for no
+other reason than that they are not suggested by the chain of successive policy steps leading up to the present''
+\cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to the next can result in messy policy
+that none are particularly fond of ("As Lindblom's sobriquet suggests, it often led to a considerable muddle''
+\cite{feeley_judicial_2000}). Consider application of the \ac{CFAA} as an example of this kind of muddle
+\cite{wolff_computer_2016}.
+
+Unfortunately, one cannot simply expand the strategic view of the incrementalist method by allowing it to look further
+into the past than the present status quo. Simplifying analysis by limiting it to marginal differences to a given
+baseline is at the heart of \ac{incrementalism}. If one tries to be a ``strategic incrementalist'' by looking into the
+past, they still have to choose a baseline from which to perform analysis. This approach is susceptible to two
+weaknesses in baseline-based reasoning that Rozenshtein described in his analysis of \ac{EA} as a \ac{wicked-prob}
+\cite{rozenshtein_wicked_2018}. First, the choice of baseline is arbitrary, yet heavily colors analysis:
+
+\begin{displayquote}
+Both the government and its critics have operated from the status-quo baseline, though from opposite directions. For the
+government, the relevant baseline is recent history---specifically, right before companies like Apple and WhatsApp
+encrypted their products. From this baseline, the government's ability to surveil has diminished. For critics of
+government surveillance, the relevant baseline is the pre-digital age, before smartphones and social media vastly
+expanded the government's surveillance capabilities. From this baseline, the technological changes underlying the
+``going dark'' problem are mere blips on the otherwise rocketing growth of the surveillance state.
+\cite{rozenshtein_wicked_2018}
+\end{displayquote}
+
+Second, unlike legal baselines, policy baselines do not carry normative force. Constant changes in the underlying
+situation mean that even optimal policy in the past is not necessarily desirable in the present. Due to these
+weaknesses, applying incrementalist methods at the strategic level does not work.
+
+A final weakness in \ac{incrementalism} is its assumption of basic agreement and political stability. The method works
+by limiting analysis to marginal comparisons of broadly similar and familiar proposals. Proposals that differ widely
+from one another or the status quo are considered irrelevant because the debating parties both share the same general
+goals and lack the ability to unilaterally impose their will. Rather a symptom of present circumstances (see
+\mysec{sec-history-current}) than an inherent weakness in the incrementalist approach, both of these assumptions are
+incorrect.
 

### 2020-10-24 Thesis-Glossary.tex
Make more writing and editing improvements
Add OODA loop diagram and notes; update outline
Write incrementalism section

@@
 }
+
+\newterm{classical-method}{classical analytical method}{
+    A policy making strategy that emphasizes rationalism and functions by setting goals, identifying problems,
+    evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct errors
+}
+
+\newterm{incrementalism}{incrementalism}{
+    A policy making strategy that functions by taking successive steps based on the status quo and chosen through
+    comparative analysis; also known as ``muddling through''
+}
+
+\newterm{muddling-through}{muddling through}{
+    See ``incrementalism''
+}
 % End terms
@@
 \newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
+\newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
+\newabbreviation{OODAloop}{OODA loop}{\ii{Observe-Orient-Decide-Act} loop}
 

### 2020-10-24 Thesis-Chapter3-Arguments.tex
Begin writing wicked problem strategy lessons

@@
 predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
-impossible to impose in a democratic system \cite{lindblom_muddling_1959}. \myfig{fig-muddling-through} illustrates this
-method.
+impossible to impose in a democratic system \cite{lindblom_muddling_1959}. The formality of the process varies;
+policymakers may use this method consciously, with considerable comparative analysis, or unconsciously, led by
+intuition. \myfig{fig-muddling-through} illustrates this method.
 
@@
 \cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to the next can result in messy policy
-that none are particularly fond of ("As Lindblom's sobriquet suggests, it often led to a considerable muddle''
-\cite{feeley_judicial_2000}). Consider application of the \ac{CFAA} as an example of this kind of muddle
-\cite{wolff_computer_2016}.
+that none are particularly fond of---``As Lindblom's sobriquet suggests, it often [leads] to a considerable muddle''
+\cite{feeley_judicial_2000}. Consider application of the \ac{CFAA} as one such muddle \cite{wolff_computer_2016}.
 
@@
 
-- Avoid pitfalls of
-  - Easy answers
-    - EA mandates, complexity denial
-  - Fallacies!
-  - Bad results with long technical tails
+The \ac{classical-method} and \ac{incrementalism} are not the only styles of policymaking. They represent perhaps two
+extremes on a spectrum of rational planning and intuitive muddling. Both have their strengths and weaknesses. One may
+intuit that a reasonable strategy is the selective use of both approaches according to the situation, and indeed this
+has also been formally suggested \cite{etzioni_scanning_1967}. However, even a combination of these methods does not
+precisely suit all classes of wicked problems. Is a problem tame? Use the \ac{classical-method}. Is it wicked, but
+strategically under control and relatively non-controversial? Use \ac{incrementalism}. Is it wicked, and lacking a
+strategic response or highly controversial? What then?
+
+There is no \ac{wicked-prob} handbook, but several sources offer advice.
+
+\newcommand{\wickedtipsstart}{\begin{itemize}}
+\newcommand{\wickedtip}[2]{ % Name, citation, description
+    \item \textbf{#1} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
+}
+\newcommand{\wickedtipsend}{\end{itemize}}
+
+\wickedtipsstart
+
+    \wickedtip{Reject Easy Answers}{
+    }
+
+    \wickedtip{Bring Everyone to the Table}{
+    }
+
+    \wickedtip{Embrace Flexible, Risk-Based Solutions}{
+    }
+
+    \wickedtip{Ensure Transparency}{
+    }
+
+\wickedtipsend
+
+Rittel's follow-up research offers another tool: argument mapping.
+
+% - Avoid pitfalls of
+%   - Easy answers
+%     - EA mandates, complexity denial
+%   - Fallacies!
+%   - Bad results with long technical tails
 
@@
 commission_tackling_2018:
-- Strategies
-  - Authoritative
-  - Competitive
-  - Collaborative
-- Approaches
-  - Linear problem solving
-    - Ineffective! Several reasons, good breakdown. Dovetails with rozenshtein_wicked_2018.
-  - Holistic problem solving: adaptive, innovative, flexible, cross-organizational
-- Avoid
-  - Narrow solutions
+% - Strategies
+%   - Authoritative
+%   - Competitive
+%   - Collaborative
+% - Approaches
+%   - Linear problem solving
+%     - Ineffective! Several reasons, good breakdown. Dovetails with rozenshtein_wicked_2018.
+%   - Holistic problem solving: adaptive, innovative, flexible, cross-organizational
+% - Avoid
+%   - Narrow solutions
 
 rozenshtein_wicked_2018:
-- Approaches
-  - Muddling through (he cites Charles E. Lindblom, just like feeley_judicial_2000!)
-  - Risk management (--finally, something familiar to security experts!)
-    - So, a broader view of security, inclusive of both public safety and information security
-    - In this model, threat actors now include terrorists, CSAM distributors, organized crime
-      - Add a new character? Evil Alice and evil Bob (Alex and Becca?)? This automatically raises to question of how to
-          identify bad actors--AI (lee_detecting_2020)? PhotoDNA (microsoft_2020)? Government orders (warrants)? Could
-          be useful.
-    - I have already suggested adding government mandates to the traditional IS threat model
-  - Imperfect solutions, like lawful hacking
-  - Meanwhile, invest in knowledge production and improved relationships (like "cross-organizational" from
-      commission_tackling_2018)
-- Avoid
-  - Easy answers ("narrow solutions" from commission_tackling_2018)
-    - Such as EA mandates or "golden age of surveillance"-based denials
+% - Approaches
+  % - Muddling through (he cites Charles E. Lindblom, just like feeley_judicial_2000!)
+  % - Risk management (--finally, something familiar to security experts!)
+  %   - So, a broader view of security, inclusive of both public safety and information security
+  %   - In this model, threat actors now include terrorists, CSAM distributors, organized crime
+  %     - Add a new character? Evil Alice and evil Bob (Alex and Becca?)? This automatically raises to question of how to
+  %         identify bad actors--AI (lee_detecting_2020)? PhotoDNA (microsoft_2020)? Government orders (warrants)? Could
+  %         be useful.
+  %   - I have already suggested adding government mandates to the traditional IS threat model
+  % - Imperfect solutions, like lawful hacking
+  % - Meanwhile, invest in knowledge production and improved relationships (like "cross-organizational" from
+  %     commission_tackling_2018)
+% - Avoid
+%   - Easy answers ("narrow solutions" from commission_tackling_2018)
+%     - Such as EA mandates or "golden age of surveillance"-based denials
 
@@
   - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
-  - Use transparency and accountability (through technical means)
+  % - Use transparency and accountability (through technical means)
 
-feeley_judicial_2000:
-- Classical analytic method ("linear" in commission_tackling_2018)
-% 1. Classical analytic method; each step has its own approaches, e.g. cost-benefit analysis (loosely based on natural
-% sciences)
-%   (missing: Observe)
-%   a. Define the problem (Orient)
-%   b. Identify a goal (Orient)
-%   c. Generate a range of alternatives for achieving that goal (Orient)
-%   d. Select the alternative that seems most promising (Decide)
-%   e. Implement the selected alternative (Act)
-%   (missing: loop)
-- Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)
-- The hermeneutic circle
+% feeley_judicial_2000:
+% - Classical analytic method ("linear" in commission_tackling_2018)
+% % 1. Classical analytic method; each step has its own approaches, e.g. cost-benefit analysis (loosely based on natural
+% % sciences)
+% %   (missing: Observe)
+% %   a. Define the problem (Orient)
+% %   b. Identify a goal (Orient)
+% %   c. Generate a range of alternatives for achieving that goal (Orient)
+% %   d. Select the alternative that seems most promising (Decide)
+% %   e. Implement the selected alternative (Act)
+% %   (missing: loop)
+% - Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)
+% - The hermeneutic circle
 

### 2020-10-25 scripts/count-citations.sh
Modify count-citations (cite) to support searching for terms

@@ -4,8 +4,20 @@ cd $(git rev-parse --show-toplevel)
 source ./scripts/sourceme.sh
 # set -x
 
+function grep_tex() {
+    grep --color=auto -rniP --include \*.tex --exclude Thesis-Glossary.tex "$1" .
+}
+
+function grep_citation() {
+    grep_tex "\\\\cite{$1}"
+}
+
+function grep_term() {
+    grep_tex "\\\\ac[slfp]?{$1}"
+}
+
 function print_number_in_tex() {
-    echo $(grep "{$1}" *.tex | wc -l)
+    echo $((grep_citation $1 || grep_term $1) | wc -l)
 }
 
 function print_number_in_argdown() {
@@ -37,6 +49,6 @@ elif (( 2 <= $# )) && [[ "-n" == "$1" ]]; then
 else
     print_count $1
     echo
-    grep --color=auto -rn --include \*.tex "\\\\cite{$1}" .
-    grep --color=auto -rn --include \*.ad "(cite:${1/_/\\\\_})" .
+    grep_citation $1 || grep_term $1
+    grep --color=auto -rni --include \*.ad "(cite:${1/_/\\\\_})" .
 fi

### 2020-10-25 Thesis-Chapter1-Introduction.tex
Separate strategies and mapping chapters

@@
 brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
-models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
-a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
-thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.
+models using data flow diagrams. In the following chapters, I analyze strategies for tacking \acp{wicked-prob},
+map the \acl{EA} arguments in detail, and create a multi-level threat model, broken down by the significantly different
+challenges posed by \acl{DAR} and \acl{DIM}. The thesis concludes with a set of policies and principles to apply to the
+problem in light of valid arguments and threats.

### 2020-10-25 Thesis-Chapter3-PolicyMaking.tex
Separate strategies and mapping chapters

--- 2020-10-24 Thesis-Chapter3-Arguments.tex
@@
-\chapter{Wicked Problems and the EA Debate}
-\label{chap-arguments}
+\chapter{Strategy for Tackling Wicked Problems}
+\label{chap-policy}
 
@@
 Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
-amount will result in a certain difference in response with a certain confidence level. Combining the tame elements in a
-cohesive solution is the ``wicked'' part. Despite certain tame elements, one cannot verify the effectiveness of a plan
-ahead of time, port a solution from a different domain, or even definitively evaluate its performance after the fact.
+amount will result in a certain difference in response with a certain statistical confidence. Refining the problem
+statement and combining the tame elements in a cohesive solution is the ``wicked'' part. Despite certain tame elements,
+one cannot verify the effectiveness of a plan ahead of time, port a solution from a different domain, or even
+definitively evaluate its performance after the fact.
 
@@
   \item \ac{EA} or other proposals cannot be objectively tested.
-  \item Every policy implementation has irreversible impacts.
+  \item Every policy implementation has irreversible effects.
   \item There is an inexhaustable set of potential solutions to the problem.
@@
 
-Cybersecurity law scholar Alan Rozenshtein argues this point at length \cite{rozenshtein_wicked_2018}. He breaks the
-root issues down into three categories. First, that there is disagreement on goals (or even basic premises). Sides can't
-agree whether things are ``going dark'' or we're in the ``golden age of surveillance'' or how much to value competing
-notions of security. Second, that information is ``uncertain and diffuse.'' There are insufficient statistics on
-encryption's effect on investigations, and though the consensus is that we are not presently capable of acceptably
-secure \ac{EA}, that is not certain to always be the case, especially as the area is under-researched. Third, that the
-problem cannot be definitively solved. Evolving values and technology mean that the subject is always up for
-renegotiation.
+Cybersecurity law scholar Alan Rozenshtein argues for treating \ac{encryption} and \ac{EA} as a \ac{wicked-prob} at
+length \cite{rozenshtein_wicked_2018}. He breaks the root issues down into three categories. First, that there is
+disagreement on goals (or even basic premises). Sides can't agree whether things are ``going dark'' or we're in the
+``golden age of surveillance'' or how much to value competing notions of security. Second, that information is
+``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations, and though the
+consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be the case,
+especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving values and
+technology mean that the subject is always up for renegotiation.
 
@@
 The \ac{classical-method} \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
-\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_science_1959}, ``traditional policy
+\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_muddling_1959}, ``traditional policy
 analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking'' \cite{commission_tackling_2018}) has already been
@@
 alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
-illustrates the approach in the context of encryption and \ac{EA}. Per its name, it has a linear flow from problem to
-solution. The refinement step reflects the fact that this method tracks outcomes and tweaks solutions over time;
-however, refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
+illustrates the approach in the context of encryption and \ac{EA}. It has a purely linear flow from problem to solution
+except for the ``refinement'' step, in which the method analyzes outcomes and corrects errors. However, it is important
+to note that refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
 solution.
@@
 that none are particularly fond of---``As Lindblom's sobriquet suggests, it often [leads] to a considerable muddle''
-\cite{feeley_judicial_2000}. Consider application of the \ac{CFAA} as one such muddle \cite{wolff_computer_2016}.
+\cite{feeley_judicial_2000}. Consider application of the \acl{CFAA} as one such muddle \cite{wolff_computer_2016}.
 
@@
 past, they still have to choose a baseline from which to perform analysis. This approach is susceptible to two
-weaknesses in baseline-based reasoning that Rozenshtein described in his analysis of \ac{EA} as a \ac{wicked-prob}
+weaknesses in baseline-based reasoning that Rozenshtein describes in his analysis of \ac{EA} as a \ac{wicked-prob}
 \cite{rozenshtein_wicked_2018}. First, the choice of baseline is arbitrary, yet heavily colors analysis:
@@
   \wickedtip{Reject Easy Answers}{
+
+Easy answers, or any solutions that artificially tame the problem, must be rejected.
+
 }
@@
 
-% Already covered defining the problem. Next will discuss goals and list alternatives. That will feed into threat
-% modeling
-
-\section{Mapping the EA Debate}
-
-\subsection{Desireable Outcomes}
-
-% See committee_decrypting_2018!
-% For a wholistic threat model, include both the threats of government (legislation) and of criminal abusers of your
-% system.
-
-\subsection{Proposed Alternatives}
-
-% See committee_decrypting_2018!
-
-commission_tackling_2018:
+% commission_tackling_2018:
 % - Strategies
@@
 
-rozenshtein_wicked_2018:
+% rozenshtein_wicked_2018:
 % - Approaches
@@
 
-matyas_incommensurability_2018:
-- Approaches to reducing incommensurability
-  - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
+% matyas_incommensurability_2018:
+% - Approaches to reducing incommensurability
+%   - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
   % - Use transparency and accountability (through technical means)
@@
 % consider balance of principles and rules.
-
-% Subject one: do we even want EA?
-%  --> no? exit here
-%  --> yes? carry on
-%  --> pragmatist? carry on
-% include fallacies argument map here
-% see rozenshtein_wicked_2018, benaloh_what_2018, soesanto_2018, ministerial_2018, fbi_2020, brantley_2018, lund_2020
-% ... group_2019, ruiz_there_2018, rozenshtein_2019
-
-% Subject two: what properties would good EA have?
-% easy for LEAs to use? hard? expensive? voluntary? unable to be used for mass surveillance? transparent (attempts are
-%   visible)? alignment of values with security (e.g. discourages bug hoarding)? adaptable (since solution won't be
-%   permanent, either technically or according to collective will as things change rozenshtein_wicked_2018)? suited to
-%   rules and principles (matyas_incommensurability_2018)?
-
-% Subject three: what options do we have?
-% map of properties onto proposals--leads into threat modeling, which will check the veracity of claims
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% In this chapter, develop the arguments employed in EA debates. Make use of argument maps to break down and visualize the
-% debate, as shown in \myfig{fig-args-top-level}.
-
-% \begin{figure}[h]
-%     \centering\CaptionFontSize
-%     \includegraphics[width=\linewidth]{arguments/build/top-level.pdf}
-%     \caption[Argument: Top Level]{Top level arguments for and against exceptional access.}
-%     \label{fig-args-top-level}
-% \end{figure}
-
-% The debate over exceptional access takes place on two fundamental levels:
-% \begin{enumerate}
-%     \item Whether cryptosystems with exceptional access represent a net benefit to society
-%     \item Whether and how to conduct exceptional access research
-% \end{enumerate}
-
-% \subsection{Arguments For}
-
-% Why EA is a benefit to society:
-% \begin{itemize}
-%     \item Absolute privacy is not a right
-%     \item Encryption gives too much power to the individual
-%     \item EA would allow more crimes to be prevented and prosecuted
-% \end{itemize}
-
-% Why to research EA:
-% \begin{itemize}
-%     \item EA done well would be more secure and privacy preserving than EA done poorly
-%     \item The threat of legislative action creates urgency
-% \end{itemize}
-
-% \subsection{Arguments Against}
-
-% Why EA is a danger to society:
-% \begin{itemize}
-%     \item Privacy is an absolute right
-%     \item Exceptional access gives too much power to the state
-%     \item Exceptional access is not possible in an acceptably secure manner
-% \end{itemize}
-
-% Why not to research EA:
-% \begin{itemize}
-%     \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
-%             them
-%     \item The technical community should not entertain delusions about what is truly possible
-% \end{itemize}
-
-% \subsection{Fallacious Arguments}
-
-% Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}. Also cited:
-% \cite{hanna_2019}.
-
-% Fallacies are as shown in \myfig{fig-args-fallacies}.
-
-% \begin{figure}[h]
-%     \centering\CaptionFontSize
-%     \myincludeargument{comprehensive.fallacies}
-%     \caption[Fallacious Arguments]{Fallacious arguments used in the exceptional access debate.}
-%     \label{fig-args-fallacies}
-% \end{figure}

### 2020-10-25 Thesis-Chapter4-Arguments.tex
Separate strategies and mapping chapters

--- 2020-10-22 Thesis-Chapter3-Arguments.tex
@@
-\chapter{Wicked Problems and the EA Debate}
+\chapter{EA Debate Argument Maps}
 \label{chap-arguments}
-
-Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
-greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
-to tackling \acp{wicked-prob} succeed and fail? How can we collectively situate ourselves to make real progress? This
-chapter seeks to answer these questions.
-
-\section{Wicked Problems}
-
-\Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
-In this section we analyze the nature of wicked problems and strategies for tackling them.
-
-\subsection{Characteristics}
-\label{wicked-characteristics}
-
-Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to classical problem solving
-methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the sweeping
-application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
-structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
-twentieth century, believers in the power of reason thought this progress would march on in the realm of public
-planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
-solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
-primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
-context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:
-
-\begin{displayquote}
-A great many barriers keep us from perfecting such a planning/governing system: theory is inadequate for decent
-forecasting; our intelligence is insufficient to our tasks; plurality of objectives held by pluralities of politics
-makes it impossible to pursue unitary aims; and so on. The difficulties attached to rationality are tenacious, and we
-have so far been unable to get untangled from their web. This is partly because the classical paradigm of science and
-engineering---the paradigm that has underlain modern professionalism---is not applicable to the problems of open
-societal systems. One reason the publics have been attacking the social professions, we believe, is that the cognitive
-and occupational styles of the professions---mimicking the cognitive style of science and the occupational style of
-engineering---have just not worked on a wide array of social problems. The lay customers are complaining because
-planners and other professionals have not succeeded in solving the problems they claimed they could solve. We shall want
-to suggest that the social professions were misled somewhere along the line into assuming they could be applied
-scientists---that they could solve problems in the ways scientists can solve their sorts of problems. The error has been
-a serious one. \cite{rittel_dilemmas_1973}
-\end{displayquote}
-
-When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
-alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation with countless
-cultures and subcultures which are effectively represented by two mutually hostile political parties, agreeing on goals
-is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that defy
-the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
-\cite{rittel_dilemmas_1973}:
-
-\begin{displayquote}
-  \begin{enumerate}
-    \item There is no definitive formulation of a wicked problem [and each formulation presupposes a solution].
-    \item Wicked problems have no stopping rule.
-    \item Solutions to wicked problems are not true-or-false, but good-or-bad.
-    \item There is no immediate and no ultimate test of a solution to a wicked problem.
-    \item Every solution to a wicked problem is a ``one-shot operation''; because there is no opportunity to learn by
-          trial-and-error, every attempt counts significantly.
-    \item Wicked problems do not have an enumerable (or an exhaustively describable) set of potential solutions, nor is
-          there a well-described set of permissible operations that may be incorporated into the plan.
-    \item Every wicked problem is essentially unique.
-    \item Every wicked problem can be considered to be a symptom of another problem.
-    \item The existence of a discrepancy representing a wicked problem can be explained in numerous ways. The choice of
-          explanation determines the nature of the problem's resolution.
-    \item The planner has no right to be wrong [i.e., they are liable for the consequences of their decisions].
-  \end{enumerate}
-\end{displayquote}
-
-Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
-infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
-analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
-conditions.
-
-Several of these characteristics are results of the fact that we have no accurate predictive model of the world and
-human behavior at large. This does not need to be expounded upon.
-
-The more important insight is one step removed: precisely because there is an inexhaustable set of potential solutions,
-the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
-proposed solution demands research and context to be added to the problem definition. This feedback from proposal to
-definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals to
-a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
-themselves.
-
-A 2018 study by the Australian government uses climate change as an example of a \ac{wicked-prob}
-\cite{commission_tackling_2018}. Say the problem is initially formulated as long-term change to the environment caused
-by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms: profligate behavior
-in consumeristic societies must be reigned in at a local and personal level; global inter-governmental coordination is
-the only solution, as individual changes will have no impact; or the situation is overblown by idealists and
-power-mongers, and technological progress and adaptive markets will handle any negative effects that come to pass
-\cite{commission_tackling_2018}. Which response is correct? One cannot tell from the problem statement. How much do
-individual choices contribute to greenhouse gas emissions? How effective are international accords? How costly will the
-changes be, and how capable is technology to respond? Evaluating the validity of each proposal requires updates to the
-problem statement itself.
-
-Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
-amount will result in a certain difference in response with a certain confidence level. Combining the tame elements in a
-cohesive solution is the ``wicked'' part. Despite certain tame elements, one cannot verify the effectiveness of a plan
-ahead of time, port a solution from a different domain, or even definitively evaluate its performance after the fact.
-
-\subsection{Encryption and EA as a Wicked Problem}
-
-Dogged by concerns over privacy, security, safety, and trust, \ac{encryption} and the presupposed solution of \ac{EA} is
-a \ac{wicked-prob}. It has each of the characteristics from Rittel's list above:
-
-\begin{enumerate}
-  \item There is no formulation that encapsulates the problem of encryption's interplay with privacy, security,
-        safety, and trust.
-  \item Balancing each value in the face of constantly evolving technology is never ending.
-  \item \ac{EA} or other proposals cannot definitively solve the problem.
-  \item \ac{EA} or other proposals cannot be objectively tested.
-  \item Every policy implementation has irreversible impacts.
-  \item There is an inexhaustable set of potential solutions to the problem.
-  \item Solutions from other domains do not apply directly to this problem.
-  \item The need for \ac{EA} or some alternative is a symptom of differing values, rapid technological change, and
-        criminal behavior.
-  \item The problem can be framed as insufficient investigatory access to digital data (presupposing \ac{EA} as the
-        solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions), and
-        more.
-  \item The decisions of regulators and technologists alike have real impacts in the world today.
-\end{enumerate}
-
-Cybersecurity law scholar Alan Rozenshtein argues this point at length \cite{rozenshtein_wicked_2018}. He breaks the
-root issues down into three categories. First, that there is disagreement on goals (or even basic premises). Sides can't
-agree whether things are ``going dark'' or we're in the ``golden age of surveillance'' or how much to value competing
-notions of security. Second, that information is ``uncertain and diffuse.'' There are insufficient statistics on
-encryption's effect on investigations, and though the consensus is that we are not presently capable of acceptably
-secure \ac{EA}, that is not certain to always be the case, especially as the area is under-researched. Third, that the
-problem cannot be definitively solved. Evolving values and technology mean that the subject is always up for
-renegotiation.
-
-Rozenshtein reflects on the diagnosis optimistically:
-
-\begin{displayquote}
-Recognizing that something is a wicked problem is not an admission of its insolubility; rather, it’s just a realistic
-appreciation of its challenges. Progress on difficult social problems reflects, almost by definition, progress on wicked
-problems, whether economic inequality, environmental degradation, or government access to data. Progress can be made,
-but it first requires a clear-eyed appreciation of the nature of the problem and the nature of its challenges.
-\cite{rozenshtein_wicked_2018}
-\end{displayquote}
-
-Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
-\acp{wicked-prob}. In the next two sections, we look at strategies for dealing with it.
-
-\section{Failure of Current Policymaking Approaches}
-
-This section describes the two common current policymaking approaches, the classical analytical method and ``muddling
-through.''
-
-\subsection{The Classical Analytical Method}
-
-The classical analytical method \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
-\cite{rittel_dilemmas_1973}, ``traditional policy analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking''
-\cite{commission_tackling_2018}) has already been introduced. It is the reason-based method that functions by setting
-goals, identifying problems, evaluating alternatives, implementing solutions, and analyzing outcomes in order to correct
-errors. \myfig{fig-classical-method} illustrates the approach in the context of encryption and \ac{EA}. Per its name, it
-has a linear flow from problem to solution. The refinement step reflects the fact that this method tracks outcomes and
-tweaks solutions over time; however, refinement is always in-kind---it represents a reinforcement, as opposed to a
-reassessment, of the chosen solution.
-
-\begin{figure}[h]
-  \centering\CaptionFontSize
-  \includegraphics[width=\linewidth]{dfds/build/OODA-classical.png}
-  \caption[The Classical Analytical Method]{The Classical Analytical Method}
-  \label{fig-classical-method}
-\end{figure}
-
-The shortcomings of the classical analytical method were discussed in \mysec{wicked-characteristics}. It fails due to
-disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
-flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.
-
-\subsection{Muddling Through}
-
-``Muddling through'' represents an alternative policy making strategy.
-
-See \myfig{fig-muddling-through}.
-
-\begin{figure}[h]
-  \centering\CaptionFontSize
-  \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
-  \caption[The ``Muddling Through'' Method]{The ``Muddling Through'' Method}
-  \label{fig-muddling-through}
-\end{figure}
-
-% Mention hermeneutics and complex adaptive systems?
-% As feeley_judicial_2000 points out, muddling through is too focused on, or even consists entirely of, the
-% implementation step. Because of the long tail effects of any implementation in a wicked scenario, and especially in
-% the context of encryption, mere muddling through is too risky. Therefore I suggest iteration. Classical methods, but
-% applied in a cyclical manner.
-
-\subsection{Lessons}
-
-- Avoid pitfalls of
-  - Easy answers
-    - EA mandates, complexity denial
-  - Fallacies!
-  - Bad results with long technical tails
-
-% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
-% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
-% decrypted! So must be very careful to avoid solution attempts that leave long scars.
-
-\section{Proposal: The OODA Loop for Policymaking}
-
-\ac{OODAloop}.
-
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
-    \caption[The OODA Loop]{The OODA Loop}
-    \label{fig-ooda-loop}
-\end{figure}
 
@@
 % See committee_decrypting_2018!
-
-commission_tackling_2018:
-- Strategies
-  - Authoritative
-  - Competitive
-  - Collaborative
-- Approaches
-  - Linear problem solving
-    - Ineffective! Several reasons, good breakdown. Dovetails with rozenshtein_wicked_2018.
-  - Holistic problem solving: adaptive, innovative, flexible, cross-organizational
-- Avoid
-  - Narrow solutions
-
-rozenshtein_wicked_2018:
-- Approaches
-  - Muddling through (he cites Charles E. Lindblom, just like feeley_judicial_2000!)
-  - Risk management (--finally, something familiar to security experts!)
-    - So, a broader view of security, inclusive of both public safety and information security
-    - In this model, threat actors now include terrorists, CSAM distributors, organized crime
-      - Add a new character? Evil Alice and evil Bob (Alex and Becca?)? This automatically raises to question of how to
-          identify bad actors--AI (lee_detecting_2020)? PhotoDNA (microsoft_2020)? Government orders (warrants)? Could
-          be useful.
-    - I have already suggested adding government mandates to the traditional IS threat model
-  - Imperfect solutions, like lawful hacking
-  - Meanwhile, invest in knowledge production and improved relationships (like "cross-organizational" from
-      commission_tackling_2018)
-- Avoid
-  - Easy answers ("narrow solutions" from commission_tackling_2018)
-    - Such as EA mandates or "golden age of surveillance"-based denials
-
-matyas_incommensurability_2018:
-- Approaches to reducing incommensurability
-  - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
-  - Use transparency and accountability (through technical means)
-
-feeley_judicial_2000:
-- Classical analytic method ("linear" in commission_tackling_2018)
-% 1. Classical analytic method; each step has its own approaches, e.g. cost-benefit analysis (loosely based on natural
-% sciences)
-%   (missing: Observe)
-%   a. Define the problem (Orient)
-%   b. Identify a goal (Orient)
-%   c. Generate a range of alternatives for achieving that goal (Orient)
-%   d. Select the alternative that seems most promising (Decide)
-%   e. Implement the selected alternative (Act)
-%   (missing: loop)
-- Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)
-- The hermeneutic circle
-
-% How to tackle wicked problems in general? Take some tips from commission_tackling_2018, rozenshtein_wicked_2018,
-% feeley_judicial_2000, and matyas_incommensurability_2018:
-% Correctly state the problem--identify the root causes. They aren't going away, there by definition is no clear long
-% term solution. Put the proposed solutions in context--none are perfect, none are permanent. For EA in particular,
-% consider balance of principles and rules.
 

### 2020-10-25 Thesis-Chapter5-ThreatModel.tex
Separate strategies and mapping chapters

--- 2020-02-28 Thesis-Chapter4-ThreatModel.tex
@@
 everything.
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
+    \caption[A DAR EA Proposal]{A DAR EA Proposal based on \cite{savage_lawful_2018}.}
+    \label{fig-dfd-lawful}
+\end{figure}
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/DIM-level-1-crypto-E2E-signal-X3DH.png}
+    \caption[The Signal X3DH Protocol]{The Signal session creation protocol \cite{marlinspike_x3dh_2016}.}
+    \label{fig-dfd-lawful}
+\end{figure}
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/DIM-level-1-crypto-E2E-signal-Double-Ratchet.png}
+    \caption[The Signal Double Ratchet Protocol]{The Signal \ac{E2EE} protocol \cite{marlinspike_2016}.}
+    \label{fig-dfd-lawful}
+\end{figure}
+
+% For analysis of threat actors, see bindenagel_discussion_2017
+
+% Generalize for DAR and DIM, then apply to specific examples. This proves that my analysis is helpful. Ex: apply to email: is a DIM _and_ DAR problem due to the server.
+
+%%%%%%%%%%%%
+% Chapter 4, threat models
+% (how to be specific enough to be useful but generic enough to apply to many systems? Focus on the EA type (DAR v DIM),
+%   then apply to specific systems)
+% DAR, DIM threat models (make sure to cover abelson_2015)
+% Basic proposals
+%  --> mappings onto threat models
+% At work: 3 example systems analyzed with this threat model
+% 1. IM (E2E IM too?) -- in motion
+% 2. Cloud storage -- at rest
+% 3. Mobile phone -- at rest, but way different scenario than cloud
+%  --> include Ozzie's CLEAR, since it was higher profile
+%%%%%%%%%%%%

### 2020-10-25 Thesis-Chapter3-PolicyMaking.tex
Finish policy lessons section

@@
 
-\subsection{The Classical Analytical Method}
+\subsection{The Classical Analytic Method}
 
@@
   \includegraphics[width=\linewidth]{dfds/build/OODA-classical.png}
-  \caption[The Classical Analytical Method]{The Classical Analytical Method}
+  \caption[The Classical Analytic Method]{The Classical Analytic Method}
   \label{fig-classical-method}
@@
 before Rittel developed the idea of ``\acp{wicked-prob},'' Lindblom nonetheless identified many of the same shortcomings
-of the analytical method and sought to formalize the process policymakers were already often using.
+of the classical method and sought to formalize the process policymakers were already often using.
 
@@
 
-There is no \ac{wicked-prob} handbook, but several sources offer advice.
+Let us define one strategy for handling \ac{wicked-prob} as corralling it back into the classes of problems we are
+better at dealing with. That means breaking it into tame sub-problems and developing agreeable strategies for its
+irreducibly wicked elements. Once that is done, we can rationally analyze and increment our way to victory. There is no
+handbook for how to do this, but several sources offer advice.
 
@@
 
-Easy answers, or any solutions that artificially tame the problem, must be rejected.
+Easy answers, or any solutions that artificially tame the problem, will not bring the matter under control. While by
+definition there is no step that will truly solve the problem, easy answers deliberately emphasize one value to the
+exclusion of others, leaving chaos in those neglected areas. As a result of ignoring root issues feeding the problem,
+such solutions are even more likely to produce unintended consequences \cite{commission_tackling_2018}.
 
@@
   \wickedtip{Bring Everyone to the Table}{
+
+Making sure every relevant group is included is important for two reasons. First, because information is ``uncertain and
+diffuse,'' generating an accurate problem statement requires diverse input \cite{rozenshtein_wicked_2018}. Second, for
+\ac{incrementalism} to work on the irreducibly wicked roots of the problem, there needs to be some degree of consensus
+on overall strategy. Consensus building is not easy among groups with differing values and priorities, but it is
+impossible without each group being represented.
+
+}
+
+  \wickedtip{Unite Problem Definition and Analysis Steps}{
+
+The failures of the \ac{classical-method} and \ac{incrementalism} teach that we must be able to think strategically
+while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
+classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
+practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
+puts it, ``The systems-approach `of the first generation' [i.e., classical analytics] is inadequate for dealing with
+wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
+process in the course of which an image of the problem and of the solution emerges gradually among the participants, as
+a product of incessant judgment, subjected to critical argument'' \cite{rittel_dilemmas_1973}.
+
 }
@@
 \wickedtip{Embrace Flexible, Risk-Based Solutions}{
+
+We know that proposals cannot be comprehensively evaluated before or even after implementation, and that each action (or
+decision not to act) has irreversible effects. \Ac{incrementalism} is the decision making model to follow here, as it is
+inherently responsive to unpredictable changes in the environment. Proposals must therefore be agile. All decisions
+involve unknowns, but risk- and uncertainty-management strategies can optimize the expected outcome and maximize the
+worst outcome \cite{sunstein_beyond_2015}.
+
 }
 
-  \wickedtip{Ensure Transparency}{
+  \wickedtip{Focus Discussion around Concrete Proposals}{
+
+This follows partially from the previous suggestion---problem definition and analysis are combined precisely because the
+problem definition depends on the nature of proposed solutions. But this advice deserves emphasis for another reason:
+consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
+due to their conflicting values and priorities, but Linblom writes encouragingly about ``the ease with which individuals
+of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to
+say, ``Labor mediators report a similar phenomenon: the contestants cannot agree on criteria for settling their disputes
+but can agree on specific proposals. Similarly, when one administrator's objective turns out to be another's means, they
+often can agree on policy'' \cite{lindblom_muddling_1959}.
+
+% So should I follow my own advice and focus on
+%   group_2019 (argues for DAR) and
+%   savage_lawful_2018 (a concrete DAR proposal)? -- and related, servan_schreiber_jje_2020
+
 }

### 2020-10-25 Thesis-Glossary.tex
Finish policy lessons section

@@
 
-\newterm{classical-method}{classical analytical method}{
+\newterm{classical-method}{classical analytic method}{
     A policy making strategy that emphasizes rationalism and functions by setting goals, identifying problems,

### 2020-10-25 Thesis-Chapter1-Introduction.tex
Termify argument maps and remove strategy notes

--- 2020-10-17 Thesis-Chapter1-Introduction.tex
@@
 This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
-models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
-a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
-thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.
+brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce \acp{argmap} and threat
+models using data flow diagrams. In the following chapters, I analyze strategies for tacking \acp{wicked-prob}, map the
+\acl{EA} arguments in detail, and create a multi-level threat model, broken down by the significantly different
+challenges posed by \acl{DAR} and \acl{DIM}. The thesis concludes with a set of policies and principles to apply to the
+problem in light of valid arguments and threats.

### 2020-10-25 Thesis-Chapter2-Background.tex
Termify argument maps and remove strategy notes

@@
 introduces cryptography, summarizes the history of encryption and the ``crypto wars,'' lists relevant encryption
-regulations, and lists prominent technical \ac{EA} proposals. It then introduces argument maps as a logical modeling
+regulations, and lists prominent technical \ac{EA} proposals. It then introduces \acp{argmap} as a logical modeling
 tool, and introduces threat modeling with data flow diagrams.
@@
 The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
-help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create argument maps.
-Argument maps are graphical representations of the logical structures and relationships between statements, premises,
-and conclusions. Argument maps are an outgrowth of the same research by Horst Rittel that originated the idea of
+help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create \acp{argmap}.
+\Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
+conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
 \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
@@
 \cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
-generates argument maps from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.
+generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.
 

### 2020-10-25 Thesis-Chapter3-PolicyMaking.tex
Termify argument maps and remove strategy notes
Introduce OODA loop

@@
 \subsection{Lessons}
+\label{sec-policy-lessons}
 
@@
 
-Rittel's follow-up research offers another tool: argument mapping.
+\section{Proposal: The OODA Loop for Wicked Problems}
 
-% - Avoid pitfalls of
-%   - Easy answers
-%     - EA mandates, complexity denial
-%   - Fallacies!
-%   - Bad results with long technical tails
+The above discussion describes the failure of well-established methods of policymaking at addressing \acp{wicked-prob}
+and properties a better method would have. Here, I propose a modified \acs{OODAloop} as an alternative policymaking
+model.
 
-% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
-% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
-% decrypted! So must be very careful to avoid solution attempts that leave long scars.
+The \ac{OODAloop} was a developed by Air Force Colonel John Boyd in the context of military strategy, but its ideas have
+penetrated many other sectors \cite{angerman_2004}. Boyd created it as a description of a successful strategy for
+countering opponents in real-time combat. There is an emphasis on speed and ``getting inside'' your opponent's loop. It
+is illustrated in \myfig{fig-ooda-loop}.
 
-\section{Proposal: The OODA Loop for Policymaking}
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/OODA-vanilla.png}
+    \caption[The OODA Loop]{The OODA Loop}
+    \label{fig-ooda-loop}
+\end{figure}
 
-\ac{OODAloop}.
+The \ac{OODAloop} serves as a good basis for contending with \acp{wicked-prob} because it has the correct fundamental
+structure. It maps very well to the \ac{classical-method} steps of problem definition (orientation), ideation and
+analysis (orientation), decision, and action, but it emphasizes the \ac{incrementalism}'s iteration and feedback.
+Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor holds up. This
+is the idea behind language like the ``war on drugs''---whether combating armed opponents or \acp{wicked-prob}, one is
+contending with an unpredictable opponent in an environment lacking complete information with an unlimited set of
+possible actions.
+
+The \ac{OODAloop} as is does not have all the properties listed in \mysec{sec-policy-lessons}, however.
+... %TODO: finish this
+See \myfig{fig-policy-ooda-loop}.
 
@@
     \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
-    \caption[The OODA Loop]{The OODA Loop}
-    \label{fig-ooda-loop}
+    \caption[The OODA Loop for Wicked Problems]{The OODA Loop for Wicked Problems}
+    \label{fig-policy-ooda-loop}
 \end{figure}
-
-% commission_tackling_2018:
-% - Strategies
-%   - Authoritative
-%   - Competitive
-%   - Collaborative
-% - Approaches
-%   - Linear problem solving
-%     - Ineffective! Several reasons, good breakdown. Dovetails with rozenshtein_wicked_2018.
-%   - Holistic problem solving: adaptive, innovative, flexible, cross-organizational
-% - Avoid
-%   - Narrow solutions
-
-% rozenshtein_wicked_2018:
-% - Approaches
-  % - Muddling through (he cites Charles E. Lindblom, just like feeley_judicial_2000!)
-  % - Risk management (--finally, something familiar to security experts!)
-  %   - So, a broader view of security, inclusive of both public safety and information security
-  %   - In this model, threat actors now include terrorists, CSAM distributors, organized crime
-  %     - Add a new character? Evil Alice and evil Bob (Alex and Becca?)? This automatically raises to question of how to
-  %         identify bad actors--AI (lee_detecting_2020)? PhotoDNA (microsoft_2020)? Government orders (warrants)? Could
-  %         be useful.
-  %   - I have already suggested adding government mandates to the traditional IS threat model
-  % - Imperfect solutions, like lawful hacking
-  % - Meanwhile, invest in knowledge production and improved relationships (like "cross-organizational" from
-  %     commission_tackling_2018)
-% - Avoid
-%   - Easy answers ("narrow solutions" from commission_tackling_2018)
-%     - Such as EA mandates or "golden age of surveillance"-based denials
-
-% matyas_incommensurability_2018:
-% - Approaches to reducing incommensurability
-%   - Reduce gray area between rules and principles (i.e. keep laws up to date with technology)
-  % - Use transparency and accountability (through technical means)
-
-% feeley_judicial_2000:
-% - Classical analytic method ("linear" in commission_tackling_2018)
-% % 1. Classical analytic method; each step has its own approaches, e.g. cost-benefit analysis (loosely based on natural
-% % sciences)
-% %   (missing: Observe)
-% %   a. Define the problem (Orient)
-% %   b. Identify a goal (Orient)
-% %   c. Generate a range of alternatives for achieving that goal (Orient)
-% %   d. Select the alternative that seems most promising (Decide)
-% %   e. Implement the selected alternative (Act)
-% %   (missing: loop)
-% - Incremental intuitive decision making ("Muddling through" in rozenshtein_wicked_2018)
-% - The hermeneutic circle
-
-% How to tackle wicked problems in general? Take some tips from commission_tackling_2018, rozenshtein_wicked_2018,
-% feeley_judicial_2000, and matyas_incommensurability_2018:
-% Correctly state the problem--identify the root causes. They aren't going away, there by definition is no clear long
-% term solution. Put the proposed solutions in context--none are perfect, none are permanent. For EA in particular,
-% consider balance of principles and rules.

### 2020-10-27 Thesis-Chapter3-PolicyMaking.tex
Refine OODA proposal

@@
 
-The \ac{OODAloop} was a developed by Air Force Colonel John Boyd in the context of military strategy, but its ideas have
-penetrated many other sectors \cite{angerman_2004}. Boyd created it as a description of a successful strategy for
-countering opponents in real-time combat. There is an emphasis on speed and ``getting inside'' your opponent's loop. It
-is illustrated in \myfig{fig-ooda-loop}.
+The \ac{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
+countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
+its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
+incident response \cite{schneier_future_2014}. There is an emphasis on speed and ``getting inside'' your opponent's
+loop. It is illustrated in \myfig{fig-ooda-loop}.
 
@@
 analysis (orientation), decision, and action, but it emphasizes the \ac{incrementalism}'s iteration and feedback.
-Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor holds up. This
-is the idea behind language like the ``war on drugs''---whether combating armed opponents or \acp{wicked-prob}, one is
-contending with an unpredictable opponent in an environment lacking complete information with an unlimited set of
-possible actions.
+Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor of facing an
+unpredictable opponent in a dynamic environment where every action counts holds up.
 
-The \ac{OODAloop} as is does not have all the properties listed in \mysec{sec-policy-lessons}, however.
-... %TODO: finish this
-See \myfig{fig-policy-ooda-loop}.
+The one modification the \ac{OODAloop} needs is unification of the problem definition and analysis steps. This is
+Rittel's ``argumentative process in the course of which an image of the problem and of the solution emerges gradually''
+\cite{rittel_dilemmas_1973}. \myfig{fig-policy-ooda-loop} illustrates the \ac{OODAloop} modified to serve as a
+policymaking strategy.
 
@@
 \end{figure}
+
+% Note on using argument maps to facilitate this argumentative process
+
+% Summarize chapter: wicked problems, why normal methods fail, how this strategy works
+
+% Preview it in action in the following chapters

### 2020-10-28 Thesis-Chapter2-Background.tex
Finish Chapter 3 draft

@@
 high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
-policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the
-application of blockchain \cite{phan_key_2017}. These approaches are not as thoroughly explored as trusted-party
-\ac{key-escrow}.
+policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
+\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
+thoroughly explored as trusted-party \ac{key-escrow}.
 

### 2020-10-28 Thesis-Chapter3-PolicyMaking.tex
Finish Chapter 3 draft

@@
 its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
-incident response \cite{schneier_future_2014}. There is an emphasis on speed and ``getting inside'' your opponent's
-loop. It is illustrated in \myfig{fig-ooda-loop}.
+incident response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your
+opponent's loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.
 
@@
 analysis (orientation), decision, and action, but it emphasizes the \ac{incrementalism}'s iteration and feedback.
-Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor of facing an
-unpredictable opponent in a dynamic environment where every action counts holds up.
+Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor holds up. One is
+facing an unpredictable opponent in a dynamic environment where every action counts. We cannot overpower
+\acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.
 
-The one modification the \ac{OODAloop} needs is unification of the problem definition and analysis steps. This is
-Rittel's ``argumentative process in the course of which an image of the problem and of the solution emerges gradually''
-\cite{rittel_dilemmas_1973}. \myfig{fig-policy-ooda-loop} illustrates the \ac{OODAloop} modified to serve as a
-policymaking strategy.
+The one modification the \ac{OODAloop} needs is unification of the problem definition and analysis steps, as one depends
+on the other. This is Rittel's ``argumentative process in the course of which an image of the problem and of the
+solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in \myfig{fig-policy-ooda-loop} as a cycle of
+``collaborative debate'' between observation and orientation. The modified \ac{OODAloop} incorporates each of the
+lessons from \mysec{sec-policy-lessons}.
 
@@
 
+Recall from \mysec{sec-arg-maps-intro} that \acp{argmap} resulted from the same research that produced the concept of
+\acp{wicked-prob}. Though formal research on the technique is thin, \acp{argmap} have proven helpful in simplifying
+complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate by helpfully slowing down
+discussion, depersonalizing conflict, and lending structure and rhythm \cite{dutoit_hypermedia_2006}. This makes them
+well suited as a tool for conducting the collaborative debate that helps problem solvers with disparate views narrow in
+on a problem definition and concrete proposals on which they can agree.
+
 % Summarize chapter: wicked problems, why normal methods fail, how this strategy works
 
-% Preview it in action in the following chapters
+\section{Summary}
+
+Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
+thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
+steps in the right direction. Both approaches have their strengths, but neither is suited to the complex and
+controversial nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation
+of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
+solutions, and crystallizes discussion around specific proposals. \Acp{argmap} are one tool that may facilitate this
+process. The result from a single iteration is an imperfect step forward based on implementing a particular proposal.
+The result from many iterations is the breakdown of the problem into sub-problems---some tame and inevitably some still
+wicked. The tame sub-problems can be addressed with the \ac{classical-method}, and the wicked sub-problems, by now
+restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.
+
+It is important to note that the same \acp{wicked-prob} is never solved twice, since each one is by definition, unique.
+However, changes in technology, culture, and current affairs may render the current strategy insufficient and require
+the process to start again. \Acp{wicked-prob} are like history: they don't repeat themselves, but they rhyme. Due to
+changing circumstances, solutions past will not work in the future, but lessons learned from the past provide useful
+experience.
+
+In the context of \ac{encryption} and \ac{EA}, this means that the outcome of \ac{the-cw1} doesn't hold precedential
+power over the debate in the second. It means that re-raising questions about the role of technology is playing society
+is legitimate. To follow this proposed strategy would mean that hardline rhetoric doesn't help \cite{ruiz_there_2018}
+\cite{geller_2019}, though sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
+helpful are collaborative efforts involving government, technical, and civil liberties representatives
+\cite{committee_decrypting_2018} \cite{group_2019}. Specific proposals around which groups can center discussion,
+whether they specify a particular form of \ac{EA} or perhaps especially if they don't, are also a necessary component
+\cite{kerr_encryption_2017} \cite{wright_crypto_2018} \cite{phan_key_2017}.
+
+In the remainder of this thesis, I will apply this strategy through the use of \acp{argmap} to document elements of the
+debate in \mychap{chap-arguments} and through creation of a threat model against which I analyze one specific proposal
+in \mychap{chap-threatmodel}.

### 2020-10-28 Thesis-Glossary.tex
Termify argument maps and remove strategy notes
Finish Chapter 3 draft

@@
 }
+
+\newterm{argmap}{argument map}{
+    A graphical representation of the logical structures and relationships between statements, premises, and
+    conclusions, used to disentangle complex arguments used in wicked problems
+}
 % End terms
@@
 \newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
-\newabbreviation{OODAloop}{OODA loop}{\ii{Observe-Orient-Decide-Act} loop}
+\newabbreviation{OODAloop}{OODA Loop}{\ii{Observe-Orient-Decide-Act} Loop}
 

### 2020-10-29 arguments/argdown.config.js
Begin arguments chapter

@@ -1,63 +1,89 @@
 // See https://argdown.org/guide/configuration-cheatsheet.html
 
 module.exports = {
-  author: 'Kevin Kredit',
-  inputPath: './build/w_citations/comprehensive.ad',
+  author: "Kevin Kredit",
+  inputPath: "./build/w_citations/comprehensive.ad",
   model: {
-    mode: 'strict',
-    removeTagsFromText: true
+    mode: "strict",
+    removeTagsFromText: true,
   },
   svg: {
-    outputDir: './build'
+    outputDir: "./build",
   },
   processes: {
     file: {
       inputPath: process.argv[process.argv.length - 1],
-      process: 'export-svg'
+      process: "export-svg",
     },
     whole: {
-      process: 'export-svg',
-      outputSuffix: ".whole"
+      process: "export-svg",
+      outputSuffix: ".whole",
+    },
+    contrib: {
+      process: "export-svg",
+      outputSuffix: ".contrib",
+      dot: {
+        graphVizSettings: {
+          // rankdir: "LR"
+        },
+      },
+      group: {
+        groupDepth: 3,
+      },
+      selection: {
+        selectedSections: [
+          "Central Issues",
+          "Response Measures",
+          "Arguments for Measures",
+          "Exceptional Access",
+        ],
+        // selectedTags: ["core", "ea-itself"],
+        // selectElementsWithoutTag: true,
+      },
     },
     ea_proposals: {
-      process: 'export-svg',
+      process: "export-svg",
       outputSuffix: ".ea_proposals",
       selection: {
-        selectedTags: ['ea-proposal', 'neutral-ea'],
-        selectElementsWithoutTag: true
-      }
+        selectedTags: ["ea-proposal", "neutral-ea"],
+        selectElementsWithoutTag: true,
+      },
     },
     pro: {
-      process: 'export-svg',
+      process: "export-svg",
       outputSuffix: ".pro",
       selection: {
-        selectedTags: ['ea-itself', 'pro-ea'],
+        selectedTags: ["ea-itself", "pro-ea"],
         selectElementsWithoutTag: true,
-      }
+      },
     },
     anti: {
-      process: 'export-svg',
+      process: "export-svg",
       outputSuffix: ".anti",
       selection: {
-        selectedTags: ['ea-itself', 'anti-ea'],
+        selectedTags: ["ea-itself", "anti-ea"],
         selectElementsWithoutTag: true,
-      }
+      },
     },
     fallacies: {
       dot: {
[…]

### 2020-10-31 Thesis-Chapter2-Background.tex
Write contributing factors map section

@@
 
-\paragraph*{\Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
+\paragraph*{The \Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
 describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
@@
 sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
-\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography
+\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography (\ac{CSAM})
 \cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
-terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on child
-pornography is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
+terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on
+\ac{CSAM} is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
 York Times} \cite{keller_internet_2019}.

### 2020-10-31 Thesis-Chapter4-Arguments.tex
Termify argument maps and remove strategy notes
Begin arguments chapter
Write contributing factors map section

@@
 
-% Already covered defining the problem. Next will discuss goals and list alternatives. That will feed into threat
-% modeling
+In this chapter we will analyze the \ac{encryption} debate using \acp{argmap}. The maps help one digest the debate, but
+there are two important shortcomings of the ones presented here. First, argument mapping is ideally just one tool in the
+collaborative process that progressively approaches the problem definition and solution. One can make an approximation
+from research, but it is not a replacement for debate arising from live discussion. Second, \acp{argmap} as portrayed in
+this thesis can be deceiving about the strength of an argument. Argument nodes show no indication of their strength or
+validity; one can formulate ten poor arguments against a single strong one, and a glance at the map will indicate that
+the case of the ten is stronger. With those qualifications in mind, let us begin by looking at the factors at the heart
+of the conflict.
 
-\section{Mapping the EA Debate}
+\section{Contributing Factors}
 
-\subsection{Desireable Outcomes}
+As previously stated, it is \ac{encryption}'s dual role as enabler of radical privacy and cornerstone of security is at
+the heart of the EA debate. Of course, there is only a conflict if one sees radical privacy as a bad thing. The right to
+privacy is both a strongly held value and an enshrined legal principle, and many use privacy-based arguments to show
+\ac{EA} as socially undesirable. The \ac{snowden} revelations \cite{landau_making_2013} only unveiled the scale of
+privacy-eroding U.S. government surveillance enabled by technological changes, terrorism-motivated policies, and weak
+oversight \cite{shamsi_2011}. Violations of privacy are concerning for those most vulnerable and for society at large.
+For example, Saudi journalist Jamal Khashoggi's private communications were surveilled by the Saudi government before he
+became a victim of politically motivated murder \cite{liebermann_2019}. At a societal level, presence of surveillance
+alone changes behavior and chills free speech \cite{rogaway_moral_2015}, while government violations of the law degrade
+institutional trust.
+
+The government, and particularly \acl{LE}, argues that encryption is handcuffing their investigational capacity. The
+\ac{FBI} has branded strong encryption as ``warrant-proof'' and states that ``the government often cannot obtain the
+electronic evidence necessary to investigate and prosecute threats to public and national safety'' \cite{fbi_2020}. The
+pro-\ac{EA} argument is typically rooted in public safety: investigators need access to evidence to fulfill their
+commission to catch and convict lawbreakers, and encryption is making that too difficult. Five-eyes nations regularly
+put out joint statements expressing their frustration with encryption \cite{ministerial_2018} \cite{goodale_2017}, and
+were joined in 2020 by India and Japan in a statement emphasizing the growing danger of \ac{CSAM} \cite{intl_2020}.
+Child pornography is one the original \ac{horsemen} that some argue are used to scare people into supporting backdoors
+\cite{schneier_scaring_2019}, but it is also a problem growing to astonishing proportions. The \textit{New York Times}'s
+2019 investigation on the matter quoted one law enforcement officer as estimating that 400,000 New Jerseyans, more than
+4\% of the state's population, could be charged with violating child exploitation material laws
+\cite{keller_internet_2019}.
+
+Unfortunately, shocking assertions such as this have not been independently confirmed or backed up with hard data.
+Intelligence agencies have a long history or misrepresenting positions and overstepping bounds
+\cite{johnson_congressional_2004} \cite{shamsi_2011}, and this has led to lack of institutional trust undercutting their
+claims that they need \ac{EA}. Indeed, the \ac{FBI} has already been caught exaggerating by several times how many
+mobile devices they could not access due to \acl{DE} \cite{devlin_2018}. As Rozenshtein says on this subject, ``It is
+impossible to know the precise extent to which encryption frustrates law-enforcement investigations, both because
+law-enforcement agencies are only beginning to collect accurate statistics, and because one can never be sure of how an
+investigation would have proceeded in the absence of encryption'' \cite{rozenshtein_wicked_2018}, but it is still
+crucial to have an accurate depiction of the problem in order to come to justified and helpful solutions. This argument
+is analyzed further in the next section.
+
+Information security is a the remaining central issue in the debate. As described in \mysec{sec-crypto-basics},
+\ac{cryptography} in various forms plays a foundational technical role in nearly every aspect of security, and
+information confidentiality itself is one of the primary properties security practitioners fight to protect. Efforts in
+the past have sought to undermine the cryptographic foundations of encryption, though this is not a feature of proposals
+today, as that is universally seen as too risky. \mysec{sec-tech-approaches} introduced several alternative technical
+approaches to \ac{EA}, but experts still proclaim loud and clear that the current state of the art is far from capable
+of providing the type of access \acl{LE} in an acceptably secure manner \cite{abelson_2015} \cite{abelson_risks_1997}.
+Security was an afterthought in early computing and networking designs, and the field of cybersecurity is still more of
+an art the a science; hobbling systems with a mandate to provide \ac{EA} would inevitably diminish security to some
+degree \cite{abelson_2015}.
+
+Finally, one may argue that \ac{EA}, depending on its implementation, could have positive auxiliary use cases. It could
+enable malware scanning, administrator access in a business setting, or password recovery. Typically these use cases
+have other solutions, however, and to combine the requirements of these uses with those of any reasonable \ac{EA}
+proposal would likely fatally weaken it.
+
+See \myfig{fig-args-contrib} for the \ac{argmap} representing this section.
+
+\begin{sidewaysfigure}
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
+  \caption[Contributing Factors to the EA Debate]{Contributing Factors to the EA Debate}
+  \label{fig-args-contrib}
+\end{sidewaysfigure}
+
+\section{Going Dark vs. The Golden Age}
+
+% Actually fund initiatives to tackle these problems!
+% pfefferkorn_2019
+% keller_internet_2019
+
+\section{Eliminating Fallacious Arguments}
+
+\section{EA and Alternatives}
+
+\section{Zooming in on EA}
+
+% - Avoid pitfalls of
+%   - Easy answers
+%     - EA mandates, complexity denial
+%   - Fallacies!
+%   - Bad results with long technical tails
+
+% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
+% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
+% decrypted! So must be very careful to avoid solution attempts that leave long scars.
 
@@
 
-\subsection{Proposed Alternatives}
+\section{Specific Proposals}
 

### 2020-10-31 arguments/argdown.config.js

@@ -22,11 +22,6 @@ module.exports = {
     contrib: {
       process: "export-svg",
       outputSuffix: ".contrib",
-      dot: {
-        graphVizSettings: {
-          // rankdir: "LR"
-        },
-      },
       group: {
         groupDepth: 3,
       },
@@ -37,8 +32,6 @@ module.exports = {
           "Arguments for Measures",
           "Exceptional Access",
         ],
-        // selectedTags: ["core", "ea-itself"],
-        // selectElementsWithoutTag: true,
       },
     },
     ea_proposals: {

### 2020-10-31 scripts/show-source.sh

@@ -11,12 +11,12 @@ if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
         readlink -f $FILE
     done
     exit 0
-elif [[ 1 == $# && -f $ORIG_DIR/$1 ]]; then
+elif [[ 1 == $# && -f $ORIG_DIR/${1%%:[0-9]*} ]]; then
     # Show a file based on relative path
-    FILE=$ORIG_DIR/$1
-elif [[ 1 == $# && -f $1 ]]; then
+    FILE=$ORIG_DIR/${1%%:[0-9]*}
+elif [[ 1 == $# && -f ${1%%:[0-9]*} ]]; then
     # Show a file based on absolute path
-    FILE=$1
+    FILE=${1%%:[0-9]*}
 elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
     # Show a file based on bibtex ref
     FILE="$(grep -riIsl "@*{$1," . | head -1)"

### 2020-11-02 Thesis-Chapter4-Arguments.tex
Create going dark and golden age arg maps

@@
 \section{Going Dark vs. The Golden Age}
+
+A significant portion of the disconnect between government and technical community representatives is disagreement over
+whether law enforcement has \ii{too much} or \ii{too little} access to data.
+
+See the ``Going Dark'' argument in \myfig{fig-arg-going-dark}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goingdark.pdf}
+  \caption{A ``Going Dark'' Argument Map}
+  \label{fig-arg-going-dark}
+\end{figure}
+
+See the ``Golden Age of Surveillance'' argument in \myfig{fig-arg-golden-age}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
+  \caption{A ``Golden Age of Surveillance'' Argument Map}
+  \label{fig-arg-golden-age}
+\end{figure}
 

### 2020-11-02 arguments/argdown.config.js

@@ -34,10 +34,51 @@ module.exports = {
         ],
       },
     },
+    goingdark: {
+      process: "export-svg",
+      outputSuffix: ".goingdark",
+      group: {
+        groupDepth: 3,
+      },
+      selection: {
+        selectedSections: [
+          "Exceptional Access",
+          "Central Issues",
+          "Going Dark Conclusions",
+          "Going Dark Argument",
+          "Going Dark Non Core",
+        ],
+      },
+    },
+    goldenage: {
+      process: "export-svg",
+      outputSuffix: ".goldenage",
+      group: {
+        groupDepth: 3,
+      },
+      selection: {
+        selectedSections: [
+          "Exceptional Access",
+          "Central Issues",
+          "Going Dark Conclusions",
+          "Golden Age Argument",
+          "Golden Age Non Core",
+        ],
+      },
+      dot: {
+        argument: { lineWidth: 125 },
+        statement: { lineWidth: 125 },
+      },
+    },
     ea_proposals: {
       process: "export-svg",
       outputSuffix: ".ea_proposals",
       selection: {
+        selectedSections: [
+          "Exceptional Access",
+          "Response Measures",
+          "Arguments for Measures",
+        ],
         selectedTags: ["ea-proposal", "neutral-ea"],
         selectElementsWithoutTag: true,
       },

### 2020-11-03 arguments/argdown.config.js
Add 'measures' map

@@ -29,7 +29,7 @@ module.exports = {
         selectedSections: [
           "Central Issues",
           "Response Measures",
-          "Arguments for Measures",
+          "Arguments for Exceptional Access",
           "Exceptional Access",
         ],
       },
@@ -70,35 +70,35 @@ module.exports = {
         statement: { lineWidth: 125 },
       },
     },
-    ea_proposals: {
-      process: "export-svg",
-      outputSuffix: ".ea_proposals",
-      selection: {
-        selectedSections: [
-          "Exceptional Access",
-          "Response Measures",
-          "Arguments for Measures",
-        ],
-        selectedTags: ["ea-proposal", "neutral-ea"],
-        selectElementsWithoutTag: true,
-      },
-    },
-    pro: {
-      process: "export-svg",
-      outputSuffix: ".pro",
-      selection: {
-        selectedTags: ["ea-itself", "pro-ea"],
-        selectElementsWithoutTag: true,
-      },
-    },
-    anti: {
-      process: "export-svg",
-      outputSuffix: ".anti",
-      selection: {
-        selectedTags: ["ea-itself", "anti-ea"],
-        selectElementsWithoutTag: true,
-      },
-    },
+    // ea_proposals: {
+    //   process: "export-svg",
+    //   outputSuffix: ".ea_proposals",
+    //   selection: {
+    //     selectedSections: [
+    //       "Exceptional Access",
+    //       "Response Measures",
+    //       "Arguments for Measures",
+    //     ],
+    //     selectedTags: ["ea-proposal", "neutral-ea"],
+    //     selectElementsWithoutTag: true,
+    //   },
+    // },
+    // pro: {
+    //   process: "export-svg",
+    //   outputSuffix: ".pro",
+    //   selection: {
+    //     selectedTags: ["ea-itself", "pro-ea"],
+    //     selectElementsWithoutTag: true,
+    //   },
+    // },
+    // anti: {
+    //   process: "export-svg",
+    //   outputSuffix: ".anti",
+    //   selection: {
+    //     selectedTags: ["ea-itself", "anti-ea"],
+    //     selectElementsWithoutTag: true,
+    //   },
+    // },
     fallacies: {
       dot: {
         graphVizSettings: {
@@ -117,6 +117,22 @@ module.exports = {
         ],
         selectedTags: ["core", "ea-itself", "fallacy", "fallacious"],
         selectElementsWithoutTag: true,
+      },
+    },
+    measures: {
+      process: "export-svg",
[…]

### 2020-11-03 arguments/argdown.config.js
Add 'classes' and 'proposals' maps

@@ -70,35 +70,6 @@ module.exports = {
         statement: { lineWidth: 125 },
       },
     },
-    // ea_proposals: {
-    //   process: "export-svg",
-    //   outputSuffix: ".ea_proposals",
-    //   selection: {
-    //     selectedSections: [
-    //       "Exceptional Access",
-    //       "Response Measures",
-    //       "Arguments for Measures",
-    //     ],
-    //     selectedTags: ["ea-proposal", "neutral-ea"],
-    //     selectElementsWithoutTag: true,
-    //   },
-    // },
-    // pro: {
-    //   process: "export-svg",
-    //   outputSuffix: ".pro",
-    //   selection: {
-    //     selectedTags: ["ea-itself", "pro-ea"],
-    //     selectElementsWithoutTag: true,
-    //   },
-    // },
-    // anti: {
-    //   process: "export-svg",
-    //   outputSuffix: ".anti",
-    //   selection: {
-    //     selectedTags: ["ea-itself", "anti-ea"],
-    //     selectElementsWithoutTag: true,
-    //   },
-    // },
     fallacies: {
       dot: {
         graphVizSettings: {
@@ -132,6 +103,38 @@ module.exports = {
           "Exceptional Access",
           "Legal Action",
           "Arguments for Measures",
+        ],
+      },
+    },
+    classes: {
+      process: "export-svg",
+      outputSuffix: ".classes",
+      group: {
+        groupDepth: 3,
+      },
+      selection: {
+        selectedSections: [
+          "Central Issues",
+          "Exceptional Access",
+          "DAR EA Classes",
+          "DIM EA Classes",
+          "Arguments for EA Types",
+        ],
+      },
+    },
+    proposals: {
+      process: "export-svg",
+      outputSuffix: ".proposals",
+      group: {
+        groupDepth: 3,
+      },
+      selection: {
+        selectedSections: [
+          "Central Issues",
+          "Exceptional Access",
+          "DAR EA Proposals",
+          "DIM EA Proposals",
+          "Arguments for EA Proposals",
         ],
       },
     },

### 2020-11-05 Thesis-Chapter1-Introduction.tex
Write going dark, golden age, and fallacies sections

@@
 The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
-access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{fbi_2020}. Human rights
+access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}. Human rights
 groups and technical leaders counter that a weakened encryption environment would fatally compromise privacy and

### 2020-11-05 Thesis-Chapter3-PolicyMaking.tex
Write going dark, golden age, and fallacies sections

@@
 disagreement on goals (or even basic premises). Sides can't agree whether things are ``going dark'' or we're in the
-``golden age of surveillance'' or how much to value competing notions of security. Second, that information is
+``golden age for surveillance'' or how much to value competing notions of security. Second, that information is
 ``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations, and though the

### 2020-11-05 Thesis-Chapter4-Arguments.tex
Write going dark, golden age, and fallacies sections

--- 2020-10-31 Thesis-Chapter4-Arguments.tex
@@
 \section{Contributing Factors}
+\label{sed-arg-contrib}
 
-As previously stated, it is \ac{encryption}'s dual role as enabler of radical privacy and cornerstone of security is at
-the heart of the EA debate. Of course, there is only a conflict if one sees radical privacy as a bad thing. The right to
-privacy is both a strongly held value and an enshrined legal principle, and many use privacy-based arguments to show
-\ac{EA} as socially undesirable. The \ac{snowden} revelations \cite{landau_making_2013} only unveiled the scale of
-privacy-eroding U.S. government surveillance enabled by technological changes, terrorism-motivated policies, and weak
-oversight \cite{shamsi_2011}. Violations of privacy are concerning for those most vulnerable and for society at large.
-For example, Saudi journalist Jamal Khashoggi's private communications were surveilled by the Saudi government before he
-became a victim of politically motivated murder \cite{liebermann_2019}. At a societal level, presence of surveillance
-alone changes behavior and chills free speech \cite{rogaway_moral_2015}, while government violations of the law degrade
-institutional trust.
+\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, it is \ac{encryption}'s dual
+role as enabler of radical privacy and cornerstone of security is at the heart of the debate. Of course, there is only a
+conflict if one sees radical privacy as a bad thing. The right to privacy is both a strongly held value and an enshrined
+legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The \ac{snowden}
+revelations \cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
+technological changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Violations of privacy are
+concerning for those most vulnerable and for society at large. For example, Saudi journalist Jamal Khashoggi's private
+communications were surveilled by the Saudi government before he became a victim of politically motivated murder
+\cite{liebermann_2019}. At a societal level, presence of surveillance alone changes behavior and chills free speech
+\cite{rogaway_moral_2015}, while government violations of the law degrade institutional trust.
+
+\begin{sidewaysfigure}
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
+  \caption[Contributing Factors to the EA Debate]{Contributing Factors to the EA Debate}
+  \label{fig-args-contrib}
+\end{sidewaysfigure}
 
@@
 
-See \myfig{fig-args-contrib} for the \ac{argmap} representing this section.
-
-\begin{sidewaysfigure}
-  \centering\CaptionFontSize
-  \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
-  \caption[Contributing Factors to the EA Debate]{Contributing Factors to the EA Debate}
-  \label{fig-args-contrib}
-\end{sidewaysfigure}
-
 \section{Going Dark vs. The Golden Age}
 
-% Actually fund initiatives to tackle these problems!
-% pfefferkorn_2019
-% keller_internet_2019
+A significant portion of the disconnect between government and technical community representatives is disagreement over
+whether law enforcement has \ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is
+well-represented by proponent former \ac{FBI} Director James Comey in a 2014 speech \cite{comey_2014} (though he did not
+coin the phrase \cite{swire_encryption_2011}). The argument is represented in \myfig{fig-arg-going-dark}, also informed
+by critiques \cite{rogaway_moral_2015}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goingdark.pdf}
+  \caption{A ``Going Dark'' Argument Map}
+  \label{fig-arg-going-dark}
+\end{figure}
+
+The ``Going Dark'' argument goes that privacy and safety are both desirable goods, but they inherently conflict, and we
+need balance. Individuals must give up some of their personal good of privacy to make room for the public good of
+safety. Encryption tips the scales too far towards privacy, upsetting the balance of the status quo. \ac{EA} is needed
+to restore the balance and enable law enforcement agencies to do their jobs and protect the public.
+
+The argument has a few blind spots. Depicting privacy and safety as competing in a zero-sum tradeoff ignores all the
+ways in which privacy directly enhances safety. Data privacy makes certain crimes more difficult, such as stalking or
+identity theft. Privacy is also a public good to the extent that it precludes mass surveillance, an argument explored
+more deeply in the ``Golden Age'' side of the argument.
+
+Most importantly, the ``Going Dark'' argument hinges on encryption being responsible for failure of law enforcement
+investigations. As described in \mysec{sed-arg-contrib}, there is a lack of data on the extent to which encryption is
+itself to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an issue. The
+highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended with the \ac{FBI} breaking into the phone but finding nothing
+of value. Tellingly, each of Comey's examples in his 2014 speech were not cases where encryption inhibited
+investigations, but where cases hypothetically \ii{could} have been inhibited if encryption was a factor
+\cite{comey_2014}.
+
+Citing lack of evidence only gets you so far, however. In most cases, it is impossible to know what would have happened
+in an investigation unimpeded by strong encryption. This is part of why it is a \ac{wicked-prob}. Indeed, when the
+encryption curtain is pulled back, there is often much to see. Unencrypted Facebook Messenger is has approximately 20\%
+market share in active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
+\cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted messaging service Encrochat
+exposed crime rings and led to a massive raids \cite{cox_2020}.
+
+Aside from encryption entirely, a 2018 study found that law enforcement's primary challenges with digital evidence were
+identifying service providers and getting the sought after data from them \cite{carter_2018}. This is in the context of
+current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging Fruit,'' and also lists
+resource limitations and training needs among top priorities. Indeed, many current initiatives are underfunded. Major
+anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels, the Justice Department
+task force has produced only two of five biennial reports, and a senior executive position within the Justice Department
+was never created \cite{keller_internet_2019}. Federal entities commissioned to improve communication between law
+enforcement and industry as well as to train law enforcement on managing digital evidence are similarly underfunded
+\cite{carter_2018}.
+
+% Golden age
+
+The argument for law enforcement having \ii{too much} access to data claims that instead of law enforcement ``Going
+Dark,'' they are living in the ``Golden Age for Surveillance'' \cite{swire_encryption_2011}. This argument goes that
+although encryption makes certain data completely unaccessible, technological change on the whole has given authorities
+much more digital evidence than encryption has taken away. Paired with the dangers of mass surveillance, encryption is
+a necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
+\myfig{fig-arg-golden-age}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
+  \caption{A ``Golden Age for Surveillance'' Argument Map}
+  \label{fig-arg-golden-age}
+\end{figure}
+
+The ``Golden Age'' argument revolves around surveillance, its undesirability, and technology's role in it. Cryptographer
+Phillip Rogaway writes at length about the negative impacts of mass surveillance in his essay ``The Moral Character of
+Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway argues that ``pervasive collection \ii{itself} chills
+free-speech and threatens liberal democracy'' (emphasis his) through the way its mere existence affects the subject. He
+argues that it is fundamentally a tool of power, and though I would add that power from the people created through
+representative political process is valid power, clear abuses have undercut reasonable confidence in these institutions.
+
+The next element is that technology has increased the government's surveillance powers. This can hardly be contested.
+Going back to Comey's ``Going Dark'' examples once more, each case he cited was cracked by digital evidence that would
+not have been available even twenty years ago \cite{comey_2014}. In response to the danger of mass surveillance and the
+scale of modern surveillance capabilities, encryption is one of the only privacy-saving tools. (Opting out may be
+possible for some, but is not a realistic option for society as a whole.)
+
+While much stronger than the ``Going Dark'' argument, the ``Golden Age for Surveillance'' argument suffers from one
+major flaw. It focuses strongly on the abilities of and dangers posed by agencies like the \ac{NSA} and ignores the
+problems of local, state, and sometimes-surprisingly-cash-strapped federal law enforcement. The boom in the existence of
+digital evidence is clear, but as shown, law enforcement already has trouble acquiring the evidence with the help of
+tech and communications companies, and does not have the funding or capability to get it on its own \cite{carter_2018}.
+Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard evidence in cases
+where it may be suspected to exist \cite{shelton_study_2006}.
+
+Still, the fact that the tools of mass surveillance are out of reach for most law enforcement must not be relied upon.
+That is, most likely, a temporary situation. A safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA}
+without mass surveillance is conceivable. \ac{EA} implementations that enable law enforcement but thwart mass
+surveillance by remote powers subvert most of the ``Golden Age'' argument. Means for \ac{EA} to do so are explored in
+future argument maps.
 
@@
 
+Analyzing debates in the \ac{argmap} format invites one to seriously consider the merits of opposing arguments. Part of
+that consideration includes identifying and eliminating fallacious arguments. In heated, partisan debates, fallacious
+arguments spread easily, and the \ac{encryption} debate has its share, shown in \myfig{fig-arg-fallacies}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.fallacies.pdf}
+  \caption{Fallacious Arguments of the EA Debate}
+  \label{fig-arg-fallacies}
+\end{figure}
+
+Let us begin with bad arguments used to attack \ac{EA}. The first is the most common: ``\ac{EA} is a bad idea because
+backdoors are insecure.'' It appears on the map as a straw man argument due to the way the term ``backdoor'' is used.
+``Backdoor'' is essentially technical shorthand for ``insecure hack that should never be used in production''; when it
+is used as a basket term for all \ac{EA} proposals, the audience is primed to think them all stupid and helplessly
+dangerous. Thus, the arguer is presenting their opponent's request for some form of \ac{EA} in the worst possible light.
+
+Another common argument is that implementing \ac{EA}---or even researching it---moves us inexorably down a slippery
+slope, where government demands will never be satisfied. This argument can be used even if one thinks judicious
+hypothetical use of \ac{EA} tools would be a good thing. Going too far and too fast is certainly a danger, but it is
+both cynical and fallacious to declare the permanent destruction of privacy as the inevitable end of \ac{EA} research.
+
+It is highly tempting to declare government calls for \ac{EA} disingenuous due to hypocritical behavior in their
+treatment of the issues. For example, if combatting \ac{CSAM} is a government priority, how come Congress hasn't funded
+it and the Justice Department hasn't prioritized it \cite{keller_internet_2019}? This question, and others like it,
+demands an answer, but it ultimately does not impact the validity (or invalidity) of pro-\ac{EA} arguments. The absence
+of government funding doesn't mean technical anti-\ac{CSAM} measures are unnecessary.
+
+The final two anti-\ac{EA} fallacies are examples of false dichotomies, presenting the situation as if there are only
+two choices. \Acp{wicked-prob} have an unlimited set of possible responses, so reducing it to two (and often, presenting
+one as a straw man) is misleading. One false dichotomy is that the only acceptable solutions are perfect \ac{EA} or no
+\ac{EA}. This leaves no room for risk-based approaches, the typical security strategy outside the realm of cryptography.
+Another false dichotomy is that regulators must either allow strong cryptography or we will live in a mass surveillance
+dystopia. This one carries hints of ``slippery slope'' and appeal to emotion as well, framed as a choice between only
+two futures. Framing the situation in all or nothing terms harms the chances for successful collaborative debate.
+
+Fallacious arguments used to support \ac{EA} rely heavily on appeal to emotion. Politicians and law enforcement agents
+commonly appeal to pity for child abuse victims or fear of terrorists, drug dealers, and kidnappers to make their point.
+Public safety is a core value in the debate and the technical community needs to and does take these concerns seriously
+\cite{schneier_scaring_2019}. These appeals become fallacious when they are used to manipulate the audience into
+supporting measures incommensurate with the problems, or to manipulate the audience into thinking that tech somehow
+actually does not care about these issues.
+
+Another argument used to subvert anti-EA arguments from experts goes along the lines of, since we put a man on the moon,
+we can surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The problem of landing an
+astronaut on the moon is different from the problem of encryption and \ac{EA} in every way except that it was hard. The
+Apollo program solved mostly tame problems, and it had enormous government backing. Resolving the encryption debate has
+neither of these factors. Many similar technological-progress based arguments suffer the same faults.
+
+Lastly, government officials often pose \ac{EA} as a problem that security experts simply have not researched enough. If
+they just ``nerd harder,'' they can find a solution \cite{schneier_2019}. This misses the point of research done up to
+this point. Encryption and \ac{EA} is a \ac{wicked-prob} for which tame, technical-only solutions will not work
+\cite{rozenshtein_wicked_2018}. Even on the technical side, experts have stated time and again that depending how you
+look at it, the technical problem is solved---\ac{EA} is easy from a purely cryptographic point of view. The hard part
+is the security of the interfaces with the human users and administers of the system \cite{abelson_risks_1997}
+\cite{abelson_2015}. This may be an area of anemic academic research, but to frame that as the primary roadblock to
+progress utterly misses the point.
+
 \section{EA and Alternatives}
 
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.measures.pdf}
+  \caption{EA and its Alternatives}
+  \label{fig-arg-measures}
+\end{figure}
+
 \section{Zooming in on EA}
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.classes.pdf}
+  \caption{Classes of EA}
+  \label{fig-arg-classes}
+\end{figure}
 
@@
 \section{Specific Proposals}
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.proposals.pdf}
+  \caption{EA Proposals}
+  \label{fig-arg-proposals}
+\end{figure}
 

### 2020-11-05 scripts/sourceme.sh

@@ -24,9 +24,14 @@ alias gen="$SCRIPTS_DIR/bib-gen.sh"
 alias cite="$SCRIPTS_DIR/count-citations.sh"
 alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
 alias fixme="(cd $BASE_DIR && grep -nE 'FIXME' *.tex)"
-alias pdf="xdg-open build/Thesis.pdf"
-alias gvsu="xdg-open build/Thesis-GVSU.pdf"
+alias pdf="xdg-open $BASE_DIR/build/Thesis.pdf"
+alias gvsu="xdg-open $BASE_DIR/build/Thesis-GVSU.pdf"
 alias start="(codium .) && (xdg-open build/Thesis.pdf) && make live"
+
+function arg {
+    xdg-open $BASE_DIR/arguments/build/$1.pdf 2>/dev/null \
+        || xdg-open $BASE_DIR/arguments/build/comprehensive.$1.pdf 2>/dev/null
+}
 
 function notes {
     pushd $BASE_DIR/reading >/dev/null

### 2020-11-05 Thesis-Chapter4-Arguments.tex
Clean up arguments chapter notes
Write EA and alternatives section

@@
 
-\begin{figure}[h]
+\begin{figure}[t!]
   \centering\CaptionFontSize
@@
 
-\begin{figure}[h]
+\begin{figure}[p!]
   \centering\CaptionFontSize
-  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
+  \includegraphics[width=0.9\linewidth]{arguments/build/comprehensive.goldenage.pdf}
   \caption{A ``Golden Age for Surveillance'' Argument Map}
@@
 that consideration includes identifying and eliminating fallacious arguments. In heated, partisan debates, fallacious
-arguments spread easily, and the \ac{encryption} debate has its share, shown in \myfig{fig-arg-fallacies}.
+arguments spread easily. \myfig{fig-arg-fallacies} shows those used in the encryption and \ac{EA} debate.
 
-\begin{figure}[h]
+\begin{figure}[t!]
   \centering\CaptionFontSize
@@
 
-\begin{figure}[h]
+Moving on from problem framing and meritless arguments, we next turn to actions that can be taken in response to the
+situation. \myfig{fig-arg-measures} breaks the actions down into the current capabilities, legal measures, and \ac{EA}.
+Current capabilities can be implemented today, and include maintaining the status quo and increasing investment in
+current programs. Legal measures require some change in legal or political direction, but do not require fundamentally
+new technical research. These include compelling passwords, which requires legal clarification, and sanctioning lawful
+hacking, which requires a strategic pivot and an oversight framework. (These non-\ac{EA} approaches were first
+introduced in \mysec{sec-alt-approaches}.) At this level, \ac{EA} is analyzed in general. Specific EA approaches are
+addressed in the next map.
+
+\begin{sidewaysfigure}
   \centering\CaptionFontSize
@@
   \label{fig-arg-measures}
-\end{figure}
+\end{sidewaysfigure}
+
+Each measure is considered with reference to five metrics. These metrics represent a generalization of metrics from
+several sources, including two separate diverse committees gathered to analyze the encryption debate
+\cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
+of security, safety, privacy, and trust. The five metrics used to analyze response measures are minimization of
+increased security risk, utility to law enforcement, protection of civil liberties---particularly important as after the
+analysis of the ``Golden Age for Surveillance'' argument---and economic impact. Economic impact is not as much of a
+primary concern as the others, but it does help determine practical viability of a proposal.
+
+Beginning with current capabilities, the first choice is to do nothing. One may come to the conclusion that the current
+state of affairs actually does represent an acceptable balance of interests. If technology and government agree on this
+approach, then \ac{the-cw2} can end much like the first. However, legislative action is a risk to be mitigated in
+itself, and if the technical community ``decides'' on this approach and government continues to reject it, we could be
+forced in an undesirable direction.
+
+We could also decide to increase investment in current technologies and investigative techniques. As already shown,
+encryption is not the greatest barrier in law enforcement's use of digital evidence \cite{carter_2018}. Even when
+encryption is involved, traditional sleuthing can often achieve access. Security expert Bruce Schneier and cyberlaw
+expert Orin Kerr have compiled a list of ``encryption workarounds'' that do not rely on \ac{EA}, including finding the
+key, guessing the key, accessing plaintext while the device is in use, and finding another copy of the plaintext
+\cite{kerr_encryption_2017}. Though probabilistic and time consuming in comparison to \ac{EA}, none of these methods
+require new legal or technical capability. These methods are effective \cite{greenberg_2018}, and paired with investment
+in current digital evidence capabilities and anti-\ac{CSAM} (for example) task forces, they represent a safe response.
+
+Compelled password disclosure is a proposal to simply require defendants to hand over passwords and \acp{PIN} to their
+devices. The legality of this approach is working its way through the courts \cite{bittenbender_2019} \cite{sobel_2019},
+as it rubs against the Fifth Amendment right against self-incrimination. Even so, its strongest argument in favor is
+that it is cheap. Schneier and Kerr point out that this approach requires the password holder to be known and available
+to investigators, the prosecution has to overcome several legal hurdles even if the basic approach is ruled
+Constitutionally legal, and even then the defendant may choose to withhold the password and be held in contempt rather
+than be charged for crime evidenced by the plaintext \cite{kerr_encryption_2017}.
+
+\Ac{lawful-hacking} is the primary alternative argued for besides investment or \ac{EA}, though it is universally seen
+as a sub-optimal compromise \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018}
+\cite{kerr_encryption_2017} \cite{soesanto_2018}. Its advantages: it is effective against those for whom \ac{EA} would
+not necessarily work, and, well, it's a technical route that isn't \ac{EA} itself. \Ac{lawful-hacking} has the advantage
+over \ac{EA} at being effective against organized crime and terror groups because while they would inevitably use
+encryption tools to evade any \ac{EA} scheme, their operational security would fail at some point, letting law
+enforcement or intelligence agencies into their networks and devices. This approach is already in use to some extent
+\cite{cox_2020}, most famously to bring down cartel kingpin El Chapo \cite{feuer_chapo_2019}.
+
+The real change that this policy would take would be a sanctioning and formalization of this strategy. One downside is
+that it would be of little use to ill-resourced departments, nullifying its usefulness for more commonplace crimes. Its
+major downside is the altered relation is would create between law enforcement and tech companies. Instead of
+cooperation, they are suddenly thrust into competition, with law enforcement incentivized to hoard rather than disclose
+vulnerabilities it discovers. Prominent proposals suggest mandating prompt disclosure of discovered vulnerabilities
+\cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016}. Theoretically, the supply of vulnerabilities is very large,
+and new ones could be found at a rate that maintains law enforcement access at constant levels while improving security
+as a positive side effect. The government would source these vulnerabilities from a combination of the public domain,
+the commercial exploit market, and a central ``Vulnerability Lab'' \cite{bellovin_lawful_2013}. However, as the \$2
+million price on the exploit market for an Apple iOS jailbreak bug shows, security really is improving
+\cite{goodin_zeroday_2019}. Law enforcement would still find itself without access, and would need to develop costly
+exploits in the lab. Mandatory vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the
+\ac{FBI}'s historical abuses of power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has
+passed \cite{keller_internet_2019} or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine
+that the government would actually give up such expensive vulnerabilities so easily when they are the pillar of their
+access strategy.
+
+The final proposal is \ac{EA} itself. Its advantage, generally speaking, is that it would provide reliable access to law
+enforcement in a reliable manner. It could serve as a compliment to \ac{lawful-hacking}, blunting the incentive to hoard
+vulnerabilities because it would be unnecessary for \ac{EA} compliant systems, which would presumably include mainstream
+technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
+and terror groups. While impact on civil liberties and mass surveillance depends on implementation, any proposal would
+be costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security
+is the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
+risk. The next map analyzes this more closely.
 
@@
 
-% - Avoid pitfalls of
-%   - Easy answers
-%     - EA mandates, complexity denial
-%   - Fallacies!
-%   - Bad results with long technical tails
-
-% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
-% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
-% decrypted! So must be very careful to avoid solution attempts that leave long scars.
-
-% See committee_decrypting_2018!
-% For a wholistic threat model, include both the threats of government (legislation) and of criminal abusers of your
-% system.
+% What properties would good EA have?
+% easy for LEAs to use? hard? expensive? voluntary? unable to be used for mass surveillance? transparent (attempts are
+%   visible)? alignment of values with security (e.g. discourages bug hoarding)? adaptable (since solution won't be
+%   permanent, either technically or according to collective will as things change rozenshtein_wicked_2018)? suited to
+%   rules and principles (matyas_incommensurability_2018)?
 
@@
 
-% See committee_decrypting_2018!
-
-% Subject one: do we even want EA?
-%  --> no? exit here
-%  --> yes? carry on
-%  --> pragmatist? carry on
-% include fallacies argument map here
-% see rozenshtein_wicked_2018, benaloh_what_2018, soesanto_2018, ministerial_2018, fbi_2020, brantley_2018, lund_2020
-% ... group_2019, ruiz_there_2018, rozenshtein_2019
-
-% Subject two: what properties would good EA have?
-% easy for LEAs to use? hard? expensive? voluntary? unable to be used for mass surveillance? transparent (attempts are
-%   visible)? alignment of values with security (e.g. discourages bug hoarding)? adaptable (since solution won't be
-%   permanent, either technically or according to collective will as things change rozenshtein_wicked_2018)? suited to
-%   rules and principles (matyas_incommensurability_2018)?
-
-% Subject three: what options do we have?
-% map of properties onto proposals--leads into threat modeling, which will check the veracity of claims
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% In this chapter, develop the arguments employed in EA debates. Make use of argument maps to break down and visualize the
-% debate, as shown in \myfig{fig-args-top-level}.
-
-% \begin{figure}[h]
-%   \centering\CaptionFontSize
-%   \includegraphics[width=\linewidth]{arguments/build/top-level.pdf}
-%   \caption[Argument: Top Level]{Top level arguments for and against exceptional access.}
-%   \label{fig-args-top-level}
-% \end{figure}
-
-% The debate over exceptional access takes place on two fundamental levels:
-% \begin{enumerate}
-%   \item Whether cryptosystems with exceptional access represent a net benefit to society
-%   \item Whether and how to conduct exceptional access research
-% \end{enumerate}
-
-% \subsection{Arguments For}
-
-% Why EA is a benefit to society:
-% \begin{itemize}
-%   \item Absolute privacy is not a right
-%   \item Encryption gives too much power to the individual
-%   \item EA would allow more crimes to be prevented and prosecuted
-% \end{itemize}
-
-% Why to research EA:
-% \begin{itemize}
-%   \item EA done well would be more secure and privacy preserving than EA done poorly
-%   \item The threat of legislative action creates urgency
-% \end{itemize}
-
-% \subsection{Arguments Against}
-
-% Why EA is a danger to society:
-% \begin{itemize}
-%   \item Privacy is an absolute right
-%   \item Exceptional access gives too much power to the state
-%   \item Exceptional access is not possible in an acceptably secure manner
-% \end{itemize}
-
-% Why not to research EA:
-% \begin{itemize}
-%   \item With both security and privacy in a poor state, research should be focused on strengthening, not weakening,
-%             them
-%   \item The technical community should not entertain delusions about what is truly possible
-% \end{itemize}
-
-% \subsection{Fallacious Arguments}
-
-% Fallacy definitions come from the helpful University of North Carolina Writing Center \cite{unc_2020}. Also cited:
-% \cite{hanna_2019}.
-
-% Fallacies are as shown in \myfig{fig-args-fallacies}.
-
-% \begin{figure}[h]
-%   \centering\CaptionFontSize
-%   \myincludeargument{comprehensive.fallacies}
-%   \caption[Fallacious Arguments]{Fallacious arguments used in the exceptional access debate.}
-%   \label{fig-args-fallacies}
-% \end{figure}
+% See: committee_decrypting_2018, group_2019, various proposals

### 2020-11-05 arguments/argdown.config.js

@@ -71,28 +71,29 @@ module.exports = {
       },
     },
     fallacies: {
+      process: "export-svg",
+      outputSuffix: ".fallacies",
       dot: {
         graphVizSettings: {
           rankdir: "LR",
         },
       },
-      process: "export-svg",
-      outputSuffix: ".fallacies",
       selection: {
         selectedSections: [
-          "Central Issues",
-          "Response Measures",
           "Exceptional Access",
           "Fallacies",
           "Fallacious Arguments",
         ],
-        selectedTags: ["core", "ea-itself", "fallacy", "fallacious"],
-        selectElementsWithoutTag: true,
       },
     },
     measures: {
       process: "export-svg",
       outputSuffix: ".measures",
+      dot: {
+        graphVizSettings: {
+          ranksep: 0.7,
+        },
+      },
       group: {
         groupDepth: 3,
       },
@@ -101,7 +102,8 @@ module.exports = {
           "Central Issues",
           "Response Measures",
           "Exceptional Access",
-          "Legal Action",
+          "Current Capabilities",
+          "Legal Measures",
           "Arguments for Measures",
         ],
       },

### 2020-11-06 arguments/argdown.config.js
Work on EA types argmaps

@@ -100,6 +100,7 @@ module.exports = {
       selection: {
         selectedSections: [
           "Central Issues",
+          "Desireable Properties",
           "Response Measures",
           "Exceptional Access",
           "Current Capabilities",
@@ -117,6 +118,7 @@ module.exports = {
       selection: {
         selectedSections: [
           "Central Issues",
+          "Desireable Properties",
           "Exceptional Access",
           "DAR EA Classes",
           "DIM EA Classes",

### 2020-11-06 Thesis-Chapter4-Arguments.tex
Work on EA types argmaps
Finish chapter 4, argument maps

@@
 
-\begin{figure}[h]
+\myfig{fig-arg-classes} lists several classes of \ac{EA} and arguments in their favor and disfavor. Classes applicable
+to both \ac{DIM} and \ac{DAR} include using weak cryptography, trusted-party key escrow, and distributed key escrow.
+Using cryptographic puzzles and adding a ghost user apply to the case of \ac{DIM}, and device key escrow applies to
+\ac{DAR}.
+
+\begin{sidewaysfigure}[p!]
   \centering\CaptionFontSize
@@
   \label{fig-arg-classes}
-\end{figure}
+\end{sidewaysfigure}
 
-% What properties would good EA have?
-% easy for LEAs to use? hard? expensive? voluntary? unable to be used for mass surveillance? transparent (attempts are
-%   visible)? alignment of values with security (e.g. discourages bug hoarding)? adaptable (since solution won't be
-%   permanent, either technically or according to collective will as things change rozenshtein_wicked_2018)? suited to
-%   rules and principles (matyas_incommensurability_2018)?
+It should be noted that the \ac{argmap} format works best when displaying the debate surrounding a single conclusion or
+proposal. That leaves room for points and counterpoints to be elaborated in detail, as opposed to creating a tangled web
+of un-elaborated connections. For the sake of brevity, this thesis considers \ac{EA}'s alternatives and variations at a
+high level. For an excellent, thorough discussion of \ac{EA}'s alternatives and variations, see the National Academies
+of Sciences, Engineering, and Medicine's 2018 report titled \ii{Decrypting the Encryption Debate: A Framework for
+Decision Makers} \cite{committee_decrypting_2018}.
 
-\section{Specific Proposals}
+The first \ac{EA} class is to simply use weak cryptographic ciphers or short encryption keys. This was the approach
+taken before \ac{the-cw1}, when strong cryptography was not highly available and subject to export controls. No party
+seriously supports this approach today, as it is irreconcilably contradictory to strong security on many levels.
 
-\begin{figure}[h]
-  \centering\CaptionFontSize
-  \includegraphics[width=\linewidth]{arguments/build/comprehensive.proposals.pdf}
-  \caption{EA Proposals}
-  \label{fig-arg-proposals}
-\end{figure}
+Trusted-party key escrow relies on the an entity or small group of entities to store a key or key recovery information.
+There are a wide variety of implementations \cite{denning_taxonomy_1996}, but by definition this class of \ac{EA}
+carries a crushing weakness: the centralization of extremely sensitive information. Proponents argue that software
+companies' protection of code-signing keys in hardware security modules is evidence that they can protect such sensitive
+data when needed \cite{ozzie_2018}, but opponents point out that such keys are accessed quite rarely used (especially
+when compared to the likely frequency of \ac{EA} requests) and yet they indeed still leak \cite{green_2018}. Once the
+escrowed keys are stolen, as they inevitably would be, every device or message encrypted with them would be at risk.
+This is an example of a case when the \ac{argmap} format doesn't do an argument justice---the argument against any
+centralized access mechanism is extremely powerful.
 
-% See: committee_decrypting_2018, group_2019, various proposals
+Distributed key escrow differs from trusted-party in that it takes strong steps to remove centralized trust. This is
+achieved by using distributed systems to make key extraction difficult and inherently public \cite{phan_key_2017}
+\cite{servan_schreiber_jje_2020}. If the distributed protocols hold up, one advantage of distribution is that this
+approach is that it is more likely to fail by making data unrecoverable than by making unintended data vulnerable,
+though more research into this area is needed. The downside of these systems is their complexity, which could make it
+costly and hard to update.
+
+One class of \ac{DIM} \ac{EA} involves using strong cryptography, but including cryptographic puzzle pieces as message
+metadata. These pieces are designed to reveal the encryption key after the surveiller spends considerable computational
+power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes \ii{mass} surveillance, but it
+also enables \ii{arbitrary} surveillance by anyone with enough computing power. Use in common cases would be impractical
+due to high cost, but use in high profile cases would be less effective due to the likelihood of \ac{EA} workarounds
+employed by sophisticated adversaries. Additionally, the calibration of puzzle difficulty is dependent on the state of
+computing. Since opponents can record data and decrypt it later when computational capabilities have changed, the data
+becomes steadily less secure over time.
+
+Another \ac{DIM} proposal is to ``conference call'' authorities in as a ghost participant in the conversation. It was
+notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson \cite{levy_robinson_2018}. Their writeup begins
+with some core ideas of this thesis:
+
+\begin{displayquote}
+In any discussion of cyber security, details matter.
+
+Unfortunately, it's the details that are missing from the discussion around lawful access to commodity end-to-end
+encrypted services and devices (often called the ``going dark'' problem). Without details, the problem is debated as a
+purely academic abstraction concerning security, liberty, and the role of government.
+
+There is a better way that doesn’t involve, on one side, various governments, and on the other side lawyers,
+philosophers, and vendors' PR departments continuing to shout at each other. If we can get all parties to look at some
+actual detail, some practices and proposals---without asking anyone to compromise on things they fundamentally believe
+in---we might get somewhere.
+\cite{levy_robinson_2018}
+\end{displayquote}
+
+As goodwill participants in the conversation, trying to spur discussion around specific proposals ``without people being
+vilified for having a point of view or daring to work on this as a problem,'' they deserve credit and thanks.
+
+Unfortunately, the proposal they share is indeed problematic. The ghost user proposal prides itself on not compromising
+strong encryption. This is a good property, but compromising a messaging app's authorization protocol is as dangerous as
+compromising its encryption \cite{callas_1_2019}. Fundamentally, it suffers from the same weakness as trusted-party key
+escrow; the introduction of an authorization vulnerability in the messaging service provider's platform amounts to a
+highly risky centralization of access capability \cite{schneier_ghost_2019}. It has the same nature, if a different
+form, as rejected key escrow proposals.
+
+Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the
+device itself. While not necessary to classify as device key escrow, most of these proposals include physical possession
+requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
+\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features precludes mass surveillance and ensure transparency to the
+device user. This approach relies on secure hardware and a method of authorizing the recovery request to the device
+itself. The authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or
+distributed systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving,
+are not totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional
+escrow \ac{EA}.
+
+Device key escrow's greatest weaknesses, however, still represent signs of progress. It is true that it relies on a
+device's secure enclave, but devices already do---it is the same hardware devices that manages device unlock and
+encryption functionality in the first place. Adding functionality increases the attack surface, but it does not add a
+completely new threat vector. It is also true that the process is subject to the same problems as traditional key escrow
+after the device has been obtained, but that is precisely the progress that has been made---\ii{after the device has
+been obtained}. Leveraging advances in secure hardware design has meaningfully changed the \ac{EA} risk profile.
+
+In 2019, Carnegie Mellon assembled an ideologically diverse group of policy and security experts to engage in the kind
+of cross-disciplinary debate this thesis advocates. After considering all the technical corners of the situation, their
+report goes into detail on why \ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated
+adversaries) focusing on \ac{DAR} in mobile phones using device key escrow requiring physical access is the most
+tractable technical case to consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed
+with a dose of lawful hacking is best suited for advanced opponents, and device key escrow represents actual progress as
+a path towards lower-risk \ac{EA}.
+
+Device key escrow still has a long way to go. There are few proposals, what few there are are high level, and questions
+of scaling and administration have not been addressed. However, as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s
+Levy and Robinson have pointed out, focusing on specific proposals has the opportunity to identify mutually agreeable
+solutions---or at least refine the problem understanding enough that the next round of debate can be better informed.
+
+It is this task that we turn to next. The following chapter analyzes one particular device-escrow proposal in moderate
+detail, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
+\cite{savage_lawful_2018}.
+
+% \section{Specific Proposals}
+%% NOTE: not worth going into this section. Just describe a few and go straight to one of the most interesting.
+
+% \begin{figure}[h]
+%   \centering\CaptionFontSize
+%   \includegraphics[width=\linewidth]{arguments/build/comprehensive.proposals.pdf}
+%   \caption{EA Proposals}
+%   \label{fig-arg-proposals}
+% \end{figure}
+
+% % See: committee_decrypting_2018, group_2019, various proposals

### 2020-11-06 Thesis-Glossary.tex
Write contributing factors map section
Finish chapter 4, argument maps

@@
 
-\newterm{horsemen}{the Four Horsemen}{
+\newterm{horsemen}{Four Horsemen}{
     Four reasons that governments and law enforcement agencies use to undercut public support for strong encryption. The
@@
 \newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
+\newabbreviation{GCHQ}{GCHQ}{(UK) Government Communications Headquarters}
 \newabbreviation{IBIS}{IBIS}{issue-based information system}

### 2020-11-06 arguments/argdown.config.js

@@ -112,6 +112,11 @@ module.exports = {
     classes: {
       process: "export-svg",
       outputSuffix: ".classes",
+      dot: {
+        graphVizSettings: {
+          ranksep: 0.7,
+        },
+      },
       group: {
         groupDepth: 3,
       },
@@ -124,23 +129,26 @@ module.exports = {
           "DIM EA Classes",
           "Arguments for EA Types",
         ],
-      },
-    },
-    proposals: {
-      process: "export-svg",
-      outputSuffix: ".proposals",
-      group: {
-        groupDepth: 3,
-      },
-      selection: {
-        selectedSections: [
-          "Central Issues",
+        excludeStatements: [
           "Exceptional Access",
-          "DAR EA Proposals",
-          "DIM EA Proposals",
-          "Arguments for EA Proposals",
         ],
       },
     },
+    // proposals: {
+    //   process: "export-svg",
+    //   outputSuffix: ".proposals",
+    //   group: {
+    //     groupDepth: 3,
+    //   },
+    //   selection: {
+    //     selectedSections: [
+    //       "Central Issues",
+    //       "Exceptional Access",
+    //       "DAR EA Proposals",
+    //       "DIM EA Proposals",
+    //       "Arguments for EA Proposals",
+    //     ],
+    //   },
+    // },
   },
 };

### 2020-11-06 Thesis-Chapter5-ThreatModel.tex
Finish Chapter 3 draft
Clean up arguments chapter notes
Prepare chapter 5 and start rearranging LDAWMSR diagram

@@
 
+\section{Developing a Threat Model}
+
+% For a wholistic threat model, include both the threats of government (legislation) and of criminal abusers of your
+% system.
+
+% For analysis of threat actors, see bindenagel_discussion_2017
+
+\section{Basic Data at Rest}
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone.png}
+    \caption{The basic encrypted mobile phone data flow diagram.}
+    \label{fig-dfd-dar}
+\end{figure}
+
+\section{The LDAWMSR Proposal}
+
 \begin{figure}[h]
@@
     \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
-    \caption[A DAR EA Proposal]{A DAR EA Proposal based on \cite{savage_lawful_2018}.}
-    \label{fig-dfd-lawful}
+    \caption{The LDAWMSR data flow diagram.}
+    \label{fig-dfd-ldawmsr}
 \end{figure}
 
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/DIM-level-1-crypto-E2E-signal-X3DH.png}
-    \caption[The Signal X3DH Protocol]{The Signal session creation protocol \cite{marlinspike_x3dh_2016}.}
-    \label{fig-dfd-lawful}
-\end{figure}
-
-\begin{figure}[h]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/DIM-level-1-crypto-E2E-signal-Double-Ratchet.png}
-    \caption[The Signal Double Ratchet Protocol]{The Signal \ac{E2EE} protocol \cite{marlinspike_2016}.}
-    \label{fig-dfd-lawful}
-\end{figure}
-
-% For analysis of threat actors, see bindenagel_discussion_2017
-
-% Generalize for DAR and DIM, then apply to specific examples. This proves that my analysis is helpful. Ex: apply to email: is a DIM _and_ DAR problem due to the server.
+\section{Discussion of Threats}
 
 %%%%%%%%%%%%
+% Old notes for reference
+% -----------------------
 % Chapter 4, threat models

### 2020-11-07 Thesis-Chapter4-Arguments.tex
Fix some typos and reshap LDAMSR diagram

@@
 escrow; the introduction of an authorization vulnerability in the messaging service provider's platform amounts to a
-highly risky centralization of access capability \cite{schneier_ghost_2019}. It has the same nature, if a different
-form, as rejected key escrow proposals.
+highly risky centralized access capability \cite{schneier_ghost_2019}. It has the same nature, if a different form, as
+rejected key escrow proposals.
 
@@
 requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
-\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features precludes mass surveillance and ensure transparency to the
+\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude mass surveillance and ensure transparency to the
 device user. This approach relies on secure hardware and a method of authorizing the recovery request to the device
@@
 Device key escrow's greatest weaknesses, however, still represent signs of progress. It is true that it relies on a
-device's secure enclave, but devices already do---it is the same hardware devices that manages device unlock and
+device's secure enclave, but devices already do---it is the same hardware device that manages device unlock and
 encryption functionality in the first place. Adding functionality increases the attack surface, but it does not add a

### 2020-11-07 Thesis-Chapter5-ThreatModel.tex
Fix some typos and reshap LDAMSR diagram
Begin threat model chapter

@@
 
-In this chapter, develop a threat model using data flow diagrams. May need separate ones for different scenarios (data
-at rest on encrypted devices, data in motion using E2E), but want at least one high level diagram encapsulating
-everything.
+This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
+``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}.
+Lacking a concise name, this proposal will be referred to by its acronym, \ldawmsR. Note that the creators of the
+proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to emphasize the role of legal
+process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an \ac{EA} proposal, but use the
+term entirely in the spirit of their paper.
+
+% TODO: remove this paragraph if I end up changing EA to LA
 
@@
 
-% For a wholistic threat model, include both the threats of government (legislation) and of criminal abusers of your
-% system.
+The essential threat modeling questions are ``what are we building'' and ``what can go wrong.'' The first step to
+answering those questions is to establish goals for the former and a list of threats in anticipation of the latter. The
+\ldawmsr proposal describes a threat model against which it was designed. This model takes cues from that one but is
+intended to be more generally applicable for any \ac{DAR} \ac{EA} proposal.
 
-% For analysis of threat actors, see bindenagel_discussion_2017
+\newcommand{\modelstart}[0]{\begin{itemize}}
+\newcommand{\modelitem}[2]{ % Name, description
+    \item \textbf{#1} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
+}
+\newcommand{\modelend}{\end{itemize}}
+
+% \newcommand{\modelstart}{\begin{itemize}}
+% \newcommand{\modelitem}[2]{\item \textbf{#1.} #2}
+% \newcommand{\modelend}{\end{itemize}}
+
+\subsection{Goals}
+
+The goals of an \ac{EA} cryptosystem are taken from the metrics of \mychap{chap-arguments}. Their justification stems
+from the preceding arguments. Each goal is listed according to priority, though none are absolute.
+
+\modelstart
+
+\modelitem{Security}{The system should minimize increased security risk with reference to the current state of the art.
+Risk is considered both with respect to individual devices and to the integrity of the entire class of devices.}
+
+\modelitem{Protection of civil liberties}{The system should include measures to prevent mass surveillance and abuse by
+political regimes that violate civil rights. This includes technologically enforcing the fundamental right to privacy
+but upon warranted search pursuant to the rule of law.}
+
+\modelitem{Transparency}{The system should be auditable by the public to enable trust that this significant power is
+being used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}
+
+\modelitem{Law enforcement utility}{The system should actually be useful to law enforcement in terms of reliability,
+speed, and cost.}
+
+\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem as well
+as the distribution of costs associated with its development and administration. Even though this is a security-oriented
+threat model, architectural decisions determine the system's economics, and economics determines the system's ultimate
+feasibility.}
+
+\modelend
+
+\subsection{Threat Actors}
+
+\ac{EA} threat models necessarily include the full cast of traditional threat actors.
+
+\modelstart
+
+\modelitem{Criminals: ordinary, organized, cyber}{Criminals represent a threat to the \ac{EA} system by trying to use it
+to gain access without authority. Ordinary thieves may want to spy on a target or steal and resell phones. They will
+have physical access but only simple tools, using ordinary device interfaces. Organized criminals will have the same
+motives but considerably more advanced technical capabilities, exerting full control over all interfaces. Cyber
+criminals have a wide spectrum of motivations and capabilities. Their significant contribution to the threat landscape
+is that they will target the backend systems that operate the \ac{EA} system in order to steal and likely sell whatever
+access capabilities they can find.}
+
+\modelitem{Insider threat}{The insider threat is someone within the \ac{EA} recovery process, whether at the law
+enforcement agency or a vendor, that seeks to abuse the system for their own purposes. They may represent the
+motivations and capabilities of any of the criminal categories, but they begin with privileged access somewhere in the
+process.}
+
+\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are always looking for means to escalate
+access and tamper with and steal information. They have world-class expertise and computing power, which they will apply
+to both individual devices and backend systems.}
+
+\modelitem{Authoritarian regimes}{States that do not respect lawful process will try to compel administers of an \ac{EA}
+system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking capability.}
+
+\modelend
+
+To make sense of an \ac{EA} proposal in the face of the stated goals and such a formidable list of threat actors, one
+must include two non-traditional threats. To do so is to step back and look at the situation from a broader view than
+security experts usually take.
+
+\modelstart
+
+\modelitem{The platform abuser}{Why bother with \ac{EA} at all? Because the device user may actually be a threat in the
+larger safety context. The end user is not the most important thing. This is a crucial point.
+
+All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of this value system is some most
+important ``thing.'' Traditional device security threat models make the end user the most important thing. This is a
+very good rule. It leads to good security priorities and more importantly, honors the user's rights and autonomy.
+
+To stray from this rule is dangerous, but---make no mistake---it is already done on a regular basis. Software that can
+force updates values the vendor's judgement over the user's (even if it is a good idea). Software that enforces Digital
+Rights Management values corporations' rights over the user's (sometimes to extremes \cite{eff_2020}). Software that
+doesn't encrypt user communications at all says the service provider's ability to scrape personal information to create
+targeted ads is more important than the user's privacy. Software that enables \ac{EA} values lawful investigatory
+process over the user's ultimate privacy. Subversion of the user's will is not a pleasant concept, but a spade is a
+spade.
+
+In the broader threat model, the platform abuser is someone who is using the device encryption for cover of illicit
+activities. They represent a threat to public safety.}
+
+\modelitem{Hawkish lawmakers}{If one's very highest priorities are security and privacy for society as a whole, then
+aggressive lawmakers that have the power to \ii{mandate} arbitrary data access represent perhaps the largest threat of
+all. Crypto-anarchists can encrypt their data above the compromised layer all they want, but the vast majority of
+individuals using default products and all above-the-table organizations will abide by U.S. law.
+
+The hawkish lawmaker is a threat to the security of a cryptosystem (\ac{EA} or not) in that they may judge it too
+inhibitive to law enforcement and outlaw its very use, mandating something weaker in its place.}
+
+\modelend
+
+\subsection{Out of Scope}
+
+The following considerations are out of scope for this threat model.
+
+\modelstart
+
+\modelitem{Encyption workarounds}{Encryption workarounds are always in play for essentially any threat actor at any
+time. Discussion for defending against those is out of scope.}
+
+\modelitem{\ac{EA} workarounds}{The user is capable of evading \ac{EA} mechanisms by encrypting data at a higher level
+in the tech stack, e.g. by encrypting files before saving them to disk. This has already been discussed, and it out of
+scope for evaluating the \ac{EA} system itself.}
+
+\modelitem{Manipulation of internal hardware}{Hyper-advanced hardware analysis and manipulation is a real threat from
+foreign intelligence agencies. Defending against these attacks is a separate problem.}
+
+\modelitem{Breakdown in rule of law}{This threat model assumes that the legal process used to obtain the warrant is
+valid. Additionally, all forms of surveillance, even when protected from mass surveillance, require oversight.
+Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
+However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
+
+\modelend
+
+% What are we building?: proposal DFDs
+% What can go wrong?: elicitation of threats
 
 \section{Basic Data at Rest}
+
+Answering ``what are we building'' and comparing it with the state of the art requires an understanding of the state of
+the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
+\ldawmsR. \myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to their
+public documentation \cite{apple_2020}.
 
@@
     \caption{The basic encrypted mobile phone data flow diagram.}
-    \label{fig-dfd-dar}
+    \label{fig-dfd-iphone}
 \end{figure}
 
-\section{The LDAWMSR Proposal}
+This \ac{DFD}, and all that follow, is at a specific level of abstraction, simplified in the spirit of ``all models are
+wrong, but some are useful.'' The \acp{DFD} are intended to be useful for communication and threat elicitation only, but
+are accurate in that capacity.
 
-\begin{figure}[h]
+As the diagram illustrates, several steps occur between the user entering her \ac{PIN} and unlocking of the device. The
+operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines it with the device-specific
+Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is finally handed to
+the Encryption Module to perform encryption and decryption at a hardware level between the operating system and actual
+encrypted storage.
+
+The \ac{UID} is not depicted in key storage because it is actually burned into the secure processor silicon, and cannot
+be read by software. The Volume key is stored encrypted by the Class key. When a user updates his \ac{PIN}, the
+encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors never
+handle any keys or secret information besides the \ac{PIN}, and due to inline encryption in the Encryption Module,
+persistent storage never handles and \ac{plaintext} data.
+
+\section{The \ldawmsr Proposal}
+
+\ldawmsr introduces device key escrow and an unexposed hardware interface that mediates \ac{EA} requests. When prompted,
+the hardware interface produces a Device ID. The law enforcement officer (assuming proper circumstances) receives this
+Device ID and reports it to her agency. The agency signs a request including the Device ID and presents it to the
+vendor's access compliance department. The vendor authenticates the request, audits the corresponding legal documents,
+and looks up the corresponding Device Seal key. This key is encrypted using the same agency's public key and returned to
+the agency. The agency decrypts the Device Seal key and issues it to the agent, who enters it through the special
+interface. The Secure Enclave Processor uses the private Device Seal key to decrypt the Class key, which was saved in
+key storage encrypted by the device's public Device Seal key, and the unlock process proceeds as normal from there. Each
+of these steps is analyzed in closer detail.
+
+\begin{figure}[p]
     \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
+    \includegraphics[width=0.81\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
     \caption{The LDAWMSR data flow diagram.}
@@
 \section{Discussion of Threats}
+
+\section{Response to Frameworks}
+
+% How does it answer (or not) committee_decrypting_2018's framework questions? What needs improvement?
 

### 2020-11-11 Thesis-Chapter2-Background.tex
Write EA and alternatives section
Begin threat model chapter
Progress on threat model

@@
 
-\mytab{table:security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
+\mytab{table-security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
 cryptographic foundations, availability indirectly relies on each other property, and authorization schemes are
@@
 \begin{table}[h]
-    \caption[Security Properties and Cryptography]{The Cryptographic Basis of Security Properties}
-    \label{table:security-and-crypto}
+    \caption{The Cryptographic Basis of Security Properties}
+    \label{table-security-and-crypto}
     \centering
-    \begin{tabular}{|l|l|}
+    \begin{tabular}{ |l|l| }
         \hline
@@
 \subsection{Alternative Approaches}
+\label{sec-alt-approaches}
 
@@
 external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
-flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is
-the primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
+flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
+primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
 \acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.

### 2020-11-11 Thesis-Chapter5-ThreatModel.tex
Progress on threat model

@@
 
-\ldawmsr introduces device key escrow and an unexposed hardware interface that mediates \ac{EA} requests. When prompted,
-the hardware interface produces a Device ID. The law enforcement officer (assuming proper circumstances) receives this
-Device ID and reports it to her agency. The agency signs a request including the Device ID and presents it to the
-vendor's access compliance department. The vendor authenticates the request, audits the corresponding legal documents,
-and looks up the corresponding Device Seal key. This key is encrypted using the same agency's public key and returned to
-the agency. The agency decrypts the Device Seal key and issues it to the agent, who enters it through the special
-interface. The Secure Enclave Processor uses the private Device Seal key to decrypt the Class key, which was saved in
-key storage encrypted by the device's public Device Seal key, and the unlock process proceeds as normal from there. Each
-of these steps is analyzed in closer detail.
+\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{EA} requests, and a vendor
+mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time vaulting,
+authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}; see the original paper
+for full detail \cite{savage_lawful_2018}.
+
+\begin{table}[h]
+  \caption{Central Concepts in the \ldawmsr Proposal}
+  \label{table-ldawmsr-concepts}
+  \begin{FlushLeft}
+    \begin{tabular}{ |l|p{6.3cm}|p{6.3cm}| }
+      \hline
+      \thead{Concept} & \thead{Implementation} & \thead{Outcome} \\ \hline
+      Self-escrow
+      & The Class key is escrowed in the device's Secure Enclave (or equivalent) using a write-only component; i.e.,
+        software can update the key, but only an unexposed hardware interface can read it.
+      & Since the Class key itself is being stored, there is no change to the security of the underlying cryptographic
+        protocols. Physical possession and partial disassembly is required, which introduces practical barriers to mass
+        surveillance.
+      \\ \hline
+      Time vaulting
+      & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period
+        (e.g., 72 hours).''
+      & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for mass surveillance.
+      \\ \hline
+      Authorization
+      & After the lockup period, the requestor must provide evidence of authorization. This comes from a shared secret
+        between the device and the manufacturer which can be obtained via legal process.
+        % Vendor authorization escrow (as opposed to _key_ escrow)
+      & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to
+        the device to unlock it.
+      \\ \hline
+      Transparency
+      & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has
+        been unlocked. The device firmware detects this modification.
+      & The device user has evidence that the device has been unlocked, preventing covert usage.
+      \\ \hline
+    \end{tabular}
+  \end{FlushLeft}
+\end{table}
+
+Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
+secure hardware offers, and \ldawmsr demonstrates this capabilities perfectly. The greatest of these capabilities are
+physical possession requirements and strong transparency. Physical possession requirements strongly mitigate mass
+surveillance and, as Savage puts it, ``provides a more intuitive compatibility with common under-standings of the
+government’s reasonable law enforcement powers (e.g., the ability to seize physical property under court order) than
+more information-centric approaches, which may appear covert by comparison'' \cite{savage_lawful_2018}. Transparency is
+also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
+detectable changes, mitigating both mass surveillance and common criminal abuse.
+
+One clear weakness in the design is the shared secret between the device and the manufacturer used in the authorization
+protocol. The key may be stored on the device, but one can access it if they have this shared secret. This is very much
+like trusted-party key escrow where the vendor is the trusted party, only they store the authentication key instead of
+the unlock key itself. Due to this similarity, one of the biggest \ac{EA} problems of centralized risk is not fully
+addressed. This is discussed further in \mysec{sec-prop-threats}.
+
+\ldawmsR's full data flow is illustrated in \myfig{fig-dfd-ldawmsr}. The proposal offers a few variations of the
+self-escrow and authorization protocols. This figure uses the asymmetric key pair approach for Class key encryption in
+the Secure Enclave, which has stricter hardware requirements; however, Savage offers a symmetric key variant that
+achieves the same purposes. The asymmetric key pair approach was chosen here because it is semantically clearer and
+easier to illustrate.
 
@@
     \centering\CaptionFontSize
-    \includegraphics[width=0.81\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
+    \includegraphics[width=0.9\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
     \caption{The LDAWMSR data flow diagram.}
@@
 
+The lawful access use case operates as follows: law enforcement obtains a warrant to conduct a search, including the
+suspect's mobile device. They obtain the phone, disassemble it to expose the \ac{EA} hardware interface, and begin the
+time-vaulted access request. After the lockout period is completed, the device offers its Device ID only. The law
+enforcement agency's digital forensics department digitally signs the Device ID and sends it to the device vendor,
+alongside the warrant and accompanying proof of legal authorization. The vendor's access compliance department audits
+the legal documents and submits the digitally signed Device ID to a secure computing environment such as a Hardware
+Security Module. The secure environment authenticates the digital signature, looks up the device's private Device Seal
+Key and authorization value, and encrypts them with the law enforcement agency's public key. These encrypted artifacts
+are returned to the agency, which decrypts and submits them to the device through the dedicated interface. First is
+the authorization value, which is hashed and compared to the stored hash. If that values match, the device accepts and
+uses the private Device Seal Key to decrypt the Class key, which was escrowed after being encrypted but the public
+Device Seal Key. Once the Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for forensic
+investigation.
+
 \section{Discussion of Threats}
+\label{sec-prop-threats}
+
+% Full lifecycle--how do the keys get in the DB!
+
+The proposal has been illustrated. What threats does it introduce? Recall from \mysec{sec-crypto-basics} that security
+can be expressed in the properties of authentication, integrity, non-repudiation, confidentiality, availability, and
+authorization. Threats are often expressed as the negative of these properties: spoofing, tampering, repudiation,
+information disclosure, denial of service, and elevation of privileged \cite{shostack_threat_2014}. This section
+analyzes \ldawmsr using these categories.
+
+\subsection{Spoofing}
+
+\begin{table}[h]
+  \caption{Spoofing Threats}
+  \label{table-ldawmsr-concepts}
+  \begin{FlushLeft}
+    \begin{tabular}{ |l|p{2cm}|p{2cm}|p{4.8cm}|p{4.8cm}| }
+      \hline
+      \thead{ID} & \thead{Identity \\ Spoofed} & \thead{With \\ Respect To} & \thead{Effect} & \thead{Mitigation}
+      \\ \hline
+      S1 & Device & Officer
+      & Planted device to unlock unauthorized device.
+      & Require it to be in the forensics department itself.
+      \\ \hline
+    \end{tabular}
+  \end{FlushLeft}
+\end{table}
+
+\ii{S1} is Eran Tromer's attack on CLEAR.
+
+\subsection{Tampering}
+
+\subsection{Repudiation}
+
+\subsection{Information Disclosure}
+
+\subsection{Denial of Service}
+
+\subsection{Elevation of Privilege}
 
@@
 
+% Or just, relationship to my goals?
+
 % How does it answer (or not) committee_decrypting_2018's framework questions? What needs improvement?
+
+% Reflection on how much progress this represents
 

### 2020-11-12 Thesis-Chapter5-ThreatModel.tex
Finish spoofing, begin tampering threats

@@
   \label{table-ldawmsr-concepts}
-  \begin{FlushLeft}
+  \FlushLeft
+  \small
   \begin{tabular}{ |l|p{6.3cm}|p{6.3cm}| }
@@
     Time vaulting
-      & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period
-        (e.g., 72 hours).''
+    & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period (e.g.,
+      72 hours).''
     & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for mass surveillance.
@@
       between the device and the manufacturer which can be obtained via legal process.
-        % Vendor authorization escrow (as opposed to _key_ escrow)
-      & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to
-        the device to unlock it.
+    & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to the
+      device to unlock it.
     \\ \hline
     Transparency
-      & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has
-        been unlocked. The device firmware detects this modification.
+    & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has been
+      unlocked. The device firmware detects this modification.
     & The device user has evidence that the device has been unlocked, preventing covert usage.
@@
   \end{tabular}
-  \end{FlushLeft}
 \end{table}
@@
 alongside the warrant and accompanying proof of legal authorization. The vendor's access compliance department audits
-the legal documents and submits the digitally signed Device ID to a secure computing environment such as a Hardware
-Security Module. The secure environment authenticates the digital signature, looks up the device's private Device Seal
-Key and authorization value, and encrypts them with the law enforcement agency's public key. These encrypted artifacts
-are returned to the agency, which decrypts and submits them to the device through the dedicated interface. First is
-the authorization value, which is hashed and compared to the stored hash. If that values match, the device accepts and
-uses the private Device Seal Key to decrypt the Class key, which was escrowed after being encrypted but the public
-Device Seal Key. Once the Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for forensic
-investigation.
+the legal documents and submits the digitally signed Device ID to a secure computing environment such as a \ac{HSM}. The
+secure environment authenticates the digital signature, looks up the device's private Device Seal Key and authorization
+token, and encrypts them with the law enforcement agency's public key. These encrypted artifacts are returned to the
+agency, which decrypts and submits them to the device through the dedicated interface. First is the authorization token,
+which is hashed and compared to the stored hash. If that values match, the device accepts and uses the private Device
+Seal Key to decrypt the Class key, which was escrowed after being encrypted but the public Device Seal Key. Once the
+Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for forensic investigation.
+
+This concludes an summarized overview of the system as described in the proposal. However, in a real system, the entire
+information lifecycle matters. In order to establish such a system, vendors and law enforcement must establish
+mechanisms to populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} is a
+\ac{DFD} for these scenarios.
+
+\begin{figure}[h!]
+  \centering\CaptionFontSize
+  \includegraphics[width=0.9\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR-setup.png}
+  \caption{The LDAWMSR maintenance data flow diagram.}
+  \label{fig-dfd-ldawmsr-setup}
+\end{figure}
 
@@
 
+To save space, the tables use shorthand for law enforcement (LE); law enforcement officer (LEO); law enforcement agency
+(LEA); Device ID (DID);
+
+% TODO: Check list of acronyms up to date here ^^
+
+\newcommand{\threattablesettings}{\FlushLeft \small}
+
+\newcommand{\threattablebegin}[3]{
+  \begin{FlushLeft}
+    \small
+    \begin{longtable}{ lp{3cm}p{5.6cm}p{5.6cm} }
+      \caption{#1 Threats}
+      \label{table-#2}
+      \\ \toprule
+      \thead{ID} & \thead{#3} & \thead{Effect} & \thead{Mitigation}
+      \\ \midrule
+      \endfirsthead
+      \caption[]{#1 Threats (continued)}
+      \\
+      \toprule
+      \thead{ID} & \thead{#3} & \thead{Effect} & \thead{Mitigation}
+      \\
+      % \midrule
+      \endhead
+      \endfoot
+      \bottomrule
+      \endlastfoot
+}
+
+\newcommand{\threattableend}{
+    \end{longtable}
+  \end{FlushLeft}
+}
+
 \subsection{Spoofing}
 
-\begin{table}[h]
-  \caption{Spoofing Threats}
-  \label{table-ldawmsr-concepts}
-  \begin{FlushLeft}
-    \begin{tabular}{ |l|p{2cm}|p{2cm}|p{4.8cm}|p{4.8cm}| }
-      \hline
-      \thead{ID} & \thead{Identity \\ Spoofed} & \thead{With \\ Respect To} & \thead{Effect} & \thead{Mitigation}
+Spoofing threats are listed in \mytab{table-spoofing}. Notably, \ii{S2} is Eran Tromer's attack \cite{tromer_2018} on
+computer scientist Ray Ozzie's related \ac{DAR} \ac{EA} protocol called CLEAR \cite{ozzie_2018}. Savage acknowledges the
+influence that CLEAR had on \ldawmsR, and specifically included a note that while this proposal is technically
+vulnerable to the same threat, the mitigations listed here would help, and furthermore the likelihood of such an attack
+is low and the difficulty high.
+
+\threattablebegin{Spoofing}{spoofing}{Spoof}
+  S1 & Device is spoofed to LE via plant
+  & Attackers plant a device to get LE to unlock it for them.
+  & Require the device to be in the forensics department itself the entire time it is unlocked.
   \\ \hline
-      S1 & Device & Officer
-      & Planted device to unlock unauthorized device.
-      & Require it to be in the forensics department itself.
+  S2 & Device is spoofed to LE via forged DID
+  & Attackers rig a device to report a different DID and record the entered unlock artifacts, allowing them into a
+    target device.
+  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage, extract its data, and destroy
+    it.
   \\ \hline
-    \end{tabular}
-  \end{FlushLeft}
-\end{table}
-
-\ii{S1} is Eran Tromer's attack on CLEAR.
+  S3 & LEO is spoofed to LEA
+  & Attackers pose as a LEO while the LEA relays back unlock artifacts.
+  & Require the device to be in the forensics department itself the entire time it is unlocked.
+  \\ \hline
+  S4 & LEA is spoofed to vendor during key exchange
+  & Attackers pose as an LEA in order to spoof future access requests.
+  & Vendor key exchange process involves thorough in-person verification.
+  \\ \hline
+  S5 & LEA is spoofed to vendor during access request
+  & Attackers pose as a LEA and submit an access request. Requires forged legal documents and a stolen LEA private key
+    (unless attacker is an insider or has executed a spoofed key exchange).
+  & Vendor independently verifies all legal documents. Vendor supports a key revocation and update process that involves
+    the same level of in-person verification.
+  \\ \hline
+  S6 & Manufacturer is spoofed to vendor
+  & Attackers pose as the manufacturer to inject malicious data into the DID-Seal database. Must be combined with other
+    attacks to be of use.
+  & Vendor authenticates data transfers between manufacturing and corporate (does not address a compromised manufacturer
+    network or an inside threat).
+\threattableend
 
 \subsection{Tampering}
+
+Tampering threats are listed in \mytab{table-tampering}.
+
+\threattablebegin{Tampering}{tampering}{Tamper}
+  T1 & Defensive user corrupts DSK, DID, or token
+  & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
+    hardware manipulation tools.
+  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
+    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
+    to this threat.)
+  \\ \hline
+  T2 & Defensive user destroys the \ac{EA} hardware interface
+  & LE attempts at perform \ac{EA} will fail due to damaged hardware.
+  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
+    associated with deploying and using the system.
+  \\ \hline
+  T3 & LE insider tampers with DID before digital signature
+  & .
+  & .
+\threattableend
 

### 2020-11-12 Thesis-Chapter5-ThreatModel.tex
Develop more threats

@@
 
+\modelitem{Supply chain attacks}{Protecting device hardware and software supply chains is necessary to prevent
+surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and software are
+considered to be as-intended, though they are assumed to contain the standard amount of bugs and side-channels.}
+
 \modelitem{Manipulation of internal hardware}{Hyper-advanced hardware analysis and manipulation is a real threat from
@@
 
-\modelitem{Breakdown in rule of law}{This threat model assumes that the legal process used to obtain the warrant is
-valid. Additionally, all forms of surveillance, even when protected from mass surveillance, require oversight.
-Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
-However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
+\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
+warrant is valid. Additionally, all forms of surveillance, even when protected from mass surveillance, require
+oversight. Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA}
+proposals. However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
 
@@
   \small
-  \begin{tabular}{ |l|p{6.3cm}|p{6.3cm}| }
+  \begin{tabular}{ |l|p{6.4cm}|p{6.4cm}| }
     \hline
@@
 
-\begin{figure}[h!]
+\begin{figure}[p!]
   \centering\CaptionFontSize
-  \includegraphics[width=0.9\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR-setup.png}
+  \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR-setup.png}
   \caption{The LDAWMSR maintenance data flow diagram.}
@@
 
-% Full lifecycle--how do the keys get in the DB!
-
 The proposal has been illustrated. What threats does it introduce? Recall from \mysec{sec-crypto-basics} that security
@@
 information disclosure, denial of service, and elevation of privileged \cite{shostack_threat_2014}. This section
-analyzes \ldawmsr using these categories.
+analyzes \ldawmsr using these categories. Threats that fit under multiple categories, such as tampering for the effect
+of repudiation, are listed only once.
 
@@
     \small
-    \begin{longtable}{ lp{3cm}p{5.6cm}p{5.6cm} }
+    \begin{longtable}{ lp{3cm}p{5.2cm}p{6cm} }
       \caption{#1 Threats}
@@
   T3 & LE insider tampers with DID before digital signature
+  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
+  & Randomize DID's to make them unguessable, meaning a target device must be in possession in order for the attacker to
+    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
+    device S/N; the \ac{HSM} would hash the value and verify a match before returning the unlock artifacts.
+  \\ \hline
+  T4 & Attacker tampers with hardware transparency mechanism
+  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
+    advanced hardware manipulation tools.
+  & Manufacture the device to require destructive disassembly to access the interface.
+\threattableend
+
+\subsection{Repudiation}
+
+Repudiation threats are listed in \mytab{table-repudiation}.
+
+\threattablebegin{Repudiation}{repudiation}{Repudiation}
+  R1 & Attacker or authoritarian government repudiates unlock attempt
+  & The user is unaware that someone tried to access their device.
+  & Manufacture the device to require destructive disassembly to access the interface, or use the same hardware
+    transparency mechanism to mark mere access attempts as well.
+  \\ \hline
+  R2 & Authoritarian government repudiates demanding vendor assistance
+  & Governments not respecting the legal framework governing the system can bully the vendor and try to force access
+    while maintaining deniability.
+  & This is close to being an out of scope threat under the ``rule of law'' exception. However, one could design the
+    system to require requests to be recorded on a public ledger. This opens up several new difficulties.
+  \\ \hline
+  R3 & LEA repudiates having unlocked a device
+  & The device is undeniably unlocked, but the LEA claims it was not them, allowing possible violations of civil
+    liberties and harming transparency.
+  & ...
+  \\ \hline
+  R4 & Vendor repudiates having responded to an access request
+  & The device is undeniably unlocked, but the vendor claims they did not provide the unlock artifacts, allowing the
+    vendor to cover up leaks or unsavory policy decisions by suggesting the device was instead hacked. Harms
+    transparency.
+  & ...
+\threattableend
+
+\subsection{Information Disclosure}
+
+Information disclosure threats are listed in \mytab{table-disclosure}.
+
+\threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
+  I1 & .
+  & .
+  & .
+  \\ \hline
+  I2 & .
   & .
@@
 
-\subsection{Repudiation}
-
-\subsection{Information Disclosure}
-
 \subsection{Denial of Service}
 
+Denial of service threats are listed in \mytab{table-dos}.
+
+\threattablebegin{Denial of Service}{dos}{Service Denied}
+  D1 & .
+  & .
+  & .
+  \\ \hline
+  D2 & .
+  & .
+  & .
+\threattableend
+
 \subsection{Elevation of Privilege}
+
+Elevation of privilege are listed in \mytab{table-privilege}.
+
+\threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
+  E1 & .
+  & .
+  & .
+  \\ \hline
+  E2 & .
+  & .
+  & .
+\threattableend
 

### 2020-11-13 Thesis-Chapter5-ThreatModel.tex
Finish STRIDE threats

@@
 
-\begin{figure}[p!]
+\begin{figure}[H]
   \centering\CaptionFontSize
@@
 authorization. Threats are often expressed as the negative of these properties: spoofing, tampering, repudiation,
-information disclosure, denial of service, and elevation of privileged \cite{shostack_threat_2014}. This section
+information disclosure, \acl{DOS}, and elevation of privileged \cite{shostack_threat_2014}. This section
 analyzes \ldawmsr using these categories. Threats that fit under multiple categories, such as tampering for the effect
@@
 
-To save space, the tables use shorthand for law enforcement (LE); law enforcement officer (LEO); law enforcement agency
-(LEA); Device ID (DID);
+To save space, the tables use shorthand for law enforcement (LE), law enforcement officer (LEO), law enforcement agency
+(LEA), Device ID (DID), Device Seal Key (DSK), \acf{DOS}
 
@@
     target device.
-  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage, extract its data, and destroy
-    it.
+  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage (to prevent telemetry), extract
+    its data, and destroy it (to prevent recovering stored data).
   \\ \hline
@@
   S6 & Manufacturer is spoofed to vendor
-  & Attackers pose as the manufacturer to inject malicious data into the DID-Seal database. Must be combined with other
+  & Attackers pose as the manufacturer to inject malicious data into the DID-DSK database. Must be combined with other
     attacks to be of use.
@@
 
-Tampering threats are listed in \mytab{table-tampering}.
+Tampering threats are listed in \mytab{table-tampering}. \ii{S2} relies on tampering similar to \ii{T1}.
 
@@
   T2 & Defensive user destroys the \ac{EA} hardware interface
-  & LE attempts at perform \ac{EA} will fail due to damaged hardware.
+  & LE attempts to perform \ac{EA} will fail due to damaged hardware.
   & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
@@
     liberties and harming transparency.
-  & ...
+  & The escrow agent could require some identifying information before unlocking, such as the LEA's public key digitally
+    signed by the vendor.
   \\ \hline
@@
     transparency.
-  & ...
+  & The same identifying information used in \ii{R3} would implicate the vendor (assuming the integrity of the vendor's
+    signing key).
 \threattableend
@@
 
-Information disclosure threats are listed in \mytab{table-disclosure}.
+Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed here, all
+attacks aimed at gaining unauthorized access to a device constitute information disclosure threats. Most spoofing
+attacks are aimed at disclosing information intended for another party.
 
 \threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
-  I1 & .
-  & .
-  & .
+  I1 & Attacker exposes the DSK, DID, the encrypted class key, or hashed token
+  & By design, none of the information stored on the phone is sensitive. The most consequential information that could
+    be stolen is the DID, which could be paired with other attacks. Disclosing the DID requires either disassembly and
+    time vaulting or advanced hardware inspection tooling.
+  & Sensitivity of this data has already been mitigated as much as possible.
   \\ \hline
-  I2 & .
-  & .
-  & .
+  I2 & Attacker intercepts unlock artifacts between LEA \ac{HSM} and device
+  & The attacker is able to unlock the target device alone. Must be combined with a spoofed DID to be of use to other
+    devices (see \ii{S2}). The could record the values to replay them after the phone has been returned.
+  & See \ii{S2}. If choosing not to destroy the device, to prevent replay attacks, a new token could be generated and
+    stored in the vendor HSM, with the new hash installed during the unlock process.
+  \\ \hline
+  I3 & Attacker steals the LEA private key
+  & The attacker is now able to spoof the LEA and decrypt returned unlock artifacts. Without the devices (which are
+    already in LE possession), the artifacts are not useful. See \ii{S4} and \ii{S5} for LEA spoofing attacks.
+  & The key is stored in an \ac{HSM}, which has extensive security measures. See \ii{S5} for a revocation and update
+    mitigation.
+  \\ \hline
+  I4 & Attacker steals the DID-DSK database
+  & This is the big prize, the largest source of centralized risk. With the database, attackers could directly
+    compromise (or more likely, sell access to) any mobile device using this system in their possession, after
+    disassembly, and after the time vault.
+  & \ii{R3} defines a mitigation that requires use of a vendor private key to create a digital signature as part of the
+    authentication process. If used, that key would have to be compromised as well for the database to be of use.
+    However, if the database is compromised, that key probably is too. Savage argues that \ac{HSM}s, though imperfect,
+    actually do provide very strong assurance. Responses could be rate-limited, and leaking an entire database once it's
+    in the \ac{HSM} could not go unnoticed. Getting the data in may be harder---see \ii{I5}.
+  \\ \hline
+  I5 & Attacker steals device unlock artifacts between manufacturing and vendor \ac{HSM}
+  & Intercepting the unlock artifacts at the source achieves the same effect as \ii{I4}, only the data set is not
+    complete. Instead of granting the ability to unlock any device, including a specific target, this is suited for
+    getting access down the line or for sale.
+  & The unlock artifacts generation process must be treated as sensitively as the \ac{HSM} itself. The artifacts should
+    be encrypted immediately and periodically transported to the corporate \ac{HSM} on physical media. This threat is
+    difficult to definitively mitigate.
 \threattableend
@@
 
-Denial of service threats are listed in \mytab{table-dos}.
+\Ac{DOS} threats are listed in \mytab{table-dos}. In addition to the threats listed, \ii{S4} can be used for \ac{DOS} by
+surreptitiously invalidating a LEA's identity key, and \ii{T1} and \ii{T2} are tampering attacks designed to deny LEAs
+from using the \ac{EA} mechanism.
 
 \threattablebegin{Denial of Service}{dos}{Service Denied}
-  D1 & .
-  & .
-  & .
-  \\ \hline
-  D2 & .
-  & .
-  & .
+  D1 & Unlock artifacts recovery
+  & \ii{S6} could be used to inject corrupt data to the DID-DSK database, potentially overwriting actual unlock artifact
+    data.
+  & Verify integrity of new data before entering it into the database, and disallow overwriting entirely.
 \threattableend
@@
 
-Elevation of privilege are listed in \mytab{table-privilege}.
+Elevation of privilege are listed in \mytab{table-privilege}. Spoofing to usurp a entity's privilege or using any attack
+to unlock a device without authorization can be seen as escalation of privilege.
 
 \threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
-  E1 & .
-  & .
-  & .
+  E1 & Foreign intelligence or organized crime recruits an insider
+  & The primary attacker gains inside access through coercion or bribery, making any other attack easier.
+  & Limit the employees, accounts, and computers involved in any element of the system. Perform background checks on
+    each employee and maintain detailed logs.
   \\ \hline
-  E2 & .
-  & .
-  & .
+  E2 & Unexposed hardware interface allows access to Secure Enclave
+  & The attacker gains access to Secure Enclave operations and data, undermining the device and achieving unlock without
+    following the \ac{EA} protocol.
+  & The time vaulting protocol and all communication over the interface must be thoroughly tested, and if possible,
+    formally proven to be logically sound. Analysis must include side channel attacks to affect the time vaulting or
+    leak internal data.
 \threattableend
+
+\section{Risk Analysis}
+
+How much risk does \ldawmsr introduce? The answer varies based on how many mitigations are implemented. If the device is
+required to be unlocked on LEA premises and destructive disassembly is chosen, several threats are significantly
+reduced. The irreducible risks are those posed by the DID-DSK database and the Secure Enclave itself.
 

### 2020-11-13 Thesis-Abstract.tex
Update abstract

@@
 Public debate has resumed over whether encryption systems should support alternative means of decryption intended for
-law enforcement use, called exceptional access (EA). Rather than a renege on a solemn promise made at the end of the
+law enforcement use, called exceptional access (EA). Rather than a renege on a resolute promise made at the end of the
 1990s ``crypto war,'' this represents a valid reassessment of optimal policy in light of changing circumstances.
@@
 technical community to stay engaged. Although the introduction of EA would inevitably introduce risk, the quality of the
-technical and regulatory approach can make a substantial difference. Furthermore, if one considers ham-fisted
-legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall
-risk.
+technical and regulatory approach can make a substantial difference. Furthermore, if one considers hard-line legislative
+action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall risk.
 
@@
 security. The emergence of strong encryption triggered the first crypto war, and its proliferation is causing the
-second. In response to both polarized and conciliatory voices, I will analyze strategies that do and do not work on
-wicked problems and promote an iterative approach to the case of encryption and EA. Along the way I will illustrate the
-components of the debate in argument maps and demonstrate the security risks with data flow diagrams, making clear
-distinctions for the separate problems of EA for data in motion and data at rest.
+second. In response to both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked
+problems and promote an iterative approach to the case of encryption and EA. Along the way I illustrate the components
+of the debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing
+on one EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''

### 2020-11-13 scripts/acronyms.sh
Shellcheck helper scripts

@@ -1,20 +1,22 @@
 #!/bin/bash
 
 set -eEuo pipefail
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
 # set -x
 
 INFILE="$BASE_DIR/notes/glossary.md"
 
 function lineof_with_offset() {
-  echo $(( $(grep -n "$1" $INFILE | cut -d: -f1) + $2 ))
+  echo $(( $(grep -n "$1" "$INFILE" | cut -d: -f1) + $2 ))
 }
 
 function cat_acronym_table() {
   START_LINE=$(lineof_with_offset "<!-- start acronyms" 4)
   END_LINE=$(lineof_with_offset "<!-- end acronyms" -2)
 
-  sed -n ${START_LINE},${END_LINE}p $INFILE
+  sed -n "$START_LINE","$END_LINE"p "$INFILE"
 }
 
 function to_lower() {
@@ -22,12 +24,13 @@ function to_lower() {
 }
 
 function convert_to_latex() {
-  TERM="$(echo $1 | cut -d\| -f 2 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
-  DEFN="$(echo $1 | cut -d\| -f 3 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
+  TERM="$(echo "$1" | cut -d\| -f 2 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
+  DEFN="$(echo "$1" | cut -d\| -f 3 | sed 's/[[:blank:]]*$//' | sed 's/^[[:blank:]]*//')"
 
+  # shellcheck disable=SC2028
   echo "\newabbreviation{$TERM}{$TERM}{$DEFN}"
 }
 
 while read -r ACR; do
     convert_to_latex "$ACR"
-done <<< $(cat_acronym_table)
+done <<< "$(cat_acronym_table)"

### 2020-11-13 scripts/argdown-preprocess.sh

@@ -2,15 +2,16 @@
 
 set -eEuo pipefail
 
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
-cd $BASE_DIR
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
+cd "$BASE_DIR"
 
 function get_ref_from_citation {
   grep "bibcite{$1}" build/Thesis.aux | cut -d'{' -f3 | cut -d'}' -f1
 }
 
 function get_citations_in_file {
-  grep -o "(cite:[a-zA-Z_0-9\\]*)" $1 | tr -d ")\\\\" | cut -d: -f2 | uniq
+  grep -o "(cite:[a-zA-Z_0-9\\]*)" "$1" | tr -d ")\\\\" | cut -d: -f2 | uniq
 }
 
 # There are two methods of resolving citations.
@@ -23,36 +24,37 @@ TEX=false
 function parse_infile_to_outfile {
   # Process citations
   # echo "IN: $1; OUT: $2"
-  cp $1 $2
+  cp "$1" "$2"
   # echo $(get_citations_in_file $1)
-  for CITE in $(get_citations_in_file $1); do
+  for CITE in $(get_citations_in_file "$1"); do
     local FULL_CITE="(cite:${CITE/_/\\\\_})"
     if ! $TEX; then
-      local REF="$(get_ref_from_citation $CITE)"
+      local REF
+      REF="$(get_ref_from_citation "$CITE")"
       if [[ "" == "$REF" ]]; then
-          printf "${RED}Bibtex citation for $CITE not found$NO_COLOR\n"
+          printf "%sBibtex citation for %s not found%s\n" "$RED" "$CITE" "$NO_COLOR"
           REF="E\\\\_AGAIN"
       fi
       # echo "CITE: $CITE; REF: $REF"
-      sed -i "s/$FULL_CITE/\\\\[$REF\\\\]/g" $2
+      sed -i "s/$FULL_CITE/\\\\[$REF\\\\]/g" "$2"
     else
       TEX_CITE="\\\\\\\\cite\\\\{${CITE/_/\\\\_}\\\\}"
-      sed -i "s/$FULL_CITE/$TEX_CITE/g" $2
+      sed -i "s/$FULL_CITE/$TEX_CITE/g" "$2"
     fi
   done
 
   # Remove preview configs
-  if grep -q "Preview configs start" $2 && grep -q "Preview configs end" $2; then
[…]

### 2020-11-13 scripts/bib-gen.sh

@@ -1,14 +1,17 @@
 #!/bin/bash
 
 set -eEuo pipefail
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
-cd $READING_DIR
+
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
+cd "$READING_DIR"
 
 OUTFILE="../Thesis.bib"
 
 function check_for_duplicate_refname() {
-  local REFNAME="$(echo "$REF" | grep ^@ | cut -d{ -f2 | cut -d, -f1)"
-  if cat $OUTFILE | grep -q "$REFNAME"; then
+  local REFNAME
+  REFNAME="$(echo "$REF" | grep ^@ | cut -d\{ -f2 | cut -d, -f1)"
+  if grep -q "$REFNAME" "$OUTFILE"; then
     echo "Duplicate reference name: $REFNAME"
     exit 1
   fi
@@ -22,12 +25,12 @@ function sanitize_language_field() {
   fi
 }
 
-echo "% This is an auto-generated bibliography file built by $(basename $0)" > $OUTFILE
+echo "% This is an auto-generated bibliography file built by $(basename "$0")" > $OUTFILE
 echo >> $OUTFILE
 
 for FILE in $(find . -type f -name "*.md"); do
-  ROWS=( $(grep -n '```' $FILE | cut -d: -f1) )
-  REF="$(sed -n $(( ${ROWS[0]} + 1 )),$(( ${ROWS[1]} - 1 ))p $FILE)"
+  ROWS=( $(grep -n '```' "$FILE" | cut -d: -f1) )
+  REF="$(sed -n $(( ROWS[0] + 1 )),$(( ROWS[1] - 1 ))p "$FILE")"
   check_for_duplicate_refname
   sanitize_language_field
   echo "$REF" >> $OUTFILE

### 2020-11-13 scripts/count-citations.sh

@@ -1,6 +1,7 @@
 #!/bin/bash
 
-cd $(git rev-parse --show-toplevel)
+cd "$(git rev-parse --show-toplevel)" || exit
+# shellcheck disable=SC1091
 source ./scripts/sourceme.sh
 # set -x
 
@@ -17,16 +18,17 @@ function grep_term() {
 }
 
 function print_number_in_tex() {
-    echo $((grep_citation $1 || grep_term $1) | wc -l)
+    (grep_citation "$1" || grep_term "$1") | wc -l
 }
 
 function print_number_in_argdown() {
-    echo $(grep "(cite:${1/_/\\\\_})" arguments/*.ad | wc -l)
+    # shellcheck disable=SC2126
+    grep "(cite:${1/_/\\\\_})" arguments/*.ad | wc -l
 }
 
 function print_colored_number() {
     (( $1 > 0 )) && COLOR=$LIGHT_GREEN || COLOR=$YELLOW
-    printf "$COLOR$1$NO_COLOR"
+    printf "$COLOR%s$NO_COLOR" "$1"
 }
 
 function pluralize() {
@@ -34,21 +36,21 @@ function pluralize() {
 }
 
 function print_count() {
-    COUNT_TEX=$(print_number_in_tex $1)
-    COUNT_ARGDOWN=$(print_number_in_argdown $1)
-    printf "$LIGHT_CYAN$1$NO_COLOR is cited $(print_colored_number $COUNT_TEX) time$(pluralize $COUNT_TEX) in writing"
-    printf " and $(print_colored_number $COUNT_ARGDOWN) time$(pluralize $COUNT_ARGDOWN) in arguments.\n"
+    COUNT_TEX=$(print_number_in_tex "$1")
+    COUNT_ARGDOWN=$(print_number_in_argdown "$1")
+    printf "$LIGHT_CYAN%s$NO_COLOR is cited $(print_colored_number "$COUNT_TEX") time$(pluralize "$COUNT_TEX") in writing" "$1"
+    echo   " and $(print_colored_number "$COUNT_ARGDOWN") time$(pluralize "$COUNT_ARGDOWN") in arguments."
 }
 
 if (( 2 <= $# )) && [[ "-c" == "$1" ]]; then
     # Print count
-    print_count $2
+    print_count "$2"
 elif (( 2 <= $# )) && [[ "-n" == "$1" ]]; then
     # Print number only
[…]

### 2020-11-13 scripts/install-deps.sh

@@ -30,7 +30,7 @@ if which snap &>/dev/null; then
     snap install drawio
 else
     URL=$(curl -sL https://github.com/jgraph/drawio-desktop/releases/latest | grep "\.deb" | head -1  | cut -d\" -f2)
-    wget -q $URL
+    wget -q "$URL"
     sudo dpkg -i ./*.deb || sudo apt-get -fy install
     rm ./*.deb
 fi

### 2020-11-13 scripts/new-source.sh

@@ -1,6 +1,7 @@
 #!/bin/bash
 
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
 
 ##################################################################### SCRIPT SETUP
 USAGE="
@@ -11,37 +12,37 @@ SOURCE - The DOI or URL of the new material
 "
 
 ORIG_DIR=$(pwd)
-cd $READING_DIR
+cd "$READING_DIR" || exit
 function cleanup() {
-    if [[ 2 == $? ]]; then
-        printf "$USAGE"
+    if (( 2 == $? )); then
+        echo "$USAGE"
     fi
 }
 trap cleanup EXIT
 
 ##################################################################### DEPENDENCIES
-[[ 0 < $(which docker | wc -l) ]] || exitprint 1 "Docker must be installed."
+which docker &>/dev/null || exitprint 1 "Docker must be installed."
 if [[ ! $(which getdoi) ]]; then
     echo "Installing scholarref tools..."
     git clone git://src.adamsgaard.dk/scholarref ../../scholarref &>/dev/null
-    pushd ../../scholarref &>/dev/null
+    pushd ../../scholarref &>/dev/null || exit
     sudo make install >/dev/null
-    popd &>/dev/null
+    popd &>/dev/null || exit
 fi
 
 ##################################################################### ARGUMENTS
 (( 2 <= $# )) || exitprint 2 "Incorrect args"
 case "$1" in
-    /*) SUBDIR=$(echo $1 | sed 's:/*$::') ;;
-    *)  SUBDIR=$ORIG_DIR/$(echo $1 | sed 's:/*$::') ;;
+    /*) SUBDIR=$(echo "$1" | sed 's:/*$::') ;;
+    *)  SUBDIR=$ORIG_DIR/$(echo "$1" | sed 's:/*$::') ;;
 esac
-[ -d $SUBDIR ] || exitprint 2 "SUBDIR=$SUBDIR must exist"
-SOURCE="${@:2}"
+[ -d "$SUBDIR" ] || exitprint 2 "SUBDIR=$SUBDIR must exist"
+SOURCE="${*:2}"
 
[…]

### 2020-11-13 scripts/show-source.sh

@@ -1,35 +1,36 @@
 #!/bin/bash
 
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
 # set -x
 ORIG_DIR=$(pwd)
-cd $READING_DIR
+cd "$READING_DIR" || exit
 
 if (( 2 <= $# )) && [[ "-a" == "$1" ]]; then
     # Show all
-    for FILE in $(eval "grep -riIsc \"${@:2}\" | grep -v :0 | sort -t: -n -k2 -r"); do
-        readlink -f $FILE
+    for FILE in $(eval "grep -riIsc \"${*:2}\" | grep -v :0 | sort -t: -n -k2 -r"); do
+        readlink -f "$FILE"
     done
     exit 0
-elif [[ 1 == $# && -f $ORIG_DIR/${1%%:[0-9]*} ]]; then
+elif (( 1 == $# )) && [[ -f $ORIG_DIR/${1%%:[0-9]*} ]]; then
     # Show a file based on relative path
     FILE=$ORIG_DIR/${1%%:[0-9]*}
-elif [[ 1 == $# && -f ${1%%:[0-9]*} ]]; then
+elif (( 1 == $# )) && [[ -f ${1%%:[0-9]*} ]]; then
     # Show a file based on absolute path
     FILE=${1%%:[0-9]*}
-elif [[ 1 == $# ]] && $(grep -riIsq "@*{$1,"); then
+elif (( 1 == $# )) && grep -riIsq "@*{$1,"; then
     # Show a file based on bibtex ref
     FILE="$(grep -riIsl "@*{$1," . | head -1)"
     FILE="$(readlink -f "$FILE")"
 else
     # Show the file with the most string matches
-    FILE_W_NUM=$(eval "grep -riIsc \"$@\" . | sort -t: -n -k2 | tail -1")
-    NUM="$(echo $FILE_W_NUM | cut -d: -f2)"
-    if (( $NUM == 0 )); then
+    FILE_W_NUM=$(eval "grep -riIsc \"$*\" . | sort -t: -n -k2 | tail -1")
+    NUM="$(echo "$FILE_W_NUM" | cut -d: -f2)"
[…]

### 2020-11-13 scripts/sourceme.sh

@@ -1,6 +1,8 @@
+# shellcheck disable=SC2148
 # Source this file
 
-export BASE_DIR="$(git rev-parse --show-toplevel)"
+BASE_DIR="$(git rev-parse --show-toplevel)"
+export BASE_DIR
 export READING_DIR="$BASE_DIR/reading"
 export SCRIPTS_DIR="$BASE_DIR/scripts"
 
@@ -13,30 +15,30 @@ export NO_COLOR='\e[0m'
 
 function exitprint() {
     echo -e "${@:2}"
-    exit $1
+    exit "$1"
 }
 export -f exitprint
 
-alias new="$SCRIPTS_DIR/new-source.sh"
-alias show="$SCRIPTS_DIR/show-source.sh"
-alias term="$SCRIPTS_DIR/term.sh"
-alias gen="$SCRIPTS_DIR/bib-gen.sh"
-alias cite="$SCRIPTS_DIR/count-citations.sh"
-alias todo="(cd $BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
-alias fixme="(cd $BASE_DIR && grep -nE 'FIXME' *.tex)"
-alias pdf="xdg-open $BASE_DIR/build/Thesis.pdf"
-alias gvsu="xdg-open $BASE_DIR/build/Thesis-GVSU.pdf"
+alias new="\$SCRIPTS_DIR/new-source.sh"
+alias show="\$SCRIPTS_DIR/show-source.sh"
+alias term="\$SCRIPTS_DIR/term.sh"
+alias gen="\$SCRIPTS_DIR/bib-gen.sh"
+alias cite="\$SCRIPTS_DIR/count-citations.sh"
+alias todo="(cd \$BASE_DIR && grep -nE '(TODO|FIXME)' *.tex)"
+alias fixme="(cd \$BASE_DIR && grep -nE 'FIXME' *.tex)"
+alias pdf="xdg-open \$BASE_DIR/build/Thesis.pdf"
+alias gvsu="xdg-open \$BASE_DIR/build/Thesis-GVSU.pdf"
 alias start="(codium .) && (xdg-open build/Thesis.pdf) && make live"
 
 function arg {
-    xdg-open $BASE_DIR/arguments/build/$1.pdf 2>/dev/null \
-        || xdg-open $BASE_DIR/arguments/build/comprehensive.$1.pdf 2>/dev/null
+    xdg-open "$BASE_DIR"/arguments/build/"$1".pdf 2>/dev/null \
[…]

### 2020-11-13 scripts/term.sh

@@ -2,8 +2,9 @@
 
 set -eEuo pipefail
 
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
-cd $BASE_DIR
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
+cd "$BASE_DIR"
 
 GLOS=Thesis-Glossary.tex
 
@@ -12,9 +13,9 @@ if grep -iEoq "newabbreviation{$1}" $GLOS; then
 else
     WHOLE="$(tr -d '\n' < $GLOS | grep -iEo "newterm{$1}{[^}]*}{[^}]*}" | tr -d '}')"
 fi
-SHORT="$(echo $WHOLE | cut -d{ -f2)"
-LONG="$(echo $WHOLE | cut -d{ -f3)"
-DEF="$(echo $WHOLE | cut -d{ -f4)"
+SHORT="$(echo "$WHOLE" | cut -d\{ -f2)"
+LONG="$(echo "$WHOLE" | cut -d\{ -f3)"
+DEF="$(echo "$WHOLE" | cut -d\{ -f4)"
 
 echo -e "$LIGHT_CYAN$LONG$NO_COLOR ($SHORT)"
 echo -e "$LIGHT_GREEN$DEF$NO_COLOR"

### 2020-11-13 scripts/update-acronyms.sh

@@ -1,23 +1,25 @@
 #!/bin/bash
 
 set -eEuo pipefail
-source $(git rev-parse --show-toplevel)/scripts/sourceme.sh
+
+# shellcheck disable=SC1090
+source "$(git rev-parse --show-toplevel)"/scripts/sourceme.sh
 #set -x
 
 TEX_FILE="$BASE_DIR/Thesis-Glossary.tex"
 
 function lineof_with_offset() {
-  echo $(( $(grep -n "$1" $TEX_FILE | cut -d: -f1) + $2 ))
+  echo $(( $(grep -n "$1" "$TEX_FILE" | cut -d: -f1) + $2 ))
 }
 
 START_LINE=$(lineof_with_offset "% Start acronyms" 0)
 END_LINE=$(lineof_with_offset "% End acronyms" 0)
-EOF_LINE=$(wc -l $TEX_FILE | awk '{print $1}')
+EOF_LINE=$(wc -l "$TEX_FILE" | awk '{print $1}')
 
-BEGINNING="$(sed -n 1,${START_LINE}p         $TEX_FILE)"
-END="$(      sed -n ${END_LINE},${EOF_LINE}p $TEX_FILE)"
+BEGINNING="$(sed -n 1,"$START_LINE"p         "$TEX_FILE")"
+END="$(      sed -n "$END_LINE","$EOF_LINE"p "$TEX_FILE")"
 
-echo "$BEGINNING"         >  $TEX_FILE
-$SCRIPTS_DIR/acronyms.sh  >> $TEX_FILE
-echo "$END"               >> $TEX_FILE
+echo "$BEGINNING"           >  "$TEX_FILE"
+"$SCRIPTS_DIR"/acronyms.sh  >> "$TEX_FILE"
+echo "$END"                 >> "$TEX_FILE"
 

### 2020-11-13 scripts/wait-till-pdf-updated.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
 
-PDF=$(git rev-parse --show-toplevel)/build/Thesis.pdf
+PDF="$(git rev-parse --show-toplevel)"/build/Thesis.pdf
 
 function getTs() {
-    stat --format=%Y $PDF || echo 0
+    stat --format=%Y "$PDF" || echo 0
 }
 
 START_TS=$(getTs)

### 2020-11-14 Thesis-Chapter5-ThreatModel.tex
Finish risk analysis

@@
 access and tamper with and steal information. They have world-class expertise and computing power, which they will apply
-to both individual devices and backend systems.}
+to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence agencies, as
+opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S., which are assumed
+to operate within---though push the boundaries of---the legal framework.)}
 
@@
 
-% What are we building?: proposal DFDs
-% What can go wrong?: elicitation of threats
-
 \section{Basic Data at Rest}
+\label{sec-basic-dar}
 
@@
   & Attackers pose as an LEA in order to spoof future access requests.
-  & Vendor key exchange process involves thorough in-person verification.
+  & Vendor key exchange process involves thorough in-person verification. Assume only a handful of digital forensics
+    laboratories nationwide are authorized to participate.
   \\ \hline
@@
   & Vendor independently verifies all legal documents. Vendor supports a key revocation and update process that involves
-    the same level of in-person verification.
+    the same level of in-person verification. Assume only a handful of digital forensics laboratories nationwide are
+    authorized to participate.
   \\ \hline
@@
   & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
-  & Randomize DID's to make them unguessable, meaning a target device must be in possession in order for the attacker to
+  & Randomize DIDs to make them unguessable, meaning a target device must be in possession in order for the attacker to
     know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
-    device S/N; the \ac{HSM} would hash the value and verify a match before returning the unlock artifacts.
+    device S/N; the \ac{HSM} would append the S/N to the DID and hash the result, verifying a match with the same stored
+    hash before returning the unlock artifacts.
   \\ \hline
@@
   & This is close to being an out of scope threat under the ``rule of law'' exception. However, one could design the
-    system to require requests to be recorded on a public ledger. This opens up several new difficulties.
+    system to require requests and responses to be recorded on a public ledger. This creates several new difficulties.
   \\ \hline
@@
     liberties and harming transparency.
-  & The escrow agent could require some identifying information before unlocking, such as the LEA's public key digitally
-    signed by the vendor.
+  & At a high level, the system should require a long paper trail before access can be granted. At a low level, the
+    escrow agent could require and store some identifying information before unlocking, such as the LEA's public key
+    digitally signed by the vendor, which provides non-repudiation assuming independent access to the device. See also
+    the mitigation in \ii{R2}.
   \\ \hline
@@
   & The device is undeniably unlocked, but the vendor claims they did not provide the unlock artifacts, allowing the
-    vendor to cover up leaks or unsavory policy decisions by suggesting the device was instead hacked. Harms
+    vendor to cover up leaks or distasteful policy decisions by suggesting the device was instead hacked. Harms
     transparency.
-  & The same identifying information used in \ii{R3} would implicate the vendor (assuming the integrity of the vendor's
-    signing key).
+  & Assuming independent access to the device, the hardware transparency mechanism should confirm whether the \ac{EA}
+    process was used, and the same identifying information used in \ii{R3} would implicate the vendor. See also the
+    mitigation in \ii{R2}.
 \threattableend
@@
   & The attacker is able to unlock the target device alone. Must be combined with a spoofed DID to be of use to other
-    devices (see \ii{S2}). The could record the values to replay them after the phone has been returned.
+    devices (see \ii{S2}). The attacker could record the values to replay them after the phone has been returned.
   & See \ii{S2}. If choosing not to destroy the device, to prevent replay attacks, a new token could be generated and
@@
   & This is the big prize, the largest source of centralized risk. With the database, attackers could directly
-    compromise (or more likely, sell access to) any mobile device using this system in their possession, after
-    disassembly, and after the time vault.
+    compromise any mobile device using this system in their possession, after disassembly, and after the time vault.
+    They may also try to sell access to the data.
   & \ii{R3} defines a mitigation that requires use of a vendor private key to create a digital signature as part of the
     authentication process. If used, that key would have to be compromised as well for the database to be of use.
-    However, if the database is compromised, that key probably is too. Savage argues that \ac{HSM}s, though imperfect,
+    However, if the database is compromised, that key probably is too. Savage argues that \acp{HSM}, though imperfect,
     actually do provide very strong assurance. Responses could be rate-limited, and leaking an entire database once it's
@@
     difficult to definitively mitigate.
+  \\ \hline
+  I6 & Attacker steals device \ac{plaintext} from LEA after unlock process
+  & Assuming the attacker doesn't care that the LEA gets access to plaintext, they could ``help'' the LE investigation
+    in order to access the data later either when it comes out as public evidence or through compromising the LEA
+    itself. This may be useful to authoritarian regimes, foreign intelligence, and organized crime.
+  & LEAs must have strong information security of their own. Disclosure of evidence is a natural side effect of
+    prosecution, so LEAs must practice discretion in choosing which devices to unlock in case mere information discovery
+    poses a great risk in itself.
 \threattableend
@@
 \threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
-  E1 & Foreign intelligence or organized crime recruits an insider
+  E1 & Foreign intelligence or organized criminals recruit an insider
   & The primary attacker gains inside access through coercion or bribery, making any other attack easier.
@@
     leak internal data.
+  \\ \hline
+  E3 & An \ac{HSM} vulnerability allows access to protected data
+  & This is a technical route to achieving \ii{I4}, disclosure of the DID-DSK database.
+  & The \ac{HSM} should be on an air-gapped network disconnected to the internet, physically guarded, and regularly
+    patched. There is no defense against an unknown vulnerability, so security ultimately comes down to incident
+    detection and response.
 \threattableend
@@
 
-How much risk does \ldawmsr introduce? The answer varies based on how many mitigations are implemented. If the device is
-required to be unlocked on LEA premises and destructive disassembly is chosen, several threats are significantly
-reduced. The irreducible risks are those posed by the DID-DSK database and the Secure Enclave itself.
+How much risk does \ldawmsr introduce? The answer lies in a range depending on how many mitigations are implemented.
+Some threats introduce inherently low risk due to their difficult nature and low pay off. Others are higher risk but can
+be reasonably mitigated. Some threats specifically target transparency, and are difficult to address without trust.
+Finally, there are some risks that are irreducible. Each of classes are discussed in turn.
 
-\section{Response to Frameworks}
+\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, and \ii{E1} \ii{E2} are each low risk either because they
+fail to produce useful results, they can be well mitigated, or they are a threat that is already present in
+state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and mitigate the
+risk posed by these threats as part of the design process, but they do not represent fatal flaws to a design. Except in
+combination with other threats, they are not worthy more concern here.
 
-% Or just, relationship to my goals?
+\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are each higher risk threats to the
+security, trustworthiness, and effectiveness of the proposal, but can be addressed through a set of related mitigations:
+require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device to be
+physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities the
+lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly is a technical measure that entirely
+precludes certain attacks; requiring the device to be on LEA premises is a procedural measure that prevents attackers
+from manipulating LE into achieving their own ends; limiting the number of LE facilities that can participate prevents
+LE impersonation attacks. Together, they make the proposal considerably harder to attack.
 
-% How does it answer (or not) committee_decrypting_2018's framework questions? What needs improvement?
+\ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
+legal process and cooperation of several parties. But what if the legal process involves secret courts, secret
+investigations, and secrecy orders on the vendor \cite{shamsi_2011}? Destructive disassembly or the weaker hardware
+transparency mechanism provide non-repudiation that the device was unlocked, if and only if one has physical access to
+the device. Technically enforced transparency into device access requests and responses would grant non-repudiation at a
+broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR} \ac{EA} proposals do
+\cite{phan_key_2017} \cite{servan_schreiber_jje_2020}. Unfortunately, building this feature into a proposal makes it
+potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.
 
-% Reflection on how much progress this represents
+Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, each representing in slightly different ways the
+disclosure of unlock artifact data, the DID-DSK database. This is the truly difficult challenge---centralizing so much
+sensitive data makes it an extremely high value target for foreign intelligence and cyber criminals. The vendor is in a
+position where they need to protect this database from the most advanced attackers in the world while regularly updating
+and accessing it. \Acp{HSM} and strong security policies would make that a difficult task, but not impossible, and one
+has to assume that the data would leak eventually. What would leaked unlock artifacts mean? It depends who steals it.
+Criminals would use it for personal and financial gain; foreign intelligence and authoritarian regimes would use it to
+spy on their enemies. In the worst-case scenario, anyone with some specialized tools and a willingness to break the law
+could get into any device under the scheme, provided they destructively disassemble it and wait through the time vault.
+
+It is important to give \ldawmsr (and similar \ac{DAR} \ac{EA} proposals) its due. The worst case scenario still
+precludes mass surveillance and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
+proposals in terms of transparency and protection of civil liberties.
+
+\section{Achievement of Goals}
+
+% Consider goals, threat actors, and loop-back to out of scope items, because they aaaaactually do matter.
+
+% Economics: vastly increases costs and barrier to entry for vendors.
 

### 2020-11-15 Thesis-Chapter5-ThreatModel.tex
Finish chapter 5

@@
 ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}.
-Lacking a concise name, this proposal will be referred to by its acronym, \ldawmsR. Note that the creators of the
-proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to emphasize the role of legal
-process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an \ac{EA} proposal, but use the
-term entirely in the spirit of their paper.
+Lacking a concise name, this proposal will be referred to by its acronym, \ldawmsR. This chapter assumes that the
+proposal is accompanied by some legal framework regulating its application and use, though it is not the focus of this
+analysis, and nothing is assumed about it in particular. Note that the creators of the proposal intentionally chose the
+phrase ``lawful access'' over ``\acl{EA}'' in order to emphasize the role of legal process and dispel the negative
+stigma that accompanies \ac{EA}. I refer to \ldawmsr as an \ac{EA} proposal, but use the term entirely in the spirit of
+their paper.
 
@@
     \small
+    \singlespacing
     \begin{longtable}{ lp{3cm}p{5.2cm}p{6cm} }
@@
 
+This chapter began with a list of goals by which \ac{DAR} \ac{EA} proposals can be measured, including security,
+protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section I measure
+\ldawmsr against each of these goals.
+
 % Consider goals, threat actors, and loop-back to out of scope items, because they aaaaactually do matter.
 
-% Economics: vastly increases costs and barrier to entry for vendors.
+\paragraph*{Security} Security is the main topic of this chapter, and has already been discussed thoroughly. \ldawmsr
+introduces significant risk in the DID-DSK database itself, but does a good job in many other respects, especially if
+the recommended mitigations are implemented. It goes from very hard to attack, when everything is working as intended,
+to simple to attack, when the database is leaked. However, even when it becomes trivial to attack, the impact is limited
+to devices in physical possession and cannot be hidden from the device user.
 
-%%%%%%%%%%%%
-% Old notes for reference
-% -----------------------
-% Chapter 4, threat models
-% (how to be specific enough to be useful but generic enough to apply to many systems? Focus on the EA type (DAR v DIM),
-%   then apply to specific systems)
-% DAR, DIM threat models (make sure to cover abelson_2015)
-% Basic proposals
-%  --> mappings onto threat models
-% At work: 3 example systems analyzed with this threat model
-% 1. IM (E2E IM too?) -- in motion
-% 2. Cloud storage -- at rest
-% 3. Mobile phone -- at rest, but way different scenario than cloud
-%  --> include Ozzie's CLEAR, since it was higher profile
-%%%%%%%%%%%%
+\paragraph*{Protection of civil liberties} This goal was defined as defending against mass surveillance and abuse by
+authoritarian regimes. This proposal does an excellent job of protecting against mass surveillance due to its
+physicality. Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian
+regimes. Prior to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign
+countries. After a leak, these regimes could get into any device they want. If one lived in a country that was willing
+to take and destroy their devices to read their contents without legal due process, this would be an unwelcome
+technology.
+
+\paragraph*{Transparency} \ldawmsr excels at transparency on the device, but fails at transparency to the public. Due to
+destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed using this method
+is irrefutable. However, this is only possible to confirm if the device is available. If it is stolen and unlocked by
+criminals or is lawfully accessed under secretive conditions, both the user and the public have no way of knowing. This
+leaves open the possibility for government abuse---though again with the caveat that physical possession requirements
+mean it cannot be abused without a trace.
+
+\paragraph*{Law enforcement utility} This proposal involves a few tradeoffs between security and law enforcement
+utility. The physical possession, time vault, and limited number of authorized law enforcement laboratory requirements
+each insert a barrier to usability. The physical possession requirement is far too important for security and civil
+liberties to sacrifice for utility. Furthermore, law enforcement, and search and seizure law generally, is geared
+towards physical evidence already. The time vault prevents against time-sensitive threats, though destructive
+disassembly (not part of the original proposal) may be enough to cover these cases as well, so the time vault could be
+decreased without significant impact. Limiting the number of parties able to use the system mitigates spoofing attacks;
+local law enforcement would have to cooperate with one of these parties, such as a regional \ac{FBI} office. (This
+really just shifts the spoofing threat up a level, though the shift is appropriate.) In the end, law enforcement still
+has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end; the onus is on the government to fund
+the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
+
+\paragraph*{Economic impact} Economic costs to the vendor fall into three categories---implementing the design, running
+the system, and absorbing lost business. Even once a proposal like \ldawmsr was refined enough to be implemented, the
+vendor would need to update hardware designs and create the software systems to get it started. Once in use, the vendor
+would face considerable costs in running it securely, considering the supreme importance of transporting and storing the
+unlock artifacts safely. Finally, the vendor would certainly lose privacy-conscious customers to vendors not under the
+system, whether small businesses, foreign competitors, or resellers offering pre-\ac{EA} devices. The scale of these
+costs relative to device manufacturer's naturally large size is unclear. Costs to the government include the \ac{FBI}
+operation and a necessary scale-up in training and resources available to state and local law enforcement. Government
+reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a \ac{DAR}
+\ac{EA} regulation as well \cite{edwards_hr4922_1994}.

### 2020-11-15 Thesis-Chapter3-PolicyMaking.tex
Clean up arguments chapter notes
Add OODA process clear out old DFDs

@@
 
+% FIXME: include this idea somewhere in this section.
+% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
+% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
+% decrypted! So must be very careful to avoid solution attempts that leave long scars.
+
 \subsection{Lessons}
@@
     \includegraphics[width=\linewidth]{dfds/build/OODA-loop.png}
-    \caption[The OODA Loop for Wicked Problems]{The OODA Loop for Wicked Problems}
+    \caption{The OODA Loop for Wicked Problems}
     \label{fig-policy-ooda-loop}
@@
 wicked. The tame sub-problems can be addressed with the \ac{classical-method}, and the wicked sub-problems, by now
-restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.
+restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}. This process is demonstrated
+in \myfig{fig-policy-ooda-process}.
+
+\begin{figure}[h]
+  \centering\CaptionFontSize
+  \includegraphics[width=\linewidth]{dfds/build/OODA-policy-process.png}
+  \caption{Using the \ac{OODAloop} to Tackle Wicked Problems}
+  \label{fig-policy-ooda-process}
+\end{figure}
 

### 2020-11-17 Thesis-Chapter4-Arguments.tex
Integrate koepke_2020 into arguments

@@
   \centering\CaptionFontSize
-  \includegraphics[width=0.9\linewidth]{arguments/build/comprehensive.goldenage.pdf}
+  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
   \caption{A ``Golden Age for Surveillance'' Argument Map}
@@
 as a positive side effect. The government would source these vulnerabilities from a combination of the public domain,
-the commercial exploit market, and a central ``Vulnerability Lab'' \cite{bellovin_lawful_2013}. However, as the \$2
-million price on the exploit market for an Apple iOS jailbreak bug shows, security really is improving
-\cite{goodin_zeroday_2019}. Law enforcement would still find itself without access, and would need to develop costly
-exploits in the lab. Mandatory vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the
-\ac{FBI}'s historical abuses of power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has
-passed \cite{keller_internet_2019} or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine
-that the government would actually give up such expensive vulnerabilities so easily when they are the pillar of their
-access strategy.
+the commercial exploit market, and a central ``Vulnerability Lab'' \cite{bellovin_lawful_2013}.
+
+The availability and cost of exploits ebbs and flows. On one hand, January 2019 sported a \$2 million price on the
+exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, an October 2020 report
+shows that law enforcement agencies can extract data from nearly every mobile phone on the market for less than \$2000 a
+unit \cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law
+enforcement would face periods where it struggles for access, and would need to develop costly exploits in the lab.
+Mandatory vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical
+abuses of power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed
+\cite{keller_internet_2019} or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the
+government would actually give up expensive vulnerabilities easily when they are the pillar of their access strategy.
 
@@
 
+% FIXME: add note about how device security actually totally sucks
+
 Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the

### 2020-11-17 Thesis-Chapter6-PathsAhead.tex
Separate strategies and mapping chapters
Finish risk analysis
Clean up conclusion notes
Integrate koepke_2020 into arguments

--- 2020-03-15 Thesis-Chapter5-PathsAhead.tex
@@
 
-In this chapter, describe actions to take in discussion, regulation, and research.
+% We have a wicked problem
+% I propose we handle it a certain way
+% Part of this is cooperative debate, which I tried to capture in argument maps
+% Another element is focusing on specific proposals, so we dove into LDAWMSR
 
-\section{Legislative Priorities}
+%% Prof. Walsh:
+%% The dynamic regulatory environment and competing governmental branches at the executive, legislative, and judicial
+%% level will continue to shape the...
 
-\begin{itemize}
-    \item Reject EA until technical capabilities are ready \cite{varia_2018}
-    \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
-    \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
-    \item Fund the research \cite{varia_2018}
-\end{itemize}
+% What can we conclude?
+%   - It may not feel like it, but (meaningful) debate and discussion counts as progress for a wicked problem
+%   - Neither side has the whole picture, but the "no EA" arguments are currently stronger
+%   - A threat model including malicious users and threatening governments makes EA make more sense
+%   - Device escrow with secure hardware represents true progress towards acceptably secure EA, so hope is not lost
+% Where do we go from here?
+%   - This was an attempt an "OO" iteration in the OODA; we need more, with involvement all around
+%   - committee_decrypting_2018 presented a framework to use to judge proposals. I have a simple version of it.
+%     here are where we are weak:
+%       - X
+%       - Y
+%       - Z
+%   - Government:
+%       - Take the low hanging fruit first, please
+%       - Fund research, don't mislead, don't be premature
+%           - If you're going to compare it to going to the moon, fund us like it!
+%       - Take on weak area XYZ
+%   - Tech:
+%       - Do research, provide clear security advice, be vigilant
+%       - Take on weak area XYZ
+%   - Either:
+%       - Know that there is no stable solution, balance rules and principles, serve your neighbor
+% see varia_2018 (maybe a good quote?)
 
-\section{Technical Research}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+% OLD NOTES for reference
+%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{itemize}
-    \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
-    \item Do the research
-    \item Focus on principles \cite{levy_robinson_2018}
-    \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
-            \cite{abelson_2015}), and clearly define their interfaces with principles
-            \cite{matyas_incommensurability_2018}
-\end{itemize}
+% In this chapter, describe actions to take in discussion, regulation, and research.
+
+% \section{Legislative Priorities}
+
+% \begin{itemize}
+%     \item Reject EA until technical capabilities are ready \cite{varia_2018}
+%     \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
+%     \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
+%         (e.g., ECPA became outdated because was too tied to specific technology \cite{shamsi_2011})
+%     \item Fund the research \cite{varia_2018} (like \cite{goss_hr2616_1999})
+% \end{itemize}
+
+% \section{Technical Research}
+
+% \begin{itemize}
+%     \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
+%     \item Do the research
+%     \item Focus on principles \cite{levy_robinson_2018}
+%     \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
+%             \cite{abelson_2015}), and clearly define their interfaces with principles
+%             \cite{matyas_incommensurability_2018}
+% \end{itemize}

### 2020-11-17 arguments/argdown.config.js

@@ -47,6 +47,7 @@ module.exports = {
           "Going Dark Conclusions",
           "Going Dark Argument",
           "Going Dark Non Core",
+          "Mobile Device Security Sucks",
         ],
       },
     },
@@ -63,6 +64,7 @@ module.exports = {
           "Going Dark Conclusions",
           "Golden Age Argument",
           "Golden Age Non Core",
+          "Mobile Device Security Sucks",
         ],
       },
       dot: {
@@ -128,6 +130,7 @@ module.exports = {
           "DAR EA Classes",
           "DIM EA Classes",
           "Arguments for EA Types",
+          "Mobile Device Security Sucks",
         ],
         excludeStatements: [
           "Exceptional Access",

### 2020-11-17 Thesis-Chapter4-Arguments.tex
Add argument that despite hackability, device EA is worth studying

@@
 adversaries) focusing on \ac{DAR} in mobile phones using device key escrow requiring physical access is the most
-tractable technical case to consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed
+tractable technical niche to consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed
 with a dose of lawful hacking is best suited for advanced opponents, and device key escrow represents actual progress as
-a path towards lower-risk \ac{EA}.
+a path towards lower-risk \ac{EA} for typical cases.
 
-Device key escrow still has a long way to go. There are few proposals, what few there are are high level, and questions
-of scaling and administration have not been addressed. However, as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s
-Levy and Robinson have pointed out, focusing on specific proposals has the opportunity to identify mutually agreeable
-solutions---or at least refine the problem understanding enough that the next round of debate can be better informed.
+A strong argument against device key escrow is that device encryption, particularly in mobile phones, may actually pose
+the smallest impediment to investigation of any encryption technology. The previously cited report into law enforcement
+agencies' ability to access mobile phones clearly demonstrates the current state of affairs: device manufacturers are
+losing the exploit arms race, private companies are commodifying the \ac{lawful-hacking} approach for agencies large and
+small, and the lack of regulatory strategy has led to pervasive data collection with little oversight
+\cite{koepke_2020}. This fact does not undermine the worthiness of research into device-oriented \ac{EA}. First of all,
+it underscores the lessons of \mychap{chap-policy}---inaction is an action when facing \acp{wicked-prob}, and in this
+case, inaction has led to an undesirable position. Second of all, the current circumstances where device manufacturers
+can't keep attackers out is temporary; when law enforcement faces real difficulty again, this issue will immediately
+resurface. Law enforcement's anticipation of this is likely why they continue to push for \ac{EA} in spite of present
+circumstances. Third of all, regardless of criminal abuse, formalized capability and process can reduce governmental
+abuse, which is the clearer threat to privacy and civil liberties.
 
-It is this task that we turn to next. The following chapter analyzes one particular device-escrow proposal in moderate
-detail, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
+Despite device key escrow's tractability as a technical niche of \ac{EA}, it still has a long way to go. There are few
+proposals, what few there are are high level, and questions of scaling and administration have not been addressed.
+However, as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific
+proposals has the opportunity to identify mutually agreeable solutions---or at least refine the problem understanding
+enough that the next round of debate can be better informed.
+
+It is this task that we turn to next. The following chapter analyzes one particular device-escrow proposal in detail,
+Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
 \cite{savage_lawful_2018}.
-
-% \section{Specific Proposals}
-%% NOTE: not worth going into this section. Just describe a few and go straight to one of the most interesting.
-
-% \begin{figure}[h]
-%   \centering\CaptionFontSize
-%   \includegraphics[width=\linewidth]{arguments/build/comprehensive.proposals.pdf}
-%   \caption{EA Proposals}
-%   \label{fig-arg-proposals}
-% \end{figure}
-
-% % See: committee_decrypting_2018, group_2019, various proposals

### 2020-11-19 Thesis-Chapter5-ThreatModel.tex
Clean up conclusion notes
Add argument that despite hackability, device EA is worth studying
Update threat model with hacking comments

@@
 
+% FIXME: add current hacking techniques.
+
 \modelitem{\ac{EA} workarounds}{The user is capable of evading \ac{EA} mechanisms by encrypting data at a higher level
-in the tech stack, e.g. by encrypting files before saving them to disk. This has already been discussed, and it out of
-scope for evaluating the \ac{EA} system itself.}
+in the tech stack, e.g. by encrypting files before saving them to disk. Additionally, investigators or attackers are
+frequently capable of bypassing current lockout mechanisms. These have already been discussed, and it out of scope for
+evaluating the \ac{EA} system itself.}
 
@@
     Self-escrow
-    & The Class key is escrowed in the device's Secure Enclave (or equivalent) using a write-only component; i.e.,
-      software can update the key, but only an unexposed hardware interface can read it.
-    & Since the Class key itself is being stored, there is no change to the security of the underlying cryptographic
-      protocols. Physical possession and partial disassembly is required, which introduces practical barriers to mass
-      surveillance.
+    & The Class key is escrowed in encrypted form in the device's Secure Enclave (or equivalent) using a write-only
+      component; i.e., software can update the key, but only an unexposed hardware interface can read it.
+    & Since the Class key itself is being stored instead of the encryption scheme being changed, there is no change to
+      the security of the underlying cryptographic protocols. Physical possession and partial disassembly is required,
+      which introduces practical barriers to mass surveillance.
     \\ \hline
@@
 broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR} \ac{EA} proposals do
-\cite{phan_key_2017} \cite{servan_schreiber_jje_2020}. Unfortunately, building this feature into a proposal makes it
-potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.
+\cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately, building this
+feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.
 

### 2020-11-20 Thesis-Chapter2-Background.tex
Address review feedback

@@
 
-There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was discovered
+There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was invented
 long ago, encryption and decryption are performed with the same key, and both parties must have this key in order to
-communicate securely. In \ac{asym-crypto}, which was discovered recently and is also called \ac{public-key-crypto},
+communicate securely. In \ac{asym-crypto}, which was invented recently and is also called \ac{public-key-crypto},
 encryption and decryption are performed with two paired keys, called the public key and the private key. The innovation
@@
 (architecture) and the security of the components themselves; likewise, the security of each component emerges from the
-architecture of its subcomponents and the security of the subcompenents themselves. At the end of this chain of
+architecture of its subcomponents and the security of the subcomponents themselves. At the end of this chain of
 analysis, the security of primitive components commonly relies on cryptography.
@@
 
-% NOTE: tried to find examples of high profile bugs or hacks that relied on crypto failures, but most are still
+% I tried to find examples of high profile bugs or hacks that relied on crypto failures, but most are still
 % architectural (exposed databases, default credentials (Mirai botnet)), human (BEC), or insecure code (buffer
@@
 communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
-the character transposition technique Julius Caesar used to protect private correspondence
+the character substitution technique Julius Caesar used to protect private correspondence
 \cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
@@
 \paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
-uncertainty, extremism, misinformation, and most recently, a pandemic that will have unknown impacts. Since legislature
-is part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
+uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term  impacts. Since legislature is
+part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
 leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
@@
 \newcommand{\law}[4]{ % Year, name, citation, description
-    \item #1: \textbf{#2} \cite{#3} \nopagebreak
+    \item #1: \textbf{#2} #3 \nopagebreak
 
@@
 
-    \law{1789}{\acf{AWA}}{congress_1789}{
+    \law{1789}{\acf{AWA}}{\cite{congress_1789}}{
 
@@
 
-    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}{
+    \law{1791}{Bill of Rights: Amendments I, IV, and V}{\cite{madison_1791}}{
 
@@
 
-    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}{
+    \law{1976}{\acf{AECA}}{\cite{morgan_hr13680_1976}}{
 
@@
 
-    \law{1978}{\acf{FISA}}{rodino_1978}{
+    \law{1978}{\acf{FISA}}{\cite{rodino_1978}}{
 
@@
 
-%     \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
+%     \law{1986}{\acf{CFAA}}{\cite{hughes_hr4718_1986}}{
 % Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
@@
 
-    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{
+    \law{1986}{\acf{ECPA}}{\cite{kastenmeier_hr4952_1986}}{
 
@@
 
-    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{
+    \law{1994}{\acf{CALEA}}{\cite{edwards_hr4922_1994}}{
 
@@
 
-    \law{2001}{\ac{USAPATRIOTAct}}{sensenbrenner_2001}{
+    \law{2001}{\ac{USAPATRIOTAct}}{\cite{sensenbrenner_2001}}{
 
@@
 
-%     \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{
-% Retroactively approved of GWB's warrantless wiretapping policy, a loose interpretation of FISA
+%     \law{2007}{\acf{PAA}}{\cite{mcconnell_s1927_2007}}{
+% Retroactively approved of George W. Bush's warrantless wiretapping policy, a loose interpretation of FISA
 % \cite{landau_making_2013}.
@@
 
-    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{
+    \law{2008}{\acf{FISAAA}}{\cite{reyes_hr6304_2008}}{
 
@@
 
-    \law{2015}{\ac{USAFREEDOMAct}}{sensenbrenner_2015}{
+    \law{2015}{\ac{USAFREEDOMAct}}{\cite{sensenbrenner_2015}}{
 
@@
 
-    \law{1993}{The \ac{clipperchip}}{press_1993}{
+    \law{1993}{The \ac{clipperchip}}{\cite{press_1993}}{
 
@@
 
-    \law{1996}{The \acf{SAFEAct}}{goodlatte_hr3011_1996}{
+    \law{1996}{The \acf{SAFEAct}}{\cite{goodlatte_hr3011_1996}}{
 
@@
 
-    \law{1996}{The \acf{PROCODEAct}}{burns_s1726_1996}{
+    \law{1996}{The \acf{PROCODEAct}}{\cite{burns_s1726_1996}}{
 
@@
 
-%     \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{
+%     \law{1997}{Encrypted Communications Privacy Act}{\cite{leahy_s376_1997}}{
 % % mix of SE protection and EA guidelines?
@@
 %     \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}
-%         {ashcroft_s2067_1998}{
+%         {\cite{ashcroft_s2067_1998}}{
 % % like above?
@@
 
-%     \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{
+%     \law{1999}{Encryption for the National Interest Act}{\cite{goss_hr2616_1999}}{
 % }
 
-%     \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{
+%     \law{1999}{Tax Relief for Responsible Encryption Act}{\cite{goss_hr2617_1999}}{
 % % ^^ these two introduced by same rep on same day
@@
 
-    \law{1999}{\ac{CESA}}{administration_CESA_1999}{
+    \law{1999}{\ac{CESA}}{\cite{administration_CESA_1999}}{
 
@@
 
-    \law{2000}{Enhancement of Privacy and Public Safety in Cyberspace Act}{leahy_s3083_2000}{
+    \law{2000}{Enhancement of Privacy and Public Safety in Cyberspace Act}{\cite{leahy_s3083_2000}}{
 
@@
 
-    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{
+    \law{2016}{Compliance with Court Orders Act (discussion draft)}{\cite{burr_2016}}{
 
@@
 
-    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{
+    \law{2018}{The Secure Data Act}{\cite{lofgren_hr5823_2018}}{
 
@@
 
-    \law{2019}{The \ac{ENCRYPTAct}}{lieu_hr4170_2019}{
+    \law{2019}{The \ac{ENCRYPTAct}}{\cite{lieu_hr4170_2019}}{
 
@@
 
-    \law{2020}{The \acf{EARNITAct}}{graham_s3398_2020}{
+    \law{2020}{The \acf{EARNITAct}}{\cite{graham_s3398_2020}}{
 
@@
 
-    \law{2020}{The \acf{LEADAct}}{graham_2020}{
+    \law{2020}{The \acf{LEADAct}}{\cite{graham_2020}}{
 
@@
 %
-% For more, read donahue_comparative_2018 (Israel and China) and budish_encryption_2018 (framework for analysis)
+% For more, read donahue_comparative_2018 (Israel and China) and \cite{budish_encryption_2018} (framework for analysis)
 %
@@
 % - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
-% - Require use of backdoored algorithms, security via obscurity (anyone with the secret knowledge can see it)
+% - Require use of backdoor-ed algorithms, security via obscurity (anyone with the secret knowledge can see it)
 % - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
-%   - Store extra in gov't DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
+%   - Store extra in govt DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
 %   - Split the key and store it multiple places
@@
 % key recovery information. Do the protocols exist? EES failed. Needs to be rock solid. (2) Dispersed key recovery
-% information. (3) A robust, antifragile, inherently secure means of protecting those keys, and perhaps decaying them over
-% time (https://en.wikipedia.org/wiki/Data_degradation).
+% information. (3) A robust, anti-fragile, inherently secure means of protecting those keys, and perhaps decaying them
+% over time (https://en.wikipedia.org/wiki/Data_degradation).
 
@@
 
-\subsection{Non-Escrow Cryptographic Recovery}
+\subsection{Non-Escrow Cryptographic Data Recovery}
 
@@
 
-% these look good: sacharoff_unlocking_2018, fredona_fifth_2016
+% these look good: \cite{sacharoff_unlocking_2018}, \cite{fredona_fifth_2016}
 % this probably has something to say: kerr_encryption_2017
@@
 \propsend
+
+% FIXME: Separate out into separate, short chapter. Also adds natural place to introduce my more "advanced" DFD syntax.
+
+% Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own
+% (short) chapter.
+% Kalafut: Argument maps are used in chapter 4. I'm not sure that you need to mention argdown at all. If you do, I would
+% only do so in the context of stating that it is the tool you are using to create argument maps.
 
@@
 \end{figure}
-
-% IDEA:
-% Use the slightly more complete IBIS notation, which includes ISSUES and POSITIONS as well as ARGUMENTS.
-% (Argdown) basically just has arguments and positions. The additional types may make the total picture clearer.
-% Either use styling with #hashtags (https://argdown.org/syntax/#hashtags with
-%                                    https://argdown.org/guide/colorizing-maps.html#the-colorscheme-setting)
-% Or add a plugin (https://argdown.org/guide/extending-argdown.html)
 

### 2020-11-20 Thesis-Chapter3-PolicyMaking.tex
Address review feedback

@@
 
-The more important insight is one step removed: precisely because there is an inexhaustable set of potential solutions,
+The more important insight is one step removed: precisely because there is an inexhaustible set of potential solutions,
 the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
@@
   \item Every policy implementation has irreversible effects.
-  \item There is an inexhaustable set of potential solutions to the problem.
+  \item There is an inexhaustible set of potential solutions to the problem.
   \item Solutions from other domains do not apply directly to this problem.
@@
 %   group_2019 (argues for DAR) and
-%   savage_lawful_2018 (a concrete DAR proposal)? -- and related, servan_schreiber_jje_2020
+%   savage_lawful_2018 (a concrete DAR proposal)? -- and related, \cite{servan_schreiber_jje_2020}
 
@@
 
-% Note on using argument maps to facilitate this argumentative process
-
 Recall from \mysec{sec-arg-maps-intro} that \acp{argmap} resulted from the same research that produced the concept of
@@
 
-% Summarize chapter: wicked problems, why normal methods fail, how this strategy works
+% TODO: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this paragraph
+% or its ideas should be at the start of the Argument Maps section.
 
@@
 power over the debate in the second. It means that re-raising questions about the role of technology is playing society
-is legitimate. To follow this proposed strategy would mean that hardline rhetoric doesn't help \cite{ruiz_there_2018}
+is legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
 \cite{geller_2019}, though sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more

### 2020-11-20 Thesis-Chapter4-Arguments.tex
Address review feedback

@@
-\chapter{EA Debate Argument Maps}
+\chapter{EA Debate Arguments}
 \label{chap-arguments}
@@
 
+% FIXME: I need to address \cite{koepke_2020} in these maps.
+
 A significant portion of the disconnect between government and technical community representatives is disagreement over
 whether law enforcement has \ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is
-well-represented by proponent former \ac{FBI} Director James Comey in a 2014 speech \cite{comey_2014} (though he did not
-coin the phrase \cite{swire_encryption_2011}). The argument is represented in \myfig{fig-arg-going-dark}, also informed
-by critiques \cite{rogaway_moral_2015}.
+well-represented by proponent former \ac{FBI} Director James Comey in a 2014 speech titled ``Going Dark''
+\cite{comey_2014} (though he did not coin the phrase \cite{swire_encryption_2011}). The argument is represented in
+\myfig{fig-arg-going-dark}, also informed by critiques \cite{rogaway_moral_2015}.
 

### 2020-11-20 Thesis-Chapter5-ThreatModel.tex
Address review feedback

@@
 
-\modelitem{The platform abuser}{Why bother with \ac{EA} at all? Because the device user may actually be a threat in the
-larger safety context. The end user is not the most important thing. This is a crucial point.
+\modelitem{The platform abuser}{The device user may actually be a threat in the larger safety context. The end user is
+not the most important thing. This is a crucial point.
 
@@
 force updates values the vendor's judgement over the user's (even if it is a good idea). Software that enforces Digital
-Rights Management values corporations' rights over the user's (sometimes to extremes \cite{eff_2020}). Software that
-doesn't encrypt user communications at all says the service provider's ability to scrape personal information to create
-targeted ads is more important than the user's privacy. Software that enables \ac{EA} values lawful investigatory
-process over the user's ultimate privacy. Subversion of the user's will is not a pleasant concept, but a spade is a
-spade.
+Rights Management values corporations' data rights over the user's (sometimes to extremes \cite{eff_2020}). Software
+that doesn't encrypt user communications at all values the service provider's ability to scrape personal information to
+create targeted ads over the user's privacy. Software that enables \ac{EA} values lawful investigatory process
+over the user's ultimate privacy. Intentional subversion of the user's will is not a comfortable thought, but it is a
+frequent reality, and the reality of \ac{EA}.
 
 In the broader threat model, the platform abuser is someone who is using the device encryption for cover of illicit
-activities. They represent a threat to public safety.}
+activities. He represents a threat to public safety.}
 
@@
 
-The hawkish lawmaker is a threat to the security of a cryptosystem (\ac{EA} or not) in that they may judge it too
+Hawkish lawmakers are a threat to the security of a cryptosystem (using \ac{EA} or not) in that they may judge it too
 inhibitive to law enforcement and outlaw its very use, mandating something weaker in its place.}

### 2020-11-20 Thesis-Chapter6-PathsAhead.tex
Write intro to final chapter

@@
 
-% We have a wicked problem
-% I propose we handle it a certain way
-% Part of this is cooperative debate, which I tried to capture in argument maps
-% Another element is focusing on specific proposals, so we dove into LDAWMSR
+The interests of security, safety, privacy, and trust are converging on the issue of encryption and \acl{EA} and
+creating a \ac{wicked-prob}. Information security is increasingly important as more data, business, and infrastructure
+goes online. The technical community is struggling with security; defending against the average attacker is hard, and
+against the best attackers is nearly impossible, even for the most motivated and well-resourced organizations.
+\Ac{cryptography} plays a fundamental role in security, yet its simultaneous role in enabling unprecedented privacy
+means it has controversial side effects.
 
-%% Prof. Walsh:
-%% The dynamic regulatory environment and competing governmental branches at the executive, legislative, and judicial
-%% level will continue to shape the...
+One of these side effects is encryption-enabled privacy's impact on public safety. Despite a shortage of quantitative
+analyses showing its impact on crime---something that would be impossible to fully quantify, after all---encryption
+decidedly withholds evidence of wrongdoing from service providers and law enforcement, enabling truly harmful activity
+to go unchecked, and decreasing public safety. This a legitimate concern and the primary motivation for \ac{EA}.
+
+However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
+stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
+ranging from identity theft, to inappropriate data mining, to mass surveillance. To the extent that encryption mitigates
+these, it increases public safety.
+
+Finally, choosing how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
+considerable questions of trust. Law enforcement agencies, intelligence agencies, technology makers, individual users,
+and the public at large are all players; each have demonstrated untrustworthiness, and different approaches disperse
+trust differently.
+
+% TODO: one paragraph here^^ or four?
+
+This complex mix of issues is what makes encryption policy a \ac{wicked-prob}. To merely formulate the problem requires
+explanation upon explanation and research upon research. Any formulation is concomitant with a solution, but the outcome
+of a solution cannot be accurately forecasted or independently evaluated. The underlying conditions are constantly in
+flux, affected by external forces as well as every policy action and inaction.
+
+In this thesis I propose a method to contend with \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes
+collaborative debate between individuals diverse in background and values, facilitated in part by argument maps, and
+focusing on specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and
+ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}.
 
@@
 %   - Government:
+%% Prof. Walsh:
+%% The dynamic regulatory environment and competing governmental branches at the executive, legislative, and judicial
+%% level will continue to shape the...
 %       - Take the low hanging fruit first, please

### 2020-11-20 Thesis-Chapter2-Background.tex
Write intro to final chapter
Clean up glossary

@@
 The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
-war, and increased government surveillance set events in motion toward conflict, but it was the 2013 \ac{snowden}
-revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
-wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
-caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
-promised a technical solution. Benefiting from advancements in underlying technologies and under pressure from both the
-human rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing
-default device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.
+war, and increased government surveillance set events in motion toward conflict, but it was the 2013 Snowden revelations
+that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
+caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
+public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
+technical solution. Benefiting from advancements in underlying technologies and under pressure from both the human
+rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing default
+device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.
 
@@
 
-The U.S. corporate response to \ac{snowden} should not be overly construed as a morally motivated defense of civil
-rights, as their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the
-pursuit of basic digital security. Recall encryption and cryptography's critical role in security as described in
+The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
+their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
+basic digital security. Recall encryption and cryptography's critical role in security as described in
 \mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
@@
 
-Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers
-in 25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope
-(Section 206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business
-records, but the \ac{snowden} leaks revealed that the government had been secretly using a non-standard definition of
-``business records'' in order ``to justify requests for domestic telephone metadata delivered in bulk, not
-individualized requests'' \cite{landau_making_2013}.
+Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers in
+25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope (Section
+206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business records, but
+the Snowden leaks revealed that the government had been secretly using a non-standard definition of ``business records''
+in order ``to justify requests for domestic telephone metadata delivered in bulk, not individualized requests''
+\cite{landau_making_2013}.
 
@@
 The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the Senate
-almost two years to the day after \ac{snowden}'s leaks began. It contains many reforms to \ac{FISA} and the use of
-\acp{NSL}. Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass
-collection of telephone metadata.
+almost two years to the day after Snowden's leaks began. It contains many reforms to \ac{FISA} and the use of \acp{NSL}.
+Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass collection of
+telephone metadata.
 
@@
 
-Though key recovery is the most discussed \ac{EA} approach, alternative technical means have been proposed.
+Though key escrow is the most discussed \ac{EA} approach, alternative technical means have been proposed.
 
@@
 
-% FIXME: Separate out into separate, short chapter. Also adds natural place to introduce my more "advanced" DFD syntax.
+% FIXME: Separate out into separate, short chapter, probably after the wicked problems chapter. Also adds natural place
+% to introduce my more "advanced" DFD syntax.
 

### 2020-11-20 Thesis-Glossary.tex
Begin threat model chapter
Finish spoofing, begin tampering threats
Finish STRIDE threats
Finish risk analysis
Address review feedback
Clean up glossary

@@
 % NOTE: Do not end with a period. The package adds this for you.
-% IDEA: auto-update glossary.md. This should be the master, because it is more fully specified
 
@@
 
-\newterm{snowden}{Snowden}{
-    Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
-    mass collection of internet communications for surveillance
-}
+% \newterm{snowden}{Snowden}{
+%     Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
+%     mass collection of internet communications for surveillance
+% }
 
@@
 }
+
+\newterm{LDAWMSR}{LDAWMSR}{
+    A reference to computer scientist and security researcher Stefan Savage's device key escrow proposal, ``Lawful
+    Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}
+}
 % End terms
+
+\newcommand{\ldawmsR}[0]{\ii{\ac{LDAWMSR}}}
+\newcommand{\ldawmsr}[0]{\ldawmsR~}
 
@@
 \newabbreviation{AECA}{AECA}{Arms Export Control Act}
+\newabbreviation{APT}{APT}{advanced persistent threat}
 \newabbreviation{AV}{AV}{anti-virus}
@@
 \newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
+\newabbreviation{CIA}{CIA}{Central Intelligence Agency}
 \newabbreviation{CP}{CP}{child pornography}
@@
 \newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
+\newabbreviation{DAR}{DAR}{data at rest}
 \newabbreviation{DE}{DE}{device encryption}
 \newabbreviation{DFD}{DFD}{data flow diagram}
+\newabbreviation{DID}{DID}{Device ID}
 \newabbreviation{DIM}{DIM}{data in motion}
-\newabbreviation{DAR}{DAR}{data at rest}
+\newabbreviation{DOS}{DOS}{denial of service}
+\newabbreviation{DSK}{DSK}{Device Seal Key}
 \newabbreviation{E2E}{E2E}{end to end}
@@
 \newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
-\newabbreviation{GCHQ}{GCHQ}{(UK) Government Communications Headquarters}
+\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters (UK)}
+\newabbreviation{HSM}{HSM}{Hardware Security Module}
 \newabbreviation{IBIS}{IBIS}{issue-based information system}
@@
 \newabbreviation{TLS}{TLS}{Transport Layer Security}
+\newabbreviation{UID}{UID}{Unique ID}
 \newabbreviation{VPN}{VPN}{virtual private network}

### 2020-11-20 Thesis-Chapter6-PathsAhead.tex
Work on conclusion

@@
-\chapter{Paths Ahead}
-\label{chap-pathsahead}
+\chapter{Paths Forward}
+\label{chap-pathsforward}
 
@@
 explanation upon explanation and research upon research. Any formulation is concomitant with a solution, but the outcome
-of a solution cannot be accurately forecasted or independently evaluated. The underlying conditions are constantly in
+of a solution cannot be accurately forecasted or evaluated in isolation. The underlying conditions are constantly in
 flux, affected by external forces as well as every policy action and inaction.
@@
 focusing on specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and
-ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}.
+ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and
+paths forward for both the technical and policymaking communities.
+
+\section{Conclusions}
 
@@
 %   - Device escrow with secure hardware represents true progress towards acceptably secure EA, so hope is not lost
+
+Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
+the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption poses a present danger to public safety
+in certain respects. The bogeymen of terrorists and criminals using encryption to evade the law are real---in the case
+of those peddling and viewing \acl{CSAM}, shockingly so \cite{keller_internet_2019}. However, the argument that
+\ac{encryption} is creating circumstances that are to blame for \acl{LE} failure is not currently strong enough to
+justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor coordination and training \cite{carter_2018}; for
+\ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
+\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Strong \ac{encryption} plays a key role
+in protecting security and privacy, public goods that other technological developments often fail to provide.
+
+Though the pro-\ac{EA} argument is ultimately not strong enough to justify an \ac{EA} mandate, incorporating it to the
+technologist's point of view teaches important lessons. One is that wholistic cryptosystem threat models need to include
+two non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and
+this power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating
+steps. Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption;
+lawmakers represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too
+dangerous. The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.
+
+Another lesson is that the current situation is temporary. This is necessarily true of \acp{wicked-prob}, but especially
+true for encryption policy in the context of rapidly evolving technology and a vulnerability arms race between device
+manufacturers and lawful hackers. The facts of the case are bound to change, and the conclusion that \ac{EA} is
+unjustified may change as well.
+
+% TODO: make a \term for mass surveillance
+
+The additions to the threat model and the instability of present circumstances mean that additional research into
+\ac{EA} is fully justified. It belongs as just a component of the effort that society should be putting towards
+addressing the broader set of issues. Fortunately, developments in secure hardware mark real progress in the search for
+acceptably secure \ac{EA}. Secure hardware technology is clearly imperfect, and alone does not resolve the numerous
+challenges that full-fledged \ac{EA} proposals must overcome, but device escrow for \acl{DAR} such as that described in
+Savage's proposal offers transparency, protection from mass surveillance, and \acl{LE} utility while making progress on
+security.
+
+These are not strong conclusions. However, when dealing with \acp{wicked-prob}, meaningful debate and discussion counts
+as progress even when it doesn't seem like it. \Acp{wicked-prob} cannot be analyzed into resolution; grasping the
+problem alone constitutes much of the work towards solving it. The conclusions above do not bring closure, but they
+should bring more clarity, and its clarity that counts.
+
 % Where do we go from here?
@@
 %       - Z
+
+\section{Paths Forward: Technology Makers}
+
+This thesis represents an ``OO'' in the \ac{OODAloop} (see \myfig{fig-policy-ooda-loop}). It is observation and
+orientation, framed around the debate and a specific proposal. Observation and orientation are just the beginning of the
+process; next it is time to decide and act. How should technology makers respond?
+
+The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
+such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is is not trying
+hard enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding
+harder will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with
+them. Tech is used to addressing tame problems, but it is time that they turn their massive operations and intellectual
+capital towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences,
+backgrounds, and values in decision making processes that steer investment and design decisions. It means creating
+products and business models that are responsive to evolving conditions. It means considering the ways their creations
+will be abused.
+
+Work on \acp{wicked-prob} spins out smaller problems, some tame and some wicked yet manageable (see
+\myfig{fig-policy-ooda-process}). Aside from making the changes to tackle \acp{wicked-prob} head on, the technical
+community should also work on the tame sub-problems, a few of which are listed here.
+
+\newcommand{\taskstart}[0]{\begin{itemize}}
+\newcommand{\taskitem}[2]{ % Name, description
+    \item \textbf{#1} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
+}
+\newcommand{\taskend}{\end{itemize}}
+
+% FIXME: flesh out.
+\taskstart
+    \taskitem{Basic security research}{
+
+\ac{EA} is only applicable when security is actually strong enough to keep attackers out in the first place.
+
+}
+
+    \taskitem{Research technical transparency mechanisms}{
+
+Technically-enforce transparency mechanisms would help with \ac{DAR} and are a necessity for \ac{DIM}.
+
+}
+
+    \taskitem{Research distributed trust mechanisms}{
+
+Centralized trust in the \ac{EA} proposals represent their greatest risk.
+
+}
+\taskend
+
+While these problems remain unsolved, technologists must continue to provide clear security advice and be vigilant
+against bad technology policy.
+
+\section{Paths Forward: Policy Makers}
+
+Government has not been silent on the issue of encryption and lawful access. The dynamics created by the executive,
+legislative, and judicial branches ensures that the regulatory environment will evolve with changing federal policies,
+new laws, and rulings that refine implementation existing law.
+
+Be strategic about \acp{wicked-prob}.
+
+% FIXME: flesh out ^^vv.
+
+\taskstart
+    \taskitem{Take the low-hanging fruit}{
+
+Implement low-risk, high-reward policies first: fund industry-LE engagement that leverages existing access laws and
+follow through on anti-CSAM legislation.
+
+}
+
+    \taskitem{Regulate current lawful hacking }{
+
+Establish rules for federal LE and guidlines for state and local that create a non-abusive use of current mobile device
+forensic tools.
+
+}
+
+    \taskitem{Fund research}{
+
+Don't compare \ac{EA} to going to the moon unless you fund it like it.
+
+}
+\taskend
+
 %   - Government:
-%% Prof. Walsh:
-%% The dynamic regulatory environment and competing governmental branches at the executive, legislative, and judicial
-%% level will continue to shape the...
 %       - Take the low hanging fruit first, please
@@
 %       - Take on weak area XYZ
-%   - Tech:
-%       - Do research, provide clear security advice, be vigilant
-%       - Take on weak area XYZ
+
 %   - Either:

### 2020-11-21 Thesis-Chapter1-Introduction.tex
Begin threat model chapter
Finish rough draft!

@@
 attempt to end the debate, but to structure and analyze it.
+
+% TODO: change EA everywhere to "lawful access"?
 
@@
 models using data flow diagrams. In the following chapters, I analyze strategies for tacking \acp{wicked-prob}, map the
-\acl{EA} arguments in detail, and create a multi-level threat model, broken down by the significantly different
-challenges posed by \acl{DAR} and \acl{DIM}. The thesis concludes with a set of policies and principles to apply to the
-problem in light of valid arguments and threats.
+\acl{EA} arguments in detail, and demonstrate the security risks with data flow diagrams and threat analysis, focusing
+on one particular EA proposal. The thesis concludes with research and policy paths forward in light of valid arguments
+and threats.

### 2020-11-21 Thesis-Chapter6-PathsAhead.tex
Finish rough draft!

@@
 
-% What can we conclude?
-%   - It may not feel like it, but (meaningful) debate and discussion counts as progress for a wicked problem
-%   - Neither side has the whole picture, but the "no EA" arguments are currently stronger
-%   - A threat model including malicious users and threatening governments makes EA make more sense
-%   - Device escrow with secure hardware represents true progress towards acceptably secure EA, so hope is not lost
-
 Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
@@
 
-% Where do we go from here?
-%   - This was an attempt an "OO" iteration in the OODA; we need more, with involvement all around
-%   - committee_decrypting_2018 presented a framework to use to judge proposals. I have a simple version of it.
-%     here are where we are weak:
-%       - X
-%       - Y
-%       - Z
-
 \section{Paths Forward: Technology Makers}
@@
 
-% FIXME: flesh out.
 \taskstart
@@
 
-\ac{EA} is only applicable when security is actually strong enough to keep attackers out in the first place.
+\ac{EA} is only applicable when security is actually strong enough to keep attackers out in the first place, which is
+not the case for current device encryption. To some extent, this is a security ergonomics problem more than a technical
+problem: cryptographers can make device encryption incredibly strong today, but users will not let them because of its
+adverse effects on usability. Solving the strong yet usable device encryption problem will make the most promising class
+of \ac{EA}, device key escrow, relevant. Improving basic security across the board will make any \ac{EA} proposal more
+tenable.
 
@@
 
-Technically-enforce transparency mechanisms would help with \ac{DAR} and are a necessity for \ac{DIM}.
+One of the largest non-technical risks of \ac{EA} is abuse by jurisdictions with poor rule of law, or where law
+insufficiently respects civil rights in the first place. This is the motive for transparency as a goal of the threat
+model. Device key escrow can achieve partial transparency through physical possession and disassembly requirements.
+Ideally, usage of \ac{EA} mechanisms would be fully auditable to the public through inherent technical mechanisms. This
+would mark an improvement for \acl{DAR} and is a necessity for progress on \acl{DIM}.
 
@@
 
-Centralized trust in the \ac{EA} proposals represent their greatest risk.
+Centralized trust in \ac{EA} proposals represent their greatest risk, whether it is in the form of a database of keys or
+a control center with the power to add ``ghost users'' to conversations. Knowing that insecurity (and trust) cannot be
+destroyed, only moved around, the goal is a distribution of trust that allows the system to function without large
+amounts of trust assigned to any single party. This would contribute to transparency and greatly reduce the risk profile
+of an \ac{EA} proposal overall.
 
@@
 
-While these problems remain unsolved, technologists must continue to provide clear security advice and be vigilant
-against bad technology policy.
+Fortunately, research in these areas is occurring. More is necessary. Some \ac{EA} proposals include built in
+transparency and distributed trust mechanisms \cite{goos_oblivious_1996} \cite{phan_key_2017}
+\cite{servan_schreiber_jje_2020}; these proposals' weakness is that the approaches are insufficiently understood to be
+trusted with the high stakes of implementation. While these problems remain unsolved, technologists must continue to
+provide clear security advice and be vigilant against bad technology policy.
 
@@
 legislative, and judicial branches ensures that the regulatory environment will evolve with changing federal policies,
-new laws, and rulings that refine implementation existing law.
+new laws, and rulings that refine implementation existing law. The judiciary will indeed shape encryption policy through
+its rulings, and is correct to be reevaluating current Fourth Amendment jurisprudence in light of rapidly changing
+circumstances. Still, the primary actors capable of consciously taking part in strategic policymaking will be the
+executive and legislative branches. How should these policymakers respond?
 
-Be strategic about \acp{wicked-prob}.
+Government must strategically approach encryption policy as a \ac{wicked-prob}. Industry has failed at analyzing the
+problem away, and government has failed at muddling the problem away. It is true that incrementalism has safely avoided
+dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a research-stifling high-tension debate
+and circumstances of poor utilization of current capabilities mixed with unregulated use of invasive \ac{lawful-hacking}
+tools. Policymakers and law enforcement agencies need to be forthcoming with the real challenges they face. They are
+important contributors to the diverse groups that can iteratively refine the problem definition and possible solutions.
+Just like the technical side, the policy side of the solutions must be responsive to evolving conditions. Most
+importantly, any initiative needs follow through, since a single \ac{OODAloop} iteration will not be enough to tame this
+issue.
 
-% FIXME: flesh out ^^vv.
+Based on the analysis performed thus far, there are a few policy changes that would be good first steps to take today.
 
@@
 
-Implement low-risk, high-reward policies first: fund industry-LE engagement that leverages existing access laws and
-follow through on anti-CSAM legislation.
+Law enforcement is often ill-equiped to handle digital evidence and coordinate with servive providers using existing
+access mechanisms \cite{carter_2018}. Justice Department task forces to combat \acl{CSAM} are underfunded and
+understaffed \cite{keller_internet_2019}. Improving the effectiveness of existing capabilities and programs carries no
+risk and guaranteed results.
 
@@
 
-    \taskitem{Regulate current lawful hacking }{
+    \taskitem{Regulate current lawful hacking}{
 
-Establish rules for federal LE and guidlines for state and local that create a non-abusive use of current mobile device
-forensic tools.
+\Ac{lawful-hacking} using commercial mobile device forensic tools has become a common practice but invokes serious
+privacy and civil rights concerns \cite{koepke_2020}. Regardless of any long-term strategy to encryption and lawful
+access policy, \ac{lawful-hacking} must be constrained. Michigan's 2020 constitutional amendment represents one positive
+development: it explicitly protects electronic data and communications from search without a warrant
+\cite{ballotpedia_2020}. Protections like this enable a non-abusive use of current \ac{lawful-hacking} capability.
 
@@
 
-Don't compare \ac{EA} to going to the moon unless you fund it like it.
+The government wants technical advancement in \ac{EA} technology, but there is very little financial incentive for
+private companies to research \ac{EA}, and the threatening climate is stifling academic research. If the federal
+government is going to compare building secure \ac{EA} to going to the moon \cite{cushing_moon_2018}, then they should
+fund it like the Apollo program. In order to be welcomed, such funding should be accompanied with a binding statute
+excluding the possibility of any premature \ac{EA} mandate.
 
@@
 
-%   - Government:
-%       - Take the low hanging fruit first, please
-%       - Fund research, don't mislead, don't be premature
-%           - If you're going to compare it to going to the moon, fund us like it!
-%       - Take on weak area XYZ
+\section{Decision and Action}
 
-%   - Either:
-%       - Know that there is no stable solution, balance rules and principles, serve your neighbor
-% see varia_2018 (maybe a good quote?)
+The goal of this thesis is to clarify encryption and \ac{EA} as a \ac{wicked-prob}, to break down the debate, and to
+evaluate the security of current \ac{EA} proposals. I aim to move the discussion forward with the ideas of how to tackle
+\acp{wicked-prob}, additions to the traditional cryptosystem threat model, and security analysis of a modern proposal.
+This study is observation and orientation; it is up to technologists, policymakers, and common people to decide and
+act. Contending with these issues is difficult on many levels, but the future of security, safety, and privacy depend on
+coming together to find a good answer where no perfect answer exists.
 

### 2020-11-21 scripts/new-source.sh

@@ -14,7 +14,7 @@ SOURCE - The DOI or URL of the new material
 ORIG_DIR=$(pwd)
 cd "$READING_DIR" || exit
 function cleanup() {
-    if (( 2 == $? )); then
+    if (( 2 < $? )); then
         echo "$USAGE"
     fi
 }

### 2020-11-21 Thesis-Chapter2-Background.tex
Fix LEAD->LAED act term

@@
 
-    \law{2020}{The \acf{LEADAct}}{\cite{graham_2020}}{
+    \law{2020}{The \acf{LAEDAct}}{\cite{graham_2020}}{
 
-Introduced by the same senator as the \ac{EARNITAct}, the \ac{LEADAct} is a direct attack on encryption. The
-\ac{LEADAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
+Introduced by the same senator as the \ac{EARNITAct}, the \ac{LAEDAct} is a direct attack on encryption. The
+\ac{LAEDAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
 developers to implement the ability to decrypt all data stored or passing through the device or software system
@@
 though it is strongly argued that \acs{CALEA} explicitly precludes this kind of demand \cite{gidari_2016}. Since the FBI
-dropped the case, the matter is not formally settled, though the consensus is that legislation such as the \ac{LEADAct}
+dropped the case, the matter is not formally settled, though the consensus is that legislation such as the \ac{LAEDAct}
 is required before courts would uphold that command.
@@
 compelled. Third, a surprise \ac{AWA} ruling could open the door to demands for access, though legislation like
-\ac{LEADAct} is more likely to have this effect.
+\ac{LAEDAct} is more likely to have this effect.
 

### 2020-11-23 Thesis-Chapter2-Background.tex
Spelling and editorial improvements

@@
 wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
-% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
 jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
@@
 leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
-community to engage productively with policy makers \cite{granick_2018}.
+community to engage productively with policymakers \cite{granick_2018}.
 
@@
 
-Among European countriues, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
+Among European countries, France and Russia join the U.K. in having access mandates \cite{lewis_2017}. France, Belgium,
 Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for France, Belgium,
@@
 Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
-information. In this arrangement, trust is put in the distrubuted system, not in any single party. Such systems offer
+information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
 high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access

### 2020-11-23 Thesis-Chapter3-PolicyMaking.tex
Spelling and editorial improvements

@@
 practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
-puts it, ``The systems-approach `of the first generation' [i.e., classical analytics] is inadequate for dealing with
+puts it, ``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
 wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
@@
 consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
-due to their conflicting values and priorities, but Linblom writes encouragingly about ``the ease with which individuals
+due to their conflicting values and priorities, but Lindblom writes encouragingly about ``the ease with which individuals
 of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to

### 2020-11-23 Thesis-Chapter5-ThreatModel.tex
Clean up glossary
Spelling and editorial improvements

@@
 
-To save space, the tables use shorthand for law enforcement (LE), law enforcement officer (LEO), law enforcement agency
-(LEA), Device ID (DID), Device Seal Key (DSK), \acf{DOS}
+To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{DID}, \acf{DSK}, and \acf{DOS}.
 
@@
 \threattablebegin{Tampering}{tampering}{Tamper}
-  T1 & Defensive user corrupts DSK, DID, or token
+  T1 & Platform abuser corrupts DSK, DID, or token
   & LE attempts at perform \ac{EA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
@@
   \\ \hline
-  T2 & Defensive user destroys the \ac{EA} hardware interface
+  T2 & Platform abuser destroys the \ac{EA} hardware interface
   & LE attempts to perform \ac{EA} will fail due to damaged hardware.
@@
 costs relative to device manufacturer's naturally large size is unclear. Costs to the government include the \ac{FBI}
-operation and a necessary scale-up in training and resources available to state and local law enforcement. Government
-reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a \ac{DAR}
-\ac{EA} regulation as well \cite{edwards_hr4922_1994}.
+laboratory cost and a necessary scale-up in training and resources available to state and local law enforcement.
+Government reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a
+\ac{DAR} \ac{EA} regulation as well \cite{edwards_hr4922_1994}.

### 2020-11-23 Thesis-Chapter6-PathsAhead.tex
Spelling and editorial improvements

@@
 \ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
-\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Strong \ac{encryption} plays a key role
-in protecting security and privacy, public goods that other technological developments often fail to provide.
+\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Additionally, strong \ac{encryption}
+plays a key role in protecting security and privacy, public goods that other technological developments often fail to
+provide.
 
@@
 
-This thesis represents an ``OO'' in the \ac{OODAloop} (see \myfig{fig-policy-ooda-loop}). It is observation and
-orientation, framed around the debate and a specific proposal. Observation and orientation are just the beginning of the
-process; next it is time to decide and act. How should technology makers respond?
-
 The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
@@
 harder will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with
-them. Tech is used to addressing tame problems, but it is time that they turn their massive operations and intellectual
+them. Tech is used to addressing tame problems, but it is time that they turn their massive financial and intellectual
 capital towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences,
@@
 
-\section{Paths Forward: Policy Makers}
+\section{Paths Forward: Policymakers}
 
@@
 circumstances. Still, the primary actors capable of consciously taking part in strategic policymaking will be the
-executive and legislative branches. How should these policymakers respond?
+executive and legislative branches.
 
-Government must strategically approach encryption policy as a \ac{wicked-prob}. Industry has failed at analyzing the
-problem away, and government has failed at muddling the problem away. It is true that incrementalism has safely avoided
-dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a research-stifling high-tension debate
-and circumstances of poor utilization of current capabilities mixed with unregulated use of invasive \ac{lawful-hacking}
-tools. Policymakers and law enforcement agencies need to be forthcoming with the real challenges they face. They are
-important contributors to the diverse groups that can iteratively refine the problem definition and possible solutions.
-Just like the technical side, the policy side of the solutions must be responsive to evolving conditions. Most
-importantly, any initiative needs follow through, since a single \ac{OODAloop} iteration will not be enough to tame this
-issue.
+Executive and legislative policymakers must strategically approach encryption policy as a \ac{wicked-prob}. Industry has
+failed at analyzing the problem away, and government has failed at muddling the problem away. It is true that
+incrementalism has safely avoided dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a
+research-stifling high-tension debate and circumstances of poor utilization of current capabilities mixed with
+unregulated use of invasive \ac{lawful-hacking} tools. Policymakers and law enforcement agencies need to be forthcoming
+with the real challenges they face. They are important contributors to the diverse groups that can iteratively refine
+the problem definition and possible solutions. Just like the technical side, the policy side of the solutions must be
+responsive to evolving conditions. Most importantly, any initiative needs follow through, since a single \ac{OODAloop}
+iteration will not be enough to tame this issue.
 
@@
 
-Law enforcement is often ill-equiped to handle digital evidence and coordinate with servive providers using existing
+Law enforcement is often ill-equipped to handle digital evidence and coordinate with service providers using existing
 access mechanisms \cite{carter_2018}. Justice Department task forces to combat \acl{CSAM} are underfunded and
@@
 \acp{wicked-prob}, additions to the traditional cryptosystem threat model, and security analysis of a modern proposal.
-This study is observation and orientation; it is up to technologists, policymakers, and common people to decide and
-act. Contending with these issues is difficult on many levels, but the future of security, safety, and privacy depend on
-coming together to find a good answer where no perfect answer exists.
+In the language of the \ac{OODAloop}, this study is observation and orientation; it is up to technologists,
+policymakers, and common people to decide and act. Contending with these issues is difficult on many levels, but the
+future of security, safety, and privacy depend on coming together to find a good answer where no perfect answer exists.
 

### 2020-11-24 Thesis-Chapter2-Background.tex
Updates based on presentation feedback

@@
 perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
-encryption of the data in transit. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}
-and \ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise
-any other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an
-attacker without detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of
-encryption for \ac{DIM}.
+encryption of the data in motion. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy} and
+\ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any
+other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an attacker without
+detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for
+\ac{DIM}.
 
@@
 \paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
-antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
+antagonistic, an assumption implicit in the ``nothing-to-hide'' argument, but discussions in recent years have shown
 that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
@@
 
+% FIXME: Engelsma: Consider moving this right after "encryption basics", as I have referred to many of these approaces
+% already.
+
 % There are several types of approaches:
@@
 
+% TODO: include a more descriptive key to understand argument maps.
+
+% - Prof. Dulimarta: Could you apply fuzzy logic to the argument map? Quantify it? -- I think this would run into limits
+%   once you start comparing unquantifiable things, but it could be used to suss out the valid arguments, and would
+%   probably be really interesting otherwise.
+
 % Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own

### 2020-11-24 Thesis-Chapter5-ThreatModel.tex
Editorial updates
Updates based on presentation feedback

@@
 % TODO: remove this paragraph if I end up changing EA to LA
+
+% TODO: Dulimarta: only looking at iPhone DAR? -- So I have to make clear that we are limiting scope here
 
@@
 handle any keys or secret information besides the \ac{PIN}, and due to inline encryption in the Encryption Module,
-persistent storage never handles and \ac{plaintext} data.
+persistent storage never handles any \ac{plaintext} data.
 
@@
 The lawful access use case operates as follows: law enforcement obtains a warrant to conduct a search, including the
-suspect's mobile device. They obtain the phone, disassemble it to expose the \ac{EA} hardware interface, and begin the
-time-vaulted access request. After the lockout period is completed, the device offers its Device ID only. The law
-enforcement agency's digital forensics department digitally signs the Device ID and sends it to the device vendor,
+suspect's mobile device. They obtain the phone, disassemble it to reveal the unexposed \ac{EA} hardware interface, and
+begin the time-vaulted access request. After the lockout period is completed, the device offers its Device ID only. The
+law enforcement agency's digital forensics department digitally signs the Device ID and sends it to the device vendor,
 alongside the warrant and accompanying proof of legal authorization. The vendor's access compliance department audits
@@
 
-To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{DID}, \acf{DSK}, and \acf{DOS}.
+To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, and
+\acf{DOS}.
 
@@
 \threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
-  I1 & Attacker exposes the DSK, DID, the encrypted class key, or hashed token
+  I1 & Attacker discloses the DSK, DID, the encrypted class key, or hashed token
   & By design, none of the information stored on the phone is sensitive. The most consequential information that could
@@
 \threattableend
+
+% TODO: Platform abuser destroys the unlock artifacts database through physical attack
 

### 2020-11-24 Thesis-Chapter6-PathsAhead.tex
Editorial updates
Updates based on presentation feedback

@@
 
-These are not strong conclusions. However, when dealing with \acp{wicked-prob}, meaningful debate and discussion counts
-as progress even when it doesn't seem like it. \Acp{wicked-prob} cannot be analyzed into resolution; grasping the
-problem alone constitutes much of the work towards solving it. The conclusions above do not bring closure, but they
-should bring more clarity, and its clarity that counts.
+% FIXME: "The conclusions I draw do not settle the issue..." Make sure wording is good.
+
+The conclusions I draw do not settle the issue. However, when dealing with \acp{wicked-prob}, meaningful debate and
+discussion counts as progress even when it doesn't seem like it. \Acp{wicked-prob} cannot be analyzed into resolution;
+grasping the problem alone constitutes much of the work towards solving it. The conclusions above do not bring closure,
+but they should bring more clarity, and its clarity that counts.
 
@@
 The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
-such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is is not trying
-hard enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding
-harder will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with
-them. Tech is used to addressing tame problems, but it is time that they turn their massive financial and intellectual
-capital towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences,
-backgrounds, and values in decision making processes that steer investment and design decisions. It means creating
-products and business models that are responsive to evolving conditions. It means considering the ways their creations
-will be abused.
+such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is not trying hard
+enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding harder
+will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with them.
+Tech is used to addressing tame problems, but it is time that they turn their massive financial and intellectual capital
+towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences, backgrounds, and
+values in decision making processes that steer investment and design decisions. It means creating products and business
+models that are responsive to evolving conditions. It means considering the ways their creations will be abused.
 
@@
 \section{Paths Forward: Policymakers}
+
+% Prof. Walsh:
+% - Liked the lessons for techies; this distills the same problems for policymakers
+% - Enjoyed perspective of the "hawkish lawmaker" and how that connects to LAED Act and fallacious argmap
+% - Would love to see this in the conclusion -- tie it all together and make sure it hits home for policymakers
 

### 2020-12-06 Thesis-Chapter2-Background.tex
Address some TODOs

@@
 uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
-part of the broader data security threat model, its behavior is important. Unfortunately, the dynamics of present
-leadership and the lack of reliable data on what law enforcement agencies need make it difficult for the technical
-community to engage productively with policymakers \cite{granick_2018}.
+part of the broader data security threat model, its behavior is important. Unfortunately, the present political dynamics
+and the lack of reliable data on what law enforcement agencies need make it difficult for the technical community to
+engage productively with policymakers \cite{granick_2018}.
 
@@
 \cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
-largely see the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
-\cite{pfefferkorn_2020}.
+largely saw the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
+\cite{pfefferkorn_2020}. In response, legislators narrowed the focus of the \ac{EARNITAct}, but only after unveiling the
+\ac{LAEDAct}
 
-% NOTE: keep up to date, including the section above mentioning the pandemic
-% IDEA: create timeline! https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
+% IDEA: create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
 
@@
 %  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
-% TODO: I _really_ need to understand these and how they're used today.
+% It would be nice to really understand these and how they're used today. committee_decrypting_2018 describes them.
 
@@
 
-Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
-% TODO: Briefly introduce these. Even just "Amended the switch-board-era"...
-electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
-access to certain data as well; some terminology has aged poorly with technological evolution, making the access
-increasingly permissive \cite{shamsi_2011}.
+Included the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act. Provided explicit
+protections for private electronic communications data, and updated definitions to contemporary technology. It granted
+law enforcement explicit access to certain data as well; some terminology has aged poorly with technological evolution,
+making the access increasingly permissive \cite{shamsi_2011}.
 
@@
 
-% TODO: as of 7/10/20, this bill is still active. Follow up.
-
 }
 
-    \law{2020}{The \acf{EARNITAct}}{\cite{graham_s3398_2020}}{
+%     \law{2020}{The \acf{EARNITAct}}{\cite{graham_s3398_2020}}{
 
-Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
-Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
-as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
-bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
-combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
-popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
-\cite{lund_2020}.
+% Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
+% Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
+% as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
+% bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
+% combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
+% popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
+% \cite{lund_2020}.
 
-% TODO: Update based on new amendment.
+% % NOTE: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
+% % CSAM focused, and the LAED Act has taken over the encryption side. A wise split (not that either bill is necessarily
+% % good.)
+% % https://www.engadget.com/earn-it-act-amendments-pass-senate-judiciary-committee-165030518.html
 
-}
+% }
 
@@
 
-Introduced by the same senator as the \ac{EARNITAct}, the \ac{LAEDAct} is a direct attack on encryption. The
-\ac{LAEDAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
-developers to implement the ability to decrypt all data stored or passing through the device or software system
-``concurrently with their transmission'' ``unless the independent actions of an unaffiliated entity make it technically
-impossible to do so.'' The act neither mandates nor recommends any technical approach to providing this capability.
-
-% TODO: how does this differ from the old "Compliance with Court Orders Act" draft?
+The \ac{LAEDAct} is a direct attack on encryption and the direct successor of the Compliance with Court Orders Act. It
+was introduced by the chair of the Senate Judiciary Committee, also sponsor of the \ac{EARNITAct} and one of the
+outspoken encryption critics during the 2019 Senate hearing on encryption and \ac{EA}. The \ac{LAEDAct} specifically
+addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application developers to implement the ability
+to decrypt all data stored or passing through the device or software system ``concurrently with their transmission''
+``unless the independent actions of an unaffiliated entity make it technically impossible to do so.'' The act neither
+mandates nor recommends any technical approach to providing this capability.
 

### 2020-12-07 Thesis-Chapter3-PolicyMaking.tex
Updates based on presentation feedback
Work on wicked vs tame descriptions

@@
           explanation determines the nature of the problem's resolution.
-    \item The planner has no right to be wrong [i.e., they are liable for the consequences of their decisions].
+    \item The [decision maker] has no right to be wrong [i.e., they are liable for the consequences of their decisions].
   \end{enumerate}
@@
 analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
-conditions.
+conditions. The ``wicked'' problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means
+that the scientific approach will be effective.
 
-Several of these characteristics are results of the fact that we have no accurate predictive model of the world and
-human behavior at large. This does not need to be expounded upon.
+Several of \acp{wicked-prob}' characteristics are results of the fact that we have no accurate predictive model of the
+world and human behavior at large. This does not need to be expounded upon.
 
@@
 the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
-proposed solution demands research and context to be added to the problem definition. This feedback from proposal to
-definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals to
-a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
+proposed solution demands research and context to be fed back into the problem definition. This feedback from proposal
+to definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals
+to a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
 themselves.
@@
 
-Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
-amount will result in a certain difference in response with a certain statistical confidence. Refining the problem
-statement and combining the tame elements in a cohesive solution is the ``wicked'' part. Despite certain tame elements,
-one cannot verify the effectiveness of a plan ahead of time, port a solution from a different domain, or even
-definitively evaluate its performance after the fact.
+\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
+change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
+solution in a way that doesn't leave vast swaths of people behind.
+
+% TODO: what am I really trying to express here?
+% ... WPs quickly grow in complexity with the number of subproblems that comprise them. Refining the problem statement
+% and combining tame solutions with wicked subproblem mitigations is the truly ``wicked'' part.
+
+% Elements of the problem may be tame---for example, one can independently confirm that reducing global carbon emissions
+% by a certain amount will result in a certain difference in response with a certain statistical confidence. Other
+% elements of the problem are wicked---for example, one cannot verify the effectiveness of a plan ahead of time, port a
+% solution from a different domain, or definitively evaluate its performance after the fact.
+
+% Refining the problem statement and combining the
+% tame elements into a cohesive solution is the ``wicked'' part.
 
@@
 
+% Prof. Engelsma
+% - A big part of the "solution" is getting people together
+% - My thesis seems aimed at tech people to get them to understand and judge threats
+% - How can I take this accessible and bridge the gap for policymakers?
+
 }
@@
 
+% TODO: Dulimarta: consider changing this heading to "Proposal: Observe-Orient-Decide-Act"
 \section{Proposal: The OODA Loop for Wicked Problems}
@@
 
+% TODO: explain the icons
+
 \begin{figure}[h]

### 2020-12-08 Thesis-Chapter1-Introduction.tex
Spelling and editorial improvements
Address some TODOs
Address more todos

@@
 are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
-attempted solution has permanent and often unintended consequences, and likely exists in a pattern of chronic policy
-failure.
+attempted solution has permanent and often unintended consequences, and is likely to exist in a pattern of chronic
+policy failure.
 
@@
 attempt to end the debate, but to structure and analyze it.
-
-% TODO: change EA everywhere to "lawful access"?
 
@@
 \principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
-proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to mass surveillance,
+proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to \ac{masssurv},
 potential for abuse, international consequences, economic impact, and need for transparency.}

### 2020-12-08 Thesis-Chapter3-PolicyMaking.tex
Address more todos

@@
 
+% Elements of the problem may be tame---for example, one can develop a new renewable energy widget and another can
+% independently confirm its efficiency. Other elements of the problem are wicked---for example, one cannot predict the
+% side effects of the deploying the widget, and there is now way to know how effort spent there could have been used
+% elsewhere. \Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves
+
 \Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
 change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
-solution in a way that doesn't leave vast swaths of people behind.
+solution in a way that doesn't leave vast swaths of people behind. \Acp{wicked-prob} quickly grow in complexity with the
+number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are a tangle of contributing factors, and
+despite the presence of tame elements, they pose a large number of difficult challenges.
 
-% TODO: what am I really trying to express here?
 % ... WPs quickly grow in complexity with the number of subproblems that comprise them. Refining the problem statement
@@
 % elements of the problem are wicked---for example, one cannot verify the effectiveness of a plan ahead of time, port a
-% solution from a different domain, or definitively evaluate its performance after the fact.
+% solution from a different domain, or definitively evaluate its performance after the fact. Refining the problem
+% statement and combining the tame elements into a cohesive solution is the ``wicked'' part.
 
-% Refining the problem statement and combining the
-% tame elements into a cohesive solution is the ``wicked'' part.
+% Elements of the problem may be tame---one can independently confirm that reducing global carbon emissions by a certain
+% amount will result in a certain difference in response with a certain statistical confidence. \Acp{wicked-prob} always
+% consist of subproblems, which may be tame or wicked in and of themselves. Refining the problem statement and combining
+% the tame elements in a cohesive solution is the ``wicked'' part. Despite certain tame elements, one cannot verify the
+% effectiveness of a plan ahead of time, port a solution from a different domain, or even definitively evaluate its
+% performance after the fact.
 
@@
 
-Rozenshtein reflects on the diagnosis optimistically:
+\Ac{encryption} technology is particularly sensitive to the property that policy implementation has irreversible
+effects. Technology deployments have long tails, and attackers have the ability to record and store data for later
+analysis. In one investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile
+device that they cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the
+only ones to benefit from a policy miscalculation could negatively impact security for a very long time.
+
+Rozenshtein reflects on the \ac{wicked-prob} diagnosis optimistically:
 
@@
 
-% FIXME: include this idea somewhere in this section.
-% Note: a problem with muddling through wicked problems is that each attempted fix leads to long lasting unintended side
-% effects. This is especially true with matters of encryption, where the data could be stored for years till it can be
-% decrypted! So must be very careful to avoid solution attempts that leave long scars.
-
 \subsection{Lessons}
@@
 Let us define one strategy for handling \ac{wicked-prob} as corralling it back into the classes of problems we are
-better at dealing with. That means breaking it into tame sub-problems and developing agreeable strategies for its
+better at dealing with. That means breaking it into tame subproblems and developing agreeable strategies for its
 irreducibly wicked elements. Once that is done, we can rationally analyze and increment our way to victory. There is no
@@
 
-% TODO: Dulimarta: consider changing this heading to "Proposal: Observe-Orient-Decide-Act"
 \section{Proposal: The OODA Loop for Wicked Problems}
@@
 
-The \ac{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
+The \acl{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
 countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
@@
 % or its ideas should be at the start of the Argument Maps section.
+% PLAN: I will move this to when I talk more about argument maps. After the refactor, they won't even be introduced yet
+% here.
 
@@
 process. The result from a single iteration is an imperfect step forward based on implementing a particular proposal.
-The result from many iterations is the breakdown of the problem into sub-problems---some tame and inevitably some still
-wicked. The tame sub-problems can be addressed with the \ac{classical-method}, and the wicked sub-problems, by now
-restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}. This process is demonstrated
-in \myfig{fig-policy-ooda-process}.
+The result from many iterations is the breakdown of the problem into subproblems---some tame and inevitably some still
+wicked. The tame subproblems can be addressed with the \ac{classical-method}, and the wicked subproblems, by now
+restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.
 
-% TODO: explain the icons
+% TODO: do I need this explanation?
+
+This process is demonstrated in \myfig{fig-policy-ooda-process}. An out of control problem (the large storm) is being
+confronted. Several iterations of the \ac{OODAloop} gradually bring it down to size, spinning out tame problems (puzzle
+pieces) addressed by classical analysis and wicked but restrained problems (small storms) controlled by
+\ac{incrementalism}.
 

### 2020-12-08 Thesis-Chapter4-Arguments.tex
Clean up glossary
Updates based on presentation feedback
Address more todos

@@
 conflict if one sees radical privacy as a bad thing. The right to privacy is both a strongly held value and an enshrined
-legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The \ac{snowden}
-revelations \cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
+legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The Snowden revelations
+\cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
 technological changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Violations of privacy are
@@
 
-% FIXME: I need to address \cite{koepke_2020} in these maps.
-
 A significant portion of the disconnect between government and technical community representatives is disagreement over
@@
 ways in which privacy directly enhances safety. Data privacy makes certain crimes more difficult, such as stalking or
-identity theft. Privacy is also a public good to the extent that it precludes mass surveillance, an argument explored
-more deeply in the ``Golden Age'' side of the argument.
+identity theft. Privacy is also a public good to the extent that it precludes \ac{masssurv}, an argument explored more
+deeply in the ``Golden Age'' side of the argument.
 
@@
 investigations, but where cases hypothetically \ii{could} have been inhibited if encryption was a factor
-\cite{comey_2014}.
+\cite{comey_2014}. More recently, an 2020 report shows that law enforcement agencies currently have the upper hand on
+mobile device encryption, and are actually able to extract data from nearly every mobile phone on the market
+\cite{koepke_2020}.
 
-Citing lack of evidence only gets you so far, however. In most cases, it is impossible to know what would have happened
-in an investigation unimpeded by strong encryption. This is part of why it is a \ac{wicked-prob}. Indeed, when the
-encryption curtain is pulled back, there is often much to see. Unencrypted Facebook Messenger is has approximately 20\%
-market share in active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
-\cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted messaging service Encrochat
-exposed crime rings and led to a massive raids \cite{cox_2020}.
+Citing lack of evidence that encryption is stymieing cases only gets you so far, however. In most cases, it is
+impossible to know what would have happened in an investigation unimpeded by strong encryption. This is part of why it
+is a \ac{wicked-prob}. Indeed, when the encryption curtain is pulled back, there is often much to see. Unencrypted
+Facebook Messenger is has approximately 20\% market share in active monthly users \cite{statista_2020} but accounts for
+65\% of \ac{CSAM} reports \cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted
+messaging service Encrochat exposed crime rings and led to a massive raids \cite{cox_2020}.
 
-Aside from encryption entirely, a 2018 study found that law enforcement's primary challenges with digital evidence were
-identifying service providers and getting the sought after data from them \cite{carter_2018}. This is in the context of
-current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging Fruit,'' and also lists
-resource limitations and training needs among top priorities. Indeed, many current initiatives are underfunded. Major
-anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels, the Justice Department
-task force has produced only two of five biennial reports, and a senior executive position within the Justice Department
-was never created \cite{keller_internet_2019}. Federal entities commissioned to improve communication between law
-enforcement and industry as well as to train law enforcement on managing digital evidence are similarly underfunded
-\cite{carter_2018}.
+Aside from the challenge of encryption entirely, a 2018 study found that law enforcement's primary difficulties with
+digital evidence were identifying service providers and getting the sought after data from them \cite{carter_2018}. This
+is in the context of current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging
+Fruit,'' and also lists resource limitations and training needs among top priorities. Indeed, many current initiatives
+are underfunded. Major anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels,
+the Justice Department task force has produced only two of five biennial reports, and a senior executive position within
+the Justice Department was never created \cite{keller_internet_2019}. Federal entities commissioned to improve
+communication between law enforcement and industry as well as to train law enforcement on managing digital evidence are
+similarly underfunded \cite{carter_2018}.
 
@@
 although encryption makes certain data completely unaccessible, technological change on the whole has given authorities
-much more digital evidence than encryption has taken away. Paired with the dangers of mass surveillance, encryption is
-a necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
+much more digital evidence than encryption has taken away. Paired with the dangers of \ac{masssurv}, encryption is a
+necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
 \myfig{fig-arg-golden-age}.
@@
 The ``Golden Age'' argument revolves around surveillance, its undesirability, and technology's role in it. Cryptographer
-Phillip Rogaway writes at length about the negative impacts of mass surveillance in his essay ``The Moral Character of
+Phillip Rogaway writes at length about the negative impacts of \ac{masssurv} in his essay ``The Moral Character of
 Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway argues that ``pervasive collection \ii{itself} chills
@@
 Going back to Comey's ``Going Dark'' examples once more, each case he cited was cracked by digital evidence that would
-not have been available even twenty years ago \cite{comey_2014}. In response to the danger of mass surveillance and the
+not have been available even twenty years ago \cite{comey_2014}. In response to the danger of \ac{masssurv} and the
 scale of modern surveillance capabilities, encryption is one of the only privacy-saving tools. (Opting out may be
@@
 problems of local, state, and sometimes-surprisingly-cash-strapped federal law enforcement. The boom in the existence of
-digital evidence is clear, but as shown, law enforcement already has trouble acquiring the evidence with the help of
-tech and communications companies, and does not have the funding or capability to get it on its own \cite{carter_2018}.
-Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard evidence in cases
-where it may be suspected to exist \cite{shelton_study_2006}.
+digital evidence is clear, but as shown, law enforcement already has trouble acquiring many forms of evidence with the
+help of tech and communications companies, and does not have the funding or capability to get it on its own
+\cite{carter_2018}. Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard
+evidence in cases where it may be suspected to exist \cite{shelton_study_2006}.
 
-Still, the fact that the tools of mass surveillance are out of reach for most law enforcement must not be relied upon.
-That is, most likely, a temporary situation. A safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA}
-without mass surveillance is conceivable. \ac{EA} implementations that enable law enforcement but thwart mass
-surveillance by remote powers subvert most of the ``Golden Age'' argument. Means for \ac{EA} to do so are explored in
-future argument maps.
+Still, the fact that the tools of \ac{masssurv} are out of reach for most law enforcement must not be relied upon. That
+is, most likely, a temporary situation. Law enforcement's present ability to access mobile devices proves this point. A
+safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA}
+implementations that enable law enforcement but thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age''
+argument. \ac{EA} proposals striving for these properties are explored in future argument maps.
 
@@
 arguments spread easily. \myfig{fig-arg-fallacies} shows those used in the encryption and \ac{EA} debate.
+
+% TODO: Dulimarta: Orange/green boxes for consistency? Decide when I make a key.
 
@@
 \ac{EA}. This leaves no room for risk-based approaches, the typical security strategy outside the realm of cryptography.
-Another false dichotomy is that regulators must either allow strong cryptography or we will live in a mass surveillance
+Another false dichotomy is that regulators must either allow strong cryptography or we will live in a \ac{masssurv}
 dystopia. This one carries hints of ``slippery slope'' and appeal to emotion as well, framed as a choice between only
@@
 The availability and cost of exploits ebbs and flows. On one hand, January 2019 sported a \$2 million price on the
-exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, an October 2020 report
-shows that law enforcement agencies can extract data from nearly every mobile phone on the market for less than \$2000 a
-unit \cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law
-enforcement would face periods where it struggles for access, and would need to develop costly exploits in the lab.
-Mandatory vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical
-abuses of power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed
-\cite{keller_internet_2019} or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the
-government would actually give up expensive vulnerabilities easily when they are the pillar of their access strategy.
+exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, the October 2020 report
+detailing law enforcement's mobile device cracking capabilities estimated the per-device cost to be only \$2000 a unit
+\cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law enforcement
+would face periods where it struggles for access, and would need to develop costly exploits in the lab. Mandatory
+vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical abuses of
+power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed \cite{keller_internet_2019}
+or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the government would actually
+give up expensive vulnerabilities easily when they are the pillar of their access strategy.
 
@@
 technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
-and terror groups. While impact on civil liberties and mass surveillance depends on implementation, any proposal would
-be costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security
-is the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
+and terror groups. While impact on civil liberties and \ac{masssurv} depends on implementation, any proposal would be
+costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security is
+the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
 risk. The next map analyzes this more closely.
@@
 
-% FIXME: add note about how device security actually totally sucks
-
 Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the
@@
 requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
-\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude mass surveillance and ensure transparency to the
-device user. This approach relies on secure hardware and a method of authorizing the recovery request to the device
-itself. The authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or
-distributed systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving,
-are not totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional
-escrow \ac{EA}.
+\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency to the device
+user. This approach relies on secure hardware and a method of authorizing the recovery request to the device itself. The
+authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or distributed
+systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving, are not
+totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional escrow
+\ac{EA}.
 

### 2020-12-08 Thesis-Chapter6-PathsAhead.tex
Address more todos

@@
 stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
-ranging from identity theft, to inappropriate data mining, to mass surveillance. To the extent that encryption mitigates
+ranging from identity theft, to inappropriate data mining, to \ac{masssurv}. To the extent that encryption mitigates
 these, it increases public safety.
@@
 
-In this thesis I propose a method to contend with \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes
-collaborative debate between individuals diverse in background and values, facilitated in part by argument maps, and
-focusing on specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and
-ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and
-paths forward for both the technical and policymaking communities.
+In this thesis I propose a method to confront \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes collaborative
+debate between individuals diverse in background and values, facilitated in part by argument maps, and focusing on
+specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and ultimately performing
+threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and paths forward for both
+the technical and policymaking communities.
 
@@
 
-% TODO: make a \term for mass surveillance
-
 The additions to the threat model and the instability of present circumstances mean that additional research into
@@
 challenges that full-fledged \ac{EA} proposals must overcome, but device escrow for \acl{DAR} such as that described in
-Savage's proposal offers transparency, protection from mass surveillance, and \acl{LE} utility while making progress on
+Savage's proposal offers transparency, protection from \ac{masssurv}, and \acl{LE} utility while making progress on
 security.
-
-% FIXME: "The conclusions I draw do not settle the issue..." Make sure wording is good.
 

### 2020-12-09 Thesis-Chapter2-Background.tex
Work on wicked vs tame descriptions
Address more todos
Restructure contents of background chapter

@@
 
-% IDEA: create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
+% Could create timeline https://tex.stackexchange.com/questions/196794/how-can-you-create-a-vertical-timeline
+
+\section{Approaches to Exceptional Access}
+\label{sec-tech-approaches}
+
+% TODO: Does this read well here? Used to be after Regulatory Environment section.
+
+% There are several types of approaches:
+%  My selections here are also rather sparse. Create another appendix? Possibly create table of
+% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
+% - Allow strong encryption (anyone with the users' private keys can see it)
+% - Disallow encryption (anyone can see it)
+% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
+% - Require use of backdoor-ed algorithms, security via obscurity (anyone with the secret knowledge can see it)
+% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
+%   - Store extra in govt DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
+%   - Split the key and store it multiple places
+%   - See denning_taxonomy_1996
+% - Use some alternative path
+%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking
+
+% Lesson: the problem is _not_ cryptographically storing and splitting key recovery information. It's keeping those keys
+%   safe! If you can sit on recorded data and wait for the keys to get leaked, which they will...
+
+% Moving forward, what are the elements that are needed for a feasible EA system? (1) White-box-secure ways of producing
+% key recovery information. Do the protocols exist? EES failed. Needs to be rock solid. (2) Dispersed key recovery
+% information. (3) A robust, anti-fragile, inherently secure means of protecting those keys, and perhaps decaying them
+% over time (https://en.wikipedia.org/wiki/Data_degradation).
+
+There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
+basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
+plaintext.
+
+\newcommand{\propsstart}[0]{\begin{itemize}}
+\newcommand{\prop}[2]{ % Name, description
+    \item \textbf{#1} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
+}
+\newcommand{\propsend}{\end{itemize}}
+
+\subsection{Types of Key Escrow}
+
+The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
+information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
+classes of key escrow.
+
+\propsstart
+    \prop{Trusted-party key escrow}{
+
+Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
+of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
+a wide variety of implementation options within this category \cite{denning_taxonomy_1996}---secret or open protocol,
+hardware-enabled or software-only, single key or split key, government or third-party escrow---the foundation of the
+effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
+a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.
+
+}
+
+    \prop{Distributed key escrow}{
+
+Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
+information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
+high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
+policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
+\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
+thoroughly explored as trusted-party \ac{key-escrow}.
+
+}
+
+    \prop{Device key escrow}{
+
+A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
+trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
+``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
+\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
+\cite{ozzie_2018}.
+
+}
+\propsend
+
+\subsection{Non-Escrow Cryptographic Data Recovery}
+
+Though key escrow is the most discussed \ac{EA} approach, alternative technical means have been proposed.
+
+\propsstart
+    \prop{Translucent cryptography}{
+
+Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
+some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
+scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
+data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
+decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
+escrow do exist \cite{bellare_translucent_1996}.
+
+}
+
+    \prop{Cryptographic ``crumple zones''}{
+
+A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
+construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
+human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
+surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
+that makes them inherently recoverable only after a large up front and marginal cost.
+
+}
+\propsend
+
+\subsection{Alternative Approaches}
+\label{sec-alt-approaches}
+
+The following approaches do not use \ac{EA} at all, but are alternative technical or legal mechanisms that can be used
+towards the same goal.
+
+\propsstart
+    \prop{\Ac{lawful-hacking}}{
+
+As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Under
+\ac{lawful-hacking}, instead of having access to plaintext data via an alternative means of decryption, law enforcement
+is allowed to perform otherwise-illegal hacking activity in order to find the key, compromise a device, or intercept
+plaintext versions of the data \cite{kerr_encryption_2017}. While some see this as a viable middle ground
+\cite{nguyen_lawful_2017}, encouraging law enforcement to exploit vulnerabilities creates mis-aligned incentives. If law
+enforcement opposes strong security in general, it could foster the exploit market while costing ordinary users money
+and security \cite{soesanto_2018}.
+
+}
+
+    \prop{Compelled password disclosure}{
+
+In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
+\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
+this approach draws debates about the Fifth Amendment right against self-incrimination \cite{bittenbender_2019}
+\cite{sobel_2019}. Of course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is
+not in custody.
+
+% these look good: \cite{sacharoff_unlocking_2018}, \cite{fredona_fifth_2016}
+% this probably has something to say: kerr_encryption_2017
+
+}
+
+    \prop{Alternative paths to plaintext}{
+
+Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
+authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
+enforcement is traditionally good at---investigation and sting operations---and is used successfully. In 2017, an
+international law enforcement operation used this technique to take down a large illicit online drug market
+\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
+
+}
+\propsend
 
@@
 describes regulations around the world.
-
-% IDEA: Create an appendix to go into greater detail?
-% Could go into these, as well as the commented-out ones below
-%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
-%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
-%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
-% It would be nice to really understand these and how they're used today. committee_decrypting_2018 describes them.
 
@@
 
-% % NOTE: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
+% % Note: based on the amendment's modifications, it is no longer about encryption. This has become more moderation and
 % % CSAM focused, and the LAED Act has taken over the encryption side. A wise split (not that either bill is necessarily
@@
 
-% Prof. Walsh:
-% Due to the dynamic nature of future judicial rulings that may expand or retract the application of exceptional access
-% it is beyond the scope of this research to comment further...
-
 \subsection{Regulations around the World}
@@
 
-% IDEA: put this in an appendix too? Or its own chapter? Too much detail for the background section.
+% Put this in an appendix too? Or its own chapter? Too much detail for the background section.
 %
@@
 
-% TODO: it is very important to know who these requirements apply to. Traditional telecoms, obviously. Facebook and
-% Google? Signal and WhatsApp? Telecoms, not platforms or apps. This is the big differentiator. This can be expanded
-% upon when I talk a bit about pen registers, etc.
+% It is very important to know who these requirements apply to. This is for telecoms, not Facebook, Google, Signal, or
+% WhatsApp. That's a little unclear, but the vagueness is a necessity, since I don't actually know this very well.
+% Consider removing the internation section altogether because it's complicated and I'm not sure I'm adding a whole lot
+% of value.
 
@@
 \countriesend
-
-\section{Approaches to Exceptional Access}
-\label{sec-tech-approaches}
-
-% FIXME: Engelsma: Consider moving this right after "encryption basics", as I have referred to many of these approaces
-% already.
-
-% There are several types of approaches:
-% IDEA: my selections here are also rather sparse. Create another appendix? (I love work). Possibly create table of
-% properties below, similar to denning_taxonomy_1996. savage_lawful_2018 has some good citations to mine.
-% - Allow strong encryption (anyone with the users' private keys can see it)
-% - Disallow encryption (anyone can see it)
-% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
-% - Require use of backdoor-ed algorithms, security via obscurity (anyone with the secret knowledge can see it)
-% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
-%   - Store extra in govt DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
-%   - Split the key and store it multiple places
-%   - See denning_taxonomy_1996
-% - Use some alternative path
-%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking
-
-% Lesson: the problem is _not_ cryptographically storing and splitting key recovery information. It's keeping those keys
-%   safe! If you can sit on recorded data and wait for the keys to get leaked, which they will...
-
-% Moving forward, what are the elements that are needed for a feasible EA system? (1) White-box-secure ways of producing
-% key recovery information. Do the protocols exist? EES failed. Needs to be rock solid. (2) Dispersed key recovery
-% information. (3) A robust, anti-fragile, inherently secure means of protecting those keys, and perhaps decaying them
-% over time (https://en.wikipedia.org/wiki/Data_degradation).
-
-There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
-basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
-plaintext.
-
-\newcommand{\propsstart}[0]{\begin{itemize}}
-\newcommand{\prop}[2]{ % Name, description
-    \item \textbf{#1} \nopagebreak
-
-    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
-}
-\newcommand{\propsend}{\end{itemize}}
-
-\subsection{Types of Key Escrow}
-
-The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
-information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
-classes of key escrow.
-
-\propsstart
-    \prop{Trusted-party key escrow}{
-
-Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
-of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
-a wide variety of implementation options within this category \cite{denning_taxonomy_1996}---secret or open protocol,
-hardware-enabled or software-only, single key or split key, government or third-party escrow---the foundation of the
-effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
-a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.
-
-}
-
-    \prop{Distributed key escrow}{
-
-Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
-information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
-high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
-policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
-\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
-thoroughly explored as trusted-party \ac{key-escrow}.
-
-}
-
-    \prop{Device key escrow}{
-
-A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
-trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
-``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
-\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
-\cite{ozzie_2018}.
-
-}
-\propsend
-
-\subsection{Non-Escrow Cryptographic Data Recovery}
-
-Though key escrow is the most discussed \ac{EA} approach, alternative technical means have been proposed.
-
-\propsstart
-    \prop{Translucent cryptography}{
-
-Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
-some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
-scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
-data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
-decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
-escrow do exist \cite{bellare_translucent_1996}.
-
-}
-
-    \prop{Cryptographic ``crumple zones''}{
-
-A recent proposal adopts the idea of crumple zones from automotive engineering---``in an emergency situation the
-construction should break a little bit in order to protect the integrity of the system as a whole and the safety of its
-human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
-surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
-that makes them inherently recoverable only after a large up front and marginal cost.
-
-}
-\propsend
-
-\subsection{Alternative Approaches}
-\label{sec-alt-approaches}
-
-The following approaches do not use \ac{EA} at all, but are alternative technical or legal mechanisms that can be used
-towards the same goal.
-
-\propsstart
-    \prop{\Ac{lawful-hacking}}{
-
-As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Under
-\ac{lawful-hacking}, instead of having access to plaintext data via an alternative means of decryption, law enforcement
-is allowed to perform otherwise-illegal hacking activity in order to find the key, compromise a device, or intercept
-plaintext versions of the data \cite{kerr_encryption_2017}. While some see this as a viable middle ground
-\cite{nguyen_lawful_2017}, encouraging law enforcement to exploit vulnerabilities creates mis-aligned incentives. If law
-enforcement opposes strong security in general, it could foster the exploit market while costing ordinary users money
-and security \cite{soesanto_2018}.
-
-}
-
-    \prop{Compelled password disclosure}{
-
-In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
-\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
-this approach draws debates about the Fifth Amendment right against self-incrimination \cite{bittenbender_2019}
-\cite{sobel_2019}. Of course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is
-not in custody.
-
-% these look good: \cite{sacharoff_unlocking_2018}, \cite{fredona_fifth_2016}
-% this probably has something to say: kerr_encryption_2017
-
-}
-
-    \prop{Alternative paths to plaintext}{
-
-Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
-authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
-enforcement is traditionally good at---investigation and sting operations---and is used successfully. In 2017, an
-international law enforcement operation used this technique to take down a large illicit online drug market
-\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
-
-}
-\propsend
-
-% FIXME: Separate out into separate, short chapter, probably after the wicked problems chapter. Also adds natural place
-% to introduce my more "advanced" DFD syntax.
-
-% TODO: include a more descriptive key to understand argument maps.
-
-% - Prof. Dulimarta: Could you apply fuzzy logic to the argument map? Quantify it? -- I think this would run into limits
-%   once you start comparing unquantifiable things, but it could be used to suss out the valid arguments, and would
-%   probably be really interesting otherwise.
-
-% Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own
-% (short) chapter.
-% Kalafut: Argument maps are used in chapter 4. I'm not sure that you need to mention argdown at all. If you do, I would
-% only do so in the context of stating that it is the tool you are using to create argument maps.
-
-\section{Argument Maps}
-\label{sec-arg-maps-intro}
-
-The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
-help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create \acp{argmap}.
-\Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
-conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
-\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
-reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
-software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
-\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
-generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.
-
-Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo}
-shows the basic structure of arguments. Statements are positions on issues. Some statements are based on values that
-cannot be addressed through argument. Statements and arguments may support or attack one another, and arguments support
-or undercut other arguments.
-
-\begin{figure}[ht]
-    \centering\CaptionFontSize
-    \myincludeargument{demo}
-    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
-    \label{fig-args-demo}
-\end{figure}
-
-\section{Threat Models and Data Flow Diagrams}
-\label{sec-threat-model-intro}
-
-Threat modeling is an important step in analyzing the security at the systems level. Using models abstracts away fine
-details and focuses on the architecture, processes, and dataflows in a system. Models assist in the understanding of
-current systems and the prevention of problems in new systems, both of which are important when facing the prospect of
-designing an \ac{EA} mechanism built on top of a complex and aging technology stack.
-
-Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
-\cite{shostack_threat_2014}. Answering the first question accurately is crucial, particularly for the field of
-cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. Answering the second
-question depends on the types of attackers to worry about, and determines the scope of threats to be considered. ``What
-are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go wrong?''
-cannot be faithfully answered without knowing what is being built.
-
-\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
-external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
-flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
-primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
-\acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.
-
-\begin{figure}[ht]
-    \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/DFD-demo.png}
-    \caption[\Acs{DFD} Demo]{A demonstrative example of a \acf{DFD}.}
-    \label{fig-dfd-demo}
-\end{figure}

### 2020-12-09 Thesis-Chapter3-PolicyMaking.tex
Restructure contents of background chapter

@@
 
-Recall from \mysec{sec-arg-maps-intro} that \acp{argmap} resulted from the same research that produced the concept of
-\acp{wicked-prob}. Though formal research on the technique is thin, \acp{argmap} have proven helpful in simplifying
-complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate by helpfully slowing down
-discussion, depersonalizing conflict, and lending structure and rhythm \cite{dutoit_hypermedia_2006}. This makes them
-well suited as a tool for conducting the collaborative debate that helps problem solvers with disparate views narrow in
-on a problem definition and concrete proposals on which they can agree.
-
-% TODO: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this paragraph
-% or its ideas should be at the start of the Argument Maps section.
-% PLAN: I will move this to when I talk more about argument maps. After the refactor, they won't even be introduced yet
-% here.
-
 \section{Summary}

### 2020-12-09 Thesis-Chapter4-Tools.tex
Restructure contents of background chapter

\chapter{Analysis Tools}
\label{chap-tools}

% FIXME: include a more descriptive key to understand argument maps.

% - Prof. Dulimarta: Could you apply fuzzy logic to the argument map? Quantify it? -- I think this would run into limits
%   once you start comparing unquantifiable things, but it could be used to suss out the valid arguments, and would
%   probably be really interesting otherwise.

% Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own
% (short) chapter.
% Kalafut: Argument maps are used in chapter 4. I'm not sure that you need to mention argdown at all. If you do, I would
% only do so in the context of stating that it is the tool you are using to create argument maps.

\section{Argument Maps}
\label{sec-arg-maps-intro}

The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create \acp{argmap}.
\Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.

Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo}
shows the basic structure of arguments. Statements are positions on issues. Some statements are based on values that
cannot be addressed through argument. Statements and arguments may support or attack one another, and arguments support
or undercut other arguments.

\begin{figure}[ht]
    \centering\CaptionFontSize
    \myincludeargument{demo}
    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
    \label{fig-args-demo}
\end{figure}

%%%%%%%%%%%%%%%%% Hack job

Recall from \mysec{sec-arg-maps-intro} that \acp{argmap} resulted from the same research that produced the concept of
\acp{wicked-prob}. Though formal research on the technique is thin, \acp{argmap} have proven helpful in simplifying
complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate by helpfully slowing down
discussion, depersonalizing conflict, and lending structure and rhythm \cite{dutoit_hypermedia_2006}. This makes them
well suited as a tool for conducting the collaborative debate that helps problem solvers with disparate views narrow in
on a problem definition and concrete proposals on which they can agree.

% FIXME: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this
% paragraph or its ideas should be at the start of the Argument Maps section.
% PLAN: I will move this to when I talk more about argument maps. After the refactor, they won't even be introduced yet
% here.

\section{Threat Models and Data Flow Diagrams}
\label{sec-threat-model-intro}

Threat modeling is an important step in analyzing the security at the systems level. Using models abstracts away fine
details and focuses on the architecture, processes, and dataflows in a system. Models assist in the understanding of
current systems and the prevention of problems in new systems, both of which are important when facing the prospect of
designing an \ac{EA} mechanism built on top of a complex and aging technology stack.

Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
\cite{shostack_threat_2014}. Answering the first question accurately is crucial, particularly for the field of
cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. Answering the second
question depends on the types of attackers to worry about, and determines the scope of threats to be considered. ``What
are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go wrong?''
cannot be faithfully answered without knowing what is being built.

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
\acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.

\begin{figure}[ht]
    \centering\CaptionFontSize
    \includegraphics[width=\linewidth]{dfds/build/DFD-demo.png}
    \caption[\Acs{DFD} Demo]{A demonstrative example of a \acf{DFD}.}
    \label{fig-dfd-demo}
\end{figure}

### 2020-12-09 Thesis-Chapter6-ThreatModel.tex
Restructure contents of background chapter

--- 2020-11-24 Thesis-Chapter5-ThreatModel.tex
@@
 This chapter consists of a threat model and analysis of a specific device key escrow \ac{EA} proposal, Stefan Savage's
-``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}.
-Lacking a concise name, this proposal will be referred to by its acronym, \ldawmsR. This chapter assumes that the
-proposal is accompanied by some legal framework regulating its application and use, though it is not the focus of this
-analysis, and nothing is assumed about it in particular. Note that the creators of the proposal intentionally chose the
-phrase ``lawful access'' over ``\acl{EA}'' in order to emphasize the role of legal process and dispel the negative
-stigma that accompanies \ac{EA}. I refer to \ldawmsr as an \ac{EA} proposal, but use the term entirely in the spirit of
-their paper.
-
-% TODO: remove this paragraph if I end up changing EA to LA
-
-% TODO: Dulimarta: only looking at iPhone DAR? -- So I have to make clear that we are limiting scope here
+``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
+proposal is specific to \acl{DAR} for mobile devices. Lacking a concise name, this proposal will be referred to by its
+acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
+application and use, though it is not the focus of this analysis, and nothing is assumed about it in particular. Note
+that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
+emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
+\ac{EA} proposal, but use the term entirely in the spirit of their paper.
 
@@
 
-\modelitem{Protection of civil liberties}{The system should include measures to prevent mass surveillance and abuse by
+\modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
 political regimes that violate civil rights. This includes technologically enforcing the fundamental right to privacy
@@
 
-% FIXME: add current hacking techniques.
-
 \modelitem{\ac{EA} workarounds}{The user is capable of evading \ac{EA} mechanisms by encrypting data at a higher level
@@
 \modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
-warrant is valid. Additionally, all forms of surveillance, even when protected from mass surveillance, require
-oversight. Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA}
-proposals. However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
+warrant is valid. Additionally, all forms of surveillance, even when protected from \ac{masssurv}, require oversight.
+Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
+However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
 
@@
 the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
-\ldawmsR. \myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to their
-public documentation \cite{apple_2020}.
+\ldawmsR. As this analysis is limited to a specific proposal, it is also limited to this reference architecture.
+\myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to Apple's public
+documentation \cite{apple_2020}.
 
@@
       the security of the underlying cryptographic protocols. Physical possession and partial disassembly is required,
-      which introduces practical barriers to mass surveillance.
+      which introduces practical barriers to \ac{masssurv}.
     \\ \hline
@@
       72 hours).''
-    & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for mass surveillance.
+    & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for \ac{masssurv}.
     \\ \hline
@@
 also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
-detectable changes, mitigating both mass surveillance and common criminal abuse.
+detectable changes, mitigating both \ac{masssurv} and common criminal abuse.
 
@@
 
-To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, and
-\acf{DOS}.
-
-% TODO: Check list of acronyms up to date here ^^
+To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, \acf{SN},
+and \acf{DOS}.
 
@@
   & Verify integrity of new data before entering it into the database, and disallow overwriting entirely.
+  \\ \hline
+  D2 & Unlock artifacts recovery
+  & Determined adversaries could physically destroy the medium storing the DID-DSK database.
+  & The \ac{HSM} should be in a physically secure and guarded facility, and the storage medium should be stable and
+    persistent.
 \threattableend
-
-% TODO: Platform abuser destroys the unlock artifacts database through physical attack
 
@@
 
-\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, and \ii{E1} \ii{E2} are each low risk either because they
-fail to produce useful results, they can be well mitigated, or they are a threat that is already present in
-state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and mitigate the
-risk posed by these threats as part of the design process, but they do not represent fatal flaws to a design. Except in
-combination with other threats, they are not worthy more concern here.
+\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low risk either
+because they fail to produce useful results, they can be reasonably mitigated, or they are a threat that is already
+present in state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and
+mitigate the risk posed by these threats as part of the design process, but they do not represent fatal flaws to a
+design. Except in combination with other threats, they are not worthy more concern here.
 
@@
 It is important to give \ldawmsr (and similar \ac{DAR} \ac{EA} proposals) its due. The worst case scenario still
-precludes mass surveillance and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
+precludes \ac{masssurv} and includes non-repudiation that access occurred. This is a significant gain for \ac{EA}
 proposals in terms of transparency and protection of civil liberties.
@@
 
-\paragraph*{Protection of civil liberties} This goal was defined as defending against mass surveillance and abuse by
-authoritarian regimes. This proposal does an excellent job of protecting against mass surveillance due to its
-physicality. Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian
-regimes. Prior to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign
-countries. After a leak, these regimes could get into any device they want. If one lived in a country that was willing
-to take and destroy their devices to read their contents without legal due process, this would be an unwelcome
-technology.
+\paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
+authoritarian regimes. This proposal does an excellent job of protecting against \ac{masssurv} due to its physicality.
+Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior
+to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After
+a leak, these regimes could get into any device they want. If one lived in a country that was willing to take and
+destroy their devices to read their contents without legal due process, this would be an unwelcome technology.
 

### 2020-12-09 Thesis-Chapter4-Tools.tex
Work on DFD section and normalize captions

@@
     \myincludeargument{demo}
-    \caption[Argdown Demo]{A demonstrative example of argument maps with Argdown.}
+    \caption{A demonstrative example of argument maps with Argdown}
     \label{fig-args-demo}
@@
 flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
-primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In \mychap{chap-threatmodel} I create
-\acp{DFD} in this style to analyze the threats already present and those that \ac{EA} would introduce.
+primary concern. \myfig{fig-dfd-dh-traditional} shows the basic elements of a \ac{DFD} of a simplified Diffie-Hellman
+key exchange.
 
-\begin{figure}[ht]
+\begin{figure}[p]
     \centering\CaptionFontSize
-    \includegraphics[width=\linewidth]{dfds/build/DFD-demo.png}
-    \caption[\Acs{DFD} Demo]{A demonstrative example of a \acf{DFD}.}
-    \label{fig-dfd-demo}
+    \includegraphics[width=0.8\linewidth]{dfds/build/diffie-hellman-traditional.png}
+    \caption{A \acf{DFD} for a simplified Diffie-Hellman key exchange}
+    \label{fig-dfd-dh-traditional}
 \end{figure}
+
+\Acp{DFD} show what data is transferred via labels on the flows. However, labels can be insufficient in complex
+\acp{DFD}, and the method has no syntax for computation or cryptographic protocols. In order to maximize their
+communicative ability in the context of encryption and \acl{EA} schemes, I introduce the additional syntax show in
+\myfig{fig-dfd-dh-updated}.
+
+\begin{figure}[p]
+    \centering\CaptionFontSize
+    \includegraphics[width=\linewidth]{dfds/build/diffie-hellman.png}
+    \caption{A Diffie-Hellman key exchange illustrated with additional syntax}
+    \label{fig-dfd-dh-updated}
+\end{figure}
+
+The new syntax illustrates the data in more detail as it is stored, transferred, and operated on. By illustrating the
+data itself, it eliminates the need for labels on lines and can be used to communicate entire protocols in a single
+% diagram. \myfig{fig-dfd-key} contains the full set of symbols. In \mychap{chap-threatmodel} I create \acp{DFD} in this
+style to analyze the threats already present and those that \ac{EA} would introduce.
+
+% \begin{figure}[p]
+%     \centering\CaptionFontSize
+%     \includegraphics[width=\linewidth]{dfds/build/dfd-key.png}
+%     \caption{A Diffie-Hellman key exchange illustrated with additional syntax}
+%     \label{fig-dfd-key}
+% \end{figure}

### 2020-12-09 Thesis-Chapter4-Tools.tex
Create DFD key

@@
 data itself, it eliminates the need for labels on lines and can be used to communicate entire protocols in a single
-% diagram. \myfig{fig-dfd-key} contains the full set of symbols. In \mychap{chap-threatmodel} I create \acp{DFD} in this
+diagram. \myfig{fig-dfd-key} contains the full set of symbols. In \mychap{chap-threatmodel} I create \acp{DFD} in this
 style to analyze the threats already present and those that \ac{EA} would introduce.
 
-% \begin{figure}[p]
-%     \centering\CaptionFontSize
-%     \includegraphics[width=\linewidth]{dfds/build/dfd-key.png}
-%     \caption{A Diffie-Hellman key exchange illustrated with additional syntax}
-%     \label{fig-dfd-key}
-% \end{figure}
+\begin{figure}[p]
+    \centering\CaptionFontSize
+    \includegraphics[width=0.85\linewidth]{dfds/build/dfd-key.png}
+    \caption{Expanded set of DFD symbols}
+    \label{fig-dfd-key}
+\end{figure}

### 2020-12-09 Thesis-Chapter1-Introduction.tex
Spruce up title, clean up Tools chapter

@@
 This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce \acp{argmap} and threat
-models using data flow diagrams. In the following chapters, I analyze strategies for tacking \acp{wicked-prob}, map the
-\acl{EA} arguments in detail, and demonstrate the security risks with data flow diagrams and threat analysis, focusing
-on one particular EA proposal. The thesis concludes with research and policy paths forward in light of valid arguments
-and threats.
+brief U.S.--focused history of encryption regulation, technology, and conflict, I analyze strategies for tacking
+\acp{wicked-prob} and introduce \acp{argmap} and threat modeling with data flow diagrams. In the following chapters, I
+map the \acl{EA} arguments in detail and demonstrate the security risks with data flow diagrams and threat analysis,
+focusing on one particular EA proposal. The thesis concludes with research and policy paths forward in light of valid
+arguments and threats.

### 2020-12-09 Thesis-Chapter2-Background.tex
Spruce up title, clean up Tools chapter

@@
 The problem of strong encryption and \acl{EA} exists in a technical, historical, and regulatory context. This chapter
-introduces cryptography, summarizes the history of encryption and the ``crypto wars,'' lists relevant encryption
-regulations, and lists prominent technical \ac{EA} proposals. It then introduces \acp{argmap} as a logical modeling
-tool, and introduces threat modeling with data flow diagrams.
+introduces cryptography, summarizes the history of encryption and the ``crypto wars,'' lists prominent technical \ac{EA}
+proposals, and lastly overviews relevant laws and regulations.
 

### 2020-12-09 Thesis-Chapter3-PolicyMaking.tex
Work on DFD section and normalize captions
Spruce up title, clean up Tools chapter

@@
   \includegraphics[width=\linewidth]{dfds/build/OODA-classical.png}
-  \caption[The Classical Analytic Method]{The Classical Analytic Method}
+  \caption{The Classical Analytic Method}
   \label{fig-classical-method}
@@
   \includegraphics[width=0.55\linewidth]{dfds/build/OODA-muddling.png}
-  \caption[The Incrementalist Method]{The Incrementalist Method}
+  \caption{Incrementalism}
   \label{fig-muddling-through}
@@
     \includegraphics[width=\linewidth]{dfds/build/OODA-vanilla.png}
-    \caption[The OODA Loop]{The OODA Loop}
+    \caption{The OODA Loop}
     \label{fig-ooda-loop}
@@
 of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
-solutions, and crystallizes discussion around specific proposals. \Acp{argmap} are one tool that may facilitate this
-process. The result from a single iteration is an imperfect step forward based on implementing a particular proposal.
-The result from many iterations is the breakdown of the problem into subproblems---some tame and inevitably some still
-wicked. The tame subproblems can be addressed with the \ac{classical-method}, and the wicked subproblems, by now
-restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.
+solutions, and crystallizes discussion around specific proposals. The result from a single iteration is an imperfect
+step forward based on implementing a particular proposal. The result from many iterations is the breakdown of the
+problem into subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
+\ac{classical-method}, and the wicked subproblems, by now restrained under a sound and agreeable strategy, can be
+addressed with \ac{incrementalism}.
 
@@
 
-In the remainder of this thesis, I will apply this strategy through the use of \acp{argmap} to document elements of the
-debate in \mychap{chap-arguments} and through creation of a threat model against which I analyze one specific proposal
-in \mychap{chap-threatmodel}.
+In the remainder of this thesis, I apply this strategy by documenting the debate in depth in \mychap{chap-arguments} and
+through creation of a threat model against which I analyze one specific proposal in \mychap{chap-threatmodel}.

### 2020-12-09 Thesis-Chapter4-Tools.tex
Spruce up title, clean up Tools chapter

@@
 
-% FIXME: include a more descriptive key to understand argument maps.
-
-% - Prof. Dulimarta: Could you apply fuzzy logic to the argument map? Quantify it? -- I think this would run into limits
-%   once you start comparing unquantifiable things, but it could be used to suss out the valid arguments, and would
-%   probably be really interesting otherwise.
-
-% Kalafut: I'm not sure that 2.5 and 2.6 really fit well as subsections of 2. Consider separating them into their own
-% (short) chapter.
-% Kalafut: Argument maps are used in chapter 4. I'm not sure that you need to mention argdown at all. If you do, I would
-% only do so in the context of stating that it is the tool you are using to create argument maps.
+Encryption policy arguments and \acl{EA} threat models are complex. \Acp{argmap} and \aclp{DFD} two methodological tools
+to structure analysis and reduce complexity. This chapter introduces these tools and their syntax as used in
+\mychap{chap-arguments} and \mychap{chap-threatmodel}.
 
@@
 conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
-\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
-reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
-software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
-\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
-generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.
-
-Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo}
-shows the basic structure of arguments. Statements are positions on issues. Some statements are based on values that
-cannot be addressed through argument. Statements and arguments may support or attack one another, and arguments support
-or undercut other arguments.
-
-\begin{figure}[ht]
-    \centering\CaptionFontSize
-    \myincludeargument{demo}
-    \caption{A demonstrative example of argument maps with Argdown}
-    \label{fig-args-demo}
-\end{figure}
-
-%%%%%%%%%%%%%%%%% Hack job
-
-Recall from \mysec{sec-arg-maps-intro} that \acp{argmap} resulted from the same research that produced the concept of
 \acp{wicked-prob}. Though formal research on the technique is thin, \acp{argmap} have proven helpful in simplifying
@@
 
-% FIXME: Kalafut: It's not clear to me that this paragraph is a good fit here. It is good content. Perhaps this
-% paragraph or its ideas should be at the start of the Argument Maps section.
-% PLAN: I will move this to when I talk more about argument maps. After the refactor, they won't even be introduced yet
-% here.
+In 1970 Rittel's research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the reasoned
+approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some software
+tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
+\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
+generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}. Argdown is used
+in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate.
 
-\section{Threat Models and Data Flow Diagrams}
+\myfig{fig-args-key} shows the basic structure of arguments. Statements are positions, values, and assertions. The
+``Proposals,'' ``Statements,'' ``Fallacies,'' and ``Values'' in the figure are classes of statements in this sense. The
+color of a statement's border indicates the class to which it belongs. Arguments are premise-conclusion structures that
+relate to statements. Arguments as classified as ``anti-EA,'' ``pro-EA,'' ``neutral,'' or ``fallacious.'' The color of
+an argument's solid background indicates the class to which it belongs. Statements and arguments may support, attack, or
+undercut one another. Supporting relationships have green arrows; attacking, red; and undercutting, purple.
+
+\begin{figure}[h]
+    \centering\CaptionFontSize
+    \myincludeargument{key}
+    \caption{A demonstrative example of argument maps with Argdown}
+    \label{fig-args-key}
+\end{figure}
+
+\section{Threat Modeling with Data Flow Diagrams}
 \label{sec-threat-model-intro}
@@
 
-\begin{figure}[p]
+\begin{figure}[h]
     \centering\CaptionFontSize

### 2020-12-10 Thesis-Abstract.tex
Editorial improvements

--- 2020-09-26 Thesis-Abstract.tex
@@
-Public debate has resumed over whether encryption systems should support alternative means of decryption intended for
-law enforcement use, called exceptional access (EA). Rather than a renege on a solemn promise made at the end of the
-1990s ``crypto war,'' this represents a valid reassessment of optimal policy in light of changing circumstances.
-Achieving proper balance between privacy and access in the context of constantly changing society and technology is a
-wicked problem that has and will evade a permanent solution. As policymakers consider next steps, it behooves the
-technical community to stay engaged. Although the introduction of EA would inevitably introduce risk, the quality of the
-technical and regulatory approach can make a substantial difference. Furthermore, if one considers ham-fisted
-legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce overall
-risk.
+Public debate has resumed on the topic of exceptional access (EA), which is alternative means of decryption intended for
+law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of the 1990s
+``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing circumstances. The
+imbalance between privacy, access, and security in the context of constantly changing society and technology is a wicked
+problem that has and will continue to evade a permanent solution. As policymakers consider next steps, it is necessary
+that the technical community to remain engaged. Although the introduction of EA would inevitably add risk, the quality
+of the technical and regulatory approach can make a substantial difference. Furthermore, if one considers hard-line
+legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce risk
+overall.
 
-The root of the conflict lies in cryptography's dual role as enabler of radical privacy and cornerstone of security. The
-emergence of strong encryption triggered the first crypto war, and its proliferation is causing the second. In the face
-of both polarized and conciliatory voices, I analyze strategies that do and do not work on wicked problems and promote
-an iterative approach to the problem of encryption and EA. Along the way I illustrate the components of the debate in
-argument maps and demonstrate the security risks with data flow diagrams, making clear distinctions for the separate
-problems of EA for data in motion and data at rest.
+The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
+security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
+In response to both polarized and conciliatory voices, this paper analyzes strategies for confronting wicked problems
+and proposes an iterative approach to the case of encryption and EA. Along the way, it illustrates the components of the
+debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing on one
+EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''

### 2020-12-10 Thesis-Chapter1-Introduction.tex
Editorial improvements

@@
 
-Certain policy issues are \acp{wicked-prob}. Originally coming from the field of design theory, the term
-``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
+Certain policy issues can be described as ``\acp{wicked-prob}.'' Originally coming from the field of design theory, the
+term ``\ac{wicked-prob}'' uses ``wicked'' not in the moral sense, but in the malignant, vicious, and tricky sense
 \cite{rittel_dilemmas_1973}. Unlike their ``tame'' counterparts which science, engineering, and traditional policymaking
-are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, or measurements. Each
+are well equipped to answer, \acp{wicked-prob} lack clear formulations, causes, resolutions, and measurements. Each
 attempted solution has permanent and often unintended consequences, and is likely to exist in a pattern of chronic
@@
 
-The conflict at the heart of the encryption and \ac{EA} debate is that the same cryptographic and design principles that
-underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
+The conflict at the heart of the encryption and \ac{EA} debate is this: the same cryptographic and design principles
+that underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
 foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
-communications and storage provides clear benefits and should be welcomed by all. In a world where information security
-is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
-privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing
-\cite{cox_2020} \cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk
-should not be ignored.
+communications and storage provides clear benefits. In a world where information security is frightfully poor yet
+increasingly important, the necessity of strong encryption cannot be understated. However, the privacy afforded by
+certain encryption technologies hampers law enforcement investigations and hides wrongdoing \cite{cox_2020}
+\cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk should not be
+ignored.
 
-The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
-access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}. Human rights
-groups and technical leaders counter that a weakened encryption environment would fatally compromise privacy and
-security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate continues,
-sides have become entrenched, and both sides have engaged in disingenuous attacks against their opponents.
+The conversation regarding encryption has reached a stalemate. Governments and law enforcement agencies frequently cite
+the need to access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}.
+Human rights groups and technical leaders counter that a weakened encryption environment would fatally compromise
+privacy and security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate
+continues, many on both sides have become entrenched in their positions and have engaged in disingenuous attacks against
+their opponents.
 
-In the long term, increased regulation of the tech industry in general is inevitable, and regulation of encryption is
-possible. Despite strong encryption's benefits, regulatory interest in the subject has not subdued. The form that
-regulation will take in the field of encryption depends on the good faith efforts made to equitably balance all the
-benefits and risks involved in deploying a cryptographic system. If the technical community acts, it can lead and shape
-legislation instead of being subjected to it.
+In the long term, increased regulation of the tech industry is inevitable and regulation of encryption is possible.
+Despite the benefits of strong encryption, regulatory interest in the subject has not subdued. The form that regulation
+will take in the field of encryption depends on the good faith efforts made to equitably balance the benefits and risks
+involved in deploying a cryptographic system. If the technical community acts, it can lead and shape legislation rather
+than be subjected to it.
 
 Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
-community fails to act but regulators move forward, everyone may become subject to technically misguided laws that do
-great harm. Bad \ac{EA} policy is just as much a threat as poor passwords. Due to this threat, as long as lawmakers
-continue pursuing \ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and
-respond to the arguments being presented.
+community fails to act but regulators move forward, everyone may become subject to technically misguided, harmful laws.
+Bad policy is just as much a threat as weak passwords. Due to this threat, as long as lawmakers continue pursuing
+\ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and respond to the
+presented arguments.
 
@@
 This paper accepts and builds on the following premises. These premises are not principles for potential \ac{EA}
-designs, but the starting place for discussing the \ac{EA} debate.
+designs, but the foundation for discussing the \ac{EA} debate.
 
@@
 
-\principle{Cybersecurity is critical}{With current levels of reliance on computer information systems, cybersecurity is
+\principle{Cybersecurity is critical}{Due to modern culture's reliance on computer information systems, cybersecurity is
 critical to the wellbeing of society. Two important elements of cybersecurity are cryptography and system architecture.
-Policy which supports security does not undercut cryptographic integrity or require high-risk architectures.}
+Policy that supports security does not undercut cryptographic integrity or require high-risk architectures.}
 
 \principle{Absolute privacy is not an absolute right}{Certain rights supersede the legitimate claims of government, but
-privacy in all contexts is not one of them. While individuals under a limited government are entitled to an expectation
-of privacy, an absolute right to privacy does not apply in all circumstances. Investigators should have access to some
-classes of data. In particular, access to certain classes of digital data is important today and will be increasingly
-important in the future.}
+privacy in all contexts does not. While individuals under a limited government are entitled to an expectation of
+privacy, an absolute right to privacy does not apply in all circumstances. Investigators should have access to some
+classes of data. In particular, access to certain classes of digital data is important today and will become
+increasingly important in the future.}
 
-\principle{\Ac{EA} is an inherently complex problem}{The factors factors that make \ac{EA} a \ac{wicked-prob} mean that
-proposed solutions must be analyzed at many levels. Some these factors are \ac{EA}'s relation to \ac{masssurv},
-potential for abuse, international consequences, economic impact, and need for transparency.}
+\principle{\Ac{EA} is an inherently complex problem}{The factors that make \ac{EA} a \ac{wicked-prob} require proposed
+solutions to be analyzed at many levels. These factors include \ac{EA}'s relation to \ac{masssurv}, potential for abuse,
+international consequences, economic impact, and need for transparency.}
 
-\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
-optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
-destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
-right balance of risk given the threats under consideration---including the threat of regulatory action.}
+\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions; they do not even have
+verifiably optimal solutions. Therefore, we cannot use perfection as the standard. To quote a famous security axiom,
+insecurity cannot be destroyed, only moved around \cite{nrc_schneider_1999}. The \Ac{EA} problem, like all security
+problems, involves finding the right balance of risk given the threats under consideration---including the threat of
+regulatory action.}
 
@@
 
-This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
-brief U.S.--focused history of encryption regulation, technology, and conflict, I analyze strategies for tacking
+This thesis aims to clarify the arguments of the debate and the nature of the technical problem. After reproducing a
+brief U.S.--focused history of encryption regulation, technology, and conflict, I analyze strategies for tackling
 \acp{wicked-prob} and introduce \acp{argmap} and threat modeling with data flow diagrams. In the following chapters, I
 map the \acl{EA} arguments in detail and demonstrate the security risks with data flow diagrams and threat analysis,
-focusing on one particular EA proposal. The thesis concludes with research and policy paths forward in light of valid
-arguments and threats.
+focusing on one EA proposal in particular. The thesis concludes with paths forward for research and policy that take the
+arguments and threats discussed into consideration.

### 2020-12-10 Thesis-Chapter2-Background.tex
Editorial improvements

@@
 
-\Ac{cryptography} is the study of techniques for communicating secretly in the presence of third parties. This is
+\Ac{cryptography} is the study of techniques used to communicate securely in the presence of third parties. This is
 performed by using a \ac{cipher} to translate between \ac{plaintext} and \ac{ciphertext}. A cipher is a tool or
-algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data. The
-processes for translating from plaintext to ciphertext and back are \ac{encryption} and \ac{decryption}. A well-designed
-cipher ensures that only those parties with the correct secret information, the \ac{key}, can perform encryption or
-decryption on the text.
+algorithm that performs the translation, plaintext is the original data, and ciphertext is the encoded data.
+\Ac{encryption} and \ac{decryption} are the processes for translating from plaintext to ciphertext and back. A
+well-designed cipher ensures that only those parties with the correct \ac{key}, or secret information, can perform
+encryption or decryption on the text.
 
@@
 
-There are two major cryptographic protocol families, symmetric and asymmetric. In \ac{sym-crypto}, which was invented
-long ago, encryption and decryption are performed with the same key, and both parties must have this key in order to
-communicate securely. In \ac{asym-crypto}, which was invented recently and is also called \ac{public-key-crypto},
-encryption and decryption are performed with two paired keys, called the public key and the private key. The innovation
-of \ac{public-key-crypto} is that the public key is not secret, and using this technique, two parties can communicate
-securely without requiring a previously agreed upon shared secret.
+There are two major cryptographic protocol families, symmetric and asymmetric. Symmetric protocols have been around for
+thousands of years, but asymmetric protocols were only invented in recent decades. In \ac{sym-crypto}, encryption and
+decryption are performed with the same key, and both parties must have this key in order to communicate securely. In
+\ac{asym-crypto}, also known as \ac{public-key-crypto}, encryption and decryption are performed with two paired keys,
+called the public key and the private key. The public key is not secret, but due to its relationship with the private
+key, it can be used to establish identity and encrypted communication. Using this technique, two parties can communicate
+privately without requiring previously agreed-upon secret information.
 
-One can also define two major cryptographic applications---securing \acf{DIM} and \acf{DAR}. Each application is a
-different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
-perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
-encryption of the data in motion. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy} and
-\ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any
-other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an attacker without
-detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for
-\ac{DIM}.
+There are also two major cryptographic applications, securing \ac{DIM} and \ac{DAR}. Each application presents different
+challenges that require different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to perform
+authentication and establish short-lived \ac{sym-crypto} session keys. The session keys perform the actual encryption of
+the \acl{DIM}. \Ac{forward-secrecy} and \ac{replay-protection} are two important properties of \ac{DIM} encryption
+protocols. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any other private
+key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an attacker without detection
+\cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of encryption for \ac{DIM}.
 
@@
 encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
-somewhere in computer memory, which often takes place with the assistance of dedicated hardware. \Ac{disk-encryption}
-for laptops and mobile devices is an example of encryption for \ac{DAR}.
+somewhere in computer memory. Secure storage often takes place with the assistance of dedicated hardware.
+\Ac{disk-encryption} for laptops and mobile devices is an example of encryption for \ac{DAR}.
 
-Data secrecy is not cryptography's only use, however. Precisely speaking, data secrecy is the purpose of encryption.
-Encryption is cryptography's ``killer app,'' but it is only a part of cryptography's usefulness to security.
+Data secrecy is not cryptography's only use, however. Data secrecy is the purpose of encryption, but encryption is only
+one application of cryptography. Although encryption is cryptography's ``killer app,'' it is just a part of
+cryptography's usefulness to security.
 
@@
 \cite{shostack_threat_2014}. These system-level properties emerge from both the composition of the system's components
-(architecture) and the security of the components themselves; likewise, the security of each component emerges from the
+(architecture) and the security of the components themselves. Likewise, the security of each component emerges from the
 architecture of its subcomponents and the security of the subcomponents themselves. At the end of this chain of
@@
 \mytab{table-security-and-crypto} connects each security property to its cryptographic basis. While lacking direct
-cryptographic foundations, availability indirectly relies on each other property, and authorization schemes are
-typically rooted in authentication.
+cryptographic foundations, the property of availability indirectly relies on the other properties, and authorization
+schemes are typically rooted in authentication.
 
@@
 catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
-information disclosure done both for their own sake and as tools in general attacks. The importance of taking these
-tools away is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
+information disclosure done both for their own sake and as tools in general attacks. The importance of neutralizing
+these tools is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
 disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
@@
 
-Encryption's dual role as enabler of privacy and cornerstone of security is at the heart of the EA debate. Cryptography
-has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary in the realm
-of security.
+Encryption's dual role as an enabler of privacy and cornerstone of security is at the heart of the EA debate.
+Cryptography has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary
+in the realm of security.
 
@@
 Computerized encryption is a new technology, but the field of cryptography is old, since demand for privacy is as old as
-communication itself. The most well known ancient example of rudimentary encryption is the Caesar cipher, named after
+communication itself. The most well-known ancient example of rudimentary encryption is the Caesar cipher, named after
 the character substitution technique Julius Caesar used to protect private correspondence
@@
 in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
-computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon's formalized the modern ``mathematical
-analysis of cryptography'' in 1949 \cite{shannon_communication_1949}, and Whitfield Diffie and Martin Hellman published
+computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon formalized the modern ``mathematical
+analysis of cryptography'' in 1949 \cite{shannon_communication_1949} and Whitfield Diffie and Martin Hellman published
 research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
-important advance. Combined with computer networking and personal computing advances in the 1980s, it for the first time
-put the power of strong encryption not just in the hands of governments and militaries, but of ordinary people.
+important advance. Combined with computer networking and personal computing advances in the 1980s, it put the power of
+strong encryption not just in the hands of governments and militaries, but of ordinary people.
 
-However, this power enables very strong privacy, and privacy is a property governments aim to regulate. The use of
-encryption for for private purposes has a contentious history. In 1587, Mary Queen of Scots was convicted of treason
-based on evidence from decrypted letters, and in 1807, prosecutors trying Aaron Burr for treason tried to force
-testimony from his secretary on the contents of encrypted messages \cite{kerr_encryption_2017}. Government's natural
-stance towards privacy makes it inherently uncomfortable with the rapid increase in availability of strong encryption.
-
-% Electronic encryption must be understood as a technological development embedded in the longer history of privacy and
-% human rights.
+However, this power enables very strong individual privacy. Government is inherently uncomfortable with individual
+privacy; therefore, the use of encryption for private purposes has a contentious history. In 1587, Mary Queen of Scots
+was convicted of treason based on evidence from decrypted letters, and in 1807, the prosecutors who tried Aaron Burr for
+treason attempted to force testimony from his secretary on the contents of encrypted messages
+\cite{kerr_encryption_2017}.
 
@@
 
-All these realities---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
+These circumstances---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
 and a blossoming computer industry foundationally reliant on cryptography---came to a head in what has become known as
@@
 
-In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
-the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
-shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
-text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
-secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
-use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
-the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
-and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
-foundational role of cryptography in security---and the potential of human rights abuses for compromised
-privacy---industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
-\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
-\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.
+In 1976, Diffie and Hellmann published their seminal research in public key cryptography. That same year, the U.S.
+Congress passed the \ac{AECA} and declared that strong cryptography is subject to export controls
+\cite{kehl_right_2015}. \Ac{the-cw1} began in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating
+access to plain text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, e-mail
+encryption software, in order that strong cryptography be ``made available to the American public before it became
+illegal to use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip}
+\cite{press_1993} with the goal of ``providing the public with strong cryptographic tools without sacrificing the
+ability of law enforcement and intelligence agencies to access unencrypted versions of those communications''
+\cite{thompson_2015}. Citing the foundational role of cryptography in security---and the potential of human rights
+abuses for compromised privacy---industrial and technical leaders reacted negatively to the initiative
+\cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users
+to subvert the \ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.
 
@@
 primarily focused on various key escrow proposals \cite{thompson_2015}. In 1996, two pro-encryption bills were
-introduced in the US Congress. In the Senate, it was S.1726, the \ac{PROCODEAct} of 1996, to abolish the export controls
-on encryption software \cite{burns_s1726_1996}. In the House, it was HR.3011, the \ac{SAFEAct} of 1996, to explicitly
-allow arbitrarily strong encryption for all legal activity \cite{goodlatte_hr3011_1996}. Indeed, the \ac{SAFEAct} went
-further, including ``provisions that would have barred the government from creating a mandatory key escrow system and
-would also have removed export restrictions on most generally available software and hardware containing encryption''
-\cite{kehl_right_2015}. Zimmermann, the author of PGP, who had by this time endured an investigation by the Customs
-Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct} \cite{zimmermann_1996}. In
-November of that year, the Clinton administration released an executive order effectively removing export controls on
-strong encryption products along the lines of the \ac{PROCODEAct}, which never came to pass \cite{clinton_1996}.
+introduced in the U.S. Congress, one in the Senate and one in the House. The Senate saw S.1726, the \ac{PROCODEAct} of
+1996, which sought to abolish the export controls on encryption software \cite{burns_s1726_1996}. The House saw HR.3011,
+the \ac{SAFEAct} of 1996, which similarly sought to remove export controls, but further sought to explicitly allow
+arbitrarily strong encryption for all legal activity and preclude mandatory \ac{EA} schemes such as the \ac{clipperchip}
+\cite{goodlatte_hr3011_1996}. Zimmermann, the author of PGP, who had by this time endured an investigation by the
+Customs Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct}
+\cite{zimmermann_1996}. In November of 1996, the Clinton administration released an executive order effectively removing
+export controls on encryption products along the lines of the \ac{PROCODEAct}, which never came to pass
+\cite{clinton_1996}.
 
-From 1996 to 1999, the \ac{SAFEAct} was proposed several times, had hearings, and gained support \cite{kehl_right_2015}.
-By this time, there was ``an overwhelming amount of evidence against moving ahead with any key escrow schemes''
-\cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost all \ac{SAFEAct}
-proposals \cite{kehl_right_2015}. This development marks the end of \ac{the-cw1}.
+From 1996 to 1999, the \ac{SAFEAct} was proposed several times, was discussed at hearings, and gained support
+\cite{kehl_right_2015}. By this time, there was ``an overwhelming amount of evidence against moving ahead with any key
+escrow schemes'' \cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost
+all \ac{SAFEAct} proposals \cite{kehl_right_2015}. This development marked the end of \ac{the-cw1}.
 
@@
 
-Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
-administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
-\ac{PROCODEAct} and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}.
-This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
-pencil of executive order.
+Two developments paved the way for \ac{the-cw2}. The first was the result of inaction: when the Clinton administration
+changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The \ac{PROCODEAct}
+and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}. This legislative
+failure meant that rather than being written in the pen of law, encryption policy was written in the pencil of executive
+order.
 
-The second development was the events of and reaction to the morning of September 11, 2001. These events occurred in the
-context of decades of all three branches of government expanding law enforcement powers. Presidents promoted the idea of
+The second development was the terror attacks of September 11, 2001. The events of that day occurred in the context of
+expanding law enforcement powers fueled for decades by all three branches of government. Presidents promoted the idea of
 being ``tough on crime'' and launched a war on drugs; legislators passed bills such as 1978's \ac{FISA}, which provided
@@
 
-In this context of expanded surveillance, the scale and audacity of the September 11 attacks seemed to justify yet
-greater investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had
-several significant impacts: it weakened the limitation for using \ac{FISA} requests on U.S. citizens; it expanded the
-scope of what could be compelled via \ac{FISA} orders; it authorized ``roving'' wiretaps; and it increased the power of
-\acp{NSL}, which can be used without judicial review to compel information from digital service providers while
+In this context of expanded surveillance, the scale of the September 11 attacks seemed to justify yet greater
+investigatory powers \cite{bloss_escalating_2007}. October of 2001 saw passage of the USA PATRIOT Act, which had several
+significant impacts: (1) it weakened the limitation for using \ac{FISA} requests on U.S. citizens; (2) it expanded the
+scope of what could be compelled via \ac{FISA} orders; (3) it authorized ``roving'' wiretaps; and (4) it increased the
+power of \acp{NSL}, which can be used without judicial review to compel information from digital service providers while
 precluding any public disclosure of the event \cite{sensenbrenner_2001} \cite{shamsi_2011}. The government was
@@
 The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
-war, and increased government surveillance set events in motion toward conflict, but it was the 2013 Snowden revelations
-that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
-caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
-public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
-technical solution. Benefiting from advancements in underlying technologies and under pressure from both the human
-rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing default
-device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.
+war. Increased government surveillance set events in motion toward conflict. These two developments prepared the way,
+but it was the 2013 Snowden revelations that constituted the crossing of the threshold. Though policies such as the U.S.
+\ac{NSA}'s warrantless wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and
+related programs that caused the real public outcry \cite{landau_making_2013}. With the public interest focused on
+digital privacy, encryption promised to be a technical solution. U.S. tech companies responded by introducing default
+device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM}.
 
@@
 
-The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
-their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
-basic digital security. Recall encryption and cryptography's critical role in security as described in
-\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
-security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
-have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.
+The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights.
+Their behavior is a matter of several practical factors, including market pressure, cultural pressure, and the
+advancement in available encryption technology \cite{treguer_us_2018}. Another factor is the pursuit of basic digital
+security. Recall encryption and cryptography's critical role in security as described in \mysec{sec-crypto-basics}. When
+lawmakers question technical leaders about their companies' encryption policies, executives have repeatedly appealed to
+the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.
 
@@
 \paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
-antagonistic, an assumption implicit in the ``nothing-to-hide'' argument, but discussions in recent years have shown
-that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
-\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the pursuit of
-law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by now
-acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
-security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.
+antagonistic, an assumption implicit in the ``nothing-to-hide'' argument. However, discussions in recent years have
+shown that the debate is more about ``Security vs. Security'' than ``Security vs. Privacy''
+\cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security. First, public security (often
+referred to as ``safety'' in this paper), which is the pursuit of law enforcement. Second, cybersecurity, which the
+pursuit of the technical community. Both sides of the debate have by now acknowledged that privacy and public security
+are not always in conflict. In fact \ac{EA}'s effects on public security vs. cybersecurity---which in return affects
+public security---may be more important \cite{schneier_2019}.
 
-\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
-locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
-to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
-objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
-\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
-February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
-phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
-\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
-wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
-jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
-make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.
+\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack resulted in the first major battle of \ac{the-cw2}. The
+suspect was killed, leaving behind a locked iPhone. The \ac{FBI} issued an order under the \acrlong{AWA} to compel Apple
+to unlock the device, which was among the first generation of Apple's fully encrypted iPhones. Apple objected on grounds
+that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices \cite{schulze_clipper_2017}.
+The case occupied U.S. District courts, the media, and the attention of the political world from February 16, 2016, when
+the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the phone through alternate
+means \cite{novet_2016}. Although the phone proved not to contain any important data \cite{schulze_clipper_2017}, the
+high profile case featured two characteristic elements of the new crypto wars---terrorism and device encryption. Shortly
+after the case ended, a pair of Republican and Democratic senators jointly released a discussion draft of a bill that
+would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the floor, and the unresolved nature of
+the court case left the legal debate unresolved.
 
-\paragraph*{The \Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}'' to
-describe four reasons that governments and law enforcement agencies use to undercut public support for strong encryption
-\cite{may_1994}. The traditional list includes terrorists, pedophiles, drug dealers, and money launderers, though
-sometimes kidnappers are substituted for money launderers \cite{schneier_scaring_2019}. Reflecting the times,
-\ac{the-cw1} emphasized drug trafficking whereas \ac{the-cw2} is emphasizing terrorism and child pornography (\ac{CSAM})
-\cite{schulze_clipper_2017}. The focus on terrorism is likely a result of culture that emerged from the September 11
-terror attacks as well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case; the focus on
-\ac{CSAM} is likely a result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New
-York Times} \cite{keller_internet_2019}.
+\paragraph*{The \Ac{horsemen}} In the 1990s, an influential encryption advocate coined the phrase ``\ac{horsemen-full}''
+to describe four reasons that governments and law enforcement agencies use to undercut public support for strong
+encryption \cite{may_1994}. The original list includes terrorists, pedophiles, drug dealers, and money launderers,
+though many substitute kidnappers for money launderers \cite{schneier_scaring_2019}. Reflecting the times, \ac{the-cw1}
+emphasized drug trafficking whereas \ac{the-cw2} emphasizes terrorism and child pornography (increasingly referred to as
+\ac{CSAM}) \cite{schulze_clipper_2017}. The focus on terrorism is likely a result of the September 11 terror attacks as
+well as the San Bernardino terror attack that launched the Apple vs. \ac{FBI} case. The focus on \ac{CSAM} is likely a
+result of the growth of the ``dark web'' and a revelatory 2019 investigation by the \textit{New York Times}
+\cite{keller_internet_2019}.
 
 \paragraph*{\ac{DIM} vs. \ac{DAR}} As noted in \mysec{sec-crypto-basics}, securing \acl{DIM} and \acl{DAR} are two
-different problems that result in different solutions. Therefore, both technical and policy proposals often split their
-recommendations along these lines \cite{group_2019} \cite{owen_law_2018}. The focus from law enforcement in the U.S. has
-pendulated between each of these applications a few times \cite{schneier_2019}. The typical \ac{DIM} application is
-\ac{E2EE} in instant messaging applications, whereas the typical \ac{DAR} application is \ac{disk-encryption} of mobile
-phones. A research group at Carnegie Mellon with participants from both sides of the debate has identified \ac{EA} for
-\ac{DAR} as the more tractable problem \cite{group_2019}.
+different challenges that require different solutions. Therefore, both technical and policy proposals often split their
+recommendations along these lines \cite{group_2019} \cite{owen_law_2018}. Discussion of \ac{EA} for \ac{DIM} typically
+involves \ac{E2EE}, whereas for \ac{DAR} it typically involves \ac{disk-encryption} of mobile phones. The focus of law
+enforcement in the U.S. has pendulated between these several times \cite{schneier_2019}. A research group at Carnegie
+Mellon with participants from both sides of the debate has identified \ac{EA} for \ac{DAR} as the more tractable problem
+\cite{group_2019}.
 
 \paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
-uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since legislature is
-part of the broader data security threat model, its behavior is important. Unfortunately, the present political dynamics
-and the lack of reliable data on what law enforcement agencies need make it difficult for the technical community to
-engage productively with policymakers \cite{granick_2018}.
+uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since government is
+part of the broader data security threat model, its behavior is important. Unfortunately, it is difficult for the
+technical community to engage productively with policymakers due to present political dynamics and the lack of reliable
+data showing what law enforcement agencies need \cite{granick_2018}.
 
 That brings us to where the debate stands today. A December 2019 U.S. Senate Judiciary Committee featured the top two
-U.S. parties both expressing anger over whether data should be ``beyond the reach of the law'' and expressed threats
-such as ``get your act together, or we will gladly get your act together for you'' \cite{geller_2019}. In March 2020,
-the Republican atop the Senate Judiciary Committee introduced S.3398, the \ac{EARNITAct} of 2020, which seeks to
-establish a 16-chair committee with the power to revoke online platforms' liability protection for user-submitted
-content if the committee determines the platform is not doing enough to fight \ac{CSAM}, a legal synonym for child
-pornography \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
-\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous; however, critics
+U.S. parties both expressing anger over whether data should be ``beyond the reach of the law.'' Senators threatened tech
+executives with statements such as ``get your act together, or we will gladly get your act together for you''
+\cite{geller_2019}. In March 2020, the Republican atop the Senate Judiciary Committee introduced S.3398, the
+\ac{EARNITAct} of 2020, which seeks to establish a 16-chair committee with the power to revoke online platforms'
+liability protection for user-submitted content if the committee determines the platform is not doing enough to fight
+\ac{CSAM} \cite{graham_s3398_2020}. As illustrated by the previously mentioned \textit{New York Times} feature
+\cite{keller_internet_2019} that likely triggered the December 2019 hearing, the problem is enormous. However, critics
 largely saw the \ac{EARNITAct} as an attempt to indirectly yet effectively outlaw \ac{E2EE} \cite{newman_2020}
@@
 
-There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
-basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
-plaintext.
+There are several technical approaches to \ac{EA}. This section organizes approaches into three
+categories---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring plaintext.
 
@@
 
-The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
-information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
-classes of key escrow.
+Key escrow is the most obvious way to implement \ac{EA}. It involves storing additional copies of the encryption keys.
+Alternatively, rather than directly storing copies, key escrow may involve storing information that can be used to
+derive additional copies. Key escrow can be subdivided into various types according to where the escrowed information is
+stored.
 
@@
 information. In this arrangement, trust is put in the distributed system, not in any single party. Such systems offer
-high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
-policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has modern blockchain
-\cite{phan_key_2017} and distributed device \cite{servan_schreiber_jje_2020} variations. These approaches are not as
-thoroughly explored as trusted-party \ac{key-escrow}.
+high data availability while making covert key recovery difficult. This approach was introduced in \ac{the-cw1}
+\cite{goos_oblivious_1996} and has modern blockchain \cite{phan_key_2017} and distributed device
+\cite{servan_schreiber_jje_2020} variations. These approaches are not as thoroughly explored as trusted-party
+\ac{key-escrow}.
 
@@
 
-A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
-trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
+The third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not
+by trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
 ``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
-\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
+\ac{DAR} and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
 \cite{ozzie_2018}.
@@
 some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
-scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
-data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
-decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
-escrow do exist \cite{bellare_translucent_1996}.
+scheme, no keys are escrowed, but law enforcement could recover a predetermined percentage of plaintext data. That
+percentage would be designed to balance privacy, security, and safety. This research, though more experimental than key
+escrow proposals, was explicitly pursued to demonstrate that alternatives to key escrow do exist
+\cite{bellare_translucent_1996}.
 
@@
 human users'' \cite{wright_crypto_2018}. Similar to the translucent approach, no keys are escrowed and only passive
-surveillance is possible. Strong cryptography is a precondition for this scheme, in which keys are generated in a manner
-that makes them inherently recoverable only after a large up front and marginal cost.
+surveillance is possible. Keys are generated in a manner that makes messages inherently recoverable, but only after
+extreme up front costs and significant marginal costs.
 
@@
 
-As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Under
-\ac{lawful-hacking}, instead of having access to plaintext data via an alternative means of decryption, law enforcement
-is allowed to perform otherwise-illegal hacking activity in order to find the key, compromise a device, or intercept
-plaintext versions of the data \cite{kerr_encryption_2017}. While some see this as a viable middle ground
-\cite{nguyen_lawful_2017}, encouraging law enforcement to exploit vulnerabilities creates mis-aligned incentives. If law
-enforcement opposes strong security in general, it could foster the exploit market while costing ordinary users money
-and security \cite{soesanto_2018}.
+\Ac{lawful-hacking} is one way to access \ac{plaintext} without \ac{EA}. Under \ac{lawful-hacking}, instead of having
+access to plaintext data via an alternative means of decryption, law enforcement is allowed to perform otherwise-illegal
+hacking activity in order to find the key, compromise a device, or intercept plaintext versions of the data
+\cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
+enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
+general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.
 
@@
 
-Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
-authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
-enforcement is traditionally good at---investigation and sting operations---and is used successfully. In 2017, an
+Simpler methods than computer hacking or legal coercion can be used: encrypted data may exist in plaintext elsewhere, or
+authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on law
+enforcement's skills in investigation and sting operations. This approach is used successfully today. In 2017, an
 international law enforcement operation used this technique to take down a large illicit online drug market
-\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
+\cite{greenberg_2018}. This approach also does not address \ac{DIM} and requires well-executed operations.
 
@@
 
-The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
+The following sections list laws which directly or indirectly affect the use of encryption. Though this thesis is
 U.S.--focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an
@@
 
-Expanded arms export controls and created the \ac{ITAR} regulations framework. Encryption products were classified as
-arms and subjected to export controls \cite{kehl_right_2015}.
+The \ac{AECA} expanded arms export controls and created the \ac{ITAR} regulations framework. Encryption products were
+classified as arms and subjected to export controls \cite{kehl_right_2015}.
 
@@
 
-Allows the executive branch ``to authorize electronic surveillance for foreign intelligence purposes without a court
-order, in some circumstances.'' \acs{FISA} applications take place in secret, and may be conducted against U.S. citizens
-if foreign intelligence is the ``primary purpose'' \cite{shamsi_2011}. \acs{FISA} has been the subject of frequent FBI
-abuse \cite{shamsi_2011} \cite{tucker_2020}.
+This statute allows the executive branch ``to authorize electronic surveillance for foreign intelligence purposes
+without a court order, in some circumstances.'' \acs{FISA} applications take place in secret, and may be conducted
+against U.S. citizens if foreign intelligence is the ``primary purpose'' \cite{shamsi_2011}. \acs{FISA} has been the
+subject of frequent FBI abuse \cite{shamsi_2011} \cite{tucker_2020}.
 
@@
 
-Included the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act. Provided explicit
-protections for private electronic communications data, and updated definitions to contemporary technology. It granted
-law enforcement explicit access to certain data as well; some terminology has aged poorly with technological evolution,
-making the access increasingly permissive \cite{shamsi_2011}.
+The \ac{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
+Provides explicit protections for private electronic communications data, and updated definitions to contemporary
+technology. It grants law enforcement explicit access to certain data as well; some terminology has aged poorly with
+technological evolution, making the access increasingly permissive \cite{shamsi_2011}.
 
 }
+% FIXME: make tense for enacted laws present
 
@@
 
-Mandated that telecommunications carriers enable the government to ``intercept all of the subscriber's wire and
-electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized to
-be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant that
-the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
+\ac{CALEA} mandates that telecommunications carriers enable the government to ``intercept all of the subscriber's wire
+and electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized
+to be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant
+that the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
 specifically excludes ``information services'' from its interception requirement which has included internet platforms,
@@
 
-\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
-led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely---encryption is an
-important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
-child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``appropriate means must be available to
-fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
-``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
-a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
+\Ac{CESA} was a legislative proposal by the Clinton administration introduced alongside the executive order that led to
+the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely: encryption is an important
+tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers, child
+pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``appropriate means must be available to fulfill
+these law enforcement objectives.'' The bill would establish a voluntary-participation third-party key escrow system
+with ``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled
+through a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
 member of Congress to officially sponsor the bill.
@@
 
-After \ac{CESA} failed, the White House proposed watered-down legislation that this time was successfully sponsored by a
-senator. This bill would have amended the \ac{ECPA} and \ac{CFAA}. It would have made relatively modest changes to
-government data access requirements. Though it received congressional sponsorship, even the sponsor was critical of
-parts of the bill, and it died in committee \cite{senate_2000}.
+After \ac{CESA} failed, the White House proposed watered-down legislation that was successfully sponsored by a senator.
+This bill would have amended the \ac{ECPA} and \ac{CFAA}. It would have made relatively modest changes to government
+data access requirements. Though it received congressional sponsorship, even the sponsor was critical of parts of the
+bill, and it died in committee \cite{senate_2000}.
 
@@
 in \ac{the-cw2}. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
-Apple to unlock the phone. Though it would not mandate a technical mechanism, the bill would have mandated that
-manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
-proposals in \ac{the-cw1} by forcing platform and application developers to comply. Previewing a common phrase in recent
-\ac{EA} arguments \cite{geller_2019}, the draft's authors both used the phrase ``above the law'' in promotion of the
-bill.
+Apple to unlock the suspect's phone. Though it would not mandate a specific technical approach, the bill would have
+mandated that manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several
+regulatory proposals in \ac{the-cw1} by forcing platform and application developers to comply. Previewing a common
+phrase in recent \ac{EA} arguments \cite{geller_2019}, the draft's authors both used the phrase ``above the law'' in
+promotion of the bill.
 
@@
 The above history and section on U.S. law encompasses the executive and legislative branches' influence on surveillance
-and cryptography. A complete picture requires understanding the judiciary branch's influence via court rulings. The
-judiciary branch influences policy through interpretation of laws and creation of legal doctrine. In this section I
-describe this influence, review how it has affected cryptography policy, and analyze where cryptographic jurisprudence
-may go from here.
+and cryptography. However, a complete portrait of government encryption policy must include the influence of the
+judiciary branch. The judiciary branch influences policy through interpretation of laws and creation of legal doctrine.
+This section describes this influence, reviews how it has affected cryptography policy, and analyzes where cryptographic
+jurisprudence may go from here.
+
+% TODO: find a "con" word that means "denied"! Contested?
 
 The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
-authoritative legal texts. However, whether it is condemned, denied, seen as a stark reality, or seen as a valid
-function, the judicial branch also establishes legal doctrine and actively enforces rulings in a manner that amounts to
-creation of public policy \cite{feeley_judicial_2000}. Due to the clear congressional authority over telecommunications
-and federal investigatory powers, the courts are most likely to exert influence in this area through interpretation of
-current law. This interpretation will both determine how present law is enforced and draw a constitutional line around
-any actions the executive or legislative branches undertake.
+authoritative legal texts. However, whether it is condemned, denied, conceded, or condoned, the judicial branch also
+establishes legal doctrine and actively enforces rulings in a manner that amounts to creation of public policy
+\cite{feeley_judicial_2000}. Due to Congress's clear authority over telecommunications and federal investigatory powers,
+the courts are most likely to exert influence in this area through interpretation of current law. This interpretation
+will both determine how present law is enforced and draw constitutional boundaries around any actions the executive or
+legislative branches undertake.
 
 Judicial influence on encryption policy is primarily mediated through the Fourth Amendment, particularly regarding the
-interpretation of a person's ``effects,'' what is ``reasonable,'' and what is a ``search.'' The Supreme Court's
-interpretation and implementation of the Amendment's crime control-- and civil liberty--balancing principles has changed
-over time. An in-depth analysis of the Court's jurisprudence from the 1960's Warren Court, through the influential 1980s
-Rehnquist Court, to the mid-2010s Roberts Court found a deep proclivity towards crime control that is slowly thawing
-\cite{gizzi_fourth_2016}. This thawing, caused by evolution of technology, of the court, and of the justices themselves,
-has been favorable for data privacy. Recent cases have increasingly sided with defendants, have caused justices to
-strain to show that they understand technology, and have increased the categories of protected digital data
-\cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter v. United States} may be the most impactful yet;
-though it only resulted in a narrow decision against unwarranted cell phone location data tracking, it does undermine
-the pre-digital third-party doctrine, which holds that Fourth Amendment protection is forfeit when the data is willingly
-given to a third-party \cite{franklin_2018}.
+interpretation of a person's ``effects,'' what is ``reasonable,'' and what defines a ``search.'' The Supreme Court's
+interpretation and implementation of the Amendment has changed over time. Legal scholars Craig Curtis and Michael Gizzi
+conducted an in-depth analysis of the Court's jurisprudence from the 1960s Warren Court, through the influential 1980s
+Rehnquist Court, to the mid-2010s Roberts Court \cite{gizzi_fourth_2016}. Their analysis revealed a deep proclivity
+towards crime control that is only recently thawing. This gradual return to decisions favoring civil liberties, caused
+by evolution of technology, of the court, and of the justices themselves, has been favorable for data privacy. Recent
+cases have increasingly sided with defendants, caused justices to strain to show that they understand technology, and
+increased the categories of protected digital data \cite{gizzi_fourth_2016}. 2018's potentially landmark \ii{Carpenter
+v. United States} may be the most impactful yet. Though the case only resulted in a narrow decision against unwarranted
+cell phone location data tracking, the Court's decision undermines the pre-digital third-party doctrine, which holds
+that Fourth Amendment protection is forfeit when the data is willingly given to a third-party \cite{franklin_2018}.
 
 As this relates to \ac{EA}, Fourth Amendment interpretation determines the threshold at which authorities require a
-warrant, not what demands they can or cannot make once they have one (e.g. demands for access to \ac{plaintext}). As
-mentioned above, the FBI argued in the San Bernardino case that the \ac{AWA} grants it authority to demand access,
-though it is strongly argued that \acs{CALEA} explicitly precludes this kind of demand \cite{gidari_2016}. Since the FBI
-dropped the case, the matter is not formally settled, though the consensus is that legislation such as the \ac{LAEDAct}
-is required before courts would uphold that command.
+warrant, not what demands they can or cannot make once they have one. As mentioned above, the FBI argued in the San
+Bernardino case that the \ac{AWA} grants it authority to demand access, though others argue that \acs{CALEA} explicitly
+precludes this kind of demand \cite{gidari_2016}. Since the FBI dropped the case, the matter is not formally settled,
+though the consensus is that legislation such as the \ac{LAEDAct} is required before courts would uphold that command.
 
-The second most salient legal issue is the Fifth Amendment right against self-incrimination, which comes up in cases of
+The second most salient legal issue is the Fifth Amendment right against self-incrimination, which arises in cases of
 compelled password disclosure. This question has not yet reached the Supreme Court, and state courts have ruled
-compelled disclosure both constitutional \cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional \cite{lee_its_2020}
-\cite{vaas_2019}.
+compelled disclosure as constitutional in some cases \cite{sobel_2019} \cite{lee_nj_2020} and unconstitutional in others
+\cite{lee_its_2020} \cite{vaas_2019}.
 
@@
 definition for wicked problems, the rules become out of sync with the principles that bade them. Recent flux in judicial
-rulings surrounding surveillance and technology, particularly around the Fourth Amendment, are evidence that the courts
-are in the process of updating interpretations of rules in light of the new facts; which principles they ultimately
-choose to favor remains to be seen, but rulings like \ii{Carpenter} show a willingness to protect privacy.
+rulings surrounding surveillance and technology, particularly around the Fourth Amendment, prove that the courts are in
+the process of updating interpretations of rules in light of the new facts; which principles they will ultimately favor
+remains to be seen, but rulings like \ii{Carpenter} show a willingness to protect privacy.
 
-Looking ahead, there are three types of cases through which the judicial branch could have a strong effect on digital
+There are three types of cases through which the judicial branch could have a strong effect on the future of digital
 privacy policy. First, Fourth Amendment cases will establish classes of protected data and thresholds for warrants.
@@
 
-This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
+This thesis focuses on U.S. policy, but \ac{EA} technologies and policies have international impacts. The following is a
 brief account of the regulatory environment in various geographical jurisdictions around the world based primarily on a
-2016 analysis by the Law Library of Congress \cite{acosta_government_2016}. Globally, a 2017 study found that more than
-half the world's internet users lived in a country where some form of plaintext recovery is mandated \cite{lewis_2017}.
+2016 analysis by the Law Library of Congress \cite{acosta_government_2016}. A separate 2017 study found that globally,
+more than half the world's internet users lived in a country where some form of plaintext recovery is mandated
+\cite{lewis_2017}.
 
@@
 
-In 1946 The U.S. formalized an intelligence sharing operation with the U.K. in the BRUSA (now UKUSA) Agreement; over the
+In 1946 The U.S. formalized an intelligence-sharing operation with the U.K. in the BRUSA (now UKUSA) Agreement. Over the
 following 10 years, Australia, Canada, and New Zealand were made full partners to this agreement, which has become known
@@
 and law enforcement but does not directly address encryption. Canada has traditionally supported strong encryption, but
-the pressure of \ac{the-cw2} is causing strain \cite{parsons_2019}.
+the pressure of \ac{the-cw2} is straining that support \cite{parsons_2019}.
 
@@
 telecommunications providers provide interfaces ``with functions that can cooperate with interception.'' Chinese
-authorities have great access to plaintext data by virtue of the political structure of the state, but also through
+authorities have great access to plaintext data by virtue of the political structure of the state, as well as through
 explicit requirements in recent anti-terrorism and cybersecurity laws \cite{lewis_2017}.

### 2020-12-10 Thesis-Chapter3-PolicyMaking.tex
Editorial improvements

@@
 
-Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
-greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
-to tackling \acp{wicked-prob} succeed and fail? How can we collectively situate ourselves to make real progress? This
-chapter seeks to answer these questions.
+Before detailing the specific arguments used in the \ac{EA} debate, it is worth achieving a better understanding of
+\acp{wicked-prob} in general. What are they, and why do the problems surrounding encryption, privacy, and \ac{EA} fall
+into this category? What approaches to tackling \acp{wicked-prob} succeed and fail? How can we strategically confront
+them and make real progress? This chapter seeks to answer these questions.
 
@@
 \Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
-In this section we analyze the nature of wicked problems and strategies for tackling them.
+This section analyzes the nature of wicked problems and approaches to tackling them.
 
@@
 
-Rittel's categorization of \acp{wicked-prob} grew out of frustration with the problems' resistance to classical problem
+Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to traditional problem
 solving methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the
 sweeping application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
-structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
-twentieth century, believers in the power of reason thought this progress would march on in the realm of public
+structures, and bestowed material wealth previously unimaginable. With these material problems largely solved in the
+twentieth century, believers in the power of reason thought this progress would continue in the realm of public
 planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
@@
 When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
-alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation with countless
-cultures and subcultures which are effectively represented by two mutually hostile political parties, agreeing on goals
-is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that defy
-the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
+alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation composed of
+many varying cultures and subcultures and politically dominated by two mutually hostile parties, agreeing on
+goals is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that
+defy the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
 \cite{rittel_dilemmas_1973}:
@@
 
-Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
-infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
-analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
-conditions. The ``wicked'' problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means
-that the scientific approach will be effective.
+In summary, (1) the \acp{wicked-prob} are impossible to definitively identify, (2) potential solutions are impossible to
+definitively evaluate, (3) every action or inaction has permanent effects, and (4) solutions are importable from one
+problem to another. The policy version of the scientific method cannot be used under these conditions. The ``wicked''
+problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means that the scientific approach
+will be effective.
 
 Several of \acp{wicked-prob}' characteristics are results of the fact that we have no accurate predictive model of the
-world and human behavior at large. This does not need to be expounded upon.
-
-The more important insight is one step removed: precisely because there is an inexhaustible set of potential solutions,
-the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
-proposed solution demands research and context to be fed back into the problem definition. This feedback from proposal
-to definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals
-to a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
-themselves.
+world and human behavior. This is self-evident. The more important insight is one step removed: precisely because there
+is an inexhaustible set of potential solutions, the problem definition---that is, the set of information required to
+produce a solution---is not self-contained. Each proposed solution demands that new research and context be fed back
+into the problem definition. This feedback from proposal to definition violates the linearity of the traditional
+approach. One cannot reason from problem statement to proposals to a solution; instead, one is forced to constantly
+refine the problem statement based on the content of proposals themselves.
 
@@
 \cite{commission_tackling_2018}. Say the problem is initially formulated as long-term change to the environment caused
-by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms: profligate behavior
-in consumeristic societies must be reigned in at a local and personal level; global inter-governmental coordination is
-the only solution, as individual changes will have no impact; or the situation is overblown by idealists and
-power-mongers, and technological progress and adaptive markets will handle any negative effects that come to pass
-\cite{commission_tackling_2018}. Which response is correct? One cannot tell from the problem statement. How much do
-individual choices contribute to greenhouse gas emissions? How effective are international accords? How costly will the
-changes be, and how capable is technology to respond? Evaluating the validity of each proposal requires updates to the
-problem statement itself.
+by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms
+\cite{commission_tackling_2018}. First, profligate behavior in consumeristic societies must be reigned in at a local and
+personal level. Second, global inter-governmental coordination is the only solution, as individual changes will have no
+impact. Third, the situation is overblown by idealists and power-mongers, and technological progress and adaptive
+markets will handle any negative effects that come to pass. Which response is correct? It is impossible to tell by
+evaluating the problem statement. How much do individual choices contribute to greenhouse gas emissions? How effective
+are international accords? How costly will the changes be, and how capable is technology to respond? Evaluating the
+validity of each proposal requires updates to the problem statement itself.
 
@@
 
-\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
-change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
-solution in a way that doesn't leave vast swaths of people behind. \Acp{wicked-prob} quickly grow in complexity with the
-number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are a tangle of contributing factors, and
-despite the presence of tame elements, they pose a large number of difficult challenges.
+\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked themselves. For example, climate change has
+many subproblems. Developing clean energy technology is a tame subproblem; implementing any solution in a way that
+doesn't leave vast swaths of people behind is a wicked subproblem. \Acp{wicked-prob} quickly grow in complexity with the
+number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are composed of a tangle of contributing
+factors, and despite the presence of tame elements, they pose a large number of difficult challenges.
 
@@
         safety, and trust.
-  \item Balancing each value in the face of constantly evolving technology is never ending.
+  \item Balancing each value in the face of constantly evolving technology is a never-ending cycle.
   \item \ac{EA} or other proposals cannot definitively solve the problem.
@@
         criminal behavior.
-  \item The problem can be framed as insufficient investigatory access to digital data (presupposing \ac{EA} as the
-        solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions), and
-        more.
-  \item The decisions of regulators and technologists alike have real impacts in the world today.
+  \item The problem can be framed many ways: as insufficient investigatory access to digital data (presupposing \ac{EA}
+        as the solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions),
+        and more.
+  \item The decisions of regulators and technologists have real impacts in the world today.
 \end{enumerate}
 
-Cybersecurity law scholar Alan Rozenshtein argues for treating \ac{encryption} and \ac{EA} as a \ac{wicked-prob} at
-length \cite{rozenshtein_wicked_2018}. He breaks the root issues down into three categories. First, that there is
-disagreement on goals (or even basic premises). Sides can't agree whether things are ``going dark'' or we're in the
-``golden age for surveillance'' or how much to value competing notions of security. Second, that information is
-``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations, and though the
-consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be the case,
-especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving values and
-technology mean that the subject is always up for renegotiation.
+Cybersecurity law scholar Alan Rozenshtein argues at length for treating \ac{encryption} and \ac{EA} as a
+\ac{wicked-prob} \cite{rozenshtein_wicked_2018}. He divides the root issues into three categories. First, there is
+disagreement on what the goals should be (and even basic premises). Sides do not agree on whether things are ``going
+dark'' or we're in the ``golden age for surveillance'' or how much to value competing notions of security. Second,
+information is ``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations.
+Though the consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be
+the case, especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving
+values and technology mean that the subject is always up for renegotiation.
 
-\Ac{encryption} technology is particularly sensitive to the property that policy implementation has irreversible
-effects. Technology deployments have long tails, and attackers have the ability to record and store data for later
-analysis. In one investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile
-device that they cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the
-only ones to benefit from a policy miscalculation could negatively impact security for a very long time.
+\Ac{encryption} technology is particularly sensitive to the irreversible effects of policy implementation. Technology
+deployments have long tails, and attackers have the ability to record and store data for later analysis. In one
+investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile device that they
+cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the only ones to
+benefit from a policy miscalculation that could negatively impact security for a very long time.
 
@@
 Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
-\acp{wicked-prob}. In the next two sections, we look at strategies for dealing with it.
+\acp{wicked-prob}. The next two sections investigate strategies for dealing with it.
 
@@
 
-The \ac{classical-method} \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
+The \ac{classical-method} \cite{feeley_judicial_2000} (also known as ``the modern-classical model of planning''
 \cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_muddling_1959}, ``traditional policy
@@
 alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
-illustrates the approach in the context of encryption and \ac{EA}. It has a purely linear flow from problem to solution
+illustrates this approach in the context of encryption and \ac{EA}. It has a purely linear flow from problem to solution
 except for the ``refinement'' step, in which the method analyzes outcomes and corrects errors. However, it is important
@@
 The shortcomings of the \ac{classical-method} were discussed in \mysec{wicked-characteristics}. It fails due to
-disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
-flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.
+disagreement over goals, the dependence of the problem definition on the potential solutions generated (violating the
+linear flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.
 
@@
 
-Lindblom's ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis. The
-alternatives selected for comparison are defined relative to the status quo and are close enough to one another to be
-able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
+Lindblom's process of ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis.
+The alternatives selected for comparison are defined relative to the status quo and are close enough to one another to
+be able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
 predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
@@
 
-Despite its strengths, \ac{incrementalism} has its weaknesses at addressing \acp{wicked-prob}. Its main weakness is its
-lack of high level strategic analysis. It is incapable of drastic change, which is sometimes necessary. By Lindblom's
-admission, it lacks a safeguard for consideration of all relevant values and may ``overlook excellent policies for no
-other reason than that they are not suggested by the chain of successive policy steps leading up to the present''
-\cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to the next can result in messy policy
-that none are particularly fond of---``As Lindblom's sobriquet suggests, it often [leads] to a considerable muddle''
-\cite{feeley_judicial_2000}. Consider application of the \acl{CFAA} as one such muddle \cite{wolff_computer_2016}.
+Despite its strengths, \ac{incrementalism} also has weaknesses in its ability to address \acp{wicked-prob}. Its main
+weakness is the absence of high level strategic analysis. Incrementalism is incapable of drastic change, which is
+sometimes necessary. By Lindblom's admission, it lacks a safeguard for consideration of all relevant values and may
+``overlook excellent policies for no other reason than that they are not suggested by the chain of successive policy
+steps leading up to the present'' \cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to
+the next can result in messy policy that none are particularly fond of---``As Lindblom's sobriquet suggests, it often
+[leads] to a considerable muddle'' \cite{feeley_judicial_2000}. Application of the \acl{CFAA} is one such muddle
+\cite{wolff_computer_2016}.
 
@@
 The \ac{classical-method} and \ac{incrementalism} are not the only styles of policymaking. They represent perhaps two
-extremes on a spectrum of rational planning and intuitive muddling. Both have their strengths and weaknesses. One may
+extremes on a spectrum of rational planning and intuitive ``muddling.'' Both have strengths and weaknesses. One may
 intuit that a reasonable strategy is the selective use of both approaches according to the situation, and indeed this
-has also been formally suggested \cite{etzioni_scanning_1967}. However, even a combination of these methods does not
-precisely suit all classes of wicked problems. Is a problem tame? Use the \ac{classical-method}. Is it wicked, but
-strategically under control and relatively non-controversial? Use \ac{incrementalism}. Is it wicked, and lacking a
-strategic response or highly controversial? What then?
+has been formally suggested \cite{etzioni_scanning_1967}. However, even a combination of these methods does not suit all
+classes of wicked problems. Is a problem tame? Use the \ac{classical-method}. Is it wicked, but strategically under
+control and relatively non-controversial? Use \ac{incrementalism}. What about when it is it wicked, lacks a strategic
+response, and is highly controversial?
 
@@
 definition there is no step that will truly solve the problem, easy answers deliberately emphasize one value to the
-exclusion of others, leaving chaos in those neglected areas. As a result of ignoring root issues feeding the problem,
-such solutions are even more likely to produce unintended consequences \cite{commission_tackling_2018}.
+exclusion of others. This leads to chaos in those neglected areas. Solutions based on easy answers produce unintended
+consequences because they ignore the root issues \cite{commission_tackling_2018}.
 
@@
 
-Making sure every relevant group is included is important for two reasons. First, because information is ``uncertain and
-diffuse,'' generating an accurate problem statement requires diverse input \cite{rozenshtein_wicked_2018}. Second, for
+Including every relevant group is important for two reasons. First, because information is ``uncertain and diffuse,''
+generating an accurate problem statement requires diverse input \cite{rozenshtein_wicked_2018}. Second, for
 \ac{incrementalism} to work on the irreducibly wicked roots of the problem, there needs to be some degree of consensus
@@
 
+% TODO:
 % Prof. Engelsma
@@
 
-The failures of the \ac{classical-method} and \ac{incrementalism} teach that we must be able to think strategically
+The failures of the \ac{classical-method} and \ac{incrementalism} reveal that we must be able to think strategically
 while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
@@
 
-This follows partially from the previous suggestion---problem definition and analysis are combined precisely because the
-problem definition depends on the nature of proposed solutions. But this advice deserves emphasis for another reason:
-consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
-due to their conflicting values and priorities, but Lindblom writes encouragingly about ``the ease with which individuals
-of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to
-say, ``Labor mediators report a similar phenomenon: the contestants cannot agree on criteria for settling their disputes
-but can agree on specific proposals. Similarly, when one administrator's objective turns out to be another's means, they
-often can agree on policy'' \cite{lindblom_muddling_1959}.
-
-% So should I follow my own advice and focus on
-%   group_2019 (argues for DAR) and
-%   savage_lawful_2018 (a concrete DAR proposal)? -- and related, \cite{servan_schreiber_jje_2020}
+Focusing on concrete proposals follows partially from the previous suggestion---problem definition and analysis are
+combined precisely because the problem definition depends on the nature of proposed solutions. But this advice deserves
+emphasis for another reason: consensus is easier to build around concrete proposals. While debate rages in the abstract,
+groups are bound to disagree due to their conflicting values and priorities. However, Lindblom writes encouragingly
+about ``the ease with which individuals of different ideologies often can agree on concrete policy'' in an example
+about congressional compromise. He goes on to say, ``Labor mediators report a similar phenomenon: the contestants cannot
+agree on criteria for settling their disputes but can agree on specific proposals. Similarly, when one administrator's
+objective turns out to be another's means, they often can agree on policy'' \cite{lindblom_muddling_1959}.
 
@@
 
-The above discussion describes the failure of well-established methods of policymaking at addressing \acp{wicked-prob}
+The discussion above describes the inability of well-established methods of policymaking to address \acp{wicked-prob}
 and properties a better method would have. Here, I propose a modified \acs{OODAloop} as an alternative policymaking
@@
 
-The \acl{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
-countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
-its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
-incident response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your
-opponent's loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.
+The \acl{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
+opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but its ideas
+have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring incident
+response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your opponent's
+loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.
 
@@
 
-The \ac{OODAloop} serves as a good basis for contending with \acp{wicked-prob} because it has the correct fundamental
+The \ac{OODAloop} serves well as a model for contending with \acp{wicked-prob} because it has the correct fundamental
 structure. It maps very well to the \ac{classical-method} steps of problem definition (orientation), ideation and
-analysis (orientation), decision, and action, but it emphasizes the \ac{incrementalism}'s iteration and feedback.
-Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor holds up. One is
-facing an unpredictable opponent in a dynamic environment where every action counts. We cannot overpower
+analysis (orientation), decision, and action; however, it also emphasizes \ac{incrementalism}'s iteration and feedback.
+Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways.
+One is facing an unpredictable opponent in a dynamic environment in which every action counts. We cannot overpower
 \acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.
 
-The one modification the \ac{OODAloop} needs is unification of the problem definition and analysis steps, as one depends
-on the other. This is Rittel's ``argumentative process in the course of which an image of the problem and of the
-solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in \myfig{fig-policy-ooda-loop} as a cycle of
-``collaborative debate'' between observation and orientation. The modified \ac{OODAloop} incorporates each of the
-lessons from \mysec{sec-policy-lessons}.
+The OODA Loop does need one modification in order to suit wicked problems. Because the problem definition step and
+analysis step depend on one another, they should be unified. This is Rittel's ``argumentative process in the course of
+which an image of the problem and of the solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in
+\myfig{fig-policy-ooda-loop} as a cycle of ``collaborative debate'' between observation and orientation. The modified
+\ac{OODAloop} incorporates each of the lessons from \mysec{sec-policy-lessons}.
 
@@
 thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
-steps in the right direction. Both approaches have their strengths, but neither is suited to the complex and
-controversial nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation
-of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
-solutions, and crystallizes discussion around specific proposals. The result from a single iteration is an imperfect
-step forward based on implementing a particular proposal. The result from many iterations is the breakdown of the
-problem into subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
+steps in the right direction. Both approaches have strengths, but neither is suited to the complex and controversial
+nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation of all
+parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based solutions,
+and crystallizes discussion around specific proposals. The result of a single iteration is an imperfect step forward
+based on implementing a particular proposal. The result of many iterations is the breakdown of the problem into
+subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
 \ac{classical-method}, and the wicked subproblems, by now restrained under a sound and agreeable strategy, can be
@@
 
-This process is demonstrated in \myfig{fig-policy-ooda-process}. An out of control problem (the large storm) is being
-confronted. Several iterations of the \ac{OODAloop} gradually bring it down to size, spinning out tame problems (puzzle
-pieces) addressed by classical analysis and wicked but restrained problems (small storms) controlled by
-\ac{incrementalism}.
+This process is demonstrated in \myfig{fig-policy-ooda-process}. An out-of-control problem (the large storm) is
+confronted in this diagram. Several iterations of the \ac{OODAloop} gradually diminish its size. This is achieved by
+spinning out tame problems (puzzle pieces) addressed by classical analysis and wicked but restrained problems (small
+storms) controlled by \ac{incrementalism}.
 
@@
 
-It is important to note that the same \acp{wicked-prob} is never solved twice, since each one is by definition, unique.
+It is important to note that the same \ac{wicked-prob} is never faced twice, since each one is by definition, unique.
 However, changes in technology, culture, and current affairs may render the current strategy insufficient and require
 the process to start again. \Acp{wicked-prob} are like history: they don't repeat themselves, but they rhyme. Due to
-changing circumstances, solutions past will not work in the future, but lessons learned from the past provide useful
-experience.
+changing circumstances, past solutions cannot be used in the future, but lessons learned from the past can be.
 
 In the context of \ac{encryption} and \ac{EA}, this means that the outcome of \ac{the-cw1} doesn't hold precedential
-power over the debate in the second. It means that re-raising questions about the role of technology is playing society
-is legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
-\cite{geller_2019}, though sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
+power over the debate in the second. It means that re-raising questions about the role of technology plays in society is
+legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
+\cite{geller_2019} but sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
 helpful are collaborative efforts involving government, technical, and civil liberties representatives

### 2020-12-10 Thesis-Chapter4-Tools.tex
Editorial improvements

@@
 
-The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
-help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create \acp{argmap}.
+The arguments used in the \ac{EA} debate are diverse and complex. Therefore, in order to advance the \ac{EA} discussion,
+it is necessary to organize the many different arguments in a cohesive manner. This thesis uses \acp{argmap} to do so.
 \Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
 conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
-\acp{wicked-prob}. Though formal research on the technique is thin, \acp{argmap} have proven helpful in simplifying
-complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate by helpfully slowing down
-discussion, depersonalizing conflict, and lending structure and rhythm \cite{dutoit_hypermedia_2006}. This makes them
-well suited as a tool for conducting the collaborative debate that helps problem solvers with disparate views narrow in
-on a problem definition and concrete proposals on which they can agree.
+\acp{wicked-prob}. Though formal research on the technique is sparse, \acp{argmap} have proven helpful in simplifying
+complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate. This is because they slow down
+discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They are
+therefore useful as a tool for conducting the collaborative debate that helps problem-solvers with disparate views
+narrow in on a problem definition and concrete proposals on which they can agree.
 
@@
 
-Threat modeling is an important step in analyzing the security at the systems level. Using models abstracts away fine
-details and focuses on the architecture, processes, and dataflows in a system. Models assist in the understanding of
-current systems and the prevention of problems in new systems, both of which are important when facing the prospect of
-designing an \ac{EA} mechanism built on top of a complex and aging technology stack.
+Threat modeling is an important step in analyzing the security at the systems level. Models abstract away fine details
+and focus on the architecture, processes, and dataflows in a system. Models assist in the understanding of current
+systems and the prevention of problems in new systems, both of which are important when facing the prospect of designing
+an \ac{EA} mechanism built on top of a complex and aging technology stack.
 
 Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
-\cite{shostack_threat_2014}. Answering the first question accurately is crucial, particularly for the field of
-cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. Answering the second
-question depends on the types of attackers to worry about, and determines the scope of threats to be considered. ``What
-are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go wrong?''
-cannot be faithfully answered without knowing what is being built.
+\cite{shostack_threat_2014}. Accurately answering the first question is crucial, particularly for the field of
+cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. The answer to the second
+question depends on the types of attackers under consideration and determines the scope of threats to be considered.
+``What are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go
+wrong?'' cannot be faithfully answered without knowing what is being built.
 
-\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
-external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
-flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of
-primary concern. \myfig{fig-dfd-dh-traditional} shows the basic elements of a \ac{DFD} of a simplified Diffie-Hellman
-key exchange.
+\Acp{DFD} were designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and external
+entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data flow, not
+control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of primary
+concern. \myfig{fig-dfd-dh-traditional} shows the basic elements of a \ac{DFD} of a simplified Diffie-Hellman key
+exchange.
 
@@
 data itself, it eliminates the need for labels on lines and can be used to communicate entire protocols in a single
-diagram. \myfig{fig-dfd-key} contains the full set of symbols. In \mychap{chap-threatmodel} I create \acp{DFD} in this
+diagram. \myfig{fig-dfd-key} contains the full set of symbols. \mychap{chap-threatmodel} includes \acp{DFD} in this
 style to analyze the threats already present and those that \ac{EA} would introduce.

### 2020-12-10 Thesis-Chapter5-Arguments.tex
Restructure contents of background chapter
Work on DFD section and normalize captions
Spruce up title, clean up Tools chapter
Editorial improvements

--- 2020-12-08 Thesis-Chapter4-Arguments.tex
@@
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
-  \caption[Contributing Factors to the EA Debate]{Contributing Factors to the EA Debate}
+  \caption{Contributing Factors to the EA Debate}
   \label{fig-args-contrib}
@@
 arguments spread easily. \myfig{fig-arg-fallacies} shows those used in the encryption and \ac{EA} debate.
-
-% TODO: Dulimarta: Orange/green boxes for consistency? Decide when I make a key.
 
@@
 Despite device key escrow's tractability as a technical niche of \ac{EA}, it still has a long way to go. There are few
-proposals, what few there are are high level, and questions of scaling and administration have not been addressed.
-However, as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific
+proposals, what few there are high level, and questions of scaling and administration have not been addressed. However,
+as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific
 proposals has the opportunity to identify mutually agreeable solutions---or at least refine the problem understanding

### 2020-12-11 Thesis-Chapter3-PolicyMaking.tex
Editorial improvements

@@
 \ac{wicked-prob} \cite{rozenshtein_wicked_2018}. He divides the root issues into three categories. First, there is
-disagreement on what the goals should be (and even basic premises). Sides do not agree on whether things are ``going
-dark'' or we're in the ``golden age for surveillance'' or how much to value competing notions of security. Second,
-information is ``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations.
-Though the consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be
-the case, especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving
-values and technology mean that the subject is always up for renegotiation.
+disagreement on what the goals should be (and even basic premises). Sides do not agree on how much to value competing
+notions of security. Regarding the basic facts, sides do not agree on whether encryption is hiding so much evidence that
+law enforcement is ``going dark,'' as one side puts it, or whether technological change overall has created a ``golden
+age for surveillance,'' as their opponents argue. Second, information is ``uncertain and diffuse.'' Comprehensive data
+regarding encryption's effect on investigations is unavailable. Although the consensus is that we are not presently
+capable of acceptably secure \ac{EA}, that consensus could change, especially considering that \ac{EA} as a field is
+under-researched. Third, that the problem cannot be definitively solved. Evolving values and technology mean that this
+policy area is always up for renegotiation.
 
 \Ac{encryption} technology is particularly sensitive to the irreversible effects of policy implementation. Technology
-deployments have long tails, and attackers have the ability to record and store data for later analysis. In one
-investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile device that they
-cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the only ones to
-benefit from a policy miscalculation that could negatively impact security for a very long time.
+deployments have long tails and attackers have the ability to record and store data for later analysis. In one
+investigation, Australian police cracked a cold case based on evidence acquired from a mobile device that they cracked
+five years after it was seized \cite{evans_arrests_2020}. In this case, it was lawful authorities that benefitted from a
+vulnerability. However, if miscalculated \ac{EA} mandates result in vulnerabilities such as this, attackers will
+benefit, not just law enforcement.
 
@@
 
-Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
-\acp{wicked-prob}. The next two sections investigate strategies for dealing with it.
+Reality must be accepted before it can be dealt with. \cite{baker_2019}. We have by now embraced the reality of
+\acp{wicked-prob}. The next two sections investigate strategies for dealing with this reality.
 
@@
 except for the ``refinement'' step, in which the method analyzes outcomes and corrects errors. However, it is important
-to note that refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
-solution.
+to note that refinement represents a reinforcement, as opposed to a reassessment, of the chosen solution.
 
@@
 \Ac{incrementalism} is an intuitive and iterative approach posed as an alternative to the \ac{classical-method}.
-\Ac{incrementalism} as a policymaking strategy is often referred to as ``\ac{muddling-through}'' after political
+\Ac{incrementalism} as a policymaking strategy is often referred to as ``\ac{muddling-through}'' from political
 scientist Charles Lindblom's classic paper defining and defending the approach \cite{lindblom_muddling_1959}. Written
@@
 Lindblom's process of ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis.
-The alternatives selected for comparison are defined relative to the status quo and are close enough to one another to
-be able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
-predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
-impossible to impose in a democratic system \cite{lindblom_muddling_1959}. The formality of the process varies;
-policymakers may use this method consciously, with considerable comparative analysis, or unconsciously, led by
-intuition. \myfig{fig-muddling-through} illustrates this method.
+The alternatives selected for comparison must be defined relative to the status quo and must be close enough to one
+another that they can be analyzed on the margin. This is done due to (a) practical necessity, due to the inability to
+predict policy outcomes, and (b) out of political realism, as non-incremental changes are usually politically impossible
+to impose in a democratic system \cite{lindblom_muddling_1959}. The formality of the process varies; policymakers may
+use this method consciously, with considerable comparative analysis, or unconsciously, led by intuition.
+\myfig{fig-muddling-through} illustrates this method.
 
@@
 \Ac{incrementalism} has several advantages over the \ac{classical-method} for handling \acp{wicked-prob}. It is rooted
-in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved,
-emphasizing iteration (``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
+in realism about the limits of rational analysis. It accepts that the problem will not be conclusively solved. Instead,
+it emphasizes iteration: ``Policy is not made once and for all; it is made and re-made endlessly. Policy-making is a
 process of successive approximation to some desired objectives in which what is desired itself continues to change under
-reconsideration.'' \cite{lindblom_muddling_1959}). Most importantly, it eschews the linear
-problem-definition--alternatives-analysis steps out of respect that this is, indeed, impossible. Lindblom argues at
-length that policy ends and means are interlinked, eventually concluding:
+reconsideration'' \cite{lindblom_muddling_1959}. Most importantly, it eschews the linear
+problem-definition--alternatives-analysis steps out of respect that the problem definition is not self-contained.
+Lindblom argues that policy ends and means are interlinked, eventually concluding:
 
@@
 ``overlook excellent policies for no other reason than that they are not suggested by the chain of successive policy
-steps leading up to the present'' \cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to
-the next can result in messy policy that none are particularly fond of---``As Lindblom's sobriquet suggests, it often
-[leads] to a considerable muddle'' \cite{feeley_judicial_2000}. Application of the \acl{CFAA} is one such muddle
-\cite{wolff_computer_2016}.
+steps leading up to the present'' \cite{lindblom_muddling_1959}. Analysis based only on the present status quo can
+result in messy policy that pleases no one---``As Lindblom's sobriquet suggests, it often [leads] to a considerable
+muddle'' \cite{feeley_judicial_2000}. Application of the \acl{CFAA} is one such muddle \cite{wolff_computer_2016}.
 
-Unfortunately, one cannot simply expand the strategic view of the incrementalist method by allowing it to look further
-into the past than the present status quo. Simplifying analysis by limiting it to marginal differences to a given
-baseline is at the heart of \ac{incrementalism}. If one tries to be a ``strategic incrementalist'' by looking into the
-past, they still have to choose a baseline from which to perform analysis. This approach is susceptible to two
-weaknesses in baseline-based reasoning that Rozenshtein describes in his analysis of \ac{EA} as a \ac{wicked-prob}
-\cite{rozenshtein_wicked_2018}. First, the choice of baseline is arbitrary, yet heavily colors analysis:
+Unfortunately, one cannot simply add high level strategic analysis to the incrementalist method by including the past in
+its analysis. \Ac{incrementalism} relies on simplifying analysis by limiting it to marginal differences to a given
+baseline. If one tries to be a ``strategic incrementalist'' by looking into the past, they still have to choose a
+baseline from which to perform analysis. This approach is susceptible to two weaknesses in baseline-based reasoning that
+Rozenshtein describes in his analysis of \ac{EA} as a \ac{wicked-prob} \cite{rozenshtein_wicked_2018}. First, the choice
+of baseline is arbitrary, yet heavily colors analysis:
 
@@
 
-Let us define one strategy for handling \ac{wicked-prob} as corralling it back into the classes of problems we are
-better at dealing with. That means breaking it into tame subproblems and developing agreeable strategies for its
-irreducibly wicked elements. Once that is done, we can rationally analyze and increment our way to victory. There is no
-handbook for how to do this, but several sources offer advice.
+\Acp{wicked-prob} were earlier described as a tangle of contributing subproblems. Using this characterization, dealing
+with a \ac{wicked-prob} requires disentangling the web, identifying the tame subproblems, and developing agreeable
+strategies for the irreducibly wicked subproblems. Once that is done, we can rationally analyze and increment our way to
+resolution. There is no handbook for how to do this, but several sources offer advice.
 
@@
 Easy answers, or any solutions that artificially tame the problem, will not bring the matter under control. While by
-definition there is no step that will truly solve the problem, easy answers deliberately emphasize one value to the
-exclusion of others. This leads to chaos in those neglected areas. Solutions based on easy answers produce unintended
-consequences because they ignore the root issues \cite{commission_tackling_2018}.
+definition there is no solution that will truly solve a \ac{wicked-prob}, easy answers deliberately emphasize one value
+to the exclusion of others. Because they neglect the root issues, solutions based on easy answers produce unintended
+consequences and chaos in those neglected areas \cite{commission_tackling_2018}.
 
@@
 
-% TODO:
-% Prof. Engelsma
+% TODO: Prof. Engelsma:
 % - A big part of the "solution" is getting people together
@@
 The failures of the \ac{classical-method} and \ac{incrementalism} reveal that we must be able to think strategically
-while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
-classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
-practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
-puts it, ``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
+while respecting the non-linear nature of \acp{wicked-prob}. One must use the high level, holistic view of the classical
+approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In practice,
+this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel puts it,
+``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
 wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
@@
 
-We know that proposals cannot be comprehensively evaluated before or even after implementation, and that each action (or
-decision not to act) has irreversible effects. \Ac{incrementalism} is the decision making model to follow here, as it is
-inherently responsive to unpredictable changes in the environment. Proposals must therefore be agile. All decisions
-involve unknowns, but risk- and uncertainty-management strategies can optimize the expected outcome and maximize the
-worst outcome \cite{sunstein_beyond_2015}.
+\Acp{wicked-prob}' potential solutions cannot be comprehensively evaluated before or even after implementation and each
+action (or decision not to act) has irreversible effects. Proposals must therefore be agile. All decisions involve
+unknowns, but risk- and uncertainty-management strategies can optimize the expected outcome and maximize the worst
+outcome \cite{sunstein_beyond_2015}.
 
@@
 
-Focusing on concrete proposals follows partially from the previous suggestion---problem definition and analysis are
-combined precisely because the problem definition depends on the nature of proposed solutions. But this advice deserves
-emphasis for another reason: consensus is easier to build around concrete proposals. While debate rages in the abstract,
+Focusing on concrete proposals follows partially from the previous lesson---problem definition and analysis are combined
+precisely because the problem definition depends on the nature of proposed solutions. But this advice deserves emphasis
+for another reason: consensus is easier to achieve for concrete proposals. Debate in the abstract can rage endlessly, as
 groups are bound to disagree due to their conflicting values and priorities. However, Lindblom writes encouragingly
-about ``the ease with which individuals of different ideologies often can agree on concrete policy'' in an example
-about congressional compromise. He goes on to say, ``Labor mediators report a similar phenomenon: the contestants cannot
-agree on criteria for settling their disputes but can agree on specific proposals. Similarly, when one administrator's
+about ``the ease with which individuals of different ideologies often can agree on concrete policy'' in an example about
+congressional compromise. He goes on to say, ``Labor mediators report a similar phenomenon: the contestants cannot agree
+on criteria for settling their disputes but can agree on specific proposals. Similarly, when one administrator's
 objective turns out to be another's means, they often can agree on policy'' \cite{lindblom_muddling_1959}.
@@
 The discussion above describes the inability of well-established methods of policymaking to address \acp{wicked-prob}
-and properties a better method would have. Here, I propose a modified \acs{OODAloop} as an alternative policymaking
-model.
+and lessons for shaping a better method. Here, I propose a modified \acs{OODAloop} as an alternative policymaking model.
 
@@
 The \ac{OODAloop} serves well as a model for contending with \acp{wicked-prob} because it has the correct fundamental
-structure. It maps very well to the \ac{classical-method} steps of problem definition (orientation), ideation and
+structure. It corresponds closely to the \ac{classical-method} steps of problem definition (orientation), ideation and
 analysis (orientation), decision, and action; however, it also emphasizes \ac{incrementalism}'s iteration and feedback.
-Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways.
-One is facing an unpredictable opponent in a dynamic environment in which every action counts. We cannot overpower
-\acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.
+Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways. In
+both cases, one faces an unpredictable opponent in a dynamic environment in which every action counts. We cannot
+overpower \acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.
 
 The OODA Loop does need one modification in order to suit wicked problems. Because the problem definition step and
-analysis step depend on one another, they should be unified. This is Rittel's ``argumentative process in the course of
+analysis step depend on one another, they must be joined. This is Rittel's ``argumentative process in the course of
 which an image of the problem and of the solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in
@@
 
-Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
-thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
+Traditional policymaking typically follows either (a) a rational approach rooted in the scientific method, aiming to be
+thorough and complete, or (b) an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
 steps in the right direction. Both approaches have strengths, but neither is suited to the complex and controversial
-nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation of all
-parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based solutions,
-and crystallizes discussion around specific proposals. The result of a single iteration is an imperfect step forward
-based on implementing a particular proposal. The result of many iterations is the breakdown of the problem into
-subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
-\ac{classical-method}, and the wicked subproblems, by now restrained under a sound and agreeable strategy, can be
-addressed with \ac{incrementalism}.
+nature of \acp{wicked-prob}. The strategy proposed here relies on the goodwill participation of all parties who engage
+in ongoing research and debate that rejects easy answers, encourages flexible risk-based solutions, and crystallizes
+discussion around specific proposals. The result of a single iteration is a bit more clarity and a small step forward.
+The result of many iterations is the breakdown of the problem into subproblems---some tame and inevitably some still
+wicked. The tame subproblems can be addressed with the \ac{classical-method}, and the wicked subproblems, by now
+restrained under a sound and agreeable strategy, can be addressed with \ac{incrementalism}.
 
-% TODO: do I need this explanation?
-
-This process is demonstrated in \myfig{fig-policy-ooda-process}. An out-of-control problem (the large storm) is
-confronted in this diagram. Several iterations of the \ac{OODAloop} gradually diminish its size. This is achieved by
-spinning out tame problems (puzzle pieces) addressed by classical analysis and wicked but restrained problems (small
-storms) controlled by \ac{incrementalism}.
+This process is demonstrated in \myfig{fig-policy-ooda-process}. An out-of-control problem (represented by the large
+storm) is confronted in this diagram. Several iterations of the \ac{OODAloop} gradually diminish its size. This is
+achieved by spinning out tame problems (represented by puzzle pieces) which are addressed with classical analysis and
+wicked but restrained problems (represented by small storms) which are controlled by \ac{incrementalism}.
 
@@
 However, changes in technology, culture, and current affairs may render the current strategy insufficient and require
-the process to start again. \Acp{wicked-prob} are like history: they don't repeat themselves, but they rhyme. Due to
-changing circumstances, past solutions cannot be used in the future, but lessons learned from the past can be.
+the process to start again. It has been said that history does not repeat itself, but it rhymes. \Acp{wicked-prob} are
+the same. Past solutions cannot be used in the future, but lessons learned can be.
 
 In the context of \ac{encryption} and \ac{EA}, this means that the outcome of \ac{the-cw1} doesn't hold precedential
-power over the debate in the second. It means that re-raising questions about the role of technology plays in society is
-legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
-\cite{geller_2019} but sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
-helpful are collaborative efforts involving government, technical, and civil liberties representatives
-\cite{committee_decrypting_2018} \cite{group_2019}. Specific proposals around which groups can center discussion,
-whether they specify a particular form of \ac{EA} or perhaps especially if they don't, are also a necessary component
-\cite{kerr_encryption_2017} \cite{wright_crypto_2018} \cite{phan_key_2017}.
+power over the debate in the second. It means that it is legitimate to re-raise questions about the role technology
+plays in society. Hard-line rhetoric doesn't help \cite{ruiz_there_2018} \cite{geller_2019}, but sincere appeals do
+\cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Collaborative efforts involving government, technical, and
+civil liberties representatives are even better \cite{committee_decrypting_2018} \cite{group_2019}. Specific proposals
+around which groups can center discussion are also necessary. This is true whether they specify a particular form of
+\ac{EA} or (perhaps especially) if they offer an alternative \cite{kerr_encryption_2017} \cite{wright_crypto_2018}
+\cite{phan_key_2017}.
 
-In the remainder of this thesis, I apply this strategy by documenting the debate in depth in \mychap{chap-arguments} and
-through creation of a threat model against which I analyze one specific proposal in \mychap{chap-threatmodel}.
+I apply this strategy in the remainder of the thesis. \mychap{chap-arguments} documents the debate in depth and
+\mychap{chap-threatmodel} defines a threat model against which I analyze a specific \ac{EA} proposal.

### 2020-12-11 Thesis-Chapter4-Tools.tex
Editorial improvements

@@
 
-Encryption policy arguments and \acl{EA} threat models are complex. \Acp{argmap} and \aclp{DFD} two methodological tools
-to structure analysis and reduce complexity. This chapter introduces these tools and their syntax as used in
-\mychap{chap-arguments} and \mychap{chap-threatmodel}.
+Encryption policy arguments and \acl{EA} threat models are complex. \Acp{argmap} and \aclp{DFD} are two methodological
+tools that can be used to structure analysis and manage complexity. This chapter introduces these tools and their
+syntax, which will be used in \mychap{chap-arguments} and \mychap{chap-threatmodel}.
 
@@
 
-The arguments used in the \ac{EA} debate are diverse and complex. Therefore, in order to advance the \ac{EA} discussion,
-it is necessary to organize the many different arguments in a cohesive manner. This thesis uses \acp{argmap} to do so.
-\Acp{argmap} are graphical representations of the logical structures and relationships between statements, premises, and
-conclusions. \Acp{argmap} are an outgrowth of the same research by Horst Rittel that originated the idea of
+In order to advance the \ac{EA} discussion, it is necessary to organize the many different arguments in a cohesive
+manner. This thesis uses \acp{argmap} to do so. \Acp{argmap} are graphical representations of the logical structures and
+relationships between statements, premises, and conclusions. They originated from the same research as the idea of
 \acp{wicked-prob}. Though formal research on the technique is sparse, \acp{argmap} have proven helpful in simplifying
-complex arguments around \acp{wicked-prob} \cite{renton_2007} and in facilitating debate. This is because they slow down
-discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They are
-therefore useful as a tool for conducting the collaborative debate that helps problem-solvers with disparate views
-narrow in on a problem definition and concrete proposals on which they can agree.
+complex arguments regarding \acp{wicked-prob} \cite{renton_2007} and in facilitating debate. This is because they slow
+down discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They
+are therefore useful as a tool for conducting the \ac{OODAloop} model's ``collaborative debate'' step.
 
-In 1970 Rittel's research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the reasoned
-approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some software
-tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
-\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
-generates \acp{argmap} from structures specified in a Markdown-based language \cite{voigt_argdown_2018}. Argdown is used
-in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA} debate.
+Argument mapping tools originated in 1970 with text-based \acp{IBIS} designed to break down \acp{wicked-prob} and
+document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \acp{IBIS} has waxed and
+waned. Some software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
+\cite{dutoit_hypermedia_2006}, which were designed to produce graphical argument representations. Argdown is a recent
+tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
+\cite{voigt_argdown_2018}. Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA}
+debate.
 
-\myfig{fig-args-key} shows the basic structure of arguments. Statements are positions, values, and assertions. The
-``Proposals,'' ``Statements,'' ``Fallacies,'' and ``Values'' in the figure are classes of statements in this sense. The
-color of a statement's border indicates the class to which it belongs. Arguments are premise-conclusion structures that
-relate to statements. Arguments as classified as ``anti-EA,'' ``pro-EA,'' ``neutral,'' or ``fallacious.'' The color of
-an argument's solid background indicates the class to which it belongs. Statements and arguments may support, attack, or
-undercut one another. Supporting relationships have green arrows; attacking, red; and undercutting, purple.
+\myfig{fig-args-key} shows the basic structure of arguments. Statements can take the form of positions, values, and
+assertions. The ``Proposals,'' ``Statements,'' ``Fallacies,'' and ``Values'' in the figure are classes of statements in
+this sense. The color of a statement's border indicates the class to which it belongs. Arguments are premise-conclusion
+structures that relate to statements; they are classified as ``anti-EA,'' ``pro-EA,'' ``neutral,'' or ``fallacious.'' An
+argument's class is distinguished by the color of its solid background. Statements and arguments may support, attack, or
+undercut one another. Green arrows represent supporting relationships; red arrows, attacking; and purple arrows,
+undercutting.
 
@@
 Threat modeling is an important step in analyzing the security at the systems level. Models abstract away fine details
-and focus on the architecture, processes, and dataflows in a system. Models assist in the understanding of current
-systems and the prevention of problems in new systems, both of which are important when facing the prospect of designing
-an \ac{EA} mechanism built on top of a complex and aging technology stack.
+and focus on the architecture, processes, and dataflows in a system. They assist in the understanding of current systems
+and the prevention of problems in new systems, both of which are important when facing the prospect of designing an
+\ac{EA} mechanism built on top of a complex and aging technology stack.
 
@@
 \Acp{DFD} were designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and external
-entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data flow, not
-control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is of primary
+entities. \Acp{DFD} are well-suited for threat modeling because security vulnerabilities tend to follow data flow, not
+control flow \cite{shostack_threat_2014}. They are particularly well-suited for \ac{EA}, as data privacy is of primary
 concern. \myfig{fig-dfd-dh-traditional} shows the basic elements of a \ac{DFD} of a simplified Diffie-Hellman key
@@
 diagram. \myfig{fig-dfd-key} contains the full set of symbols. \mychap{chap-threatmodel} includes \acp{DFD} in this
-style to analyze the threats already present and those that \ac{EA} would introduce.
+style to analyze threats that EA would introduce.
 

### 2020-12-11 Thesis-Chapter5-Arguments.tex
Editorial improvements

@@
 
-In this chapter we will analyze the \ac{encryption} debate using \acp{argmap}. The maps help one digest the debate, but
-there are two important shortcomings of the ones presented here. First, argument mapping is ideally just one tool in the
-collaborative process that progressively approaches the problem definition and solution. One can make an approximation
-from research, but it is not a replacement for debate arising from live discussion. Second, \acp{argmap} as portrayed in
-this thesis can be deceiving about the strength of an argument. Argument nodes show no indication of their strength or
-validity; one can formulate ten poor arguments against a single strong one, and a glance at the map will indicate that
-the case of the ten is stronger. With those qualifications in mind, let us begin by looking at the factors at the heart
-of the conflict.
+This chapter will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
+debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
+as tools in the collaborative pursuit of the problem definition and solution. One can approximate the debate based on
+research, but it is not a replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively
+portray the strength of an argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments
+were presented against one indisputable argument, the side with more arguments would, regardless of their integrity,
+appear stronger. With those qualifications in mind, let us begin by examining the factors at the center of the conflict.
 
@@
 
-\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, it is \ac{encryption}'s dual
-role as enabler of radical privacy and cornerstone of security is at the heart of the debate. Of course, there is only a
-conflict if one sees radical privacy as a bad thing. The right to privacy is both a strongly held value and an enshrined
-legal principle, and many use privacy-based arguments to show \ac{EA} as socially undesirable. The Snowden revelations
-\cite{landau_making_2013} only unveiled the scale of privacy-eroding U.S. government surveillance enabled by
-technological changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Violations of privacy are
-concerning for those most vulnerable and for society at large. For example, Saudi journalist Jamal Khashoggi's private
-communications were surveilled by the Saudi government before he became a victim of politically motivated murder
-\cite{liebermann_2019}. At a societal level, presence of surveillance alone changes behavior and chills free speech
-\cite{rogaway_moral_2015}, while government violations of the law degrade institutional trust.
+\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, \ac{encryption}'s dual
+contributions to information security and radical privacy are central to the debate. Of course, there is only conflict
+when radical privacy is perceived as a negative. The right to privacy is both a strongly held value and an enshrined
+legal principle. Many use privacy concerns to argue that \ac{EA} is socially undesirable. The Snowden revelations
+\cite{landau_making_2013} unveiled the scale of privacy-eroding U.S. government surveillance enabled by technological
+changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract worry
+to many, but for the most vulnerable, they are frighteningly concrete. For example, Saudi journalist Jamal Khashoggi's
+fell
+% TODO: bookmark
+
+private communications were surveilled by the Saudi government before he became a victim of politically motivated murder
+\cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free speech
+\cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.
 
@@
 
-The government, and particularly \acl{LE}, argues that encryption is handcuffing their investigational capacity. The
-\ac{FBI} has branded strong encryption as ``warrant-proof'' and states that ``the government often cannot obtain the
-electronic evidence necessary to investigate and prosecute threats to public and national safety'' \cite{fbi_2020}. The
-pro-\ac{EA} argument is typically rooted in public safety: investigators need access to evidence to fulfill their
-commission to catch and convict lawbreakers, and encryption is making that too difficult. Five-eyes nations regularly
-put out joint statements expressing their frustration with encryption \cite{ministerial_2018} \cite{goodale_2017}, and
-were joined in 2020 by India and Japan in a statement emphasizing the growing danger of \ac{CSAM} \cite{intl_2020}.
-Child pornography is one the original \ac{horsemen} that some argue are used to scare people into supporting backdoors
-\cite{schneier_scaring_2019}, but it is also a problem growing to astonishing proportions. The \textit{New York Times}'s
-2019 investigation on the matter quoted one law enforcement officer as estimating that 400,000 New Jerseyans, more than
-4\% of the state's population, could be charged with violating child exploitation material laws
+\Acl{LE}, argues that encryption handcuffs its investigational capacity. The \ac{FBI} brands strong encryption as
+``warrant-proof'' and states that ``the government often cannot obtain the electronic evidence necessary to investigate
+and prosecute threats to public and national safety'' \cite{fbi_2020}. The pro-\ac{EA} argument is typically rooted in
+public safety: encryption makes it too difficult for investigators to access the evidence needed to catch and convict
+law-breakers. Five-eyes nations have regularly published joint statements expressing their frustration with encryption
+\cite{ministerial_2018} \cite{goodale_2017}. They were joined in 2020 by India and Japan in a statement emphasizing the
+growing danger of \ac{CSAM} \cite{intl_2020}. Child pornography is one of the original \ac{horsemen} used to scare
+people into supporting backdoors \cite{schneier_scaring_2019}, but it is also a problem growing to astonishing
+proportions. A 2019 investigation by the \textit{New York Times} quoted a law officer's estimate that 400,000 New
+Jerseyans, more than 4\% of the state's population, have violated child exploitation material laws
 \cite{keller_internet_2019}.
 
-Unfortunately, shocking assertions such as this have not been independently confirmed or backed up with hard data.
-Intelligence agencies have a long history or misrepresenting positions and overstepping bounds
-\cite{johnson_congressional_2004} \cite{shamsi_2011}, and this has led to lack of institutional trust undercutting their
-claims that they need \ac{EA}. Indeed, the \ac{FBI} has already been caught exaggerating by several times how many
-mobile devices they could not access due to \acl{DE} \cite{devlin_2018}. As Rozenshtein says on this subject, ``It is
-impossible to know the precise extent to which encryption frustrates law-enforcement investigations, both because
-law-enforcement agencies are only beginning to collect accurate statistics, and because one can never be sure of how an
-investigation would have proceeded in the absence of encryption'' \cite{rozenshtein_wicked_2018}, but it is still
-crucial to have an accurate depiction of the problem in order to come to justified and helpful solutions. This argument
-is analyzed further in the next section.
+Unfortunately, these shocking claims have not been independently confirmed with hard data. Intelligence agencies have
+historically misrepresented statistics and overstepped their bounds \cite{johnson_congressional_2004}
+\cite{shamsi_2011}. This undercuts their claims that they need \ac{EA} and has diminished institutional trust. For
+example, the \ac{FBI} has already been found to exaggerate the number of mobile devices it could not access due to
+\acl{DE} \cite{devlin_2018}. As Rozenshtein points out, ``It is impossible to know the precise extent to which
+encryption frustrates law-enforcement investigations, both because law-enforcement agencies are only beginning to
+collect accurate statistics, and because one can never be sure of how an investigation would have proceeded in the
+absence of encryption'' \cite{rozenshtein_wicked_2018}. However, it is still crucial to have an accurate depiction of
+the problem in order to come to justified and helpful solutions. This argument is analyzed further in the next section.
 
-Information security is a the remaining central issue in the debate. As described in \mysec{sec-crypto-basics},
-\ac{cryptography} in various forms plays a foundational technical role in nearly every aspect of security, and
-information confidentiality itself is one of the primary properties security practitioners fight to protect. Efforts in
-the past have sought to undermine the cryptographic foundations of encryption, though this is not a feature of proposals
-today, as that is universally seen as too risky. \mysec{sec-tech-approaches} introduced several alternative technical
-approaches to \ac{EA}, but experts still proclaim loud and clear that the current state of the art is far from capable
-of providing the type of access \acl{LE} in an acceptably secure manner \cite{abelson_2015} \cite{abelson_risks_1997}.
-Security was an afterthought in early computing and networking designs, and the field of cybersecurity is still more of
-an art the a science; hobbling systems with a mandate to provide \ac{EA} would inevitably diminish security to some
-degree \cite{abelson_2015}.
+Information security is the remaining central issue in the debate. As described in \mysec{sec-crypto-basics},
+\ac{cryptography} plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought
+to compromise the cryptographic foundations of encryption, though this is not a feature of proposals today, as that is
+universally seen as too risky. \mysec{sec-tech-approaches} introduced several alternative technical approaches to
+\ac{EA}, but experts still proclaim that the current state of the art is far from capable of providing the type of
+access \acl{LE} in an acceptably secure manner \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an
+afterthought in early computing and networking designs, and the field of cybersecurity is still equal parts art and
+science; mandated \ac{EA} would inevitably diminish security to some degree \cite{abelson_2015}.
 
-Finally, one may argue that \ac{EA}, depending on its implementation, could have positive auxiliary use cases. It could
-enable malware scanning, administrator access in a business setting, or password recovery. Typically these use cases
-have other solutions, however, and to combine the requirements of these uses with those of any reasonable \ac{EA}
-proposal would likely fatally weaken it.
+Finally, it is arguable that \ac{EA} could have positive alternative applications. It could enable malware scanning,
+password recovery, or administrator access in a business setting. Typically these applications have other solutions,
+however, and to combine the requirements of these uses with those of any reasonable \ac{EA} proposal would likely
+fatally weaken it.
 
@@
 whether law enforcement has \ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is
-well-represented by proponent former \ac{FBI} Director James Comey in a 2014 speech titled ``Going Dark''
-\cite{comey_2014} (though he did not coin the phrase \cite{swire_encryption_2011}). The argument is represented in
-\myfig{fig-arg-going-dark}, also informed by critiques \cite{rogaway_moral_2015}.
+well-represented by proponent former \ac{FBI} Director James Comey in his 2014 speech titled ``Going Dark''
+\cite{comey_2014} (though he did not coin the phrase \cite{swire_encryption_2011}). The argument is presented in
+\myfig{fig-arg-going-dark} including several critiques \cite{rogaway_moral_2015}.
 
@@
 
-The ``Going Dark'' argument goes that privacy and safety are both desirable goods, but they inherently conflict, and we
-need balance. Individuals must give up some of their personal good of privacy to make room for the public good of
-safety. Encryption tips the scales too far towards privacy, upsetting the balance of the status quo. \ac{EA} is needed
-to restore the balance and enable law enforcement agencies to do their jobs and protect the public.
+According to the "Going Dark" argument, privacy and safety are both desirable goods, but they conflict with one another;
+therefore, they must be balanced. Individuals must sacrifice some of their personal good of privacy to in honor of the
+public good of safety. Widespread encryption tips the scales too far towards privacy, upsetting the balance we had
+before. \ac{EA} restores the balance and enables law enforcement agencies to fulfill their duty to protect the public.
 
-The argument has a few blind spots. Depicting privacy and safety as competing in a zero-sum tradeoff ignores all the
-ways in which privacy directly enhances safety. Data privacy makes certain crimes more difficult, such as stalking or
-identity theft. Privacy is also a public good to the extent that it precludes \ac{masssurv}, an argument explored more
-deeply in the ``Golden Age'' side of the argument.
+The ``Going Dark'' argument has a few shortcomings. It oversimplifies the relationship between privacy and safety by
+depicting it as a zero-sum conflict. In reality, privacy does not always diminish safety, but sometimes enhances safety.
+For example, data privacy makes crimes such as stalking and identify more difficult. The ``Going Dark'' argument also
+categorizes privacy as a personal good and safety as a public good. However, privacy can also be a public good when it
+precludes \ac{masssurv}. This argument is explored more deeply in the ``Golden Age for Surveillance'' map.
 
-Most importantly, the ``Going Dark'' argument hinges on encryption being responsible for failure of law enforcement
-investigations. As described in \mysec{sed-arg-contrib}, there is a lack of data on the extent to which encryption is
-itself to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an issue. The
-highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended with the \ac{FBI} breaking into the phone but finding nothing
-of value. Tellingly, each of Comey's examples in his 2014 speech were not cases where encryption inhibited
-investigations, but where cases hypothetically \ii{could} have been inhibited if encryption was a factor
-\cite{comey_2014}. More recently, an 2020 report shows that law enforcement agencies currently have the upper hand on
-mobile device encryption, and are actually able to extract data from nearly every mobile phone on the market
-\cite{koepke_2020}.
+Most importantly, the strength of the ``Going Dark'' argument rests on law enforcement investigations failing due to
+encryption. As described in \mysec{sed-arg-contrib}, there is no conclusive data regarding the extent to which
+encryption is to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an
+issue. The highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended when the \ac{FBI} broke into the phone and found
+nothing of value. Tellingly, the examples cited in Comey's 2014 speech were not cases in which encryption inhibited
+investigations, but were cases in which encryption \ii{could have} inhibited investigations, had it been a factor.
+\cite{comey_2014}. More recently, a 2020 report titled ``Mass Extraction: The Widespread Power of U.S. Law Enforcement
+to Search Mobile Phones'' revealed that law enforcement agencies currently have the upper hand on mobile device
+encryption and are actually able to extract data from nearly every mobile phone on the market \cite{koepke_2020}.
 
 Citing lack of evidence that encryption is stymieing cases only gets you so far, however. In most cases, it is
-impossible to know what would have happened in an investigation unimpeded by strong encryption. This is part of why it
-is a \ac{wicked-prob}. Indeed, when the encryption curtain is pulled back, there is often much to see. Unencrypted
-Facebook Messenger is has approximately 20\% market share in active monthly users \cite{statista_2020} but accounts for
-65\% of \ac{CSAM} reports \cite{keller_internet_2019}, and European law enforcement's infiltration of encrypted
-messaging service Encrochat exposed crime rings and led to a massive raids \cite{cox_2020}.
+impossible to know what would have happened without the interference of strong encryption. Inability to perform
+controlled tests is a characteristic of \ac{wicked-prob}. When controlled settings are unachievable, another way to
+gather data is by observing naturally-occurring conditions where everything but the variable of interest is the same.
+These conditions exist for encryption in messaging application ecosystem. Messaging applications WhatsApp, Facebook
+Messenger, and Telegram all have large user bases have approximate feature parity. Among these services, only Facebook
+Messenger does not employ \ac{E2EE}. What effect does that have on investigation? Globally, Facebook Messenger has
+approximately 20\% market share as measured by active monthly users \cite{statista_2020} but accounts for 65\% of
+\ac{CSAM} reports \cite{keller_internet_2019}.
+
+% TODO: I still really want to use this!
+% European law enforcement's infiltration of encrypted messaging service Encrochat exposed crime rings and led to a
+% massive raids \cite{cox_2020}.
 
@@
 
-The argument for law enforcement having \ii{too much} access to data claims that instead of law enforcement ``Going
-Dark,'' they are living in the ``Golden Age for Surveillance'' \cite{swire_encryption_2011}. This argument goes that
-although encryption makes certain data completely unaccessible, technological change on the whole has given authorities
-much more digital evidence than encryption has taken away. Paired with the dangers of \ac{masssurv}, encryption is a
-necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument is mapped in
-\myfig{fig-arg-golden-age}.
+The argument for law enforcement having \ii{too much} access to data claims that law enforcement is not ``Going Dark'';
+rather, they are experiencing the ``Golden Age for Surveillance'' \cite{swire_encryption_2011}. This argument posits
+that although encryption makes certain data completely unaccessible, technological change on the whole has given
+authorities much more digital evidence than encryption has taken away. Paired with the dangers of \ac{masssurv},
+encryption is a necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument
+is mapped in \myfig{fig-arg-golden-age}.
 
@@
 
-The ``Golden Age'' argument revolves around surveillance, its undesirability, and technology's role in it. Cryptographer
+The ``Golden Age'' argument focuses on surveillance, its undesirability, and technology's role in it. Cryptographer
 Phillip Rogaway writes at length about the negative impacts of \ac{masssurv} in his essay ``The Moral Character of
-Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway argues that ``pervasive collection \ii{itself} chills
-free-speech and threatens liberal democracy'' (emphasis his) through the way its mere existence affects the subject. He
-argues that it is fundamentally a tool of power, and though I would add that power from the people created through
-representative political process is valid power, clear abuses have undercut reasonable confidence in these institutions.
+Cryptographic Work'' \cite{rogaway_moral_2015}. Rogaway asserts that ``pervasive collection \ii{itself} chills
+free-speech and threatens liberal democracy.'' He argues that surveillance is fundamentally a tool of power, and for
+this reason it should be resisted. Though power from the people created through representative political process is
+valid, clear abuses have undercut reasonable confidence in these institutions.
 
-The next element is that technology has increased the government's surveillance powers. This can hardly be contested.
-Going back to Comey's ``Going Dark'' examples once more, each case he cited was cracked by digital evidence that would
-not have been available even twenty years ago \cite{comey_2014}. In response to the danger of \ac{masssurv} and the
-scale of modern surveillance capabilities, encryption is one of the only privacy-saving tools. (Opting out may be
-possible for some, but is not a realistic option for society as a whole.)
+After establishing the danger of surveillance, the argument notes how technology has increased the government's
+surveillance abilities. This can hardly be contested. Going back to Comey's ``Going Dark'' examples once more, each case
+he cited was cracked by digital evidence that would not have been available even twenty years ago \cite{comey_2014}. In
+response to the danger of \ac{masssurv} and the scale of modern surveillance capabilities, encryption is one of the only
+privacy-saving tools. (Opting out may be possible for some, but is not a realistic option for society as a whole.)
 
-While much stronger than the ``Going Dark'' argument, the ``Golden Age for Surveillance'' argument suffers from one
-major flaw. It focuses strongly on the abilities of and dangers posed by agencies like the \ac{NSA} and ignores the
-problems of local, state, and sometimes-surprisingly-cash-strapped federal law enforcement. The boom in the existence of
-digital evidence is clear, but as shown, law enforcement already has trouble acquiring many forms of evidence with the
-help of tech and communications companies, and does not have the funding or capability to get it on its own
-\cite{carter_2018}. Exacerbating the situation is the ``tech effect,'' which is increasing jurors' expectations of hard
-evidence in cases where it may be suspected to exist \cite{shelton_study_2006}.
+While stronger than the ``Going Dark'' argument, the ``Golden Age for Surveillance'' argument suffers from a significant
+flaw. It focuses on the abilities and dangers of agencies like the \ac{NSA} while ignoring the problems of
+resource-strapped local, state, and federal law enforcement. It is obvious that the amount of digital evidence is
+increasing. However, law enforcement already has trouble acquiring many forms of evidence with the help of tech and
+communications companies, and does not have the funding or capability to get it on its own \cite{carter_2018}.
+Exacerbating the situation is the ``tech effect,'' which refers to jurors' increased expectations of hard evidence in cases
+where it may be suspected to exist \cite{shelton_study_2006}.
 
-Still, the fact that the tools of \ac{masssurv} are out of reach for most law enforcement must not be relied upon. That
-is, most likely, a temporary situation. Law enforcement's present ability to access mobile devices proves this point. A
-safer criticism of the ``Golden Age'' framing of \ac{EA} is that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA}
-implementations that enable law enforcement but thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age''
-argument. \ac{EA} proposals striving for these properties are explored in future argument maps.
+Still, law enforcement's inability to access the tools of \ac{masssurv} cannot be taken for granted. It will likely
+change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
+in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement but
+thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age'' argument. \ac{EA} proposals that strive for these
+properties will be explored in \mysec{sec-ea-types}.
 
@@
 
-Analyzing debates in the \ac{argmap} format invites one to seriously consider the merits of opposing arguments. Part of
-that consideration includes identifying and eliminating fallacious arguments. In heated, partisan debates, fallacious
-arguments spread easily. \myfig{fig-arg-fallacies} shows those used in the encryption and \ac{EA} debate.
+Using \acp{argmap} to analyze debates allows for consideration of opposing arguments. Part of that consideration
+includes identifying and eliminating fallacious arguments. In heated partisan debates, fallacious arguments spread
+easily. \myfig{fig-arg-fallacies} shows those most commonly used in the encryption and \ac{EA} debate.
 
@@
 
-Let us begin with bad arguments used to attack \ac{EA}. The first is the most common: ``\ac{EA} is a bad idea because
-backdoors are insecure.'' It appears on the map as a straw man argument due to the way the term ``backdoor'' is used.
-``Backdoor'' is essentially technical shorthand for ``insecure hack that should never be used in production''; when it
-is used as a basket term for all \ac{EA} proposals, the audience is primed to think them all stupid and helplessly
-dangerous. Thus, the arguer is presenting their opponent's request for some form of \ac{EA} in the worst possible light.
+The map begins with fallacious arguments used to attack \ac{EA}. The first is the most common: ``\ac{EA} is a bad idea
+because backdoors are insecure.'' It appears on the map as a straw man argument due to the way the term ``backdoor'' is
+used. ``Backdoor'' is essentially technical shorthand for ``insecure hack that should never be used in production'';
+when it is used as a basket term for all \ac{EA} proposals, the audience is primed to consider them all ill-conceived
+and hopelessly dangerous. Thus, the arguer is presenting their opponent's argument for any degree of \ac{EA} in the
+weakest possible fashion.
 
-Another common argument is that implementing \ac{EA}---or even researching it---moves us inexorably down a slippery
-slope, where government demands will never be satisfied. This argument can be used even if one thinks judicious
-hypothetical use of \ac{EA} tools would be a good thing. Going too far and too fast is certainly a danger, but it is
-both cynical and fallacious to declare the permanent destruction of privacy as the inevitable end of \ac{EA} research.
+Another common argument is that implementing \ac{EA}---or even researching it---inexorably moves tech policy down a
+slippery slope, in which government demands will never be satisfied. This argument can be used to prevent research from
+those who think judicious hypothetical use of \ac{EA} tools could be positive. Going too far and too fast is certainly
+dangerous, but it is both cynical and misleading to declare the permanent destruction of privacy as the inevitable end
+of \ac{EA} research.
 
-It is highly tempting to declare government calls for \ac{EA} disingenuous due to hypocritical behavior in their
-treatment of the issues. For example, if combatting \ac{CSAM} is a government priority, how come Congress hasn't funded
-it and the Justice Department hasn't prioritized it \cite{keller_internet_2019}? This question, and others like it,
-demands an answer, but it ultimately does not impact the validity (or invalidity) of pro-\ac{EA} arguments. The absence
-of government funding doesn't mean technical anti-\ac{CSAM} measures are unnecessary.
+It is tempting to declare government requests for \ac{EA} disingenuous due to their hypocritical behavior towards some
+of these issues. For example, if combatting \ac{CSAM} is a government priority, why has Congress underfunded the laws
+they've passed, and why has the Justice Department understaffed their task force for it \cite{keller_internet_2019}?
+These questions demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The
+absence of government funding does not mean technical anti-\ac{CSAM} measures are unnecessary.
 
-The final two anti-\ac{EA} fallacies are examples of false dichotomies, presenting the situation as if there are only
-two choices. \Acp{wicked-prob} have an unlimited set of possible responses, so reducing it to two (and often, presenting
-one as a straw man) is misleading. One false dichotomy is that the only acceptable solutions are perfect \ac{EA} or no
-\ac{EA}. This leaves no room for risk-based approaches, the typical security strategy outside the realm of cryptography.
-Another false dichotomy is that regulators must either allow strong cryptography or we will live in a \ac{masssurv}
-dystopia. This one carries hints of ``slippery slope'' and appeal to emotion as well, framed as a choice between only
-two futures. Framing the situation in all or nothing terms harms the chances for successful collaborative debate.
+The final two anti-\ac{EA} fallacies are examples of false dichotomies; they present the situation as if there are only
+two outcomes. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading. One
+false dichotomy is that the only acceptable solutions are either perfect \ac{EA} or no \ac{EA}. This leaves no room for
+risk-based approaches, the typical security strategy outside the realm of cryptography. Another false dichotomy is that
+regulators must either allow strong cryptography or we will live in a \ac{masssurv} dystopia. This one carries hints of
+``slippery slope'' and appeal to emotion as well, framed as a choice between only two futures. Framing the situation in
+all or nothing terms reduces the chance of a successful collaborative debate.
 
-Fallacious arguments used to support \ac{EA} rely heavily on appeal to emotion. Politicians and law enforcement agents
-commonly appeal to pity for child abuse victims or fear of terrorists, drug dealers, and kidnappers to make their point.
-Public safety is a core value in the debate and the technical community needs to and does take these concerns seriously
-\cite{schneier_scaring_2019}. These appeals become fallacious when they are used to manipulate the audience into
-supporting measures incommensurate with the problems, or to manipulate the audience into thinking that tech somehow
-actually does not care about these issues.
+Some arguments used to support \ac{EA} rely on appeal to emotion. Politicians and law enforcement agents often appeal to
+pity for child abuse victims or fear of terrorists, drug dealers, and kidnappers to make their point. Public safety is a
+central issue in the debate. The technical community does take these concerns seriously \cite{schneier_scaring_2019}.
+These appeals become fallacious when they are used to manipulate the audience into supporting solutions incommensurate
+with the problems or to manipulate the audience into believing that the technical community does not care about these
+issues.
 
 Another argument used to subvert anti-EA arguments from experts goes along the lines of, since we put a man on the moon,
-we can surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The problem of landing an
-astronaut on the moon is different from the problem of encryption and \ac{EA} in every way except that it was hard. The
-Apollo program solved mostly tame problems, and it had enormous government backing. Resolving the encryption debate has
-neither of these factors. Many similar technological-progress based arguments suffer the same faults.
+we can surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The only similarity between the
+challenge of landing an astronaut on the moon and the challenge of building a cryptosystem with secure EA is that they
+are difficult. Once the political commitment was made, the Apollo program solved mostly tame problems and had enormous
+government backing. Resolving the encryption debate has neither of these benefits. Many similar arguments comparing
+\ac{EA} to other feats of technological progress suffer the same faults.
 
-Lastly, government officials often pose \ac{EA} as a problem that security experts simply have not researched enough. If
-they just ``nerd harder,'' they can find a solution \cite{schneier_2019}. This misses the point of research done up to
-this point. Encryption and \ac{EA} is a \ac{wicked-prob} for which tame, technical-only solutions will not work
-\cite{rozenshtein_wicked_2018}. Even on the technical side, experts have stated time and again that depending how you
-look at it, the technical problem is solved---\ac{EA} is easy from a purely cryptographic point of view. The hard part
-is the security of the interfaces with the human users and administers of the system \cite{abelson_risks_1997}
-\cite{abelson_2015}. This may be an area of anemic academic research, but to frame that as the primary roadblock to
-progress utterly misses the point.
+Lastly, government officials often portray \ac{EA} as a problem that security experts simply have not researched enough.
+If experts just ``nerd harder,'' as some put it, they can find a solution \cite{schneier_2019}. This misses the point of
+research done up to this point. Encryption and \ac{EA} is a \ac{wicked-prob} for which tame, technical-only solutions
+will not work \cite{rozenshtein_wicked_2018}. Even on the technical side, experts have stated time and again that,
+depending how you look at it, the technical problem is solved---\ac{EA} is easy from a purely cryptographic point of
+view. The hard part is the security of the interfaces with the human users and administers of the system
+\cite{abelson_risks_1997} \cite{abelson_2015}. This may not be an area of strong academic research, but to frame that
+as the primary roadblock to progress utterly misses the point.
 
@@
 Moving on from problem framing and meritless arguments, we next turn to actions that can be taken in response to the
-situation. \myfig{fig-arg-measures} breaks the actions down into the current capabilities, legal measures, and \ac{EA}.
-Current capabilities can be implemented today, and include maintaining the status quo and increasing investment in
-current programs. Legal measures require some change in legal or political direction, but do not require fundamentally
-new technical research. These include compelling passwords, which requires legal clarification, and sanctioning lawful
-hacking, which requires a strategic pivot and an oversight framework. (These non-\ac{EA} approaches were first
-introduced in \mysec{sec-alt-approaches}.) At this level, \ac{EA} is analyzed in general. Specific EA approaches are
-addressed in the next map.
+situation. \myfig{fig-arg-measures} divides potential solutions into the categories of current capabilities, legal
+measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the status quo and
+increasing investment in current programs. Legal measures require change the text or application of the law, but do not
+require fundamentally new technical capabilities. These include compelling passwords, which requires legal
+clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
+non-\ac{EA} approaches were first introduced in \mysec{sec-alt-approaches}.) At this level, \ac{EA} is analyzed in
+general. Specific EA approaches are addressed in the next map.
 
@@
 \cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
-of security, safety, privacy, and trust. The five metrics used to analyze response measures are minimization of
-increased security risk, utility to law enforcement, protection of civil liberties---particularly important as after the
-analysis of the ``Golden Age for Surveillance'' argument---and economic impact. Economic impact is not as much of a
-primary concern as the others, but it does help determine practical viability of a proposal.
+of security, safety, privacy, and trust. The five metrics used to analyze response measures are (1) minimization of
+increased security risk, (2) utility to law enforcement, (3) protection of civil liberties---particularly important as
+after the analysis of the ``Golden Age for Surveillance'' argument---(4) transparency, and (5) economic impact. Economic
+impact is not as central an issue as the others, but it does help determine practical viability of a proposal.
 
-Beginning with current capabilities, the first choice is to do nothing. One may come to the conclusion that the current
-state of affairs actually does represent an acceptable balance of interests. If technology and government agree on this
-approach, then \ac{the-cw2} can end much like the first. However, legislative action is a risk to be mitigated in
-itself, and if the technical community ``decides'' on this approach and government continues to reject it, we could be
-forced in an undesirable direction.
+Beginning with current capabilities, the first choice is to do nothing. Some may conclude that the current state of
+affairs actually does represent an acceptable balance of interests. If technology and government agree on this approach,
+then \ac{the-cw2} can end much like the first. However, legislative action itself is a risk to be mitigated. If the
+technical community ``decides'' on this approach and government continues to reject it, we could be forced in an
+undesirable direction.
 
-We could also decide to increase investment in current technologies and investigative techniques. As already shown,
-encryption is not the greatest barrier in law enforcement's use of digital evidence \cite{carter_2018}. Even when
-encryption is involved, traditional sleuthing can often achieve access. Security expert Bruce Schneier and cyberlaw
-expert Orin Kerr have compiled a list of ``encryption workarounds'' that do not rely on \ac{EA}, including finding the
-key, guessing the key, accessing plaintext while the device is in use, and finding another copy of the plaintext
-\cite{kerr_encryption_2017}. Though probabilistic and time consuming in comparison to \ac{EA}, none of these methods
-require new legal or technical capability. These methods are effective \cite{greenberg_2018}, and paired with investment
-in current digital evidence capabilities and anti-\ac{CSAM} (for example) task forces, they represent a safe response.
+We could also decide to increase investment in current technologies and investigative techniques. As previously
+discussed, encryption is not the most significant barrier in law enforcement's use of digital evidence
+\cite{carter_2018}. Even when encryption is involved, traditional sleuthing can often lead to access. Security expert
+Bruce Schneier and cyberlaw expert Orin Kerr have compiled a list of ``encryption workarounds'' that do not rely on
+\ac{EA}. These include finding the key, guessing the key, accessing plaintext while the device is in use, and finding
+another copy of the plaintext \cite{kerr_encryption_2017}. Though probabilistic and time-consuming in comparison to
+\ac{EA}, these methods do not require new legal or technical capability. These methods are effective
+\cite{greenberg_2018}; investment in current digital evidence capabilities and anti-\ac{CSAM} task forces, they are also
+safe.
 
-Compelled password disclosure is a proposal to simply require defendants to hand over passwords and \acp{PIN} to their
-devices. The legality of this approach is working its way through the courts \cite{bittenbender_2019} \cite{sobel_2019},
-as it rubs against the Fifth Amendment right against self-incrimination. Even so, its strongest argument in favor is
-that it is cheap. Schneier and Kerr point out that this approach requires the password holder to be known and available
-to investigators, the prosecution has to overcome several legal hurdles even if the basic approach is ruled
-Constitutionally legal, and even then the defendant may choose to withhold the password and be held in contempt rather
-than be charged for crime evidenced by the plaintext \cite{kerr_encryption_2017}.
+Compelled password disclosure is a potential solution that simply requires defendants to surrender passwords and
+\acp{PIN} to their devices. Questions regarding the legality of this approach have been working their way through the
+courts \cite{bittenbender_2019} \cite{sobel_2019}, as may be interpreted to conflict with the Fifth Amendment right
+against self-incrimination. Even so, affordability is the strongest argument in its favor. Schneier and Kerr point out
+that this approach requires the password-holder to be available to investigators, the prosecution has to overcome
+several legal hurdles even if the basic approach is ruled Constitutionally legal, and even then the defendant may choose
+to withhold the password and be held in contempt rather than be charged for crime evidenced by the plaintext
+\cite{kerr_encryption_2017}.
 
-\Ac{lawful-hacking} is the primary alternative argued for besides investment or \ac{EA}, though it is universally seen
-as a sub-optimal compromise \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018}
-\cite{kerr_encryption_2017} \cite{soesanto_2018}. Its advantages: it is effective against those for whom \ac{EA} would
-not necessarily work, and, well, it's a technical route that isn't \ac{EA} itself. \Ac{lawful-hacking} has the advantage
-over \ac{EA} at being effective against organized crime and terror groups because while they would inevitably use
-encryption tools to evade any \ac{EA} scheme, their operational security would fail at some point, letting law
-enforcement or intelligence agencies into their networks and devices. This approach is already in use to some extent
-\cite{cox_2020}, most famously to bring down cartel kingpin El Chapo \cite{feuer_chapo_2019}.
+\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA}, though it is universally seen as a
+sub-optimal compromise \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018}
+\cite{kerr_encryption_2017} \cite{soesanto_2018}. One explanation for \ac{lawful-hacking}'s support is that it's a
+technical route that isn't \ac{EA} itself. Another is its effectiveness against those for whom \ac{EA} would not
+necessarily work. \Ac{lawful-hacking} has the advantage over \ac{EA} at being effective against organized crime and
+terror groups because while they would inevitably use encryption tools to evade any \ac{EA} scheme, their operational
+security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
+This approach is already in use to some extent \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
+\cite{feuer_chapo_2019}.
 
@@
 \section{Zooming in on EA}
+\label{sec-ea-types}
 

### 2020-12-12 Thesis-Chapter4-Tools.tex
Editorial improvements

@@
 
-Argument mapping tools originated in 1970 with text-based \acp{IBIS} designed to break down \acp{wicked-prob} and
-document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \acp{IBIS} has waxed and
-waned. Some software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
-\cite{dutoit_hypermedia_2006}, which were designed to produce graphical argument representations. Argdown is a recent
-tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
+% TODO: can probably cut more
+Argument mapping tools originated in 1970 with text-based \acp{IBIS} \cite{kunz_issues_1970}. Interest in \acp{IBIS} has
+waxed and waned. Some software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and
+Compendium \cite{dutoit_hypermedia_2006}, which were designed to produce graphical argument representations. Argdown is
+a recent tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
 \cite{voigt_argdown_2018}. Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA}
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
     \centering\CaptionFontSize

### 2020-12-12 Thesis-Chapter5-Arguments.tex
Editorial improvements

@@
 debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
-as tools in the collaborative pursuit of the problem definition and solution. One can approximate the debate based on
-research, but it is not a replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively
-portray the strength of an argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments
-were presented against one indisputable argument, the side with more arguments would, regardless of their integrity,
-appear stronger. With those qualifications in mind, let us begin by examining the factors at the center of the conflict.
+as tools in the collaborative pursuit of the problem definition and solution. It is not a replacement for debate arising
+from live discussion. Second, \acp{argmap} may deceptively portray the strength of an argument. Argument nodes do not
+indicate their strength or validity. If ten unsound arguments were presented against one indisputable argument, the side
+with more arguments would, regardless of their integrity, appear stronger. With those qualifications in mind, let us
+begin by examining the factors at the center of the conflict.
 
@@
 
-\Acl{LE}, argues that encryption handcuffs its investigational capacity. The \ac{FBI} brands strong encryption as
+\Acl{LE} argues that encryption handcuffs its investigational capacity. The \ac{FBI} brands strong encryption as
 ``warrant-proof'' and states that ``the government often cannot obtain the electronic evidence necessary to investigate
@@
 \cite{comey_2014} (though he did not coin the phrase \cite{swire_encryption_2011}). The argument is presented in
-\myfig{fig-arg-going-dark} including several critiques \cite{rogaway_moral_2015}.
+\myfig{fig-arg-going-dark}.
 
@@
 According to the "Going Dark" argument, privacy and safety are both desirable goods, but they conflict with one another;
-therefore, they must be balanced. Individuals must sacrifice some of their personal good of privacy to in honor of the
-public good of safety. Widespread encryption tips the scales too far towards privacy, upsetting the balance we had
+therefore, they must be balanced. Individuals must sacrifice some of their personal good of privacy to allow for the
+public good of safety. Widespread encryption tips the scales too far towards privacy, upsetting the balance that existed
 before. \ac{EA} restores the balance and enables law enforcement agencies to fulfill their duty to protect the public.
@@
 
-Most importantly, the strength of the ``Going Dark'' argument rests on law enforcement investigations failing due to
+Most importantly, the ``Going Dark'' argument relies on the premise that law enforcement investigations fail due to
 encryption. As described in \mysec{sed-arg-contrib}, there is no conclusive data regarding the extent to which
@@
 Citing lack of evidence that encryption is stymieing cases only gets you so far, however. In most cases, it is
-impossible to know what would have happened without the interference of strong encryption. Inability to perform
-controlled tests is a characteristic of \ac{wicked-prob}. When controlled settings are unachievable, another way to
-gather data is by observing naturally-occurring conditions where everything but the variable of interest is the same.
-These conditions exist for encryption in messaging application ecosystem. Messaging applications WhatsApp, Facebook
-Messenger, and Telegram all have large user bases have approximate feature parity. Among these services, only Facebook
-Messenger does not employ \ac{E2EE}. What effect does that have on investigation? Globally, Facebook Messenger has
-approximately 20\% market share as measured by active monthly users \cite{statista_2020} but accounts for 65\% of
-\ac{CSAM} reports \cite{keller_internet_2019}.
+impossible to know what would have happened without the interference of strong encryption. Inability to gather data from
+controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable, another way to
+gather data is by observing naturally-occurring conditions in which everything but the variable of interest remain are
+constant. These conditions exist for encryption in messaging application ecosystem. Messaging applications WhatsApp,
+Facebook Messenger, and Telegram all have large user bases and similar features. Among these services, only Facebook
+Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market share as measured by
+active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports \cite{keller_internet_2019}. This
+indicates that encryption does indeed affect investigations.
 
@@
 free-speech and threatens liberal democracy.'' He argues that surveillance is fundamentally a tool of power, and for
-this reason it should be resisted. Though power from the people created through representative political process is
-valid, clear abuses have undercut reasonable confidence in these institutions.
+this reason it should be resisted. Though power from the people, created through representative political process, is
+valid, clear abuses have undercut confidence in these institutions.
 
@@
 in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement but
-thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age'' argument. \ac{EA} proposals that strive for these
-properties will be explored in \mysec{sec-ea-types}.
+thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age'' argument. \ac{EA} proposals that strive for
+these properties will be explored in \mysec{sec-ea-types}.
 
@@
 Another common argument is that implementing \ac{EA}---or even researching it---inexorably moves tech policy down a
-slippery slope, in which government demands will never be satisfied. This argument can be used to prevent research from
-those who think judicious hypothetical use of \ac{EA} tools could be positive. Going too far and too fast is certainly
-dangerous, but it is both cynical and misleading to declare the permanent destruction of privacy as the inevitable end
-of \ac{EA} research.
+slippery slope, in which government demands will never be satisfied. This argument can be used to prevent research by
+those who think judicious use of \ac{EA} tools could be positive. Going too far and too fast is certainly dangerous, but
+it is both cynical and misleading to declare the permanent destruction of privacy as the inevitable end of \ac{EA}
+research.
 
@@
 of these issues. For example, if combatting \ac{CSAM} is a government priority, why has Congress underfunded the laws
-they've passed, and why has the Justice Department understaffed their task force for it \cite{keller_internet_2019}?
-These questions demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The
-absence of government funding does not mean technical anti-\ac{CSAM} measures are unnecessary.
+it's passed, and why has the Justice Department understaffed their its task force \cite{keller_internet_2019}? These
+questions demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The absence of
+government funding does not mean technical anti-\ac{CSAM} measures are unnecessary.
 
 The final two anti-\ac{EA} fallacies are examples of false dichotomies; they present the situation as if there are only
-two outcomes. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading. One
+two options. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading. One
 false dichotomy is that the only acceptable solutions are either perfect \ac{EA} or no \ac{EA}. This leaves no room for
@@
 are difficult. Once the political commitment was made, the Apollo program solved mostly tame problems and had enormous
-government backing. Resolving the encryption debate has neither of these benefits. Many similar arguments comparing
+government backing. Resolving the encryption debate has neither of these advantages. Many similar arguments comparing
 \ac{EA} to other feats of technological progress suffer the same faults.
@@
 measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the status quo and
-increasing investment in current programs. Legal measures require change the text or application of the law, but do not
-require fundamentally new technical capabilities. These include compelling passwords, which requires legal
+increasing investment in current programs. Legal measures require a change in the text or application of the law, but do
+not require fundamentally new technical capabilities. These include compelling passwords, which requires legal
 clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
@@
 
-Each measure is considered with reference to five metrics. These metrics represent a generalization of metrics from
-several sources, including two separate diverse committees gathered to analyze the encryption debate
+Each measure is considered according to five metrics. These metrics represent a generalization of metrics from several
+sources, including two separate diverse committees gathered to analyze the encryption debate
 \cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
@@
 \acp{PIN} to their devices. Questions regarding the legality of this approach have been working their way through the
-courts \cite{bittenbender_2019} \cite{sobel_2019}, as may be interpreted to conflict with the Fifth Amendment right
+courts \cite{bittenbender_2019} \cite{sobel_2019} because of the apparent conflict with the Fifth Amendment right
 against self-incrimination. Even so, affordability is the strongest argument in its favor. Schneier and Kerr point out
@@
 
-\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA}, though it is universally seen as a
-sub-optimal compromise \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018}
-\cite{kerr_encryption_2017} \cite{soesanto_2018}. One explanation for \ac{lawful-hacking}'s support is that it's a
-technical route that isn't \ac{EA} itself. Another is its effectiveness against those for whom \ac{EA} would not
-necessarily work. \Ac{lawful-hacking} has the advantage over \ac{EA} at being effective against organized crime and
-terror groups because while they would inevitably use encryption tools to evade any \ac{EA} scheme, their operational
+\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA} \cite{bellovin_lawful_2013}
+\cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018} \cite{kerr_encryption_2017} \cite{soesanto_2018}.
+\Ac{lawful-hacking}'s support may be due in part to the fact that it is a technical approach that isn't \ac{EA} itself.
+However, its effectiveness against criminals for whom \ac{EA} would not work is its obvious benefit. This is because,
+while organized crime and terror groups would still use encryption tools to evade an \ac{EA} scheme, their operational
 security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
-This approach is already in use to some extent \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
+This approach is already in use \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
 \cite{feuer_chapo_2019}.
 
-The real change that this policy would take would be a sanctioning and formalization of this strategy. One downside is
-that it would be of little use to ill-resourced departments, nullifying its usefulness for more commonplace crimes. Its
-major downside is the altered relation is would create between law enforcement and tech companies. Instead of
-cooperation, they are suddenly thrust into competition, with law enforcement incentivized to hoard rather than disclose
-vulnerabilities it discovers. Prominent proposals suggest mandating prompt disclosure of discovered vulnerabilities
-\cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016}. Theoretically, the supply of vulnerabilities is very large,
-and new ones could be found at a rate that maintains law enforcement access at constant levels while improving security
-as a positive side effect. The government would source these vulnerabilities from a combination of the public domain,
-the commercial exploit market, and a central ``Vulnerability Lab'' \cite{bellovin_lawful_2013}.
+The real change that \ac{lawful-hacking} would take would be a sanctioning and formalization of this strategy.
+% TODO: tie to paragraph above?
 
-The availability and cost of exploits ebbs and flows. On one hand, January 2019 sported a \$2 million price on the
-exploit market for an Apple iOS jailbreak bug \cite{goodin_zeroday_2019}. On the other hand, the October 2020 report
-detailing law enforcement's mobile device cracking capabilities estimated the per-device cost to be only \$2000 a unit
-\cite{koepke_2020}. As the exploit arms race continues, exploitability will be inconsistent. In the end, law enforcement
-would face periods where it struggles for access, and would need to develop costly exploits in the lab. Mandatory
-vulnerability disclosure is key to a \ac{lawful-hacking} proposal; however, based on the \ac{FBI}'s historical abuses of
-power \cite{shamsi_2011} and Congress's distaste for funding even initiatives it has passed \cite{keller_internet_2019}
-or providing strong oversight \cite{johnson_congressional_2004}, one cannot imagine that the government would actually
-give up expensive vulnerabilities easily when they are the pillar of their access strategy.
+Despite its strength's \ac{lawful-hacking} is a sub-optimal compromise. Because it requires time and specialized skills,
+it is unuseful for small departments and commonplace crimes. It would also alter the relationship between law
+enforcement and tech companies. Law enforcement agencies would be forced into competition with tech companies if they
+were incentivized to conceal the vulnerabilities they discover. Prominent proposals suggest mandating disclosure of
+discovered vulnerabilities to mitigate this issue \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016}.
+Theoretically, the supply of vulnerabilities is very large, so even though the disclosed vulnerabilities would be
+disclosed and patched, new ones could be found at a rate that maintains law enforcement access. Government could source
+these vulnerabilities from a combination of the public domain, the commercial exploit market, and a central
+``Vulnerability Lab'' \cite{bellovin_lawful_2013}.
 
-The final proposal is \ac{EA} itself. Its advantage, generally speaking, is that it would provide reliable access to law
-enforcement in a reliable manner. It could serve as a compliment to \ac{lawful-hacking}, blunting the incentive to hoard
-vulnerabilities because it would be unnecessary for \ac{EA} compliant systems, which would presumably include mainstream
+The availability and cost of exploits fluctuates. For example, in January 2019, an Apple iOS jailbreak bug was valued at
+\$2 million on the exploit market \cite{goodin_zeroday_2019}. In contrast, a 2020 report revealed that cracking mobile
+devices cost law enforcement a mere \$2000 per unit \cite{koepke_2020}. As the exploit arms race continues,
+exploitability will be inconsistent. In the end, law enforcement would face periods where it struggles for access, and
+would need to develop costly exploits in the lab. Mandatory vulnerability disclosure is necessary to avoid competition
+between law enforcement and tech companies. However, based on the \ac{FBI}'s historical abuses of power
+\cite{shamsi_2011} and Congress's reluctance to fund initiatives it has passed \cite{keller_internet_2019} and failure
+to provide strong oversight \cite{johnson_congressional_2004}, it seems unlikely that government would readily forfeit
+expensive vulnerabilities.
+
+The final proposal is \ac{EA} itself. \ac{EA} would provide a standardized method for law enforcement to reliably access
+\ac{plaintext}. It could serve as a compliment to \ac{lawful-hacking}by eliminating the incentive to conceal
+vulnerabilities because it would be unnecessary for \ac{EA}-compliant systems, which would presumably include mainstream
 technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
-and terror groups. While impact on civil liberties and \ac{masssurv} depends on implementation, any proposal would be
-costly to introduce and threaten U.S. product competitiveness due to perceived insecurity. Indeed, risk to security is
-the largest weakness of \ac{EA}. Any proposal has to overcome significant challenges to provide acceptable levels of
-risk. The next map analyzes this more closely.
+and terror groups. While \ac{EA}'s impact on civil liberties and \ac{masssurv} would depend on its implementation, any
+proposal would be costly to introduce and would threaten U.S. product competitiveness due to perceived insecurity. Risk
+to security is the largest weakness of \ac{EA}. Any proposal must overcome significant challenges to minimize risk. The
+next map analyzes this more closely.
 
@@
 
-\myfig{fig-arg-classes} lists several classes of \ac{EA} and arguments in their favor and disfavor. Classes applicable
-to both \ac{DIM} and \ac{DAR} include using weak cryptography, trusted-party key escrow, and distributed key escrow.
-Using cryptographic puzzles and adding a ghost user apply to the case of \ac{DIM}, and device key escrow applies to
-\ac{DAR}.
+\myfig{fig-arg-classes} lists several classes of \ac{EA} and arguments in their favor and disfavor. The \ac{EA} classes
+applicable to both \ac{DIM} and \ac{DAR} include weak cryptography, trusted-party key escrow, and distributed key
+escrow. Using cryptographic puzzles and adding a ghost user apply to the case of \ac{DIM}, and device key escrow applies
+to \ac{DAR}.
 
@@
 
-It should be noted that the \ac{argmap} format works best when displaying the debate surrounding a single conclusion or
-proposal. That leaves room for points and counterpoints to be elaborated in detail, as opposed to creating a tangled web
-of un-elaborated connections. For the sake of brevity, this thesis considers \ac{EA}'s alternatives and variations at a
-high level. For an excellent, thorough discussion of \ac{EA}'s alternatives and variations, see the National Academies
-of Sciences, Engineering, and Medicine's 2018 report titled \ii{Decrypting the Encryption Debate: A Framework for
-  Decision Makers} \cite{committee_decrypting_2018}.
+% TODO: move this up? Applies to previous map as well.
+It should be noted that the \ac{argmap} format works best when it displays the debate surrounding just one conclusion or
+proposal. This allows adequate room for elaboration on points and counterpoints. When a map includes multiple
+conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
+of brevity, this thesis compares several \ac{EA} variations in a single map. For an thorough discussion of \ac{EA}'s
+alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report titled
+\ii{Decrypting the Encryption Debate: A Framework for Decision Makers} \cite{committee_decrypting_2018}.
 
 The first \ac{EA} class is to simply use weak cryptographic ciphers or short encryption keys. This was the approach
-taken before \ac{the-cw1}, when strong cryptography was not highly available and subject to export controls. No party
-seriously supports this approach today, as it is irreconcilably contradictory to strong security on many levels.
+taken before \ac{the-cw1}, when strong cryptography was not readily available and subjected to export controls. No party
+seriously supports this approach today, as it contradicts security on many levels.
 
-Trusted-party key escrow relies on the an entity or small group of entities to store a key or key recovery information.
-There are a wide variety of implementations \cite{denning_taxonomy_1996}, but by definition this class of \ac{EA}
-carries a crushing weakness: the centralization of extremely sensitive information. Proponents argue that software
-companies' protection of code-signing keys in hardware security modules is evidence that they can protect such sensitive
-data when needed \cite{ozzie_2018}, but opponents point out that such keys are accessed quite rarely used (especially
-when compared to the likely frequency of \ac{EA} requests) and yet they indeed still leak \cite{green_2018}. Once the
-escrowed keys are stolen, as they inevitably would be, every device or message encrypted with them would be at risk.
-This is an example of a case when the \ac{argmap} format doesn't do an argument justice---the argument against any
-centralized access mechanism is extremely powerful.
+Trusted-party key escrow relies on an entity or a small group of entities to store a key or key recovery information.
+There are a wide variety of implementations \cite{denning_taxonomy_1996}. This class of \ac{EA} has a crushing weakness:
+it centralizes extremely sensitive information. Some argue software companies are capable of protecting such sensitive
+data, citing their current practice of storing code-signing keys in hardware security modules \cite{ozzie_2018}.
+However, others point out that such keys are quite rarely used (especially when compared to the likely frequency of
+\ac{EA} requests), and yet they still leak \cite{green_2018}. Once the escrowed keys were stolen, as they inevitably
+would be, every device or message encrypted with them would be at risk. This is an example of a case when the
+\ac{argmap} format doesn't do an argument justice---the argument against any centralized access mechanism is extremely
+powerful.
 
 Distributed key escrow differs from trusted-party in that it takes strong steps to remove centralized trust. This is
-achieved by using distributed systems to make key extraction difficult and inherently public \cite{phan_key_2017}
-\cite{servan_schreiber_jje_2020}. If the distributed protocols hold up, one advantage of distribution is that this
-approach is that it is more likely to fail by making data unrecoverable than by making unintended data vulnerable,
-though more research into this area is needed. The downside of these systems is their complexity, which could make it
-costly and hard to update.
+achieved by using distributed systems to make key extraction difficult and publicly observable \cite{phan_key_2017}
+\cite{servan_schreiber_jje_2020}. One advantage of distribution is that this approach is that it is more likely to fail
+by making data unrecoverable than by making unintended data vulnerable, though more research into this area is needed.
+Unfortunately, these systems would likely be costly to deploy and difficult to update due to their complexity.
 

### 2020-12-12 Thesis-Chapter7-PathsAhead.tex
Editorial improvements

--- 2020-12-08 Thesis-Chapter6-PathsAhead.tex
@@
 
-The interests of security, safety, privacy, and trust are converging on the issue of encryption and \acl{EA} and
-creating a \ac{wicked-prob}. Information security is increasingly important as more data, business, and infrastructure
-goes online. The technical community is struggling with security; defending against the average attacker is hard, and
-against the best attackers is nearly impossible, even for the most motivated and well-resourced organizations.
-\Ac{cryptography} plays a fundamental role in security, yet its simultaneous role in enabling unprecedented privacy
-means it has controversial side effects.
+The interests of security, safety, privacy, and trust converge on the issue of encryption and \acl{EA}, creating a
+\ac{wicked-prob}. The importance of information security is increasing as more data, business, and infrastructure goes
+online. The technical community is struggles with security. Even for the most well-equipped organizations, thwarting the
+average attacker is difficult, and the best attacker, impossible. \Ac{cryptography} plays a fundamental role in
+security, yet its simultaneous role in enabling unprecedented privacy means it has controversial side effects.
 
 One of these side effects is encryption-enabled privacy's impact on public safety. Despite a shortage of quantitative
-analyses showing its impact on crime---something that would be impossible to fully quantify, after all---encryption
-decidedly withholds evidence of wrongdoing from service providers and law enforcement, enabling truly harmful activity
-to go unchecked, and decreasing public safety. This a legitimate concern and the primary motivation for \ac{EA}.
+analyses showing its impact on crime, encryption certainly does allow evidence of wrongdoing to be hidden from service
+providers and law enforcement. This enables harmful activity to go unchecked and decreases public safety, which is a
+legitimate concern and the primary motivation for \ac{EA}.
 
 However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
-stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling threats
-ranging from identity theft, to inappropriate data mining, to \ac{masssurv}. To the extent that encryption mitigates
-these, it increases public safety.
+stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling identity
+theft, inappropriate data mining, \ac{masssurv}, and other threats. To the extent that encryption mitigates these, it
+increases public safety.
 
-Finally, choosing how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
+Finally, deciding how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
 considerable questions of trust. Law enforcement agencies, intelligence agencies, technology makers, individual users,
@@
 
-% TODO: one paragraph here^^ or four?
-
-This complex mix of issues is what makes encryption policy a \ac{wicked-prob}. To merely formulate the problem requires
-explanation upon explanation and research upon research. Any formulation is concomitant with a solution, but the outcome
-of a solution cannot be accurately forecasted or evaluated in isolation. The underlying conditions are constantly in
-flux, affected by external forces as well as every policy action and inaction.
+This complex mix of issues makes encryption policy a \ac{wicked-prob}. Even just the problem requires explanation upon
+explanation and research upon research. Any formulation is concomitant with a solution, but the outcome of a solution
+cannot be accurately evaluated in isolation. The underlying conditions constantly change due to external forces and
+policy actions (or lack thereof).
 
@@
 Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
-the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption poses a present danger to public safety
-in certain respects. The bogeymen of terrorists and criminals using encryption to evade the law are real---in the case
-of those peddling and viewing \acl{CSAM}, shockingly so \cite{keller_internet_2019}. However, the argument that
-\ac{encryption} is creating circumstances that are to blame for \acl{LE} failure is not currently strong enough to
-justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor coordination and training \cite{carter_2018}; for
-\ac{DAR}, commercial \ac{lawful-hacking} tools currently have the upper hand on mobile phone manufacturers, meaning
-\acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Additionally, strong \ac{encryption}
-plays a key role in protecting security and privacy, public goods that other technological developments often fail to
-provide.
+the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption does pose certain threats to public
+safety. Terrorists and criminals can and do use encryption to evade the law; this has allowed for shocking levels of
+\acf{CSAM} to be peddled \cite{keller_internet_2019}. However, the argument that \ac{encryption} is causing \acl{LE}
+failure is not currently strong enough to justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor
+coordination and training \cite{carter_2018}; for \ac{DAR}, commercial \ac{lawful-hacking} tools currently have the
+upper hand on mobile phone manufacturers, meaning \acl{LE} can get into almost any mobile device they acquire
+\cite{koepke_2020}. Additionally, strong \ac{encryption} is necessary to protect security and privacy---two public goods
+that other technological developments often fail to provide.
 
 Though the pro-\ac{EA} argument is ultimately not strong enough to justify an \ac{EA} mandate, incorporating it to the
-technologist's point of view teaches important lessons. One is that wholistic cryptosystem threat models need to include
-two non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and
-this power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating
-steps. Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption;
-lawmakers represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too
-dangerous. The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.
+technologist's point of view teaches important lessons. One is that holistic cryptosystem threat models must include two
+non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and this
+power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating steps.
+Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption; lawmakers
+represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too dangerous.
+The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.
 

### 2020-12-13 Thesis-Abstract.tex
Editorial improvements

@@
-Public debate has resumed on the topic of exceptional access (EA), which is alternative means of decryption intended for
-law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of the 1990s
-``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing circumstances. The
-imbalance between privacy, access, and security in the context of constantly changing society and technology is a wicked
-problem that has and will continue to evade a permanent solution. As policymakers consider next steps, it is necessary
-that the technical community to remain engaged. Although the introduction of EA would inevitably add risk, the quality
-of the technical and regulatory approach can make a substantial difference. Furthermore, if one considers hard-line
-legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed EA may reduce risk
-overall.
+Public debate has resumed on the topic of exceptional access (EA), which refers to alternative means of decryption
+intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of
+the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
+circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
+technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
+steps, it is necessary that the technical community remain engaged. Although any EA framework would increase risk, the
+magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
+considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
+EA may reduce risk overall.
 

### 2020-12-13 Thesis-Chapter5-Arguments.tex
Editorial improvements

@@
 
-Despite its strength's \ac{lawful-hacking} is a sub-optimal compromise. Because it requires time and specialized skills,
-it is unuseful for small departments and commonplace crimes. It would also alter the relationship between law
+Despite its strengths \ac{lawful-hacking} is a sub-optimal compromise. Because it requires time and specialized skills,
+it is not useful for small departments and commonplace crimes. It would also alter the relationship between law
 enforcement and tech companies. Law enforcement agencies would be forced into competition with tech companies if they
@@
 discovered vulnerabilities to mitigate this issue \cite{bellovin_lawful_2013} \cite{hennessey_lawful_2016}.
-Theoretically, the supply of vulnerabilities is very large, so even though the disclosed vulnerabilities would be
-disclosed and patched, new ones could be found at a rate that maintains law enforcement access. Government could source
-these vulnerabilities from a combination of the public domain, the commercial exploit market, and a central
-``Vulnerability Lab'' \cite{bellovin_lawful_2013}.
+Theoretically, there are enough vulnerabilities that, even though these discovered vulnerabilities would be disclosed and
+patched, new ones could be found quickly enough to enable steady levels of access. Government could source these
+vulnerabilities from a combination of the public domain, the commercial exploit market, and a central ``Vulnerability
+Lab'' \cite{bellovin_lawful_2013}.
 
@@
 devices cost law enforcement a mere \$2000 per unit \cite{koepke_2020}. As the exploit arms race continues,
-exploitability will be inconsistent. In the end, law enforcement would face periods where it struggles for access, and
-would need to develop costly exploits in the lab. Mandatory vulnerability disclosure is necessary to avoid competition
-between law enforcement and tech companies. However, based on the \ac{FBI}'s historical abuses of power
-\cite{shamsi_2011} and Congress's reluctance to fund initiatives it has passed \cite{keller_internet_2019} and failure
-to provide strong oversight \cite{johnson_congressional_2004}, it seems unlikely that government would readily forfeit
-expensive vulnerabilities.
+exploitability will be inconsistent. If law enforcement relies on \ac{lawful-hacking} as its strategy, it would face
+periods where it struggles for access, and would need to develop costly exploits in the lab. Mandatory vulnerability
+disclosure is necessary to avoid competition between law enforcement and tech companies. However, based on the
+\ac{FBI}'s historical abuses of power \cite{shamsi_2011} and Congress's reluctance to fund initiatives it has passed
+\cite{keller_internet_2019} and failure to provide strong oversight \cite{johnson_congressional_2004}, it seems unlikely
+that government would readily forfeit expensive vulnerabilities.
 
@@
 metadata. These pieces are designed to reveal the encryption key after the surveiller spends considerable computational
-power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes \ii{mass} surveillance, but it
-also enables \ii{arbitrary} surveillance by anyone with enough computing power. Use in common cases would be impractical
-due to high cost, but use in high profile cases would be less effective due to the likelihood of \ac{EA} workarounds
-employed by sophisticated adversaries. Additionally, the calibration of puzzle difficulty is dependent on the state of
-computing. Since opponents can record data and decrypt it later when computational capabilities have changed, the data
-becomes steadily less secure over time.
+power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass surveillance, but it also
+enables arbitrary surveillance by anyone with enough computing power. Using this class of \ac{DIM} \ac{EA} in common
+cases would be impractical due to its high cost. However, it would be less effective in high-profile cases due to the
+ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, the calibration of puzzle difficulty
+is dependent on the state of computing. Since opponents can record data and decrypt it later when computational
+capabilities have changed, the data becomes steadily less secure over time.
 
-Another \ac{DIM} proposal is to ``conference call'' authorities in as a ghost participant in the conversation. It was
+Another \ac{DIM} proposal is to ``conference call'' authorities in as ghost participants in the conversation. This was
 notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson \cite{levy_robinson_2018}. Their writeup begins
@@
 
-As goodwill participants in the conversation, trying to spur discussion around specific proposals ``without people being
-vilified for having a point of view or daring to work on this as a problem,'' they deserve credit and thanks.
-
-Unfortunately, the proposal they share is indeed problematic. The ghost user proposal prides itself on not compromising
-strong encryption. This is a good property, but compromising a messaging app's authorization protocol is as dangerous as
-compromising its encryption \cite{callas_1_2019}. Fundamentally, it suffers from the same weakness as trusted-party key
-escrow; the introduction of an authorization vulnerability in the messaging service provider's platform amounts to a
-highly risky centralized access capability \cite{schneier_ghost_2019}. It has the same nature, if a different form, as
-rejected key escrow proposals.
+Although this proposal was raised with good intentions, it is nevertheless problematic. The ghost user proposal prides
+itself on not compromising strong encryption. Although this is a good property, compromising a messaging app's
+authorization protocol is as dangerous as compromising its encryption \cite{callas_1_2019}. Fundamentally, this proposal
+suffers from the same weakness as trusted-party key escrow---it introduces an authorization vulnerability in the
+messaging service provider's platform, resulting in dangerous centralized access capability \cite{schneier_ghost_2019}.
+It has the same nature, if a different form, as rejected key escrow proposals.
 
@@
 authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or distributed
-systems. This is where the device escrow shows its weaknesses. Secure hardware enclaves, though improving, are not
-totally reliable, and authentication to the hardware device is susceptible to the same attacks as traditional escrow
-\ac{EA}.
+systems. This is where the weakness of the device escrow is revealed. Secure hardware enclaves, though improving, are
+not totally reliable. Therefore, authentication to the hardware device is susceptible to the same attacks as traditional
+escrow \ac{EA}.
 
-Device key escrow's greatest weaknesses, however, still represent signs of progress. It is true that it relies on a
-device's secure enclave, but devices already do---it is the same hardware device that manages device unlock and
-encryption functionality in the first place. Adding functionality increases the attack surface, but it does not add a
+Device key escrow's greatest weaknesses, however, are still signs of progress. It is true that it relies on a device's
+secure enclave, but devices already do---it is the same hardware device that manages device unlock and encryption
+functionality in the first place. Additional functionality increases the attack surface, but it does not add a
 completely new threat vector. It is also true that the process is subject to the same problems as traditional key escrow
 after the device has been obtained, but that is precisely the progress that has been made---\ii{after the device has
-been obtained}. Leveraging advances in secure hardware design has meaningfully changed the \ac{EA} risk profile.
+been obtained}. Advances in secure hardware design have meaningfully changed the \ac{EA} risk profile.
 
 In 2019, Carnegie Mellon assembled an ideologically diverse group of policy and security experts to engage in the kind
-of cross-disciplinary debate this thesis advocates. After considering all the technical corners of the situation, their
-report goes into detail on why \ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated
-adversaries) focusing on \ac{DAR} in mobile phones using device key escrow requiring physical access is the most
-tractable technical niche to consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed
-with a dose of lawful hacking is best suited for advanced opponents, and device key escrow represents actual progress as
-a path towards lower-risk \ac{EA} for typical cases.
+of cross-disciplinary research. After describing every technical branch of the situation, their report goes into detail
+on why \ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated adversaries) focusing on
+\ac{DAR} in mobile phones using device key escrow requiring physical access is the most tractable technical niche to
+consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed with a dose of lawful hacking
+is best suited for advanced opponents, and device key escrow represents actual progress as a path towards lower-risk
+\ac{EA} for typical cases.
 

### 2020-12-13 Thesis-Chapter6-ThreatModel.tex
Work on DFD section and normalize captions
Editorial improvements

@@
 ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
-proposal is specific to \acl{DAR} for mobile devices. Lacking a concise name, this proposal will be referred to by its
-acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
+proposal is specific to \acl{DAR} for mobile devices. Due to the length of its title, this proposal will be referred to
+by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
 application and use, though it is not the focus of this analysis, and nothing is assumed about it in particular. Note
@@
 emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
-\ac{EA} proposal, but use the term entirely in the spirit of their paper.
+\ac{EA} proposal, but use the term in the spirit of their paper.
 
@@
 
-The essential threat modeling questions are ``what are we building'' and ``what can go wrong.'' The first step to
+``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The first step to
 answering those questions is to establish goals for the former and a list of threats in anticipation of the latter. The
@@
 The goals of an \ac{EA} cryptosystem are taken from the metrics of \mychap{chap-arguments}. Their justification stems
-from the preceding arguments. Each goal is listed according to priority, though none are absolute.
+from the preceding arguments. Each goal is listed according to priority, though no goal has absolute priority over the
+others.
 
@@
 
-\modelitem{Security}{The system should minimize increased security risk with reference to the current state of the art.
-Risk is considered both with respect to individual devices and to the integrity of the entire class of devices.}
+\modelitem{Security}{The system should minimize security risk with reference to the current state of the art. Risk is
+considered with respect to both individual devices and the entire class of devices.}
 
 \modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
-political regimes that violate civil rights. This includes technologically enforcing the fundamental right to privacy
-but upon warranted search pursuant to the rule of law.}
+authoritarian regimes. This includes technologically enforcing the fundamental right to privacy but upon warranted
+search pursuant to the rule of law.}
 
-\modelitem{Transparency}{The system should be auditable by the public to enable trust that this significant power is
-being used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}
+\modelitem{Transparency}{The system should be auditable by the public to enable verification that its power is being
+used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}
 
-\modelitem{Law enforcement utility}{The system should actually be useful to law enforcement in terms of reliability,
-speed, and cost.}
+\modelitem{Law enforcement utility}{The system should be useful to law enforcement in terms of reliability, speed, and
+cost.}
 
-\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem as well
-as the distribution of costs associated with its development and administration. Even though this is a security-oriented
-threat model, architectural decisions determine the system's economics, and economics determines the system's ultimate
+\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem and the
+distribution of costs associated with its development and administration. Even though this is a security-oriented threat
+model, architectural decisions determine the system's economics, and economics determines the system's ultimate
 feasibility.}
@@
 
-\ac{EA} threat models necessarily include the full cast of traditional threat actors.
+\ac{EA} threat models include the full cast of traditional threat actors.
 
@@
 
-\modelitem{Criminals: ordinary, organized, cyber}{Criminals represent a threat to the \ac{EA} system by trying to use it
-to gain access without authority. Ordinary thieves may want to spy on a target or steal and resell phones. They will
-have physical access but only simple tools, using ordinary device interfaces. Organized criminals will have the same
-motives but considerably more advanced technical capabilities, exerting full control over all interfaces. Cyber
-criminals have a wide spectrum of motivations and capabilities. Their significant contribution to the threat landscape
-is that they will target the backend systems that operate the \ac{EA} system in order to steal and likely sell whatever
-access capabilities they can find.}
+\modelitem{Criminals: ordinary, organized, cyber}{Criminals threaten the \ac{EA} system by using it to access plaintext
+without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical
+access but only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but
+considerably more advanced technical capabilities, exerting full control over all interfaces. Cyber criminals have a
+wide spectrum of motivations and capabilities. Their significant contribution to the threat landscape is that they will
+target the backend systems that operate the \ac{EA} system in order to steal and likely sell whatever access
+capabilities they can find.}
 
-\modelitem{Insider threat}{The insider threat is someone within the \ac{EA} recovery process, whether at the law
-enforcement agency or a vendor, that seeks to abuse the system for their own purposes. They may represent the
-motivations and capabilities of any of the criminal categories, but they begin with privileged access somewhere in the
-process.}
+\modelitem{Insider threat}{The insider threat involves someone within the \ac{EA} recovery process, perhaps at the law
+enforcement agency or a vendor, that seeks to abuse the system for their own purposes. They may have the same
+motivations and skills of any other criminal category, but they have the advantage of privileged access at some stage in
+the process.}
 
-\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are always looking for means to escalate
-access and tamper with and steal information. They have world-class expertise and computing power, which they will apply
-to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence agencies, as
-opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S., which are assumed
-to operate within---though push the boundaries of---the legal framework.)}
+\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are constantly looking for ways to escalate
+privilege and tamper with or steal information. They have world-class expertise and computing power, which they can
+apply to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence
+agencies, as opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S.,
+which are assumed to operate within---though push the boundaries of---the legal framework.)}
 
-\modelitem{Authoritarian regimes}{States that do not respect lawful process will try to compel administers of an \ac{EA}
-system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking capability.}
+\modelitem{Authoritarian regimes}{States that do not respect lawful process will try to compel administrators of an
+\ac{EA} system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking
+capability.}
 
@@
 To make sense of an \ac{EA} proposal in the face of the stated goals and such a formidable list of threat actors, one
-must include two non-traditional threats. To do so is to step back and look at the situation from a broader view than
-security experts usually take.
+must include two non-traditional threats. To do so is to step back and view the situation with a broader perspective
+than security experts usually take.
 
@@
 
-\modelitem{The platform abuser}{The device user may actually be a threat in the larger safety context. The end user is
-not the most important thing. This is a crucial point.
+\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
+The end user is not the most important thing. This is a crucial point.
 
 All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of this value system is some most
-important ``thing.'' Traditional device security threat models make the end user the most important thing. This is a
-very good rule. It leads to good security priorities and more importantly, honors the user's rights and autonomy.
+important ``thing.'' Traditional device security threat models define this ``thing'' as the end user. This leads to good
+security priorities and honors the user's rights and autonomy.
 
-To stray from this rule is dangerous, but---make no mistake---it is already done on a regular basis. Software that can
-force updates values the vendor's judgement over the user's (even if it is a good idea). Software that enforces Digital
-Rights Management values corporations' data rights over the user's (sometimes to extremes \cite{eff_2020}). Software
-that doesn't encrypt user communications at all values the service provider's ability to scrape personal information to
-create targeted ads over the user's privacy. Software that enables \ac{EA} values lawful investigatory process
-over the user's ultimate privacy. Intentional subversion of the user's will is not a comfortable thought, but it is a
-frequent reality, and the reality of \ac{EA}.
+To stray from this rule is dangerous, but is already frequent practice. Software that forces updates values the vendor's
+judgement over the user's (even if it is a good idea). Software that enforces Digital Rights Management values
+corporations' data rights over the user's (sometimes to extremes \cite{eff_2020}). Software that does not encrypt user
+communications at all values the service provider's ability to scrape personal information over the user's privacy.
+Software that enables \ac{EA} values lawful investigatory process over the user's ultimate privacy. Intentional
+subversion of the user's will is not a comfortable thought, but it is a frequent reality, and the reality of \ac{EA}.
 
-In the broader threat model, the platform abuser is someone who is using the device encryption for cover of illicit
-activities. He represents a threat to public safety.}
+In the broader threat model, the platform abuser uses encryption to hide illicit activities. He is a threat to public
+safety.}
 
@@
 aggressive lawmakers that have the power to \ii{mandate} arbitrary data access represent perhaps the largest threat of
-all. Crypto-anarchists can encrypt their data above the compromised layer all they want, but the vast majority of
-individuals using default products and all above-the-table organizations will abide by U.S. law.
+all. Crypto-anarchists can encrypt their data above the compromised layer all they want, but above-the-table
+organizations and most individuals who use default products will abide by U.S. law.
 
-Hawkish lawmakers are a threat to the security of a cryptosystem (using \ac{EA} or not) in that they may judge it too
-inhibitive to law enforcement and outlaw its very use, mandating something weaker in its place.}
+Hawkish lawmakers threaten the security of a cryptosystem (using \ac{EA} or not) because they may judge the system as
+too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something weaker in its place.}
 
@@
 
-The following considerations are out of scope for this threat model.
+The following considerations are beyond the scope of this threat model.
 
@@
 
-\modelitem{\ac{EA} workarounds}{The user is capable of evading \ac{EA} mechanisms by encrypting data at a higher level
-in the tech stack, e.g. by encrypting files before saving them to disk. Additionally, investigators or attackers are
-frequently capable of bypassing current lockout mechanisms. These have already been discussed, and it out of scope for
-evaluating the \ac{EA} system itself.}
+\modelitem{\ac{EA} workarounds}{The user can evade \ac{EA} mechanisms by encrypting data at a higher level in the tech
+stack, for example by encrypting files before saving them to disk. Additionally, investigators or attackers can often
+bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the \ac{EA}
+system itself.}
 
-\modelitem{Supply chain attacks}{Protecting device hardware and software supply chains is necessary to prevent
-surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and software are
-considered to be as-intended, though they are assumed to contain the standard amount of bugs and side-channels.}
+\modelitem{Supply chain attacks}{It is necessary to protecting device hardware and software supply chains in order to
+prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
+software are considered to be as-intended, though they are assumed to contain the standard amount of bugs and
+side-channels.}
 
@@
     \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone.png}
-    \caption{The basic encrypted mobile phone data flow diagram.}
+    \caption{The basic encrypted mobile phone data flow diagram}
     \label{fig-dfd-iphone}
@@
     \includegraphics[width=0.9\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
-    \caption{The LDAWMSR data flow diagram.}
+    \caption{The LDAWMSR data flow diagram}
     \label{fig-dfd-ldawmsr}
@@
   \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR-setup.png}
-  \caption{The LDAWMSR maintenance data flow diagram.}
+  \caption{The LDAWMSR maintenance data flow diagram}
   \label{fig-dfd-ldawmsr-setup}

### 2020-12-13 Thesis-Chapter7-PathsAhead.tex
Editorial improvements

@@
 average attacker is difficult, and the best attacker, impossible. \Ac{cryptography} plays a fundamental role in
-security, yet its simultaneous role in enabling unprecedented privacy means it has controversial side effects.
+security, yet it remains a controversial technology due to its simultaneous role in enabling unprecedented privacy.
 
-One of these side effects is encryption-enabled privacy's impact on public safety. Despite a shortage of quantitative
-analyses showing its impact on crime, encryption certainly does allow evidence of wrongdoing to be hidden from service
-providers and law enforcement. This enables harmful activity to go unchecked and decreases public safety, which is a
-legitimate concern and the primary motivation for \ac{EA}.
+First, proponents of \acl{EA} argue that encryption-enabled privacy as harmful effects on public safety. Despite a
+shortage of quantitative analyses showing its impact on crime, encryption certainly does allow evidence of wrongdoing to
+be hidden from service providers and law enforcement. This enables harmful activity to go unchecked and decreases public
+safety, which is a legitimate concern and the primary motivation for \ac{EA}.
 
-However, at the same time as encryption is enabling strong privacy, the broader technological path society is taking is
-stripping it away. The ubiquity and insecurity of networked computers has been a bear to privacy, enabling identity
-theft, inappropriate data mining, \ac{masssurv}, and other threats. To the extent that encryption mitigates these, it
-increases public safety.
+However, while encryption adds privacy, technological change as a whole is stripping it away. The ubiquity of modern
+computing has created terabytes of data on every individual; access laws, business models, and insecurity have exposed
+this data to huge audiences. This trend has enabled \ac{masssurv}, inappropriate data mining, identity theft, and other
+threats. When encryption mitigates these threats, it increases public safety.
 
-Finally, deciding how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---involves
-considerable questions of trust. Law enforcement agencies, intelligence agencies, technology makers, individual users,
-and the public at large are all players; each have demonstrated untrustworthiness, and different approaches disperse
-trust differently.
+Finally, deciding how to respond to these issues---whether with inaction, \ac{EA}, or some alternative---requires
+determining what institutions are worthy of trust. Potential solutions assign different levels of trust to law
+enforcement agencies, intelligence agencies, technology makers, individual users, and the public at large. Each of these
+groups have demonstrated untrustworthiness, but all paths forward require putting trust somewhere.
 
-This complex mix of issues makes encryption policy a \ac{wicked-prob}. Even just the problem requires explanation upon
-explanation and research upon research. Any formulation is concomitant with a solution, but the outcome of a solution
-cannot be accurately evaluated in isolation. The underlying conditions constantly change due to external forces and
-policy actions (or lack thereof).
+This complex mix of issues makes encryption policy a \ac{wicked-prob}. Even formulating the problem requires several
+iterations of research into history, arguments, and technology. Every formulation presupposes a solution, but the
+effectiveness of a solution cannot be accurately evaluated. Meanwhile, the underlying conditions constantly change due
+to external forces and policy actions (or lack thereof).
 
-In this thesis I propose a method to confront \acp{wicked-prob}, a modified \ac{OODAloop} that emphasizes collaborative
-debate between individuals diverse in background and values, facilitated in part by argument maps, and focusing on
-specific proposals. I try to capture that here by surveying the debate, mapping the arguments, and ultimately performing
-threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions and paths forward for both
-the technical and policymaking communities.
+In this thesis, I propose a method to confront \acp{wicked-prob} based on a modified \ac{OODAloop}. The method
+emphasizes collaborative debate between diverse individuals, facilitated in part by argument maps and focused on
+specific proposals. The second half of the thesis applies this method by surveying the debate, mapping the arguments,
+and ultimately performing threat model analysis of a specific \ac{EA} proposal for \acl{DAR}. I now turn to conclusions
+and paths forward for both the technical and policymaking communities.
 
@@
 
-Regarding the arguments analyzed in \mychap{chap-arguments}, while neither side of the \ac{EA} debate fully incorporates
-the whole picture, the anti-\ac{EA} argument is currently stronger. Encryption does pose certain threats to public
-safety. Terrorists and criminals can and do use encryption to evade the law; this has allowed for shocking levels of
-\acf{CSAM} to be peddled \cite{keller_internet_2019}. However, the argument that \ac{encryption} is causing \acl{LE}
-failure is not currently strong enough to justify an \ac{EA} mandate. For \ac{DIM}, the primary issues are poor
-coordination and training \cite{carter_2018}; for \ac{DAR}, commercial \ac{lawful-hacking} tools currently have the
-upper hand on mobile phone manufacturers, meaning \acl{LE} can get into almost any mobile device they acquire
-\cite{koepke_2020}. Additionally, strong \ac{encryption} is necessary to protect security and privacy---two public goods
-that other technological developments often fail to provide.
+Based on the arguments analyzed in \mychap{chap-arguments}, neither side of the \ac{EA} debate addresses all of the root
+issues. However, given the wide gap between the requests of government officials and the maturity of \ac{EA} technology,
+the anti-\ac{EA} argument is currently stronger. Encryption does pose certain threats to public safety. Terrorists and
+criminals can and do use encryption to evade the law; in particular, this has allowed for shocking levels of \acf{CSAM}
+to be peddled \cite{keller_internet_2019}. However, the argument that \ac{encryption} is causing \acl{LE} failure is
+unsubstantiated at this time. For \acl{DIM}, poor coordination and training hamper law enforcement more than encryption
+\cite{carter_2018}. For \acl{DAR}, commercial \ac{lawful-hacking} tools currently outdo mobile phone manufacturers,
+meaning \acl{LE} can get into almost any mobile device they acquire \cite{koepke_2020}. Additionally, strong
+\ac{encryption} is a necessary tool in protecting security and privacy---two public goods that other technological
+developments often fail to provide.
 
-Though the pro-\ac{EA} argument is ultimately not strong enough to justify an \ac{EA} mandate, incorporating it to the
-technologist's point of view teaches important lessons. One is that holistic cryptosystem threat models must include two
-non-traditional threat actors, the malicious user and threatening lawmaker. Encryption lends power to its user, and this
-power can be abused. Technology makers should treat abuse of their creations as a threat, and take mitigating steps.
-Governments are naturally sensitive to shifts in power, and thus are interested in the impacts of encryption; lawmakers
-represent a unique threat in that they can outlaw certain types of encryption entirely if they deem them too dangerous.
-The threat of an \ac{EA} mandate should be treated as a true threat in the security modeling sense.
+Though the pro-\ac{EA} argument does not justify an \ac{EA} mandate, it can teach technologists some important lessons.
+First, holistic cryptosystem threat models must include two non-traditional threat actors, the malicious user and
+threatening lawmaker. Encryption lends power to its user, and this power can be abused. Technology makers must include
+abusive use in their threat modeling. Governments are inherently sensitive to changes in personal privacy; lawmakers
+therefore represent a unique threat because they can outlaw uses of encryption entirely if they deem them too dangerous.
+The danger of an \ac{EA} mandate must also be included in the cryptosystem threat model.
 
-Another lesson is that the current situation is temporary. This is necessarily true of \acp{wicked-prob}, but especially
-true for encryption policy in the context of rapidly evolving technology and a vulnerability arms race between device
-manufacturers and lawful hackers. The facts of the case are bound to change, and the conclusion that \ac{EA} is
-unjustified may change as well.
+Second, \ac{EA} history teaches that the current situation is temporary. This is true of all \acp{wicked-prob}, but
+especially for encryption policy due to rapidly evolving technology and a vulnerability arms race between device
+manufacturers and lawful hackers. The circumstances are sure to change, and the conclusion that \ac{EA} is unjustified
+may change as well.
 
-The additions to the threat model and the instability of present circumstances mean that additional research into
-\ac{EA} is fully justified. It belongs as just a component of the effort that society should be putting towards
-addressing the broader set of issues. Fortunately, developments in secure hardware mark real progress in the search for
-acceptably secure \ac{EA}. Secure hardware technology is clearly imperfect, and alone does not resolve the numerous
-challenges that full-fledged \ac{EA} proposals must overcome, but device escrow for \acl{DAR} such as that described in
-Savage's proposal offers transparency, protection from \ac{masssurv}, and \acl{LE} utility while making progress on
-security.
+% Research is just part of the effort that society should put towards addressing the root issues.
 
-The conclusions I draw do not settle the issue. However, when dealing with \acp{wicked-prob}, meaningful debate and
-discussion counts as progress even when it doesn't seem like it. \Acp{wicked-prob} cannot be analyzed into resolution;
-grasping the problem alone constitutes much of the work towards solving it. The conclusions above do not bring closure,
-but they should bring more clarity, and its clarity that counts.
+Further research into \ac{EA} is justified by the additions to the threat model and the instability of present
+circumstances. Fortunately, developments in secure hardware mark real progress in the search for acceptably secure
+\ac{EA}. Secure hardware technology is not perfect; on its own, it does not resolve the numerous challenges that
+full-fledged \ac{EA} proposals must overcome. However, \ac{EA} proposals leveraging secure hardware, such as device
+escrow for \acl{DAR}, represent progress. They offer transparency, protection from \ac{masssurv}, and \acl{LE} utility
+while remaining somewhat secure.
+
+The conclusions presented here do not settle the issue. However, when dealing with \acp{wicked-prob}, meaningful
+discussion can be considered progress even if it does not reach concrete resolution. Much of the work involved in
+confronting \acp{wicked-prob} is understanding the problem itself. The conclusions above do not bring closure, but
+clarity.
 
@@
 
-The technical community needs to accept and deal with the enormous role technology is playing in current policy issues,
-such as encryption, automation, and misinformation. Government sometimes suggests that the tech world is not trying hard
-enough, and that ``nerding harder'' will result in solutions. The tech world is absolutely right that nerding harder
-will not work on these issues---because they are \acp{wicked-prob}---but that does not mean they cannot deal with them.
-Tech is used to addressing tame problems, but it is time that they turn their massive financial and intellectual capital
-towards strategies designed for \acp{wicked-prob}. This means including people of diverse experiences, backgrounds, and
-values in decision making processes that steer investment and design decisions. It means creating products and business
-models that are responsive to evolving conditions. It means considering the ways their creations will be abused.
+It is the responsibility of the technical community to accept and address the substantial role technology plays in
+current policy issues, such as encryption, automation, and misinformation. Government sometimes suggests that the tech
+community simply needs to try harder. Because these are not tame problems, the tech world is absolutely right that
+simply trying harder will not work---but that does not mean they cannot deal with them. The tech industry is optimized
+to address tame problems, but it must now turn its financial and intellectual capital towards strategies designed for
+\acp{wicked-prob}. In order to do so, the tech industry must involve people of diverse backgrounds and values in design
+and investment decisions. Designers must consider the ways in which their products could be abused and business models
+must be responsive to evolving conditions.
 
-Work on \acp{wicked-prob} spins out smaller problems, some tame and some wicked yet manageable (see
-\myfig{fig-policy-ooda-process}). Aside from making the changes to tackle \acp{wicked-prob} head on, the technical
-community should also work on the tame sub-problems, a few of which are listed here.
+Addressing \acp{wicked-prob} reveals smaller subproblems, some tame and some also wicked (see
+\myfig{fig-policy-ooda-process}). In addition to making changes to tackle \acp{wicked-prob} directly, the technical
+community should also address the tame subproblems. Some suggestions for doing so are listed below.
 
@@
 
-\ac{EA} is only applicable when security is actually strong enough to keep attackers out in the first place, which is
-not the case for current device encryption. To some extent, this is a security ergonomics problem more than a technical
-problem: cryptographers can make device encryption incredibly strong today, but users will not let them because of its
-adverse effects on usability. Solving the strong yet usable device encryption problem will make the most promising class
-of \ac{EA}, device key escrow, relevant. Improving basic security across the board will make any \ac{EA} proposal more
-tenable.
+\ac{EA} is only worthwhile when security is strong enough to keep attackers out in the first place, which is not the
+case for current device encryption. This is a security ergonomics problem more than a technical problem---cryptographers
+are already capable of creating stronger device encryption, but users would rebel due to its adverse effects on
+usability. Solving the strong-yet-usable device encryption problem will make the most promising class of \ac{EA}, device
+key escrow, relevant. Improving security in general will make any \ac{EA} proposal more tenable.
 
@@
 
-One of the largest non-technical risks of \ac{EA} is abuse by jurisdictions with poor rule of law, or where law
-insufficiently respects civil rights in the first place. This is the motive for transparency as a goal of the threat
-model. Device key escrow can achieve partial transparency through physical possession and disassembly requirements.
-Ideally, usage of \ac{EA} mechanisms would be fully auditable to the public through inherent technical mechanisms. This
-would mark an improvement for \acl{DAR} and is a necessity for progress on \acl{DIM}.
+One of the largest non-technical risks of \ac{EA} is its potential abuse by jurisdictions with poor rule of law. This is
+the reason transparency is a goal of the threat model. Device key escrow can achieve partial transparency through
+physical possession and disassembly requirements. Ideally, usage of \ac{EA} mechanisms would be fully auditable to the
+public through inherent technical mechanisms. This could be improved for \acl{DAR} and must be improved for \acl{DIM}.
 
@@
 
-Centralized trust in \ac{EA} proposals represent their greatest risk, whether it is in the form of a database of keys or
-a control center with the power to add ``ghost users'' to conversations. Knowing that insecurity (and trust) cannot be
-destroyed, only moved around, the goal is a distribution of trust that allows the system to function without large
-amounts of trust assigned to any single party. This would contribute to transparency and greatly reduce the risk profile
-of an \ac{EA} proposal overall.
+Most \ac{EA} proposals would still result in a concentrated center of risk, whether it would take the form of a database
+of keys or of a control center with the ability to add ``ghost users'' to conversations. As explained in
+\mysec{sec-premises}, insecurity cannot be destroyed, it can only be moved around. Distributed trust mechanisms avoid
+assigning a large amount of trust to any single party. This contributes to both transparency and security.
 
@@
 
-Fortunately, research in these areas is occurring. More is necessary. Some \ac{EA} proposals include built in
-transparency and distributed trust mechanisms \cite{goos_oblivious_1996} \cite{phan_key_2017}
-\cite{servan_schreiber_jje_2020}; these proposals' weakness is that the approaches are insufficiently understood to be
-trusted with the high stakes of implementation. While these problems remain unsolved, technologists must continue to
-provide clear security advice and be vigilant against bad technology policy.
+Fortunately, this research is already happening to some degree, but more is necessary. Some \ac{EA} proposals include
+built-in transparency and distributed trust mechanisms \cite{goos_oblivious_1996} \cite{phan_key_2017}
+\cite{servan_schreiber_jje_2020}. Like device key escrow, approaches such as these hold promise, but must be better
+understood before they could be implemented. Even while these problems remain unsolved, technologists must provide clear
+security advice and speak out against bad policy.
 
@@
 
-% Prof. Walsh:
+% TODO: Prof. Walsh:
 % - Liked the lessons for techies; this distills the same problems for policymakers
@@
 
-Government has not been silent on the issue of encryption and lawful access. The dynamics created by the executive,
+Policymakers have not been silent on the issue of encryption and lawful access. The influence of the executive,
 legislative, and judicial branches ensures that the regulatory environment will evolve with changing federal policies,
-new laws, and rulings that refine implementation existing law. The judiciary will indeed shape encryption policy through
-its rulings, and is correct to be reevaluating current Fourth Amendment jurisprudence in light of rapidly changing
-circumstances. Still, the primary actors capable of consciously taking part in strategic policymaking will be the
-executive and legislative branches.
+laws, and rulings. The judiciary will shape encryption policy through its rulings and by reevaluating Fourth Amendment
+jurisprudence in light of rapidly changing circumstances. Still, the primary actors capable of consciously taking part
+in strategic policymaking will be the executive and legislative branches.
 
 Executive and legislative policymakers must strategically approach encryption policy as a \ac{wicked-prob}. Industry has
-failed at analyzing the problem away, and government has failed at muddling the problem away. It is true that
-incrementalism has safely avoided dangerous outcomes such as blanket \ac{EA} mandates, but it has also resulted in a
-research-stifling high-tension debate and circumstances of poor utilization of current capabilities mixed with
-unregulated use of invasive \ac{lawful-hacking} tools. Policymakers and law enforcement agencies need to be forthcoming
-with the real challenges they face. They are important contributors to the diverse groups that can iteratively refine
-the problem definition and possible solutions. Just like the technical side, the policy side of the solutions must be
-responsive to evolving conditions. Most importantly, any initiative needs follow through, since a single \ac{OODAloop}
-iteration will not be enough to tame this issue.
+failed to resolve the problem with analysis, and government has failed to resolve the problem with \ac{incrementalism}.
+\Ac{incrementalism} has safely avoided dangerous outcomes such as blanket \ac{EA} mandates. However, it has also
+resulted in a research-stifling debate, the poor utilization of current capabilities, and the unregulated use of
+invasive \ac{lawful-hacking} tools. While policymakers and law enforcement agencies have made important contributions to
+the debate, they need to share data more openly. Just like the technical aspects, the policy aspects of the solutions
+must be responsive to evolving conditions.
 
-Based on the analysis performed thus far, there are a few policy changes that would be good first steps to take today.
+Based on the analysis performed thus far, the following steps could be taken today.
 
@@
 
-Law enforcement is often ill-equipped to handle digital evidence and coordinate with service providers using existing
-access mechanisms \cite{carter_2018}. Justice Department task forces to combat \acl{CSAM} are underfunded and
-understaffed \cite{keller_internet_2019}. Improving the effectiveness of existing capabilities and programs carries no
-risk and guaranteed results.
+Law enforcement is ill-equipped to access digital evidence with existing mechanisms \cite{carter_2018}. Justice
+Department task forces for combatting \acl{CSAM} are underfunded and understaffed \cite{keller_internet_2019}.
+Investment in existing programs will likely produce results without introducing risk.
 
@@
 
-\Ac{lawful-hacking} using commercial mobile device forensic tools has become a common practice but invokes serious
-privacy and civil rights concerns \cite{koepke_2020}. Regardless of any long-term strategy to encryption and lawful
-access policy, \ac{lawful-hacking} must be constrained. Michigan's 2020 constitutional amendment represents one positive
-development: it explicitly protects electronic data and communications from search without a warrant
-\cite{ballotpedia_2020}. Protections like this enable a non-abusive use of current \ac{lawful-hacking} capability.
+\Ac{lawful-hacking} using commercial mobile device forensic tools has become a common practice; this invokes serious
+privacy and civil rights concerns \cite{koepke_2020}. Regardless of long-term encryption policy strategy,
+\ac{lawful-hacking} must be regulated. Michigan's 2020 constitutional amendment represents one positive development: it
+explicitly protects electronic data and communications from search without a warrant \cite{ballotpedia_2020}.
+Protections like this can prevent abuse of current \ac{lawful-hacking} capability.
 
@@
 
-The government wants technical advancement in \ac{EA} technology, but there is very little financial incentive for
-private companies to research \ac{EA}, and the threatening climate is stifling academic research. If the federal
-government is going to compare building secure \ac{EA} to going to the moon \cite{cushing_moon_2018}, then they should
-fund it like the Apollo program. In order to be welcomed, such funding should be accompanied with a binding statute
-excluding the possibility of any premature \ac{EA} mandate.
+Lawmakers request research into \ac{EA} technology, but private companies lack financial incentive and universities lack
+interest due to the threatening climate. The federal government cannot compare building secure \ac{EA} to going to the
+moon if it does not provide the same level of funding. In order to be welcomed by the academic community, such funding
+must be accompanied by a binding statute excluding the possibility of a premature \ac{EA} mandate.
 
@@
 
-The goal of this thesis is to clarify encryption and \ac{EA} as a \ac{wicked-prob}, to break down the debate, and to
-evaluate the security of current \ac{EA} proposals. I aim to move the discussion forward with the ideas of how to tackle
-\acp{wicked-prob}, additions to the traditional cryptosystem threat model, and security analysis of a modern proposal.
-In the language of the \ac{OODAloop}, this study is observation and orientation; it is up to technologists,
-policymakers, and common people to decide and act. Contending with these issues is difficult on many levels, but the
-future of security, safety, and privacy depend on coming together to find a good answer where no perfect answer exists.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% OLD NOTES for reference
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% In this chapter, describe actions to take in discussion, regulation, and research.
-
-% \section{Legislative Priorities}
-
-% \begin{itemize}
-%     \item Reject EA until technical capabilities are ready \cite{varia_2018}
-%     \item Be forthcoming with data so that we know what to work on \cite{devlin_2018}
-%     \item Don't prescribe technical methods, but be clear about principles \cite{matyas_incommensurability_2018}
-%         (e.g., ECPA became outdated because was too tied to specific technology \cite{shamsi_2011})
-%     \item Fund the research \cite{varia_2018} (like \cite{goss_hr2616_1999})
-% \end{itemize}
-
-% \section{Technical Research}
-
-% \begin{itemize}
-%     \item Acknowledge the moral character of cryptography \cite{rogaway_moral_2015}
-%     \item Do the research
-%     \item Focus on principles \cite{levy_robinson_2018}
-%     \item Be clear on definitions \cite{varia_2018}, but don't expect govt to provide those (unlike
-%             \cite{abelson_2015}), and clearly define their interfaces with principles
-%             \cite{matyas_incommensurability_2018}
-% \end{itemize}
+The goal of this thesis is to establish encryption and \ac{EA} as a \ac{wicked-prob}, to analyze the debate, and to
+evaluate the security of current \ac{EA} technology. I aim to advance the discussion with a strategy to tackle
+\acp{wicked-prob}, additions to the cryptosystem threat model, and detailed analysis of a modern proposal. The strategy
+is based on the \acl{OODAloop}. In those terms, this paper serves to \ii{observe} and \ii{orient}. Technologists,
+policymakers, and common people must \ii{decide} and \ii{act}. Confronting these issues may be difficult, but the future
+of security, safety, and privacy depends on finding an acceptable solution where no perfect solution exists.

### 2020-12-16 Thesis-Chapter2-Background.tex
Address some TODOs

@@
 
-A short law designed to grant federal courts the right to ``issue all writs necessary or appropriate in aid of their
-respective jurisdictions and agreeable to the usages and principles of law,'' creating a legal framework for providing
-data to law enforcement. The \acs{AWA} was famously cited in the Apple vs. FBI case, though its applicability to \ac{EA}
-in general is unsettled \cite{matyas_incommensurability_2018}.
+The \acs{AWA} is a short law designed to grant federal courts the right to ``issue all writs necessary or appropriate in
+aid of their respective jurisdictions and agreeable to the usages and principles of law,'' creating a legal framework
+for providing data to law enforcement. The \acs{AWA} was famously cited in the Apple vs. FBI case, though its
+applicability to \ac{EA} in general is unsettled \cite{matyas_incommensurability_2018}.
 
@@
 
-The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: protection from
-unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.
+Three Bill of Rights amendments are frequently cited in the \ac{EA} debate. I: freedom of speech; IV: protection from
+unreasonable searches and seizures and requirements for warrants; and V: the right against self-incrimination.
 
@@
 
-The \ac{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
+\acs{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
 Provides explicit protections for private electronic communications data, and updated definitions to contemporary
@@
 }
-% FIXME: make tense for enacted laws present
 
@@
 
-\ac{CALEA} mandates that telecommunications carriers enable the government to ``intercept all of the subscriber's wire
+\acs{CALEA} mandates that telecommunications carriers enable the government to ``intercept all of the subscriber's wire
 and electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized
@@
 
-% TODO: find a "con" word that means "denied"! Contested?
-
 The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
-authoritative legal texts. However, whether it is condemned, denied, conceded, or condoned, the judicial branch also
+authoritative legal texts. However, whether it is contested, conceded, condemned, or condoned, the judicial branch also
 establishes legal doctrine and actively enforces rulings in a manner that amounts to creation of public policy

### 2020-12-16 Thesis-Chapter4-Tools.tex
Address some TODOs

@@
 
-% TODO: can probably cut more
-Argument mapping tools originated in 1970 with text-based \acp{IBIS} \cite{kunz_issues_1970}. Interest in \acp{IBIS} has
-waxed and waned. Some software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and
-Compendium \cite{dutoit_hypermedia_2006}, which were designed to produce graphical argument representations. Argdown is
-a recent tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
+Argument mapping tools originated in 1970 with text-based \acp{IBIS} \cite{kunz_issues_1970}. Graphical software mapping
+tools include the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006}. Argdown is a
+recent tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
 \cite{voigt_argdown_2018}. Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA}

### 2020-12-16 Thesis-Chapter5-Arguments.tex
Address some TODOs

@@
 changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract worry
-to many, but for the most vulnerable, they are frighteningly concrete. For example, Saudi journalist Jamal Khashoggi's
-fell
-% TODO: bookmark
-
-private communications were surveilled by the Saudi government before he became a victim of politically motivated murder
-\cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free speech
-\cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.
+to many, but for the most vulnerable, they are frighteningly concrete. For example, mobile phone surveillance malware
+enabled the Saudi government to capture and murder journalist Jamal Khashoggi for expressing dissent against the crown
+prince \cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free
+speech \cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.
 
@@
 
-% TODO: I still really want to use this!
+% TODO: Possibly find a new spot for this Encrochat story
 % European law enforcement's infiltration of encrypted messaging service Encrochat exposed crime rings and led to a
@@
 This approach is already in use \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
-\cite{feuer_chapo_2019}.
-
-The real change that \ac{lawful-hacking} would take would be a sanctioning and formalization of this strategy.
-% TODO: tie to paragraph above?
+\cite{feuer_chapo_2019}. Because it is already used, deliberately choosing \ac{lawful-hacking} as the primary strategy
+for accessing plaintext would mean officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.
 

### 2020-12-16 Thesis-Chapter2-Background.tex
Editorial improvements

@@
 The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
-authoritative legal texts. However, whether it is contested, conceded, condemned, or condoned, the judicial branch also
+authoritative legal texts. However, whether it is contested or conceded, condemned or condoned, the judicial branch also
 establishes legal doctrine and actively enforces rulings in a manner that amounts to creation of public policy

### 2020-12-16 Thesis-Chapter5-Arguments.tex
Editorial improvements

@@
 \cite{landau_making_2013} unveiled the scale of privacy-eroding U.S. government surveillance enabled by technological
-changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract worry
-to many, but for the most vulnerable, they are frighteningly concrete. For example, mobile phone surveillance malware
+changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract concern
+to many, but for those most vulnerable, they are frighteningly concrete. For example, mobile phone surveillance malware
 enabled the Saudi government to capture and murder journalist Jamal Khashoggi for expressing dissent against the crown
-prince \cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free
-speech \cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.
+prince \cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and suppresses
+free speech \cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.
 
@@
 \ac{cryptography} plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought
-to compromise the cryptographic foundations of encryption, though this is not a feature of proposals today, as that is
-universally seen as too risky. \mysec{sec-tech-approaches} introduced several alternative technical approaches to
-\ac{EA}, but experts still proclaim that the current state of the art is far from capable of providing the type of
-access \acl{LE} in an acceptably secure manner \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an
-afterthought in early computing and networking designs, and the field of cybersecurity is still equal parts art and
-science; mandated \ac{EA} would inevitably diminish security to some degree \cite{abelson_2015}.
+to compromise the cryptographic foundations of encryption; however, this is seen as too risky today.
+\mysec{sec-tech-approaches} introduced several alternative technical approaches to \ac{EA}, but experts still make clear
+that current systems are incapable of securely providing the level of access \acl{LE} wants \cite{abelson_2015}
+\cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs, and the field of
+cybersecurity is still equal parts art and science; mandated \ac{EA} would put a heavy burden on a field still finding
+its legs.
+% TODO: too colloquial?
 
@@
 password recovery, or administrator access in a business setting. Typically these applications have other solutions,
-however, and to combine the requirements of these uses with those of any reasonable \ac{EA} proposal would likely
-fatally weaken it.
+however, and unnecessary requirements would weaken any \ac{EA} system.
 
@@
 
-A significant portion of the disconnect between government and technical community representatives is disagreement over
-whether law enforcement has \ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is
-well-represented by proponent former \ac{FBI} Director James Comey in his 2014 speech titled ``Going Dark''
-\cite{comey_2014} (though he did not coin the phrase \cite{swire_encryption_2011}). The argument is presented in
-\myfig{fig-arg-going-dark}.
+Much of the disconnect between government and the technical community is disagreement over whether law enforcement has
+\ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is well-represented by proponent
+former \ac{FBI} Director James Comey in his 2014 speech titled ``Going Dark'' \cite{comey_2014} (though he did not coin
+the phrase \cite{swire_encryption_2011}). The argument is presented in \myfig{fig-arg-going-dark}.
 
@@
 encryption. As described in \mysec{sed-arg-contrib}, there is no conclusive data regarding the extent to which
-encryption is to blame for failed cases. One can certainly understand the argument, but the lack of evidence is an
-issue. The highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended when the \ac{FBI} broke into the phone and found
-nothing of value. Tellingly, the examples cited in Comey's 2014 speech were not cases in which encryption inhibited
-investigations, but were cases in which encryption \ii{could have} inhibited investigations, had it been a factor.
-\cite{comey_2014}. More recently, a 2020 report titled ``Mass Extraction: The Widespread Power of U.S. Law Enforcement
-to Search Mobile Phones'' revealed that law enforcement agencies currently have the upper hand on mobile device
-encryption and are actually able to extract data from nearly every mobile phone on the market \cite{koepke_2020}.
+encryption is to blame for failed cases. The highest-profile case of \ac{the-cw2}, Apple vs. FBI, ended when the
+\ac{FBI} broke into the phone and found nothing of value. Tellingly, the examples cited in Comey's 2014 speech were not
+cases in which encryption inhibited investigations, but were cases in which encryption \ii{could have} inhibited
+investigations, had it been a factor. \cite{comey_2014}. More recently, a 2020 report titled ``Mass Extraction: The
+Widespread Power of U.S. Law Enforcement to Search Mobile Phones'' revealed that current hacking tools are stronger than
+mobile device encryption's defenses, and law enforcement is actually able to extract data from nearly every mobile phone
+on the market \cite{koepke_2020}. Without conclusive data showing the extent that encryption hinders investigations, the
+``Going Dark'' argument weakens.
 
-Citing lack of evidence that encryption is stymieing cases only gets you so far, however. In most cases, it is
+Lack of evidence for an argument does not prove that it is incorrect. Because law enforcement has the data, the burden
+of proof lies with them to show that encryption is causing investigatory failure. However, in most cases, it is
 impossible to know what would have happened without the interference of strong encryption. Inability to gather data from
@@
 
-Aside from the challenge of encryption entirely, a 2018 study found that law enforcement's primary difficulties with
-digital evidence were identifying service providers and getting the sought after data from them \cite{carter_2018}. This
-is in the context of current legal frameworks and access capabilities. The report is fittingly titled ``Low-Hanging
-Fruit,'' and also lists resource limitations and training needs among top priorities. Indeed, many current initiatives
-are underfunded. Major anti-\ac{CSAM} legislation passed in 2008, but has been funded at only 50\% of authorized levels,
-the Justice Department task force has produced only two of five biennial reports, and a senior executive position within
-the Justice Department was never created \cite{keller_internet_2019}. Federal entities commissioned to improve
-communication between law enforcement and industry as well as to train law enforcement on managing digital evidence are
-similarly underfunded \cite{carter_2018}.
+Encryption is not the only challenge that law enforcement faces with digital evidence. A 2018 study found that
+identifying service providers and getting data from them using existing mechanisms were even greater impediments
+\cite{carter_2018}. The report, titled ``Low-Hanging Fruit,'' also lists resource limitations and poor training as top
+issues. Indeed, many current initiatives are underfunded. The \ii{New York Times}' 2019 investigation into \ac{CSAM}
+found chronic lack of commitment from government initiatives---major legislation passed in 2008, but has been funded at
+only 50\% of authorized levels; a commissioned Justice Department task force has produced only two of five biennial
+reports; and a senior executive position within the Justice Department to target \ac{CSAM} was never created
+\cite{keller_internet_2019}. Federal entities aiming to improve training and coordination between law enforcement and
+industry are similarly underfunded \cite{carter_2018}.
 
@@
 authorities much more digital evidence than encryption has taken away. Paired with the dangers of \ac{masssurv},
-encryption is a necessary defense, and any \ac{EA} implementation would compromise encryption's protection. The argument
-is mapped in \myfig{fig-arg-golden-age}.
+encryption is a necessary defense. \Ac{EA} would compromise encryption's defense against \ac{masssurv}, and must
+therefore be opposed. The argument is mapped in \myfig{fig-arg-golden-age}.
 
@@
 After establishing the danger of surveillance, the argument notes how technology has increased the government's
-surveillance abilities. This can hardly be contested. Going back to Comey's ``Going Dark'' examples once more, each case
-he cited was cracked by digital evidence that would not have been available even twenty years ago \cite{comey_2014}. In
-response to the danger of \ac{masssurv} and the scale of modern surveillance capabilities, encryption is one of the only
-privacy-saving tools. (Opting out may be possible for some, but is not a realistic option for society as a whole.)
+surveillance abilities. This statement as strong evidence. In Comey's ``Going Dark'' speech, it was digital evidence
+which would not have existed twenty years ago that cracked each case \cite{comey_2014}. Encryption is one of the only
+privacy-saving technologies to emerge alongside the boom in privacy-eroding technologies. Opting out of the
+technological boom is possible for some, but is not a realistic option for society as a whole. Due to the dangers of
+\ac{masssurv}, \ac{encryption} must be embraced.
 
@@
 resource-strapped local, state, and federal law enforcement. It is obvious that the amount of digital evidence is
-increasing. However, law enforcement already has trouble acquiring many forms of evidence with the help of tech and
-communications companies, and does not have the funding or capability to get it on its own \cite{carter_2018}.
-Exacerbating the situation is the ``tech effect,'' which refers to jurors' increased expectations of hard evidence in cases
-where it may be suspected to exist \cite{shelton_study_2006}.
+increasing. However, as seen in ``Going Dark'' argument, law enforcement has considerable struggles with digital
+evidence, even aside from encryption. Exacerbating the problem for justice system is the ``tech effect,'' which refers
+to jurors' increased expectations of digital evidence in cases where they suspect it exists \cite{shelton_study_2006}.
 
@@
 change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
-in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement but
-thwart \ac{masssurv} by remote powers subvert most of the ``Golden Age'' argument. \ac{EA} proposals that strive for
-these properties will be explored in \mysec{sec-ea-types}.
+in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
+thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
+in \mysec{sec-ea-types}.
 
@@
 Another common argument is that implementing \ac{EA}---or even researching it---inexorably moves tech policy down a
-slippery slope, in which government demands will never be satisfied. This argument can be used to prevent research by
-those who think judicious use of \ac{EA} tools could be positive. Going too far and too fast is certainly dangerous, but
-it is both cynical and misleading to declare the permanent destruction of privacy as the inevitable end of \ac{EA}
-research.
+slippery slope, in which government demands will never be satisfied. This argument can be used to prevent anyone from
+seeking a middle ground. Going too far and too fast is certainly dangerous, but it is both cynical and misleading to
+declare the permanent destruction of privacy as the inevitable end of \ac{EA} research.
 
@@
 The final two anti-\ac{EA} fallacies are examples of false dichotomies; they present the situation as if there are only
-two options. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading. One
-false dichotomy is that the only acceptable solutions are either perfect \ac{EA} or no \ac{EA}. This leaves no room for
-risk-based approaches, the typical security strategy outside the realm of cryptography. Another false dichotomy is that
-regulators must either allow strong cryptography or we will live in a \ac{masssurv} dystopia. This one carries hints of
-``slippery slope'' and appeal to emotion as well, framed as a choice between only two futures. Framing the situation in
-all or nothing terms reduces the chance of a successful collaborative debate.
+two options. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading.
+Holding either perfect \ac{EA} or no \ac{EA} at all as the only acceptable solutions is one false dichotomy. This leaves
+no room for risk-based approaches, the typical security strategy outside the realm of cryptography. Suggesting that
+regulators must either allow strong cryptography or we will live in a \ac{masssurv} dystopia is another false dichotomy.
+This leans towards ``slippery slope'' and appeal to emotion fallacies as well. Framing the situation in all or nothing
+terms reduces the chance of a successful collaborative debate.
 
-Some arguments used to support \ac{EA} rely on appeal to emotion. Politicians and law enforcement agents often appeal to
-pity for child abuse victims or fear of terrorists, drug dealers, and kidnappers to make their point. Public safety is a
-central issue in the debate. The technical community does take these concerns seriously \cite{schneier_scaring_2019}.
-These appeals become fallacious when they are used to manipulate the audience into supporting solutions incommensurate
-with the problems or to manipulate the audience into believing that the technical community does not care about these
-issues.
+Some arguments used to support \ac{EA} rely on appeal to emotion. Politicians and law enforcement agents often argue for
+\ac{EA} by appealing to pity for child abuse victims or appealing to fear of terrorists, drug dealers, and kidnappers.
+Public safety is a central issue in the debate. The technical community does take these concerns seriously
+\cite{schneier_scaring_2019}. These appeals become fallacious when they are used to manipulate the audience into
+supporting solutions incommensurate with the problems or to manipulate the audience into believing that the technical
+community does not care about these issues.
 
-Another argument used to subvert anti-EA arguments from experts goes along the lines of, since we put a man on the moon,
-we can surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The only similarity between the
-challenge of landing an astronaut on the moon and the challenge of building a cryptosystem with secure EA is that they
-are difficult. Once the political commitment was made, the Apollo program solved mostly tame problems and had enormous
+When experts argue against \ac{EA} due to security challenges, some respond that since we put a man on the moon, we can
+surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The only similarity between the challenge
+of landing an astronaut on the moon and the challenge of building a cryptosystem with secure EA is that they are
+difficult. Once the political commitment was made, the Apollo program solved mostly tame problems and had enormous
 government backing. Resolving the encryption debate has neither of these advantages. Many similar arguments comparing
@@
 Lastly, government officials often portray \ac{EA} as a problem that security experts simply have not researched enough.
-If experts just ``nerd harder,'' as some put it, they can find a solution \cite{schneier_2019}. This misses the point of
-research done up to this point. Encryption and \ac{EA} is a \ac{wicked-prob} for which tame, technical-only solutions
-will not work \cite{rozenshtein_wicked_2018}. Even on the technical side, experts have stated time and again that,
-depending how you look at it, the technical problem is solved---\ac{EA} is easy from a purely cryptographic point of
-view. The hard part is the security of the interfaces with the human users and administers of the system
-\cite{abelson_risks_1997} \cite{abelson_2015}. This may not be an area of strong academic research, but to frame that
-as the primary roadblock to progress utterly misses the point.
+If experts just ``nerd harder,'' as some put it, they can find a solution \cite{schneier_2019}. Concluding that
+encryption policy can be settled merely by more technical research misses the point of the research done up to this
+point. Encryption policy is a \ac{wicked-prob} for which tame, technical-only solutions will not work
+\cite{rozenshtein_wicked_2018}. Even on the technical side, research has concluded that, depending how you look at it,
+the technical problem is solved. \Ac{EA} is easy from a purely cryptographic point of view; the interfaces between the
+cryptographic and human portions of the system pose the real difficulty \cite{abelson_risks_1997} \cite{abelson_2015}.
+This may not be an area of strong academic inquiry, but to frame lack of technical research as the primary roadblock to
+encryption policy progress utterly misses the point.
 
@@
 
-Moving on from problem framing and meritless arguments, we next turn to actions that can be taken in response to the
-situation. \myfig{fig-arg-measures} divides potential solutions into the categories of current capabilities, legal
-measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the status quo and
-increasing investment in current programs. Legal measures require a change in the text or application of the law, but do
-not require fundamentally new technical capabilities. These include compelling passwords, which requires legal
-clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
-non-\ac{EA} approaches were first introduced in \mysec{sec-alt-approaches}.) At this level, \ac{EA} is analyzed in
-general. Specific EA approaches are addressed in the next map.
+Previous maps have framed the debate, presented leading arguments, and called-out fallacious arguments. The next map
+explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
+capabilities, legal measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the
+status quo and increasing investment in current programs. Legal measures require a change in the text or application of
+the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
+legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
+non-\ac{EA} approaches were first introduced in \mysec{sec-alt-approaches}.) In this section, \ac{EA} is compared to its
+alternatives. \ac{EA} variations are compared in the next map.
+
+It should be noted that the \ac{argmap} format works best when it displays the debate surrounding just one conclusion or
+proposal. This allows adequate room for elaboration on points and counterpoints. When a map includes multiple
+conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
+of brevity, this thesis compares \ac{EA}'s alternatives and variations in single maps. For a thorough discussion of both
+\ac{EA}'s alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report
+titled \ii{Decrypting the Encryption Debate: A Framework for Decision Makers} \cite{committee_decrypting_2018}.
 
@@
 
-Each measure is considered according to five metrics. These metrics represent a generalization of metrics from several
-sources, including two separate diverse committees gathered to analyze the encryption debate
+In \myfig{fig-arg-measures}, response measures are considered according to five metrics. These metrics represent are
+based on several sources, including two separate diverse committees gathered to analyze the encryption debate
 \cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
@@
 
-Beginning with current capabilities, the first choice is to do nothing. Some may conclude that the current state of
+Potential solutions are divided into the categories of current capabilities, legal measures, and exceptional access. The
+first option, in the category of current capabilities, is to do nothing. Some may conclude that the current state of
 affairs actually does represent an acceptable balance of interests. If technology and government agree on this approach,
 then \ac{the-cw2} can end much like the first. However, legislative action itself is a risk to be mitigated. If the
-technical community ``decides'' on this approach and government continues to reject it, we could be forced in an
-undesirable direction.
+technical community settles on this approach and government continues to reject it, the government could force its will
+and move encryption policy in a dangerous direction.
 
@@
 courts \cite{bittenbender_2019} \cite{sobel_2019} because of the apparent conflict with the Fifth Amendment right
-against self-incrimination. Even so, affordability is the strongest argument in its favor. Schneier and Kerr point out
-that this approach requires the password-holder to be available to investigators, the prosecution has to overcome
-several legal hurdles even if the basic approach is ruled Constitutionally legal, and even then the defendant may choose
-to withhold the password and be held in contempt rather than be charged for crime evidenced by the plaintext
-\cite{kerr_encryption_2017}.
+against self-incrimination. Affordability is the strongest argument in compelled password disclosure's favor. Schneier
+and Kerr point out several weaknesses with this approach: prosecution must overcome the aforementioned legal hurdles,
+the password-holder must be available to investigators, and the password-holder must choose to provide the password
+rather than be held in contempt of court \cite{kerr_encryption_2017}. Being found in contempt of court would often be
+preferable to conviction of the original crime.
 
@@
 security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
-This approach is already in use \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
-\cite{feuer_chapo_2019}. Because it is already used, deliberately choosing \ac{lawful-hacking} as the primary strategy
-for accessing plaintext would mean officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.
+This approach has already been used \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
+\cite{feuer_chapo_2019}. Because it is already used, using \ac{lawful-hacking} as the primary strategy would require
+officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.
 
-Despite its strengths \ac{lawful-hacking} is a sub-optimal compromise. Because it requires time and specialized skills,
+Despite its strengths, \ac{lawful-hacking} is a sub-optimal compromise. Because it requires time and specialized skills,
 it is not useful for small departments and commonplace crimes. It would also alter the relationship between law
@@
 The final proposal is \ac{EA} itself. \ac{EA} would provide a standardized method for law enforcement to reliably access
-\ac{plaintext}. It could serve as a compliment to \ac{lawful-hacking}by eliminating the incentive to conceal
-vulnerabilities because it would be unnecessary for \ac{EA}-compliant systems, which would presumably include mainstream
-technology. In that situation, \ac{EA} would combat common crime and \ac{lawful-hacking} would combat organized crime
-and terror groups. While \ac{EA}'s impact on civil liberties and \ac{masssurv} would depend on its implementation, any
-proposal would be costly to introduce and would threaten U.S. product competitiveness due to perceived insecurity. Risk
-to security is the largest weakness of \ac{EA}. Any proposal must overcome significant challenges to minimize risk. The
-next map analyzes this more closely.
+\ac{plaintext}. It could compliment \ac{lawful-hacking}by eliminating the incentive to conceal vulnerabilities, because
+hacking would be unnecessary for \ac{EA}-compliant systems. If used together, \ac{EA} would combat common crime and
+\ac{lawful-hacking} would combat organized crime and terror groups. \Ac{EA}'s impact on civil liberties and
+\ac{masssurv} would depend on its implementation. Any proposal would be costly to deploy and would threaten U.S. product
+competitiveness due to perceived insecurity. Increased security risk is the largest weakness of \ac{EA}. Any proposal
+must overcome significant challenges to minimize risk. The next map analyzes this more closely.
 
@@
 
-\myfig{fig-arg-classes} lists several classes of \ac{EA} and arguments in their favor and disfavor. The \ac{EA} classes
-applicable to both \ac{DIM} and \ac{DAR} include weak cryptography, trusted-party key escrow, and distributed key
-escrow. Using cryptographic puzzles and adding a ghost user apply to the case of \ac{DIM}, and device key escrow applies
-to \ac{DAR}.
+\myfig{fig-arg-classes} maps the arguments surrounding several classes of \ac{EA}. Weak cryptography, trusted-party key
+escrow, and distributed key escrow apply to both \ac{DAR} and \ac{DIM}. Device key escrow applies only to \ac{DAR}, and
+cryptographic puzzles and ghost users apply only to \ac{DIM}.
 
@@
 
-% TODO: move this up? Applies to previous map as well.
-It should be noted that the \ac{argmap} format works best when it displays the debate surrounding just one conclusion or
-proposal. This allows adequate room for elaboration on points and counterpoints. When a map includes multiple
-conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
-of brevity, this thesis compares several \ac{EA} variations in a single map. For an thorough discussion of \ac{EA}'s
-alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report titled
-\ii{Decrypting the Encryption Debate: A Framework for Decision Makers} \cite{committee_decrypting_2018}.
-
-The first \ac{EA} class is to simply use weak cryptographic ciphers or short encryption keys. This was the approach
-taken before \ac{the-cw1}, when strong cryptography was not readily available and subjected to export controls. No party
-seriously supports this approach today, as it contradicts security on many levels.
+The first type of \ac{EA} is simply use of weak cryptographic ciphers or short encryption keys. This was the approach
+taken before \ac{the-cw1}, when strong cryptography was not readily available and was subjected to export controls. No
+party seriously supports this approach today, as far too many attackers would have the capability to abuse it.
 
 Trusted-party key escrow relies on an entity or a small group of entities to store a key or key recovery information.
-There are a wide variety of implementations \cite{denning_taxonomy_1996}. This class of \ac{EA} has a crushing weakness:
-it centralizes extremely sensitive information. Some argue software companies are capable of protecting such sensitive
-data, citing their current practice of storing code-signing keys in hardware security modules \cite{ozzie_2018}.
-However, others point out that such keys are quite rarely used (especially when compared to the likely frequency of
-\ac{EA} requests), and yet they still leak \cite{green_2018}. Once the escrowed keys were stolen, as they inevitably
-would be, every device or message encrypted with them would be at risk. This is an example of a case when the
-\ac{argmap} format doesn't do an argument justice---the argument against any centralized access mechanism is extremely
-powerful.
+There are a wide variety of implementations from the days of \ac{the-cw1} \cite{denning_taxonomy_1996}. This type of
+\ac{EA} has a crushing weakness: it centralizes extremely sensitive information, which results in high risk. Some argue
+software companies are capable of protecting such sensitive data, citing their current practice of storing code-signing
+keys in \acp{HSM} \cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be
+updated and accessed frequently. Despite code-signing keys being rarely used and protected in \acp{HSM}, they they still
+leak \cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This
+is a case where the \ac{argmap} format fails to reflect an argument's strength---the argument against any centralized
+access mechanism is powerful.
 
@@
 achieved by using distributed systems to make key extraction difficult and publicly observable \cite{phan_key_2017}
-\cite{servan_schreiber_jje_2020}. One advantage of distribution is that this approach is that it is more likely to fail
-by making data unrecoverable than by making unintended data vulnerable, though more research into this area is needed.
-Unfortunately, these systems would likely be costly to deploy and difficult to update due to their complexity.
+\cite{servan_schreiber_jje_2020}. One advantage of distribution over centralization is that it is more likely to fail
+safely. System failure would more likely result in \ii{no one} to be able to recover plaintext instead of \ii{anyone}.
+More research in this area is needed. Unfortunately, these systems would likely be costly to deploy and difficult to
+update due to their complexity.
 
-One class of \ac{DIM} \ac{EA} involves using strong cryptography, but including cryptographic puzzle pieces as message
-metadata. These pieces are designed to reveal the encryption key after the surveiller spends considerable computational
-power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass surveillance, but it also
-enables arbitrary surveillance by anyone with enough computing power. Using this class of \ac{DIM} \ac{EA} in common
-cases would be impractical due to its high cost. However, it would be less effective in high-profile cases due to the
-ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, the calibration of puzzle difficulty
-is dependent on the state of computing. Since opponents can record data and decrypt it later when computational
-capabilities have changed, the data becomes steadily less secure over time.
+One type of \ac{DIM} \ac{EA} involves using strong cryptography, but including cryptographic ``puzzle pieces'' in
+message metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
+considerable computational power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass
+surveillance, but it also enables arbitrary surveillance by anyone with enough computing power. Using this type of
+\ac{DIM} \ac{EA} in common cases would be impractical due to its high cost. However, it would be less effective in
+high-profile cases due to the ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, puzzle
+difficulty would be calibrated based on ever-changing standards of computational capabilities. Since opponents can
+record data and decrypt it later when capabilities have changed, the data becomes steadily less secure over time.
 
-Another \ac{DIM} proposal is to ``conference call'' authorities in as ghost participants in the conversation. This was
-notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson \cite{levy_robinson_2018}. Their writeup begins
-with some core ideas of this thesis:
+Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
+a silent listener on a conference call. This was notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson
+\cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:
 
@@
 
-Although this proposal was raised with good intentions, it is nevertheless problematic. The ghost user proposal prides
-itself on not compromising strong encryption. Although this is a good property, compromising a messaging app's
-authorization protocol is as dangerous as compromising its encryption \cite{callas_1_2019}. Fundamentally, this proposal
-suffers from the same weakness as trusted-party key escrow---it introduces an authorization vulnerability in the
-messaging service provider's platform, resulting in dangerous centralized access capability \cite{schneier_ghost_2019}.
-It has the same nature, if a different form, as rejected key escrow proposals.
+Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
+proposal consider the uncompromised use of strong encryption as its great strength. Although this is a good property,
+compromising a messaging app's authorization protocol is as dangerous as compromising its encryption
+\cite{callas_1_2019}. Fundamentally, this proposal suffers from the same weakness as trusted-party key escrow---it
+introduces an authorization vulnerability in the messaging service provider's platform, resulting in dangerous
+centralized access capability \cite{schneier_ghost_2019}. It has a different form, but same nature as rejected key
+escrow proposals.
 
-Finally, one \ac{DAR}-specific \ac{EA} class is device key escrow. In this scheme, key recovery material exists on the
-device itself. While not necessary to classify as device key escrow, most of these proposals include physical possession
-requirements and additionally ``brick'' the device or make it absolutely clear that \ac{EA} has been performed
-\cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency to the device
-user. This approach relies on secure hardware and a method of authorizing the recovery request to the device itself. The
-authentication step can be paired with other \ac{EA} strategies, such as engaging with trusted-parties or distributed
-systems. This is where the weakness of the device escrow is revealed. Secure hardware enclaves, though improving, are
-not totally reliable. Therefore, authentication to the hardware device is susceptible to the same attacks as traditional
+Finally, one type of \ac{DAR} \ac{EA} is device key escrow. In this scheme, key recovery material exists on the device
+itself. Most of these proposals include physical possession requirements and make it clear when \ac{EA} has been
+performed \cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency to
+the device user. This approach relies on secure hardware and a device unlock authentication process. The authentication
+process can be paired with other \ac{EA} strategies, such as storing keys with trusted-parties or distributed systems.
+This is where the weakness of the device escrow is revealed. First, secure hardware enclaves, though improving, are not
+totally reliable. Second, authentication to the hardware device is susceptible to the same attacks as traditional
 escrow \ac{EA}.
@@
 Device key escrow's greatest weaknesses, however, are still signs of progress. It is true that it relies on a device's
-secure enclave, but devices already do---it is the same hardware device that manages device unlock and encryption
+secure enclave, but this is not a new risk. The same hardware device already manages device unlock and encryption
 functionality in the first place. Additional functionality increases the attack surface, but it does not add a
 completely new threat vector. It is also true that the process is subject to the same problems as traditional key escrow
-after the device has been obtained, but that is precisely the progress that has been made---\ii{after the device has
-been obtained}. Advances in secure hardware design have meaningfully changed the \ac{EA} risk profile.
+after the device has been obtained, but this is still an improvement. Device key escrow requires the device to be
+obtained and cannot be used surreptitiously. Thus, advances in secure hardware design have meaningfully changed the
+\ac{EA} risk profile.
 
 In 2019, Carnegie Mellon assembled an ideologically diverse group of policy and security experts to engage in the kind
-of cross-disciplinary research. After describing every technical branch of the situation, their report goes into detail
-on why \ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated adversaries) focusing on
-\ac{DAR} in mobile phones using device key escrow requiring physical access is the most tractable technical niche to
-consider \cite{group_2019}. By now it should be clear why: traditional investigation mixed with a dose of lawful hacking
-is best suited for advanced opponents, and device key escrow represents actual progress as a path towards lower-risk
-\ac{EA} for typical cases.
+of cross-disciplinary research. After describing every technical branch of the situation, their report identified
+\ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated adversaries) focusing on \ac{DAR} in
+mobile phones using device key escrow requiring physical access as the most tractable technical niche to consider
+\cite{group_2019}. By now it should be clear why: traditional investigation combined with lawful hacking is best suited
+for advanced opponents, while device key escrow represents actual progress towards low-risk \ac{EA} for common crimes.
 
-A strong argument against device key escrow is that device encryption, particularly in mobile phones, may actually pose
-the smallest impediment to investigation of any encryption technology. The previously cited report into law enforcement
-agencies' ability to access mobile phones clearly demonstrates the current state of affairs: device manufacturers are
-losing the exploit arms race, private companies are commodifying the \ac{lawful-hacking} approach for agencies large and
-small, and the lack of regulatory strategy has led to pervasive data collection with little oversight
-\cite{koepke_2020}. This fact does not undermine the worthiness of research into device-oriented \ac{EA}. First of all,
-it underscores the lessons of \mychap{chap-policy}---inaction is an action when facing \acp{wicked-prob}, and in this
-case, inaction has led to an undesirable position. Second of all, the current circumstances where device manufacturers
-can't keep attackers out is temporary; when law enforcement faces real difficulty again, this issue will immediately
-resurface. Law enforcement's anticipation of this is likely why they continue to push for \ac{EA} in spite of present
-circumstances. Third of all, regardless of criminal abuse, formalized capability and process can reduce governmental
-abuse, which is the clearer threat to privacy and civil liberties.
+The fact that device encryption, particularly in mobile phones, may actually pose the smallest impediment to
+investigation of any encryption technology is a strong argument against device key escrow. The ``Mass Extraction''
+report clearly describes the current state of affairs: device manufacturers are losing the exploit arms race, private
+companies are commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory
+strategy is leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs
+underscores the lessons of \mychap{chap-policy}---inaction is an action when facing \acp{wicked-prob}, and in this case,
+inaction has led to an undesirable position. This situation does not undermine the worthiness of research into
+device-oriented \ac{EA}. First, the current advantage hackers hold over device manufacturers is temporary; when law
+enforcement faces difficulty accessing devices again, this issue will immediately resurface. This is likely why law
+enforcement continues to push for \ac{EA} despite present circumstances. Second, formalized capability and process may
+reduce government abuse, which threatens privacy and civil liberties more significantly than criminal abuse.
 
-Despite device key escrow's tractability as a technical niche of \ac{EA}, it still has a long way to go. There are few
-proposals, what few there are high level, and questions of scaling and administration have not been addressed. However,
-as \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific
-proposals has the opportunity to identify mutually agreeable solutions---or at least refine the problem understanding
-enough that the next round of debate can be better informed.
+Despite device key escrow's tractability as a technical niche of \ac{EA}, it is not ready for deployment. There are few
+proposals, the few that exist lack detail, and matters of scaling and administration are unresolved. However, as
+\ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific proposals
+may help to identify mutually agreeable solutions---or at least to refine the debate so that the next round of inquiry
+can be better informed.
 
-It is this task that we turn to next. The following chapter analyzes one particular device-escrow proposal in detail,
-Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
+The next chapter focuses on one such proposal. It defines a threat model and analyzes Stefan Savage's 2018 device escrow
+proposal, titled, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
 \cite{savage_lawful_2018}.

### 2020-12-16 Thesis-Chapter6-ThreatModel.tex
Editorial improvements

@@
 
-\modelitem{Transparency}{The system should be auditable by the public to enable verification that its power is being
+\modelitem{Transparency}{The system should be auditable by the public to allow for verification that its power is being
 used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}
@@
 \modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are constantly looking for ways to escalate
-privilege and tamper with or steal information. They have world-class expertise and computing power, which they can
-apply to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence
+their privilege and tamper with or steal information. They have world-class expertise and computing power, which they
+can apply to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence
 agencies, as opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S.,
@@
 
-Hawkish lawmakers threaten the security of a cryptosystem (using \ac{EA} or not) because they may judge the system as
-too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something weaker in its place.}
+Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{EA} or not) because they may judge the
+system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something weaker in its
+place.}
 

### 2020-12-17 Thesis-Chapter2-Background.tex
Editorial improvements

@@
 as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
-these countries, the member's encryption policies are relevant to one another.
+these countries, the member's encryption policies are relevant to one another.\\
 

### 2020-12-17 Thesis-Chapter5-Arguments.tex
Editorial improvements

@@
 to compromise the cryptographic foundations of encryption; however, this is seen as too risky today.
-\mysec{sec-tech-approaches} introduced several alternative technical approaches to \ac{EA}, but experts still make clear
-that current systems are incapable of securely providing the level of access \acl{LE} wants \cite{abelson_2015}
-\cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs, and the field of
+\mysec{sec-tech-approaches} introduced several alternative technical approaches to \ac{EA}, but experts still emphasize
+that current systems cannot securely provide the level of access \acl{LE} asks for \cite{abelson_2015}
+\cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs. The field of
 cybersecurity is still equal parts art and science; mandated \ac{EA} would put a heavy burden on a field still finding
 its legs.
-% TODO: too colloquial?
 
@@
 password recovery, or administrator access in a business setting. Typically these applications have other solutions,
-however, and unnecessary requirements would weaken any \ac{EA} system.
+however, and unnecessary requirements would weaken \ac{EA} systems.
 
@@
 
-Much of the disconnect between government and the technical community is disagreement over whether law enforcement has
-\ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is well-represented by proponent
-former \ac{FBI} Director James Comey in his 2014 speech titled ``Going Dark'' \cite{comey_2014} (though he did not coin
-the phrase \cite{swire_encryption_2011}). The argument is presented in \myfig{fig-arg-going-dark}.
+Much of the disconnect between government and the technical community is the result of disagreement over whether law
+enforcement has \ii{too little} or \ii{too much} access to data. The argument for \ii{too little} is well-represented by
+proponent former \ac{FBI} Director James Comey in his 2014 speech titled ``Going Dark'' \cite{comey_2014} (though he did
+not coin the phrase \cite{swire_encryption_2011}). The argument is presented in \myfig{fig-arg-going-dark}.
 
@@
 Widespread Power of U.S. Law Enforcement to Search Mobile Phones'' revealed that current hacking tools are stronger than
-mobile device encryption's defenses, and law enforcement is actually able to extract data from nearly every mobile phone
-on the market \cite{koepke_2020}. Without conclusive data showing the extent that encryption hinders investigations, the
-``Going Dark'' argument weakens.
+the defenses provided by mobile device encryption and that law enforcement is actually able to extract data from nearly
+every mobile phone on the market \cite{koepke_2020}. The lack of conclusive data showing the extent that encryption
+hinders investigations is a weakness of the ``Going Dark'' argument.
 
-Lack of evidence for an argument does not prove that it is incorrect. Because law enforcement has the data, the burden
-of proof lies with them to show that encryption is causing investigatory failure. However, in most cases, it is
-impossible to know what would have happened without the interference of strong encryption. Inability to gather data from
-controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable, another way to
-gather data is by observing naturally-occurring conditions in which everything but the variable of interest remain are
-constant. These conditions exist for encryption in messaging application ecosystem. Messaging applications WhatsApp,
-Facebook Messenger, and Telegram all have large user bases and similar features. Among these services, only Facebook
-Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market share as measured by
-active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports \cite{keller_internet_2019}. This
-indicates that encryption does indeed affect investigations.
+However, lack of evidence for an argument does not prove that it is incorrect. Because law enforcement has the data, the
+burden of proof is on them---they must show that encryption is causing investigatory failure. Unfortunately, in most
+cases, it is impossible to know what would have happened without the interference of strong encryption. Inability to
+gather data from controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable,
+another way to gather data is by observing naturally-occurring conditions in which everything but the variable of
+interest remain are constant. These conditions exist for encryption in messaging application ecosystem. Messaging
+applications WhatsApp, Facebook Messenger, and Telegram all have large user bases and similar features. Among these
+services, only Facebook Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market
+share as measured by active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
+\cite{keller_internet_2019}. This indicates that encryption does indeed affect investigations.
 
@@
 Encryption is not the only challenge that law enforcement faces with digital evidence. A 2018 study found that
-identifying service providers and getting data from them using existing mechanisms were even greater impediments
+identifying service providers and getting data from them using existing mechanisms were even greater challenges
 \cite{carter_2018}. The report, titled ``Low-Hanging Fruit,'' also lists resource limitations and poor training as top
 issues. Indeed, many current initiatives are underfunded. The \ii{New York Times}' 2019 investigation into \ac{CSAM}
-found chronic lack of commitment from government initiatives---major legislation passed in 2008, but has been funded at
-only 50\% of authorized levels; a commissioned Justice Department task force has produced only two of five biennial
-reports; and a senior executive position within the Justice Department to target \ac{CSAM} was never created
-\cite{keller_internet_2019}. Federal entities aiming to improve training and coordination between law enforcement and
+uncovered a chronic lack of commitment from government initiatives \cite{keller_internet_2019}. Major legislation passed
+in 2008, but has been funded at only 50\% of authorized levels; a commissioned Justice Department task force has
+produced only two of five biennial reports; and a senior executive position within the Justice Department to target
+\ac{CSAM} was never created. Federal entities aiming to improve training and coordination between law enforcement and
 industry are similarly underfunded \cite{carter_2018}.
@@
 
-After establishing the danger of surveillance, the argument notes how technology has increased the government's
-surveillance abilities. This statement as strong evidence. In Comey's ``Going Dark'' speech, it was digital evidence
-which would not have existed twenty years ago that cracked each case \cite{comey_2014}. Encryption is one of the only
-privacy-saving technologies to emerge alongside the boom in privacy-eroding technologies. Opting out of the
+After establishing the danger of surveillance, the argument notes that technology has increased the government's
+surveillance abilities. This statement is supported by strong evidence. In Comey's ``Going Dark'' speech, it was digital
+evidence which would not have existed twenty years ago that cracked each case \cite{comey_2014}. Encryption is one of
+the only privacy-saving technologies to emerge alongside the boom in privacy-eroding technologies. Opting out of the
 technological boom is possible for some, but is not a realistic option for society as a whole. Due to the dangers of
@@
 resource-strapped local, state, and federal law enforcement. It is obvious that the amount of digital evidence is
-increasing. However, as seen in ``Going Dark'' argument, law enforcement has considerable struggles with digital
-evidence, even aside from encryption. Exacerbating the problem for justice system is the ``tech effect,'' which refers
-to jurors' increased expectations of digital evidence in cases where they suspect it exists \cite{shelton_study_2006}.
+increasing. However, as seen in ``Going Dark'' argument, law enforcement often struggles with digital evidence for
+reasons unrelated to encryption. Exacerbating the problem for justice system is the ``tech effect,'' which refers to
+jurors' increased expectations of digital evidence in cases where they suspect it exists \cite{shelton_study_2006}.
 
@@
 Another common argument is that implementing \ac{EA}---or even researching it---inexorably moves tech policy down a
-slippery slope, in which government demands will never be satisfied. This argument can be used to prevent anyone from
+slippery slope in which government demands will never be satisfied. This argument can be used to discourage those
 seeking a middle ground. Going too far and too fast is certainly dangerous, but it is both cynical and misleading to
@@
 two options. \Acp{wicked-prob} have unlimited sets of potential responses, so reducing them to two is misleading.
-Holding either perfect \ac{EA} or no \ac{EA} at all as the only acceptable solutions is one false dichotomy. This leaves
-no room for risk-based approaches, the typical security strategy outside the realm of cryptography. Suggesting that
-regulators must either allow strong cryptography or we will live in a \ac{masssurv} dystopia is another false dichotomy.
-This leans towards ``slippery slope'' and appeal to emotion fallacies as well. Framing the situation in all or nothing
-terms reduces the chance of a successful collaborative debate.
+Arguers create a false dichotomy when they present either perfect \ac{EA} or no \ac{EA} at all as the only acceptable
+solutions. This leaves no room for risk-based approaches, the typical security strategy outside the realm of
+cryptography. Opponents of \ac{EA} create another false dichotomy when they suggest that regulators must allow strong
+cryptography or we will live in a \ac{masssurv} dystopia. This argument has elements of the ``slippery slope'' and
+appeal to emotion fallacies as well. Framing the situation in all-or-nothing terms reduces the chance of a successful
+collaborative debate.
 
@@
 surely create secure \ac{EA} \cite{cushing_moon_2018}. This is a weak analogy. The only similarity between the challenge
-of landing an astronaut on the moon and the challenge of building a cryptosystem with secure EA is that they are
+of landing an astronaut on the moon and the challenge of building a cryptosystem with secure EA is that they are both
 difficult. Once the political commitment was made, the Apollo program solved mostly tame problems and had enormous
@@
 Lastly, government officials often portray \ac{EA} as a problem that security experts simply have not researched enough.
-If experts just ``nerd harder,'' as some put it, they can find a solution \cite{schneier_2019}. Concluding that
-encryption policy can be settled merely by more technical research misses the point of the research done up to this
-point. Encryption policy is a \ac{wicked-prob} for which tame, technical-only solutions will not work
-\cite{rozenshtein_wicked_2018}. Even on the technical side, research has concluded that, depending how you look at it,
-the technical problem is solved. \Ac{EA} is easy from a purely cryptographic point of view; the interfaces between the
-cryptographic and human portions of the system pose the real difficulty \cite{abelson_risks_1997} \cite{abelson_2015}.
-This may not be an area of strong academic inquiry, but to frame lack of technical research as the primary roadblock to
-encryption policy progress utterly misses the point.
+If experts just ``nerd harder,'' as some put it, they will be able to find a solution \cite{schneier_2019}. However,
+this notion ignores the conclusions of research done thus far. Encryption policy is a \ac{wicked-prob} for which tame,
+technical-only solutions will not work \cite{rozenshtein_wicked_2018}. Even on the technical side, research has
+concluded that \ac{EA} is easy from a purely cryptographic point of view. The interfaces between the cryptographic and
+human portions of the system pose the real difficulty \cite{abelson_risks_1997} \cite{abelson_2015}. This may not be an
+area of strong academic inquiry, but to frame lack of technical research as the primary roadblock to encryption policy
+progress misses the point.
 
@@
 
-Previous maps have framed the debate, presented leading arguments, and called-out fallacious arguments. The next map
+The last four maps framed the debate, presented leading arguments, and identified fallacious arguments. The next map
 explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
@@
 
-In \myfig{fig-arg-measures}, response measures are considered according to five metrics. These metrics represent are
-based on several sources, including two separate diverse committees gathered to analyze the encryption debate
+In \myfig{fig-arg-measures}, response measures are considered according to five metrics. These metrics are based on
+several sources, including two separate diverse committees gathered to analyze the encryption debate
 \cite{committee_decrypting_2018} \cite{group_2019} \cite{varia_2018}. They also align closely to the four central issues
@@
 courts \cite{bittenbender_2019} \cite{sobel_2019} because of the apparent conflict with the Fifth Amendment right
-against self-incrimination. Affordability is the strongest argument in compelled password disclosure's favor. Schneier
+against self-incrimination. Affordability is the strongest argument in favor of compelled password disclosure. Schneier
 and Kerr point out several weaknesses with this approach: prosecution must overcome the aforementioned legal hurdles,
-the password-holder must be available to investigators, and the password-holder must choose to provide the password
-rather than be held in contempt of court \cite{kerr_encryption_2017}. Being found in contempt of court would often be
-preferable to conviction of the original crime.
+the password-holder must be available to investigators, and the password-holder must elect to provide the password
+\cite{kerr_encryption_2017}. Failing to provide the password may result in being found in contempt of court, but this
+would often be preferable to conviction of the original crime.
 
@@
 The final proposal is \ac{EA} itself. \ac{EA} would provide a standardized method for law enforcement to reliably access
-\ac{plaintext}. It could compliment \ac{lawful-hacking}by eliminating the incentive to conceal vulnerabilities, because
-hacking would be unnecessary for \ac{EA}-compliant systems. If used together, \ac{EA} would combat common crime and
+\ac{plaintext}. It could compliment \ac{lawful-hacking} by eliminating the incentive to conceal vulnerabilities because
+hacking would be unnecessary for \ac{EA}-compliant systems. When used together, \ac{EA} would combat common crime and
 \ac{lawful-hacking} would combat organized crime and terror groups. \Ac{EA}'s impact on civil liberties and
 \ac{masssurv} would depend on its implementation. Any proposal would be costly to deploy and would threaten U.S. product
-competitiveness due to perceived insecurity. Increased security risk is the largest weakness of \ac{EA}. Any proposal
-must overcome significant challenges to minimize risk. The next map analyzes this more closely.
+competitiveness due to perceived insecurity. Increased security risk is \ac{EA}'s largest weakness. Any proposal must
+overcome significant challenges to minimize risk. The next map analyzes this more closely.
 
@@
 
-The first type of \ac{EA} is simply use of weak cryptographic ciphers or short encryption keys. This was the approach
-taken before \ac{the-cw1}, when strong cryptography was not readily available and was subjected to export controls. No
-party seriously supports this approach today, as far too many attackers would have the capability to abuse it.
+The first type of \ac{EA} is simply the use of weak cryptographic ciphers or short encryption keys. This was the
+approach taken before \ac{the-cw1}, when strong cryptography was not readily available and was subjected to export
+controls. No party seriously supports this approach today, as far too many attackers would have the capability to abuse
+it.
 
 Trusted-party key escrow relies on an entity or a small group of entities to store a key or key recovery information.
-There are a wide variety of implementations from the days of \ac{the-cw1} \cite{denning_taxonomy_1996}. This type of
-\ac{EA} has a crushing weakness: it centralizes extremely sensitive information, which results in high risk. Some argue
-software companies are capable of protecting such sensitive data, citing their current practice of storing code-signing
-keys in \acp{HSM} \cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be
-updated and accessed frequently. Despite code-signing keys being rarely used and protected in \acp{HSM}, they they still
-leak \cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This
-is a case where the \ac{argmap} format fails to reflect an argument's strength---the argument against any centralized
-access mechanism is powerful.
+There were a wide variety of trusted-party key escrow implementations created during \ac{the-cw1}
+\cite{denning_taxonomy_1996}. This type of \ac{EA} has a crushing weakness: it centralizes extremely sensitive
+information, which heightens the risk of a disastrous data breach. Some argue software companies are capable of
+protecting such sensitive data, citing their current practice of storing code-signing keys in \acp{HSM}
+\cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be updated and accessed
+frequently. Although code-signing keys are rarely used and are protected in \acp{HSM}, they still leak
+\cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This is an
+example of the \ac{argmap}'s sometimes imbalanced portrayal of an argument's strength---in this case, the map
+underrepresents the effectiveness of the security-based argument against centralized access mechanisms.
 
@@
 
-One type of \ac{DIM} \ac{EA} involves using strong cryptography, but including cryptographic ``puzzle pieces'' in
+One type of \ac{DIM} \ac{EA} involves uses strong cryptography but includes extra cryptographic ``puzzle pieces'' in
 message metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
@@
 high-profile cases due to the ability of sophisticated adversaries to employ \ac{EA} workarounds. Additionally, puzzle
-difficulty would be calibrated based on ever-changing standards of computational capabilities. Since opponents can
-record data and decrypt it later when capabilities have changed, the data becomes steadily less secure over time.
+difficulty would be calibrated based on the changing standards of computational capability. Since opponents can record
+data and decrypt it later when capabilities have changed, the data becomes steadily less secure over time.
 
@@
 Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
-proposal consider the uncompromised use of strong encryption as its great strength. Although this is a good property,
-compromising a messaging app's authorization protocol is as dangerous as compromising its encryption
+proposal consider the uncompromised use of strong encryption as the proposal's great strength. Although this is a good
+property, compromising a messaging app's authorization protocol is as dangerous as compromising its encryption
 \cite{callas_1_2019}. Fundamentally, this proposal suffers from the same weakness as trusted-party key escrow---it
@@
 Finally, one type of \ac{DAR} \ac{EA} is device key escrow. In this scheme, key recovery material exists on the device
-itself. Most of these proposals include physical possession requirements and make it clear when \ac{EA} has been
-performed \cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency to
-the device user. This approach relies on secure hardware and a device unlock authentication process. The authentication
+itself. Most device key escrow proposals include physical possession requirements and make it clear when \ac{EA} has
+been performed \cite{savage_lawful_2018} \cite{ozzie_2018}. Such features preclude \ac{masssurv} and ensure transparency
+to the device user. This approach relies on secure hardware and a device unlock authorization process. The authorization
 process can be paired with other \ac{EA} strategies, such as storing keys with trusted-parties or distributed systems.
 This is where the weakness of the device escrow is revealed. First, secure hardware enclaves, though improving, are not
-totally reliable. Second, authentication to the hardware device is susceptible to the same attacks as traditional
-escrow \ac{EA}.
+totally reliable. Second, authorization to the hardware device is susceptible to the same attacks as traditional escrow
+\ac{EA}.
 
-Device key escrow's greatest weaknesses, however, are still signs of progress. It is true that it relies on a device's
-secure enclave, but this is not a new risk. The same hardware device already manages device unlock and encryption
-functionality in the first place. Additional functionality increases the attack surface, but it does not add a
-completely new threat vector. It is also true that the process is subject to the same problems as traditional key escrow
-after the device has been obtained, but this is still an improvement. Device key escrow requires the device to be
-obtained and cannot be used surreptitiously. Thus, advances in secure hardware design have meaningfully changed the
-\ac{EA} risk profile.
+Device key escrow's greatest weaknesses, however, are still signs of progress. While it does rely on a device's secure
+enclave, this is not a new risk. The same hardware device already manages device unlock and encryption functionality.
+Additional functionality increases the attack surface, but it does not add a completely new threat vector. While the
+authorization process is subject to the same problems as traditional key escrow after the device has been obtained, this
+is still an improvement over previous designs. Device key escrow requires the device to be obtained and cannot be used
+surreptitiously. Thus, advances in secure hardware design have meaningfully changed the \ac{EA} risk profile.
 
 In 2019, Carnegie Mellon assembled an ideologically diverse group of policy and security experts to engage in the kind
-of cross-disciplinary research. After describing every technical branch of the situation, their report identified
-\ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated adversaries) focusing on \ac{DAR} in
-mobile phones using device key escrow requiring physical access as the most tractable technical niche to consider
-\cite{group_2019}. By now it should be clear why: traditional investigation combined with lawful hacking is best suited
-for advanced opponents, while device key escrow represents actual progress towards low-risk \ac{EA} for common crimes.
+of cross-disciplinary research. After describing every technical branch of the situation, their report identified the
+most promising branch: \ac{EA} for domestic law enforcement (i.e., common cases instead of sophisticated adversaries)
+focusing on \ac{DAR} in mobile phones using device key escrow requiring physical access \cite{group_2019}. This is
+because, as explained above, traditional investigation combined with lawful hacking is best suited for advanced
+opponents, while device key escrow represents actual progress towards low-risk \ac{EA} for common crimes.
 
-The fact that device encryption, particularly in mobile phones, may actually pose the smallest impediment to
-investigation of any encryption technology is a strong argument against device key escrow. The ``Mass Extraction''
-report clearly describes the current state of affairs: device manufacturers are losing the exploit arms race, private
-companies are commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory
-strategy is leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs
-underscores the lessons of \mychap{chap-policy}---inaction is an action when facing \acp{wicked-prob}, and in this case,
-inaction has led to an undesirable position. This situation does not undermine the worthiness of research into
-device-oriented \ac{EA}. First, the current advantage hackers hold over device manufacturers is temporary; when law
-enforcement faces difficulty accessing devices again, this issue will immediately resurface. This is likely why law
-enforcement continues to push for \ac{EA} despite present circumstances. Second, formalized capability and process may
-reduce government abuse, which threatens privacy and civil liberties more significantly than criminal abuse.
+Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
+encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
+describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
+commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
+leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
+lessons of \mychap{chap-policy}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has
+led to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This
+is because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
+accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
+\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
+privacy and civil liberties more significantly than criminal abuse.
 
 Despite device key escrow's tractability as a technical niche of \ac{EA}, it is not ready for deployment. There are few
-proposals, the few that exist lack detail, and matters of scaling and administration are unresolved. However, as
+proposals, the few that exist lack detail, and the matters of scaling and administration are unresolved. However, as
 \ac{incrementalism}'s Lindblom and even \ac{GCHQ}'s Levy and Robinson have pointed out, focusing on specific proposals
@@
 
-The next chapter focuses on one such proposal. It defines a threat model and analyzes Stefan Savage's 2018 device escrow
-proposal, titled, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
+The next chapter focuses on one specific device key escrow proposal. It defines a threat model and analyzes Stefan
+Savage's 2018 proposal, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
 \cite{savage_lawful_2018}.

### 2020-12-17 Thesis-Chapter6-ThreatModel.tex
Editorial improvements

@@
 proposal is specific to \acl{DAR} for mobile devices. Due to the length of its title, this proposal will be referred to
-by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by some legal framework regulating its
-application and use, though it is not the focus of this analysis, and nothing is assumed about it in particular. Note
-that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
-emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
-\ac{EA} proposal, but use the term in the spirit of their paper.
+by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by a legal framework regulating its
+application and use, though the legal framework is not the focus of this analysis.
+
+% Note that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
+% emphasize the role of legal process and dispel the negative stigma that accompanies \ac{EA}. I refer to \ldawmsr as an
+% \ac{EA} proposal, but use \ac{EA} in ... [that spirit].
 
@@
 
-``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The first step to
-answering those questions is to establish goals for the former and a list of threats in anticipation of the latter. The
-\ldawmsr proposal describes a threat model against which it was designed. This model takes cues from that one but is
-intended to be more generally applicable for any \ac{DAR} \ac{EA} proposal.
+``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The process of
+answering these questions begins with setting goals, identifying threat actors, and declaring scope. The \ldawmsr
+proposal describes a threat model against which the proposal was designed. This threat model takes cues from the
+\ldawmsr model, but is intended to be more generally applicable for any \ac{DAR} \ac{EA} proposal.
 
@@
 
-The goals of an \ac{EA} cryptosystem are taken from the metrics of \mychap{chap-arguments}. Their justification stems
-from the preceding arguments. Each goal is listed according to priority, though no goal has absolute priority over the
+The goals of an \ac{EA} cryptosystem are taken from the five metrics used to judge \ac{EA} alternatives and variations
+in \mychap{chap-arguments}. Each goal is listed according to priority, though no goal has absolute priority over the
 others.
@@
 
-\modelitem{Security}{The system should minimize security risk with reference to the current state of the art. Risk is
-considered with respect to both individual devices and the entire class of devices.}
+\modelitem{Security}{The system should minimize security risk relative to the current baseline. Risk is considered with
+respect to both individual devices and the entire class of devices.}
 
 \modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
-authoritarian regimes. This includes technologically enforcing the fundamental right to privacy but upon warranted
-search pursuant to the rule of law.}
+authoritarian regimes. This includes technologically enforcing the fundamental right to privacy outside the context of
+warranted search under the rule of law.}
 
 \modelitem{Transparency}{The system should be auditable by the public to allow for verification that its power is being
-used responsibly. In the case of \ac{DAR}, the subject of the search should be aware of the access.}
+used responsibly. In the case of \ac{DAR}, access should be apparent to the data owner.}
 
@@
 distribution of costs associated with its development and administration. Even though this is a security-oriented threat
-model, architectural decisions determine the system's economics, and economics determines the system's ultimate
-feasibility.}
+model, a system's design affects its economics, and its economics determines its ultimate feasibility.}
 
@@
 
-\ac{EA} threat models include the full cast of traditional threat actors.
+\ac{EA} threat models include the full slate of traditional threat actors.
 
@@
 \modelitem{Criminals: ordinary, organized, cyber}{Criminals threaten the \ac{EA} system by using it to access plaintext
-without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical
-access but only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but
-considerably more advanced technical capabilities, exerting full control over all interfaces. Cyber criminals have a
-wide spectrum of motivations and capabilities. Their significant contribution to the threat landscape is that they will
-target the backend systems that operate the \ac{EA} system in order to steal and likely sell whatever access
-capabilities they can find.}
+without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical access but
+only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but considerably
+more advanced technical capabilities, including full control over all interfaces. Cyber criminals have a wide variety of
+of motivations and capabilities. They will target the backend systems that operate the \ac{EA} system in order to abuse
+or sell whatever access capabilities they acquire.}
 
-\modelitem{Insider threat}{The insider threat involves someone within the \ac{EA} recovery process, perhaps at the law
-enforcement agency or a vendor, that seeks to abuse the system for their own purposes. They may have the same
-motivations and skills of any other criminal category, but they have the advantage of privileged access at some stage in
-the process.}
+\modelitem{Insider threat}{Insider threats involve attackers with roles inside the \ac{EA} recovery process who seek to
+abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
+provider. They may have the same motivations and skills of any other criminal category, but they have the advantage of
+privileged access at some stage in the process.}
 
@@
 
-\modelitem{Authoritarian regimes}{States that do not respect lawful process will try to compel administrators of an
+\modelitem{Authoritarian regimes}{Governments that do not adhere to rule of law will try to compel administrators of an
 \ac{EA} system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking
@@
 
-To make sense of an \ac{EA} proposal in the face of the stated goals and such a formidable list of threat actors, one
-must include two non-traditional threats. To do so is to step back and view the situation with a broader perspective
-than security experts usually take.
+Two non-traditional threats must be considered to make sense of an \ac{EA} proposal in the face of the stated goals and
+such a formidable list of threat actors. Understanding these threats requires viewing security with a broader than usual
+perspective.
 
@@
 \modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
-The end user is not the most important thing. This is a crucial point.
+Therefore, the end user's will is not the platform's top priority.\\
 
-All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of this value system is some most
-important ``thing.'' Traditional device security threat models define this ``thing'' as the end user. This leads to good
-security priorities and honors the user's rights and autonomy.
+All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is one top
+priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
+decisions and honors the user's rights and autonomy.\\
 
-To stray from this rule is dangerous, but is already frequent practice. Software that forces updates values the vendor's
-judgement over the user's (even if it is a good idea). Software that enforces Digital Rights Management values
-corporations' data rights over the user's (sometimes to extremes \cite{eff_2020}). Software that does not encrypt user
-communications at all values the service provider's ability to scrape personal information over the user's privacy.
-Software that enables \ac{EA} values lawful investigatory process over the user's ultimate privacy. Intentional
-subversion of the user's will is not a comfortable thought, but it is a frequent reality, and the reality of \ac{EA}.
+It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the vendor's update
+policy over the user's preferences. Digital Rights Management software values corporations' data rights over the user's
+(sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's ability to collect
+personal information over the user's privacy. \Acl{EA} software values lawful investigation over the user's absolute
+privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it is a frequent
+reality.\\
 
-In the broader threat model, the platform abuser uses encryption to hide illicit activities. He is a threat to public
-safety.}
+In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is a threat to
+public safety.}
 
-\modelitem{Hawkish lawmakers}{If one's very highest priorities are security and privacy for society as a whole, then
-aggressive lawmakers that have the power to \ii{mandate} arbitrary data access represent perhaps the largest threat of
-all. Crypto-anarchists can encrypt their data above the compromised layer all they want, but above-the-table
-organizations and most individuals who use default products will abide by U.S. law.
+\modelitem{Hawkish lawmakers}{In the pursuit of security and privacy for society as a whole, aggressive lawmakers that
+have the power to \ii{mandate} arbitrary data access present perhaps the largest threat of all. Crypto-anarchists will
+encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
+use default products will abide by U.S. law.\\
 
@@
 
-\modelitem{Encyption workarounds}{Encryption workarounds are always in play for essentially any threat actor at any
-time. Discussion for defending against those is out of scope.}
+\modelitem{Encyption workarounds}{Encryption workarounds, such as finding a copy of plaintext or guessing a password,
+are always a risk. Defending against such workarounds is not the role of the \ac{EA} system itself.}
 
@@
 prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
-software are considered to be as-intended, though they are assumed to contain the standard amount of bugs and
-side-channels.}
+software are assumed to be uncompromised, aside from the standard amount of bugs and side-channels.}
 
-\modelitem{Manipulation of internal hardware}{Hyper-advanced hardware analysis and manipulation is a real threat from
-foreign intelligence agencies. Defending against these attacks is a separate problem.}
+\modelitem{Manipulation of internal hardware}{Foreign intelligence agencies may have the capability to perform
+hyper-advanced hardware analysis and manipulation. Defending against these attacks is not the role of the \ac{EA} system
+itself.}
 
-\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes that the legal process used to obtain the
-warrant is valid. Additionally, all forms of surveillance, even when protected from \ac{masssurv}, require oversight.
-Transparency via auditability, preferably ensured via technical mechanism, is a goal for good \ac{EA} proposals.
-However, rule of law is an absolute prerequisite for which no \ac{EA} mechanism can compensate.}
+\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes legitimacy of the legal process used to
+obtain the warrant. All forms of government search require oversight. Transparency via auditability, preferably ensured
+via technical mechanism, should be a goal for \ac{EA} proposals. However, rule of law is an absolute prerequisite for
+which no \ac{EA} mechanism can compensate.}
 
@@
 
-Answering ``what are we building'' and comparing it with the state of the art requires an understanding of the state of
-the art. Apple iPhones are among the most secure consumer mobile phones, and serve as the point of reference for
-\ldawmsR. As this analysis is limited to a specific proposal, it is also limited to this reference architecture.
-\myfig{fig-dfd-iphone} is a \ac{DFD} depicting the iPhone's unlock and decryption process according to Apple's public
-documentation \cite{apple_2020}.
+Understanding the \ldawmsr proposal and comparing it to current mobile device encryption technology requires an
+understanding of current technology. Apple iPhones are among the most secure consumer mobile phones and serve as the
+point of reference for \ldawmsR. Because this analysis is limited to a specific proposal, it is also limited to this
+reference architecture. In \myfig{fig-dfd-iphone}, a \ac{DFD} depicts the iPhone's unlock and decryption process
+according to Apple's public documentation \cite{apple_2020}.
 
@@
 
-This \ac{DFD}, and all that follow, is at a specific level of abstraction, simplified in the spirit of ``all models are
-wrong, but some are useful.'' The \acp{DFD} are intended to be useful for communication and threat elicitation only, but
-are accurate in that capacity.
+This \ac{DFD}, and all that follow, presents the system at a specific level of abstraction. The system representation is
+simplified in the spirit of ``all models are wrong, but some are useful.'' The \acp{DFD} exclude some details, but are
+accurate for the purposes of communication and threat elicitation.
 
-As the diagram illustrates, several steps occur between the user entering her \ac{PIN} and unlocking of the device. The
-operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines it with the device-specific
-Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is finally handed to
-the Encryption Module to perform encryption and decryption at a hardware level between the operating system and actual
-encrypted storage.
+As the diagram illustrates, several steps occur between the \ac{PIN} being entered and the device being unlocked. The
+operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines the \ac{PIN} with the
+device-specific Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is
+passed to the Encryption Module. The Encryption Module performs encryption and decryption at a hardware level between
+the operating system and the encrypted storage.
 
-The \ac{UID} is not depicted in key storage because it is actually burned into the secure processor silicon, and cannot
-be read by software. The Volume key is stored encrypted by the Class key. When a user updates his \ac{PIN}, the
-encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors never
-handle any keys or secret information besides the \ac{PIN}, and due to inline encryption in the Encryption Module,
-persistent storage never handles any \ac{plaintext} data.
+The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
+software. The Volume key is encrypted by the Class key before it is stored. When a user updates his \ac{PIN}, the
+encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
+not handle any keys or secret information besides the \ac{PIN}. Due to inline encryption in the Encryption Module,
+persistent storage does not handle \ac{plaintext} data.
 
@@
 
-\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{EA} requests, and a vendor
-mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time vaulting,
-authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}; see the original paper
-for full detail \cite{savage_lawful_2018}.
+\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{EA} requests, and a
+vendor-mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time
+vaulting, authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}.
 
@@
 Device key escrow is considered the most promising technical \ac{EA} direction because of the unique capabilities that
-secure hardware offers, and \ldawmsr demonstrates this capabilities perfectly. The greatest of these capabilities are
-physical possession requirements and strong transparency. Physical possession requirements strongly mitigate mass
-surveillance and, as Savage puts it, ``provides a more intuitive compatibility with common under-standings of the
-government’s reasonable law enforcement powers (e.g., the ability to seize physical property under court order) than
-more information-centric approaches, which may appear covert by comparison'' \cite{savage_lawful_2018}. Transparency is
-also rooted in the physicality of the approach, wherein the device cannot be unlocked without making irreversible and
-detectable changes, mitigating both \ac{masssurv} and common criminal abuse.
+secure hardware offers---namely, physical possession requirements and strong transparency. The physical possession
+requirement mitigates the risk of \ac{masssurv}. As Savage adds, it also ``provides a more intuitive compatibility with
+common under-standings of the government’s reasonable law enforcement powers (e.g., the ability to seize physical
+property under court order) than more information-centric approaches, which may appear covert by comparison''
+\cite{savage_lawful_2018}. Transparency is also rooted in the physicality of the approach, wherein the device cannot be
+unlocked without making irreversible and detectable changes, reducing the risk of both \ac{masssurv} and common criminal
+abuse.
 
-One clear weakness in the design is the shared secret between the device and the manufacturer used in the authorization
-protocol. The key may be stored on the device, but one can access it if they have this shared secret. This is very much
-like trusted-party key escrow where the vendor is the trusted party, only they store the authentication key instead of
-the unlock key itself. Due to this similarity, one of the biggest \ac{EA} problems of centralized risk is not fully
-addressed. This is discussed further in \mysec{sec-prop-threats}.
+One clear weakness in the design is its reliance on the secret used in the authorization protocol shared between the
+device and the manufacturer. Although the key is stored on the device, anyone could access the key if they know the
+shared secret. This is similar to trusted-party key escrow. The vendor is the trusted party, and they store the
+authorization key rather than the encryption key itself. Due to this similarity, the problem of centralized risk is not
+fully addressed. This is discussed further in \mysec{sec-prop-threats}.
 
@@
 self-escrow and authorization protocols. This figure uses the asymmetric key pair approach for Class key encryption in
-the Secure Enclave, which has stricter hardware requirements; however, Savage offers a symmetric key variant that
-achieves the same purposes. The asymmetric key pair approach was chosen here because it is semantically clearer and
-easier to illustrate.
+the Secure Enclave, which creates strict hardware requirements. However, Savage offers a symmetric key variant that
+achieves the same purposes. The asymmetric key variant was chosen for analysis because it is semantically simpler.
 
@@
 
-The lawful access use case operates as follows: law enforcement obtains a warrant to conduct a search, including the
-suspect's mobile device. They obtain the phone, disassemble it to reveal the unexposed \ac{EA} hardware interface, and
-begin the time-vaulted access request. After the lockout period is completed, the device offers its Device ID only. The
-law enforcement agency's digital forensics department digitally signs the Device ID and sends it to the device vendor,
-alongside the warrant and accompanying proof of legal authorization. The vendor's access compliance department audits
-the legal documents and submits the digitally signed Device ID to a secure computing environment such as a \ac{HSM}. The
-secure environment authenticates the digital signature, looks up the device's private Device Seal Key and authorization
-token, and encrypts them with the law enforcement agency's public key. These encrypted artifacts are returned to the
-agency, which decrypts and submits them to the device through the dedicated interface. First is the authorization token,
-which is hashed and compared to the stored hash. If that values match, the device accepts and uses the private Device
-Seal Key to decrypt the Class key, which was escrowed after being encrypted but the public Device Seal Key. Once the
-Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for forensic investigation.
+When law enforcement needs access to a device, it first obtains a warrant to conduct the search. The warrant uniquely
+identifies the suspect's mobile device. Law enforcement obtains the device, disassembles it to reveal the unexposed
+\ac{EA} hardware interface, and begins the time-vaulted access request. After the lockout period is completed, the
+device offers its Device ID only. The law enforcement agency's digital forensics department digitally signs the Device
+ID and sends it to the device vendor, alongside the warrant and accompanying proof of legal authorization. The vendor's
+access compliance department audits the legal documents and submits the digitally signed Device ID to a secure computing
+environment such as a \ac{HSM}. The secure environment authenticates the digital signature, looks up the device's
+private Device Seal Key and authorization token, and encrypts them with the law enforcement agency's public key. These
+encrypted artifacts are returned to the agency, which decrypts and submits them to the device through the dedicated
+interface. First, the authorization token is hashed and compared to the stored hash. If the hash matches, the device
+accepts and uses the private Device Seal Key to decrypt the Class key, which was escrowed after being encrypted but the
+public Device Seal Key. Once the Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for
+forensic investigation.
 
-This concludes an summarized overview of the system as described in the proposal. However, in a real system, the entire
-information lifecycle matters. In order to establish such a system, vendors and law enforcement must establish
-mechanisms to populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} is a
-\ac{DFD} for these scenarios.
+This above process is a summary of the system in the \ldawmsr proposal. However, complete proposals must account for the
+entire information lifecycle. In order to establish such a system, vendors and law enforcement must build mechanisms to
+populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} depicts a \ac{DFD}
+for these scenarios.
 
@@
 
-The proposal has been illustrated. What threats does it introduce? Recall from \mysec{sec-crypto-basics} that security
-can be expressed in the properties of authentication, integrity, non-repudiation, confidentiality, availability, and
-authorization. Threats are often expressed as the negative of these properties: spoofing, tampering, repudiation,
-information disclosure, \acl{DOS}, and elevation of privileged \cite{shostack_threat_2014}. This section
-analyzes \ldawmsr using these categories. Threats that fit under multiple categories, such as tampering for the effect
-of repudiation, are listed only once.
+The next step in threat modeling is to elicit threats based on the system's goals, threat actors, scope, and design.
+Recall from \mysec{sec-crypto-basics} that security can be expressed in terms of authentication, integrity,
+non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
+properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privileged
+\cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
+multiple categories, such as tampering for the effect of repudiation, are listed only once.
 
@@
 computer scientist Ray Ozzie's related \ac{DAR} \ac{EA} protocol called CLEAR \cite{ozzie_2018}. Savage acknowledges the
-influence that CLEAR had on \ldawmsR, and specifically included a note that while this proposal is technically
-vulnerable to the same threat, the mitigations listed here would help, and furthermore the likelihood of such an attack
-is low and the difficulty high.
+influence that CLEAR had on \ldawmsR, and specifically included a note that this proposal is technically vulnerable to
+the same threat. This threat is not particularly severe, as the difficulty of such an attack is so high that it is
+unlikely to be attempted. Even so, the mitigations listed here would also help.
 
@@
 
-Tampering threats are listed in \mytab{table-tampering}. \ii{S2} relies on tampering similar to \ii{T1}.
+Tampering threats are listed in \mytab{table-tampering}. Note that execution of \ii{S2} above relies on a tampering
+attack such as \ii{T1}.
 
@@
 
-Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed here, all
-attacks aimed at gaining unauthorized access to a device constitute information disclosure threats. Most spoofing
-attacks are aimed at disclosing information intended for another party.
+Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed, any attack
+aimed at gaining unauthorized access to a device is an information disclosure threat. Most spoofing attacks aim to
+disclose information intended for another party.
 
@@
 
-How much risk does \ldawmsr introduce? The answer lies in a range depending on how many mitigations are implemented.
-Some threats introduce inherently low risk due to their difficult nature and low pay off. Others are higher risk but can
-be reasonably mitigated. Some threats specifically target transparency, and are difficult to address without trust.
-Finally, there are some risks that are irreducible. Each of classes are discussed in turn.
+The amount of risk \ldawmsr introduces depends on the number of suggested mitigations that are implemented. Some threats
+introduce inherently little risk due to their difficult nature and low pay off. Others introduce greater risk but can be
+reasonably mitigated. Some threats specifically target transparency, and therefore are difficult to address without
+trust. Finally, there are some risks that cannot be safely mitigated. Each of these classes of threats are discussed in
+turn.
 
-\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low risk either
-because they fail to produce useful results, they can be reasonably mitigated, or they are a threat that is already
-present in state-of-the-art designs as depicted in \mysec{sec-basic-dar}. It is crucial for a design to consider and
-mitigate the risk posed by these threats as part of the design process, but they do not represent fatal flaws to a
-design. Except in combination with other threats, they are not worthy more concern here.
+\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low-risk. They
+either they fail to produce useful results, can be reasonably mitigated, or are a threat that is already present in the
+baseline depicted in \mysec{sec-basic-dar}. Although these threats are not fatal flaws to a design, it is important for
+a design to consider and mitigate the risk they pose. Unless they appear in combination with other threats, they are do
+not deserve more discussion.
 
-\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are each higher risk threats to the
-security, trustworthiness, and effectiveness of the proposal, but can be addressed through a set of related mitigations:
-require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device to be
-physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities the
-lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly is a technical measure that entirely
-precludes certain attacks; requiring the device to be on LEA premises is a procedural measure that prevents attackers
-from manipulating LE into achieving their own ends; limiting the number of LE facilities that can participate prevents
-LE impersonation attacks. Together, they make the proposal considerably harder to attack.
+\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
+security, trustworthiness, and effectiveness of the proposal. However, they can be addressed through a set of related
+requirements: require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device
+to be physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities
+the lawful right to use the \ac{EA} mechanism. Requiring destructive disassembly entirely precludes certain attacks.
+Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
+Limiting the number of law enforcement facilities that can participate in the \ac{EA} system prevents LE impersonation
+attacks. Together, these mitigations make the proposal more resistant to attack.
 
 \ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
-legal process and cooperation of several parties. But what if the legal process involves secret courts, secret
-investigations, and secrecy orders on the vendor \cite{shamsi_2011}? Destructive disassembly or the weaker hardware
-transparency mechanism provide non-repudiation that the device was unlocked, if and only if one has physical access to
-the device. Technically enforced transparency into device access requests and responses would grant non-repudiation at a
-broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR} \ac{EA} proposals do
-\cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately, building this
-feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider it an anti-feature.
+legal process and cooperation of several parties. Even in the U.S., the legal process can involve secret courts, secret
+investigation, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is difficult to
+achieve. A destructive disassembly requirement provides non-repudiation that the device was unlocked, if and only if one
+has physical access to the device. Technically enforced transparency into device access requests and responses would
+grant non-repudiation at a broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR}
+\ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately,
+building this feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider too much
+transparency to be an anti-feature.
 
-Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, each representing in slightly different ways the
-disclosure of unlock artifact data, the DID-DSK database. This is the truly difficult challenge---centralizing so much
-sensitive data makes it an extremely high value target for foreign intelligence and cyber criminals. The vendor is in a
-position where they need to protect this database from the most advanced attackers in the world while regularly updating
-and accessing it. \Acp{HSM} and strong security policies would make that a difficult task, but not impossible, and one
-has to assume that the data would leak eventually. What would leaked unlock artifacts mean? It depends who steals it.
-Criminals would use it for personal and financial gain; foreign intelligence and authoritarian regimes would use it to
-spy on their enemies. In the worst-case scenario, anyone with some specialized tools and a willingness to break the law
-could get into any device under the scheme, provided they destructively disassemble it and wait through the time vault.
+Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
+data, the DID-DSK database. These threats create a large amount of risk that cannot be safely mitigated. To store so
+much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
+vendor needs to protect this database from the most advanced attackers in the world while regularly updating and
+accessing it. \Acp{HSM} and robust security policies would make that a difficult task. The data will almost certainly
+leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would use the
+data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on their
+enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
+compliant device after dissembling it and waiting through the time vault.
 
@@
 This chapter began with a list of goals by which \ac{DAR} \ac{EA} proposals can be measured, including security,
-protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section I measure
-\ldawmsr against each of these goals.
+protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section \ldawmsr is
+measured against each of these goals.
 
-% Consider goals, threat actors, and loop-back to out of scope items, because they aaaaactually do matter.
-
-\paragraph*{Security} Security is the main topic of this chapter, and has already been discussed thoroughly. \ldawmsr
-introduces significant risk in the DID-DSK database itself, but does a good job in many other respects, especially if
-the recommended mitigations are implemented. It goes from very hard to attack, when everything is working as intended,
-to simple to attack, when the database is leaked. However, even when it becomes trivial to attack, the impact is limited
-to devices in physical possession and cannot be hidden from the device user.
+\paragraph*{Security} Security has already been discussed at length in this chapter. \ldawmsr introduces significant
+risk in the DID-DSK database itself, but performs well in many other respects, especially if the recommended mitigations
+are implemented. It goes from very hard to attack, when everything is working as intended, to simple to attack, when the
+database is leaked. However, even when it becomes trivial to attack, the attackers are limited to devices in their
+physical possession and cannot hide access from their victims.
 
 \paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
-authoritarian regimes. This proposal does an excellent job of protecting against \ac{masssurv} due to its physicality.
-Any power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior
-to a leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After
-a leak, these regimes could get into any device they want. If one lived in a country that was willing to take and
-destroy their devices to read their contents without legal due process, this would be an unwelcome technology.
+authoritarian regimes. \ldawmsr does an excellent job of protecting against \ac{masssurv} due to its physicality. Any
+power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior to a
+leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After a
+leak, these regimes could get into any device they want. If one lived in a country that was willing to take and destroy
+their devices to read their contents without legal due process, this would be an unwelcome technology.
 
-\paragraph*{Transparency} \ldawmsr excels at transparency on the device, but fails at transparency to the public. Due to
-destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed using this method
-is irrefutable. However, this is only possible to confirm if the device is available. If it is stolen and unlocked by
-criminals or is lawfully accessed under secretive conditions, both the user and the public have no way of knowing. This
-leaves open the possibility for government abuse---though again with the caveat that physical possession requirements
-mean it cannot be abused without a trace.
+\paragraph*{Transparency} \ldawmsr provides transparency to the device user, but does not provide transparency to the
+public. Due to destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed
+using this method is irrefutable. However, this can only be confirmed if the device is available. Neither the user nor
+the public can detect if a device is lawfully accessed under secretive conditions or is stolen and unlocked by
+criminals. This leaves allows the potential for government abuse---though again with the caveat that physical possession
+requirements mean it cannot be abused without a trace.
 
-\paragraph*{Law enforcement utility} This proposal involves a few tradeoffs between security and law enforcement
-utility. The physical possession, time vault, and limited number of authorized law enforcement laboratory requirements
-each insert a barrier to usability. The physical possession requirement is far too important for security and civil
-liberties to sacrifice for utility. Furthermore, law enforcement, and search and seizure law generally, is geared
-towards physical evidence already. The time vault prevents against time-sensitive threats, though destructive
-disassembly (not part of the original proposal) may be enough to cover these cases as well, so the time vault could be
-decreased without significant impact. Limiting the number of parties able to use the system mitigates spoofing attacks;
-local law enforcement would have to cooperate with one of these parties, such as a regional \ac{FBI} office. (This
-really just shifts the spoofing threat up a level, though the shift is appropriate.) In the end, law enforcement still
-has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end; the onus is on the government to fund
-the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
+\paragraph*{Law enforcement utility} \ldawmsr performs a few tradeoffs between security and law enforcement utility. The
+requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
+authorized law enforcement laboratories each create barriers to usability. The physical possession requirement is far
+too important for security and civil liberties to sacrifice for utility. Furthermore, law enforcement, and search and
+seizure law generally, is geared towards physical evidence already. The time vault prevents against time-sensitive
+threats, though destructive disassembly (not part of the original proposal) may be enough to cover these cases as well,
+so the time vault could be decreased without significant impact. Limiting the number of laboratories able to use the
+system reduces the risk of spoofing attacks; local law enforcement would have to cooperate with one of these special
+laboratories, such as a regional \ac{FBI} office. (This really just shifts the spoofing threat up a level, though the
+shift is appropriate.) In the end, law enforcement still has a reliable way to get into lawfully seized devices.
+\ldawmsr upholds its end, but the government must fund the \ac{FBI} operation and cultivate cooperation between local
+and federal agencies.
 
-\paragraph*{Economic impact} Economic costs to the vendor fall into three categories---implementing the design, running
-the system, and absorbing lost business. Even once a proposal like \ldawmsr was refined enough to be implemented, the
-vendor would need to update hardware designs and create the software systems to get it started. Once in use, the vendor
-would face considerable costs in running it securely, considering the supreme importance of transporting and storing the
-unlock artifacts safely. Finally, the vendor would certainly lose privacy-conscious customers to vendors not under the
-system, whether small businesses, foreign competitors, or resellers offering pre-\ac{EA} devices. The scale of these
-costs relative to device manufacturer's naturally large size is unclear. Costs to the government include the \ac{FBI}
-laboratory cost and a necessary scale-up in training and resources available to state and local law enforcement.
-Government reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be applied to a
-\ac{DAR} \ac{EA} regulation as well \cite{edwards_hr4922_1994}.
+\paragraph*{Economic impact} Expenses to the vendor fall into three categories---the costs of implementing the design,
+running the system, and absorbing lost business. Even if a proposal like \ldawmsr was refined enough to be implemented,
+the vendor would need to update hardware designs and create the software systems to get it started. Once in use, the
+vendor would face considerable costs in running it securely, due to the difficulty or safely accessing and updating the
+unlock artifacts. Finally, the vendor would certainly lose privacy-conscious customers to small businesses, foreign
+competitors, or resellers offering pre-\ac{EA} devices. The scale of these costs relative to device manufacturer's large
+size is unclear. Costs to the government include the \ac{FBI} laboratory cost and the cost of law enforcement resources
+and training. Government reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be
+applied to a \ac{DAR} \ac{EA} regulation as well \cite{edwards_hr4922_1994}.

### 2020-12-17 Thesis-Chapter6-ThreatModel.tex
Work on risk analysis

@@
 \ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low-risk. They
-either they fail to produce useful results, can be reasonably mitigated, or are a threat that is already present in the
+either fail to produce useful results, can be reasonably mitigated, or are a threat that is already present in the
 baseline depicted in \mysec{sec-basic-dar}. Although these threats are not fatal flaws to a design, it is important for
-a design to consider and mitigate the risk they pose. Unless they appear in combination with other threats, they are do
+a designer to consider and mitigate the risk they pose. Unless they appear in combination with other threats, they do
 not deserve more discussion.
@@
 \ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
-security, trustworthiness, and effectiveness of the proposal. However, they can be addressed through a set of related
+security, trustworthiness, and effectiveness of the proposal. However, they can be mitigated through a set of related
 requirements: require destructive disassembly of the device to access the \ac{EA} hardware interface, require the device
@@
 Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
-Limiting the number of law enforcement facilities that can participate in the \ac{EA} system prevents LE impersonation
-attacks. Together, these mitigations make the proposal more resistant to attack.
+Limiting the number of LE facilities that can participate in the \ac{EA} process prevents LE impersonation attacks.
+Together, these mitigations make the proposal more resistant to attacks.
 
 \ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
-legal process and cooperation of several parties. Even in the U.S., the legal process can involve secret courts, secret
-investigation, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is difficult to
-achieve. A destructive disassembly requirement provides non-repudiation that the device was unlocked, if and only if one
-has physical access to the device. Technically enforced transparency into device access requests and responses would
-grant non-repudiation at a broader level. \ldawmsr does not include a mechanism for the feature, though other \ac{DAR}
-\ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}. Unfortunately,
-building this feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may consider too much
-transparency to be an anti-feature.
+legal processes and cooperation of several parties. Even in the U.S., the legal process may involve secret courts,
+secret investigations, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is
+difficult to achieve. A destructive disassembly requirement provides non-repudiation that the device was unlocked, if
+and only if one has physical access to the device. Technically enforced transparency into device access requests and
+responses would grant non-repudiation at a broader level. \ldawmsr does not include a mechanism for the feature, though
+other \ac{DAR} \ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}.
+Unfortunately, building this feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may
+consider too much transparency to be an anti-feature.
 
 Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
-data, the DID-DSK database. These threats create a large amount of risk that cannot be safely mitigated. To store so
+data (the DID-DSK database). These threats create a large amount of risk that cannot be safely mitigated. Storing so
 much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
@@
 enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
-compliant device after dissembling it and waiting through the time vault.
+compliant device after disassembling it and waiting through the time vault.
 
@@
 using this method is irrefutable. However, this can only be confirmed if the device is available. Neither the user nor
-the public can detect if a device is lawfully accessed under secretive conditions or is stolen and unlocked by
-criminals. This leaves allows the potential for government abuse---though again with the caveat that physical possession
-requirements mean it cannot be abused without a trace.
+the public can detect if a device has been unlocked under secretive conditions, whether lawfully or criminally. This
+leaves allows the potential for government abuse---though again with the caveat that physical possession requirements
+mean it cannot be abused without a trace.
 
-\paragraph*{Law enforcement utility} \ldawmsr performs a few tradeoffs between security and law enforcement utility. The
+\paragraph*{Law enforcement utility} \ldawmsr includes a few tradeoffs between security and law enforcement utility. The
 requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
-authorized law enforcement laboratories each create barriers to usability. The physical possession requirement is far
-too important for security and civil liberties to sacrifice for utility. Furthermore, law enforcement, and search and
+authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is far too
+important for security and civil liberties to sacrifice for utility. Furthermore, law enforcement, and search and
 seizure law generally, is geared towards physical evidence already. The time vault prevents against time-sensitive
@@
 the vendor would need to update hardware designs and create the software systems to get it started. Once in use, the
-vendor would face considerable costs in running it securely, due to the difficulty or safely accessing and updating the
+vendor would face considerable costs in running it securely due to the difficulty of safely accessing and updating the
 unlock artifacts. Finally, the vendor would certainly lose privacy-conscious customers to small businesses, foreign
@@
 size is unclear. Costs to the government include the \ac{FBI} laboratory cost and the cost of law enforcement resources
-and training. Government reimbursement for vendor compliance costs has precedent in laws like \acs{CALEA}, and could be
-applied to a \ac{DAR} \ac{EA} regulation as well \cite{edwards_hr4922_1994}.
+and training. \acs{CALEA} includes provisions for government reimbursement of vendor compliance costs, and a similar
+provision could be applied to a \ac{DAR} \ac{EA} regulation \cite{edwards_hr4922_1994}.

### 2020-12-18 Thesis-Chapter6-ThreatModel.tex
Editorial improvements

@@
 
-The amount of risk \ldawmsr introduces depends on the number of suggested mitigations that are implemented. Some threats
-introduce inherently little risk due to their difficult nature and low pay off. Others introduce greater risk but can be
-reasonably mitigated. Some threats specifically target transparency, and therefore are difficult to address without
-trust. Finally, there are some risks that cannot be safely mitigated. Each of these classes of threats are discussed in
-turn.
+The amount of risk \ldawmsr introduces depends on the number of suggested mitigations that are implemented. Some
+possible attacks introduce inherently little risk due to their difficult nature and low pay off. Others introduce
+greater risk but can be reasonably addressed. Some threats specifically target transparency, and therefore are difficult
+to address without trust. Finally, there are some risks that cannot be safely mitigated. Each of these classes of
+threats are discussed in turn.
 
 \ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low-risk. They
-either fail to produce useful results, can be reasonably mitigated, or are a threat that is already present in the
-baseline depicted in \mysec{sec-basic-dar}. Although these threats are not fatal flaws to a design, it is important for
-a designer to consider and mitigate the risk they pose. Unless they appear in combination with other threats, they do
-not deserve more discussion.
+either are preventable, unuseful, or already present in the baseline depicted in \mysec{sec-basic-dar}. Although these
+threats are not fatal flaws to a design, it is important for a designer to consider and mitigate the risk they pose.
+Unless they appear in combination with other threats, they do not deserve more discussion.
 
@@
 secret investigations, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is
-difficult to achieve. A destructive disassembly requirement provides non-repudiation that the device was unlocked, if
-and only if one has physical access to the device. Technically enforced transparency into device access requests and
-responses would grant non-repudiation at a broader level. \ldawmsr does not include a mechanism for the feature, though
-other \ac{DAR} \ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020} \cite{goldwasser_public_2017}.
-Unfortunately, building this feature into a proposal makes it potentially vulnerable the hawkish lawmaker, who may
-consider too much transparency to be an anti-feature.
+difficult to achieve. The destructive disassembly requirement enables transparency at a device level. Auditable logs of
+device access \ii{requests and responses} would enable transparency at a system level. \ldawmsr does not create such a
+log, though other \ac{DAR} \ac{EA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020}
+\cite{goldwasser_public_2017}. Unfortunately, building this feature into a proposal makes it potentially vulnerable the
+hawkish lawmaker, who may consider too much transparency to be an anti-feature.
 
@@
 much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
-vendor needs to protect this database from the most advanced attackers in the world while regularly updating and
+vendor needs to protect this database from the most advanced attackers in the world while also regularly updating and
 accessing it. \Acp{HSM} and robust security policies would make that a difficult task. The data will almost certainly
@@
 risk in the DID-DSK database itself, but performs well in many other respects, especially if the recommended mitigations
-are implemented. It goes from very hard to attack, when everything is working as intended, to simple to attack, when the
-database is leaked. However, even when it becomes trivial to attack, the attackers are limited to devices in their
-physical possession and cannot hide access from their victims.
+are implemented. When the unlock artifacts are disclosed, the system becomes simple to attack. However, even in these
+conditions, attackers are limited to devices in their physical possession and cannot hide access from their victims.
 
 \paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
-authoritarian regimes. \ldawmsr does an excellent job of protecting against \ac{masssurv} due to its physicality. Any
-power cannot abuse this technology at scale. It is not as strong at protecting against authoritarian regimes. Prior to a
-leak, vendors would face intense and possibly coercive legal pressure to unlock devices in foreign countries. After a
-leak, these regimes could get into any device they want. If one lived in a country that was willing to take and destroy
-their devices to read their contents without legal due process, this would be an unwelcome technology.
+authoritarian regimes. \ldawmsr successfully protects against \ac{masssurv} due to the physical possession and
+destructive disassembly requirements. It is not as effective against authoritarian regimes, which would put intense and
+possibly coercive legal pressure on vendors to unlock devices. In the case of an unlock artifacts leak, regimes could
+broadly abuse the system. This technology would be unwelcome to those living in countries that do not respect rule of
+law.
 
 \paragraph*{Transparency} \ldawmsr provides transparency to the device user, but does not provide transparency to the
-public. Due to destructive disassembly (or the hardware transparency mechanism), whether a device has been accessed
-using this method is irrefutable. However, this can only be confirmed if the device is available. Neither the user nor
-the public can detect if a device has been unlocked under secretive conditions, whether lawfully or criminally. This
-leaves allows the potential for government abuse---though again with the caveat that physical possession requirements
-mean it cannot be abused without a trace.
+public. As discussed in the previous section, the destructive disassembly requirement would result in transparency at
+the device level. However, access to a device can only be confirmed if the device is available. Neither the user nor the
+public can detect if a device has been unlocked under secretive conditions, whether lawfully or criminally. This creates
+potential for government abuse.
 
@@
 requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
-authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is far too
-important for security and civil liberties to sacrifice for utility. Furthermore, law enforcement, and search and
-seizure law generally, is geared towards physical evidence already. The time vault prevents against time-sensitive
-threats, though destructive disassembly (not part of the original proposal) may be enough to cover these cases as well,
-so the time vault could be decreased without significant impact. Limiting the number of laboratories able to use the
-system reduces the risk of spoofing attacks; local law enforcement would have to cooperate with one of these special
-laboratories, such as a regional \ac{FBI} office. (This really just shifts the spoofing threat up a level, though the
-shift is appropriate.) In the end, law enforcement still has a reliable way to get into lawfully seized devices.
-\ldawmsr upholds its end, but the government must fund the \ac{FBI} operation and cultivate cooperation between local
-and federal agencies.
+authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is too
+important for security to be compromised for utility. Additionally, search and seizure law is already designed for
+physical evidence; maintaining the possession requirement makes the system operate within well-understood norms. The
+time vault prevents against time-sensitive attacks, though destructive disassembly (not part of the original proposal)
+may be enough to cover these cases. The time vault period could be decreased without significant impact. Limiting the
+number of laboratories that can use the system reduces the risk of spoofing attacks; local law enforcement would have to
+cooperate with one of these special laboratories, such as a regional \ac{FBI} office, to unlock a device. In the end,
+law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end, but the
+government must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
 

### 2020-12-18 Thesis-Chapter2-Background.tex
Update forms, small editorial tweaks

--- 2020-12-16 Thesis-Chapter2-Background.tex
@@
 introduces cryptography, summarizes the history of encryption and the ``crypto wars,'' lists prominent technical \ac{EA}
-proposals, and lastly overviews relevant laws and regulations.
+proposals, and overviews relevant laws and regulations.
 
@@
 called the public key and the private key. The public key is not secret, but due to its relationship with the private
-key, it can be used to establish identity and encrypted communication. Using this technique, two parties can communicate
-privately without requiring previously agreed-upon secret information.
+key, it can be used to establish identity and initiate encrypted communications. Using this technique, two parties can
+communicate privately without requiring previously agreed-upon secret information.
 
@@
 challenges that require different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to perform
-authentication and establish short-lived \ac{sym-crypto} session keys. The session keys perform the actual encryption of
-the \acl{DIM}. \Ac{forward-secrecy} and \ac{replay-protection} are two important properties of \ac{DIM} encryption
+authentication and to establish short-lived \ac{sym-crypto} session keys. The session keys perform the actual encryption
+of the \acl{DIM}. \Ac{forward-secrecy} and \ac{replay-protection} are two important properties of \ac{DIM} encryption
 protocols. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise any other private
@@
 \Ac{DAR} by nature must use long-lived keys for encryption. Instead of being negotiated and randomly generated at
-encryption-time, as \ac{DIM} session keys are, these keys are either derived from a user-entered password or are stored
+encryption-time, as \ac{DIM} session keys are, these keys are either derived from user-entered passwords or are stored
 somewhere in computer memory. Secure storage often takes place with the assistance of dedicated hardware.
@@
 (architecture) and the security of the components themselves. Likewise, the security of each component emerges from the
-architecture of its subcomponents and the security of the subcomponents themselves. At the end of this chain of
+architecture of its subcomponents and the security of the subcomponents themselves. At the bottom of this chain of
 analysis, the security of primitive components commonly relies on cryptography.
@@
 catastrophic effects. Compromised digital certificates, hashes, or encryption would enable spoofing, tampering, and
-information disclosure done both for their own sake and as tools in general attacks. The importance of neutralizing
-these tools is clear when such attacks include interference with elections \cite{mueller_2018}, multi-billion dollar
-disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt care of patients
-\cite{goodin_ransomware_2019}.
+information disclosure attacks. The attacks could be done for their own sake or as steps in larger attack chains. The
+importance of neutralizing these attacks is clear when they result in interference with elections \cite{mueller_2018},
+multi-billion dollar disruptions of business \cite{greenberg_2018_notpetya}, and hospital network ransoms that disrupt
+care of patients \cite{goodin_ransomware_2019}.
 
@@
 
-Encryption's dual role as an enabler of privacy and cornerstone of security is at the heart of the EA debate.
-Cryptography has an absolute, mathematical power---a power that is objectionable in the realm of privacy, but necessary
-in the realm of security.
+Cryptography has an absolute, mathematical power---a power that is necessary in the realm of security, but objectionable
+in the realm of privacy. Encryption's dual role as an enabler of privacy and cornerstone of security is at the heart of
+the EA debate.
 
@@
 
-The section provides a brief overview of encryption from ancient to modern times.
+The section provides a brief history of encryption from ancient to modern times.
 
@@
 \cite{luciano_cryptology_1987}. Significant developments include the first formal cryptographic study by Arab scholars
-in the eighth century, advancements made out of necessity in the twentieth century's world wars, and the application of
-computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon formalized the modern ``mathematical
-analysis of cryptography'' in 1949 \cite{shannon_communication_1949} and Whitfield Diffie and Martin Hellman published
-research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of \ac{public-key-crypto} was an
-important advance. Combined with computer networking and personal computing advances in the 1980s, it put the power of
-strong encryption not just in the hands of governments and militaries, but of ordinary people.
+in the eighth century, advancements made out of necessity during the twentieth century's world wars, and the application
+of computers to cryptographic problems \cite{kahn_codebreakers_1996}. Claude Shannon formalized the modern
+``mathematical analysis of cryptography'' in 1949 \cite{shannon_communication_1949} and Whitfield Diffie and Martin
+Hellman published research on \ac{public-key-crypto} in 1976 \cite{diffie_new_1976}. The discovery of
+\ac{public-key-crypto} was an important advance. Combined with computer networking and personal computing advances in
+the 1980s, it put the power of strong encryption not just in the hands of governments and militaries, but of ordinary
+people.
 
@@
 \cite{thompson_2015}. Citing the foundational role of cryptography in security---and the potential of human rights
-abuses for compromised privacy---industrial and technical leaders reacted negatively to the initiative
+abuses resulting from compromised privacy---industrial and technical leaders reacted negatively to the initiative
 \cite{kehl_right_2015} \cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users
@@
 the \ac{SAFEAct} of 1996, which similarly sought to remove export controls, but further sought to explicitly allow
-arbitrarily strong encryption for all legal activity and preclude mandatory \ac{EA} schemes such as the \ac{clipperchip}
-\cite{goodlatte_hr3011_1996}. Zimmermann, the author of PGP, who had by this time endured an investigation by the
-Customs Service for publishing his work, testified before the Senate in favor of the \ac{PROCODEAct}
-\cite{zimmermann_1996}. In November of 1996, the Clinton administration released an executive order effectively removing
-export controls on encryption products along the lines of the \ac{PROCODEAct}, which never came to pass
-\cite{clinton_1996}.
+arbitrarily strong encryption for all legal activity and to preclude mandatory \ac{EA} schemes such as the
+\ac{clipperchip} \cite{goodlatte_hr3011_1996}. Zimmermann, the author of PGP, who had by this time endured an
+investigation by the Customs Service for publishing his work, testified before the Senate in favor of the
+\ac{PROCODEAct} \cite{zimmermann_1996}. In November of 1996, the Clinton administration released an executive order
+effectively removing export controls on encryption products along the lines of the \ac{PROCODEAct}, which never came to
+pass \cite{clinton_1996}.
 
@@
 \cite{kehl_right_2015}. By this time, there was ``an overwhelming amount of evidence against moving ahead with any key
-escrow schemes'' \cite{thompson_2015}, and in 1999 the Clinton administration abruptly changed course, adopting almost
-all \ac{SAFEAct} proposals \cite{kehl_right_2015}. This development marked the end of \ac{the-cw1}.
+escrow schemes'' \cite{thompson_2015}, and in 1999 the Clinton administration changed course again, adopting almost all
+\ac{SAFEAct} proposals \cite{kehl_right_2015}. This development marked the end of \ac{the-cw1}.
 
@@
 and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}. This legislative
-failure meant that rather than being written in the pen of law, encryption policy was written in the pencil of executive
-order.
+failure meant that encryption policy was written not in the pen of law, but in the pencil of executive order.
 
@@
 pursuit of the technical community. Both sides of the debate have by now acknowledged that privacy and public security
-are not always in conflict. In fact \ac{EA}'s effects on public security vs. cybersecurity---which in return affects
-public security---may be more important \cite{schneier_2019}.
+are not always in conflict. In fact, it is the relationship between public security and cybersecurity---which in return
+affects public security---that be more important \cite{schneier_2019}.
 
 \paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack resulted in the first major battle of \ac{the-cw2}. The
-suspect was killed, leaving behind a locked iPhone. The \ac{FBI} issued an order under the \acrlong{AWA} to compel Apple
-to unlock the device, which was among the first generation of Apple's fully encrypted iPhones. Apple objected on grounds
-that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices \cite{schulze_clipper_2017}.
-The case occupied U.S. District courts, the media, and the attention of the political world from February 16, 2016, when
-the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the phone through alternate
-means \cite{novet_2016}. Although the phone proved not to contain any important data \cite{schulze_clipper_2017}, the
-high profile case featured two characteristic elements of the new crypto wars---terrorism and device encryption. Shortly
-after the case ended, a pair of Republican and Democratic senators jointly released a discussion draft of a bill that
-would have forced Apple to comply \cite{burr_2016}. The draft did not make it to the floor, and the unresolved nature of
-the court case left the legal debate unresolved.
+attacker was killed, leaving behind a locked iPhone. The \ac{FBI} issued an order under the \acrlong{AWA} to compel
+Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones. Apple objected on
+grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
+\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the attention of the political world
+from February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to
+the phone through alternate means \cite{novet_2016}. Although the phone proved not to contain any important data
+\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
+wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
+jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
+make it to the Senate floor, and the unresolved nature of the court case left the legal debate unresolved.
 
@@
 
-\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amongst a political landscape marked by
-uncertainty, extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since government is
-part of the broader data security threat model, its behavior is important. Unfortunately, it is difficult for the
-technical community to engage productively with policymakers due to present political dynamics and the lack of reliable
-data showing what law enforcement agencies need \cite{granick_2018}.
+\paragraph*{Volatile Politics} Finally, \ac{the-cw2} is taking place amidst a political landscape marked by uncertainty,
+extremism, misinformation, and a pandemic that will have unknown long-term impacts. Since government is part of the
+broader data security threat model, its behavior is important. Unfortunately, it is difficult for the technical
+community to engage productively with policymakers due to present political dynamics and the lack of reliable data
+demonstrating the extent of law enforcement's difficulties \cite{granick_2018}.
 
@@
 \cite{pfefferkorn_2020}. In response, legislators narrowed the focus of the \ac{EARNITAct}, but only after unveiling the
-\ac{LAEDAct}
+\ac{LAEDAct}---a direct attack on strong encryption calling for broad \acl{EA} capabilities.
 
@@
 a wide variety of implementation options within this category \cite{denning_taxonomy_1996}---secret or open protocol,
-hardware-enabled or software-only, single key or split key, government or third-party escrow---the foundation of the
+hardware-enabled or software-only, single- key or split-key, government or third-party escrow---the foundation of the
 effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
-a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.
+a secret-protocol hardware-enabled split-key government escrow system \cite{denning_taxonomy_1996}.
 
@@
 
-The third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not
-by trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
+The third type of \ac{key-escrow} is device-based. In device \ac{key-escrow}, key information is held not by trusted
+parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
 ``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
@@
 \cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
-enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
+enforcement to exploit vulnerabilities can create misaligned incentives. If law enforcement opposes strong security in
 general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.
@@
 
-Simpler methods than computer hacking or legal coercion can be used: encrypted data may exist in plaintext elsewhere, or
-authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on law
-enforcement's skills in investigation and sting operations. This approach is used successfully today. In 2017, an
+Simpler methods than computer hacking or legal coercion can be used: encrypted data may exist in plaintext copies
+elsewhere, or authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on
+law enforcement's skills in investigation and sting operations. This approach is used successfully today. In 2017, an
 international law enforcement operation used this technique to take down a large illicit online drug market
@@
 
-\acs{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
-Provides explicit protections for private electronic communications data, and updated definitions to contemporary
-technology. It grants law enforcement explicit access to certain data as well; some terminology has aged poorly with
-technological evolution, making the access increasingly permissive \cite{shamsi_2011}.
+\acs{ECPA} includes the Stored Communications Act and Pen Register Act, and amends the switchboard-era Wiretap Act.
+\acs{ECPA} provides explicit protections for private electronic communications data, and updated definitions to
+contemporary technology. It grants law enforcement explicit access to certain data as well; some terminology has aged
+poorly with technological evolution, making the access increasingly permissive \cite{shamsi_2011}.
 
@@
 Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers in
-25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope (Section
-206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business records, but
-the Snowden leaks revealed that the government had been secretly using a non-standard definition of ``business records''
-in order ``to justify requests for domestic telephone metadata delivered in bulk, not individualized requests''
-\cite{landau_making_2013}.
+the 25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expanded \acs{FISA} order scope
+(Section 206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business
+records, but the Snowden leaks revealed that the government had been secretly using a non-standard definition of
+``business records'' in order ``to justify requests for domestic telephone metadata delivered in bulk, not
+individualized requests'' \cite{landau_making_2013}.
 
@@
 
-The \ac{LAEDAct} is a direct attack on encryption and the direct successor of the Compliance with Court Orders Act. It
-was introduced by the chair of the Senate Judiciary Committee, also sponsor of the \ac{EARNITAct} and one of the
+The \ac{LAEDAct} is a direct attack on encryption and the spiritual successor of the Compliance with Court Orders Act.
+It was introduced by the chair of the Senate Judiciary Committee, also sponsor of the \ac{EARNITAct} and one of the
 outspoken encryption critics during the 2019 Senate hearing on encryption and \ac{EA}. The \ac{LAEDAct} specifically
@@
 v. United States} may be the most impactful yet. Though the case only resulted in a narrow decision against unwarranted
-cell phone location data tracking, the Court's decision undermines the pre-digital third-party doctrine, which holds
+cell phone location data tracking, the Court's decision undermines the pre-digital era third-party doctrine, which holds
 that Fourth Amendment protection is forfeit when the data is willingly given to a third-party \cite{franklin_2018}.
@@
 As this relates to \ac{EA}, Fourth Amendment interpretation determines the threshold at which authorities require a
-warrant, not what demands they can or cannot make once they have one. As mentioned above, the FBI argued in the San
-Bernardino case that the \ac{AWA} grants it authority to demand access, though others argue that \acs{CALEA} explicitly
-precludes this kind of demand \cite{gidari_2016}. Since the FBI dropped the case, the matter is not formally settled,
-though the consensus is that legislation such as the \ac{LAEDAct} is required before courts would uphold that command.
+warrant, but does not specify what demands they can or cannot make once they have one. As mentioned above, the FBI
+argued in the San Bernardino case that the \ac{AWA} grants it authority to demand access to \ac{plaintext}, though
+others argue that \acs{CALEA} explicitly precludes this kind of demand \cite{gidari_2016}. Since the FBI dropped the
+case, the matter is not formally settled, though the consensus is that legislation such as the \ac{LAEDAct} is required
+before courts would uphold such a command.
 
@@
 
-That judicial influence on privacy and technology is changing is to be expected. To explain the fluctuating state of
-legal interpretation as it relates to \ac{EA}, Matyas et al. make the helpful distinction between legal rules and
+It is to be expected that judicial influence on privacy and technology is evolving. To explain the fluctuating state of
+legal interpretation as it relates to \ac{EA}, Vashek Matyas et al. make the helpful distinction between legal rules and
 principles \cite{matyas_incommensurability_2018}:
@@
 as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
-these countries, the member's encryption policies are relevant to one another.
+these countries, the member's encryption policies are relevant to each another.\\
 

### 2020-12-18 Thesis-Chapter6-ThreatModel.tex
Update forms, small editorial tweaks

@@
 
-\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are constantly looking for ways to escalate
+\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are constantly looking for ways to elevate
 their privilege and tamper with or steal information. They have world-class expertise and computing power, which they
@@
 Elevation of privilege are listed in \mytab{table-privilege}. Spoofing to usurp a entity's privilege or using any attack
-to unlock a device without authorization can be seen as escalation of privilege.
+to unlock a device without authorization can be seen as elevation of privilege.
 

### 2020-12-18 Thesis-Chapter7-PathsForward.tex
Minor editorial improvements

--- 2020-12-13 Thesis-Chapter7-PathsAhead.tex
@@
 \ac{wicked-prob}. The importance of information security is increasing as more data, business, and infrastructure goes
-online. The technical community is struggles with security. Even for the most well-equipped organizations, thwarting the
+online. The technical community struggles with security. Even for the most well-equipped organizations, thwarting the
 average attacker is difficult, and the best attacker, impossible. \Ac{cryptography} plays a fundamental role in
@@
 especially for encryption policy due to rapidly evolving technology and a vulnerability arms race between device
-manufacturers and lawful hackers. The circumstances are sure to change, and the conclusion that \ac{EA} is unjustified
-may change as well.
+manufacturers and lawful hackers. Circumstances are sure to change, and the conclusion that \ac{EA} is unjustified may
+change as well.
 
@@
 
-Further research into \ac{EA} is justified by the additions to the threat model and the instability of present
+Further research into \ac{EA} is warranted due to the additions to the threat model and the instability of present
 circumstances. Fortunately, developments in secure hardware mark real progress in the search for acceptably secure
@@
 current policy issues, such as encryption, automation, and misinformation. Government sometimes suggests that the tech
-community simply needs to try harder. Because these are not tame problems, the tech world is absolutely right that
+world simply needs to try harder. Because these are not tame problems, the tech world is absolutely right that
 simply trying harder will not work---but that does not mean they cannot deal with them. The tech industry is optimized
@@
 \acp{wicked-prob}. In order to do so, the tech industry must involve people of diverse backgrounds and values in design
-and investment decisions. Designers must consider the ways in which their products could be abused and business models
+and investment decisions. Designers must consider the ways in which their products could be abused, and business models
 must be responsive to evolving conditions.
@@
 
-\ac{EA} is only worthwhile when security is strong enough to keep attackers out in the first place, which is not the
-case for current device encryption. This is a security ergonomics problem more than a technical problem---cryptographers
-are already capable of creating stronger device encryption, but users would rebel due to its adverse effects on
-usability. Solving the strong-yet-usable device encryption problem will make the most promising class of \ac{EA}, device
-key escrow, relevant. Improving security in general will make any \ac{EA} proposal more tenable.
+\ac{EA} is only worthwhile when basic security renders lawful hacking insufficient, which is not the case for current
+device encryption. This is a security ergonomics problem more than a technical problem---cryptographers are already
+capable of creating stronger device encryption, but users would rebel due to its adverse effects on usability. Solving
+the strong-yet-usable device encryption problem will make the most promising class of \ac{EA}, device key escrow,
+relevant. Improving security in general will make any \ac{EA} proposal more tenable.
 
@@
 physical possession and disassembly requirements. Ideally, usage of \ac{EA} mechanisms would be fully auditable to the
-public through inherent technical mechanisms. This could be improved for \acl{DAR} and must be improved for \acl{DIM}.
+public through inherent technical mechanisms. This can be improved for \acl{DAR} and must be improved for \acl{DIM}.
 
@@
 
-Fortunately, this research is already happening to some degree, but more is necessary. Some \ac{EA} proposals include
+Fortunately, this research is already happening to some degree, though more is necessary. Some \ac{EA} proposals include
 built-in transparency and distributed trust mechanisms \cite{goos_oblivious_1996} \cite{phan_key_2017}
@@
 
-The goal of this thesis is to establish encryption and \ac{EA} as a \ac{wicked-prob}, to analyze the debate, and to
-evaluate the security of current \ac{EA} technology. I aim to advance the discussion with a strategy to tackle
-\acp{wicked-prob}, additions to the cryptosystem threat model, and detailed analysis of a modern proposal. The strategy
-is based on the \acl{OODAloop}. In those terms, this paper serves to \ii{observe} and \ii{orient}. Technologists,
-policymakers, and common people must \ii{decide} and \ii{act}. Confronting these issues may be difficult, but the future
-of security, safety, and privacy depends on finding an acceptable solution where no perfect solution exists.
+The goal of this thesis is to establish encryption policy as a \ac{wicked-prob}, to analyze the debate, and to evaluate
+the security of current \ac{EA} technology. I aim to advance the discussion with a strategy to tackle \acp{wicked-prob},
+additions to the cryptosystem threat model, and detailed analysis of a modern proposal. The strategy is based on the
+\acl{OODAloop}. In those terms, this paper serves to \ii{observe} and \ii{orient}. Technologists, policymakers, and
+common people must \ii{decide} and \ii{act}. Confronting these issues may be difficult, but the future of security,
+safety, and privacy depends on finding an acceptable solution where no perfect solution exists.

### 2020-12-19 Thesis-Abstract.tex
Editorial tweaks

@@
-Public debate has resumed on the topic of exceptional access (EA), which refers to alternative means of decryption
-intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at the end of
-the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
+\parboxindent Public debate has resumed on the topic of exceptional access (EA), which refers to alternative means of
+decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
+the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
 circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
@@
 
-The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
-security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
-In response to both polarized and conciliatory voices, this paper analyzes strategies for confronting wicked problems
-and proposes an iterative approach to the case of encryption and EA. Along the way, it illustrates the components of the
-debate in argument maps and demonstrate the security risks with data flow diagrams and threat analysis, focusing on one
-EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk.''
+\parboxindent The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a
+cornerstone of security. The emergence of strong encryption incited the first crypto war, and its proliferation is
+causing the second. In response to both polarized and conciliatory voices, this paper analyzes strategies for
+confronting wicked problems and proposes an iterative approach to the case of encryption and EA. Along the way, it
+illustrates the components of the debate in argument maps and demonstrate the security risks with data flow diagrams and
+threat analysis, focusing on one EA proposal in particular, Stefan Savage's ``Lawful Device Access without Mass
+Surveillance Risk.''

### 2020-12-19 Thesis-Chapter2-Background.tex
Editorial tweaks

@@
 as ``Five Eyes'' \cite{nsa_2020}. Due to the close relationship between the intelligence and law enforcement agencies of
-these countries, the member's encryption policies are relevant to each another.\\
+these countries, the member's encryption policies are relevant to each another.
 
@@
 
-The U.K. government has the explicit ability to force decryption, perform \ac{lawful-hacking}, and mandate \ac{EA}
-though ``technical capability notices'' introduced in 2016's Investigatory Powers Act \cite{legislature_2016}. Australia
-has weaker powers through a variety of laws, and in 2018 passed the Assistance and Access Act which created very similar
-``technical capability notices'' \cite{ag_2018}. Canadian law requires cooperation between telecommunications providers
-and law enforcement but does not directly address encryption. Canada has traditionally supported strong encryption, but
-the pressure of \ac{the-cw2} is straining that support \cite{parsons_2019}.
+\parboxindent The U.K. government has the explicit ability to force decryption, perform \ac{lawful-hacking}, and mandate
+\ac{EA} though ``technical capability notices'' introduced in 2016's Investigatory Powers Act \cite{legislature_2016}.
+Australia has weaker powers through a variety of laws, and in 2018 passed the Assistance and Access Act which created
+very similar ``technical capability notices'' \cite{ag_2018}. Canadian law requires cooperation between
+telecommunications providers and law enforcement but does not directly address encryption. Canada has traditionally
+supported strong encryption, but the pressure of \ac{the-cw2} is straining that support \cite{parsons_2019}.
 

### 2020-12-19 Thesis-Chapter3-PolicyMaking.tex
Update forms, small editorial tweaks
Editorial tweaks

@@
 capable of acceptably secure \ac{EA}, that consensus could change, especially considering that \ac{EA} as a field is
-under-researched. Third, that the problem cannot be definitively solved. Evolving values and technology mean that this
-policy area is always up for renegotiation.
+under-researched. Third, the problem cannot be definitively solved. Evolving values and technology mean that this policy
+area is always up for renegotiation.
 
@@
 process of successive approximation to some desired objectives in which what is desired itself continues to change under
-reconsideration'' \cite{lindblom_muddling_1959}. Most importantly, it eschews the linear
-problem-definition--alternatives-analysis steps out of respect that the problem definition is not self-contained.
-Lindblom argues that policy ends and means are interlinked, eventually concluding:
+reconsideration'' \cite{lindblom_muddling_1959}. Most importantly, it eschews the linear flow from problem definition to
+alternatives analysis out of respect that the problem definition is not self-contained. Lindblom argues that policy ends
+and means are interlinked, eventually concluding:
 
@@
 
-The \acl{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
+The \ac{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
 opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but its ideas
 have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring incident
-response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your opponent's
-loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.
+response \cite{schneier_future_2014}. The model emphasizes fast cycle times and ``getting inside'' your opponent's loop
+as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.
 
@@
 both cases, one faces an unpredictable opponent in a dynamic environment in which every action counts. We cannot
-overpower \acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.
+overpower \acp{wicked-prob} through force of reason, but perhaps we can through feedback and agility.
 
@@
   \includegraphics[width=\linewidth]{dfds/build/OODA-policy-process.png}
-  \caption{Using the \ac{OODAloop} to Tackle Wicked Problems}
+  \caption{Using the \acs{OODAloop} to Tackle Wicked Problems}
   \label{fig-policy-ooda-process}

### 2020-12-19 Thesis-Chapter4-Tools.tex
Update forms, small editorial tweaks
Editorial tweaks

@@
 \acp{wicked-prob}. Though formal research on the technique is sparse, \acp{argmap} have proven helpful in simplifying
-complex arguments regarding \acp{wicked-prob} \cite{renton_2007} and in facilitating debate. This is because they slow
+complex arguments regarding \acp{wicked-prob} and in facilitating debate \cite{renton_2007}. This is because they slow
 down discussion, depersonalize conflict, and lend structure and rhythm to meetings \cite{dutoit_hypermedia_2006}. They
@@
     \includegraphics[width=0.8\linewidth]{dfds/build/diffie-hellman-traditional.png}
-    \caption{A \acf{DFD} for a simplified Diffie-Hellman key exchange}
+    \caption{A \acs{DFD} for a simplified Diffie-Hellman key exchange}
     \label{fig-dfd-dh-traditional}

### 2020-12-19 Thesis-Chapter5-Arguments.tex
Editorial improvements
Editorial tweaks

@@
 debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
-as tools in the collaborative pursuit of the problem definition and solution. It is not a replacement for debate arising
-from live discussion. Second, \acp{argmap} may deceptively portray the strength of an argument. Argument nodes do not
-indicate their strength or validity. If ten unsound arguments were presented against one indisputable argument, the side
-with more arguments would, regardless of their integrity, appear stronger. With those qualifications in mind, let us
-begin by examining the factors at the center of the conflict.
+as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
+replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of an
+argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against one
+indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger. With
+those qualifications in mind, let us begin by examining the factors at the center of the conflict.
 
@@
 depicting it as a zero-sum conflict. In reality, privacy does not always diminish safety, but sometimes enhances safety.
-For example, data privacy makes crimes such as stalking and identify more difficult. The ``Going Dark'' argument also
-categorizes privacy as a personal good and safety as a public good. However, privacy can also be a public good when it
-precludes \ac{masssurv}. This argument is explored more deeply in the ``Golden Age for Surveillance'' map.
+For example, data privacy makes crimes such as stalking and identity theft more difficult. The ``Going Dark'' argument
+also categorizes privacy as a personal good and safety as a public good. However, privacy can also be a public good when
+it counters \ac{masssurv}. This argument is explored more deeply in the ``Golden Age for Surveillance'' map.
 
@@
 burden of proof is on them---they must show that encryption is causing investigatory failure. Unfortunately, in most
-cases, it is impossible to know what would have happened without the interference of strong encryption. Inability to
-gather data from controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable,
-another way to gather data is by observing naturally-occurring conditions in which everything but the variable of
-interest remain are constant. These conditions exist for encryption in messaging application ecosystem. Messaging
-applications WhatsApp, Facebook Messenger, and Telegram all have large user bases and similar features. Among these
-services, only Facebook Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market
-share as measured by active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports
-\cite{keller_internet_2019}. This indicates that encryption does indeed affect investigations.
+cases, it is impossible to know what would have happened without the interference of encryption. Inability to gather
+data from controlled tests is a characteristic of \acp{wicked-prob}. When controlled settings are unachievable, another
+way to gather data is by observing naturally-occurring conditions in which everything but the variable of interest are
+constant. These conditions exist for encryption in the messaging application ecosystem. Messaging applications WhatsApp,
+Facebook Messenger, and Telegram all have large user bases and similar features. Among these services, only Facebook
+Messenger does not employ \ac{E2EE}. Globally, Facebook Messenger has approximately 20\% market share as measured by
+active monthly users \cite{statista_2020} but accounts for 65\% of \ac{CSAM} reports \cite{keller_internet_2019}. This
+indicates that encryption does indeed affect investigations.
 
@@
 Encryption is not the only challenge that law enforcement faces with digital evidence. A 2018 study found that
-identifying service providers and getting data from them using existing mechanisms were even greater challenges
+identifying and cooperating with service providers through existing legal frameworks were even greater challenges
 \cite{carter_2018}. The report, titled ``Low-Hanging Fruit,'' also lists resource limitations and poor training as top
@@
 of these issues. For example, if combatting \ac{CSAM} is a government priority, why has Congress underfunded the laws
-it's passed, and why has the Justice Department understaffed their its task force \cite{keller_internet_2019}? These
-questions demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The absence of
-government funding does not mean technical anti-\ac{CSAM} measures are unnecessary.
+it's passed, and why has the Justice Department understaffed its task force \cite{keller_internet_2019}? These questions
+demand answers, but they do not change the validity (or invalidity) of pro-\ac{EA} arguments. The absence of government
+funding does not mean technical anti-\ac{CSAM} measures are unnecessary.
 
@@
 after the analysis of the ``Golden Age for Surveillance'' argument---(4) transparency, and (5) economic impact. Economic
-impact is not as central an issue as the others, but it does help determine practical viability of a proposal.
+impact is not as central an issue as the others, but it does affect the practical viability of a proposal.
 
@@
 \ac{EA}, these methods do not require new legal or technical capability. These methods are effective
-\cite{greenberg_2018}; investment in current digital evidence capabilities and anti-\ac{CSAM} task forces, they are also
-safe.
+\cite{greenberg_2018}; due to their conservative nature, they are also safe.
 
@@
 This approach has already been used \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
-\cite{feuer_chapo_2019}. Because it is already used, using \ac{lawful-hacking} as the primary strategy would require
+\cite{feuer_chapo_2019}. Because it is already used, choosing \ac{lawful-hacking} as the primary strategy would require
 officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.
@@
 \cite{ozzie_2018}. However, such keys are rarely used in comparison to \ac{EA} keys, which would be updated and accessed
-frequently. Although code-signing keys are rarely used and are protected in \acp{HSM}, they still leak
+frequently. Additionally, although code-signing keys are rarely used and are protected in \acp{HSM}, they still leak
 \cite{green_2018}. Once the escrowed keys leak, every device or message encrypted with them would be at risk. This is an
@@
 
-One type of \ac{DIM} \ac{EA} involves uses strong cryptography but includes extra cryptographic ``puzzle pieces'' in
-message metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
+One type of \ac{DIM} \ac{EA} uses strong cryptography but includes extra cryptographic ``puzzle pieces'' in message
+metadata. A surveiller could reconstruct the encryption key from the puzzle pieces, but only after expending
 considerable computational power \cite{bellare_translucent_1996} \cite{wright_crypto_2018}. This approach precludes mass
@@
 Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
-a silent listener on a conference call. This was notably suggested by \ac{GCHQ} officials Ian Levy and Crispin Robinson
-\cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:
+a silent listener on a conference call. This was notably suggested by UK \ac{GCHQ} officials Ian Levy and Crispin
+Robinson \cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:
 
@@
 \ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
-privacy and civil liberties more significantly than criminal abuse.
+privacy and civil liberties more systematically than criminal abuse.
 

### 2020-12-19 Thesis-Chapter6-ThreatModel.tex
Editorial tweaks

@@
 
-\modelitem{Insider threat}{Insider threats involve attackers with roles inside the \ac{EA} recovery process who seek to
+\modelitem{Insider threat}{Insider threats are attackers with roles inside the \ac{EA} recovery process who seek to
 abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
@@
 \modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
-Therefore, the end user's will is not the platform's top priority.\\
+Therefore, the end user's will is not the platform's top priority.
 
-All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is one top
-priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
-decisions and honors the user's rights and autonomy.\\
+\parboxindent All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is
+one top priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
+decisions and honors the user's rights and autonomy.
 
-It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the vendor's update
-policy over the user's preferences. Digital Rights Management software values corporations' data rights over the user's
-(sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's ability to collect
-personal information over the user's privacy. \Acl{EA} software values lawful investigation over the user's absolute
-privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it is a frequent
-reality.\\
+\parboxindent It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the
+vendor's update policy over the user's preferences. Digital Rights Management software values corporations' data rights
+over the user's (sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's
+ability to collect personal information over the user's privacy. \Acl{EA} systems value lawful investigation over the
+user's absolute privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it
+is a frequent reality.
 
-In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is a threat to
-public safety.}
+\parboxindent In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is
+a threat to public safety.}
 
@@
 encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
-use default products will abide by U.S. law.\\
+use default products will abide by U.S. law.
 
-Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{EA} or not) because they may judge the
-system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something weaker in its
-place.}
+\parboxindent Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{EA} or not) because they
+may judge the system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something
+weaker in its place.}
 
@@
 \modelitem{\ac{EA} workarounds}{The user can evade \ac{EA} mechanisms by encrypting data at a higher level in the tech
-stack, for example by encrypting files before saving them to disk. Additionally, investigators or attackers can often
-bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the \ac{EA}
-system itself.}
+stack, for example by encrypting files in software before saving them to disk. Additionally, investigators or attackers
+can often bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the
+\ac{EA} system itself.}
 
-\modelitem{Supply chain attacks}{It is necessary to protecting device hardware and software supply chains in order to
+\modelitem{Supply chain attacks}{It is necessary to protect device hardware and software supply chains in order to
 prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
@@
 The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
-software. The Volume key is encrypted by the Class key before it is stored. When a user updates his \ac{PIN}, the
+software. The Volume key is encrypted by the Class key before it is stored. When a user updates his or her \ac{PIN}, the
 encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
@@
 non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
-properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privileged
+properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privilege
 \cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
@@
 Limiting the number of LE facilities that can participate in the \ac{EA} process prevents LE impersonation attacks.
-Together, these mitigations make the proposal more resistant to attacks.
+Together, these mitigations make the proposal more resistant to attack.
 
@@
 vendor needs to protect this database from the most advanced attackers in the world while also regularly updating and
-accessing it. \Acp{HSM} and robust security policies would make that a difficult task. The data will almost certainly
-leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would use the
-data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on their
-enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
+accessing it. \Acp{HSM} and robust security policies would make that a difficult task. However, the data will almost
+certainly leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would
+use the data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on
+their enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
 compliant device after disassembling it and waiting through the time vault.
@@
 cooperate with one of these special laboratories, such as a regional \ac{FBI} office, to unlock a device. In the end,
-law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr upholds its end, but the
-government must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
+law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr does its part, but the government
+must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
 
@@
 and training. \acs{CALEA} includes provisions for government reimbursement of vendor compliance costs, and a similar
-provision could be applied to a \ac{DAR} \ac{EA} regulation \cite{edwards_hr4922_1994}.
+provision could be applied to \ac{DAR} \ac{EA} regulation \cite{edwards_hr4922_1994}.

### 2020-12-19 Thesis-Glossary.tex
Fix LEAD->LAED act term
Address more todos
Editorial tweaks

@@
     when security researcher Matt Blaze discovered flaws that allowed users to subvert the Clipper Chip mechanisms
+}
+
+\newterm{masssurv}{mass surveillance}{
+    Ubiquitous data collection and analysis by remote, centralized bodies
 }
@@
 \newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
-\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters (UK)}
+\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters}
 \newabbreviation{HSM}{HSM}{Hardware Security Module}
@@
 \newabbreviation{EARNITAct}{EARN IT Act}{Eliminating Abusive and Rampant Neglect of Interactive Technologies Act}
-\newabbreviation{LEADAct}{LEAD Act}{Lawful Access to Encrypted Data Act}
+\newabbreviation{LAEDAct}{LAED Act}{Lawful Access to Encrypted Data Act}
 \newabbreviation{OODAloop}{OODA Loop}{\ii{Observe-Orient-Decide-Act} Loop}
+\newabbreviation{SN}{S/N}{serial number}
 

### 2021-01-30 articles/tops-acm/manuscript.tex
Add ACM template and samples

%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[manuscript,screen,review]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
%% TODO: update with real info.
\setcopyright{acmcopyright}
\copyrightyear{2021}
\acmYear{2021}
\acmDOI{10.1145/1122445.1122456}

%%
%% These commands are for a JOURNAL article.
%% TODO: update with real info.
\acmJournal{TOPS}
\acmVolume{0}
\acmNumber{1}
\acmArticle{2}
\acmMonth{3}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Threat Modeling for a Lawful Device Access Proposal}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Kevin Kredit}
\email{k.kredit.us@ieee.org}
% \orcid{1234-5678-9012}
\affiliation{%
  \institution{Grand Valley State University}
  \streetaddress{318C DeVos Center, 401 W. Fulton Street}
  \city{Grand Rapids}
  \state{Michigan}
  \country{USA}
  \postcode{49504}
}
% \author{G.K.M. Tobin}
% \authornote{Both authors contributed equally to this research.}
% \authornotemark[1]
% \email{webmaster@marysville-ohio.com}
% \affiliation{%
% }

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Kredit}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  Abstract text.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
  <concept>
    <concept_id>10002978.10003006.10011608</concept_id>
    <concept_desc>Security and privacy~Information flow control</concept_desc>
    <concept_significance>500</concept_significance>
  </concept>
  <concept>
    <concept_id>10002978.10003029.10003032</concept_id>
    <concept_desc>Security and privacy~Social aspects of security and privacy</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
    <concept>
    <concept_id>10003456.10003462.10003487</concept_id>
    <concept_desc>Social and professional topics~Surveillance</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
  <concept>
    <concept_id>10003456.10003462.10003588.10003589</concept_id>
    <concept_desc>Social and professional topics~Governmental regulations</concept_desc>
    <concept_significance>300</concept_significance>
  </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Information flow control}
\ccsdesc[300]{Security and privacy~Social aspects of security and privacy}
\ccsdesc[300]{Social and professional topics~Surveillance}
\ccsdesc[300]{Social and professional topics~Governmental regulations}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{encryption, exceptional access, public policy, threat modeling}

\maketitle

\section{Introduction}

Begin article text.

\end{document}
\endinput

### 2021-02-11 articles/tots-ieee/Article.tex
Update TOTS IEEE article, copy in chapter 5

--- 2020-12-19 Thesis-Chapter5-Arguments.tex
@@
-\chapter{EA Debate Arguments}
-\label{chap-arguments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SETUP %
+\documentclass[conference]{IEEEtran}
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+
+%%%%%%%%%%% Non-standard packages
+\usepackage[hyphens]{url}
+\usepackage[acronym,toc,nonumberlist]{glossaries}
+\usepackage[shortcuts=ac]{glossaries-extra}
+\usepackage{rotating}
+\usepackage{csquotes}
+
+%%%%%%%%%%% Commands
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\def\tt#1{\mbox{\texttt{#1}}}
+
+\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}
+
+%% Mine
+\def\tt#1{\mbox{\texttt{#1}}}
+\def\bb#1{\mbox{\textbf{#1}}}
+\def\ii#1{\mbox{\textit{#1}}}
+\newcommand{\myfig}[1]{Figure~\ref{#1}}
+\newcommand{\mysec}[1]{Section~\ref{#1}}
+
+%%%%%%%%%%% Inputs
+\input{../../Thesis-Glossary}
+
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %
+
+\title{Confronting Wicked Crypto}
+
+\author{\IEEEauthorblockN{Kevin Kredit}
+\IEEEauthorblockA{\textit{GVSU ACS Department} \\
+Grand Rapids, Michigan \\
+Email: k.kredit.us@ieee.org}
+\and
+% FIXME: Prof. Kalafut, update as appropriate
+% Is there a way to indicate that you are faculty and I am not? Label yourself with PhD, perhaps?
+\IEEEauthorblockN{Andrew Kalafut, Ph.D.}
+\IEEEauthorblockA{\textit{GVSU ACS Department} \\
+Grand Rapids, Michigan \\
+Email: kalafuta@gvsu.edu}
+}
+
+\maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %
+
+\begin{abstract}
+
+% NOTE: I think we should switch terminology to "Lawful access" for the paper (as opposed to EA)
+
+% NOTE: The first paragraph of the abstract was copied. The second I tweaked.
+
+Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
+decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
+the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
+circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
+technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
+steps, it is necessary that the technical community remain engaged. Although any LA framework would increase risk, the
+magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
+considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
+LA may reduce risk overall.
+
+The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a
+cornerstone of security. The emergence of strong encryption incited the first crypto war, and its proliferation is
+causing the second. In order to make progress in the debate, this paper frames encryption and LA as a wicked problem and
+analyze the debate using argument maps.
+
+\end{abstract}
+
+\begin{IEEEkeywords}
+encryption, exceptional access, policy, wicked problems
+\end{IEEEkeywords}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PAPER %
+
+\section{Introduction}
+\label{sec-intro}
+
+% TODO: add actual intro, not just chapter intro. Take some from paper intro, but also briefly discuss wicked problems
 
@@
 
-\begin{sidewaysfigure}
-  \centering\CaptionFontSize
+\begin{sidewaysfigure*}
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
@@
   \label{fig-args-contrib}
-\end{sidewaysfigure}
+\end{sidewaysfigure*}
 
@@
 
-Information security is the remaining central issue in the debate. As described in \mysec{sec-crypto-basics},
-\ac{cryptography} plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought
-to compromise the cryptographic foundations of encryption; however, this is seen as too risky today.
-\mysec{sec-tech-approaches} introduced several alternative technical approaches to \ac{EA}, but experts still emphasize
-that current systems cannot securely provide the level of access \acl{LE} asks for \cite{abelson_2015}
-\cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs. The field of
-cybersecurity is still equal parts art and science; mandated \ac{EA} would put a heavy burden on a field still finding
-its legs.
+Information security is the remaining central issue in the debate. As described in \mysec{sec-intro}, \ac{cryptography}
+plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
+cryptographic foundations of encryption; however, this is seen as too risky today. \mysec{sec-intro} introduced several
+alternative technical approaches to \ac{EA}, but experts still emphasize that current systems cannot securely provide
+the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
+early computing and networking designs. The field of cybersecurity is still equal parts art and science; mandated
+\ac{EA} would put a heavy burden on a field still finding its legs.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
+% TODO: fix \mysec{sec-intro} reference in above paragraph (there are two instances)
 
@@
 
-\begin{figure}[t!]
-  \centering\CaptionFontSize
+\begin{figure*}[t!]
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.goingdark.pdf}
@@
   \label{fig-arg-going-dark}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[p!]
-  \centering\CaptionFontSize
+\begin{figure*}[p!]
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
@@
   \label{fig-arg-golden-age}
-\end{figure}
+\end{figure*}
 
@@
 thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
-in \mysec{sec-ea-types}.
+in \mysec{sec-intro}.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
 
@@
 
-\begin{figure}[t!]
-  \centering\CaptionFontSize
+\begin{figure*}[t!]
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.fallacies.pdf}
@@
   \label{fig-arg-fallacies}
-\end{figure}
+\end{figure*}
 
@@
 legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
-non-\ac{EA} approaches were first introduced in \mysec{sec-alt-approaches}.) In this section, \ac{EA} is compared to its
+non-\ac{EA} approaches were first introduced in \mysec{sec-intro}.) In this section, \ac{EA} is compared to its
 alternatives. \ac{EA} variations are compared in the next map.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
 
@@
 
-\begin{sidewaysfigure}
-  \centering\CaptionFontSize
+\begin{sidewaysfigure*}
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.measures.pdf}
@@
   \label{fig-arg-measures}
-\end{sidewaysfigure}
+\end{sidewaysfigure*}
 
@@
 
-\begin{sidewaysfigure}[p!]
-  \centering\CaptionFontSize
+\begin{sidewaysfigure*}[p!]
+  \centering
   \includegraphics[width=\linewidth]{arguments/build/comprehensive.classes.pdf}
@@
   \label{fig-arg-classes}
-\end{sidewaysfigure}
+\end{sidewaysfigure*}
 
@@
 leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
-lessons of \mychap{chap-policy}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has
-led to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This
-is because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
+lessons of \mysec{sec-intro}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has led
+to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This is
+because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
 accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
@@
 privacy and civil liberties more systematically than criminal abuse.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
 
@@
 \cite{savage_lawful_2018}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %
+
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,Article} % IEEEfull or IEEEabrv
+
+\end{document}

### 2021-02-11 articles/tops-acm/manuscript.tex
Update TOPS ACM article, copy in chapter 6

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SETUP %
 %%
@@
 
+%%%%%%%%%%% Non-standard packages
+% TODO: check if these are allowed
+% \usepackage[hyphens]{url}
+\usepackage[acronym,toc,nonumberlist]{glossaries}
+\usepackage[shortcuts=ac]{glossaries-extra}
+\usepackage{rotating}
+\usepackage{csquotes}
+\usepackage{ragged2e}\usepackage{makecell}
+
+%%%%%%%%%%% Commands
+\def\tt#1{\mbox{\texttt{#1}}}
+\def\bb#1{\mbox{\textbf{#1}}}
+\def\ii#1{\mbox{\textit{#1}}}
+\newcommand{\myfig}[1]{Figure~\ref{#1}}
+\newcommand{\mytab}[1]{Table~\ref{#1}}
+\newcommand{\mysec}[1]{Section~\ref{#1}}
+\newcommand{\parboxindent}{~~~~~~}
+
+% To use makecell like normal (lines between header cols)
+% \renewcommand\theadfont{\bfseries}
+% To not need makecell (no newlines in headers)
+% \newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}
+% To modify makecell (no lines between header cols)
+\renewcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries \makecell{#1}}}
+
+%%%%%%%%%%% Inputs
+\input{../../Thesis-Glossary}
+
 %%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %
 %%
@@
 }
+% FIXME: Prof. Kalafut, update as appropriate
+% Is there a way to indicate that you are faculty and I am not? Label yourself with PhD, perhaps?
+\author{Andrew Kalafut, Ph.D.}
+\email{kalafuta@gvsu.edu}
+% \orcid{1234-5678-9012}
+\affiliation{%
+  \institution{Grand Valley State University}
+  \streetaddress{318C DeVos Center, 401 W. Fulton Street}
+  \city{Grand Rapids}
+  \state{Michigan}
+  \country{USA}
+  \postcode{49504}
+}
+% TODO: decide whether to add an author note of some kind
 % \author{G.K.M. Tobin}
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Kredit}
+\renewcommand{\shortauthors}{Kredit and Kalafut}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %
 %%
@@
 \begin{abstract}
-  Abstract text.
+
+% NOTE: I think we should switch terminology to "Lawful access" for the paper (as opposed to EA)
+
+% NOTE: The first paragraph of the abstract was copied. The second I tweaked.
+
+Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
+decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
+the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
+circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
+technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
+steps, it is necessary that the technical community remain engaged. Although any LA framework would increase risk, the
+magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
+considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
+LA may reduce risk overall.
+
+The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a cornerstone of
+security. The emergence of strong encryption incited the first crypto war, and its proliferation is causing the second.
+In complex debates, progress is made by focusing on concrete proposals. This paper explores a previous LA proposal,
+Stefan Savage's ``Lawful Device Access without Mass Surveillance Risk'' using augmented data flow diagrams and threat
+analysis.
+
 \end{abstract}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FRONT MATTER %
 %%
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{encryption, exceptional access, public policy, threat modeling}
+\keywords{encryption, lawful access, public policy, threat modeling}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PAPER %
 \section{Introduction}
+\label{sec-intro}
 
-Begin article text.
+% TODO: add actual intro, not just chapter intro. Take some from paper intro, but also introduce the diagram syntax
+
+This chapter consists of a threat model and analysis of a specific device key escrow \ac{LA} proposal, Stefan Savage's
+``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}. The
+proposal is specific to \acl{DAR} for mobile devices. Due to the length of its title, this proposal will be referred to
+by its acronym, \ldawmsR. This chapter assumes that the proposal is accompanied by a legal framework regulating its
+application and use, though the legal framework is not the focus of this analysis.
+
+% Note that the creators of the proposal intentionally chose the phrase ``lawful access'' over ``\acl{EA}'' in order to
+% emphasize the role of legal process and dispel the negative stigma that accompanies \ac{LA}. I refer to \ldawmsr as an
+% \ac{LA} proposal, but use \ac{LA} in ... [that spirit].
+
+\section{Developing a Threat Model}
+
+``What are we building?'' and ``What can go wrong?'' are the essential threat modeling questions. The process of
+answering these questions begins with setting goals, identifying threat actors, and declaring scope. The \ldawmsr
+proposal describes a threat model against which the proposal was designed. This threat model takes cues from the
+\ldawmsr model, but is intended to be more generally applicable for any \ac{DAR} \ac{LA} proposal.
+
+\newcommand{\modelstart}[0]{\begin{itemize}}
+\newcommand{\modelitem}[2]{ % Name, description
+    \item \textbf{#1} \nopagebreak
+
+    \vspace{0.5\baselineskip} \parbox{\linewidth}{#2} \vspace{0.5\baselineskip}
+}
+\newcommand{\modelend}{\end{itemize}}
+
+% \newcommand{\modelstart}{\begin{itemize}}
+% \newcommand{\modelitem}[2]{\item \textbf{#1.} #2}
+% \newcommand{\modelend}{\end{itemize}}
+
+\subsection{Goals}
+
+The goals of an \ac{LA} cryptosystem are taken from the five metrics used to judge \ac{LA} alternatives and variations
+in \mysec{sec-intro}. Each goal is listed according to priority, though no goal has absolute priority over the others.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
+
+\modelstart
+
+\modelitem{Security}{The system should minimize security risk relative to the current baseline. Risk is considered with
+respect to both individual devices and the entire class of devices.}
+
+\modelitem{Protection of civil liberties}{The system should include measures to prevent \ac{masssurv} and abuse by
+authoritarian regimes. This includes technologically enforcing the fundamental right to privacy outside the context of
+warranted search under the rule of law.}
+
+\modelitem{Transparency}{The system should be auditable by the public to allow for verification that its power is being
+used responsibly. In the case of \ac{DAR}, access should be apparent to the data owner.}
+
+\modelitem{Law enforcement utility}{The system should be useful to law enforcement in terms of reliability, speed, and
+cost.}
+
+\modelitem{Economic impact}{The system should consider global competitiveness of devices using the cryptosystem and the
+distribution of costs associated with its development and administration. Even though this is a security-oriented threat
+model, a system's design affects its economics, and its economics determines its ultimate feasibility.}
+
+\modelend
+
+\subsection{Threat Actors}
+
+\ac{LA} threat models include the full slate of traditional threat actors.
+
+\modelstart
+
+\modelitem{Criminals: ordinary, organized, cyber}{Criminals threaten the \ac{LA} system by using it to access plaintext
+without authority. Ordinary thieves may spy on targets or steal and resell phones. They will have physical access but
+only simple tools that use ordinary device interfaces. Organized criminals will have the same motives but considerably
+more advanced technical capabilities, including full control over all interfaces. Cyber criminals have a wide variety of
+of motivations and capabilities. They will target the backend systems that operate the \ac{LA} system in order to abuse
+or sell whatever access capabilities they acquire.}
+
+\modelitem{Insider threat}{Insider threats are attackers with roles inside the \ac{LA} recovery process who seek to
+abuse the system for their own purposes. They could work in law enforcement, a device vendor, or a digital service
+provider. They may have the same motivations and skills of any other criminal category, but they have the advantage of
+privileged access at some stage in the process.}
+
+\modelitem{Foreign intelligence}{Foreign intelligence agencies and \acp{APT} are constantly looking for ways to elevate
+their privilege and tamper with or steal information. They have world-class expertise and computing power, which they
+can apply to both individual devices and backend systems. (Here I am referring to \ii{foreign} foreign intelligence
+agencies, as opposed to \ii{domestic} foreign intelligence agencies such as the \acs{CIA} or \acs{NSA} in the U.S.,
+which are assumed to operate within---though push the boundaries of---the legal framework.)}
+
+\modelitem{Authoritarian regimes}{Governments that do not adhere to rule of law will try to compel administrators of an
+\ac{LA} system to provide access upon demand. Legal coercion may or may not be combined with state-backed hacking
+capability.}
+
+\modelend
+
+Two non-traditional threats must be considered to make sense of an \ac{LA} proposal in the face of the stated goals and
+such a formidable list of threat actors. Understanding these threats requires viewing security with a broader than usual
+perspective.
+
+\modelstart
+
+\modelitem{The platform abuser}{The device user him or herself may actually be a threat in the larger safety context.
+Therefore, the end user's will is not the platform's top priority.
+
+\parboxindent All technologies have embedded values \cite{rogaway_moral_2015}, and at the top of each value system is
+one top priority. Traditional device security threat models prioritize the will of end user. This leads to robust design
+decisions and honors the user's rights and autonomy.
+
+\parboxindent It is dangerous to stray from this rule, but many systems already do. Auto-updating software values the
+vendor's update policy over the user's preferences. Digital Rights Management software values corporations' data rights
+over the user's (sometimes to extremes \cite{eff_2020}). Unencrypted e-mail software values the e-mail provider's
+ability to collect personal information over the user's privacy. \Acl{EA} systems value lawful investigation over the
+user's absolute privacy. Intentional subversion of the user's will violates a security practitioner's instincts, but it
+is a frequent reality.
+
+\parboxindent In the broader threat model, the platform abuser uses encryption to hide illicit activities. He or she is
+a threat to public safety.}
+
+\modelitem{Hawkish lawmakers}{In the pursuit of security and privacy for society as a whole, aggressive lawmakers that
+have the power to \ii{mandate} arbitrary data access present perhaps the largest threat of all. Crypto-anarchists will
+encrypt their data above the compromised layer regardless, but above-the-table organizations and most individuals who
+use default products will abide by U.S. law.
+
+\parboxindent Hawkish lawmakers threaten the security of a cryptosystem (whether it uses \ac{LA} or not) because they
+may judge the system as too inhibitive to law enforcement. Therefore, they may outlaw its use and mandate something
+weaker in its place.}
+
+\modelend
+
+\subsection{Out of Scope}
+
+The following considerations are beyond the scope of this threat model.
+
+\modelstart
+
+\modelitem{Encyption workarounds}{Encryption workarounds, such as finding a copy of plaintext or guessing a password,
+are always a risk. Defending against such workarounds is not the role of the \ac{LA} system itself.}
+
+\modelitem{\ac{LA} workarounds}{The user can evade \ac{LA} mechanisms by encrypting data at a higher level in the tech
+stack, for example by encrypting files in software before saving them to disk. Additionally, investigators or attackers
+can often bypass current lockout mechanisms. This has already been discussed, and is not relevant when evaluating the
+\ac{LA} system itself.}
+
+\modelitem{Supply chain attacks}{It is necessary to protect device hardware and software supply chains in order to
+prevent surreptitious, trivially exploitable backdoors. For the purposes of this threat model, device hardware and
+software are assumed to be uncompromised, aside from the standard amount of bugs and side-channels.}
+
+\modelitem{Manipulation of internal hardware}{Foreign intelligence agencies may have the capability to perform
+hyper-advanced hardware analysis and manipulation. Defending against these attacks is not the role of the \ac{LA} system
+itself.}
+
+\modelitem{Breakdown in (or absence of) rule of law}{This threat model assumes legitimacy of the legal process used to
+obtain the warrant. All forms of government search require oversight. Transparency via auditability, preferably ensured
+via technical mechanism, should be a goal for \ac{LA} proposals. However, rule of law is an absolute prerequisite for
+which no \ac{LA} mechanism can compensate.}
+
+\modelend
+
+\section{Basic Data at Rest}
+\label{sec-basic-dar}
+
+Understanding the \ldawmsr proposal and comparing it to current mobile device encryption technology requires an
+understanding of current technology. Apple iPhones are among the most secure consumer mobile phones and serve as the
+point of reference for \ldawmsR. Because this analysis is limited to a specific proposal, it is also limited to this
+reference architecture. In \myfig{fig-dfd-iphone}, a \ac{DFD} depicts the iPhone's unlock and decryption process
+according to Apple's public documentation \cite{apple_2020}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone.png}
+    \caption{The basic encrypted mobile phone data flow diagram}
+    \label{fig-dfd-iphone}
+\end{figure}
+
+This \ac{DFD}, and all that follow, presents the system at a specific level of abstraction. The system representation is
+simplified in the spirit of ``all models are wrong, but some are useful.'' The \acp{DFD} exclude some details, but are
+accurate for the purposes of communication and threat elicitation.
+
+As the diagram illustrates, several steps occur between the \ac{PIN} being entered and the device being unlocked. The
+operating system forwards the \ac{PIN} to the Secure Enclave Processor, which combines the \ac{PIN} with the
+device-specific Hardware \ac{UID} to generate the Class key. The Class key is used to decrypt the Volume key, which is
+passed to the Encryption Module. The Encryption Module performs encryption and decryption at a hardware level between
+the operating system and the encrypted storage.
+
+The \ac{UID} is not depicted in key storage because it is burned into the secure processor silicon and cannot be read by
+software. The Volume key is encrypted by the Class key before it is stored. When a user updates his or her \ac{PIN}, the
+encrypted Volume key in storage is replaced with the copy encrypted with the new Class key. Application processors do
+not handle any keys or secret information besides the \ac{PIN}. Due to inline encryption in the Encryption Module,
+persistent storage does not handle \ac{plaintext} data.
+
+\section{The \ldawmsr Proposal}
+
+\ldawmsr introduces device key escrow, an unexposed hardware interface that mediates \ac{LA} requests, and a
+vendor-mediated authorization process. To achieve its goals, the proposal relies on the concepts of self-escrow, time
+vaulting, authorization, and transparency. These concepts are summarized in \mytab{table-ldawmsr-concepts}.
+
+\begin{table}[h]
+  \caption{Central Concepts in the \ldawmsr Proposal}
+  \label{table-ldawmsr-concepts}
+  \FlushLeft
+  \small
+  \begin{tabular}{ |l|p{6.4cm}|p{6.4cm}| }
+    \hline
+    \thead{Concept} & \thead{Implementation} & \thead{Outcome} \\ \hline
+    Self-escrow
+    & The Class key is escrowed in encrypted form in the device's Secure Enclave (or equivalent) using a write-only
+      component; i.e., software can update the key, but only an unexposed hardware interface can read it.
+    & Since the Class key itself is being stored instead of the encryption scheme being changed, there is no change to
+      the security of the underlying cryptographic protocols. Physical possession and partial disassembly is required,
+      which introduces practical barriers to \ac{masssurv}.
+    \\ \hline
+    Time vaulting
+    & The hardware interface responds to requests only after proof of sustained possession for a ``lockup period (e.g.,
+      72 hours).''
+    & The waiting period precludes ``sneak and peak'' attacks and further reduces the utility for \ac{masssurv}.
+    \\ \hline
+    Authorization
+    & After the lockup period, the requestor must provide evidence of authorization. This comes from a shared secret
+      between the device and the manufacturer which can be obtained via legal process.
+    & Physical possession is now not enough; law enforcement (or attackers) must provide this secret information to the
+      device to unlock it.
+    \\ \hline
+    Transparency
+    & In addition to the device disassembly, the escrow agent modifies the device (e.g., burns a fuse) when it has been
+      unlocked. The device firmware detects this modification.
+    & The device user has evidence that the device has been unlocked, preventing covert usage.
+    \\ \hline
+  \end{tabular}
+\end{table}
+
+Device key escrow is considered the most promising technical \ac{LA} direction because of the unique capabilities that
+secure hardware offers---namely, physical possession requirements and strong transparency. The physical possession
+requirement mitigates the risk of \ac{masssurv}. As Savage adds, it also ``provides a more intuitive compatibility with
+common under-standings of the government’s reasonable law enforcement powers (e.g., the ability to seize physical
+property under court order) than more information-centric approaches, which may appear covert by comparison''
+\cite{savage_lawful_2018}. Transparency is also rooted in the physicality of the approach, wherein the device cannot be
+unlocked without making irreversible and detectable changes, reducing the risk of both \ac{masssurv} and common criminal
+abuse.
+
+One clear weakness in the design is its reliance on the secret used in the authorization protocol shared between the
+device and the manufacturer. Although the key is stored on the device, anyone could access the key if they know the
+shared secret. This is similar to trusted-party key escrow. The vendor is the trusted party, and they store the
+authorization key rather than the encryption key itself. Due to this similarity, the problem of centralized risk is not
+fully addressed. This is discussed further in \mysec{sec-prop-threats}.
+
+\ldawmsR's full data flow is illustrated in \myfig{fig-dfd-ldawmsr}. The proposal offers a few variations of the
+self-escrow and authorization protocols. This figure uses the asymmetric key pair approach for Class key encryption in
+the Secure Enclave, which creates strict hardware requirements. However, Savage offers a symmetric key variant that
+achieves the same purposes. The asymmetric key variant was chosen for analysis because it is semantically simpler.
+
+\begin{figure}[p]
+    \centering
+    \includegraphics[width=0.9\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR.png}
+    \caption{The LDAWMSR data flow diagram}
+    \label{fig-dfd-ldawmsr}
+\end{figure}
+
+When law enforcement needs access to a device, it first obtains a warrant to conduct the search. The warrant uniquely
+identifies the suspect's mobile device. Law enforcement obtains the device, disassembles it to reveal the unexposed
+\ac{LA} hardware interface, and begins the time-vaulted access request. After the lockout period is completed, the
+device offers its Device ID only. The law enforcement agency's digital forensics department digitally signs the Device
+ID and sends it to the device vendor, alongside the warrant and accompanying proof of legal authorization. The vendor's
+access compliance department audits the legal documents and submits the digitally signed Device ID to a secure computing
+environment such as a \ac{HSM}. The secure environment authenticates the digital signature, looks up the device's
+private Device Seal Key and authorization token, and encrypts them with the law enforcement agency's public key. These
+encrypted artifacts are returned to the agency, which decrypts and submits them to the device through the dedicated
+interface. First, the authorization token is hashed and compared to the stored hash. If the hash matches, the device
+accepts and uses the private Device Seal Key to decrypt the Class key, which was escrowed after being encrypted but the
+public Device Seal Key. Once the Class key is decrypted, the escrow agent burns a fuse, and the device is unlocked for
+forensic investigation.
+
+This above process is a summary of the system in the \ldawmsr proposal. However, complete proposals must account for the
+entire information lifecycle. In order to establish such a system, vendors and law enforcement must build mechanisms to
+populate and update the device and \ac{HSM} key stores and databases. \myfig{fig-dfd-ldawmsr-setup} depicts a \ac{DFD}
+for these scenarios.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{dfds/build/DAR-level-1-crypto-iphone-LDAMSR-setup.png}
+  \caption{The LDAWMSR maintenance data flow diagram}
+  \label{fig-dfd-ldawmsr-setup}
+\end{figure}
+
+\section{Discussion of Threats}
+\label{sec-prop-threats}
+
+The next step in threat modeling is to elicit threats based on the system's goals, threat actors, scope, and design.
+Recall from \mysec{sec-crypto-basics} that security can be expressed in terms of authentication, integrity,
+non-repudiation, confidentiality, availability, and authorization. Threats are often expressed as the opposite of these
+properties: spoofing, tampering, repudiation, information disclosure, \acl{DOS}, and elevation of privilege
+\cite{shostack_threat_2014}. This section analyzes \ldawmsr according to these categories. Threats that could fall into
+multiple categories, such as tampering for the effect of repudiation, are listed only once.
+
+To save space, the tables use shorthand for \acf{LE}, \acf{LEO}, \acf{LEA}, \acf{HSM}, \acf{DID}, \acf{DSK}, \acf{SN},
+and \acf{DOS}.
+
+\newcommand{\threattablesettings}{\FlushLeft \small}
+
+\newcommand{\threattablebegin}[3]{
+  \begin{FlushLeft}
+    \small
+    \singlespacing
+    \begin{longtable}{ lp{3cm}p{5.2cm}p{6cm} }
+      \caption{#1 Threats}
+      \label{table-#2}
+      \\ \toprule
+      \thead{ID} & \thead{#3} & \thead{Effect} & \thead{Mitigation}
+      \\ \midrule
+      \endfirsthead
+      \caption[]{#1 Threats (continued)}
+      \\
+      \toprule
+      \thead{ID} & \thead{#3} & \thead{Effect} & \thead{Mitigation}
+      \\
+      % \midrule
+      \endhead
+      \endfoot
+      \bottomrule
+      \endlastfoot
+}
+
+\newcommand{\threattableend}{
+    \end{longtable}
+  \end{FlushLeft}
+}
+
+\subsection{Spoofing}
+
+Spoofing threats are listed in \mytab{table-spoofing}. Notably, \ii{S2} is Eran Tromer's attack \cite{tromer_2018} on
+computer scientist Ray Ozzie's related \ac{DAR} \ac{LA} protocol called CLEAR \cite{ozzie_2018}. Savage acknowledges the
+influence that CLEAR had on \ldawmsR, and specifically included a note that this proposal is technically vulnerable to
+the same threat. This threat is not particularly severe, as the difficulty of such an attack is so high that it is
+unlikely to be attempted. Even so, the mitigations listed here would also help.
+
+\threattablebegin{Spoofing}{spoofing}{Spoof}
+  S1 & Device is spoofed to LE via plant
+  & Attackers plant a device to get LE to unlock it for them.
+  & Require the device to be in the forensics department itself the entire time it is unlocked.
+  \\ \hline
+  S2 & Device is spoofed to LE via forged DID
+  & Attackers rig a device to report a different DID and record the entered unlock artifacts, allowing them into a
+    target device.
+  & Add a hardware integrity check or alternatively unlock the device in a Faraday cage (to prevent telemetry), extract
+    its data, and destroy it (to prevent recovering stored data).
+  \\ \hline
+  S3 & LEO is spoofed to LEA
+  & Attackers pose as a LEO while the LEA relays back unlock artifacts.
+  & Require the device to be in the forensics department itself the entire time it is unlocked.
+  \\ \hline
+  S4 & LEA is spoofed to vendor during key exchange
+  & Attackers pose as an LEA in order to spoof future access requests.
+  & Vendor key exchange process involves thorough in-person verification. Assume only a handful of digital forensics
+    laboratories nationwide are authorized to participate.
+  \\ \hline
+  S5 & LEA is spoofed to vendor during access request
+  & Attackers pose as a LEA and submit an access request. Requires forged legal documents and a stolen LEA private key
+    (unless attacker is an insider or has executed a spoofed key exchange).
+  & Vendor independently verifies all legal documents. Vendor supports a key revocation and update process that involves
+    the same level of in-person verification. Assume only a handful of digital forensics laboratories nationwide are
+    authorized to participate.
+  \\ \hline
+  S6 & Manufacturer is spoofed to vendor
+  & Attackers pose as the manufacturer to inject malicious data into the DID-DSK database. Must be combined with other
+    attacks to be of use.
+  & Vendor authenticates data transfers between manufacturing and corporate (does not address a compromised manufacturer
+    network or an inside threat).
+\threattableend
+
+\subsection{Tampering}
+
+Tampering threats are listed in \mytab{table-tampering}. Note that execution of \ii{S2} above relies on a tampering
+attack such as \ii{T1}.
+
+\threattablebegin{Tampering}{tampering}{Tamper}
+  T1 & Platform abuser corrupts DSK, DID, or token
+  & LE attempts at perform \ac{LA} will fail due to corrupt data. Requires compromise of the Secure Enclave or advanced
+    hardware manipulation tools.
+  & Since these values are known at manufacturing time and never change, they can be made read-only. (The symmetric key
+    variant of \ldawmsr requires updating the symmetric equivalent of the public DSK, and is therefore more vulnerable
+    to this threat.)
+  \\ \hline
+  T2 & Platform abuser destroys the \ac{LA} hardware interface
+  & LE attempts to perform \ac{LA} will fail due to damaged hardware.
+  & Manufacture the device to require destructive disassembly to access the interface. This increases the cost
+    associated with deploying and using the system.
+  \\ \hline
+  T3 & LE insider tampers with DID before digital signature
+  & LEA receives an unauthorized device's unlock artifacts. Can be combined with other attacks to compromise a device.
+  & Randomize DIDs to make them unguessable, meaning a target device must be in possession in order for the attacker to
+    know what DID to use. Input to the vendor's \ac{HSM} could also include data from the legal documents, such as
+    device S/N; the \ac{HSM} would append the S/N to the DID and hash the result, verifying a match with the same stored
+    hash before returning the unlock artifacts.
+  \\ \hline
+  T4 & Attacker tampers with hardware transparency mechanism
+  & The device user is unaware that the device was unlocked. Highly implementation dependent, but certainly requires
+    advanced hardware manipulation tools.
+  & Manufacture the device to require destructive disassembly to access the interface.
+\threattableend
+
+\subsection{Repudiation}
+
+Repudiation threats are listed in \mytab{table-repudiation}.
+
+\threattablebegin{Repudiation}{repudiation}{Repudiation}
+  R1 & Attacker or authoritarian government repudiates unlock attempt
+  & The user is unaware that someone tried to access their device.
+  & Manufacture the device to require destructive disassembly to access the interface, or use the same hardware
+    transparency mechanism to mark mere access attempts as well.
+  \\ \hline
+  R2 & Authoritarian government repudiates demanding vendor assistance
+  & Governments not respecting the legal framework governing the system can bully the vendor and try to force access
+    while maintaining deniability.
+  & This is close to being an out of scope threat under the ``rule of law'' exception. However, one could design the
+    system to require requests and responses to be recorded on a public ledger. This creates several new difficulties.
+  \\ \hline
+  R3 & LEA repudiates having unlocked a device
+  & The device is undeniably unlocked, but the LEA claims it was not them, allowing possible violations of civil
+    liberties and harming transparency.
+  & At a high level, the system should require a long paper trail before access can be granted. At a low level, the
+    escrow agent could require and store some identifying information before unlocking, such as the LEA's public key
+    digitally signed by the vendor, which provides non-repudiation assuming independent access to the device. See also
+    the mitigation in \ii{R2}.
+  \\ \hline
+  R4 & Vendor repudiates having responded to an access request
+  & The device is undeniably unlocked, but the vendor claims they did not provide the unlock artifacts, allowing the
+    vendor to cover up leaks or distasteful policy decisions by suggesting the device was instead hacked. Harms
+    transparency.
+  & Assuming independent access to the device, the hardware transparency mechanism should confirm whether the \ac{LA}
+    process was used, and the same identifying information used in \ii{R3} would implicate the vendor. See also the
+    mitigation in \ii{R2}.
+\threattableend
+
+\subsection{Information Disclosure}
+
+Information disclosure threats are listed in \mytab{table-disclosure}. In addition to the threats listed, any attack
+aimed at gaining unauthorized access to a device is an information disclosure threat. Most spoofing attacks aim to
+disclose information intended for another party.
+
+\threattablebegin{Information Disclosure}{disclosure}{Information \\ Disclosed}
+  I1 & Attacker discloses the DSK, DID, the encrypted class key, or hashed token
+  & By design, none of the information stored on the phone is sensitive. The most consequential information that could
+    be stolen is the DID, which could be paired with other attacks. Disclosing the DID requires either disassembly and
+    time vaulting or advanced hardware inspection tooling.
+  & Sensitivity of this data has already been mitigated as much as possible.
+  \\ \hline
+  I2 & Attacker intercepts unlock artifacts between LEA \ac{HSM} and device
+  & The attacker is able to unlock the target device alone. Must be combined with a spoofed DID to be of use to other
+    devices (see \ii{S2}). The attacker could record the values to replay them after the phone has been returned.
+  & See \ii{S2}. If choosing not to destroy the device, to prevent replay attacks, a new token could be generated and
+    stored in the vendor HSM, with the new hash installed during the unlock process.
+  \\ \hline
+  I3 & Attacker steals the LEA private key
+  & The attacker is now able to spoof the LEA and decrypt returned unlock artifacts. Without the devices (which are
+    already in LE possession), the artifacts are not useful. See \ii{S4} and \ii{S5} for LEA spoofing attacks.
+  & The key is stored in an \ac{HSM}, which has extensive security measures. See \ii{S5} for a revocation and update
+    mitigation.
+  \\ \hline
+  I4 & Attacker steals the DID-DSK database
+  & This is the big prize, the largest source of centralized risk. With the database, attackers could directly
+    compromise any mobile device using this system in their possession, after disassembly, and after the time vault.
+    They may also try to sell access to the data.
+  & \ii{R3} defines a mitigation that requires use of a vendor private key to create a digital signature as part of the
+    authentication process. If used, that key would have to be compromised as well for the database to be of use.
+    However, if the database is compromised, that key probably is too. Savage argues that \acp{HSM}, though imperfect,
+    actually do provide very strong assurance. Responses could be rate-limited, and leaking an entire database once it's
+    in the \ac{HSM} could not go unnoticed. Getting the data in may be harder---see \ii{I5}.
+  \\ \hline
+  I5 & Attacker steals device unlock artifacts between manufacturing and vendor \ac{HSM}
+  & Intercepting the unlock artifacts at the source achieves the same effect as \ii{I4}, only the data set is not
+    complete. Instead of granting the ability to unlock any device, including a specific target, this is suited for
+    getting access down the line or for sale.
+  & The unlock artifacts generation process must be treated as sensitively as the \ac{HSM} itself. The artifacts should
+    be encrypted immediately and periodically transported to the corporate \ac{HSM} on physical media. This threat is
+    difficult to definitively mitigate.
+  \\ \hline
+  I6 & Attacker steals device \ac{plaintext} from LEA after unlock process
+  & Assuming the attacker doesn't care that the LEA gets access to plaintext, they could ``help'' the LE investigation
+    in order to access the data later either when it comes out as public evidence or through compromising the LEA
+    itself. This may be useful to authoritarian regimes, foreign intelligence, and organized crime.
+  & LEAs must have strong information security of their own. Disclosure of evidence is a natural side effect of
+    prosecution, so LEAs must practice discretion in choosing which devices to unlock in case mere information discovery
+    poses a great risk in itself.
+\threattableend
+
+\subsection{Denial of Service}
+
+\Ac{DOS} threats are listed in \mytab{table-dos}. In addition to the threats listed, \ii{S4} can be used for \ac{DOS} by
+surreptitiously invalidating a LEA's identity key, and \ii{T1} and \ii{T2} are tampering attacks designed to deny LEAs
+from using the \ac{LA} mechanism.
+
+\threattablebegin{Denial of Service}{dos}{Service Denied}
+  D1 & Unlock artifacts recovery
+  & \ii{S6} could be used to inject corrupt data to the DID-DSK database, potentially overwriting actual unlock artifact
+    data.
+  & Verify integrity of new data before entering it into the database, and disallow overwriting entirely.
+  \\ \hline
+  D2 & Unlock artifacts recovery
+  & Determined adversaries could physically destroy the medium storing the DID-DSK database.
+  & The \ac{HSM} should be in a physically secure and guarded facility, and the storage medium should be stable and
+    persistent.
+\threattableend
+
+\subsection{Elevation of Privilege}
+
+Elevation of privilege are listed in \mytab{table-privilege}. Spoofing to usurp a entity's privilege or using any attack
+to unlock a device without authorization can be seen as elevation of privilege.
+
+\threattablebegin{Elevation of Privilege}{privilege}{Privilege Gained}
+  E1 & Foreign intelligence or organized criminals recruit an insider
+  & The primary attacker gains inside access through coercion or bribery, making any other attack easier.
+  & Limit the employees, accounts, and computers involved in any element of the system. Perform background checks on
+    each employee and maintain detailed logs.
+  \\ \hline
+  E2 & Unexposed hardware interface allows access to Secure Enclave
+  & The attacker gains access to Secure Enclave operations and data, undermining the device and achieving unlock without
+    following the \ac{LA} protocol.
+  & The time vaulting protocol and all communication over the interface must be thoroughly tested, and if possible,
+    formally proven to be logically sound. Analysis must include side channel attacks to affect the time vaulting or
+    leak internal data.
+  \\ \hline
+  E3 & An \ac{HSM} vulnerability allows access to protected data
+  & This is a technical route to achieving \ii{I4}, disclosure of the DID-DSK database.
+  & The \ac{HSM} should be on an air-gapped network disconnected to the internet, physically guarded, and regularly
+    patched. There is no defense against an unknown vulnerability, so security ultimately comes down to incident
+    detection and response.
+\threattableend
+
+\section{Risk Analysis}
+
+The amount of risk \ldawmsr introduces depends on the number of suggested mitigations that are implemented. Some
+possible attacks introduce inherently little risk due to their difficult nature and low pay off. Others introduce
+greater risk but can be reasonably addressed. Some threats specifically target transparency, and therefore are difficult
+to address without trust. Finally, there are some risks that cannot be safely mitigated. Each of these classes of
+threats are discussed in turn.
+
+\ii{S6}, \ii{T1}, \ii{T3}, \ii{I1}, \ii{I3}, \ii{I6}, \ii{D1}, \ii{D2}, and \ii{E1} \ii{E2} are each low-risk. They
+either are preventable, unuseful, or already present in the baseline depicted in \mysec{sec-basic-dar}. Although these
+threats are not fatal flaws to a design, it is important for a designer to consider and mitigate the risk they pose.
+Unless they appear in combination with other threats, they do not deserve more discussion.
+
+\ii{S1}, \ii{S2}, \ii{S3}, \ii{S4}, \ii{S5}, \ii{T2}, \ii{T4}, \ii{R1}, and \ii{I2} are all higher-risk threats to the
+security, trustworthiness, and effectiveness of the proposal. However, they can be mitigated through a set of related
+requirements: require destructive disassembly of the device to access the \ac{LA} hardware interface, require the device
+to be physically present on LEA premises the entire duration it is unlocked, and grant only a handful of LE facilities
+the lawful right to use the \ac{LA} mechanism. Requiring destructive disassembly entirely precludes certain attacks.
+Requiring the device to be on LEA premises prevents attackers from manipulating LE into achieving the attackers' ends.
+Limiting the number of LE facilities that can participate in the \ac{LA} process prevents LE impersonation attacks.
+Together, these mitigations make the proposal more resistant to attack.
+
+\ii{R2}, \ii{R3}, and \ii{R4} are repudiation attacks that belie the system's potential for abuse. The system relies on
+legal processes and cooperation of several parties. Even in the U.S., the legal process may involve secret courts,
+secret investigations, and secrecy orders on the vendor \cite{shamsi_2011}. Under these conditions, transparency is
+difficult to achieve. The destructive disassembly requirement enables transparency at a device level. Auditable logs of
+device access \ii{requests and responses} would enable transparency at a system level. \ldawmsr does not create such a
+log, though other \ac{DAR} \ac{LA} proposals do \cite{phan_key_2017} \cite{servan_schreiber_jje_2020}
+\cite{goldwasser_public_2017}. Unfortunately, building this feature into a proposal makes it potentially vulnerable the
+hawkish lawmaker, who may consider too much transparency to be an anti-feature.
+
+Finally, there are the threats of \ii{I4}, \ii{I5}, and \ii{E3}, which each represent the disclosure of unlock artifact
+data (the DID-DSK database). These threats create a large amount of risk that cannot be safely mitigated. Storing so
+much sensitive data in a central database creates a high-value target for foreign intelligence and cyber criminals. The
+vendor needs to protect this database from the most advanced attackers in the world while also regularly updating and
+accessing it. \Acp{HSM} and robust security policies would make that a difficult task. However, the data will almost
+certainly leak eventually. The effects of such an event would depend on who stole the unlock artifacts. Criminals would
+use the data for personal and financial gain; foreign intelligence and authoritarian regimes would use it to spy on
+their enemies. In the worst-case scenario, anyone with specialized tools and disregard for the law could get into any
+compliant device after disassembling it and waiting through the time vault.
+
+It is important to give \ldawmsr (and similar \ac{DAR} \ac{LA} proposals) its due. The worst case scenario still
+precludes \ac{masssurv} and includes non-repudiation that access occurred. This is a significant gain for \ac{LA}
+proposals in terms of transparency and protection of civil liberties.
+
+\section{Achievement of Goals}
+
+This chapter began with a list of goals by which \ac{DAR} \ac{LA} proposals can be measured, including security,
+protection of civil liberties, transparency, law enforcement utility, and economic impact. In this section \ldawmsr is
+measured against each of these goals.
+
+\paragraph*{Security} Security has already been discussed at length in this chapter. \ldawmsr introduces significant
+risk in the DID-DSK database itself, but performs well in many other respects, especially if the recommended mitigations
+are implemented. When the unlock artifacts are disclosed, the system becomes simple to attack. However, even in these
+conditions, attackers are limited to devices in their physical possession and cannot hide access from their victims.
+
+\paragraph*{Protection of civil liberties} This goal was defined as defending against \ac{masssurv} and abuse by
+authoritarian regimes. \ldawmsr successfully protects against \ac{masssurv} due to the physical possession and
+destructive disassembly requirements. It is not as effective against authoritarian regimes, which would put intense and
+possibly coercive legal pressure on vendors to unlock devices. In the case of an unlock artifacts leak, regimes could
+broadly abuse the system. This technology would be unwelcome to those living in countries that do not respect rule of
+law.
+
+\paragraph*{Transparency} \ldawmsr provides transparency to the device user, but does not provide transparency to the
+public. As discussed in the previous section, the destructive disassembly requirement would result in transparency at
+the device level. However, access to a device can only be confirmed if the device is available. Neither the user nor the
+public can detect if a device has been unlocked under secretive conditions, whether lawfully or criminally. This creates
+potential for government abuse.
+
+\paragraph*{Law enforcement utility} \ldawmsr includes a few tradeoffs between security and law enforcement utility. The
+requirements to have physical possession of the device, to wait through the time vault, and to limit the number of
+authorized law enforcement laboratories all create barriers to usability. The physical possession requirement is too
+important for security to be compromised for utility. Additionally, search and seizure law is already designed for
+physical evidence; maintaining the possession requirement makes the system operate within well-understood norms. The
+time vault prevents against time-sensitive attacks, though destructive disassembly (not part of the original proposal)
+may be enough to cover these cases. The time vault period could be decreased without significant impact. Limiting the
+number of laboratories that can use the system reduces the risk of spoofing attacks; local law enforcement would have to
+cooperate with one of these special laboratories, such as a regional \ac{FBI} office, to unlock a device. In the end,
+law enforcement still has a reliable way to get into lawfully seized devices. \ldawmsr does its part, but the government
+must fund the \ac{FBI} operation and cultivate cooperation between local and federal agencies.
+
+\paragraph*{Economic impact} Expenses to the vendor fall into three categories---the costs of implementing the design,
+running the system, and absorbing lost business. Even if a proposal like \ldawmsr was refined enough to be implemented,
+the vendor would need to update hardware designs and create the software systems to get it started. Once in use, the
+vendor would face considerable costs in running it securely due to the difficulty of safely accessing and updating the
+unlock artifacts. Finally, the vendor would certainly lose privacy-conscious customers to small businesses, foreign
+competitors, or resellers offering pre-\ac{LA} devices. The scale of these costs relative to device manufacturer's large
+size is unclear. Costs to the government include the \ac{FBI} laboratory cost and the cost of law enforcement resources
+and training. \acs{CALEA} includes provisions for government reimbursement of vendor compliance costs, and a similar
+provision could be applied to \ac{DAR} \ac{LA} regulation \cite{edwards_hr4922_1994}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %
+
+\bibliographystyle{acm}
+\bibliography{manuscript}
 

### 2021-02-23 scripts/install-deps.sh
Removed packages that do not exist in ubuntu 20. Added missing dependencies. Adjusted to work on wsl

@@ -13,20 +13,20 @@ sudo apt-get install -y \
     texlive-latex-extra \
     texlive-science \
     texlive-xetex \
-    texlive-generic-extra \
+    texlive-plain-generic \
     texlive-publishers \
     latexmk \
     inkscape \
     libcanberra-gtk-module \
     libcanberra-gtk3-module \
-    npm
+    inkscape
 
 curl -sL https://deb.nodesource.com/setup_10.x | bash -
 sudo apt-get install -y --no-install-recommends \
     nodejs
 sudo npm install -g @argdown/cli
 
-if which snap &>/dev/null; then
+if which snap &>/dev/null && !grep microsoft /proc/version &>/dev/null; then
     snap install drawio
 else
     URL=$(curl -sL https://github.com/jgraph/drawio-desktop/releases/latest | grep "\.deb" | head -1  | cut -d\" -f2)

### 2021-02-23 articles/tots-ieee/Article.tex
Update TOPS ACM article, copy in chapter 6
changed author format to match journal requirements. Added placeholders for author biographies

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SETUP %
-\documentclass[conference]{IEEEtran}
+\documentclass{IEEEtran}
 \usepackage{cite}
@@
 %%%%%%%%%%% Commands
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-\def\tt#1{\mbox{\texttt{#1}}}
-
-\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}
-
-%% Mine
 \def\tt#1{\mbox{\texttt{#1}}}
@@
 
-\author{\IEEEauthorblockN{Kevin Kredit}
-\IEEEauthorblockA{\textit{GVSU ACS Department} \\
-Grand Rapids, Michigan \\
-Email: k.kredit.us@ieee.org}
-\and
+\author{Kevin Kredit and Andrew Kalafut\thanks{K. Kredit was with Grand Valley State University. A. Kalafut is with Grand Valley State University}}
+
+%\IEEEauthorblockN{Kevin Kredit}
+%\IEEEauthorblockA{\textit{GVSU ACS Department} \\
+%Grand Rapids, Michigan \\
+%Email: k.kredit.us@ieee.org}
+%\and
 % FIXME: Prof. Kalafut, update as appropriate
 % Is there a way to indicate that you are faculty and I am not? Label yourself with PhD, perhaps?
-\IEEEauthorblockN{Andrew Kalafut, Ph.D.}
-\IEEEauthorblockA{\textit{GVSU ACS Department} \\
-Grand Rapids, Michigan \\
-Email: kalafuta@gvsu.edu}
-}
+%\IEEEauthorblockN{Andrew Kalafut, Ph.D.}
+%\IEEEauthorblockA{\textit{GVSU ACS Department} \\
+%Grand Rapids, Michigan \\
+%Email: kalafuta@gvsu.edu}
+%}
 
@@
 
+\begin{IEEEbiography}{Kevin Kredit}
+  Placeholder biography for Kevin. Please fill in.
+\end{IEEEbiography}
+
+\begin{IEEEbiography}{Andrew Kalafut}
+  Placeholder biography for prof. kalafut. Will fill in later 
+\end{IEEEbiography}
+
 \end{document}

### 2021-03-03 articles/tops-acm/manuscript.tex
Add DFD syntax to TOPS article

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PAPER %
-\section{Introduction}
+
+% TODO: add actual intro.
+
+\section{Threat Modeling with Data Flow Diagrams}
+\label{sec-threat-model-intro}
+
+Threat modeling is an important step in analyzing the security at the systems level. Models abstract away fine details
+and focus on the architecture, processes, and dataflows in a system. They assist in the understanding of current systems
+and the prevention of problems in new systems, both of which are important when facing the prospect of designing an
+\ac{EA} mechanism built on top of a complex and aging technology stack.
+
+Threat modeling begins with the questions, ``What are you building?'' and ``What can go wrong?''
+\cite{shostack_threat_2014}. Accurately answering the first question is crucial, particularly for the field of
+cryptography, which hinges on precise definitions of security requirements \cite{varia_2018}. The answer to the second
+question depends on the types of attackers under consideration and determines the scope of threats to be considered.
+``What are you building?'' hasn't been asked enough in the current phase of the \ac{EA} debate, and ``What can go
+wrong?'' cannot be faithfully answered without knowing what is being built.
+
+\Acp{DFD} were designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and external
+entities. \Acp{DFD} are well-suited for threat modeling because security vulnerabilities tend to follow data flow, not
+control flow \cite{shostack_threat_2014}. They are particularly well-suited for \ac{EA}, as data privacy is of primary
+concern. \myfig{fig-dfd-dh-traditional} shows the basic elements of a \ac{DFD} of a simplified Diffie-Hellman key
+exchange.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{dfds/build/diffie-hellman-traditional.png}
+    \caption{A \acs{DFD} for a simplified Diffie-Hellman key exchange}
+    \label{fig-dfd-dh-traditional}
+\end{figure}
+
+\Acp{DFD} show what data is transferred via labels on the flows. However, labels can be insufficient in complex
+\acp{DFD}, and the method has no syntax for computation or cryptographic protocols. In order to maximize their
+communicative ability in the context of encryption and \acl{EA} schemes, I introduce the additional syntax show in
+\myfig{fig-dfd-dh-updated}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{dfds/build/diffie-hellman.png}
+    \caption{A Diffie-Hellman key exchange illustrated with additional syntax}
+    \label{fig-dfd-dh-updated}
+\end{figure}
+
+The new syntax illustrates the data in more detail as it is stored, transferred, and operated on. By illustrating the
+data itself, it eliminates the need for labels on lines and can be used to communicate entire protocols in a single
+diagram. \myfig{fig-dfd-key} contains the full set of symbols. \mysec{sec-intro} includes \acp{DFD} in this style to
+analyze threats that EA would introduce.
+
+% TODO: fix \mysec{sec-intro} reference in above paragraph
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=0.85\linewidth]{dfds/build/dfd-key.png}
+    \caption{Expanded set of DFD symbols}
+    \label{fig-dfd-key}
+\end{figure}
+
+\section{Analysis Introduction}
 \label{sec-intro}
 
-% TODO: add actual intro, not just chapter intro. Take some from paper intro, but also introduce the diagram syntax
+% TODO: add actual intro, not just chapter intro. Take some from Thesis intro.
 

### 2021-04-06 articles/tots-ieee/Article.tex
modifications to make paper stand-alone, including pulling in paragraphs from thesis intro

@@
 
-\title{Confronting Wicked Crypto}
+\title{Lawful Access: Arguments and Alternatives}
 
@@
 % NOTE: The first paragraph of the abstract was copied. The second I tweaked.
-
 Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
@@
 
-This chapter will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
-debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
+In encryption policy, \ac{EA} is alternative means of
+decryption intended for law enforcement use. Characterized by a dynamic technological environment, disagreement about
+underlying values, and resistance to a clear solution, the debate on \ac{EA} will not go away.
+
+The conflict at the heart of the encryption and \ac{EA} debate is this: the same cryptographic and design principles
+that underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
+foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
+communications and storage provides clear benefits. In a world where information security is frightfully poor yet
+increasingly important, the necessity of strong encryption cannot be understated. However, the privacy afforded by
+certain encryption technologies hampers law enforcement investigations and hides wrongdoing \cite{cox_2020}
+\cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk should not be
+ignored.
+
+The conversation regarding encryption has reached a stalemate. Governments and law enforcement agencies frequently cite
+the need to access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}.
+Human rights groups and technical leaders counter that a weakened encryption environment would fatally compromise
+privacy and security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate
+continues, many on both sides have become entrenched in their positions and have engaged in disingenuous attacks against
+their opponents.
+
+In the long term, increased regulation of the tech industry is inevitable and regulation of encryption is possible.
+Despite the benefits of strong encryption, regulatory interest in the subject has not subdued. The form that regulation
+will take in the field of encryption depends on the good faith efforts made to equitably balance the benefits and risks
+involved in deploying a cryptographic system. If the technical community acts, it can lead and shape legislation rather
+than be subjected to it.
+
+Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
+community fails to act but regulators move forward, everyone may become subject to technically misguided, harmful laws.
+Bad policy is just as much a threat as weak passwords. Due to this threat, as long as lawmakers continue pursuing
+\ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and respond to the
+presented arguments.
+
+This paper will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
+debate, although two important shortcomings of the maps presented here must be acknowledged. First, argument maps should be used simply
 as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
@@
 
-Information security is the remaining central issue in the debate. As described in \mysec{sec-intro}, \ac{cryptography}
+Information security is the remaining central issue in the debate. \ac{cryptography}
 plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
-cryptographic foundations of encryption; however, this is seen as too risky today. \mysec{sec-intro} introduced several
-alternative technical approaches to \ac{EA}, but experts still emphasize that current systems cannot securely provide
+cryptographic foundations of encryption; however, this is seen as too risky today.
+Although there are several alternative technical approaches to \ac{EA}, experts still emphasize that current systems cannot securely provide
 the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
@@
 \ac{EA} would put a heavy burden on a field still finding its legs.
-
-% TODO: fix \mysec{sec-intro} reference in above paragraph
-% TODO: fix \mysec{sec-intro} reference in above paragraph (there are two instances)
 
@@
 
-\section{Going Dark vs. The Golden Age}
+\section{The privacy vs safety debate}
 
@@
 in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
-thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
-in \mysec{sec-intro}.
-
-% TODO: fix \mysec{sec-intro} reference in above paragraph
+thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument.
 
@@
 the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
-legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
-non-\ac{EA} approaches were first introduced in \mysec{sec-intro}.) In this section, \ac{EA} is compared to its
+legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. In this section, \ac{EA} is compared to its
 alternatives. \ac{EA} variations are compared in the next map.
-
-% TODO: fix \mysec{sec-intro} reference in above paragraph
 
@@
 conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
-of brevity, this thesis compares \ac{EA}'s alternatives and variations in single maps. For a thorough discussion of both
+of brevity, this paper compares \ac{EA}'s alternatives and variations in single maps. For a thorough discussion of both
 \ac{EA}'s alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report
@@
 a silent listener on a conference call. This was notably suggested by UK \ac{GCHQ} officials Ian Levy and Crispin
-Robinson \cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:
-
-\begin{displayquote}
-In any discussion of cyber security, details matter.
-
-Unfortunately, it's the details that are missing from the discussion around lawful access to commodity end-to-end
-encrypted services and devices (often called the ``going dark'' problem). Without details, the problem is debated as a
-purely academic abstraction concerning security, liberty, and the role of government.
-
-There is a better way that doesn’t involve, on one side, various governments, and on the other side lawyers,
-philosophers, and vendors' PR departments continuing to shout at each other. If we can get all parties to look at some
-actual detail, some practices and proposals---without asking anyone to compromise on things they fundamentally believe
-in---we might get somewhere.
-\cite{levy_robinson_2018}
-\end{displayquote}
-
-Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
+Robinson \cite{levy_robinson_2018}. Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
 proposal consider the uncompromised use of strong encryption as the proposal's great strength. Although this is a good
@@
 commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
-leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
-lessons of \mysec{sec-intro}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has led
-to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This is
+leading to pervasive data collection with little oversight \cite{koepke_2020}. 
+This situation does not negate the value of research into device-oriented \ac{EA}. This is
 because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
@@
 privacy and civil liberties more systematically than criminal abuse.
-
-% TODO: fix \mysec{sec-intro} reference in above paragraph
 
@@
 
-The next chapter focuses on one specific device key escrow proposal. It defines a threat model and analyzes Stefan
-Savage's 2018 proposal, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
-\cite{savage_lawful_2018}.
+%The next chapter focuses on one specific device key escrow proposal. It defines a threat model and analyzes Stefan
+%Savage's 2018 proposal, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
+%\cite{savage_lawful_2018}.
 

### 2021-04-23 articles/tots-ieee/Article.tex
ToTS: Updates to bio and comments
- Don't use 'Lawful Access' term
- Avoid 'wicked problems' deep dive to save time
- Justify some paragraphs to column 100
- Update own bio

@@
 
-% NOTE: I think we should switch terminology to "Lawful access" for the paper (as opposed to EA)
-
 % NOTE: The first paragraph of the abstract was copied. The second I tweaked.
@@
 
-% TODO: add actual intro, not just chapter intro. Take some from paper intro, but also briefly discuss wicked problems
+% TODO: add actual intro, not just chapter intro. Take some from paper intro. Introduce topic as complex and
+% multifaceted, but avoid actual "wicked problem" terminology for sake of space.
 
-In encryption policy, \ac{EA} is alternative means of
-decryption intended for law enforcement use. Characterized by a dynamic technological environment, disagreement about
-underlying values, and resistance to a clear solution, the debate on \ac{EA} will not go away.
+In encryption policy, \ac{EA} is alternative means of decryption intended for law enforcement use. Characterized by a
+dynamic technological environment, disagreement about underlying values, and resistance to a clear solution, the debate
+on \ac{EA} will not go away.
 
@@
 
-This paper will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
-debate, although two important shortcomings of the maps presented here must be acknowledged. First, argument maps should be used simply
-as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
-replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of an
-argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against one
-indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger. With
-those qualifications in mind, let us begin by examining the factors at the center of the conflict.
+This paper will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the debate,
+although two important shortcomings of the maps presented here must be acknowledged. First, argument maps should be used
+simply as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are
+not a replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of
+an argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against
+one indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger.
+With those qualifications in mind, let us begin by examining the factors at the center of the conflict.
 
@@
 
-Information security is the remaining central issue in the debate. \ac{cryptography}
-plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
-cryptographic foundations of encryption; however, this is seen as too risky today.
-Although there are several alternative technical approaches to \ac{EA}, experts still emphasize that current systems cannot securely provide
-the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
-early computing and networking designs. The field of cybersecurity is still equal parts art and science; mandated
-\ac{EA} would put a heavy burden on a field still finding its legs.
+Information security is the remaining central issue in the debate. \ac{cryptography} plays a foundational role in nearly
+every aspect of security. Past \ac{EA} regulation efforts sought to compromise the cryptographic foundations of
+encryption; however, this is seen as too risky today. Although there are several alternative technical approaches to
+\ac{EA}, experts still emphasize that current systems cannot securely provide the level of access \acl{LE} asks for
+\cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs.
+The field of cybersecurity is still equal parts art and science; mandated \ac{EA} would put a heavy burden on a field
+still finding its legs.
 
@@
 \begin{IEEEbiography}{Kevin Kredit}
-  Placeholder biography for Kevin. Please fill in.
+  B.S Engineering from Calvin University, M.S. Applied Computer Science from Grand Valley State University.
+  % NOTE: I'm not really sure what these are supposed to contain. I tried to view some example papers, but couldn't get
+  % access.
 \end{IEEEbiography}
